### 2018-04-26 Gender homophilly project.Rmd
first commit

---
title: "Homophily in science: do researchers preferentially collaborate with same-gendered colleagues?"
subtitle: Online Supplementary Material
author: "Luke Holman and Claire Morandin"
output:
  html_document: 
    toc: true # show a table of contents
    toc_float: true # make table of contents float at the side of the page
    depth: 3  # upto three depths of headings (specified by #, ## and ###)
    number_sections: false  # set to true if you want to number sections at each table header
    theme: lumen # Luke's favourite fonts and colours
    code_folding: hide # Add some buttons to show/hide the code
---

## Setting up and defining custom functions

### Set the working directory, load up R packages and data
Remember to re-set the working directory and install the packages first if needed.

```{r warning=FALSE, message=FALSE, results='hide'}
library(dplyr) # 'data wrangling' 
library(stringr) # string manipulation
library(ggplot2) # plots
library(parallel) # parallel computing
library(lme4) # mixed models
library(lmerTest) # mixed models
library(pander) # nice Markdown tables
library(kableExtra) # scrollable table
library(tidyr) # spread and gather functions
library(gridExtra) # arranging ggplots
library(mgcv) # generalised additive models
library(ggbeeswarm) # beeswarm plots
library(Cairo) # for exporting the alpha characters via ggsave

# Using a database allows quick access to the whole PubMed dataset without loading the whole thing to memory.
pubmed_db <- src_sqlite("~/Desktop/Gender Pubmed mining project/data for analysis/pubmed_gender_data.sqlite3", create = F)
pubmed_sqlite <- tbl(pubmed_db, "papers") # information about each individual paper
journals_sqlite <- tbl(pubmed_db, "journals") # information about each journal
```

```{r echo=FALSE, results='hide'}
library(knitr) # set up cacheing to save time when re-building the html document - chunks are not re-evaluated if the code has not changed
figure.counter <- 2  # Counters for numbering figures and tables
supp.figure.counter <- 1
table.counter <- 1
supp.table.counter <- 1
```
`r opts_chunk$set(cache=TRUE)`

### Function to retrieve data from the SQLite database
Importantly, this function discards all papers where we do not know the gender of every single author. This is conservative but greatly simplifies the test -- we'd have to be careful how we included papers where some author's genders were unknown. Our failures to identify author gender are almost certainly randomly distributed across papers with respect to gender homophily, this omission makes our tests less sensitive but should not bias their outcomes.

The function returns all the data that we need to run the test on every journal for which we found more than `minimum.n` suitable papers that were published in the last `nYears`.
```{r}
# This function converts dates into a different format, namely time before 'the present' in years. 
# 'The present' is defined as 20/08/2016 which is the date that Luke downloaded a local copy of the Medline database.
convert.dates <- function(dates) {
  dates <- as.numeric(as.Date(dates, "%d_%m_%Y")) - as.numeric(as.Date("20_08_2016","%d_%m_%Y")) # Convert the date to days before the present
  return(dates / 365.25)  # Convert from days to years
}

retrieve.authorship.data <- function(journals, minimum.n, start.of.time.period, end.of.time.period, file.name, over.write = FALSE){

  if(over.write | !(file.name %in% list.files())){ # Only do the test if over.write = TRUE or there is no datafile in the working directory

    # Declare variables to hold the output
    journal.names <- rep(NA, 10^5)
    count.data <- list()
    counter <- 0

    # Loop over all the journals, try to get the data, and if there is enough, store the data we need as a list entry in 'count.data'
    for(i in 1:length(journals)){

      # Get the relevant data for the focal journal out of the database using dplyr
      # the gender.95 column contains author genders in a format like "MFU" - 
      # MFU means the first author was male, second female, and third was of unknown gender. 
      # Genders were only recorded as M or F if the person's given name is associated with one gender >95% of the time.
      focal <- pubmed_sqlite %>% 
        filter(journal == journals[i]) %>% # restrict to focal journal
        select(gender.95, date) %>%        # get these columns
        collect(n = Inf) %>%               # get the data into memory
        filter(nchar(gender.95) > 1)       # only keep papers with >1 author

      # Convert the data column into 'years before present', to the nearest day
      focal$date <- convert.dates(focal$date)

      # Restrict the analysis to a specified period of time. My logic is that if the gender ratio changes over time, it will give the false impression of homophily, as explained in Bergstrom et al.'s comment here: http://www.michaeleisen.org/blog/?p=1931
      # For example for papers from the last one year in the dataset, set start.of.time.period = 1 and end.of.time.period = 0
      # For papers from between 5 and 6 yeras ago, set start.of.time.period = 6 and end.of.time.period = 5
      focal <- focal %>% 
        filter(date > (start.of.time.period * -1), 
               date < (end.of.time.period * -1)) %>% 
        select(-date)

      # Proceed if the sample size is big enough
      if(nrow(focal) >= minimum.n){

        focal <- (focal %>% as.data.frame())[,1] # Get the gender column as a character vector
        focal <- focal[!str_detect(focal, "U")] # to keep it simple, let's only include papers where we know all the authors' genders

        # And if the sample size is STILL big enough...
        if(length(focal) >= minimum.n){

          # Make a summary of the number of M and F authors on each paper, and the gender of the first and last author
          n.char <- nchar(focal)
          counts <- data.frame(nMales = str_count(focal, "M"), 
                               nFemales = str_count(focal, "F"),
                               firstF = substr(focal, 1, 1) == "F",
                               lastF = substr(focal, n.char, n.char) == "F")

          rm(focal) # Explicitly discard 'focal', in case this helps to reduce memory footprint

          counter <- counter + 1 # Save the data and increment the counter
          count.data[[counter]] <- counts
          journal.names[counter] <- journals[i]
        }

        else rm(focal) # Explicitly discard 'focal' anyway
      } 
    } 
    names(count.data) <- journal.names[!is.na(journal.names)] # Name each entry in the list with the title of the journal the data are from

    saveRDS(count.data, file = file.name)
  }
}
```

### Define a function to calculate Bergstrom et al.'s 'coefficient of homophily', $\alpha$
These two functions estimate $\alpha$ for a set of papers' author lists. The version `find.alpha.CIs()` performs bootstrapping resampling of the dataset to obtain the 95% confidence limits on $\alpha$. `find.alpha()` calculates $\alpha$ without bootstrapping, making it much faster (used later when we are calculating the null distribution of $\alpha$).

$\alpha$ is defined as _p_ - _q_, and is termed "the coefficient of homophily". _p_ is the probability that a randomly-chosen co-author of a **male** author is a man, and _q_ is the probability that a randomly-chosen co-author of a **female** author is a man. Thus, positive $\alpha$ means that men coauthor with men, and women coauthor with women, more often than expected under the null model that coauthorship are random with respect to gender. Negative $\alpha$ means the opposite: men and women coauthor more often than expected.

For more information about $\alpha$, see this working paper by Bergstrom _et al._ http://eigenfactor.org/gender/assortativity/measuring_homophily.pdf. Also, see the comments section of this blog post: http://www.michaeleisen.org/blog/?p=1931. Bergstrom _et al._ point out that $\alpha$ is equivalent to the Peason correlation coefficient between the genders of authors on a paper.

```{r}
# The objects nMales and nFemales should give the number of male and female authors on a set of papers
# The function is written assuming that all papers have two or more authors of known gender. 
find.alpha <- function(nMales, nFemales){

  # Number of authors per paper, minus one
  nAuthors <- nMales + nFemales - 1

  # If you were to pick a SECOND random author from each paper, 
  # after picking a MALE the first time, what is the probability that the author is male?
  p.second.pick.is.male.M <- (nMales - 1) / nAuthors

  # If you were to pick a SECOND random author from each paper, 
  # after picking a FEMALE the first time, what is the probability that the author is male?
  p.second.pick.is.male.F <- nMales / nAuthors

  # If the probability of picking a male was zero (or one) the first time, 
  # make sure it's still zero (or one) for the second pick
  p.second.pick.is.male.M[nMales == 0] <- 0 
  p.second.pick.is.male.F[nMales == 0] <- 0 
  p.second.pick.is.male.M[nFemales == 0] <- 1 
  p.second.pick.is.male.F[nFemales == 0] <- 1

  # Find average proportion of male authors across all the men (p),
  # and average proportion of male authors across all the women (q). Subtract to get alpha
  return(sum(nMales * p.second.pick.is.male.M / sum(nMales)) - 
         sum(nFemales * p.second.pick.is.male.F / sum(nFemales)))
}

# This version uses bootstrapping to also estimate the 95% confidence limits on alpha
# Note that papers with lots of authors are more likely to be picked, ensuring that 
# each individual author is equally likely to end up in the sample. So if
# 2-author vs 10-author papers have different amounts of homophilly, or 
# different gender ratios, the CIs should not be biased
find.alpha.CIs <- function(nMales, nFemales, nBoots, correction = NULL){

  counts <- data.frame(nMales = nMales, nFemales = nFemales)
  num.papers <- nrow(counts)
  num.authors <- rowSums(counts)

  # Resample 'num.papers' papers with replacement, 'nBoots' times. 
  # Papers with lots of authors are more likely to be picked
  boot.alpha <- counts %>% 
    sample_n(size = num.papers * nBoots,   
             replace = TRUE,
             weight = num.authors) %>%
    mutate(replicate = rep(1:nBoots, each = num.papers)) %>%
    group_by(replicate) %>%     # Calculate alpha on each random sample
    summarise(alpha = find.alpha(nMales, nFemales)) %>% .$alpha

  # Optionally, apply a correction to get alpha' (e.g. if alpha is expected to 
  # be -0.05 under the null, add 0.05 to get alpha')
  if(!is.null(correction)){
    boot.alpha <- boot.alpha - correction
    alpha <- find.alpha(nMales, nFemales) - correction
  }

  else alpha <- find.alpha(nMales, nFemales)

  # If there are any re-samples entirely male/female (usually v rare), they generate NaNs. Remove them
  boot.alpha <- boot.alpha[!is.nan(boot.alpha)]

  c(alpha, as.numeric(quantile(boot.alpha, probs = c(0.025, 0.975)))) 
}

# This version calculates the difference in prop. male co-authors between male and female first (or last) authors
# The is.female column is boolean, and identifies whether each paper's first (or last) author is female
find.alpha.specific.author <- function(nMales, nFemales, is.female){

  # Find the proportion of male coauthors for each first (or last) author
  nMales[!is.female] <- nMales[!is.female] - 1 # If first/last author is male, don't count him among coauthors
  nFemales[is.female] <- nFemales[is.female] - 1 # If first/last author is female, don't count her among coauthors
  prop.male <- nMales / (nMales + nFemales) # proportion of male coauthors

  return(sum(prop.male[!is.female]) / length(prop.male[!is.female]) - 
         sum(prop.male[is.female])  / length(prop.male[is.female]))
}

# Same, but it also gets the 95% CIs by bootstrapping
# Here, the first authors (or last authors) are the unit of replication, and there is always 1 per paper.
# So, no need to sample papers with lots of authors more frequently
find.alpha.specific.author.CIs <- function(nMales, nFemales, 
                                           nBoots, is.female, 
                                           correction = NULL){

  counts <- data.frame(nMales = nMales, 
                       nFemales = nFemales, 
                       is.female = is.female)
  num.papers <- nrow(counts)

  boot.alpha <- counts %>% 
    sample_n(size = num.papers * nBoots,   
             replace = TRUE) %>%
    mutate(replicate = rep(1:nBoots, each = num.papers)) %>%
    group_by(replicate) %>%  # Calculate alpha on each random sample
    summarise(alpha = find.alpha.specific.author(nMales, nFemales, is.female)) %>% .$alpha

  # Optionally, apply a correction to get alpha'
  if(!is.null(correction)){
    boot.alpha <- boot.alpha - correction
    alpha <- find.alpha.specific.author(nMales, nFemales, is.female) - correction
  }

  else alpha <- find.alpha.specific.author(nMales, nFemales, is.female)

  # If any re-samples chanced to be entirely male/female, they generate NaNs. Remove them
  boot.alpha <- boot.alpha[!is.nan(boot.alpha)]

  c(alpha, as.numeric(quantile(boot.alpha, probs = c(0.025, 0.975)))) 
}
```

### Define function to calculate $\alpha$ for a given journal
For each journal, the function calculates $\alpha$, and its 95% confidence limits. It also generates a two-tailed p-value for the observed $\alpha$, under the null hypothesis that authors assort randomly with respect to gender. Small p-values mean that the observed set of papers shows significantly more, or less, gender homophily than expected under the null.

There is a good reason to test whether $\alpha$ is significantly greater (or lower) than expected under the null, rather than simply checking whether the 95% CIs on $\alpha$ overlap zero. The reason is that the null expectation for $\alpha$ is not centred zero for small datasets. This issue was summed up clearly by Prof. Carl Bergstrom in a comment on a blog post (http://www.michaeleisen.org/blog/?p=1931). Prof. Bergstrom wrote:

"Consider a tiny field with four authors, two men and two women, and one paper in each of the six two-author combinations. Now if you pick a man author at random, you are twice likely to pick the man in one of the four man-woman papers as you are to pick a man in the man-man paper. Therefore _p_ = 1/3. But if pick a randomly chosen woman, again you’re more likely to pick a man-woman combination than the one woman-woman combination, and _q_ = 2/3. As a result, $\alpha$ = -1/3 even though the authorships seem to be distributed without gender bias."

So, $\alpha$ is sensitive to the size of the dataset being tested (at least for very small datasets). Our solution is to simulate many fake datasets, where authors are distributed randomly with respect to gender, but the datasets are otherwise identical (in terms of numbers of authorships by men and women, the number of papers, and the distribution of authors per paper). This is what this function does, when calculating whether $\alpha$ is greater/lower than expected under random assortment by gender.

The function also returns a 'corrected' version of $\alpha$, which is simply $\alpha$ minus the null expected value of $\alpha$ for this particular dataset (calculated by our simulation). Because the null expected $\alpha$ is generally close to zero, the corrected version of $\alpha$ is usually almost identical. However, the corrected version can deviate a little (as much as 0.05) for the smallest datasets.

```{r}
# - if a vector is entered for is.female, then the test calculates alpha for a specific authorship position (i.e. first or last); otherwise, it calculates alpha for all authorship positions
# - the 'nYears' argument can be used to restrict the data to papers published nYears before the present. I chose nYears = 1, because if the gender ratio of authors changes over time (which it does), it will generate spurious coauthorship gender homophily 
# - 'minimum.n' refers to the number of papers - the function will quit without performing the test if we did not get enough papers that fit the criteria (i.e. papers published within NYears of the present, with at least 2 authors, where we know all authors' genders with >95% confidence)
gender.grouping.test.one.journal <- function(nMales, nFemales, is.female = NULL, journal, nBoots, nSimulations){
  print(journal)

  # First count number (and distribution) of authors that we will swap around
  total.males <- sum(nMales)
  total.females <- sum(nFemales)
  authors.per.paper <- nMales + nFemales

  # If we are looking at first or last authors, we will not be swapping them, so subtract accordingly
  if(!is.null(is.female)){ 
    num.male.focal.authors <- sum(!is.female) 
    num.female.focal.authors <- sum(is.female)
    # Quit unless there is a mixture of male and female first/last authors
    if(num.male.focal.authors == 0 | num.female.focal.authors == 0) return(NULL) 
    total.males <- total.males - num.male.focal.authors
    total.females <- total.females - num.female.focal.authors
    authors.per.paper <- authors.per.paper - 1
  }

  # Now, let's simulate the expected alpha under the null hypothesis that authors 
  # publish together randomly with respect to gender.
  # Simulate 'nSimulations' fake datasets, which have the same number of papers, 
  # and the same distribution of author list lengths per papers, as the real data,
  # but male and female authors are randomly grouped across papers.
  # The code here is ugly, but is vectorised for speed. Essentially I generate 
  # num.authors * nSimulations random numbers between zero and one, convert them 
  # to ranks, and use these ranks as index positions in 'pasted.gender'. This 
  # allows all the random numbers to be generated in one call, saving time. 
  # Then, it's just a matter of prettifying tapply's output using unname and unlist
  pasted.gender <- c(rep(1, total.males), rep(0, total.females))
  num.papers <- length(nMales)
  num.authors <- total.males + total.females
  simulated.data <- data.frame(
    gender = pasted.gender[unlist(unname(tapply(runif(num.authors * nSimulations), 
                                                rep(1:nSimulations, each = num.authors), rank)))],
    paperID = rep(unlist(mapply(rep, 1:num.papers, each = authors.per.paper)), nSimulations),
    replicate = rep(1:nSimulations, each = num.authors))
  rm(pasted.gender)

  # Calculate number of males and females on each simulated paper, in each replicate
  simulated.data <- simulated.data %>% 
    group_by(replicate, paperID) %>% 
    summarise(nFemales = sum(gender == 0), 
              nMales = sum(gender == 1)) %>% 
    select(-paperID)

  # Find alpha for each simulation replicate

  # If we are doing ALL the author positions, use find.alpha()
  if(is.null(is.female)){
    simulated.alpha.under.null <- simulated.data %>%
      group_by(replicate) %>%
      summarise(alpha = find.alpha(nMales, nFemales)) %>% .$alpha

    median.alpha.under.null <- median(simulated.alpha.under.null)

    # Calculate the observed alpha and its 95% CIs, both with and 
    # without adjusting for the fact that the null expected alpha can be non-zero
    observed.alpha <- find.alpha.CIs(nMales, nFemales, nBoots = nBoots)
    observed.alpha.adjusted <- find.alpha.CIs(nMales, 
                                              nFemales, nBoots = nBoots, 
                                              correction = median.alpha.under.null)
  }

  else{ # Otherwise, use find.alpha.specific.author()
    simulated.data$is.female <- rep(is.female, nSimulations) # Add the first/last authors back in
    simulated.data$nMales[!simulated.data$is.female] <- 
      simulated.data$nMales[!simulated.data$is.female] + 1
    simulated.data$nFemales[simulated.data$is.female] <- 
      simulated.data$nFemales[simulated.data$is.female] + 1

    simulated.alpha.under.null <- simulated.data %>%
      group_by(replicate) %>%
      summarise(alpha = find.alpha.specific.author(nMales, nFemales, is.female)) %>% .$alpha

    median.alpha.under.null <- median(simulated.alpha.under.null)

    observed.alpha <- find.alpha.specific.author.CIs(
      nMales, nFemales, is.female, nBoots = nBoots)
    observed.alpha.adjusted <- find.alpha.specific.author.CIs(
      nMales, nFemales, is.female, nBoots = nBoots, correction = median.alpha.under.null)
  }

  # The 2-tailed p-value is the proportion of null-simulated alpha values with 
  # a value at least as far from zero as the observed alpha
  # Thus, a significant p-value means that the observed alpha is larger OR smaller than expected
  two.tail.p.value <- sum(abs(simulated.alpha.under.null) >= 
                            abs(observed.alpha[1])) / length(simulated.alpha.under.null)

  # Return the pertinent results
  data.frame(journal = journal,
             n.useable.papers = num.papers,
             n.authors = num.authors,
             gender.ratio.of.sample = 100 * total.females /
               (total.females + total.males),
             observed.alpha = observed.alpha[1],
             lower.95.CI = observed.alpha[2],
             upper.95.CI = observed.alpha[3],
             alpha.adjusted = observed.alpha.adjusted[1],
             lower.95.CI2 = observed.alpha.adjusted[2],
             upper.95.CI2 = observed.alpha.adjusted[3],
             two.tail.p.value = two.tail.p.value)
}
```

### Define a function for parallelisation
Runs `gender.grouping.test.one.journal()` on all the journals in 'journals', in parallel across multiple CPUs.
```{r}
gender.grouping.test.many.journals <- function(count.data, authors, output.file, nBoots, nSimulations, number.of.cores, over.write = FALSE){

  # Only do the test if over.write = TRUE or there is no datafile in the working directory
  if(over.write | !(output.file %in% list.files())){ 
    journals <- names(count.data)
    nJournals <- length(count.data)
    # Set up a cluster
    cl <- makePSOCKcluster(number.of.cores)
    setDefaultCluster(cl)
    clusterExport(NULL, c("find.alpha",
                          "find.alpha.CIs", 
                          "find.alpha.specific.author",
                          "find.alpha.specific.author.CIs",
                          "gender.grouping.test.one.journal",
                          "pubmed_sqlite"))
    clusterEvalQ(NULL, library(dplyr))

    # Run the desired test on each journal, in parallel
    if(authors == "all"){
      results <- parLapply(NULL, 1:nJournals, function(i) 
        gender.grouping.test.one.journal(count.data[[journals[i]]]$nMales,
                                         count.data[[journals[i]]]$nFemales,
                                         NULL,
                                         journal = journals[i],
                                         nBoots = nBoots,
                                         nSimulations = nSimulations))
    }

    else if(authors == "first"){
      results <- parLapply(NULL, 1:nJournals, function(i) 
        gender.grouping.test.one.journal(count.data[[journals[i]]]$nMales,
                                         count.data[[journals[i]]]$nFemales,
                                         count.data[[journals[i]]]$firstF,
                                         journal = journals[i],
                                         nBoots = nBoots,
                                         nSimulations = nSimulations))
    }

    else if(authors == "last"){
      results <- parLapply(NULL, 1:nJournals, function(i) 
        gender.grouping.test.one.journal(count.data[[journals[i]]]$nMales,
                                         count.data[[journals[i]]]$nFemales,
                                         count.data[[journals[i]]]$lastF,
                                         journal = journals[i],
                                         nBoots = nBoots,
                                         nSimulations = nSimulations))
    }

    setDefaultCluster(NULL) # close cluster and write the results to disk
    stopCluster(cl = cl)
    write.csv(do.call("rbind", results), file = output.file, row.names = FALSE)
  }
}
```

## Sample size information

**Table S`r supp.table.counter`**: Sample sizes for the two datasets, which comprise papers published in the timeframes August 2005 - August 2006, and August 2015 - August 2016.

```{r echo = FALSE, results = "hide"}
supp.table.counter <- supp.table.counter + 1
```

```{r}
minimum.n <- 50   # Only include journals with at least 50 papers in the specified range of time
unique.journals <- pubmed_sqlite %>% 
  select(journal) %>% distinct() %>% 
  collect(n = Inf) %>% .$journal # Get a list of all the journals in the database

# Get all the data we need out of the database, and save as an RDS file - for papers in the last 12 months
retrieve.authorship.data(unique.journals, 
                         minimum.n = minimum.n, 
                         start.of.time.period = 1, 
                         end.of.time.period = 0, 
                         file.name = "last.year.rds", 
                         over.write = FALSE)

# Get all the data we need out of the database, and save as an RDS file - for papers that are between 10 and 11 years old
retrieve.authorship.data(unique.journals, 
                         minimum.n = minimum.n, 
                         start.of.time.period = 11, 
                         end.of.time.period = 10, 
                         file.name = "older.papers.rds", 
                         over.write = FALSE)

# Get metadata about each journal (research discipline, ISI impact factor, and the country that publishes it)
# see Holman et al. in prep, for information on how journals were assigned to disciplines
journal.data <- journals_sqlite %>% collect(n=Inf) %>% 
  rename(journal = short.title, 
         discipline = Discipline, 
         country = journal.country) %>% 
  select(journal, discipline, country, IF) %>% 
  as.data.frame %>% 
  distinct(journal, .keep_all = T)

make.sample.size.table <- function(filename){

  count.data <- readRDS(paste(filename, ".rds", sep = "")) 
  data.frame(Quantity = c("Number of disciplines", 
                          "Number of journals", 
                          "Number of papers",
                          "Number of authors",
                          "Median number of papers per journal", 
                          "Median number of authors per journal",
                          "Median number of authors per paper"),
             Value = c(journal.data %>% filter(journal %in% names(count.data)) %>% 
                         select(discipline) %>% 
                         distinct() %>% nrow(),
                       length(count.data),
                       nrow(do.call("rbind", count.data)[,1:2]),
                       sum(do.call("rbind", count.data)[,1:2]),
                       median(sapply(count.data, nrow)),
                       median(sapply(count.data, function(x) sum(x[,1]) + sum(x[,2]))),
                       median(rowSums(do.call("rbind", count.data)[,1:2])))) 
}

make.sample.size.table("older.papers") %>% 
  rename(`n (2005-2006)` = Value) %>% 
  left_join(make.sample.size.table("last.year") %>% 
              rename(`n (2015-2016)` = Value), by = "Quantity") %>% 
    pander(split.cell = 40, split.table = Inf)
```

## Calculating $\alpha$'

### Run the test on every journal for which we have sufficient data
Here, 'sufficient data' means that there are at least 50 papers, for which there are 2+ authors, and we know all authors' genders, and these papers were published in the last year. We calculate $\alpha$ for three types of authors:

1. All authors; this alpha statistic is positive if the average man author has a co-author gender ratio that is more male-biased than the average woman author.
2. First authors; this alpha statistic is positive if the average man first author has a co-author gender ratio that is more male-biased than the average woman first author.
3. Last authors; this alpha statistic is positive if the average man last author has a co-author gender ratio that is more male-biased than the average woman last author.

```{r}
# Calculate alpha across all authorship positions for all suitable NEW papers
gender.grouping.test.many.journals(readRDS("last.year.rds"),
                                   "all",
                                   "homophily.results.all.csv",
                                   nBoots = 1000, 
                                   nSimulations = 1000, 
                                   number.of.cores = 8,
                                   over.write = FALSE)

# Calculate alpha for first authors for all suitable NEW papers
gender.grouping.test.many.journals(readRDS("last.year.rds"),
                                   "first",
                                   "homophily.results.first.authors.csv",
                                   nBoots = 1000, 
                                   nSimulations = 1000, 
                                   number.of.cores = 8,
                                   over.write = FALSE)

# Calculate alpha for last authors for all suitable NEW papers
gender.grouping.test.many.journals(readRDS("last.year.rds"),
                                   "last",
                                   "homophily.results.last.authors.csv",
                                   nBoots = 1000, 
                                   nSimulations = 1000, 
                                   number.of.cores = 8,
                                   over.write = FALSE)

# Calculate alpha across all authorship positions for all suitable OLD papers
gender.grouping.test.many.journals(readRDS("older.papers.rds"),
                                   "all",
                                   "older.homophily.results.all.csv",
                                   nBoots = 1000, 
                                   nSimulations = 1000, 
                                   number.of.cores = 8,
                                   over.write = FALSE)

# Calculate alpha for first authors for all suitable OLD papers
gender.grouping.test.many.journals(readRDS("older.papers.rds"),
                                   "first",
                                   "older.homophily.results.first.authors.csv",
                                   nBoots = 1000, 
                                   nSimulations = 1000, 
                                   number.of.cores = 8,
                                   over.write = FALSE)

# Calculate alpha for last authors for all suitable OLD papers
gender.grouping.test.many.journals(readRDS("older.papers.rds"),
                                   "last",
                                   "older.homophily.results.last.authors.csv",
                                   nBoots = 1000, 
                                   nSimulations = 1000, 
                                   number.of.cores = 8,
                                   over.write = FALSE)

```

### Add additional metadata to the $\alpha$ results
Here, we assign extra information to each journal - research discipline (see Holman _et al._ in prep), ISI impact factor (where available), and the country that publishes the journal (where known; information taken from PubMed).
```{r message=FALSE, warning=FALSE}
# Calculate the impact factor of each journal relative to other journals in its discipline (using residuals from a mixed model)
journal.data$logIF <- log10(journal.data$IF) # Take the log10 of the impact factor
journal.data$standardised.IF <- NA
# Calculate residual log10 IF for each journal (i.e. IF relative to the field that it is in)
journal.data$standardised.IF[!is.na(journal.data$logIF)] <- 
  resid(lmer(logIF ~ (1|discipline), data = journal.data[!is.na(journal.data$logIF), ]))

# Function to neatly merge the journal-specific data with the gender homophily results
# Also add another column of p-values, which are False Discovery Rate-corrected using B-H method
merge.and.clean <- function(homophily.filepath, journal.data){
  left_join(read.csv(homophily.filepath, stringsAsFactors = FALSE), journal.data, by = "journal") %>% 
    select(discipline, 
           journal, 
           standardised.IF, 
           country,
           n.useable.papers, 
           n.authors, 
           gender.ratio.of.sample, 
           observed.alpha, 
           lower.95.CI, 
           upper.95.CI, 
           alpha.adjusted, 
           lower.95.CI2, 
           upper.95.CI2, 
           two.tail.p.value) %>% 
    arrange(-n.useable.papers) %>% # Sort by sample size
    mutate(adjusted.p.value = p.adjust(two.tail.p.value, method = "BH")) 
}

# Load up all the homophily results, and merge them with the journal-specific information
gender.homophily.all <- merge.and.clean("homophily.results.all.csv", journal.data)
gender.homophily.first <- merge.and.clean("homophily.results.first.authors.csv", journal.data)
gender.homophily.last <- merge.and.clean("homophily.results.last.authors.csv", journal.data)

gender.homophily.all.old <- merge.and.clean("older.homophily.results.all.csv", journal.data)
gender.homophily.first.old <- merge.and.clean("older.homophily.results.first.authors.csv", journal.data)
gender.homophily.last.old <- merge.and.clean("older.homophily.results.last.authors.csv", journal.data)

# Bind all 5 datasets into a single one, with a column called 'Subset' to identify them
master.dataset <- rbind(gender.homophily.all,
                        gender.homophily.first,
                        gender.homophily.last,
                        gender.homophily.all.old,
                        gender.homophily.first.old,
                        gender.homophily.last.old) %>%
  mutate(Subset = unlist(mapply(rep, c("All authors", "First authors", "Last authors"), 
                                each = c(nrow(gender.homophily.all), 
                                         nrow(gender.homophily.first),
                                         nrow(gender.homophily.last),
                                         nrow(gender.homophily.all.old),
                                         nrow(gender.homophily.first.old),
                                         nrow(gender.homophily.last.old)))),
         Age = unlist(mapply(rep, c("New", "Old"), 
                                each = c(sum(nrow(gender.homophily.all), 
                                             nrow(gender.homophily.first), 
                                             nrow(gender.homophily.last)),
                                         sum(nrow(gender.homophily.all.old), 
                                             nrow(gender.homophily.first.old), 
                                             nrow(gender.homophily.last.old)))))) %>%
  select(discipline, journal, Subset, Age, standardised.IF, n.useable.papers, n.authors, 
         gender.ratio.of.sample, observed.alpha, lower.95.CI, upper.95.CI, alpha.adjusted,
         lower.95.CI2, upper.95.CI2, two.tail.p.value, adjusted.p.value) %>%
  rename(unadjusted.p.value = two.tail.p.value)

# Save and export a neat and tidy version for the supplementary material
neat <- master.dataset
names(neat) <- gsub("[.]", " ", names(neat))
names(neat) <- gsub("Subset", "Author position", names(neat))
names(neat) <- paste(toupper(substring(names(neat), 1, 1)), substring(names(neat), 2), sep="")
names(neat) <- gsub("N", "n", names(neat))
names(neat) <- gsub("2", "", names(neat))
for(i in which(sapply(neat, is.numeric))) neat[,i] <- round(neat[,i], 3)
write.csv(neat, file = "Supplementary dataset S1.csv", row.names = FALSE)
rm(neat)
```

### Calculating gender homophily for each combination of author country and journal
One reason that we might see 'spurious' homophily is that the gender ratio of resaerchers varies across countries. For example, in our dartaset, most Serbian author are women and most Japanese authors are men. Because there are realtive few collaborations between people from these countries (relative to within country-collaborations), we will tend to see more msotly-male and mostly-female papers than expected even if people assort randomly within countries. So, if we split the data by country of author affiliation, and we still see homophily, this suggests that variation across countries cannot fully explain our results.

The following code restricts our dataset to only include papers where we know the country of affiliation for all authors on the paper, and all authors are from the same country (this is much simpler than trying to account for multi-country papers). Then, for each journal-country combination where we have at least 50 papers, we calculate the coefficient of homophily just as before.

```{r}
gender.grouping.test.by.journal.and.country <- function(){

  country.data <- pubmed_sqlite %>% # Go to the big database
    select(pmid, date, gender.95, country) %>%   # Get the pubmed ID, gender, and author country columns
    filter(!is.na(country) & country != "NA") %>% # Throw out papers with no country
    collect(n=Inf) %>%                        # Bring the data into memory
    filter(!grepl("NA", country),             # Throw out papers with no country (again)
           !grepl("U", gender.95),            # Throw out papers where we don't know all author genders
           nchar(gender.95) > 1) %>%          # Throw out papers with only one author
    mutate(date = convert.dates(date)) %>%    # Convert the dates to years-before-present
    filter(date > -1) %>% select(-date)       # Only keep papers from August 2015 - August 2016
  unique.split.countries <- lapply(strsplit(country.data$country, 
                                            split = "_"), unique)
  to.include <- sapply(unique.split.countries, length) == 1
  country.data <- data.frame(pmid = country.data$pmid[to.include],
                             gender.95 = country.data$gender.95[to.include],
                             country = unlist(unique.split.countries[which(to.include)]))
  rm(list = c("unique.split.countries", "to.include")) # free up memory

  # Merge in information on the journal and discipline of all these suitable papers
  country.data <- country.data %>%          
    left_join(pubmed_sqlite %>% select(pmid, journal, discipline) %>%
                collect(n=Inf), by = "pmid") %>% 
    select(-pmid)

  # Restrict the dataset to only journal-country combinations 
  # for which we have at least 'minimum.n' papers
  country.data <- country.data %>% 
    group_by(journal, country) %>% 
    summarise(n.papers = n()) %>% 
    filter(n.papers >= minimum.n) %>% select(-n.papers) %>% 
    left_join(country.data, by = c("journal", "country")) %>%
    mutate(gender.95 = as.character(gender.95))

  country.data$nMales <- str_count(country.data$gender.95, "M")
  country.data$nFemales <- str_count(country.data$gender.95, "F")

  # First find all the combinations of country and journal
  combinations <- (country.data %>% group_by(journal, country) %>% 
                     summarise(n=n()))[,1:2] %>% as.data.frame()
  # Now loop over them, using multiple cores
  output <- do.call("rbind", mclapply(1:nrow(combinations), function(i){
    focal <- country.data %>% filter(country == combinations$country[i], 
                                     journal == combinations$journal[i])
    data.frame(country = combinations$country[i],
               journal = combinations$journal[i],
               gender.grouping.test.one.journal(focal$nMales, 
                                                focal$nFemales, 
                                                is.female = NULL, 
                                                "Doing one", 
                                                nBoots=1000, 
                                                nSimulations=1000)[,-1]
    )}))

  # Function to fix capitalisation of country names
  capitalise.countries <- function(countries){ 
    countries <- as.character(
      sapply(as.character(countries), function(x) {
               s <- strsplit(x, " ")[[1]]
               paste(toupper(substring(s, 1, 1)), substring(s, 2), sep="", collapse=" ")
             })) 
    str_replace(str_replace(countries, " And ", " and "), "Usa", "USA")
  }

  output$country <- capitalise.countries(output$country)
  output
}

if(!("country.data.csv" %in% list.files())){ # Write the data, assuming it's not already been done
  country.data <- gender.grouping.test.by.journal.and.country(country.data)
  write.csv(country.data, file = "country.data.csv", row.names = FALSE)
}

country.data <- read.csv("country.data.csv", stringsAsFactors = FALSE) %>%
  mutate(p.adjusted = p.adjust(two.tail.p.value, method = "BH")) %>%
  left_join(journal.data %>% select(journal, discipline), by = "journal") # %>% 
  # left_join(master.dataset %>% 
  #             filter(Age == "New" & Subset == "All authors") %>%
  #             select(journal, alpha.adjusted) %>%
  #             rename(journal.alpha = alpha.adjusted),
  #           by = "journal")
```

### Checking that adjusted and un-adjusted $\alpha$ are very similar
```{r}
master.dataset %>%
  ggplot(aes(x = observed.alpha, y = alpha.adjusted)) + 
  geom_abline(intercept = 0, slope = 1, colour = "darkgrey") + 
    geom_point(size=0.6, alpha=0.4) + 
  ylab("Adjusted measure of homophily (\u03B1')") + 
  xlab("Unadjusted measure of homophily (\u03B1)") + 
  theme_minimal()
```
<br></br> 
**Figure S`r supp.figure.counter`**: There is a very strong correlation between the values of $\alpha$ and $\alpha'$ calculated for each journal, though in a handful of cases the difference is considerable. The deviation between $\alpha$ and $\alpha'$ is greatest for journals for which we have a small sample size (see next figure).

```{r echo=FALSE, results='hide'}
supp.figure.counter <- supp.figure.counter + 1
```

### Un-adjusted $\alpha$ is downwardly biased for small datasets
As expected, unadjusted $\alpha$ is slightly downwardly biased for small datasets, because the expected value of $\alpha$ under the null is less than 0. Thus, our adjusted $\alpha'$ statistic, which is the observed $\alpha$ minus the value of $\alpha$ expected under the null, tends to be substantially higher than $\alpha$ for small datasets. All of our results do not differ qualitatively if we use the unadjusted $\alpha$, but we have elected to present the results for $\alpha'$ in light of the issues with unadjusted $\alpha$ for small datasets.
```{r}
master.dataset %>%
  filter(Age == "New") %>%
  ggplot(aes(x = log10(n.useable.papers), y = alpha.adjusted - observed.alpha)) + 
  geom_point(alpha = 0.3, size = 0.5) + 
  facet_wrap(~Subset) +
  ylab("\u03B1' - \u03B1") + xlab("Log10 sample size\n(number of papers)") + 
  theme_minimal()
```
<br></br> 
**Figure S`r supp.figure.counter`**: For journals for which we recovered a small number of papers (<100), the unadjusted metric $\alpha$ was downwardly biased. This fits our expectations: because authors cannot be their own co-authors, small datasets will tend to produce negative estimates of $\alpha$ even if authors assort randomly with respect to gender (see main text). This suggests that $\alpha'$ is a more useful measure of homophily and heterophily, especially for small samples.

```{r echo=FALSE, results='hide'}
supp.figure.counter <- supp.figure.counter + 1
```

## Results

### Distribution of gender homophily and heterophily across journals
Most journals show gender homophily (i.e. $\alpha' > 0$), not heterophily, meaning that people are more likely to co-publish with colleagues of the same gender.

```{r fig.width=5, fig.height=8}
figure2 <- master.dataset %>%
  filter(Age == "New") %>% 
  mutate(Significant = 1 * (adjusted.p.value < 0.05),
         Significant = replace(Significant, Significant == 0, "No"),
         Significant = replace(Significant, Significant == 1, "Yes")) 
n.sig.homo <- sum(figure2$Significant[figure2$Subset == "All authors"] == "Yes" & 
                    figure2$alpha.adjusted[figure2$Subset == "All authors"] > 0)
n.sig.hetero <- sum(figure2$Significant[figure2$Subset == "All authors"] == "Yes" & 
                    figure2$alpha.adjusted[figure2$Subset == "All authors"] < 0)
n.journals <- sum(figure2$Subset == "All authors")

figure2 <- figure2 %>%
  ggplot(aes(x = alpha.adjusted, fill = Significant)) + 
  geom_vline(xintercept = 0, linetype = 2) +
  geom_histogram(binwidth = 0.01, colour = "black") + 
  scale_fill_manual(values = c("skyblue", "white")) +
  xlab("Coefficient of homophily (\u03B1')") + 
  ylab("Number of journals") +
  facet_wrap(~Subset, scales = "free_y", ncol=1) + 
  theme_minimal(12) + 
  theme(legend.position = "top")
ggsave(plot = figure2, filename = "figures/figure2.pdf",  width = 5, height = 8, device = cairo_pdf)
figure2
```
<br></br> 
**Figure `r figure.counter`**: Of the `r n.journals` journals for which we had adequate data in 2015-2016, `r n.sig.homo` showed statistically significant evidence of homophily (denoted by $\alpha' > 0$), and `r n.sig.hetero` showed statistically significant evidence of heterophily ($\alpha' < 0$), after adjusting p-values using Benjamini-Hochberg false discovery rate correction. The white area shows the number of journals for which homophily was significantly stronger than expected under the null hypothesis (p < 0.05), while the blue area shows all the remainder. Patterns were similar whether $\alpha'$ was calculated for all authors, for first authors only, or for last authors only.

```{r echo=FALSE, results='hide'}
figure.counter <- figure.counter + 1

# Save these figures for manuscript too:
saveRDS(c("n.journals" = n.journals, "n.sig.homo" = n.sig.homo, "n.sig.hetero" = n.sig.hetero), file ="manuscript/fig2_numbers.RData")
```

**Table S`r supp.table.counter`**: Number of journals showing statistically significant homophily or heterophily, in two one-year periods. The significance threshold was p < 0.05, and p-values were adjused using Benjamini-Hochberg false discovery rate correction. Note that the power of our test is lower for the 2005-2006 data because fewer papers were recovered per journal: thus, it is not meaningful to compare the % significant journals (i.e. 11% vs 24%) between the two time periods.

```{r echo = FALSE, results = "hide"}
supp.table.counter <- supp.table.counter + 1
```

```{r}
data.frame(Year = c("2015-2016", "2005-2006"), 
           rbind(master.dataset %>% 
                   filter(adjusted.p.value < 0.05, Age == "New") %>% 
                   mutate(homophily = as.logical(ceiling(alpha.adjusted))) %>% 
                   .$homophily %>% table() %>% unname(),
                 master.dataset %>% 
                   filter(adjusted.p.value < 0.05, Age == "Old") %>% 
                   mutate(homophily = as.logical(ceiling(alpha.adjusted))) %>% 
                   .$homophily %>% table() %>% unname()), 
           Total.n = c(sum(master.dataset$Age == "New" & master.dataset$Subset == "All authors"), 
                       sum(master.dataset$Age == "Old" & master.dataset$Subset == "All authors"))) %>% 
  rename(`Significant heterophily` = X1, `Significant homophily` = X2, `Year range` = Year,
         `Total n journals` = Total.n) %>%
  pander(split.cell = 40, split.table = Inf)
```

### Gender homophily has increased slightly over the last ten years
Although the effect is weak, we show here that the average $\alpha$ across journals was lower in 2005-2006 than it was in 2015-2016. One possible reason is that researchers have become more likely to preferentially select same-gendered colleagues over time. Another hypothesis is an increase in the Wahlund effect with time: for example, journals might have become more multi-disciplinary, or started publishing papers from a more diverse range of countries.

#### Figure showing the average change in $\alpha$ across journals
```{r}
# For each journal where we have observations of alpha' at
# both time points, calculate the change in alpha'
old.vs.new <- master.dataset %>% 
  filter(Subset == "All authors") %>%
  group_by(journal) %>%
  summarise(alpha.old = alpha.adjusted[2], 
            alpha.new = alpha.adjusted[1],
            SE.old = (upper.95.CI2[2] - lower.95.CI2[2]) / 2 * 1.96,
            SE.new = (upper.95.CI2[1] - lower.95.CI2[1]) / 2 * 1.96,
            lower883.old = alpha.old - SE.old * 1.190,
            upper883.old = alpha.old + SE.old * 1.190,
            lower883.new = alpha.new - SE.new * 1.190,
            upper883.new = alpha.new + SE.new * 1.190,
            Significant = "No",
            Significant = replace(Significant, lower883.new > upper883.old, "Yes"),
            Significant = replace(Significant, lower883.old > upper883.new, "Yes")) %>%
  filter(!is.na(alpha.old))

old.vs.new %>%
  ggplot(aes(x = alpha.new - alpha.old, fill = Significant)) + 
  geom_vline(xintercept = 0, linetype = 2) +
  geom_histogram(binwidth = 0.01, colour = "black", fill = "skyblue") + 
  scale_fill_manual(values = c("skyblue", "white")) +
  xlab("Change in \u03B1' over the last ten years") + 
  ylab("Number of journals") +
  theme_minimal()
```
<br></br>
**Figure S`r supp.figure.counter`**: Histogram showing the distribution of differences in $\alpha'$ between the 2015-16 and 2005-6 samples, where positive numbers indicate an increase in $\alpha'$ with time. The mean is slightly greater than 0, indicating a mild increase in average $\alpha'$ with time.

```{r echo=FALSE, results='hide'}
supp.figure.counter <- supp.figure.counter + 1
```

#### Model of the average change in $\alpha$
Journals had slightly lower values of $\alpha'$ ten years ago than they do today. The difference in mean $\alpha'$ values has an effect size of Cohen's d = 0.09, which is considered small. The effect is statistically significant, since we have a large sample of journals. The model includes 'journal' and 'discipline' as random effects, preventing pseudoreplication caused by measuring the same journals twice (i.e. once in each time period), and allowing us to test whether the discipline of the journal explains any variance in $\alpha'$. Both random factors explain only around 1% of the variance. 
```{r}
summary(lmer(scale(alpha.adjusted) ~ Age + (1|journal) + (1|discipline), # Note that the response variable gets scaled, to make the effect size interpretable as Cohen's d
             weights = 1 / ((upper.95.CI2 - lower.95.CI2)/2*1.96), # Weight each observation of alpha by its precision
             data = master.dataset %>% 
               filter(Subset == "All authors", journal %in% old.vs.new$journal)))
```

### Gender homophily is higher for first authors than last authors
One might expect $\alpha'$ to be stronger among last authors if homophily arises because senior scientists (who tend to be last authors) preferentially select or attract junior colleagues (e.g. graduate students and postdocs) of the same gender.

Instead, we found that $\alpha'$ was slightly higher (Cohen's d = 0.06) when calculated for first authors rather than for last authors, across journals (using the 2015-2016 dataset). Thus, our results imply that first authors tend to be the ones driving the observed gender homophily.

#### Model of the difference in $\alpha'$ for first and last authors
```{r showWarnings = FALSE}
first.vs.last.data <- master.dataset %>%
  filter(Subset != "All authors", Age == "New")

summary(lmer(scale(alpha.adjusted) ~ Subset + (1|journal) + (1|discipline), 
             weights = 1 / ((upper.95.CI2 - lower.95.CI2)/2*1.96),
             data = first.vs.last.data))
```

#### Plot of the difference in $\alpha'$ for first and last authors
```{r}
first.vs.last.data <- first.vs.last.data %>%
  select(journal, Subset, alpha.adjusted) %>%
  spread(Subset, alpha.adjusted) %>%
  mutate(diff = `First authors` -  `Last authors`) %>%
  filter(!is.na(diff))

first.vs.last.plot <- first.vs.last.data %>%
  ggplot(aes(x = diff)) +
  geom_vline(xintercept = 0, linetype = 2) +
  geom_histogram(binwidth = 0.02, colour = "black", fill = "skyblue") +
  xlab("\u03B1' for first authors - \u03B1' for last authors") +
  ylab("Number of journals")

first.vs.last.plot
```
<br></br>
**Figure S`r supp.figure.counter`**: Histogram showing the difference between $\alpha'$ calculated for first and last authors. Positive values mean that $\alpha'$ was higher when calculated for first authors, and negative values mean $\alpha'$ was higher when calculated for last authors. The mean is very slightly higher than zero, indicated that $\alpha'$ tends to be higher for first authors.

```{r echo=FALSE, results='hide'}
supp.figure.counter <- supp.figure.counter + 1
```

### Variance in gender homophily across disciplines
When we compared the $\alpha'$ values of journals from different research disciplines, we found that the strength of homophily varied between disciplines, but not a lot. That is, the variance within disciplines was substantially greater than the variance between disciplines (see above linear mixed model: discipline explains about 4% of the variance in $\alpha'$).

#### Figure showing the variance in $\alpha'$ between disciplines
```{r fig.width=8, fig.height=12}
# There is plenty of homophily, and basically no heterophily at all. 
# The disciplines that contain more diverse topics (e.g. "Multidisciplinary") show most homophily, as expected. 
# Still, there is homophily in many comparatively narrow disciplines too.
disc.summary <- master.dataset %>% 
  filter(!is.na(alpha.adjusted), 
         Age == "New", 
         Subset == "All authors") %>% 
  group_by(discipline) %>% 
  summarise(Mean.gender.ratio = mean(gender.ratio.of.sample) %>% round(1),
            Mean.adjusted.alpha = mean(alpha.adjusted) %>% round(3),
            SE.of.adjusted.alpha = (sd(alpha.adjusted)/sqrt(length(alpha.adjusted))) %>% round(3), 
            Median.adjusted.alpha = median(alpha.adjusted) %>% round(3), 
            n.journals = n(),
            n.significant.homophily = sum(alpha.adjusted > 0 & unadjusted.p.value < 0.05), 
            n.significant.heterophily = sum(alpha.adjusted < 0 & unadjusted.p.value < 0.05),
            n.FDR.significant.homophily = sum(alpha.adjusted > 0 & adjusted.p.value < 0.05), 
            n.FDR.significant.heterophily = sum(alpha.adjusted < 0 & adjusted.p.value < 0.05)) %>% 
  arrange(-Mean.adjusted.alpha) %>% 
  mutate(n.significant.homophily = paste(n.significant.homophily, 
                                         " (", n.FDR.significant.homophily, ")", sep = ""),
         n.significant.heterophily = paste(n.significant.heterophily, 
                                           " (", n.FDR.significant.heterophily, ")", sep = ""),
         discipline = factor(discipline, levels = discipline)) %>%
  select(-n.FDR.significant.homophily, -n.FDR.significant.heterophily) %>% 
  rename(Discipline = discipline) %>%
  as.data.frame()

disc.summary$n.significant.homophily[disc.summary$n.significant.homophily == "0 (0)"] <- 0
disc.summary$n.significant.heterophily[disc.summary$n.significant.heterophily == "0 (0)"] <- 0
names(disc.summary) <- gsub("[.]", " ", names(disc.summary))

gender.homophily.all %>% 
  filter(discipline %in% disc.summary$Discipline[disc.summary$`n journals` > 10]) %>%
  mutate(Discipline = factor(discipline, levels = rev(disc.summary$Discipline))) %>%
  ggplot(aes(x = Discipline, y = alpha.adjusted, fill = Discipline)) +
  geom_hline(yintercept = 0, linetype = 2) + 
  geom_boxplot(outlier.size = 0.5, outlier.shape = NA) + 
  geom_quasirandom(aes(colour = Discipline), alpha = 0.3, bandwidth = 0.25) + 
  ylab("Coefficient of homophily (\u03B1')") + 
  coord_flip() + 
  theme_minimal() + 
  theme(legend.position = "none")
```
<br></br>
**Figure `r figure.counter`**: Distribution of the $\alpha'$ values across journals, split by the research discipline of the journal. The points show the $\alpha'$ for each journal, and the boxes show quartiles 1, 2 and 3. The boxplot whiskers show the distance of the farthest point that is within 1.5 times the interquartile range from the edge of the box.

```{r echo=FALSE, results='hide'}
figure.counter <- figure.counter + 1
```

#### Table showing the difference in $\alpha$ between disciplines

**Table S`r supp.table.counter`**: The table shows similar information to Figure S`r supp.figure.counter`, namely the distribution of the $\alpha'$ values across journals, split by the research discipline. The gender ratio column shows the percentage of women authors in the sample used to calculate $\alpha'$, across all authorship positions. In the last two columns, the numbers outside parentheses give the number of journals that deviate statistically significantly from zero, while the numbers inside parentheses give the number that remain significant after false discovery rate correction.

```{r echo=FALSE, results='hide'}
supp.table.counter <- supp.table.counter + 1
```

```{r}
disc.summary %>% filter(`n journals` > 9) %>% 
  pander(split.cell = 40, split.table = Inf)
```

### Gender homophily covaries with author gender ratio
Journals with a male-biased author gender ratio showed less author gender homophily, while journals with an even or female-biased author gender ratio showed more homophily. The pattern was similar whether we calculated $\alpha$ for all authors, first authors only, or last authors only. 
```{r fig.width=6, fig.height=5}
gender.ratio <- master.dataset %>% 
  filter(Age == "New") %>% 
  ggplot(aes(x = (gender.ratio.of.sample - 50), y = alpha.adjusted)) + 
  geom_hline(yintercept = 0, linetype = 2) +
  geom_vline(xintercept = 0, linetype = 2) +
  geom_point(alpha = 0.4, size = 0.7) + 
  stat_smooth(method = "gam", formula = y ~ s(x, k = 3), colour = "black", fill = "skyblue") + 
  xlab("Deviation from gender parity for all authors in this journal") + 
  ylab("Coefficient of homophily (\u03B1')") +
  facet_wrap(~Subset) + 
  theme_minimal() +
  theme(legend.position = "none") 
ggsave(plot = gender.ratio, filename = "figures/figure3.pdf", device = cairo_pdf, width=7, height=5)
gender.ratio
```
<br></br>
**Figure `r figure.counter`**: There is a weakly positive, non-linear relationship between the gender ratio of authors publishing in a journal, and the coefficient of homophily ($\alpha'$). Specifically, journals with 50% women authors or higher tended to have more same-sex coauthorships than did journals with predominantly men authors.  This relationship held whether $\alpha'$ was calculated for all authors, first authors only, or last authors only. A negative value on the x axis denotes an excess of men authors, a positive value denotes an excess of women authors, and zero denotes gender parity. The lines were fitted using generalised additive models with the smoothing parameter $k$ set to 3.

```{r echo=FALSE, results='hide'}
figure.counter <- figure.counter + 1
```

Here are the results of three generalised additive models (GAMs) fit to the data in the three panels of Figure `r figure.counter - 1`:
```{r}
gam(alpha.adjusted ~ s(gender.ratio.of.sample), data = master.dataset %>% 
      filter(Age == "New", Subset == "All authors") %>% 
      mutate(gender.ratio.of.sample = gender.ratio.of.sample - 50)) %>% summary()
gam(alpha.adjusted ~ s(gender.ratio.of.sample), data = master.dataset %>% 
      filter(Age == "New", Subset == "First authors") %>% 
      mutate(gender.ratio.of.sample = gender.ratio.of.sample - 50)) %>% summary()
gam(alpha.adjusted ~ s(gender.ratio.of.sample), data = master.dataset %>% 
      filter(Age == "New", Subset == "Last authors") %>% 
      mutate(gender.ratio.of.sample = gender.ratio.of.sample - 50)) %>% summary()
```

### Gender homophily is negatively correlated with journal impact factor

After correcting for differences in journal impact factor between disciplines (using residuals from a mixed model), there was a negative relationship between impact factor and the coefficient of homophily, $\alpha$. This means that journals that have a high impact factor, relative to journals in their discipline, show less gender homophily than do low-impact journals.

```{r fig.width=6, fig.height=5}
model <- summary(lm(alpha.adjusted ~ standardised.IF, data = master.dataset %>% 
                      filter(Age == "New", 
                             Subset == "All authors", 
                             !is.na(standardised.IF))))
R2 <- model$adj.r.squared %>% 
        round(3) %>% format(nsmall = 3)

model$coefficients %>% pander(split.cell = 40, split.table = Inf) # The model coefficients and p-values

gender.homophily.all %>% 
  filter(!is.na(standardised.IF)) %>% 
  ggplot(aes(x = standardised.IF, y = alpha.adjusted)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_vline(xintercept = 0, linetype = 2) +
  stat_smooth(method = "lm", colour = "black", fill = "skyblue") +
  geom_point(size = 0.5, alpha = 0.9) + 
  xlab("Standardised journal impact factor") + 
  ylab("Coefficient of homophily (\u03B1')") + 
  theme_minimal()
```
<br></br>
**Figure `r figure.counter`**: Journal impact factor (expressed relative to the average for the discipline) is negatively correlated with $\alpha'$. The relationship is weak ($R^2$ = `r R2`), but the results suggest that journals with strong homophily tend to have lower impact factors than journals with weak homophily in the same discipline.

```{r echo=FALSE, results='hide'}
figure.counter <- figure.counter + 1
```

### Gender homophilly is also observed within countries

#### Histogram of $\alpha'$ for unique journal-country combinations

```{r}
country.histo <- country.data %>%
  mutate(Significant = 1 * (p.adjusted < 0.05),
         Significant = replace(Significant, Significant == 0, "No"),
         Significant = replace(Significant, Significant == 1, "Yes")) %>%
  ggplot(aes(x = alpha.adjusted, fill = Significant)) + 
  geom_vline(xintercept = 0, linetype = 2) +
  geom_histogram(binwidth = 0.02, colour = "black") + 
  scale_fill_manual(values = c("skyblue", "white")) +
  xlab("Coefficient of homophily (\u03B1')") + 
  ylab("Number of journals") +
  theme_minimal() + 
  theme(legend.position = "top")
ggsave(plot = country.histo, filename = "figures/figure4.pdf", device = cairo_pdf, width = 6, height = 5)
country.histo
```
<br></br>
**Figure S`r supp.figure.counter`**: Histogram of $\alpha'$ for 325 unique combinations of journal and country, using data from August 2015 - August 2016. The white areas denote combinations for which $\alpha'$ differs significantly from zero (p < 0.05, following false discovery rate correction).

```{r echo=FALSE, results='hide'}
supp.figure.counter <- supp.figure.counter + 1
```

#### Sample sizes for the journal-country dataset

```{r}
tibble(
  `Number of journal-country combinations` = nrow(country.data),
  `Median number of papers per journal` = median(country.data$n.useable.papers),
  `Median number of authors per journal` = median(country.data$n.authors)
) %>% as.data.frame() %>% pander(split.cell = 40, split.table = Inf)
```

```{r}
summary(lme4::lmer(alpha.adjusted ~ (1|journal) + (1|country), weights = 1/SE, data = country.data %>% 
               mutate(SE = (upper.95.CI2 - alpha.adjusted)/1.96) %>% filter(upper.95.CI2 > alpha.adjusted)))

paste("% variance associated with country:", round(100 * 0.001069 / (0.001069 + 0.001787 + 0.070926), 2))
paste("% variance associated with journal:", round(100 * 0.001787 / (0.001069 + 0.001787 + 0.070926), 2))
```

### Journal-country combinations with significantly non-zero $\alpha'$

```{r fig.height= 9}
x <- nrow(country.data[country.data$p.adjusted < 0.05, ])
country.data %>% 
  filter(p.adjusted < 0.05) %>%
  filter(discipline != "Multidisciplinary") %>%
  arrange(lower.95.CI2) %>% 
  mutate(y = paste(country, journal, sep = " - "), 
         y = (factor(y, y))) %>%
  ggplot(aes(alpha.adjusted, y, colour = y)) + 
  geom_errorbarh(aes(xmin = lower.95.CI2, xmax = upper.95.CI2), height = 0) +
  geom_point() + 
  xlab("Coefficient of homophily (\u03B1') \u00B1 95% CIs") + 
  ylab("Country - journal combination") + 
  theme_minimal() + 
  theme(legend.position = "none")
```
<br></br>
**Fig S`r supp.figure.counter`**: Plot showing the `r nrow(country.data[country.data$p.adjusted < 0.05, ])` combinations of journal and author country of affiliation for which $\alpha'$ is significantly higher than expected.

```{r echo=FALSE, results='hide'}
supp.figure.counter <- supp.figure.counter + 1
```

### Estimating inflation in $\alpha'$ due to country effects

Using the country data, we can estimate whether the Wahlund effect created by gender differences between countries produced 'spurious' positive $\alpha'$ values in our dataset.

For journal $i$, we take the average of the differences $\alpha'_i - \alpha'_{i,j}$, where $\alpha'_i$ is the coefficient of homophily calculated for all papers from journal $i$, and $\alpha'_{i,j}$ is the coefficient of homophily calculated for papers from journal $i$, by authors from country $j$. A positive value implies that pooling authors from different countries gives a higher estimate of homophily, consistent with a Wahlund effect.

Plotting a histogram of these average differences for the `r length(unique(country.data$journal))` journals shows that the mean and median difference is essentially zero, illustrating that $\alpha'$ is not strongly inflated by Wahlund effects caused by differences in author gender ratio between countries.

```{r}
mean.inflation.due.to.country <- country.data %>%
  group_by(journal) %>%
  summarise(mean.inflation.due.to.country = mean(journal.alpha - alpha.adjusted))

avg <- mean(mean.inflation.due.to.country$mean.inflation.due.to.country)

txt <- data.frame(x = avg + 0.035, y = 56, label = paste("Mean =", round(avg,3)))

mean.inflation.due.to.country %>%
  ggplot(aes(mean.inflation.due.to.country)) + 
  geom_histogram(binwidth = 0.02, colour = "black", fill = "skyblue") + 
  geom_vline(xintercept = avg, linetype = 2) + 
  geom_text(data=txt, aes(x=x, y=y, label=label)) + 
  theme_minimal() + 
  xlab("Estimated inflation in \u03B1' due to differences in gender ratio between countries") + 
  ylab("Frequency")
```
<br></br>
**Fig S`r supp.figure.counter`**: Histogram showing the estimated degree to which $\alpha'$ is inflated by inter-country differences in author gender ratio, across the `r length(unique(country.data$journal))` journals for which we had adequate data after restricting the analysis by country. The average inflation in $\alpha'$ is negligible, suggesting that Wahlund effects resulting from inter-country differences have neglible effect on our estimates of gender homophily.

```{r echo=FALSE, results='hide'}
supp.figure.counter <- supp.figure.counter + 1
```

### 2018-04-26 manuscript/Manuscript.Rmd
first commit

---
title: "Do researchers preferentially collaborate with same-gendered colleagues?"
author:
- Luke Holman* and Claire Morandin
- '*\textit{luke.holman@unimelb.edu.au} \vspace{5mm}'
output:
  pdf_document: default
  html_document: default
bibliography: references.bib
fontsize: 12pt
header-includes: 
  \usepackage[all]{hypcap} 
  \usepackage{amsmath} 
  \usepackage{booktabs}
  \usepackage{caption} 
  \usepackage[labelfont=bf]{caption} 
  \usepackage{titling} 
  \pretitle{\begin{flushleft}\LARGE}
  \posttitle{\par\end{flushleft}\vskip 0.5em} 
  \preauthor{\begin{flushleft}\large \lineskip 0.5em} 
  \postauthor{\par\end{flushleft}} 
  \predate{\begin{flushleft}\large} 
  \postdate{\par\end{flushleft}}
  \usepackage{fancyhdr} 
  \pagestyle{fancy} 
  \fancyhead[LO,LE]{\textsl{\leftmark}} 
  \rhead[]{Conference gender ratios} 
  \usepackage[noabbrev,capitalise]{cleveref}
  \usepackage{titlefoot} 
  \usepackage{amssymb} 
  \usepackage{rotating}
csl: the-american-naturalist.csl
abstract: Evidence suggests that women in academia are hindered by conscious and unconscious biases, and that many female researchers feel excluded from formal and informal opportunities for research collaboration. In addition to helping redress gender imbalance in the academic workforce, increasing women’s access to collaboration helps scientific progress by drawing on more of the available human capital. Here, we test whether researchers preferentially collaborate with same-gendered colleagues, expanding on earlier work by using more stringent methods and a larger dataset. We reaffirm that researchers tend to co-publish with colleagues of the same gender, and show that this ‘gender homophily’ is slightly stronger today than it was 10 years ago. Contrary to our expectations, we found no evidence that homophily is driven mostly by senior academics, and no evidence that homophily is strongest in fields where women are in the minority. Interestingly, homophily was negatively correlated with journal impact factor (standardised by research discipline), as predicted if mixed-gender teams produce better research. \vspace{5mm}
  \par\noindent \textbf{Keywords:} Gender bias, Homophily, Scientific collaboration, Test mining, Women in STEM
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

\maketitle\unmarkedfntext{*School of BioSciences, The University of Melbourne, Victoria, Australia}
\newpage

# Introduction
Women are substantially underrepresented in the workforce in many branches of science, technology, engineering, mathematics, and medicine (STEMM), and they face additional challenges and inequities relative to men [e.g. @Lariviere_2013; @West_2013; @Elsevier_report; @Holman_2018]. For example, women tend to occupy more junior positions (Wutte 2007, Reuben, Sapienza & Zingales, 2014) with lower salaries (Trower and Chait 2002, Umback 2007), receive less grant money (Hosek et al. 2005, OER 2005), are promoted more slowly (Hopkins et al. 2013; Long et al. 1993, Rosenfeld 1991, Zuckerman 1987), and are allocated fewer resources (European Commission 2009) and less research funding (Feldt 1986, Stack 2004, Larivière et al. 2011) than men. Experimental studies have demonstrated that researchers regard women's achievements less favourably than identical achievements by a man (Moss et al. 2012; Knobloch et al. 2013). Multiple studies have found that women publish fewer papers per year than men (Cole and Zuckerman 1984, Long 1992, Larivière et al. 2013, Astronomy one in Nature) [check whether they control for career stage].

Because publishing, networking and collaboration are all instrumental to academic career advancement (Lee and Boheman 2005, Hopkins et al. 2012), numerous studies have tested for gender differences in these areas (reviewed in Table S1 in Holman et al. 2018; see also XXX). For example, studies have shown that women tend to be less involved in international collaboration than men (Lemoine 1995; Lewison 2001; Webster 2001; Larivière et al. 2011, Bozeman and Corley 2004), collaborate less with researchers from their own university department (Websters 2001), and have less prestigious collaborations compared to men (Long 1990), but also fewer collaborations in general (Fuchs Von Stebut and Allmendinger 2001). Possible explanations for the smaller average professional networks of women and include a greater average amount of family obligations (Long 1990, Wright et al. 2003, Reskin 1978), lower participation of women in international conferences (Lewison 2001, Eisen 2014, Martin 2014), and reduced receipt of research travel funds by women (Bozeman and Corley 2004).

A high, steadily increasing proportion of research papers is written by more than one author (West et al. 2013), making collaboration a key predictor of publication output, and in turn, a key predictor of career prospects (Cohn et al 2014, Jordan et al. 2008, Tower et al. 2007). Previous studies suggest that mixed-gender (Campbell et al. 2013) or otherwise diverse (Hong and Page 2004) teams tend to produce better results on collaborative tasks. For these reasons, several studies have tested for gender differences in collaboration frequency or pattern by examining the author lists of published research. To our knowledge, every study of this question has concluded that men publish with other men, and women with women, more often than expected if collaborators assort randomly with regards to gender (Ferber and Teiman, 1980; McDowell and Smith 1992; Ghiasi et al. 2015; Crow and Smykla 2015, Zettler et al. 2016, Fahmy and Young, 2016, Jadidi et al. 2017; Teele and Thelen 2017, Araujo and Fontainha 2017a,b). This phenomenon is often termed ‘gender homophily’.

Unfortunately, these results all derive from studies that used sampling strategies that could overstate the strength of homophily, making the true extent of homophily difficult to ascertain from present data (\autoref{wahlund_plot}). Essentially, whenever samples are taken from two or more discrete sets of literature, which vary in the author gender ratio and which are largely not connected by collaboration, the number of same-gendered coauthors will be inflated. This can make it appear as though authors preferentially publish with same-gendered colleagues if no gender preferences exist, or even if researchers prefer opposite-gendered colleagues. For example, a sample of bioinformatics and cell biology papers might contain an excess of mostly-male and mostly-female author lists, simply because researchers preferentially work with colleagues from the same discipline, and because more women working in cell biology than in bioinformatics [@Holman_2018]. We name this issue the Wahlund effect, by analogy with the Wahlund effect in population genetics (Wahlund 1928), whereby spatial differences in allele frequencies create an excess of homozygotes, even if alleles are randomly assorted within each subpopulation.

In the present study, we test whether academics tend to co-publish with same-gendered collaborators, while controlling for the Wahlund effect as strictly as possible. Our study uses a recently-published dataset describing the gender of 35.5m authors from 9.15m articles indexed on PubMed [@Holman_2018]. Holman et al. [-@Holman_2018] reported large differences in the gender ratio of authors across research disciplines, journals, countries, and across the years 2002-2016. We therefore tested for gender homophily after restricting the analysis to particular journals (i.e. research specialties), time periods, and countries.

# Methods
## Obtaining author lists and assigning gender
We used the dataset collected from PubMed by Holman et al. [-@Holman_2018]. Briefly, Holman et al. created the dataset by downloading every single article indexed on PubMed, from the oldest articles up to the present (which was August 2016), and attempting to assign a gender to each author based on their given name, using the web service genderize.io. Each journal was assigned to one of >100 scientific disciplines (e.g. “Nephrology” or “Cell Biology”), using PubMed’s journal categorisations where available, or manual methods otherwise. Because the present study focuses on co-authorship, all single-author papers were removed from the dataset. To simplify the statistical analysis, we also discarded all papers for which we could not determine the gender of every author with at least 95% certainty. To mitigate Wahlund effects caused by variation in the gender ratio of researchers over time (see below), we also discarded all papers except those that were published 0-1 or 10-11 years ago, as of August 2016. Lastly, we excluded journals with fewer than 50 suitable papers. This left us with a dataset from August 2015 - August 2016 containing 276,879 papers with 1,311,213 authorships, and a dataset from August 2005 - August 2006 containing 151,652 papers and 647,634 authorships. These papers came from 2077 and 1192 journals respectively, which were grouped into 107 and 101 research disciplines. There was a median of 88 (87) papers per journal, 419 (371) authors per journal, and 4 (4) authors per paper (the first number is for the 2015-6 dataset, and the bracketed number for the 2005-6 dataset).

## Calculating $\alpha$, the coefficient of homophily
Inspired by Bergstrom et al. (http://www.eigenfactor.org/gender/assortativity/measuring_homophily.pdf), we defined the coefficient of homophily as $\alpha = p - q$, where $p$ is the probability that a randomly-chosen co-author of a male author is a man, and $q$ is the probability that a randomly-chosen co-author of a female author is a man. Therefore, $\alpha$ > 0 suggests that same-gender authors publish together more often than expected if authors assort randomly with respect to gender (homophily), while $\alpha$ < 0 suggests that opposite-gender authors publish together more often than expected (heterophily), and $\alpha$ = 0 suggests random assortment with respect to gender.

To estimate $\alpha$ for a particular subset of the scientific literature, we estimated $p$ as the average proportion of men’s co-authors who are men (averaged across all papers with at least one man author), and $q$ as the average proportion of women’s co-authors who are men (averaged across all papers with at least one woman author). To estimate the 95% confidence intervals on $\alpha$ for a given set of $n$ papers, we sampled $n$ papers with replacement 1000 times, estimated $\alpha$ on each sample, and recorded the 95% quantiles of the 1000 resulting estimates.

As well as calculating $\alpha$ for all authors, we calculated $\alpha$ for first and last authors only. $\alpha$ was again defined as $p - q$, but this time $p$ was estimated as the average proportion of male co-authors on papers with a male first/last author, and $q$ was estimated as the average proportion of male co-authors on papers with female first/last authors. We did not calculate $\alpha$ for other authorship positions (e.g. second or third authors) because this would necessitate reducing the dataset to papers with a sufficiently long authorship list, complicating comparison with other authorship positions that were calculated using a different sample.

Our test assumes that the expected value of $\alpha$ is zero if authors randomly assort, but for small datasets this is not always true (as pointed out by Carl T. Bergstrom in a blog post, http://www.eigenfactor.org/gender/assortativity/note_to_eisen.rtf). To borrow Prof. Bergstrom’s example, consider a small research specialty comprising just two men and two women researchers, who have together produced six two-author papers: one in each of the six possible combinations. For these six papers, $\alpha$ = $-\frac{1}{3}$, even though same- and opposite-gendered coauthors were selected in equal proportion to their frequency in the pool of possible collaborators. To control for the fact that the null expectation for $\alpha$ is not zero for every dataset (particularly small datasets), we devised an adjusted version of the coefficient of homophily, which we term $\alpha'$. Every time we calculated $\alpha$ for a set of papers, we also determined the expected value of $\alpha$ under the null hypothesis that authors assort randomly with respect to gender. This was accomplished by randomly shuffling authors across papers 1000 times, recalculating $\alpha$, and taking the median. We then calculated $\alpha'$ by subtracting the null expectation for $\alpha$ from the observed value. We also used the null-simulated $\alpha$ values to calculate a two-tailed p-value for the observed value of $\alpha$; the p-value was defined as the proportion of null simulations for which $|\alpha_{null}| > |\alpha_{obs}|$. We applied false discovery rate correction to each set of p-values to account for multiple testing (Benjamini and Hochberg 1995).

As expected, $\alpha'$ was usually almost identical to $\alpha$ (Figure S1), but $\alpha$ was downwardly biased relative to $\alpha'$ for small datasets (Figure S2). Additionally, the correlation between $\alpha'$ and sample size was negligible ($R^2$ < 0.01), suggesting that $\alpha'$ effectively removed the dependence of $\alpha$ on sample size. We therefore used the adjusted statistic $\alpha'$ in all our analyses.

## Minimising the Wahlund effect: discipline and time period
As discussed in the Introduction, the Wahlund effect can give the appearance of gender homophily (reflected in our study as $\alpha'$ > 0) whenever one has lumped together different ‘subpopulations’ of literature that differ in author gender ratio (Figure 1).

To minimise bias in $\alpha'$ due to the Wahlund effect, we restricted each set of papers to a single research specialty to the greatest extent allowed by our data. We therefore decided to only calculate $\alpha'$ for individual journals, since papers from the same journal typically focus on one or a few closely related topics. Although some journals, e.g. PLoS ONE, publish research from diverse disciplines with very different author gender ratios [@Holman_2018], calculating $\alpha'$ for these highly multidisciplinary journals is still useful as a contrast. The difference in $\alpha'$ between highly multidisciplinary and more specialised journals, e.g. PLoS ONE versus PLoS Computational Biology, gives a rough estimate of the maximum extent to which multidisciplinarity can inflate $\alpha'$.

As well as varying between disciplines, the gender ratio of authors has changed markedly over time [@Holman_2018]. Because the gender ratio was more male-biased in the past, $\alpha'$ would be inflated if we calculated it for a sample of papers published over a lengthy time frame. To minimise this effect, we only sampled papers from two one-year periods (namely 2005-6 and 2015-16). The median change per year in % (fe)male authors across journals is < 0.5% [@Holman_2018], and so restricting our dataset to a single year should prevent temporal changes in gender ratio from noticeably inflating our estimates of $\alpha'$.

## Minimising the Wahlund effect: author country of affiliation
A Wahlund effect could arise even if one calculates $\alpha'$ for a single discipline and time period, because of variation in the gender ratio of researchers from different countries. For example, Holman et al. [-@Holman_2018] showed that authors based in Serbia are more than twice as likely to be women as are authors based in Japan. Therefore, a dataset containing a mix of papers from teams of authors based in these two countries would probably contain an excess of same-sex coauthorships, even if collaboration were random with respect to gender within each country.

To address this, we also analysed every combination of journal and author country of affiliation for which we had enough data (meaning 50 or more papers published in 2015-16). For simplicity, we restricted the dataset to only include papers for which Holman et al. [-@Holman_2018] had successfully identified the country of affiliation for all authors on the paper, and all authors shared the same country of affiliation. Restricting the dataset in this fashion produced enough data to measure $\alpha'$ for 325 combinations of journal and country (median: 70 papers and 273 authors per combination).

## Calculating standardised journal impact factor
We obtained the 3-year impact factor for each journal from Clarivate Analytics. To account for large differences in impact factor between disciplines, we took the the residuals from a model with log10 impact factor as the response and the research discipline of the journal as a random effect. Thus, journals with a positive standardised impact factor have a higher mean number of citations than the average for journals in their discipline. We then tested whether our estimates of gender homophily covary with journal impact factor using Spearman rank correlation.

## Statistical models
Previous authors (e.g. Bonham & Stefan 2017) have hypothesised that senior scientists preferentially recruit staff and students of the same gender, and/or that junior researchers preferentially select same-gendered mentors. In the majority of disciplines, authorship conventions mean that the first-listed author is often an early-career researcher, while the author listed last is more likely to be a senior researcher leading a research team (Wren et al. 2007). Assuming that senior researchers are the main drivers of homophily and that there are enough papers with three or more authors, we predict that the last author’s gender will be the strongest predictor of the remaining authors’ genders (i.e. the gender of the last author will be more salient than that of the first author, or any other authorship position). This is because the first author’s gender would simply be an imperfect correlate of the true causal effect, while the last author’s gender would be the causal effect itself.

To test whether $\alpha'$ for last authors tends to be higher than $\alpha'$ for first authors for any given dataset, we used a linear mixed model implemented in the ``lme4`` and ``lmerTest`` packages for R, with _authorship position_ (first or last) as a fixed factor, and _journal_ and _research discipline_ as crossed random effects. The response variable was $\alpha'$, and we weighted each observation by the inverse of the standard error from our estimate of $\alpha'$, meaning that more accurate measurements of $\alpha'$ had more influence on the results.

We used a very similar model to test for a difference in $\alpha'$ between the 2005-6 and the 2015-16 datasets. There were two differences: instead of authorship position, we fit year range as a two-level fixed factor, and used $\alpha'$ estimated for all authors (not first/last authors) as the response variable.

The relationship between the gender ratio of authors publishing in a journal and its $\alpha'$ value appeared nonlinear (Figure 4). We therefore fit a generalised additive model with thin plate regression spline smoothing, implemented using the ``mgcv`` package for R.

# Results
## Most disciplines show evidence of gender homophily
Figure 2 shows the distribution of $\alpha'$ estimates in 2015-2016 across all journals for which we recovered sufficient data, when $\alpha'$ was calculated for all authors, first authors only, or last authors only. The great majority of journals had $\alpha'$ > 0, and for many of these the corrected two-tailed p-values suggested that $\alpha'$ was significantly greater than zero (1469/2077 journals were significant in 2015-16, and 404/1192 in 2005-6). Only 2/2077 journals had statistically significantly heterophily (i.e. $\alpha'$ < 0) in 2015-16, and 1/1192 in 2005-6. The remaining 606 or 787 journals (in 2015 and 2005 respectively) had a value of $\alpha'$ not significantly different from zero, such that we could not reject the null hypothesis of random assortment with respect to gender.

$\alpha'$ was statistically significantly higher in the sample from 2015-16 relative to 2005-6, but the difference in means was small (Figure SX; Effect of the fixed factor ‘time period’ in a linear mixed model of the data for all author positions: Cohen’s $d$ = 0.09  0.04, $t_{948}$ = 2.51, p = 0.012). This means that gender homophily was slightly stronger in 2015-6 relative to 2005-6.
When comparing pairs of $\alpha'$ values estimated for the first and last authors for the same journals, we found that $\alpha'$ tended to be higher for the first authors than for the last authors (Figure SX; Effect of the fixed factor ‘Authorship position’ in a linear mixed model: Cohen’s d = 0.07 ± 0.02, $t_{1988}$ = 4.48, p < 0.0001). This suggests that the gender of the first author was a slightly stronger predictor of the remaining authors’ genders than the gender of the last author, i.e. the opposite of our prediction based on the hypothesis that senior scientists are causally response for homophily.

## Possible differences in the strength of homophily between disciplines
Figure 3 illustrates the variance in journal homophily values ($\alpha'$) across scientific disciplines. All disciplines had a positive average $\alpha'$, although homophily appeared somewhat stronger in some disciplines than others (e.g. mean $\alpha'$ was 0.12±0.02 for Urology journals, and 0.03±0.01 for Veterinary Medicine journals; Table S3). However, there was little evidence for strong differences in $\alpha'$ between disciplines: the random factor ‘discipline’ explained at most 1% of the variance in $\alpha'$ in the two linear mixed models described in the previous section (see mixed models in Online Supplementary Material). This implies that the processes causing positive $\alpha'$ are similarly strong in all the disciplines we examined.

Reassuringly, we found no support for our prediction that journals publishing on a wide range of topics should have high $\alpha'$ values due to the Wahlund effect (Figure 3). For example, the journal category “Multidisciplinary” - which includes journals like PLoS ONE, Nature, Science, and PNAS - does not have notably elevated $\alpha'$. This result suggests that our estimates of homophily were not greatly inflated by the presence of disparate research topics (with variable author gender ratios) in the same journal.

## Relationship between gender homophily and gender ratio
One might predict that researchers would be more likely to seek out same-gendered colleagues in the most strongly gender-biased disciplines in our dataset (e.g. Surgery and Nursing), relative to disciplines with a comparatively gender-balanced workforce (e.g. Psychiatry). However, we instead found evidence of the reverse: there was a positive, non-linear relationship between the overall gender ratio of all authors publishing in the journal  [as estimated by @Holman_2018], and the estimated value of $\alpha'$ for all authors and for first authors (Figure 4). That is, journals with a balanced or female-biased author gender ratio tended to have higher $\alpha'$ than journals with a male-biased author gender ratio (GAM smooth terms p < 0.001, see Online Supplementary Material). The relationship was not statistically significant when $\alpha'$ was calculated for last authors (GAM, p = 0.142), though the trend appeared similar (Figure 4).

## Relationship between journal impact factor and gender homophily
We observed a noisy but statistically significant linear relationship between discipline-standardised journal impact factor and $\alpha'$, such that journals with a high impact factor for their discipline had weaker gender homophily (Figure 5; linear regression: $R^2$ = 0.043, $t_{1415}$ = -8.0, p < 0.0001).

## Analysis correcting for differences between countries
When we restricted the analysis to only include authors with affiliations from a single country, we found statistically significant homophily for 72 of the 325 journal-country combinations tested (64 unique journals and 18 unique countries), and no significant evidence of heterophily (Figures S5-S6). Additionally, the values of $\alpha'$ calculated for each journal-country combination were mostly very similar to the $\alpha'$ values calculated for the journal as a whole (i.e. when pooling papers from different countries); the average difference in $\alpha'$ was 0.002 (Figure S7). These results suggest that our main analysis’ findings of widespread homophily are not driven by a Wahlund effect resulting from gender differences between countries.

# Discussion
We found evidence that researchers tend to publish with same-gendered coauthors, even after implementing stringent controls for ‘spurious’ homophily resulting from Wahlund effects (Figure 1). Our data therefore concur with previous conclusions (Ferber and Teiman 1980; McDowell and Smith 1992; Crow and Smykla 2015; Ghiasi et al. 2015; Zettler et al. 2016; Fahmy and Young 2016; Jadihi et al. 2017; Araujo and Fontainha 2017a; Teele and Thelen 2017) that research collaborations occur between same-gendered colleagues substantially more often than expected if collaborators are selected without regard to their gender. Moreover, we found that gender homophily is not unique to particular disciplines, but rather appears to be similar in strength across many diverse fields of STEMM.

Academic publishing is one of the most (if not the most) important tool for a researcher to display their work, and is frequently used as a success indicator for being promoted, hired as a faculty member, or being invited as plenary speaker. Our finding that researchers are more likely to publish with same-gendered coauthors implies that co-authorship listing is not a random process, but instead creates additional gender imbalance. Homophily could make it harder for women to find collaborators in disciplines where women are in the minority [i.e. almost all STEMM fields; @Holman_2018]. Because women make up for as little as 30% of the STEMM scientist community (REF), they are therefore less likely to be picked as coauthors, and, as a result, gender homophily creates further gender imbalance and disadvantages for women. Men tend to publish more in high impact journals (REF) as first author, and generate more citations than papers lead by first author women (REF). As a consequence, gender homophily is also likely to contribute to lower publication rates for women (McDowell and Smith 1992), and lower scientific success (REF).

Why does gender homophily exists in science? Collaboration often leads to co-authorship publications, citations and additional professional recognitions, therefore first author should always be considering collaborators that are most likely to enhance their productivity and in the long term also their academic success, rather than pick co-authors according to their gender. First of all, it is likely that decades ago, the probability to meet a woman colleague working on the same area of expertise was slimmer than now. However, the number of women in STEM is increasing, and it is not likely to be an obstacle to collaboration nowadays.

# Acknowledgements
We are grateful to XXXX.

# References
<div id="refs"></div>

\newpage
# Figures
![The Wahlund effect can make it appear that authors prefer to publish with same-gendered colleagues, even when no such preference exists. Here, the circles represent male (blue) and female (orange) authors, and coauthors are linked with lines. Across the whole set of ten papers, there is an apparent excess of same-gender collaborations. Specifically, there are six same-gender papers and only four mixed-gender papers, which is less than the $10\times2\times0.5\times0.5 = 5$ mixed-gender papers we would expect under the null hypothesis that authors assort randomly with respect to gender. However, within each subset, there is no evidence that authors prefer to publish with same-gendered individuals. The Wahlund effect will tend to inflate the frequency of same-sex coauthors whenever the data is composed of two or more disconnected subsets of literature with different author gender ratios; these subsets could be research discipline, old versus new papers, or papers from authors in different countries. \label{wahlund_plot}](../figures/Figure 1 - Wahlund figure.png)

\newpage
```{r echo=FALSE, results='hide'}
fig2 <- readRDS("fig2_numbers.RData")
n.journals <- fig2[1]
n.sig.homo <- fig2[2]
n.sig.hetero <- fig2[3]
```
![Of the `r n.journals` journals for which we had adequate data in 2015-2016, `r n.sig.homo` showed statistically significant evidence of homophily (denoted by $\alpha' > 0$), and `r n.sig.hetero` showed statistically significant evidence of heterophily ($\alpha' < 0$), after adjusting p-values using Benjamini-Hochberg false discovery rate correction. The white area shows the number of journals for which homophily was significantly stronger than expected under the null hypothesis (p < 0.05), while the blue area shows all the remainder. Patterns were similar whether $\alpha'$ was calculated for all authors, for first authors only, or for last authors only. \label{alpha_histograms}](../figures/figure2.pdf)

\newpage
![There is a weakly positive, non-linear relationship between the gender ratio of authors publishing in a journal, and the coefficient of homophily ($\alpha'$). Specifically, journals with 50% women authors or higher tended to have more same-sex coauthorships than did journals with predominantly men authors.  This relationship held whether $\alpha'$ was calculated for all authors, first authors only, or last authors only. A negative value on the x axis denotes an excess of men authors, a positive value denotes an excess of women authors, and zero denotes gender parity. The lines were fitted using generalised additive models with the smoothing parameter $k$ set to 3. \label{alpha_gender_ratio}](../figures/figure3.pdf)

\newpage
![Histogram of $\alpha'$ for 325 unique combinations of journal and country, using data from August 2015 - August 2016. The white areas denote combinations for which $\alpha'$ differs significantly from zero (p < 0.05, following false discovery rate correction). \label{alpha_country}](../figures/figure4.pdf)

<!-- Sensitivity analysis of A -->
<!-- Note: using Latex to avoid an issue where the heading appears below the first figure. -->

<!-- \begin{figure}[!htbp] -->
<!--   \centering -->
<!--   \includegraphics[width=1\textwidth]{../Figures/pubSensA.pdf} -->
<!--   \caption{(\textbf{A}): Sensitivity analysis of Eq. \ref{eq-1}, a growth function subject to the Allee effect. (\textbf{B}): Sensitivity analysis of Eq. \ref{eq-1}, but with a reproductive trade-off across densities mediated by Eq. \ref{eq-1.1}. In both cases, $A_i$ determines the Allee threshold, or the critical density below which a population begins to shrink (here seen for values of $E(W_{ix})$ below the dotted line).} -->
<!--   \label{figSensA} -->
<!-- \end{figure} -->

<!-- ![(**A**): Mean $\bar{A}_x$ (the mean value of $A_i$, an Allee threshold governing trait, for all individuals located in the same patch) and mean density for invasions (n = 20) after 150 generations. More recently colonised patches exhibit greater levels of resistance to the Allee effect, as shown by lower mean values of $\bar{A}_x$. (**B**): Invasions with a reproductive trade-off. As the trade-off influences carrying capacity, core and frontier patches support differing numbers of invaders.\label{figScen1ADiff}](../Figures/pubScen1ADiffAlt.pdf) -->

### 2018-04-27 manuscript/Manuscript.Rmd
Edited MS

@@
 @Lariviere_2013; @West_2013; @Elsevier_report; @Holman_2018].
-For example, women tend to occupy more junior positions (Wutte 2007, Reuben, Sapienza & Zingales, 2014) with lower salaries (Trower and Chait 2002, Umback 2007), receive less grant money (Hosek et al.
+On average, women occupy more junior positions (Wutte 2007, Reuben, Sapienza & Zingales, 2014) with lower salaries (Trower and Chait 2002, Umback 2007), receive less grant money (Hosek et al.
 2005, OER 2005), are promoted more slowly (Hopkins et al.
@@
 2011) than men.
-Experimental studies have demonstrated that researchers regard women's achievements less favourably than identical achievements by a man (Moss et al.
+Experimental studies have demonstrated that researchers regard women's achievements less favourably than identical achievements by men (Moss et al.
 2012; Knobloch et al.
@@
 2007).
-Previous studies suggest that mixed-gender (Campbell et al.
-2013) or otherwise diverse (Hong and Page 2004) teams tend to produce better results on collaborative tasks.
-For these reasons, several studies have tested for gender differences in collaboration frequency or pattern by examining the author lists of published research.
+Additionally, empirical studies suggest that mixed-gender (Campbell et al.
+2013) or otherwise diverse (Hong and Page 2004) teams can produce better results on collaborative tasks.
+For reasons such as these, several studies have tested for gender differences in collaboration frequency or pattern by examining the author lists of published research.
 To our knowledge, every study of this question has concluded that men publish with other men, and women with women, more often than expected if collaborators assort randomly with regards to gender (Ferber and Teiman, 1980; McDowell and Smith 1992; Ghiasi et al.
@@
 2017; Teele and Thelen 2017, Araujo and Fontainha 2017a,b).
-This phenomenon is often termed ‘gender homophily’.
+This preference for same-gendered colleagues is often termed ‘gender homophily’.
 
-Unfortunately, these results all derive from studies that used sampling strategies that could overstate the strength of homophily, making the true extent of homophily difficult to ascertain from present data (\autoref{wahlund_plot}).
-Essentially, whenever samples are taken from two or more discrete sets of literature, which vary in the author gender ratio and which are largely not connected by collaboration, the number of same-gendered coauthors will be inflated.
+However, we believe that all of these studies used sampling strategies that could overstate the strength of homophily, making the true extent of homophily difficult to ascertain from present data.
+Essentially, whenever samples are taken from two or more discrete sets of literature that vary in the author gender ratio and that are largely not connected by collaboration, the number of same-gendered coauthors will be inflated (\autoref{wahlund_plot}).
 This can make it appear as though authors preferentially publish with same-gendered colleagues if no gender preferences exist, or even if researchers prefer opposite-gendered colleagues.
-For example, a sample of bioinformatics and cell biology papers might contain an excess of mostly-male and mostly-female author lists, simply because researchers preferentially work with colleagues from the same discipline, and because more women working in cell biology than in bioinformatics [@Holman_2018].
-We name this issue the Wahlund effect, by analogy with the Wahlund effect in population genetics (Wahlund 1928), whereby spatial differences in allele frequencies create an excess of homozygotes, even if alleles are randomly assorted within each subpopulation.
+For example, a sample of bioinformatics and cell biology papers might contain an excess of mostly-male and mostly-female author lists, simply because researchers preferentially work with colleagues from the same discipline, and because more women work in cell biology than in bioinformatics [@Holman_2018].
+We name this issue the Wahlund effect, by analogy with the Wahlund effect in population genetics (Wahlund 1928), whereby spatial differences in allele frequencies create an excess of homozygotes across the whole population, even if alleles assort randomly within each subpopulation.
 
@@
 # Discussion
-We found evidence that researchers tend to publish with same-gendered coauthors, even after implementing stringent controls for ‘spurious’ homophily resulting from Wahlund effects (Figure 1).
+We found evidence that researchers tend to publish with same-gendered coauthors, even after implementing stringent controls for ‘spurious’ homophily resulting from Wahlund effects (\autoref{wahlund_plot}).
 Our data therefore concur with previous conclusions (Ferber and Teiman 1980; McDowell and Smith 1992; Crow and Smykla 2015; Ghiasi et al.
@@
 2017; Araujo and Fontainha 2017a; Teele and Thelen 2017) that research collaborations occur between same-gendered colleagues substantially more often than expected if collaborators are selected without regard to their gender.
-Moreover, we found that gender homophily is not unique to particular disciplines, but rather appears to be similar in strength across many diverse fields of STEMM.
+Moreover, we found that gender homophily is not unique to particular disciplines, but rather appears to be similarly strong across many diverse fields of STEMM.
 

### 2018-04-27 Gender homophilly project.Rmd
More writing

@@
 ---
-title: "Homophily in science: do researchers preferentially collaborate with same-gendered colleagues?"
+title: "Do researchers preferentially collaborate with same-gendered colleagues?"
 subtitle: Online Supplementary Material
@@
 library(knitr) # set up cacheing to save time when re-building the html document - chunks are not re-evaluated if the code has not changed
+opts_chunk$set(warning=FALSE, message=FALSE) # Suppress warnings in the HTML output
+
 figure.counter <- 2  # Counters for numbering figures and tables
@@
 # each individual author is equally likely to end up in the sample. So if
-# 2-author vs 10-author papers have different amounts of homophilly, or 
+# 2-author vs 10-author papers have different amounts of homophily, or 
 # different gender ratios, the CIs should not be biased
@@
 
-```{r}
+```{r get_data}
 minimum.n <- 50   # Only include journals with at least 50 papers in the specified range of time
@@
 
-make.sample.size.table("older.papers") %>% 
+sample.size.table <- make.sample.size.table("older.papers") %>% 
   rename(`n (2005-2006)` = Value) %>% 
   left_join(make.sample.size.table("last.year") %>% 
-              rename(`n (2015-2016)` = Value), by = "Quantity") %>% 
-    pander(split.cell = 40, split.table = Inf)
+              rename(`n (2015-2016)` = Value), by = "Quantity")
+write.csv(sample.size.table, file = "manuscript/sample.size.table.csv", row.names = FALSE)
+
+sample.size.table %>% pander(split.cell = 40, split.table = Inf)
 ```
@@
 
-```{r}
+```{r get_alpha}
 # Calculate alpha across all authorship positions for all suitable NEW papers
@@
 ```{r fig.width=5, fig.height=8}
-figure2 <- master.dataset %>%
+figure2A <- master.dataset %>%
   filter(Age == "New") %>% 
@@
          Significant = replace(Significant, Significant == 1, "Yes")) 
-n.sig.homo <- sum(figure2$Significant[figure2$Subset == "All authors"] == "Yes" & 
-                    figure2$alpha.adjusted[figure2$Subset == "All authors"] > 0)
-n.sig.hetero <- sum(figure2$Significant[figure2$Subset == "All authors"] == "Yes" & 
-                    figure2$alpha.adjusted[figure2$Subset == "All authors"] < 0)
-n.journals <- sum(figure2$Subset == "All authors")
-
-figure2 <- figure2 %>%
+n.sig.homo <- with(figure2A, sum(Significant[Subset == "All authors"] == "Yes" & 
+                                   alpha.adjusted[Subset == "All authors"] > 0))
+n.sig.hetero <- with(figure2A, sum(Significant[Subset == "All authors"] == "Yes" & 
+                                     alpha.adjusted[Subset == "All authors"] < 0))
+n.journals <- sum(figure2A$Subset == "All authors")
+
+figure2A <- figure2A %>%
   ggplot(aes(x = alpha.adjusted, fill = Significant)) + 
   geom_vline(xintercept = 0, linetype = 2) +
-  geom_histogram(binwidth = 0.01, colour = "black") + 
+  geom_area(binwidth = 0.01, colour = "black", stat = "bin") + 
   scale_fill_manual(values = c("skyblue", "white")) +
@@
   theme_minimal(12) + 
-  theme(legend.position = "top")
-ggsave(plot = figure2, filename = "figures/figure2.pdf",  width = 5, height = 8, device = cairo_pdf)
-figure2
+  theme(legend.position = c(0.9, 0.96))
+figure2A
 ```
 <br></br>
-**Figure `r figure.counter`**: Of the `r n.journals` journals for which we had adequate data in 2015-2016, `r n.sig.homo` showed statistically significant evidence of homophily (denoted by $\alpha' > 0$), and `r n.sig.hetero` showed statistically significant evidence of heterophily ($\alpha' < 0$), after adjusting p-values using Benjamini-Hochberg false discovery rate correction.
-The white area shows the number of journals for which homophily was significantly stronger than expected under the null hypothesis (p < 0.05), while the blue area shows all the remainder.
-Patterns were similar whether $\alpha'$ was calculated for all authors, for first authors only, or for last authors only.
+**Figure 2A**: See the legend for Figure 2 below.
 
 ```{r echo=FALSE, results='hide'}
-figure.counter <- figure.counter + 1
-
 # Save these figures for manuscript too:
@@
 
-gender.homophily.all %>% 
+figure2B <- gender.homophily.all %>% 
   filter(discipline %in% disc.summary$Discipline[disc.summary$`n journals` > 10]) %>%
   mutate(Discipline = factor(discipline, levels = rev(disc.summary$Discipline))) %>%
-  ggplot(aes(x = Discipline, y = alpha.adjusted, fill = Discipline)) +
+  ggplot(aes(x = Discipline, y = alpha.adjusted)) +
   geom_hline(yintercept = 0, linetype = 2) + 
-  geom_boxplot(outlier.size = 0.5, outlier.shape = NA) + 
   geom_quasirandom(aes(colour = Discipline), alpha = 0.3, bandwidth = 0.25) + 
-  ylab("Coefficient of homophily (\u03B1')") + 
+  geom_boxplot(outlier.size = 0.5, outlier.shape = NA, fill = NA) +
+  ylab(NULL) + 
   coord_flip() + 
@@
   theme(legend.position = "none")
+
+grid.arrange(figure2A + xlab(NULL), figure2B, nrow = 1, bottom = "Coefficient of homophily (\u03B1')")
+
+ggsave(grid.arrange(figure2A + xlab(NULL), figure2B, nrow = 1, bottom = "Coefficient of homophily (\u03B1')"), filename = "figures/figure2.pdf", height = 8, width = 7, device = cairo_pdf)
 ```
 <br></br>
-**Figure `r figure.counter`**: Distribution of the $\alpha'$ values across journals, split by the research discipline of the journal.
-The points show the $\alpha'$ for each journal, and the boxes show quartiles 1, 2 and 3.
+**Figure `r figure.counter`**: Of the `r n.journals` journals for which we had adequate data in 2015-2016, `r n.sig.homo` showed statistically significant evidence of homophily (denoted by $\alpha' > 0$), and `r n.sig.hetero` showed statistically significant evidence of heterophily ($\alpha' < 0$), after adjusting p-values using Benjamini-Hochberg false discovery rate correction.
+In the left panels, the white area shows the number of journals for which homophily was significantly stronger than expected under the null hypothesis (p < 0.05), while the blue area shows the remainder.
+Patterns were similar whether $\alpha'$ was calculated for all authors, for first authors only, or for last authors only.
+The right panel shows the distribution of the $\alpha'$ values across journals, split by the research discipline of the journal.
+The points show the $\alpha'$ for individual journals, and the boxes show quartiles 1, 2 and 3.
 The boxplot whiskers show the distance of the farthest point that is within 1.5 times the interquartile range from the edge of the box.
@@
   geom_point(alpha = 0.4, size = 0.7) + 
-  stat_smooth(method = "gam", formula = y ~ s(x, k = 3), colour = "black", fill = "skyblue") + 
+  stat_smooth(method = "gam", formula = y ~ s(x, k = 3), colour = "tomato", fill = "skyblue") + 
   xlab("Deviation from gender parity for all authors in this journal") + 
@@
 
-gender.homophily.all %>% 
+figure4 <- gender.homophily.all %>% 
   filter(!is.na(standardised.IF)) %>% 
@@
   geom_vline(xintercept = 0, linetype = 2) +
-  stat_smooth(method = "lm", colour = "black", fill = "skyblue") +
+  stat_smooth(method = "lm", colour = "tomato", fill = "skyblue") +
   geom_point(size = 0.5, alpha = 0.9) + 
@@
   theme_minimal()
+ggsave(figure4, file = "figures/figure4.pdf", height = 5, width = 5, device = cairo_pdf)
+saveRDS(R2, "manuscript/R2_for_figure4.RData")
+figure4
 ```
@@
 
-### Gender homophilly is also observed within countries
+### Gender homophily is also observed within countries
 
@@
   theme(legend.position = "top")
-ggsave(plot = country.histo, filename = "figures/figure4.pdf", device = cairo_pdf, width = 6, height = 5)
 country.histo

### 2018-04-27 manuscript/Manuscript.Rmd
More writing

@@
 abstract: Evidence suggests that women in academia are hindered by conscious and unconscious biases, and that many female researchers feel excluded from formal and informal opportunities for research collaboration.
-In addition to helping redress gender imbalance in the academic workforce, increasing women’s access to collaboration helps scientific progress by drawing on more of the available human capital.
-Here, we test whether researchers preferentially collaborate with same-gendered colleagues, expanding on earlier work by using more stringent methods and a larger dataset.
-We reaffirm that researchers tend to co-publish with colleagues of the same gender, and show that this ‘gender homophily’ is slightly stronger today than it was 10 years ago.
+In addition to ensuring fairness and helping redress gender imbalance in the academic workforce, increasing women’s access to collaboration could help scientific progress by drawing on more of the available human capital.
+Here, we test whether researchers preferentially collaborate with same-gendered colleagues, using more stringent methods and a larger dataset than in past work.
+Our results reaffirm that researchers co-publish with colleagues of the same gender, and show that this ‘gender homophily’ is slightly stronger today than it was 10 years ago.
 Contrary to our expectations, we found no evidence that homophily is driven mostly by senior academics, and no evidence that homophily is strongest in fields where women are in the minority.
 Interestingly, homophily was negatively correlated with journal impact factor (standardised by research discipline), as predicted if mixed-gender teams produce better research.
-\vspace{5mm}
-  \par\noindent \textbf{Keywords:} Gender bias, Homophily, Scientific collaboration, Test mining, Women in STEM
+\vspace{5mm} \par\noindent \textbf{Keywords:} Gender bias, Homophily, Scientific collaboration, Text mining, Women in STEM.
 ---
@@
 knitr::opts_chunk$set(echo = FALSE)
+
+# Load some little files with numbers to be inserted into the MS
+sample.size <- read.csv("sample.size.table.csv", stringsAsFactors = FALSE)
+sample.size[,2] <- prettyNum(sample.size[,2], big.mark = ",")
+sample.size[,3] <- prettyNum(sample.size[,3], big.mark = ",")
+names(sample.size)[2:3] <- c("n2005", "n2015")
+
+fig2 <- readRDS("fig2_numbers.RData") # For figure 2
+n.journals <- fig2[1]
+n.sig.homo <- fig2[2]
+n.sig.hetero <- fig2[3]
+
+R2 <- readRDS("R2_for_figure4.RData") # For Fig 4
 ```
@@
 [-@Holman_2018].
-Briefly, Holman et al.
-created the dataset by downloading every single article indexed on PubMed, from the oldest articles up to the present (which was August 2016), and attempting to assign a gender to each author based on their given name, using the web service genderize.io.
-Each journal was assigned to one of >100 scientific disciplines (e.g.
-“Nephrology” or “Cell Biology”), using PubMed’s journal categorisations where available, or manual methods otherwise.
+Briefly, that dataset was created by downloading every single article indexed on PubMed, from the oldest articles up to the present (which was August 2016), and attempting to assign a gender to each author based on their given name, using the web service genderize.io.
+Each journal was assigned to one of 107 scientific disciplines (e.g.
+“Nephrology” or “Cell Biology”), using PubMed’s journal categorisations where available.
 Because the present study focuses on co-authorship, all single-author papers were removed from the dataset.
 To simplify the statistical analysis, we also discarded all papers for which we could not determine the gender of every author with at least 95% certainty.
-To mitigate Wahlund effects caused by variation in the gender ratio of researchers over time (see below), we also discarded all papers except those that were published 0-1 or 10-11 years ago, as of August 2016.
+To mitigate Wahlund effects caused by variation in the gender ratio of researchers over time (see below), we also discarded all papers except those that were published 0-1 or 10-11 years before the PubMed data were collected (i.e.
+20 August 2016).
 Lastly, we excluded journals with fewer than 50 suitable papers.
-This left us with a dataset from August 2015 - August 2016 containing 276,879 papers with 1,311,213 authorships, and a dataset from August 2005 - August 2006 containing 151,652 papers and 647,634 authorships.
-These papers came from 2077 and 1192 journals respectively, which were grouped into 107 and 101 research disciplines.
-There was a median of 88 (87) papers per journal, 419 (371) authors per journal, and 4 (4) authors per paper (the first number is for the 2015-6 dataset, and the bracketed number for the 2005-6 dataset).
+This left us with a dataset from August 2015 - August 2016 containing `r sample.size$n2015[3]` papers with `r sample.size$n2015[4]` authorships, and a dataset from August 2005 - August 2006 containing `r sample.size$n2005[3]` papers and `r sample.size$n2005[4]` authorships.
+These papers came from `r sample.size$n2015[2]` and `r sample.size$n2005[2]` journals respectively, which were grouped into `r sample.size$n2015[1]` and `r sample.size$n2005[1]` research disciplines.
+There was a median of `r sample.size$n2015[5]` (`r sample.size$n2005[5]`) papers per journal, `r sample.size$n2015[6]` (`r sample.size$n2005[6]`) authors per journal, and `r sample.size$n2015[7]` (`r sample.size$n2005[7]`) authors per paper (the first number is for 2015-6, and the bracketed number for 2005-6).
 
@@
 Inspired by Bergstrom et al.
-(http://www.eigenfactor.org/gender/assortativity/measuring_homophily.pdf), we defined the coefficient of homophily as $\alpha = p - q$, where $p$ is the probability that a randomly-chosen co-author of a male author is a man, and $q$ is the probability that a randomly-chosen co-author of a female author is a man.
-Therefore, $\alpha$ > 0 suggests that same-gender authors publish together more often than expected if authors assort randomly with respect to gender (homophily), while $\alpha$ < 0 suggests that opposite-gender authors publish together more often than expected (heterophily), and $\alpha$ = 0 suggests random assortment with respect to gender.
+(http://www.eigenfactor.org/gender/assortativity/measuring_homophily.pdf), we defined the coefficient of homophily as $\alpha = p - q$, where $p$ is the probability that a randomly-chosen co-author of a _male_ author is a man and $q$ is the probability that a randomly-chosen co-author of a _female_ author is a man.
+Therefore, $\alpha$ > 0 suggests that same-gender authors publish together more often than expected if authors assort randomly with respect to gender (homophily), $\alpha$ < 0 suggests that opposite-gender authors publish together more often than expected (heterophily), and $\alpha$ = 0 suggests random assortment with respect to gender.
 
@@
 
-Our test assumes that the expected value of $\alpha$ is zero if authors randomly assort, but for small datasets this is not always true (as pointed out by Carl T.
+Our test assumes that the expected value of $\alpha$ is zero if authors randomly assort, but for small datasets this assumption is not always true (as pointed out by Carl T.
 Bergstrom in a blog post, http://www.eigenfactor.org/gender/assortativity/note_to_eisen.rtf).
 To borrow Prof.
-Bergstrom’s example, consider a small research specialty comprising just two men and two women researchers, who have together produced six two-author papers: one in each of the six possible combinations.
+Bergstrom’s example, consider a small research specialty comprising just two men and two women researchers, who have together produced six two-author papers: one paper in each of the six possible two-author combinations.
 For these six papers, $\alpha$ = $-\frac{1}{3}$, even though same- and opposite-gendered coauthors were selected in equal proportion to their frequency in the pool of possible collaborators.
-To control for the fact that the null expectation for $\alpha$ is not zero for every dataset (particularly small datasets), we devised an adjusted version of the coefficient of homophily, which we term $\alpha'$.
+To control for the fact that the null expectation for $\alpha$ is not necessarily zero (particularly for very small datasets), we devised an adjusted version of the coefficient of homophily, which we term $\alpha'$.
 Every time we calculated $\alpha$ for a set of papers, we also determined the expected value of $\alpha$ under the null hypothesis that authors assort randomly with respect to gender.
@@
 ## Minimising the Wahlund effect: discipline and time period
-As discussed in the Introduction, the Wahlund effect can give the appearance of gender homophily (reflected in our study as $\alpha'$ > 0) whenever one has lumped together different ‘subpopulations’ of literature that differ in author gender ratio (Figure 1).
+As discussed in the Introduction, the Wahlund effect can give the appearance of gender homophily (reflected in our study as $\alpha'$ > 0) whenever one lumps together different ‘subpopulations’ of academic papers that differ in their author gender ratios (Figure 1).
 
 To minimise bias in $\alpha'$ due to the Wahlund effect, we restricted each set of papers to a single research specialty to the greatest extent allowed by our data.
-We therefore decided to only calculate $\alpha'$ for individual journals, since papers from the same journal typically focus on one or a few closely related topics.
+Specifically, we only calculated $\alpha'$ for individual journals, since papers from the same journal typically focus on one or a few closely related topics.
 Although some journals, e.g.
-PLoS ONE, publish research from diverse disciplines with very different author gender ratios [@Holman_2018], calculating $\alpha'$ for these highly multidisciplinary journals is still useful as a contrast.
+_PLoS ONE_, publish research from diverse disciplines with very different author gender ratios [@Holman_2018], calculating $\alpha'$ for these highly multidisciplinary journals is still useful as a contrast.
 The difference in $\alpha'$ between highly multidisciplinary and more specialised journals, e.g.
-PLoS ONE versus PLoS Computational Biology, gives a rough estimate of the maximum extent to which multidisciplinarity can inflate $\alpha'$.
+_PLoS ONE_ versus _PLoS Computational Biology_, gives a rough estimate of the extent to which multidisciplinarity can inflate $\alpha'$.
 
 As well as varying between disciplines, the gender ratio of authors has changed markedly over time [@Holman_2018].
-Because the gender ratio was more male-biased in the past, $\alpha'$ would be inflated if we calculated it for a sample of papers published over a lengthy time frame.
+Because the gender ratio was more male-biased in the past, $\alpha'$ would be inflated if we calculated it for a sample of papers published over a long time frame.
 To minimise this effect, we only sampled papers from two one-year periods (namely 2005-6 and 2015-16).
@@
 
-To address this, we also analysed every combination of journal and author country of affiliation for which we had enough data (meaning 50 or more papers published in 2015-16).
+To address this, we also analysed every combination of journal and author country of affiliation for which we had enough data (i.e.
+50 or more papers published in 2015-16).
 For simplicity, we restricted the dataset to only include papers for which Holman et al.
@@
 We obtained the 3-year impact factor for each journal from Clarivate Analytics.
-To account for large differences in impact factor between disciplines, we took the the residuals from a model with log10 impact factor as the response and the research discipline of the journal as a random effect.
+To account for large differences in impact factor between disciplines, we took the the residuals from a model with $Log_{10}$ impact factor as the response and the research discipline of the journal as a random effect.
 Thus, journals with a positive standardised impact factor have a higher mean number of citations than the average for journals in their discipline.
-We then tested whether our estimates of gender homophily covary with journal impact factor using Spearman rank correlation.
+We then used Spearman rank correlation to test whether our estimates of gender homophily covaried with journal impact factor.
 
-## Statistical models
+## Statistical analysis
 Previous authors (e.g.
@@
 The response variable was $\alpha'$, and we weighted each observation by the inverse of the standard error from our estimate of $\alpha'$, meaning that more accurate measurements of $\alpha'$ had more influence on the results.
-
-We used a very similar model to test for a difference in $\alpha'$ between the 2005-6 and the 2015-16 datasets.
-There were two differences: instead of authorship position, we fit year range as a two-level fixed factor, and used $\alpha'$ estimated for all authors (not first/last authors) as the response variable.
+We used a similar model to test for a difference in $\alpha'$ between the 2005-6 and the 2015-16 datasets, with two differences: instead of authorship position, we fit year range as a two-level fixed factor, and used $\alpha'$ estimated for all authors (not first/last authors) as the response variable.
 
@@
 # Results
-## Most disciplines show evidence of gender homophily
-Figure 2 shows the distribution of $\alpha'$ estimates in 2015-2016 across all journals for which we recovered sufficient data, when $\alpha'$ was calculated for all authors, first authors only, or last authors only.
+## Gender homophily by discipline, time period, and authorship position
+\autoref{alpha_histograms} shows the distribution of $\alpha'$ estimates in 2015-2016 across all journals for which we recovered sufficient data, when $\alpha'$ was calculated for all authors, first authors only, or last authors only.
 The great majority of journals had $\alpha'$ > 0, and for many of these the corrected two-tailed p-values suggested that $\alpha'$ was significantly greater than zero (1469/2077 journals were significant in 2015-16, and 404/1192 in 2005-6).
@@
 
-$\alpha'$ was statistically significantly higher in the sample from 2015-16 relative to 2005-6, but the difference in means was small (Figure SX; Effect of the fixed factor ‘time period’ in a linear mixed model of the data for all author positions: Cohen’s $d$ = 0.09  0.04, $t_{948}$ = 2.51, p = 0.012).
-This means that gender homophily was slightly stronger in 2015-6 relative to 2005-6.
-When comparing pairs of $\alpha'$ values estimated for the first and last authors for the same journals, we found that $\alpha'$ tended to be higher for the first authors than for the last authors (Figure SX; Effect of the fixed factor ‘Authorship position’ in a linear mixed model: Cohen’s d = 0.07 ± 0.02, $t_{1988}$ = 4.48, p < 0.0001).
+$\alpha'$ was significantly higher in the sample from 2015-16 relative to 2005-6, but the difference in means was small (Figure S3; Effect of the fixed factor ‘time period’ in a linear mixed model of the data for all author positions: Cohen’s $d$ = 0.09 ± 0.04, $t_{948}$ = 2.51, p = 0.012).
+This result illustrates that gender homophily was slightly stronger in 2015-6 relative to 2005-6.
+
+When comparing pairs of $\alpha'$ values estimated for the first and last authors for the same journals, we found that $\alpha'$ tended to be higher for first authors than for last authors (Figure S4; Effect of the fixed factor ‘Authorship position’ in a linear mixed model: Cohen’s d = 0.07 ± 0.02, $t_{1988}$ = 4.48, p < 0.0001).
 This suggests that the gender of the first author was a slightly stronger predictor of the remaining authors’ genders than the gender of the last author, i.e.
-the opposite of our prediction based on the hypothesis that senior scientists are causally response for homophily.
+the opposite of our prediction based on the hypothesis that senior scientists are causally responsible for homophily.
 
 ## Possible differences in the strength of homophily between disciplines
-Figure 3 illustrates the variance in journal homophily values ($\alpha'$) across scientific disciplines.
+\label{alpha_gender_ratio} illustrates the variance in journal homophily values ($\alpha'$) across scientific disciplines.
 All disciplines had a positive average $\alpha'$, although homophily appeared somewhat stronger in some disciplines than others (e.g.
-mean $\alpha'$ was 0.12±0.02 for Urology journals, and 0.03±0.01 for Veterinary Medicine journals; Table S3).
+mean $\alpha'$ was 0.12 ± 0.02 for Urology journals, and 0.03 ± 0.01 for Veterinary Medicine journals; Table S3).
 However, there was little evidence for strong differences in $\alpha'$ between disciplines: the random factor ‘discipline’ explained at most 1% of the variance in $\alpha'$ in the two linear mixed models described in the previous section (see mixed models in Online Supplementary Material).
@@
 
-Reassuringly, we found no support for our prediction that journals publishing on a wide range of topics should have high $\alpha'$ values due to the Wahlund effect (Figure 3).
+Reassuringly, we found no support for our prediction that journals publishing on a wide range of topics should have high $\alpha'$ values due to the Wahlund effect \label{alpha_gender_ratio}.
 For example, the journal category “Multidisciplinary” - which includes journals like PLoS ONE, Nature, Science, and PNAS - does not have notably elevated $\alpha'$.
@@
 ## Relationship between gender homophily and gender ratio
-One might predict that researchers would be more likely to seek out same-gendered colleagues in the most strongly gender-biased disciplines in our dataset (e.g.
-Surgery and Nursing), relative to disciplines with a comparatively gender-balanced workforce (e.g.
+One might predict that researchers would be more likely to seek out same-gendered colleagues in strongly gender-biased disciplines (e.g.
+Surgery or Nursing), relative to disciplines with a comparatively gender-balanced workforce (e.g.
 Psychiatry).
-However, we instead found evidence of the reverse: there was a positive, non-linear relationship between the overall gender ratio of all authors publishing in the journal  [as estimated by @Holman_2018], and the estimated value of $\alpha'$ for all authors and for first authors (Figure 4).
+However, we instead found evidence of the reverse: there was a positive, non-linear relationship between the overall gender ratio of all authors publishing in a particular journal [as estimated by @Holman_2018], and the estimated value of $\alpha'$ for all authors and for first authors \label{alpha_country}.
 That is, journals with a balanced or female-biased author gender ratio tended to have higher $\alpha'$ than journals with a male-biased author gender ratio (GAM smooth terms p < 0.001, see Online Supplementary Material).
-The relationship was not statistically significant when $\alpha'$ was calculated for last authors (GAM, p = 0.142), though the trend appeared similar (Figure 4).
+The relationship was not statistically significant when $\alpha'$ was calculated for last authors (GAM, p = 0.142), though the trend appeared similar \label{alpha_country}.
 
 ## Relationship between journal impact factor and gender homophily
-We observed a noisy but statistically significant linear relationship between discipline-standardised journal impact factor and $\alpha'$, such that journals with a high impact factor for their discipline had weaker gender homophily (Figure 5; linear regression: $R^2$ = 0.043, $t_{1415}$ = -8.0, p < 0.0001).
+We observed a noisy but statistically significant linear relationship between discipline-standardised journal impact factor and $\alpha'$, such that journals with a high impact factor for their discipline had weaker gender homophily than did journals with a low impact factor for their discipline (Figure 5; linear regression: $R^2$ = `r R2`, $t_{1415}$ = -8.0, p < 0.0001).
 
@@
 We found evidence that researchers tend to publish with same-gendered coauthors, even after implementing stringent controls for ‘spurious’ homophily resulting from Wahlund effects (\autoref{wahlund_plot}).
-Our data therefore concur with previous conclusions (Ferber and Teiman 1980; McDowell and Smith 1992; Crow and Smykla 2015; Ghiasi et al.
+Our data therefore reaffirm earlier conclusions (Ferber and Teiman 1980; McDowell and Smith 1992; Crow and Smykla 2015; Ghiasi et al.
 2015; Zettler et al.
@@
 2017; Araujo and Fontainha 2017a; Teele and Thelen 2017) that research collaborations occur between same-gendered colleagues substantially more often than expected if collaborators are selected without regard to their gender.
-Moreover, we found that gender homophily is not unique to particular disciplines, but rather appears to be similarly strong across many diverse fields of STEMM.
+Many journals had $\alpha'$ values over 0.1, meaning that the gender ratio of the average man's coauthors is skewed towards men by >10%, relative to the gender ratio of a woman's coauthors (or equivalently, that the gender ratio of women's coauthors is >10% more female than that of men's).
+Few journals had $\alpha'$ values below zero, and fewer still had significantly negative $\alpha'$ values, suggesting that researchers rarely seek out collaborators of the opposite gender (or at least, that they do so less often than they seek out same-gendered colleagues).
+
+We found no evidence that gender homophily is restricted to particular disciplines: instead, there was evidence of homophily across many diverse research disciplines.
+We also found that gender homophily was marginally stronger in 2015-2016, relative to 2005-2006.
+This result suggests that recent increases in the number of women working in STEMM, as well as recent efforts to recruit, retain, and celebrate women in STEMM, have (if anything) strengthened gender homophily rather than weakening it.
+
+Gender homophily was slightly stronger in research disciplines with a female-biased gender ratio, as compared to male-biased or gender-balanced fields.
+One interpretation of this result is that men are more likely to seek out other men in fields in which men are in the minority, relative to the homophily shown by women in fields in which women are in the minority.
+
+We also found that journals with a high impact factor (relative to other journals in the same research discipline) tended to have weaker gender homophily than did low-impact journals.
+If one accepts the contentious claim that papers published in high-impact journals tend to be of higher quality than those in low-impact journals [e.g.
+@Garfield_2006], then this result provides correlational support for the hypothesis that mixed-gender teams produce better research than single-gender teams (Campbell et al.
+2013).
+
+Contrary to our expectations, we found no evidence that homophily is driven mostly by senior academics, and no evidence that homophily is strongest in fields where women are in the minority.
+Interestingly, homophily was negatively correlated with journal impact factor (standardised by research discipline), as predicted if mixed-gender teams produce better research.
 
@@
 # Figures
-![The Wahlund effect can make it appear that authors prefer to publish with same-gendered colleagues, even when no such preference exists.
-Here, the circles represent male (blue) and female (orange) authors, and coauthors are linked with lines.
+![The Wahlund effect can make it appear as if authors prefer to publish with same-gendered colleagues, even if no such preference exists.
+Here, coloured circles represent male and female authors, and coauthors are linked with lines.
 Across the whole set of ten papers, there is an apparent excess of same-gender collaborations.
-Specifically, there are six same-gender papers and only four mixed-gender papers, which is less than the $10\times2\times0.5\times0.5 = 5$ mixed-gender papers we would expect under the null hypothesis that authors assort randomly with respect to gender.
+Specifically, there are six same-gender papers and only four mixed-gender papers, which is fewer than the $10\times2\times0.5\times0.5 = 5$ mixed-gender papers we would expect under the null hypothesis that authors assort randomly with respect to gender.
 However, within each subset, there is no evidence that authors prefer to publish with same-gendered individuals.
-The Wahlund effect will tend to inflate the frequency of same-sex coauthors whenever the data is composed of two or more disconnected subsets of literature with different author gender ratios; these subsets could be research discipline, old versus new papers, or papers from authors in different countries.
-\label{wahlund_plot}](../figures/Figure 1 - Wahlund figure.png)
+The Wahlund effect will tend to inflate the frequency of same-sex coauthors whenever the data is composed of two or more disconnected subsets of literature with different author gender ratios; these subsets could be research disciplines, older versus newer papers, or papers from authors in different countries.
+\label{wahlund_plot}](../figures/Figure 1 - Wahlund figure.pdf)
 
 \newpage
-```{r echo=FALSE, results='hide'}
-fig2 <- readRDS("fig2_numbers.RData")
-n.journals <- fig2[1]
-n.sig.homo <- fig2[2]
-n.sig.hetero <- fig2[3]
-```
 ![Of the `r n.journals` journals for which we had adequate data in 2015-2016, `r n.sig.homo` showed statistically significant evidence of homophily (denoted by $\alpha' > 0$), and `r n.sig.hetero` showed statistically significant evidence of heterophily ($\alpha' < 0$), after adjusting p-values using Benjamini-Hochberg false discovery rate correction.
@@
 This relationship held whether $\alpha'$ was calculated for all authors, first authors only, or last authors only.
-A negative value on the x axis denotes an excess of men authors, a positive value denotes an excess of women authors, and zero denotes gender parity.
+A negative value on the x-axis denotes an excess of men authors, a positive value denotes an excess of women authors, and zero denotes gender parity.
 The lines were fitted using generalised additive models with the smoothing parameter $k$ set to 3.
@@
 \newpage
-![Histogram of $\alpha'$ for 325 unique combinations of journal and country, using data from August 2015 - August 2016.
-The white areas denote combinations for which $\alpha'$ differs significantly from zero (p < 0.05, following false discovery rate correction).
-\label{alpha_country}](../figures/figure4.pdf)
-
-<!-- Sensitivity analysis of A -->
-<!-- Note: using Latex to avoid an issue where the heading appears below the first figure. -->
-
-<!-- \begin{figure}[!htbp] -->
-<!--   \centering -->
-<!--   \includegraphics[width=1\textwidth]{../Figures/pubSensA.pdf} -->
-<!--   \caption{(\textbf{A}): Sensitivity analysis of Eq.
-\ref{eq-1}, a growth function subject to the Allee effect.
-(\textbf{B}): Sensitivity analysis of Eq.
-\ref{eq-1}, but with a reproductive trade-off across densities mediated by Eq.
-\ref{eq-1.1}.
-In both cases, $A_i$ determines the Allee threshold, or the critical density below which a population begins to shrink (here seen for values of $E(W_{ix})$ below the dotted line).} -->
-<!--   \label{figSensA} -->
-<!-- \end{figure} -->
-
-<!-- ![(**A**): Mean $\bar{A}_x$ (the mean value of $A_i$, an Allee threshold governing trait, for all individuals located in the same patch) and mean density for invasions (n = 20) after 150 generations.
-More recently colonised patches exhibit greater levels of resistance to the Allee effect, as shown by lower mean values of $\bar{A}_x$.
-(**B**): Invasions with a reproductive trade-off.
-As the trade-off influences carrying capacity, core and frontier patches support differing numbers of invaders.\label{figScen1ADiff}](../Figures/pubScen1ADiffAlt.pdf) -->
+![Journal impact factor (expressed relative to the average for the discipline) is negatively correlated with $\alpha'$.
+The relationship is noisy ($R^2$ = `r R2`), but the results suggest that journals with strong homophily tend to have lower impact factors than journals with weak homophily in the same discipline.
+\label{impact_factor}](../figures/figure5.pdf)

### 2018-04-28 manuscript/Manuscript.Rmd
Tweaks

@@
 ## Possible differences in the strength of homophily between disciplines
-\label{alpha_gender_ratio} illustrates the variance in journal homophily values ($\alpha'$) across scientific disciplines.
+\autoref{alpha_histograms} illustrates the variance in journal homophily values ($\alpha'$) across scientific disciplines.
 All disciplines had a positive average $\alpha'$, although homophily appeared somewhat stronger in some disciplines than others (e.g.
-mean $\alpha'$ was 0.12 ± 0.02 for Urology journals, and 0.03 ± 0.01 for Veterinary Medicine journals; Table S3).
-However, there was little evidence for strong differences in $\alpha'$ between disciplines: the random factor ‘discipline’ explained at most 1% of the variance in $\alpha'$ in the two linear mixed models described in the previous section (see mixed models in Online Supplementary Material).
+mean $\alpha'$ was 0.12 ± 0.02 for Urology journals, and 0.03 ± 0.01 for Veterinary Medicine journals; \autoref{alpha_histograms}, Table S3).
+However, there was little evidence for strong differences in $\alpha'$ between disciplines: the random factor ‘discipline’ explained at most 1% of the variance in $\alpha'$ in the two linear mixed models described in the previous section (see \autoref{alpha_histograms} and mixed models in Online Supplementary Material).
 This implies that the processes causing positive $\alpha'$ are similarly strong in all the disciplines we examined.
 
-Reassuringly, we found no support for our prediction that journals publishing on a wide range of topics should have high $\alpha'$ values due to the Wahlund effect \label{alpha_gender_ratio}.
-For example, the journal category “Multidisciplinary” - which includes journals like PLoS ONE, Nature, Science, and PNAS - does not have notably elevated $\alpha'$.
-This result suggests that our estimates of homophily were not greatly inflated by the presence of disparate research topics (with variable author gender ratios) in the same journal.
+Reassuringly, we found no support for our prediction that journals publishing on a wide range of topics should have higher $\alpha'$ values than more specialised journals, due to the Wahlund effect.
+Specifically, the journal category “Multidisciplinary” - which includes journals like _PLoS ONE_, _Nature_, _Science_, and _PNAS_ - did not have notably elevated $\alpha'$ (\autoref{alpha_histograms}).
+This result suggests that our estimates of homophily are not notably inflated by the presence of disparate research topics (with variable author gender ratios) being published within individual journals.
 
@@
 Psychiatry).
-However, we instead found evidence of the reverse: there was a positive, non-linear relationship between the overall gender ratio of all authors publishing in a particular journal [as estimated by @Holman_2018], and the estimated value of $\alpha'$ for all authors and for first authors \label{alpha_country}.
+However, we instead found evidence of the reverse: there was a positive, non-linear relationship between the overall gender ratio of all authors publishing in a particular journal [as estimated by @Holman_2018], and the estimated value of $\alpha'$ for all authors and for first authors (\autoref{alpha_gender_ratio}).
 That is, journals with a balanced or female-biased author gender ratio tended to have higher $\alpha'$ than journals with a male-biased author gender ratio (GAM smooth terms p < 0.001, see Online Supplementary Material).
-The relationship was not statistically significant when $\alpha'$ was calculated for last authors (GAM, p = 0.142), though the trend appeared similar \label{alpha_country}.
+The relationship was not statistically significant when $\alpha'$ was calculated for last authors (GAM, p = 0.142), though the trend appeared similar (\autoref{alpha_gender_ratio}).
 
 ## Relationship between journal impact factor and gender homophily
-We observed a noisy but statistically significant linear relationship between discipline-standardised journal impact factor and $\alpha'$, such that journals with a high impact factor for their discipline had weaker gender homophily than did journals with a low impact factor for their discipline (Figure 5; linear regression: $R^2$ = `r R2`, $t_{1415}$ = -8.0, p < 0.0001).
+We observed a noisy but statistically significant linear relationship between discipline-standardised journal impact factor and $\alpha'$, such that journals with a high impact factor for their discipline had weaker gender homophily than did journals with a low impact factor for their discipline (\autoref{impact_factor}; linear regression: $R^2$ = `r R2`, $t_{1415}$ = -8.0, p < 0.0001).
 
@@
 The relationship is noisy ($R^2$ = `r R2`), but the results suggest that journals with strong homophily tend to have lower impact factors than journals with weak homophily in the same discipline.
-\label{impact_factor}](../figures/figure5.pdf)
+\label{impact_factor}](../figures/figure4.pdf)

### 2018-05-01 manuscript/Manuscript.Rmd
Intro

@@
 # Introduction
-Women are substantially underrepresented in the workforce in many branches of science, technology, engineering, mathematics, and medicine (STEMM), and they face additional challenges and inequities relative to men [e.g.
+Women are substantially underrepresented in many branches of the workforce in science, technology, engineering, mathematics, and medicine (STEMM), and they face additional challenges and inequities relative to men [e.g.
 @Lariviere_2013; @West_2013; @Elsevier_report; @Holman_2018].
@@
 2013).
-Multiple studies have found that women publish fewer papers per year than men (Cole and Zuckerman 1984, Long 1992, Larivière et al.
-2013, Astronomy one in Nature) [check whether they control for career stage].
 
@@
 2018; see also XXX).
-For example, studies have shown that women tend to be less involved in international collaboration than men (Lemoine 1995; Lewison 2001; Webster 2001; Larivière et al.
-2011, Bozeman and Corley 2004), collaborate less with researchers from their own university department (Websters 2001), and have less prestigious collaborations compared to men (Long 1990), but also fewer collaborations in general (Fuchs Von Stebut and Allmendinger 2001).
-Possible explanations for the smaller average professional networks of women and include a greater average amount of family obligations (Long 1990, Wright et al.
+For example, studies have shown that relative to men, women tend to be less involved in international collaboration (Lemoine 1995; Lewison 2001; Webster 2001; Larivière et al.
+2011, Bozeman and Corley 2004), collaborate less with researchers from their own university department (Websters 2001), have less prestigious collaborations (Long 1990), and fewer collaborations in general (Fuchs Von Stebut and Allmendinger 2001).
+Possible explanations for the smaller average professional networks of women include a greater average amount of family obligations (Long 1990, Wright et al.
 2003, Reskin 1978), lower participation of women in international conferences (Lewison 2001, Eisen 2014, Martin 2014), and reduced receipt of research travel funds by women (Bozeman and Corley 2004).
@@
 
-However, we believe that all of these studies used sampling strategies that could overstate the strength of homophily, making the true extent of homophily difficult to ascertain from present data.
-Essentially, whenever samples are taken from two or more discrete sets of literature that vary in the author gender ratio and that are largely not connected by collaboration, the number of same-gendered coauthors will be inflated (\autoref{wahlund_plot}).
-This can make it appear as though authors preferentially publish with same-gendered colleagues if no gender preferences exist, or even if researchers prefer opposite-gendered colleagues.
+However, we hypothesise that these studies of gender homophily were hindered by a sampling issue that we term the Wahlund effect (\autoref{wahlund_plot}), making the true extent of homophily difficult to ascertain from present data.
+Essentially, whenever samples are taken from two or more discrete sets of literature, which vary in the author gender ratio and which are largely not connected by collaboration, the number of same-gendered coauthors will be inflated.
+This can make it appear as though authors preferentially publish with same-gendered colleagues if no gender preferences exist, or even if researchers preferentially select opposite-gendered colleagues.
 For example, a sample of bioinformatics and cell biology papers might contain an excess of mostly-male and mostly-female author lists, simply because researchers preferentially work with colleagues from the same discipline, and because more women work in cell biology than in bioinformatics [@Holman_2018].
-We name this issue the Wahlund effect, by analogy with the Wahlund effect in population genetics (Wahlund 1928), whereby spatial differences in allele frequencies create an excess of homozygotes across the whole population, even if alleles assort randomly within each subpopulation.
+We name this issue after the Wahlund effect in population genetics (Wahlund 1928), whereby spatial differences in allele frequencies create an excess of homozygotes across the whole population, even if alleles assort randomly within each subpopulation.
 

### 2018-05-01 manuscript/Manuscript.Rmd
writing

@@
 ## Obtaining author lists and assigning gender
-We used the dataset collected from PubMed by Holman et al.
+We used the dataset of PubMed author lists from Holman et al.
 [-@Holman_2018].
@@
 For simplicity, we restricted the dataset to only include papers for which Holman et al.
-[-@Holman_2018] had successfully identified the country of affiliation for all authors on the paper, and all authors shared the same country of affiliation.
+[-@Holman_2018] had identified the country of affiliation for all authors on the paper, and all authors shared the same country of affiliation.
 Restricting the dataset in this fashion produced enough data to measure $\alpha'$ for 325 combinations of journal and country (median: 70 papers and 273 authors per combination).
@@
 
-$\alpha'$ was significantly higher in the sample from 2015-16 relative to 2005-6, but the difference in means was small (Figure S3; Effect of the fixed factor ‘time period’ in a linear mixed model of the data for all author positions: Cohen’s $d$ = 0.09 ± 0.04, $t_{948}$ = 2.51, p = 0.012).
-This result illustrates that gender homophily was slightly stronger in 2015-6 relative to 2005-6.
+$\alpha'$ was significantly higher in the sample from 2015-16 relative to 2005-6, but the difference in means was small (Figure S3; Effect of the fixed factor ‘time period’ in a linear mixed model of the data for all author positions: Cohen’s $d$ = 0.09±0.04, $t_{948}$ = 2.51, p = 0.012).
 
-When comparing pairs of $\alpha'$ values estimated for the first and last authors for the same journals, we found that $\alpha'$ tended to be higher for first authors than for last authors (Figure S4; Effect of the fixed factor ‘Authorship position’ in a linear mixed model: Cohen’s d = 0.07 ± 0.02, $t_{1988}$ = 4.48, p < 0.0001).
+When comparing pairs of $\alpha'$ values estimated for the first and last authors for the same journals, we found that $\alpha'$ tended to be higher for first authors than for last authors (Figure S4; Effect of the fixed factor ‘Authorship position’ in a linear mixed model: Cohen’s $d$ = 0.07±0.02, $t_{1988}$ = 4.48, p < 0.0001).
 This suggests that the gender of the first author was a slightly stronger predictor of the remaining authors’ genders than the gender of the last author, i.e.
@@
 All disciplines had a positive average $\alpha'$, although homophily appeared somewhat stronger in some disciplines than others (e.g.
-mean $\alpha'$ was 0.12 ± 0.02 for Urology journals, and 0.03 ± 0.01 for Veterinary Medicine journals; \autoref{alpha_histograms}, Table S3).
+mean $\alpha'$ was 0.12±0.02 for Urology journals, and 0.03±0.01 for Veterinary Medicine journals; \autoref{alpha_histograms}, Table S3).
 However, there was little evidence for strong differences in $\alpha'$ between disciplines: the random factor ‘discipline’ explained at most 1% of the variance in $\alpha'$ in the two linear mixed models described in the previous section (see \autoref{alpha_histograms} and mixed models in Online Supplementary Material).
@@
 
-Reassuringly, we found no support for our prediction that journals publishing on a wide range of topics should have higher $\alpha'$ values than more specialised journals, due to the Wahlund effect.
-Specifically, the journal category “Multidisciplinary” - which includes journals like _PLoS ONE_, _Nature_, _Science_, and _PNAS_ - did not have notably elevated $\alpha'$ (\autoref{alpha_histograms}).
-This result suggests that our estimates of homophily are not notably inflated by the presence of disparate research topics (with variable author gender ratios) being published within individual journals.
+There was no indication that journals publishing on a wide range of topics have higher $\alpha'$ values than more specialised journals, due to the Wahlund effect.
+For example, the journal category “Multidisciplinary” - which includes journals like _PLoS ONE_, _Nature_, _Science_, and _PNAS_ - did not have notably elevated $\alpha'$ (\autoref{alpha_histograms}).
+This result suggests that our estimates of homophily, and estimates from earlier studies, are not notably inflated by the presence of disparate research topics (with variable author gender ratios) being published within individual journals.
 
 ## Relationship between gender homophily and gender ratio
-One might predict that researchers would be more likely to seek out same-gendered colleagues in strongly gender-biased disciplines (e.g.
+We next tested whether researchers are more or less likely to seek out same-gendered colleagues in strongly gender-biased disciplines (e.g.
 Surgery or Nursing), relative to disciplines with a comparatively gender-balanced workforce (e.g.
 Psychiatry).
-However, we instead found evidence of the reverse: there was a positive, non-linear relationship between the overall gender ratio of all authors publishing in a particular journal [as estimated by @Holman_2018], and the estimated value of $\alpha'$ for all authors and for first authors (\autoref{alpha_gender_ratio}).
-That is, journals with a balanced or female-biased author gender ratio tended to have higher $\alpha'$ than journals with a male-biased author gender ratio (GAM smooth terms p < 0.001, see Online Supplementary Material).
+We found a positive, non-linear relationship between the overall gender ratio of all authors publishing in a particular journal [as estimated by @Holman_2018], and the estimated value of $\alpha'$ for all authors and for first authors (\autoref{alpha_gender_ratio}).
+Journals with a balanced or female-biased author gender ratio tended to have higher $\alpha'$ than journals with a male-biased author gender ratio (GAM smooth terms p < 0.001, see Online Supplementary Material).
 The relationship was not statistically significant when $\alpha'$ was calculated for last authors (GAM, p = 0.142), though the trend appeared similar (\autoref{alpha_gender_ratio}).
@@
 when pooling papers from different countries); the average difference in $\alpha'$ was 0.002 (Figure S7).
-These results suggest that our main analysis’ findings of widespread homophily are not driven by a Wahlund effect resulting from gender differences between countries.
+These results suggest that our findings of widespread homophily in the main analysis cannot solely be driven by a Wahlund effect resulting from gender differences between countries.
 
 # Discussion
-We found evidence that researchers tend to publish with same-gendered coauthors, even after implementing stringent controls for ‘spurious’ homophily resulting from Wahlund effects (\autoref{wahlund_plot}).
+We found evidence that researchers tend to publish with same-gendered coauthors, despite implementing stringent controls for ‘spurious’ homophily resulting from Wahlund effects (\autoref{wahlund_plot}).
 Our data therefore reaffirm earlier conclusions (Ferber and Teiman 1980; McDowell and Smith 1992; Crow and Smykla 2015; Ghiasi et al.
@@
 2017; Araujo and Fontainha 2017a; Teele and Thelen 2017) that research collaborations occur between same-gendered colleagues substantially more often than expected if collaborators are selected without regard to their gender.
-Many journals had $\alpha'$ values over 0.1, meaning that the gender ratio of the average man's coauthors is skewed towards men by >10%, relative to the gender ratio of a woman's coauthors (or equivalently, that the gender ratio of women's coauthors is >10% more female than that of men's).
-Few journals had $\alpha'$ values below zero, and fewer still had significantly negative $\alpha'$ values, suggesting that researchers rarely seek out collaborators of the opposite gender (or at least, that they do so less often than they seek out same-gendered colleagues).
+Many journals had $\alpha'$ values over 0.1, meaning that the gender ratio of the average man's coauthors is skewed towards men by >10%, relative to the gender ratio of a woman's coauthors.
+Few journals had $\alpha'$ values below zero, and fewer still had significantly negative $\alpha'$ values, suggesting that it is comparatively rare for researchers to preferentially select opposite-gendered collaborators.
 
@@
 We also found that gender homophily was marginally stronger in 2015-2016, relative to 2005-2006.
-This result suggests that recent increases in the number of women working in STEMM, as well as recent efforts to recruit, retain, and celebrate women in STEMM, have (if anything) strengthened gender homophily rather than weakening it.
+This result suggests that recent increases in the number of women working in STEMM, as well as recent efforts to recruit, retain, and celebrate women in STEMM, have co-occurred with rise in the frequency of same-gender collaborations, not a fall.
+We speculate that this trend might result, at least in part, from the increasing number of women working in senior positions in STEMM over the last 10 years (DATA?).
+Because many co-authorships involve a junior and a senior researcher (e.g.
+a student and professor), and the gender ratio is typically more skewed among senior staff (Shaw and Stanton ref), there will be more opposite-gender collaborations than expected if the gender ratio were the same at both career stages.
+Thus, the observed increase in homophily might be due to the greater numbers of women working in leading roles.
 

### 2018-05-22 manuscript/Manuscript.tex
Manuscript_Claire.Rmd

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Do researchers preferentially collaborate with same-gendered colleagues?},
            pdfauthor={Luke Holman* and Claire Morandin; *},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{Do researchers preferentially collaborate with same-gendered colleagues?}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{Luke Holman* and Claire Morandin \\ *\textit{luke.holman@unimelb.edu.au} \vspace{5mm}}
  \preauthor{\centering\large\emph}
  \postauthor{\par}
  \date{}
  \predate{}\postdate{}

\usepackage[all]{hypcap} \usepackage{amsmath} \usepackage{booktabs}
\usepackage{caption} \usepackage[labelfont=bf]{caption}
\usepackage{titling} \pretitle{\begin{flushleft}\LARGE}
\posttitle{\par\end{flushleft}\vskip 0.5em}
\preauthor{\begin{flushleft}\large \lineskip 0.5em}
\postauthor{\par\end{flushleft}} \predate{\begin{flushleft}\large}
\postdate{\par\end{flushleft}} \usepackage{fancyhdr} \pagestyle{fancy}
\fancyhead[LO,LE]{\textsl{\leftmark}} \rhead[]{Conference gender ratios}
\usepackage[noabbrev,capitalise]{cleveref} \usepackage{titlefoot}
\usepackage{amssymb} \usepackage{rotating}

\begin{document}
\maketitle
\begin{abstract}
Evidence suggests that women in academia are hindered by conscious and
unconscious biases, and that many female researchers feel excluded from
formal and informal opportunities for research collaboration. In
addition to ensuring fairness and helping redress gender imbalance in
the academic workforce, increasing women???s access to collaboration
could help scientific progress by drawing on more of the available human
capital. Here, we test whether researchers preferentially collaborate
with same-gendered colleagues, using more stringent methods and a larger
dataset than in past work. Our results reaffirm that researchers
co-publish with colleagues of the same gender, and show that this
???gender homophily??? is slightly stronger today than it was 10 years
ago. Contrary to our expectations, we found no evidence that homophily
is driven mostly by senior academics, and no evidence that homophily is
strongest in fields where women are in the minority. Interestingly,
homophily was negatively correlated with journal impact factor
(standardised by research discipline), as predicted if mixed-gender
teams produce better research. \vspace{5mm}
\par\noindent \textbf{Keywords:} Gender bias, Homophily, Scientific
collaboration, Text mining, Women in STEM.
\end{abstract}

\maketitle

\unmarkedfntext{*School of BioSciences, The University of Melbourne, Victoria, Australia}
\newpage

\section{Introduction}\label{introduction}

Women are substantially underrepresented in many branches of the
workforce in science, technology, engineering, mathematics, and medicine
(STEMM), and they face additional challenges and inequities relative to
men (e.g. Larivière et al. 2013; West et al. 2013; Elsevier Report 2017;
Holman et al. 2018). On average, women occupy more junior positions
(Wutte 2007, Reuben, Sapienza \& Zingales, 2014) with lower salaries
(Trower and Chait 2002, Umback 2007), receive less grant money (Hosek et
al. 2005, OER 2005), are promoted more slowly (Hopkins et al. 2013; Long
et al. 1993, Rosenfeld 1991, Zuckerman 1987), and are allocated fewer
resources (European Commission 2009) and less research funding (Feldt
1986, Stack 2004, Larivi??re et al. 2011) than men. Experimental studies
have demonstrated that researchers regard women's achievements less
favourably than identical achievements by men (Moss et al. 2012;
Knobloch et al. 2013).

Because publishing, networking and collaboration are all instrumental to
academic career advancement (Lee and Boheman 2005, Hopkins et al. 2012),
numerous studies have tested for gender differences in these areas
(reviewed in Table S1 in Holman et al. 2018; see also XXX). For example,
studies have shown that relative to men, women tend to be less involved
in international collaboration (Lemoine 1995; Lewison 2001; Webster
2001; Larivi??re et al. 2011, Bozeman and Corley 2004), collaborate less
with researchers from their own university department (Websters 2001),
have less prestigious collaborations (Long 1990), and fewer
collaborations in general (Fuchs Von Stebut and Allmendinger 2001).
Possible explanations for the smaller average professional networks of
women include a greater average amount of family obligations (Long 1990,
Wright et al. 2003, Reskin 1978), lower participation of women in
international conferences (Lewison 2001, Eisen 2014, Martin 2014), and
reduced receipt of research travel funds by women (Bozeman and Corley
2004).

A high, steadily increasing proportion of research papers is written by
more than one author (West et al. 2013), making collaboration a key
predictor of publication output, and in turn, a key predictor of career
prospects (Cohn et al 2014, Jordan et al. 2008, Tower et al. 2007).
Additionally, empirical studies suggest that mixed-gender (Campbell et
al. 2013) or otherwise diverse (Hong and Page 2004) teams can produce
better results on collaborative tasks. For reasons such as these,
several studies have tested for gender differences in collaboration
frequency or pattern by examining the author lists of published
research. To our knowledge, every study of this question has concluded
that men publish with other men, and women with women, more often than
expected if collaborators assort randomly with regards to gender (Ferber
and Teiman, 1980; McDowell and Smith 1992; Ghiasi et al. 2015; Crow and
Smykla 2015, Zettler et al. 2016, Fahmy and Young, 2016, Jadidi et al.
2017; Teele and Thelen 2017, Araujo and Fontainha 2017a,b). This
preference for same-gendered colleagues is often termed ???gender
homophily???.

However, we hypothesise that these studies of gender homophily were
hindered by a sampling issue that we term the Wahlund effect
(\autoref{wahlund_plot}), making the true extent of homophily difficult
to ascertain from present data. Essentially, whenever samples are taken
from two or more discrete sets of literature, which vary in the author
gender ratio and which are largely not connected by collaboration, the
number of same-gendered coauthors will be inflated. This can make it
appear as though authors preferentially publish with same-gendered
colleagues if no gender preferences exist, or even if researchers
preferentially select opposite-gendered colleagues. For example, a
sample of bioinformatics and cell biology papers might contain an excess
of mostly-male and mostly-female author lists, simply because
researchers preferentially work with colleagues from the same
discipline, and because more women work in cell biology than in
bioinformatics (Holman et al. 2018). We name this issue after the
Wahlund effect in population genetics (Wahlund 1928), whereby spatial
differences in allele frequencies create an excess of homozygotes across
the whole population, even if alleles assort randomly within each
subpopulation.

In the present study, we test whether academics tend to co-publish with
same-gendered collaborators, while controlling for the Wahlund effect as
strictly as possible. Our study uses a recently-published dataset
describing the gender of 35.5m authors from 9.15m articles indexed on
PubMed (Holman et al. 2018). Holman et al. (2018) reported large
differences in the gender ratio of authors across research disciplines,
journals, countries, and across the years 2002-2016. We therefore tested
for gender homophily after restricting the analysis to particular
journals (i.e.~research specialties), time periods, and countries.

\section{Methods}\label{methods}

\subsection{Obtaining author lists and assigning
gender}\label{obtaining-author-lists-and-assigning-gender}

We used the dataset of PubMed author lists from Holman et al. (2018).
Briefly, that dataset was created by downloading every single article
indexed on PubMed, from the oldest articles up to the present (which was
August 2016), and attempting to assign a gender to each author based on
their given name, using the web service genderize.io. Each journal was
assigned to one of 107 scientific disciplines (e.g. ???Nephrology??? or
???Cell Biology???), using PubMed???s journal categorisations where
available. Because the present study focuses on co-authorship, all
single-author papers were removed from the dataset. To simplify the
statistical analysis, we also discarded all papers for which we could
not determine the gender of every author with at least 95\% certainty.
To mitigate Wahlund effects caused by variation in the gender ratio of
researchers over time (see below), we also discarded all papers except
those that were published 0-1 or 10-11 years before the PubMed data were
collected (i.e.~20 August 2016). Lastly, we excluded journals with fewer
than 50 suitable papers. This left us with a dataset from August 2015 -
August 2016 containing 276,879 papers with 1,311,213 authorships, and a
dataset from August 2005 - August 2006 containing 151,652 papers and
647,634 authorships. These papers came from 2,116 and 1,192 journals
respectively, which were grouped into 107 and 101 research disciplines.
There was a median of 87 (87) papers per journal, 413 (371) authors per
journal, and 4 (4) authors per paper (the first number is for 2015-6,
and the bracketed number for 2005-6).

\subsection{\texorpdfstring{Calculating \(\alpha\), the coefficient of
homophily}{Calculating \textbackslash{}alpha, the coefficient of homophily}}\label{calculating-alpha-the-coefficient-of-homophily}

Inspired by Bergstrom et al.
(\url{http://www.eigenfactor.org/gender/assortativity/measuring_homophily.pdf}),
we defined the coefficient of homophily as \(\alpha = p - q\), where
\(p\) is the probability that a randomly-chosen co-author of a
\emph{male} author is a man and \(q\) is the probability that a
randomly-chosen co-author of a \emph{female} author is a man. Therefore,
\(\alpha\) \textgreater{} 0 suggests that same-gender authors publish
together more often than expected if authors assort randomly with
respect to gender (homophily), \(\alpha\) \textless{} 0 suggests that
opposite-gender authors publish together more often than expected
(heterophily), and \(\alpha\) = 0 suggests random assortment with
respect to gender.

To estimate \(\alpha\) for a particular subset of the scientific
literature, we estimated \(p\) as the average proportion of men???s
co-authors who are men (averaged across all papers with at least one man
author), and \(q\) as the average proportion of women???s co-authors who
are men (averaged across all papers with at least one woman author). To
estimate the 95\% confidence intervals on \(\alpha\) for a given set of
\(n\) papers, we sampled \(n\) papers with replacement 1000 times,
estimated \(\alpha\) on each sample, and recorded the 95\% quantiles of
the 1000 resulting estimates.

As well as calculating \(\alpha\) for all authors, we calculated
\(\alpha\) for first and last authors only. \(\alpha\) was again defined
as \(p - q\), but this time \(p\) was estimated as the average
proportion of male co-authors on papers with a male first/last author,
and \(q\) was estimated as the average proportion of male co-authors on
papers with female first/last authors. We did not calculate \(\alpha\)
for other authorship positions (e.g.~second or third authors) because
this would necessitate reducing the dataset to papers with a
sufficiently long authorship list, complicating comparison with other
authorship positions that were calculated using a different sample.

Our test assumes that the expected value of \(\alpha\) is zero if
authors randomly assort, but for small datasets this assumption is not
always true (as pointed out by Carl T. Bergstrom in a blog post,
\url{http://www.eigenfactor.org/gender/assortativity/note_to_eisen.rtf}).
To borrow Prof.~Bergstrom???s example, consider a small research
specialty comprising just two men and two women researchers, who have
together produced six two-author papers: one paper in each of the six
possible two-author combinations. For these six papers, \(\alpha\) =
\(-\frac{1}{3}\), even though same- and opposite-gendered coauthors were
selected in equal proportion to their frequency in the pool of possible
collaborators. To control for the fact that the null expectation for
\(\alpha\) is not necessarily zero (particularly for very small
datasets), we devised an adjusted version of the coefficient of
homophily, which we term \(\alpha'\). Every time we calculated
\(\alpha\) for a set of papers, we also determined the expected value of
\(\alpha\) under the null hypothesis that authors assort randomly with
respect to gender. This was accomplished by randomly shuffling authors
across papers 1000 times, recalculating \(\alpha\), and taking the
median. We then calculated \(\alpha'\) by subtracting the null
expectation for \(\alpha\) from the observed value. We also used the
null-simulated \(\alpha\) values to calculate a two-tailed p-value for
the observed value of \(\alpha\); the p-value was defined as the
proportion of null simulations for which
\(|\alpha_{null}| > |\alpha_{obs}|\). We applied false discovery rate
correction to each set of p-values to account for multiple testing
(Benjamini and Hochberg 1995).

As expected, \(\alpha'\) was usually almost identical to \(\alpha\)
(Figure S1), but \(\alpha\) was downwardly biased relative to
\(\alpha'\) for small datasets (Figure S2). Additionally, the
correlation between \(\alpha'\) and sample size was negligible (\(R^2\)
\textless{} 0.01), suggesting that \(\alpha'\) effectively removed the
dependence of \(\alpha\) on sample size. We therefore used the adjusted
statistic \(\alpha'\) in all our analyses.

\subsection{Minimising the Wahlund effect: discipline and time
period}\label{minimising-the-wahlund-effect-discipline-and-time-period}

As discussed in the Introduction, the Wahlund effect can give the
appearance of gender homophily (reflected in our study as \(\alpha'\)
\textgreater{} 0) whenever one lumps together different
???subpopulations??? of academic papers that differ in their author
gender ratios (Figure 1).

To minimise bias in \(\alpha'\) due to the Wahlund effect, we restricted
each set of papers to a single research specialty to the greatest extent
allowed by our data. Specifically, we only calculated \(\alpha'\) for
individual journals, since papers from the same journal typically focus
on one or a few closely related topics. Although some journals, e.g.
\emph{PLoS ONE}, publish research from diverse disciplines with very
different author gender ratios (Holman et al. 2018), calculating
\(\alpha'\) for these highly multidisciplinary journals is still useful
as a contrast. The difference in \(\alpha'\) between highly
multidisciplinary and more specialised journals, e.g. \emph{PLoS ONE}
versus \emph{PLoS Computational Biology}, gives a rough estimate of the
extent to which multidisciplinarity can inflate \(\alpha'\).

As well as varying between disciplines, the gender ratio of authors has
changed markedly over time (Holman et al. 2018). Because the gender
ratio was more male-biased in the past, \(\alpha'\) would be inflated if
we calculated it for a sample of papers published over a long time
frame. To minimise this effect, we only sampled papers from two one-year
periods (namely 2005-6 and 2015-16). The median change per year in \%
(fe)male authors across journals is \textless{} 0.5\% (Holman et al.
2018), and so restricting our dataset to a single year should prevent
temporal changes in gender ratio from noticeably inflating our estimates
of \(\alpha'\).

\subsection{Minimising the Wahlund effect: author country of
affiliation}\label{minimising-the-wahlund-effect-author-country-of-affiliation}

A Wahlund effect could arise even if one calculates \(\alpha'\) for a
single discipline and time period, because of variation in the gender
ratio of researchers from different countries. For example, Holman et
al. (2018) showed that authors based in Serbia are more than twice as
likely to be women as are authors based in Japan. Therefore, a dataset
containing a mix of papers from teams of authors based in these two
countries would probably contain an excess of same-sex coauthorships,
even if collaboration were random with respect to gender within each
country.

To address this, we also analysed every combination of journal and
author country of affiliation for which we had enough data (i.e.~50 or
more papers published in 2015-16). For simplicity, we restricted the
dataset to only include papers for which Holman et al. (2018) had
identified the country of affiliation for all authors on the paper, and
all authors shared the same country of affiliation. Restricting the
dataset in this fashion produced enough data to measure \(\alpha'\) for
325 combinations of journal and country (median: 70 papers and 273
authors per combination).

\subsection{Calculating standardised journal impact
factor}\label{calculating-standardised-journal-impact-factor}

We obtained the 3-year impact factor for each journal from Clarivate
Analytics. To account for large differences in impact factor between
disciplines, we took the the residuals from a model with \(Log_{10}\)
impact factor as the response and the research discipline of the journal
as a random effect. Thus, journals with a positive standardised impact
factor have a higher mean number of citations than the average for
journals in their discipline. We then used Spearman rank correlation to
test whether our estimates of gender homophily covaried with journal
impact factor.

\subsection{Statistical analysis}\label{statistical-analysis}

Previous authors (e.g.~Bonham \& Stefan 2017) have hypothesised that
senior scientists preferentially recruit staff and students of the same
gender, and/or that junior researchers preferentially select
same-gendered mentors. In the majority of disciplines, authorship
conventions mean that the first-listed author is often an early-career
researcher, while the author listed last is more likely to be a senior
researcher leading a research team (Wren et al. 2007). Assuming that
senior researchers are the main drivers of homophily and that there are
enough papers with three or more authors, we predict that the last
author???s gender will be the strongest predictor of the remaining
authors??? genders (i.e.~the gender of the last author will be more
salient than that of the first author, or any other authorship
position). This is because the first author???s gender would simply be
an imperfect correlate of the true causal effect, while the last
author???s gender would be the causal effect itself.

To test whether \(\alpha'\) for last authors tends to be higher than
\(\alpha'\) for first authors for any given dataset, we used a linear
mixed model implemented in the \texttt{lme4} and \texttt{lmerTest}
packages for R, with \emph{authorship position} (first or last) as a
fixed factor, and \emph{journal} and \emph{research discipline} as
crossed random effects. The response variable was \(\alpha'\), and we
weighted each observation by the inverse of the standard error from our
estimate of \(\alpha'\), meaning that more accurate measurements of
\(\alpha'\) had more influence on the results. We used a similar model
to test for a difference in \(\alpha'\) between the 2005-6 and the
2015-16 datasets, with two differences: instead of authorship position,
we fit year range as a two-level fixed factor, and used \(\alpha'\)
estimated for all authors (not first/last authors) as the response
variable.

The relationship between the gender ratio of authors publishing in a
journal and its \(\alpha'\) value appeared nonlinear (Figure 4). We
therefore fit a generalised additive model with thin plate regression
spline smoothing, implemented using the \texttt{mgcv} package for R.

\section{Results}\label{results}

\subsection{Gender homophily by discipline, time period, and authorship
position}\label{gender-homophily-by-discipline-time-period-and-authorship-position}

\autoref{alpha_histograms} shows the distribution of \(\alpha'\)
estimates in 2015-2016 across all journals for which we recovered
sufficient data, when \(\alpha'\) was calculated for all authors, first
authors only, or last authors only. The great majority of journals had
\(\alpha'\) \textgreater{} 0, and for many of these the corrected
two-tailed p-values suggested that \(\alpha'\) was significantly greater
than zero (1469/2077 journals were significant in 2015-16, and 404/1192
in 2005-6). Only 2/2077 journals had statistically significantly
heterophily (i.e. \(\alpha'\) \textless{} 0) in 2015-16, and 1/1192 in
2005-6. The remaining 606 or 787 journals (in 2015 and 2005
respectively) had a value of \(\alpha'\) not significantly different
from zero, such that we could not reject the null hypothesis of random
assortment with respect to gender.

\(\alpha'\) was significantly higher in the sample from 2015-16 relative
to 2005-6, but the difference in means was small (Figure S3; Effect of
the fixed factor ???time period??? in a linear mixed model of the data
for all author positions: Cohen???s \(d\) = 0.09??0.04, \(t_{948}\) =
2.51, p = 0.012).

When comparing pairs of \(\alpha'\) values estimated for the first and
last authors for the same journals, we found that \(\alpha'\) tended to
be higher for first authors than for last authors (Figure S4; Effect of
the fixed factor ???Authorship position??? in a linear mixed model:
Cohen???s \(d\) = 0.07??0.02, \(t_{1988}\) = 4.48, p \textless{}
0.0001). This suggests that the gender of the first author was a
slightly stronger predictor of the remaining authors??? genders than the
gender of the last author, i.e.~the opposite of our prediction based on
the hypothesis that senior scientists are causally responsible for
homophily.

\subsection{Possible differences in the strength of homophily between
disciplines}\label{possible-differences-in-the-strength-of-homophily-between-disciplines}

\autoref{alpha_histograms} illustrates the variance in journal homophily
values (\(\alpha'\)) across scientific disciplines. All disciplines had
a positive average \(\alpha'\), although homophily appeared somewhat
stronger in some disciplines than others (e.g.~mean \(\alpha'\) was
0.12??0.02 for Urology journals, and 0.03??0.01 for Veterinary Medicine
journals; \autoref{alpha_histograms}, Table S3). However, there was
little evidence for strong differences in \(\alpha'\) between
disciplines: the random factor ???discipline??? explained at most 1\% of
the variance in \(\alpha'\) in the two linear mixed models described in
the previous section (see \autoref{alpha_histograms} and mixed models in
Online Supplementary Material). This implies that the processes causing
positive \(\alpha'\) are similarly strong in all the disciplines we
examined.

There was no indication that journals publishing on a wide range of
topics have higher \(\alpha'\) values than more specialised journals,
due to the Wahlund effect. For example, the journal category
???Multidisciplinary??? - which includes journals like \emph{PLoS ONE},
\emph{Nature}, \emph{Science}, and \emph{PNAS} - did not have notably
elevated \(\alpha'\) (\autoref{alpha_histograms}). This result suggests
that our estimates of homophily, and estimates from earlier studies, are
not notably inflated by the presence of disparate research topics (with
variable author gender ratios) being published within individual
journals.

\subsection{Relationship between gender homophily and gender
ratio}\label{relationship-between-gender-homophily-and-gender-ratio}

We next tested whether researchers are more or less likely to seek out
same-gendered colleagues in strongly gender-biased disciplines
(e.g.~Surgery or Nursing), relative to disciplines with a comparatively
gender-balanced workforce (e.g.~Psychiatry). We found a positive,
non-linear relationship between the overall gender ratio of all authors
publishing in a particular journal (as estimated by Holman et al. 2018),
and the estimated value of \(\alpha'\) for all authors and for first
authors (\autoref{alpha_gender_ratio}). Journals with a balanced or
female-biased author gender ratio tended to have higher \(\alpha'\) than
journals with a male-biased author gender ratio (GAM smooth terms p
\textless{} 0.001, see Online Supplementary Material). The relationship
was not statistically significant when \(\alpha'\) was calculated for
last authors (GAM, p = 0.142), though the trend appeared similar
(\autoref{alpha_gender_ratio}).

\subsection{Relationship between journal impact factor and gender
homophily}\label{relationship-between-journal-impact-factor-and-gender-homophily}

We observed a noisy but statistically significant linear relationship
between discipline-standardised journal impact factor and \(\alpha'\),
such that journals with a high impact factor for their discipline had
weaker gender homophily than did journals with a low impact factor for
their discipline (\autoref{impact_factor}; linear regression: \(R^2\) =
0.043, \(t_{1415}\) = -8.0, p \textless{} 0.0001).

\subsection{Analysis correcting for differences between
countries}\label{analysis-correcting-for-differences-between-countries}

When we restricted the analysis to only include authors with
affiliations from a single country, we found statistically significant
homophily for 72 of the 325 journal-country combinations tested (64
unique journals and 18 unique countries), and no significant evidence of
heterophily (Figures S5-S6). Additionally, the values of \(\alpha'\)
calculated for each journal-country combination were mostly very similar
to the \(\alpha'\) values calculated for the journal as a whole
(i.e.~when pooling papers from different countries); the average
difference in \(\alpha'\) was 0.002 (Figure S7). These results suggest
that our findings of widespread homophily in the main analysis cannot
solely be driven by a Wahlund effect resulting from gender differences
between countries.

\section{Discussion}\label{discussion}

We found evidence that researchers tend to publish with same-gendered
coauthors, despite implementing stringent controls for ???spurious???
homophily resulting from Wahlund effects (\autoref{wahlund_plot}). Our
data therefore reaffirm earlier conclusions (Ferber and Teiman 1980;
McDowell and Smith 1992; Crow and John Ortiz (2015); Ghiasi et al. 2015;
Zettler et al. 2016; Fahmy and Young 2016; Jadihi et al. 2017; Araujo
and Fontainha 2017a; Teele and Thelen 2017) that research collaborations
occur between same-gendered colleagues substantially more often than
expected if collaborators are selected without regard to their gender.
Many journals had \(\alpha'\) values over 0.1, meaning that the gender
ratio of the average man's coauthors is skewed towards men by
\textgreater{}10\%, relative to the gender ratio of a woman's coauthors.
Few journals had \(\alpha'\) values below zero, and fewer still had
significantly negative \(\alpha'\) values, suggesting that it is
comparatively rare for researchers to preferentially select
opposite-gendered collaborators.

We found no evidence that gender homophily is restricted to particular
disciplines: instead, there was evidence of homophily across many
diverse research disciplines. We also found that gender homophily was
marginally stronger in 2015-2016, relative to 2005-2006. This result
suggests that recent increases in the number of women working in STEMM,
as well as recent efforts to recruit, retain, and celebrate women in
STEMM, have co-occurred with rise in the frequency of same-gender
collaborations, not a fall. We speculate that this trend might result,
at least in part, from the increasing number of women working in senior
positions in STEMM over the last 10 years (DATA?). Because many
co-authorships involve a junior and a senior researcher (e.g.~a student
and professor), and the gender ratio is typically more skewed among
senior staff (Shaw and Stanton ref), there will be more opposite-gender
collaborations than expected if the gender ratio were the same at both
career stages. Thus, the observed increase in homophily might be due to
the greater numbers of women working in leading roles.

Gender homophily was slightly stronger in research disciplines with a
female-biased gender ratio, as compared to male-biased or
gender-balanced fields. One interpretation of this result is that men
are more likely to seek out other men in fields in which men are in the
minority, relative to the homophily shown by women in fields in which
women are in the minority.

We also found that journals with a high impact factor (relative to other
journals in the same research discipline) tended to have weaker gender
homophily than did low-impact journals. If one accepts the contentious
claim that papers published in high-impact journals tend to be of higher
quality than those in low-impact journals (e.g. Garfield 2006), then
this result provides correlational support for the hypothesis that
mixed-gender teams produce better research than single-gender teams
(Campbell et al. 2013).

Contrary to our expectations, we found no evidence that homophily is
driven mostly by senior academics, and no evidence that homophily is
strongest in fields where women are in the minority. Interestingly,
homophily was negatively correlated with journal impact factor
(standardised by research discipline), as predicted if mixed-gender
teams produce better research.

Academic publishing is one of the most (if not the most) important tool
for a researcher to display their work, and is frequently used as a
success indicator for being promoted, hired as a faculty member, or
being invited as plenary speaker. Our finding that researchers are more
likely to publish with same-gendered coauthors implies that
co-authorship listing is not a random process, but instead creates
additional gender imbalance. Homophily could make it harder for women to
find collaborators in disciplines where women are in the minority
(i.e.~almost all STEMM fields; Holman et al. 2018). Because women make
up for as little as 30\% of the STEMM scientist community (REF), they
are therefore less likely to be picked as coauthors, and, as a result,
gender homophily creates further gender imbalance and disadvantages for
women. Men tend to publish more in high impact journals (REF) as first
author, and generate more citations than papers lead by first author
women (REF). As a consequence, gender homophily is also likely to
contribute to lower publication rates for women (McDowell and Smith
1992), and lower scientific success (REF).

Why does gender homophily exists in science? Collaboration often leads
to co-authorship publications, citations and additional professional
recognitions, therefore first author should always be considering
collaborators that are most likely to enhance their productivity and in
the long term also their academic success, rather than pick co-authors
according to their gender. First of all, it is likely that decades ago,
the probability to meet a woman colleague working on the same area of
expertise was slimmer than now. However, the number of women in STEM is
increasing, and it is not likely to be an obstacle to collaboration
nowadays.

\section{Acknowledgements}\label{acknowledgements}

We are grateful to XXXX.

\section{References}\label{references}

\hypertarget{refs}{}
\hypertarget{ref-Crow_2015}{}
Crow, M. S., and S. John Ortiz. 2015. The effect of gender‐sorting on
propensity to coauthor: Implications for academic promotion. American
Journal of Criminal Justice 40:441--455.

\hypertarget{ref-Elsevier_report}{}
Elsevier Report. 2017. Gender in the global research landscape.
elsevier.com/research-intelligence/resource-library/gender-report.

\hypertarget{ref-Garfield_2006}{}
Garfield, E. 2006. The history and meaning of the journal impact factor.
JAMA 295:90--93.

\hypertarget{ref-Holman_2018}{}
Holman, L., D. Stuart Fox, and C. E. Hauser. 2018. The gender gap in
science: How long until women are equally represented? PLoS Biology
16:e2004956.

\hypertarget{ref-Lariviere_2013}{}
Larivière, V., C. Ni, Y. Gingras, B. Cronin, and C. R. Sugimoto. 2013.
Bibliometrics: global gender disparities in science. Nature
504:211--213.

\hypertarget{ref-West_2013}{}
West, J. D., J. Jacquet, M. M. King, S. J. Correll, and C. T. Bergstrom.
2013. The role of gender in scholarly authorship. PLoS ONE 8:e66212.

\newpage

\section{Figures}\label{figures}

\begin{figure}
\centering
\includegraphics{../figures/Figure 1 - Wahlund figure.pdf}
\caption{The Wahlund effect can make it appear as if authors prefer to
publish with same-gendered colleagues, even if no such preference
exists. Here, coloured circles represent male and female authors, and
coauthors are linked with lines. Across the whole set of ten papers,
there is an apparent excess of same-gender collaborations. Specifically,
there are six same-gender papers and only four mixed-gender papers,
which is fewer than the \(10\times2\times0.5\times0.5 = 5\) mixed-gender
papers we would expect under the null hypothesis that authors assort
randomly with respect to gender. However, within each subset, there is
no evidence that authors prefer to publish with same-gendered
individuals. The Wahlund effect will tend to inflate the frequency of
same-sex coauthors whenever the data is composed of two or more
disconnected subsets of literature with different author gender ratios;
these subsets could be research disciplines, older versus newer papers,
or papers from authors in different countries. \label{wahlund_plot}}
\end{figure}

\newpage

\begin{figure}
\centering
\includegraphics{../figures/figure2.pdf}
\caption{Of the 2077 journals for which we had adequate data in
2015-2016, 830 showed statistically significant evidence of homophily
(denoted by \(\alpha' > 0\)), and 1 showed statistically significant
evidence of heterophily (\(\alpha' < 0\)), after adjusting p-values
using Benjamini-Hochberg false discovery rate correction. The white area
shows the number of journals for which homophily was significantly
stronger than expected under the null hypothesis (p \textless{} 0.05),
while the blue area shows all the remainder. Patterns were similar
whether \(\alpha'\) was calculated for all authors, for first authors
only, or for last authors only. \label{alpha_histograms}}
\end{figure}

\newpage

\begin{figure}
\centering
\includegraphics{../figures/figure3.pdf}
\caption{There is a weakly positive, non-linear relationship between the
gender ratio of authors publishing in a journal, and the coefficient of
homophily (\(\alpha'\)). Specifically, journals with 50\% women authors
or higher tended to have more same-sex coauthorships than did journals
with predominantly men authors. This relationship held whether
\(\alpha'\) was calculated for all authors, first authors only, or last
authors only. A negative value on the x-axis denotes an excess of men
authors, a positive value denotes an excess of women authors, and zero
denotes gender parity. The lines were fitted using generalised additive
models with the smoothing parameter \(k\) set to 3.
\label{alpha_gender_ratio}}
\end{figure}

\newpage

\begin{figure}
\centering
\includegraphics{../figures/figure4.pdf}
\caption{Journal impact factor (expressed relative to the average for
the discipline) is negatively correlated with \(\alpha'\). The
relationship is noisy (\(R^2\) = 0.043), but the results suggest that
journals with strong homophily tend to have lower impact factors than
journals with weak homophily in the same discipline.
\label{impact_factor}}
\end{figure}

\end{document}

### 2018-05-22 manuscript/Manuscript_Claire.Rmd
Manuscript_Claire.Rmd

--- 2018-05-01 manuscript/Manuscript.Rmd
@@
 abstract: Evidence suggests that women in academia are hindered by conscious and unconscious biases, and that many female researchers feel excluded from formal and informal opportunities for research collaboration.
-In addition to ensuring fairness and helping redress gender imbalance in the academic workforce, increasing women’s access to collaboration could help scientific progress by drawing on more of the available human capital.
+In addition to ensuring fairness and helping redress gender imbalance in the academic workforce, increasing women???s access to collaboration could help scientific progress by drawing on more of the available human capital.
 Here, we test whether researchers preferentially collaborate with same-gendered colleagues, using more stringent methods and a larger dataset than in past work.
-Our results reaffirm that researchers co-publish with colleagues of the same gender, and show that this ‘gender homophily’ is slightly stronger today than it was 10 years ago.
+Our results reaffirm that researchers co-publish with colleagues of the same gender, and show that this ???gender homophily???
+is slightly stronger today than it was 10 years ago.
 Contrary to our expectations, we found no evidence that homophily is driven mostly by senior academics, and no evidence that homophily is strongest in fields where women are in the minority.
@@
 @Lariviere_2013; @West_2013; @Elsevier_report; @Holman_2018].
-On average, women occupy more junior positions (Wutte 2007, Reuben, Sapienza & Zingales, 2014) with lower salaries (Trower and Chait 2002, Umback 2007), receive less grant money (Hosek et al.
-2005, OER 2005), are promoted more slowly (Hopkins et al.
-2013; Long et al.
-1993, Rosenfeld 1991, Zuckerman 1987), and are allocated fewer resources (European Commission 2009) and less research funding (Feldt 1986, Stack 2004, Larivière et al.
-2011) than men.
-Experimental studies have demonstrated that researchers regard women's achievements less favourably than identical achievements by men (Moss et al.
-2012; Knobloch et al.
-2013).
+On average, women occupy more junior positions [@Wutte_2007; @Reuben_2014] with lower salaries [@Trower_2002; @Umbach_2007], receive less grant money [@Hosek_2005; @OER_2008], are promoted more slowly [@Zuckerman_1987; @Rosenfeld_1991; @Long_1993; @Hopkins_2013], and are allocated fewer resources [@ODorchai_2009] and less research funding [@Feldt_1986, @Stack_2004, @Lariviere_2011] than men.
+Experimental studies have demonstrated that researchers regard women's achievements less favourably than identical achievements by men [@Moss_2012; @Knobloch_2013].
 
-Because publishing, networking and collaboration are all instrumental to academic career advancement (Lee and Boheman 2005, Hopkins et al.
-2012), numerous studies have tested for gender differences in these areas (reviewed in Table S1 in Holman et al.
-2018; see also XXX).
-For example, studies have shown that relative to men, women tend to be less involved in international collaboration (Lemoine 1995; Lewison 2001; Webster 2001; Larivière et al.
-2011, Bozeman and Corley 2004), collaborate less with researchers from their own university department (Websters 2001), have less prestigious collaborations (Long 1990), and fewer collaborations in general (Fuchs Von Stebut and Allmendinger 2001).
-Possible explanations for the smaller average professional networks of women include a greater average amount of family obligations (Long 1990, Wright et al.
-2003, Reskin 1978), lower participation of women in international conferences (Lewison 2001, Eisen 2014, Martin 2014), and reduced receipt of research travel funds by women (Bozeman and Corley 2004).
+Because publishing, networking and collaboration are all instrumental to scientific productivity and thus, academic career advancement [@Lee_2005; @Wuchty_2007;@Abramo_2009;@Lariviere_2015], numerous studies have tested for gender differences in these areas [reviewed in Table S1 in @Holman_2018; see also @Bozeman_2011; @Abramo_2013; @Badar_2013; @Long_1992].
+For example, studies have shown that relative to men, women tend to be less involved in international collaboration [@Lewison_2001; @Webster_2001; @Bozeman_2004; @Lariviere_2011; @Abramo_2013], collaborate less with researchers from their own university department [@Webster_2001], have less prestigious collaborations [@Long_1990], and fewer collaborations in general [@Fuchs_2001].
+Possible explanations for the smaller average professional networks of women include a greater average amount of family obligations [@Reskin_1978; @Long_1990; @Wright_2003], lower participation of women in international conferences [@Lewison_2001; Eisen 2014; @Martin_2014], and reduced receipt of research travel funds by women [@Bozeman_2004].
 
-A high, steadily increasing proportion of research papers is written by more than one author (West et al.
-2013), making collaboration a key predictor of publication output, and in turn, a key predictor of career prospects (Cohn et al 2014, Jordan et al.
-2008, Tower et al.
-2007).
-Additionally, empirical studies suggest that mixed-gender (Campbell et al.
-2013) or otherwise diverse (Hong and Page 2004) teams can produce better results on collaborative tasks.
+A high, steadily increasing proportion of research papers is written by more than one author [@West_2013], making collaboration a key predictor of publication output, and in turn, a key predictor of career prospects [@Tower_2007; @Jordan_2008; @Cohn_2014].
+Additionally, empirical studies suggest that mixed-gender [@Bear_2011; @Campbell_2013] or otherwise diverse [@Hong_2004] teams can produce better results on collaborative tasks.
 For reasons such as these, several studies have tested for gender differences in collaboration frequency or pattern by examining the author lists of published research.
-To our knowledge, every study of this question has concluded that men publish with other men, and women with women, more often than expected if collaborators assort randomly with regards to gender (Ferber and Teiman, 1980; McDowell and Smith 1992; Ghiasi et al.
-2015; Crow and Smykla 2015, Zettler et al.
-2016, Fahmy and Young, 2016, Jadidi et al.
-2017; Teele and Thelen 2017, Araujo and Fontainha 2017a,b).
-This preference for same-gendered colleagues is often termed ‘gender homophily’.
+To our knowledge, every study of this question has concluded that men publish with other men, and women with women, more often than expected if collaborators assort randomly with regards to gender [@Ferber_1980; @McDowell_1992; @Ghiasi_2015; @Crow_2015;@Fahmy_2017; @Zettler_2017;@Jadidi_2017; @Teele_2017; @Araujo_2017a; @Araujo_2017b].
+This preference for same-gendered colleagues is often termed ???gender homophily???.
 
@@
 For example, a sample of bioinformatics and cell biology papers might contain an excess of mostly-male and mostly-female author lists, simply because researchers preferentially work with colleagues from the same discipline, and because more women work in cell biology than in bioinformatics [@Holman_2018].
-We name this issue after the Wahlund effect in population genetics (Wahlund 1928), whereby spatial differences in allele frequencies create an excess of homozygotes across the whole population, even if alleles assort randomly within each subpopulation.
+We name this issue after the Wahlund effect in population genetics [@Wahlund_1928], whereby spatial differences in allele frequencies create an excess of homozygotes across the whole population, even if alleles assort randomly within each subpopulation.
 
@@
 Each journal was assigned to one of 107 scientific disciplines (e.g.
-“Nephrology” or “Cell Biology”), using PubMed’s journal categorisations where available.
+???Nephrology???
+or ???Cell Biology???), using PubMed???s journal categorisations where available.
 Because the present study focuses on co-authorship, all single-author papers were removed from the dataset.
@@
 
-To estimate $\alpha$ for a particular subset of the scientific literature, we estimated $p$ as the average proportion of men’s co-authors who are men (averaged across all papers with at least one man author), and $q$ as the average proportion of women’s co-authors who are men (averaged across all papers with at least one woman author).
+To estimate $\alpha$ for a particular subset of the scientific literature, we estimated $p$ as the average proportion of men???s co-authors who are men (averaged across all papers with at least one man author), and $q$ as the average proportion of women???s co-authors who are men (averaged across all papers with at least one woman author).
 To estimate the 95% confidence intervals on $\alpha$ for a given set of $n$ papers, we sampled $n$ papers with replacement 1000 times, estimated $\alpha$ on each sample, and recorded the 95% quantiles of the 1000 resulting estimates.
@@
 To borrow Prof.
-Bergstrom’s example, consider a small research specialty comprising just two men and two women researchers, who have together produced six two-author papers: one paper in each of the six possible two-author combinations.
+Bergstrom???s example, consider a small research specialty comprising just two men and two women researchers, who have together produced six two-author papers: one paper in each of the six possible two-author combinations.
 For these six papers, $\alpha$ = $-\frac{1}{3}$, even though same- and opposite-gendered coauthors were selected in equal proportion to their frequency in the pool of possible collaborators.
@@
 We also used the null-simulated $\alpha$ values to calculate a two-tailed p-value for the observed value of $\alpha$; the p-value was defined as the proportion of null simulations for which $|\alpha_{null}| > |\alpha_{obs}|$.
-We applied false discovery rate correction to each set of p-values to account for multiple testing (Benjamini and Hochberg 1995).
+We applied false discovery rate correction to each set of p-values to account for multiple testing [@Benjamini_1995].
 
@@
 ## Minimising the Wahlund effect: discipline and time period
-As discussed in the Introduction, the Wahlund effect can give the appearance of gender homophily (reflected in our study as $\alpha'$ > 0) whenever one lumps together different ‘subpopulations’ of academic papers that differ in their author gender ratios (Figure 1).
+As discussed in the Introduction, the Wahlund effect can give the appearance of gender homophily (reflected in our study as $\alpha'$ > 0) whenever one lumps together different ???subpopulations???
+of academic papers that differ in their author gender ratios (Figure 1).
 
@@
 Previous authors (e.g.
-Bonham & Stefan 2017) have hypothesised that senior scientists preferentially recruit staff and students of the same gender, and/or that junior researchers preferentially select same-gendered mentors.
-In the majority of disciplines, authorship conventions mean that the first-listed author is often an early-career researcher, while the author listed last is more likely to be a senior researcher leading a research team (Wren et al.
-2007).
-Assuming that senior researchers are the main drivers of homophily and that there are enough papers with three or more authors, we predict that the last author’s gender will be the strongest predictor of the remaining authors’ genders (i.e.
+@Bonham_2017) have hypothesised that senior scientists preferentially recruit staff and students of the same gender, and/or that junior researchers preferentially select same-gendered mentors.
+In the majority of disciplines, authorship conventions mean that the first-listed author is often an early-career researcher, while the author listed last is more likely to be a senior researcher leading a research team [@Wren_2007].
+Assuming that senior researchers are the main drivers of homophily and that there are enough papers with three or more authors, we predict that the last author???s gender will be the strongest predictor of the remaining authors???
+genders (i.e.
 the gender of the last author will be more salient than that of the first author, or any other authorship position).
-This is because the first author’s gender would simply be an imperfect correlate of the true causal effect, while the last author’s gender would be the causal effect itself.
+This is because the first author???s gender would simply be an imperfect correlate of the true causal effect, while the last author???s gender would be the causal effect itself.
 
@@
 
-$\alpha'$ was significantly higher in the sample from 2015-16 relative to 2005-6, but the difference in means was small (Figure S3; Effect of the fixed factor ‘time period’ in a linear mixed model of the data for all author positions: Cohen’s $d$ = 0.09±0.04, $t_{948}$ = 2.51, p = 0.012).
+$\alpha'$ was significantly higher in the sample from 2015-16 relative to 2005-6, but the difference in means was small (Figure S3; Effect of the fixed factor ???time period???
+in a linear mixed model of the data for all author positions: Cohen???s $d$ = 0.09??0.04, $t_{948}$ = 2.51, p = 0.012).
 
-When comparing pairs of $\alpha'$ values estimated for the first and last authors for the same journals, we found that $\alpha'$ tended to be higher for first authors than for last authors (Figure S4; Effect of the fixed factor ‘Authorship position’ in a linear mixed model: Cohen’s $d$ = 0.07±0.02, $t_{1988}$ = 4.48, p < 0.0001).
-This suggests that the gender of the first author was a slightly stronger predictor of the remaining authors’ genders than the gender of the last author, i.e.
+When comparing pairs of $\alpha'$ values estimated for the first and last authors for the same journals, we found that $\alpha'$ tended to be higher for first authors than for last authors (Figure S4; Effect of the fixed factor ???Authorship position???
+in a linear mixed model: Cohen???s $d$ = 0.07??0.02, $t_{1988}$ = 4.48, p < 0.0001).
+This suggests that the gender of the first author was a slightly stronger predictor of the remaining authors???
+genders than the gender of the last author, i.e.
 the opposite of our prediction based on the hypothesis that senior scientists are causally responsible for homophily.
@@
 All disciplines had a positive average $\alpha'$, although homophily appeared somewhat stronger in some disciplines than others (e.g.
-mean $\alpha'$ was 0.12±0.02 for Urology journals, and 0.03±0.01 for Veterinary Medicine journals; \autoref{alpha_histograms}, Table S3).
-However, there was little evidence for strong differences in $\alpha'$ between disciplines: the random factor ‘discipline’ explained at most 1% of the variance in $\alpha'$ in the two linear mixed models described in the previous section (see \autoref{alpha_histograms} and mixed models in Online Supplementary Material).
+mean $\alpha'$ was 0.12??0.02 for Urology journals, and 0.03??0.01 for Veterinary Medicine journals; \autoref{alpha_histograms}, Table S3).
+However, there was little evidence for strong differences in $\alpha'$ between disciplines: the random factor ???discipline???
+explained at most 1% of the variance in $\alpha'$ in the two linear mixed models described in the previous section (see \autoref{alpha_histograms} and mixed models in Online Supplementary Material).
 This implies that the processes causing positive $\alpha'$ are similarly strong in all the disciplines we examined.
@@
 There was no indication that journals publishing on a wide range of topics have higher $\alpha'$ values than more specialised journals, due to the Wahlund effect.
-For example, the journal category “Multidisciplinary” - which includes journals like _PLoS ONE_, _Nature_, _Science_, and _PNAS_ - did not have notably elevated $\alpha'$ (\autoref{alpha_histograms}).
+For example, the journal category ???Multidisciplinary???
+- which includes journals like _PLoS ONE_, _Nature_, _Science_, and _PNAS_ - did not have notably elevated $\alpha'$ (\autoref{alpha_histograms}).
 This result suggests that our estimates of homophily, and estimates from earlier studies, are not notably inflated by the presence of disparate research topics (with variable author gender ratios) being published within individual journals.
@@
 # Discussion
-We found evidence that researchers tend to publish with same-gendered coauthors, despite implementing stringent controls for ‘spurious’ homophily resulting from Wahlund effects (\autoref{wahlund_plot}).
-Our data therefore reaffirm earlier conclusions (Ferber and Teiman 1980; McDowell and Smith 1992; Crow and Smykla 2015; Ghiasi et al.
-2015; Zettler et al.
-2016; Fahmy and Young 2016; Jadihi et al.
-2017; Araujo and Fontainha 2017a; Teele and Thelen 2017) that research collaborations occur between same-gendered colleagues substantially more often than expected if collaborators are selected without regard to their gender.
+We found evidence that researchers tend to publish with same-gendered coauthors, despite implementing stringent controls for ???spurious???
+homophily resulting from Wahlund effects (\autoref{wahlund_plot}).
+Our data therefore reaffirm earlier conclusions [@Ferber_1980; @McDowell_1992; @Crow_2015; @Ghiasi_2015; @Araujo_2017a; @Fahmy_2017; @Jadidi_2017; @Zettler_2017; @Teele_2017] that research collaborations occur between same-gendered colleagues substantially more often than expected if collaborators are selected without regard to their gender.
 Many journals had $\alpha'$ values over 0.1, meaning that the gender ratio of the average man's coauthors is skewed towards men by >10%, relative to the gender ratio of a woman's coauthors.
@@
 
-We found no evidence that gender homophily is restricted to particular disciplines: instead, there was evidence of homophily across many diverse research disciplines.
+We hypothesized that disciplines with strong gender-biases (e.g.
+physics and computer science for men and nursing, midwifery for women [@Holman_2018]) would display a higher level of gender-homophily due to the increased difficulty to find an opposite-gender co-author.
+However, we found no evidence that gender homophily is restricted to particular disciplines: instead, there was evidence of homophily across many diverse research disciplines (FIG).
+Gender homophily is thus a global concern rather than an artifact of the existing gender gap (REF).
+Additionally, we found that gender homophily was slightly stronger in research disciplines with a female-biased gender ratio, as compared to male-biased or gender-balanced fields (whether it is in the first or last author position).
+One interpretation of this result is that men are more likely to seek out other men in fields in which men are in the minority, relative to the homophily shown by women in fields in which women are in the minority.
+This male-driven homophily could be, in part, due to the shortage of opportunities for women to participate in international research projects (@Elsevier report),  or international conferences (REF) and therefore less likely to collaborate across gender or fields.
+
 We also found that gender homophily was marginally stronger in 2015-2016, relative to 2005-2006.
 This result suggests that recent increases in the number of women working in STEMM, as well as recent efforts to recruit, retain, and celebrate women in STEMM, have co-occurred with rise in the frequency of same-gender collaborations, not a fall.
-We speculate that this trend might result, at least in part, from the increasing number of women working in senior positions in STEMM over the last 10 years (DATA?).
+We speculate that this trend might result, at least in part, from the increasing number of women working in senior positions in STEMM over the last 10 years [@Bendels_2018; @Long_2015; @McKenzie_2017].
+Coupled with an ever increase proportion of women in the first author position [@West_2013], this trend leads to an increase in same-sex collaboration.
 Because many co-authorships involve a junior and a senior researcher (e.g.
-a student and professor), and the gender ratio is typically more skewed among senior staff (Shaw and Stanton ref), there will be more opposite-gender collaborations than expected if the gender ratio were the same at both career stages.
-Thus, the observed increase in homophily might be due to the greater numbers of women working in leading roles.
-
-Gender homophily was slightly stronger in research disciplines with a female-biased gender ratio, as compared to male-biased or gender-balanced fields.
-One interpretation of this result is that men are more likely to seek out other men in fields in which men are in the minority, relative to the homophily shown by women in fields in which women are in the minority.
+a student and professor), and the gender ratio is typically more skewed among senior staff [@Shaw_2012, @West_2013], there will be more opposite-gender collaborations than expected if the gender ratio were the same at both career stages.
+Thus, the observed increase in homophily might partially be due to the greater numbers of women working in leading roles.
+However, gender homophily during collaboration and co-authoring of scientific manuscript may also impact negatively women careers as there is still a smaller proportion of women in science and thus, making it harder to find female collaborators (Bentley and Adamson 2003).
 
@@
 If one accepts the contentious claim that papers published in high-impact journals tend to be of higher quality than those in low-impact journals [e.g.
-@Garfield_2006], then this result provides correlational support for the hypothesis that mixed-gender teams produce better research than single-gender teams (Campbell et al.
-2013).
-
+@Garfield_2006], then this result provides correlational support for the hypothesis that mixed-gender teams produce better research than single-gender teams [@Campbell_2013].
+However, journal impact factors tend to be negatively associated with the proportion of women [@Holman_2018.
+REF].
 Contrary to our expectations, we found no evidence that homophily is driven mostly by senior academics, and no evidence that homophily is strongest in fields where women are in the minority.
@@
 almost all STEMM fields; @Holman_2018].
-Because women make up for as little as 30% of the STEMM scientist community (REF), they are therefore less likely to be picked as coauthors, and, as a result, gender homophily creates further gender imbalance and disadvantages for women.
-Men tend to publish more in high impact journals (REF) as first author, and generate more citations than papers lead by first author women (REF).
-As a consequence, gender homophily is also likely to contribute to lower publication rates for women (McDowell and Smith 1992), and lower scientific success (REF).
+Because women make up for as little as 30% of the STEMM scientist community [@Holman_2018], they are therefore less likely to be picked as coauthors, and, as a result, gender homophily creates further gender imbalance and disadvantages for women.
+Men tend to publish more in high impact journals as first author [@West_2013], and generate more citations than papers lead by first author women [@Long_1992; @Lariviere_2013].
+And as a consequence, gender homophily is also likely to contribute to lower publication rates for women [@McDowell_1992], and lower scientific success (REF).
 
-Why does gender homophily exists in science?
+Although the proportion of women in science has significantly increased over the last decades (REF), women still tend to leave their academic careers behind considerably more than their male counterpart (D??barre et al 2018).
+Lack of scientific recognition (Lincolm et al- 2012), research funding (Bozeman and Corley 2004), travel opportunities, invitation to speak at scientific events (Isbell et al.
+2012; Schroeder et al- 2013; Klein et al- 2017), discrimination during hiring (Moss-Racusin et al- 2012; Reuben et al.
+2014), when writing a recommendation letter (Trix and Psenka 2003; Schmader et al- 2007), or personal reasons (e.g.
+child rearing or family ties (Zimmer et al- 2006)), all in all this concentration of small disadvantages drives women out of science.
+Whether their decision to leave academia is deliberate or not, it will influence the future of the scientific community, as mentoring, networking, and supporting younger women scientists, are especially important for women in academia.
+
+Why does gender homophily exist in science?
 Collaboration often leads to co-authorship publications, citations and additional professional recognitions, therefore first author should always be considering collaborators that are most likely to enhance their productivity and in the long term also their academic success, rather than pick co-authors according to their gender.
@@
 # Acknowledgements
-We are grateful to XXXX.
+We are grateful to XXXX.This work was supported by the Academy of Finland (284666 to the Centre of Excellence in Biological Interactions).
 

### 2018-05-23 manuscript/Manuscript_Claire.Rmd
manuscript/

@@
 author:
-- Luke Holman* and Claire Morandin
+- Luke Holman* and Claire Morandin+
 - '*\textit{luke.holman@unimelb.edu.au} \vspace{5mm}'
@@
 \maketitle\unmarkedfntext{*School of BioSciences, The University of Melbourne, Victoria, Australia}
+\maketitle\unmarkedfntext{+Organismal and Evolutionary Biology Research Programme,
+Faculty of Biological and Environmental Sciences, University of Helsinki, Finland}
 \newpage
@@
 # Discussion
-We found evidence that researchers tend to publish with same-gendered coauthors, despite implementing stringent controls for ???spurious???
+We found evidence that researchers tend to publish more with same-gendered coauthors, despite implementing stringent controls for ???spurious???
 homophily resulting from Wahlund effects (\autoref{wahlund_plot}).
-Our data therefore reaffirm earlier conclusions [@Ferber_1980; @McDowell_1992; @Crow_2015; @Ghiasi_2015; @Araujo_2017a; @Fahmy_2017; @Jadidi_2017; @Zettler_2017; @Teele_2017] that research collaborations occur between same-gendered colleagues substantially more often than expected if collaborators are selected without regard to their gender.
+Our data therefore reaffirm earlier conclusions [@Ferber_1980; @McDowell_1992; @Bentley_2003; @Crow_2015; @Ghiasi_2015; @Araujo_2017a; @Fahmy_2017; @Jadidi_2017; @Teele_2017; @Zettler_2017] that research collaborations occur between same-gendered colleagues substantially more often than expected if collaborators are selected without regard to their gender.
 Many journals had $\alpha'$ values over 0.1, meaning that the gender ratio of the average man's coauthors is skewed towards men by >10%, relative to the gender ratio of a woman's coauthors.
@@
 
-We hypothesized that disciplines with strong gender-biases (e.g.
-physics and computer science for men and nursing, midwifery for women [@Holman_2018]) would display a higher level of gender-homophily due to the increased difficulty to find an opposite-gender co-author.
-However, we found no evidence that gender homophily is restricted to particular disciplines: instead, there was evidence of homophily across many diverse research disciplines (FIG).
-Gender homophily is thus a global concern rather than an artifact of the existing gender gap (REF).
-Additionally, we found that gender homophily was slightly stronger in research disciplines with a female-biased gender ratio, as compared to male-biased or gender-balanced fields (whether it is in the first or last author position).
+As some disciplines exhibit strong gender-biases (e.g.
+physics and computer science with predominantly men co-authors, and nursing, midwifery for women [@Holman_2018]), we hypothesized that these disciplines would also display a higher level of gender-homophily due to the increased difficulty to find an opposite-gender co-author.
+However, we found no evidence that gender homophily is restricted to particular disciplines: instead, there was evidence of homophily across many diverse research disciplines (Figure 2).
+This finding may be indicative that gender homophily in science is a comprehensive concern, rather than an artifact of the existing gender gap, and not limited to stereotypical disciplines.
+Furthermore, we found that gender homophily was slightly stronger in research disciplines with a female-biased gender ratio, as compared to male-biased or gender-balanced fields (whether it is in the first or last author position).
 One interpretation of this result is that men are more likely to seek out other men in fields in which men are in the minority, relative to the homophily shown by women in fields in which women are in the minority.
-This male-driven homophily could be, in part, due to the shortage of opportunities for women to participate in international research projects (@Elsevier report),  or international conferences (REF) and therefore less likely to collaborate across gender or fields.
 
@@
 We speculate that this trend might result, at least in part, from the increasing number of women working in senior positions in STEMM over the last 10 years [@Bendels_2018; @Long_2015; @McKenzie_2017].
-Coupled with an ever increase proportion of women in the first author position [@West_2013], this trend leads to an increase in same-sex collaboration.
 Because many co-authorships involve a junior and a senior researcher (e.g.
@@
 Thus, the observed increase in homophily might partially be due to the greater numbers of women working in leading roles.
-However, gender homophily during collaboration and co-authoring of scientific manuscript may also impact negatively women careers as there is still a smaller proportion of women in science and thus, making it harder to find female collaborators (Bentley and Adamson 2003).
 
-We also found that journals with a high impact factor (relative to other journals in the same research discipline) tended to have weaker gender homophily than did low-impact journals.
+Despite the fact that journal impact factors tend to be negatively associated with the proportion of women [@Holman_2018], we found that journals with a high impact factor (relative to other journals in the same research discipline) tended to have weaker gender homophily than did low-impact journals (Figure 4).
 If one accepts the contentious claim that papers published in high-impact journals tend to be of higher quality than those in low-impact journals [e.g.
 @Garfield_2006], then this result provides correlational support for the hypothesis that mixed-gender teams produce better research than single-gender teams [@Campbell_2013].
-However, journal impact factors tend to be negatively associated with the proportion of women [@Holman_2018.
-REF].
-Contrary to our expectations, we found no evidence that homophily is driven mostly by senior academics, and no evidence that homophily is strongest in fields where women are in the minority.
-Interestingly, homophily was negatively correlated with journal impact factor (standardised by research discipline), as predicted if mixed-gender teams produce better research.
 
-Academic publishing is one of the most (if not the most) important tool for a researcher to display their work, and is frequently used as a success indicator for being promoted, hired as a faculty member, or being invited as plenary speaker.
+Why does gender homophily exist in science?
+Collaboration often leads to co-authorship publications, citations and additional professional recognitions, therefore first author should always be considering collaborators that are most likely to enhance their productivity and in the long term also their academic success, rather than pick co-authors according to their gender.
+First of all, it is likely that decades ago, the probability to meet a woman colleague working on the same area of expertise was slimmer than now.
+Even though the number of women in STEM is increasing, and it is not likely to be an obstacle to collaboration nowadays; we cannot forgo that female representation in science is still not equal to that of males.
+Second, the most basic source of homophily is space, scientists are more likely to collaborate with colleagues closer to them in geographic location (@McPherson_2011).
+In the case of women, personal considerations (e.g.
+family commitments) represent the principal restrictions on their international mobility.
+As a results, women are less likely to originate collaborations with foreign colleagues [@Lemoine_1992; @Lewison_2001; @ Webster_2001; @ Lariviere_2011].
+Lastly, men and women think differently on average (REF), and are more likely to initiate collaborations based on similarity of thoughts to progress their scientific output (REF).
+
+Academic publishing is one of the most (if not the most) important tool for a researcher to display their work, and is frequently used as a success indicator for being promoted and hired as a faculty member.
 Our finding that researchers are more likely to publish with same-gendered coauthors implies that co-authorship listing is not a random process, but instead creates additional gender imbalance.
@@
 Men tend to publish more in high impact journals as first author [@West_2013], and generate more citations than papers lead by first author women [@Long_1992; @Lariviere_2013].
-And as a consequence, gender homophily is also likely to contribute to lower publication rates for women [@McDowell_1992], and lower scientific success (REF).
+And as a consequence, gender homophily is also likely to contribute to lower publication rates for women [@McDowell_1992].
 
-Although the proportion of women in science has significantly increased over the last decades (REF), women still tend to leave their academic careers behind considerably more than their male counterpart (D??barre et al 2018).
-Lack of scientific recognition (Lincolm et al- 2012), research funding (Bozeman and Corley 2004), travel opportunities, invitation to speak at scientific events (Isbell et al.
-2012; Schroeder et al- 2013; Klein et al- 2017), discrimination during hiring (Moss-Racusin et al- 2012; Reuben et al.
-2014), when writing a recommendation letter (Trix and Psenka 2003; Schmader et al- 2007), or personal reasons (e.g.
-child rearing or family ties (Zimmer et al- 2006)), all in all this concentration of small disadvantages drives women out of science.
-Whether their decision to leave academia is deliberate or not, it will influence the future of the scientific community, as mentoring, networking, and supporting younger women scientists, are especially important for women in academia.
+Although the proportion of women in science has significantly increased over the last decades, women still tend to leave their academic careers behind considerably more than their male counterpart [@Debarre_2018] .
+Lack of scientific recognition [@Lincoln_2012], lack of research funding [@Bozeman_2004], lack of invitation to speak at scientific events[@Isbell_2012; @Schroeder_2013; @Klein_2017], discrimination during hiring [@MossRacusin_2017; @Reuben_2014], discrimination when writing a recommendation letter [@Trix_2003; @Schmader_2007], all in all this concentration of small disadvantages drives women out of science.
+Whether their decision to leave academia is deliberate or not, it will negatively impact science overall, as mentoring, networking, and supporting younger women scientists are especially important for women in academia.
 
-Why does gender homophily exist in science?
-Collaboration often leads to co-authorship publications, citations and additional professional recognitions, therefore first author should always be considering collaborators that are most likely to enhance their productivity and in the long term also their academic success, rather than pick co-authors according to their gender.
-First of all, it is likely that decades ago, the probability to meet a woman colleague working on the same area of expertise was slimmer than now.
-However, the number of women in STEM is increasing, and it is not likely to be an obstacle to collaboration nowadays.
+Altogether, our findings suggest that researchers preferably collaborate with same-gendered colleagues in STEMM research.
+Despites using a more stringent analysis methods (e.g.
+Wahlund effects) and the largest dataset to date, our results were in accord with longstanding and recent research.
+Gender homophily in science not only reinforces the less favorable status of women in science, but also creates long term problems (see above).
+It is important to explore the causes and implement measures to redress the gender disparity.
+We advocate for the introduction and implementation of gender policies in order to support opposite sex collaboration, to give men and women a chance to collaborate across gender, fields, and to promote scientific and innovative excellence.
 

### 2018-05-23 manuscript/Manuscript_Claire.Rmd
manuscript/

@@
 # Discussion
-We found evidence that researchers tend to publish more with same-gendered coauthors, despite implementing stringent controls for ???spurious???
-homophily resulting from Wahlund effects (\autoref{wahlund_plot}).
-Our data therefore reaffirm earlier conclusions [@Ferber_1980; @McDowell_1992; @Bentley_2003; @Crow_2015; @Ghiasi_2015; @Araujo_2017a; @Fahmy_2017; @Jadidi_2017; @Teele_2017; @Zettler_2017] that research collaborations occur between same-gendered colleagues substantially more often than expected if collaborators are selected without regard to their gender.
-Many journals had $\alpha'$ values over 0.1, meaning that the gender ratio of the average man's coauthors is skewed towards men by >10%, relative to the gender ratio of a woman's coauthors.
-Few journals had $\alpha'$ values below zero, and fewer still had significantly negative $\alpha'$ values, suggesting that it is comparatively rare for researchers to preferentially select opposite-gendered collaborators.
+We found evidence for researchers tending to publish more constantly with same-gendered coauthors, despite implementing stringent controls for ???spurious???
+homophily, resulting from Wahlund effects (\autoref{wahlund_plot}).
+Therefore, our data reaffirm earlier conclusions [@Ferber_1980; @McDowell_1992; @Bentley_2003; @Crow_2015; @Ghiasi_2015; @Araujo_2017a; @Fahmy_2017; @Jadidi_2017; @Teele_2017; @Zettler_2017] that research collaborations between same-gendered colleagues occur substantially more frequently than expected by random, not considering their gender.
+Many journals had $\alpha'$ values over 0.1, implying that the gender ratio of the average man's coauthors is skewed towards men by >10%, relative to the gender ratio of a woman's coauthors.
+Only few journals had $\alpha'$ values below zero, and even fewer had significantly negative $\alpha'$ values, suggesting that it is comparatively rare for researchers to preferentially select opposite-gendered co-authors.
 
 As some disciplines exhibit strong gender-biases (e.g.
-physics and computer science with predominantly men co-authors, and nursing, midwifery for women [@Holman_2018]), we hypothesized that these disciplines would also display a higher level of gender-homophily due to the increased difficulty to find an opposite-gender co-author.
+physics and computer science with predominantly male co-authors, and nursing and midwifery female [@Holman_2018]), we hypothesized that these disciplines would also display a higher level of gender-homophily due to the increased difficulty to find an opposite-gender co-author.
 However, we found no evidence that gender homophily is restricted to particular disciplines: instead, there was evidence of homophily across many diverse research disciplines (Figure 2).
 This finding may be indicative that gender homophily in science is a comprehensive concern, rather than an artifact of the existing gender gap, and not limited to stereotypical disciplines.
-Furthermore, we found that gender homophily was slightly stronger in research disciplines with a female-biased gender ratio, as compared to male-biased or gender-balanced fields (whether it is in the first or last author position).
-One interpretation of this result is that men are more likely to seek out other men in fields in which men are in the minority, relative to the homophily shown by women in fields in which women are in the minority.
+Furthermore, we found that gender homophily was slightly stronger in research disciplines dominated by a female-biased gender ratio, compared to male-biased or gender-balanced fields (regardless of their position).
+One interpretation of this result is that men are more likely to seek out to publish with other men in fields where men are in minority, compared to the homophily displayed by women in fields where females are in in minority.
 
 We also found that gender homophily was marginally stronger in 2015-2016, relative to 2005-2006.
-This result suggests that recent increases in the number of women working in STEMM, as well as recent efforts to recruit, retain, and celebrate women in STEMM, have co-occurred with rise in the frequency of same-gender collaborations, not a fall.
-We speculate that this trend might result, at least in part, from the increasing number of women working in senior positions in STEMM over the last 10 years [@Bendels_2018; @Long_2015; @McKenzie_2017].
+This suggests that recent increases in the number of women working in STEMM, as well as recent efforts to recruit, retain, and celebrate women in STEMM, have co-occurred with a rising frequency of same-gender collaborations.
+We speculate that this trend might result, at least in part, from the increasing number of women working in senior positions in STEMM over the past decade [@Bendels_2018; @Long_2015; @McKenzie_2017].
 Because many co-authorships involve a junior and a senior researcher (e.g.
 a student and professor), and the gender ratio is typically more skewed among senior staff [@Shaw_2012, @West_2013], there will be more opposite-gender collaborations than expected if the gender ratio were the same at both career stages.
-Thus, the observed increase in homophily might partially be due to the greater numbers of women working in leading roles.
+Thus, the observed increase in homophily might partially be due to the greater number of women working in leading roles.
 
@@
 Why does gender homophily exist in science?
-Collaboration often leads to co-authorship publications, citations and additional professional recognitions, therefore first author should always be considering collaborators that are most likely to enhance their productivity and in the long term also their academic success, rather than pick co-authors according to their gender.
-First of all, it is likely that decades ago, the probability to meet a woman colleague working on the same area of expertise was slimmer than now.
-Even though the number of women in STEM is increasing, and it is not likely to be an obstacle to collaboration nowadays; we cannot forgo that female representation in science is still not equal to that of males.
+Collaboration often leads to co-authorship publications, citations and additional professional recognition, therefore first authors should always be considering collaborators that are most likely to enhance their productivity and in the long term also their academic success, rather than pick co-authors according to their gender.
+First of all, it is likely that decades ago, the probability to meet a female colleague working on the same area of expertise was slimmer than now.
+Even though the number of women in STEMM is increasing, i.e.
+it is not likely to be an obstacle to collaboration anymore; we cannot forgo that female representation in science is still not equal to that of males.
 Second, the most basic source of homophily is space, scientists are more likely to collaborate with colleagues closer to them in geographic location (@McPherson_2011).
@@
 family commitments) represent the principal restrictions on their international mobility.
-As a results, women are less likely to originate collaborations with foreign colleagues [@Lemoine_1992; @Lewison_2001; @ Webster_2001; @ Lariviere_2011].
-Lastly, men and women think differently on average (REF), and are more likely to initiate collaborations based on similarity of thoughts to progress their scientific output (REF).
+As a result women are less likely to originate collaborations with foreign colleagues [@Lemoine_1992; @Lewison_2001; @ Webster_2001; @ Lariviere_2011].
+Lastly, men and women think differently on average (REF) and are more likely to initiate collaborations based on similarity of thoughts to progress their scientific output (REF).
 
-Academic publishing is one of the most (if not the most) important tool for a researcher to display their work, and is frequently used as a success indicator for being promoted and hired as a faculty member.
-Our finding that researchers are more likely to publish with same-gendered coauthors implies that co-authorship listing is not a random process, but instead creates additional gender imbalance.
+Academic publishing is one of the most (if not the most) important tool for a researcher to display their work and is frequently used as an indicator of success for being promoted and hired as a faculty member.
+Our finding that researchers are more likely to publish with same-gendered co-authors implies that co-authorship listing is not a random process, but instead creates additional gender imbalance.
 Homophily could make it harder for women to find collaborators in disciplines where women are in the minority [i.e.
 almost all STEMM fields; @Holman_2018].
-Because women make up for as little as 30% of the STEMM scientist community [@Holman_2018], they are therefore less likely to be picked as coauthors, and, as a result, gender homophily creates further gender imbalance and disadvantages for women.
+Because women make up for as little as 30% of the STEMM scientist community [@Holman_2018], they are less likely to be picked as coauthors, and, as a result, gender homophily creates further gender imbalance and disadvantages for women.
 Men tend to publish more in high impact journals as first author [@West_2013], and generate more citations than papers lead by first author women [@Long_1992; @Lariviere_2013].
-And as a consequence, gender homophily is also likely to contribute to lower publication rates for women [@McDowell_1992].
+As a consequence, gender homophily is also likely to contribute to lower publication rates for women [@McDowell_1992].
 
-Although the proportion of women in science has significantly increased over the last decades, women still tend to leave their academic careers behind considerably more than their male counterpart [@Debarre_2018] .
-Lack of scientific recognition [@Lincoln_2012], lack of research funding [@Bozeman_2004], lack of invitation to speak at scientific events[@Isbell_2012; @Schroeder_2013; @Klein_2017], discrimination during hiring [@MossRacusin_2017; @Reuben_2014], discrimination when writing a recommendation letter [@Trix_2003; @Schmader_2007], all in all this concentration of small disadvantages drives women out of science.
+Although the proportion of women in science has significantly increased over the last decades, women still tend to leave their academic careers behind considerably more often than their male counterparts [@Debarre_2018].
+Lack of scientific recognition [@Lincoln_2012], lack of research funding [@Bozeman_2004], lack of invitation to speak at scientific events [@Isbell_2012; @Schroeder_2013; @Klein_2017], discrimination during hiring [@MossRacusin_2017; @Reuben_2014] and discrimination when writing a recommendation letter [@Trix_2003; @Schmader_2007], are only some of the small disadvantages adding up to drive women out of science.
 Whether their decision to leave academia is deliberate or not, it will negatively impact science overall, as mentoring, networking, and supporting younger women scientists are especially important for women in academia.
@@
 Altogether, our findings suggest that researchers preferably collaborate with same-gendered colleagues in STEMM research.
-Despites using a more stringent analysis methods (e.g.
-Wahlund effects) and the largest dataset to date, our results were in accord with longstanding and recent research.
-Gender homophily in science not only reinforces the less favorable status of women in science, but also creates long term problems (see above).
-It is important to explore the causes and implement measures to redress the gender disparity.
+Despites using a more stringent analysis method (e.g.
+Wahlund effects) as well as the largest dataset to date, our results were in accord with longstanding and recent research, while adding new details, reasoning and insight for the current situation.
+Gender homophily in science not only reinforces the less favorable status of women in science, but also creates long term problems as described above.
+It is paramount to explore the causes and implement measures to readdress the gender disparity.
 We advocate for the introduction and implementation of gender policies in order to support opposite sex collaboration, to give men and women a chance to collaborate across gender, fields, and to promote scientific and innovative excellence.

### 2018-05-25 Gender homophilly project.Rmd
Added simulation

@@
 ```
+
+## Theoretical expectations for $\alpha$ when there is a gender gap between career stages
+
+This section makes Figure 6, by deriving the expected $\alpha$ under various assumptions, assuming that there is (potentially) a gender gap between career stages but no direct choice of collaborators by gender.
+
+```{r}
+simulate <- function(params){
+
+  # Attach the vectors of paramters from the dataframe input
+  GR.stage1 <- params$GR.stage1
+  GR.stage2 <- params$GR.stage2
+  prop.student.prof <- params$prop.student.prof
+  prop.student.student <- params$prop.student.student
+
+  # This is the proportion of each type of collaboration:
+  # early-early, early-late, late-late respectively
+  p1.1 <- (1 - prop.student.prof) * prop.student.student
+  p1.2 <- prop.student.prof
+  p2.2 <- (1 - prop.student.prof) * (1-prop.student.student)
+
+  # given the gender ratio at career stages a and b, find the proportion of collaborations that are man-man, woman-woman, or mixed gender
+  man.man <- function(GR_a, GR_b) (1-GR_a) * (1-GR_b)
+  woman.woman <- function(GR_a, GR_b) GR_a * GR_b
+  mixed <- function(GR_a, GR_b) (1-GR_a) * GR_b + (1-GR_b) * GR_a
+
+  # Find the frequencies of the 3 kinds of collaboration
+  man.man <- p1.1 * man.man(GR.stage1, GR.stage1) +
+    p1.2 * man.man(GR.stage1, GR.stage2) +
+    p2.2 * man.man(GR.stage2, GR.stage2)
+
+  woman.woman <- p1.1 * woman.woman(GR.stage1, GR.stage1) +
+    p1.2 * woman.woman(GR.stage1, GR.stage2) +
+    p2.2 * woman.woman(GR.stage2, GR.stage2)
+
+  mixed <- p1.1 * mixed(GR.stage1, GR.stage1) +
+    p1.2 * mixed(GR.stage1, GR.stage2) +
+    p2.2 * mixed(GR.stage2, GR.stage2)
+
+  # Find p and q, which are used to calcualte alpha
+
+  # p is proportion of men coauthoring with men
+  p <- 1 - mixed / (2 * man.man + mixed)
+  # q is proportion of women coauthoring with men
+  q <- mixed / (2 * woman.woman + mixed)
+
+  data.frame(alpha = p - q, 
+             GR.stage1,
+             GR.stage2,
+             prop.student.prof,
+             prop.student.student)
+}
+
+results <- expand.grid(
+  GR.stage1 = seq(0.01, 0.7, length = 201),
+  GR.stage2 = seq(0.01, 0.7, length = 201),
+  prop.student.prof = c(0.01, 0.25, 0.5, 0.75, .99), 
+  prop.student.student = c(0.01, 0.25, 0.5, 0.75, .99)) %>% 
+  filter(GR.stage1 >= GR.stage2) %>%
+  simulate() %>% 
+  mutate(GR.stage1 = round(GR.stage1 * 100),
+         GR.stage2 = round(GR.stage2 * 100),
+         prop.student.prof = prop.student.prof * 100,
+         prop.student.student = prop.student.student * 100) %>%
+  filter(!duplicated(paste(GR.stage1,GR.stage2,prop.student.prof,prop.student.student)))
+
+simulation.figure <- results %>%
+  ggplot(aes(GR.stage1, GR.stage2, z = alpha, fill = alpha)) + 
+  facet_grid(prop.student.prof ~ prop.student.student) + 
+  geom_tile() + 
+  #stat_contour(binwidth = 0.1, colour = "grey20", size = 0.3) + 
+  scale_fill_distiller(name = "\u03B1",
+                       palette = "RdYlBu",
+                       limits = c(-1,1) * max(abs(results$alpha))) +
+  scale_x_continuous(expand = c(0,0), limits = c(0,70))+
+  scale_y_continuous(expand = c(0,0), limits = c(0,70))+
+  xlab("% women among early-career authors") + 
+  ylab("% women among established authors") +
+  theme_bw() + 
+  theme(panel.grid.minor = element_blank(),
+        strip.background = element_blank(),
+        legend.position = "bottom") + 
+  ggtitle("Rows show % between-stage collaborations ", 
+          subtitle = "Columns show % within-stage collaborations that involve two early-career researchers")
+
+simulation.figure
+ggsave(simulation.figure, file = "figures/figure5.pdf", 
+       height = 5.6, width = 5)#, device = cairo_pdf)
+```
+
+### R session info
+```{r}
+sessionInfo() %>% pander()
+```

### 2018-05-25 manuscript/Manuscript_Claire.Rmd
Added simulation

@@
 author:
-- Luke Holman* and Claire Morandin+
+- Luke Holman* and Claire Morandin&sect;
 - '*\textit{luke.holman@unimelb.edu.au} \vspace{5mm}'
@@
 header-includes: 
+  \usepackage{microtype} 
   \usepackage[all]{hypcap} 
@@
 abstract: Evidence suggests that women in academia are hindered by conscious and unconscious biases, and that many female researchers feel excluded from formal and informal opportunities for research collaboration.
-In addition to ensuring fairness and helping redress gender imbalance in the academic workforce, increasing women???s access to collaboration could help scientific progress by drawing on more of the available human capital.
+In addition to ensuring fairness and helping redress gender imbalance in the academic workforce, increasing women's access to collaboration could help scientific progress by drawing on more of the available human capital.
 Here, we test whether researchers preferentially collaborate with same-gendered colleagues, using more stringent methods and a larger dataset than in past work.
-Our results reaffirm that researchers co-publish with colleagues of the same gender, and show that this ???gender homophily???
-is slightly stronger today than it was 10 years ago.
+Our results reaffirm that researchers co-publish with colleagues of the same gender, and show that this 'gender homophily' is slightly stronger today than it was 10 years ago.
 Contrary to our expectations, we found no evidence that homophily is driven mostly by senior academics, and no evidence that homophily is strongest in fields where women are in the minority.
@@
 \maketitle\unmarkedfntext{*School of BioSciences, The University of Melbourne, Victoria, Australia}
-\maketitle\unmarkedfntext{+Organismal and Evolutionary Biology Research Programme,
+\maketitle\unmarkedfntext{§Organismal and Evolutionary Biology Research Programme,
 Faculty of Biological and Environmental Sciences, University of Helsinki, Finland}
@@
 # Introduction
-Women are substantially underrepresented in many branches of the workforce in science, technology, engineering, mathematics, and medicine (STEMM), and they face additional challenges and inequities relative to men [e.g.
+Women are substantially underrepresented in many branches of the workforce in science, technology, engineering, mathematics, and medicine (STEMM), and face additional challenges and inequities relative to men [e.g.
 @Lariviere_2013; @West_2013; @Elsevier_report; @Holman_2018].
 On average, women occupy more junior positions [@Wutte_2007; @Reuben_2014] with lower salaries [@Trower_2002; @Umbach_2007], receive less grant money [@Hosek_2005; @OER_2008], are promoted more slowly [@Zuckerman_1987; @Rosenfeld_1991; @Long_1993; @Hopkins_2013], and are allocated fewer resources [@ODorchai_2009] and less research funding [@Feldt_1986, @Stack_2004, @Lariviere_2011] than men.
-Experimental studies have demonstrated that researchers regard women's achievements less favourably than identical achievements by men [@Moss_2012; @Knobloch_2013].
+Experimental studies have shown that people regard women's research achievements less favourably than identical achievements by men [@Moss_2012; @Knobloch_2013].
 
-Because publishing, networking and collaboration are all instrumental to scientific productivity and thus, academic career advancement [@Lee_2005; @Wuchty_2007;@Abramo_2009;@Lariviere_2015], numerous studies have tested for gender differences in these areas [reviewed in Table S1 in @Holman_2018; see also @Bozeman_2011; @Abramo_2013; @Badar_2013; @Long_1992].
+Because publishing, networking and collaboration are all instrumental to scientific productivity and thus, academic career advancement [@Lee_2005; @Wuchty_2007; @Abramo_2009; @Lariviere_2015], dozens of studies have tested for gender differences in these areas [reviewed in Table S1 of @Holman_2018; see also @Bozeman_2011; @Abramo_2013; @Badar_2013; @Long_1992].
 For example, studies have shown that relative to men, women tend to be less involved in international collaboration [@Lewison_2001; @Webster_2001; @Bozeman_2004; @Lariviere_2011; @Abramo_2013], collaborate less with researchers from their own university department [@Webster_2001], have less prestigious collaborations [@Long_1990], and fewer collaborations in general [@Fuchs_2001].
-Possible explanations for the smaller average professional networks of women include a greater average amount of family obligations [@Reskin_1978; @Long_1990; @Wright_2003], lower participation of women in international conferences [@Lewison_2001; Eisen 2014; @Martin_2014], and reduced receipt of research travel funds by women [@Bozeman_2004].
+These gender differences in collaboration practice presumably have multiple concurrent causes, which might include implicit and explicit gender bias [REFS], differential family obligations [@Reskin_1978; @Long_1990; @Wright_2003], concerns relating to sexual harassment [REFS], lower participation of women in international conferences [@Lewison_2001; Eisen 2014; @Martin_2014], and reduced receipt of research travel funds by women [@Bozeman_2004].
 
-A high, steadily increasing proportion of research papers is written by more than one author [@West_2013], making collaboration a key predictor of publication output, and in turn, a key predictor of career prospects [@Tower_2007; @Jordan_2008; @Cohn_2014].
-Additionally, empirical studies suggest that mixed-gender [@Bear_2011; @Campbell_2013] or otherwise diverse [@Hong_2004] teams can produce better results on collaborative tasks.
+A high, steadily increasing proportion of research papers is written by more than one author [@West_2013], making collaboration a key predictor of publication output, and thus of career prospects [@Tower_2007; @Jordan_2008; @Cohn_2014].
+Additionally, empirical studies suggest that mixed-gender [@Bear_2011; @Campbell_2013] or otherwise diverse [@Hong_2004] teams can produce better results on collaborative tasks than less diverse teams.
 For reasons such as these, several studies have tested for gender differences in collaboration frequency or pattern by examining the author lists of published research.
-To our knowledge, every study of this question has concluded that men publish with other men, and women with women, more often than expected if collaborators assort randomly with regards to gender [@Ferber_1980; @McDowell_1992; @Ghiasi_2015; @Crow_2015;@Fahmy_2017; @Zettler_2017;@Jadidi_2017; @Teele_2017; @Araujo_2017a; @Araujo_2017b].
-This preference for same-gendered colleagues is often termed ???gender homophily???.
+To our knowledge, every published study of this question has concluded that men co-publish with other men, and women with women, more often than expected if collaborators assort randomly with respect to gender [@Ferber_1980; @McDowell_1992; @Ghiasi_2015; @Crow_2015;@Fahmy_2017; @Zettler_2017; @Jadidi_2017; @Teele_2017; @Araujo_2017a; @Araujo_2017b].
+This pattern of assortative publishing has been termed 'gender homophily'.
 
-However, we hypothesise that these studies of gender homophily were hindered by a sampling issue that we term the Wahlund effect (\autoref{wahlund_plot}), making the true extent of homophily difficult to ascertain from present data.
-Essentially, whenever samples are taken from two or more discrete sets of literature, which vary in the author gender ratio and which are largely not connected by collaboration, the number of same-gendered coauthors will be inflated.
-This can make it appear as though authors preferentially publish with same-gendered colleagues if no gender preferences exist, or even if researchers preferentially select opposite-gendered colleagues.
-For example, a sample of bioinformatics and cell biology papers might contain an excess of mostly-male and mostly-female author lists, simply because researchers preferentially work with colleagues from the same discipline, and because more women work in cell biology than in bioinformatics [@Holman_2018].
-We name this issue after the Wahlund effect in population genetics [@Wahlund_1928], whereby spatial differences in allele frequencies create an excess of homozygotes across the whole population, even if alleles assort randomly within each subpopulation.
+However, we believe that these prior studies of gender homophily were hindered by a largely unacknowledged statistical issue that we term the Wahlund effect (\autoref{wahlund_plot}), meaning that the true extent of homophily is difficult to ascertain from current data.
+Essentially, whenever coauthorship data are sampled from two or more discrete sets of literature, which vary in the author gender ratio and which are largely not connected by collaboration, the number of same-gendered coauthors will be inflated.
+This can give the impression that authors preferentially publish with same-gendered colleagues if no gender preferences exist, or even if researchers actually prefer opposite-gendered colleagues.
+For example, a sample of literature containing bioinformatics and cell biology papers will probably contain an excess of mostly-male and mostly-female author lists, simply because researchers preferentially work with colleagues from the same discipline, and because the author gender ratio is more male-biased in bioinformatics than in cell biology [@Holman_2018].
+We chose the name by analogy with the Wahlund effect in population genetics [@Wahlund_1928], whereby spatial differences in allele frequencies create an excess of homozygotes across the whole population, even if alleles are randomly assorted within each subpopulation (original idea in BERGSTROM REF).
 
@@
 Each journal was assigned to one of 107 scientific disciplines (e.g.
-???Nephrology???
-or ???Cell Biology???), using PubMed???s journal categorisations where available.
+_Nephrology_ or _Cell Biology_), using PubMed's journal categorisations where available.
 Because the present study focuses on co-authorship, all single-author papers were removed from the dataset.
 To simplify the statistical analysis, we also discarded all papers for which we could not determine the gender of every author with at least 95% certainty.
-To mitigate Wahlund effects caused by variation in the gender ratio of researchers over time (see below), we also discarded all papers except those that were published 0-1 or 10-11 years before the PubMed data were collected (i.e.
-20 August 2016).
+To mitigate Wahlund effects caused by variation in the gender ratio of researchers over time (see below), we also discarded all papers except those that were published either 0-1 or 10-11 years before the PubMed data were collected (i.e.
+$20^{th}$ August 2016).
 Lastly, we excluded journals with fewer than 50 suitable papers.
@@
 (http://www.eigenfactor.org/gender/assortativity/measuring_homophily.pdf), we defined the coefficient of homophily as $\alpha = p - q$, where $p$ is the probability that a randomly-chosen co-author of a _male_ author is a man and $q$ is the probability that a randomly-chosen co-author of a _female_ author is a man.
-Therefore, $\alpha$ > 0 suggests that same-gender authors publish together more often than expected if authors assort randomly with respect to gender (homophily), $\alpha$ < 0 suggests that opposite-gender authors publish together more often than expected (heterophily), and $\alpha$ = 0 suggests random assortment with respect to gender.
+Therefore, $\alpha > 0$ suggests that same-gender authors publish together more often than expected if authors assort randomly with respect to gender (homophily), $\alpha < 0$ suggests that opposite-gender authors publish together more often than expected (heterophily), and $\alpha = 0$ suggests random assortment with respect to gender.
 
-To estimate $\alpha$ for a particular subset of the scientific literature, we estimated $p$ as the average proportion of men???s co-authors who are men (averaged across all papers with at least one man author), and $q$ as the average proportion of women???s co-authors who are men (averaged across all papers with at least one woman author).
+To estimate $\alpha$ for a particular subset of the scientific literature, we estimated $p$ as the average proportion of men's co-authors who are men (averaged across all papers with at least one man author), and $q$ as the average proportion of women's co-authors who are men (averaged across all papers with at least one woman author).
 To estimate the 95% confidence intervals on $\alpha$ for a given set of $n$ papers, we sampled $n$ papers with replacement 1000 times, estimated $\alpha$ on each sample, and recorded the 95% quantiles of the 1000 resulting estimates.
 
-As well as calculating $\alpha$ for all authors, we calculated $\alpha$ for first and last authors only.
-$\alpha$ was again defined as $p - q$, but this time $p$ was estimated as the average proportion of male co-authors on papers with a male first/last author, and $q$ was estimated as the average proportion of male co-authors on papers with female first/last authors.
+As well as calculating $\alpha$ for all authors, we calculated $\alpha$ for first or last authors only.
+$\alpha$ was again defined as $p - q$, but this time $p$ was estimated as the average proportion of male co-authors on papers with a male first (or last) author, and $q$ was estimated as the average proportion of male co-authors on papers with female first (or last) authors.
 We did not calculate $\alpha$ for other authorship positions (e.g.
@@
 To borrow Prof.
-Bergstrom???s example, consider a small research specialty comprising just two men and two women researchers, who have together produced six two-author papers: one paper in each of the six possible two-author combinations.
-For these six papers, $\alpha$ = $-\frac{1}{3}$, even though same- and opposite-gendered coauthors were selected in equal proportion to their frequency in the pool of possible collaborators.
+Bergstrom's example, consider a small research specialty comprising just two men and two women researchers, who have together produced six two-author papers: one paper in each of the six possible two-author combinations.
+For these six papers, $\alpha = -\frac{1}{3}$, even though same- and opposite-gendered coauthors were selected in equal proportion to their frequency in the pool of possible collaborators.
 To control for the fact that the null expectation for $\alpha$ is not necessarily zero (particularly for very small datasets), we devised an adjusted version of the coefficient of homophily, which we term $\alpha'$.
 Every time we calculated $\alpha$ for a set of papers, we also determined the expected value of $\alpha$ under the null hypothesis that authors assort randomly with respect to gender.
-This was accomplished by randomly shuffling authors across papers 1000 times, recalculating $\alpha$, and taking the median.
+This was accomplished by randomly permuting authors across papers 1000 times, recalculating $\alpha$, and taking the median.
 We then calculated $\alpha'$ by subtracting the null expectation for $\alpha$ from the observed value.
 We also used the null-simulated $\alpha$ values to calculate a two-tailed p-value for the observed value of $\alpha$; the p-value was defined as the proportion of null simulations for which $|\alpha_{null}| > |\alpha_{obs}|$.
-We applied false discovery rate correction to each set of p-values to account for multiple testing [@Benjamini_1995].
+We applied false discovery rate correction to each set of p-values to account for multiple testing following Benjamini and Hochberg [-@Benjamini_1995].
 
 As expected, $\alpha'$ was usually almost identical to $\alpha$ (Figure S1), but $\alpha$ was downwardly biased relative to $\alpha'$ for small datasets (Figure S2).
-Additionally, the correlation between $\alpha'$ and sample size was negligible ($R^2$ < 0.01), suggesting that $\alpha'$ effectively removed the dependence of $\alpha$ on sample size.
-We therefore used the adjusted statistic $\alpha'$ in all our analyses.
+Additionally, the correlation between $\alpha'$ and sample size was negligible ($R^2 < 0.01$), suggesting that our calculation og $\alpha'$ effectively removed the dependence of $\alpha$ on sample size.
+We therefore use the adjusted statistic $\alpha'$ in all analyses.
 
 ## Minimising the Wahlund effect: discipline and time period
-As discussed in the Introduction, the Wahlund effect can give the appearance of gender homophily (reflected in our study as $\alpha'$ > 0) whenever one lumps together different ???subpopulations???
-of academic papers that differ in their author gender ratios (Figure 1).
+As discussed in the Introduction and shown in \autoref{wahlund_plot}, the Wahlund effect can give the appearance of gender homophily (reflected in our study as $\alpha' > 0$) whenever one lumps together different 'subpopulations' of academic papers that differ in their author gender ratios.
 
@@
 For example, Holman et al.
-[-@Holman_2018] showed that authors based in Serbia are more than twice as likely to be women as are authors based in Japan.
-Therefore, a dataset containing a mix of papers from teams of authors based in these two countries would probably contain an excess of same-sex coauthorships, even if collaboration were random with respect to gender within each country.
+[-@Holman_2018] showed that PubMed-indexed authors based in Serbia are more than twice as likely to be women as are authors based in Japan.
+Therefore, a dataset containing a mix of papers from teams of authors based in these two countries would contain an excess of same-sex coauthorships, even if collaboration were random with respect to gender within each country.
 
-To address this, we also analysed every combination of journal and author country of affiliation for which we had enough data (i.e.
+To address this issue, we also analysed every combination of journal and author country of affiliation for which we had enough data (i.e.
 50 or more papers published in 2015-16).
@@
 In the majority of disciplines, authorship conventions mean that the first-listed author is often an early-career researcher, while the author listed last is more likely to be a senior researcher leading a research team [@Wren_2007].
-Assuming that senior researchers are the main drivers of homophily and that there are enough papers with three or more authors, we predict that the last author???s gender will be the strongest predictor of the remaining authors???
-genders (i.e.
+Assuming that senior researchers are the main drivers of homophily and that there are enough papers with three or more authors, we predict that the last author's gender will be the strongest predictor of the remaining authors' genders (i.e.
 the gender of the last author will be more salient than that of the first author, or any other authorship position).
-This is because the first author???s gender would simply be an imperfect correlate of the true causal effect, while the last author???s gender would be the causal effect itself.
+This is because the first author's gender would simply be an imperfect correlate of the true causal effect, while the last author's gender would be the causal effect itself.
 
@@
 The response variable was $\alpha'$, and we weighted each observation by the inverse of the standard error from our estimate of $\alpha'$, meaning that more accurate measurements of $\alpha'$ had more influence on the results.
-We used a similar model to test for a difference in $\alpha'$ between the 2005-6 and the 2015-16 datasets, with two differences: instead of authorship position, we fit year range as a two-level fixed factor, and used $\alpha'$ estimated for all authors (not first/last authors) as the response variable.
+We used a similar model to test for a difference in $\alpha'$ between the 2005-6 and the 2015-16 datasets, with two differences: we fit year range as a two-level fixed factor (instead of authorship position), and we used $\alpha'$ estimated for all authors (not first/last authors) as the response variable.
 
-The relationship between the gender ratio of authors publishing in a journal and its $\alpha'$ value appeared nonlinear (Figure 4).
+The relationship between the gender ratio of authors publishing in a journal and its $\alpha'$ value appeared nonlinear (see Results).
 We therefore fit a generalised additive model with thin plate regression spline smoothing, implemented using the ``mgcv`` package for R.
+
+## Theoretical expectations for $\alpha$ when there is a gender gap between career stages
+In many STEMM subjects, the gender ratio is more skewed among established researchers relative to early-career researchers [@Shaw_2012; @Holman_2018].
+We hypothesised that this skew could potentially create Wahlund effects and 'reverse' Wahlund effects.
+For example, imagine that the majority of collaborations are between students and professors, and that the gender ratio is skewed in opposite directions at each career stage: we will then see an excess of mixed-gender coauthorships (heterophily, $\alpha < 0$), even if gender has no causal effect on the selection of coauthors.
+Similarly, a hypothetical field in which students work only with students, and professors with professors, will have apparent gender homophily ($\alpha > 0$).
+
+Because we can think of no tractable way to control for this issue using our dataset (which contains no information on career stage), we instead decided to derive the theoretical expectations for $\alpha$ when there is a difference in gender ratio across career stages.
+For simplicity, we assume there are only two career stages, though the results are likely to be general.
+Under the null model that gender has no causal effect on collaboration, we calculated $\alpha$ for various combinations of the four free parameters: the gender ratios for early- and late-career researchers, and the relative frequency of collaborations between early-early, early-late, and late-late collaborations.
+We then used the theoretical expectations for $\alpha$ to properly qualify our main conclusions (see Results).
+The R code used to derive the theoretical expectations can be found in the Online Supplementary Material.
+
+## Data availability and reproducibility
+The Online Supplementary Material is an HTML document prepared with R Markdown, which contains R scripts used to produce all data, figures and tables.
+The input data from Holman et al.
+[-@Holman_2018] is archived at https://osf.io/bt9ya/.
 
@@
 \autoref{alpha_histograms} shows the distribution of $\alpha'$ estimates in 2015-2016 across all journals for which we recovered sufficient data, when $\alpha'$ was calculated for all authors, first authors only, or last authors only.
-The great majority of journals had $\alpha'$ > 0, and for many of these the corrected two-tailed p-values suggested that $\alpha'$ was significantly greater than zero (1469/2077 journals were significant in 2015-16, and 404/1192 in 2005-6).
+The great majority of journals had $\alpha' > 0$, and for many of these the corrected two-tailed p-values suggested that $\alpha'$ was significantly greater than zero (1469/2077 journals were significant in 2015-16, and 404/1192 in 2005-6).
 Only 2/2077 journals had statistically significantly heterophily (i.e.
-$\alpha'$ < 0) in 2015-16, and 1/1192 in 2005-6.
+$\alpha' < 0$) in 2015-16, and 1/1192 in 2005-6.
 The remaining 606 or 787 journals (in 2015 and 2005 respectively) had a value of $\alpha'$ not significantly different from zero, such that we could not reject the null hypothesis of random assortment with respect to gender.
 
-$\alpha'$ was significantly higher in the sample from 2015-16 relative to 2005-6, but the difference in means was small (Figure S3; Effect of the fixed factor ???time period???
-in a linear mixed model of the data for all author positions: Cohen???s $d$ = 0.09??0.04, $t_{948}$ = 2.51, p = 0.012).
+$\alpha'$ was significantly higher in the sample from 2015-16 relative to 2005-6, but the difference in means was small (Figure S3; Effect of the fixed factor 'Time period' in a linear mixed model of the data for all author positions: Cohen's $d$ = $0.09{\pm}0.04$, $t_{948}$ = 2.51, p = 0.012).
 
-When comparing pairs of $\alpha'$ values estimated for the first and last authors for the same journals, we found that $\alpha'$ tended to be higher for first authors than for last authors (Figure S4; Effect of the fixed factor ???Authorship position???
-in a linear mixed model: Cohen???s $d$ = 0.07??0.02, $t_{1988}$ = 4.48, p < 0.0001).
-This suggests that the gender of the first author was a slightly stronger predictor of the remaining authors???
-genders than the gender of the last author, i.e.
+When comparing pairs of $\alpha'$ values estimated for the first and last authors for the same journals, we found that $\alpha'$ tended to be higher for first authors than for last authors (Figure S4; Effect of the fixed factor 'Authorship position' in a linear mixed model: Cohen's $d$ = $0.07{\pm}0.02$, $t_{1988}$ = 4.48, p < 0.0001).
+This suggests that the gender of the first author was a slightly stronger predictor of the remaining authors' genders than the gender of the last author, i.e.
 the opposite of our prediction based on the hypothesis that senior scientists are causally responsible for homophily.
 
-## Possible differences in the strength of homophily between disciplines
+## Variance in homophily between disciplines
 \autoref{alpha_histograms} illustrates the variance in journal homophily values ($\alpha'$) across scientific disciplines.
 All disciplines had a positive average $\alpha'$, although homophily appeared somewhat stronger in some disciplines than others (e.g.
-mean $\alpha'$ was 0.12??0.02 for Urology journals, and 0.03??0.01 for Veterinary Medicine journals; \autoref{alpha_histograms}, Table S3).
-However, there was little evidence for strong differences in $\alpha'$ between disciplines: the random factor ???discipline???
-explained at most 1% of the variance in $\alpha'$ in the two linear mixed models described in the previous section (see \autoref{alpha_histograms} and mixed models in Online Supplementary Material).
+mean $\alpha'$ was $0.12{\pm}0.02$ for Urology journals, and $0.03{\pm}0.01$ for Veterinary Medicine journals; \autoref{alpha_histograms}, Table S3).
+However, there was little evidence for strong differences in $\alpha'$ between disciplines: the random factor 'Discipline' explained at most 1% of the variance in $\alpha'$ in the two linear mixed models described in the previous section (see \autoref{alpha_histograms} and mixed models in Online Supplementary Material).
 This implies that the processes causing positive $\alpha'$ are similarly strong in all the disciplines we examined.
@@
 There was no indication that journals publishing on a wide range of topics have higher $\alpha'$ values than more specialised journals, due to the Wahlund effect.
-For example, the journal category ???Multidisciplinary???
-- which includes journals like _PLoS ONE_, _Nature_, _Science_, and _PNAS_ - did not have notably elevated $\alpha'$ (\autoref{alpha_histograms}).
+For example, the journal category 'Multidisciplinary' - which includes journals like _PLoS ONE_, _Nature_, _Science_, and _PNAS_ - did not have notably elevated $\alpha'$ (\autoref{alpha_histograms}).
 This result suggests that our estimates of homophily, and estimates from earlier studies, are not notably inflated by the presence of disparate research topics (with variable author gender ratios) being published within individual journals.
@@
 
-## Analysis correcting for differences between countries
+## Analysis accounting for differences in author gender ratio between countries
 When we restricted the analysis to only include authors with affiliations from a single country, we found statistically significant homophily for 72 of the 325 journal-country combinations tested (64 unique journals and 18 unique countries), and no significant evidence of heterophily (Figures S5-S6).
@@
 
+## Theoretical expectations for $\alpha$ when there is a gender gap between career stages
+As shown in \autoref{simulation}, we confirmed that $\alpha$ is expecte to be non-zero, even if collaborators are selected in a gender-blind fashion, if there is a gender gap between career stages.
+The deviation of $\alpha$ from zero depends on the relative frequencies of collaboration within and between career stages.
+When >50% of collaborations were between early and established researchers, we observed gender heterophily ($\alpha < 0$).
+Conversely, when >50% of collaborations occured within career stages, we observed gender homophily ($\alpha > 0$).
+In a few parameter spaces (shown in red; \autoref{simulation}), $\alpha$ was quite high, and overlapped with the values that we estimated (\autoref{alpha_histograms}).
+
+\autoref{simulation} suggests that our main conclusions are likely to be robust to this career stage issue.
+If A) the gender ratio is highly skewed across career stages (e.g.
+50% vs 5% women among early- vs late-career researchers), and B) collaborations between early and established researchers are very rare (e.g.
+1% of the total), then $\alpha$ can approach the levels shown in \autoref{alpha_histograms} even if collaborators are selected randomly with respect to gender.
+The first of these conditions can be rejected for the majority of research fields [@Shaw_2012; @Holman_2018], and the second seems improbable because it is very common for early-career researchers to co-publish with an established mentor.
+However, one can get $\alpha > 0$ for plausible combinations of parameters, e.g.
+a moderate gender gap coupled with a moderate excess of within career stage collaboration; thus, this effect may explain some of the observed homophily.
+
+Additionally, if there is a large gender gap between career stages and collaboration between early-career and established researchers are commonplace, then the baseline expectation for $\alpha$ is less than zero.
+Therefore, for disciplines where these two conditions are met, our results will under-estimate the extent to which researchers preferentially select same-gendered collaborators.
+
 # Discussion
-We found evidence for researchers tending to publish more constantly with same-gendered coauthors, despite implementing stringent controls for ???spurious???
-homophily, resulting from Wahlund effects (\autoref{wahlund_plot}).
-Therefore, our data reaffirm earlier conclusions [@Ferber_1980; @McDowell_1992; @Bentley_2003; @Crow_2015; @Ghiasi_2015; @Araujo_2017a; @Fahmy_2017; @Jadidi_2017; @Teele_2017; @Zettler_2017] that research collaborations between same-gendered colleagues occur substantially more frequently than expected by random, not considering their gender.
-Many journals had $\alpha'$ values over 0.1, implying that the gender ratio of the average man's coauthors is skewed towards men by >10%, relative to the gender ratio of a woman's coauthors.
-Only few journals had $\alpha'$ values below zero, and even fewer had significantly negative $\alpha'$ values, suggesting that it is comparatively rare for researchers to preferentially select opposite-gendered co-authors.
+We found evidence that researchers publish with same-gendered coauthors more often than expected under random assortment, even after implementing stringent controls for 'spurious' homophily resulting from Wahlund effects (\autoref{wahlund_plot}).
+Our data therfore reaffirm and refine the results of earlier studies finding evidence of gender homophily [@Ferber_1980; @McDowell_1992; @Bentley_2003; @Crow_2015; @Ghiasi_2015; @Araujo_2017a; @Fahmy_2017; @Jadidi_2017; @Teele_2017; @Zettler_2017].
+The excess of same-gender coauthorships was quite strong: many journals had $\alpha' > 0.1$, indicating that the gender ratio of the average man's coauthors is skewed towards men by >10%, relative to that of the average woman.
+Few journals had $\alpha'$ values below zero, and almost no journals showed statistically significant gender heterophily after controlling for multiple testing.
+The results therefore suggest that it is comparatively rare for researchers to preferentially select opposite-gendered co-authors.
 
-As some disciplines exhibit strong gender-biases (e.g.
-physics and computer science with predominantly male co-authors, and nursing and midwifery female [@Holman_2018]), we hypothesized that these disciplines would also display a higher level of gender-homophily due to the increased difficulty to find an opposite-gender co-author.
-However, we found no evidence that gender homophily is restricted to particular disciplines: instead, there was evidence of homophily across many diverse research disciplines (Figure 2).
-This finding may be indicative that gender homophily in science is a comprehensive concern, rather than an artifact of the existing gender gap, and not limited to stereotypical disciplines.
-Furthermore, we found that gender homophily was slightly stronger in research disciplines dominated by a female-biased gender ratio, compared to male-biased or gender-balanced fields (regardless of their position).
-One interpretation of this result is that men are more likely to seek out to publish with other men in fields where men are in minority, compared to the homophily displayed by women in fields where females are in in minority.
+An important limitation of our study (and all other similar studies) is that we cannot reliably ascertain what caused the excess of same-gender coauthorships.
+As well as conscious or unconcious choice by researchers, there could be uncontrolled-for Wahlund effects.
+We suspect the contribution of these to be minor, since we found positive $\alpha'$ after controlling for three obvious sources of Wahlund effect, but it is not possible to be sure.
+
+Some disciplines exhibit especially skewed gender ratios; e.g.
+most authors in physics and computer science are men, while most authors in nursing are women [@Holman_2018].
+We hypothesised that these gender-imbalanced disciplines might show the strongest gender homophily, e.g.
+because one needs to exercise stronger choice (with respect to gender) if one enjoys working with same-gendered collaborators, if one's field has a large gender gap.
+By contrast, researchers in fields without a gender gap will have many same-gender collaborators just by collaborating randomly with respect to gender.
+Contrary to this hypothesis, we found no evidence that gender homophily is restricted to particular disciplines: $\alpha'$ was similarly high across the board \autoref{alpha_histograms}.
+Moreover, $\alpha'$ was _lower_ for journals with a male-biased author gender ratio relative to those with an even gener ratio, suggesting that women are _less_ likely to preferentially seek out women colleagues in male-biased research disciplines.
+Interestingly though, $\alpha'$ was highest for journals with a female-biased author gender ratio.
+One interpretation of this result is that men are more likely to preferentially select male collaborators in fields where men are a minority, relative to the homophily displayed by women in fields where women are the minority.
+However, this latter result is only tentatively supported by our data, since there are so few journals on PubMed with mostly-female authors \autoref{alpha_gender_ratio}.
 
@@
 This suggests that recent increases in the number of women working in STEMM, as well as recent efforts to recruit, retain, and celebrate women in STEMM, have co-occurred with a rising frequency of same-gender collaborations.
-We speculate that this trend might result, at least in part, from the increasing number of women working in senior positions in STEMM over the past decade [@Bendels_2018; @Long_2015; @McKenzie_2017].
+Although this trend might reflect a change in the gender preferences of researchers seeking collaborators, there are other (probably more likely) explanations that could explain this result.
+Specifically, we speculate that this trend results from the increasing number of women working in senior positions in STEMM over the past decade [e.g.
+@Long_2015; @Bendels_2018; @McKenzie_2017].
 Because many co-authorships involve a junior and a senior researcher (e.g.
-a student and professor), and the gender ratio is typically more skewed among senior staff [@Shaw_2012, @West_2013], there will be more opposite-gender collaborations than expected if the gender ratio were the same at both career stages.
-Thus, the observed increase in homophily might partially be due to the greater number of women working in leading roles.
+a student and professor), and the gender ratio is typically more male-biased among senior staff than junior staff [@Shaw_2012; @Holman_2018], there will be an excess of opposite-gender collaborations relative to the expectation when there is no difference in gender ratio across career stages.
+As the gender ratio equalises across career stages, this effect will weaken, causing $\alpha'$ to rise.
 
-Despite the fact that journal impact factors tend to be negatively associated with the proportion of women [@Holman_2018], we found that journals with a high impact factor (relative to other journals in the same research discipline) tended to have weaker gender homophily than did low-impact journals (Figure 4).
+We found that journals with a high impact factor (relative to other journals in the same research discipline) tended to have weaker gender homophily than did low-impact journals \autoref{impact_factor}.
 If one accepts the contentious claim that papers published in high-impact journals tend to be of higher quality than those in low-impact journals [e.g.
-@Garfield_2006], then this result provides correlational support for the hypothesis that mixed-gender teams produce better research than single-gender teams [@Campbell_2013].
+@Garfield_2006], then this result provides non-experimental support for the hypothesis that mixed-gender teams produce better research than single-gender teams [@Campbell_2013].
 
 Why does gender homophily exist in science?
-Collaboration often leads to co-authorship publications, citations and additional professional recognition, therefore first authors should always be considering collaborators that are most likely to enhance their productivity and in the long term also their academic success, rather than pick co-authors according to their gender.
+Collaboration often leads to co-authorship, publications, citations and additional professional recognition, therefore first authors should always be considering collaborators that are most likely to enhance their productivity and in the long term also their academic success, rather than pick co-authors according to their gender.
 First of all, it is likely that decades ago, the probability to meet a female colleague working on the same area of expertise was slimmer than now.
@@
 # Acknowledgements
-We are grateful to XXXX.This work was supported by the Academy of Finland (284666 to the Centre of Excellence in Biological Interactions).
+We are grateful to XXXX. CM was supported by the Academy of Finland (284666 to the Centre of Excellence in Biological Interactions).
 
@@
 \label{impact_factor}](../figures/figure4.pdf)
+
+\newpage
+![When there is a difference in gender ratio between early-career and established researchers, and collaboration is non-random with respect to career stage, the null expectation for $\alpha$ deviates from zero.
+An excess of collaborations between career stages gives the appearance of gender heterophily (lower rows, blue areas), while an excess of within-career stage collaborations produced apparent gender homophily (upper rows, red areas).
+However, the conditions required for strong gender homophily are quite restrictive, making it unlikely that this issue explains all of the homophily observed in Figure 2.
+\label{simulation}](../figures/figure5.pdf)

### 2018-05-25 manuscript/Manuscript_Claire.Rmd
Figures in text

@@
 csl: the-american-naturalist.csl
-abstract: Evidence suggests that women in academia are hindered by conscious and unconscious biases, and that many female researchers feel excluded from formal and informal opportunities for research collaboration.
+abstract: Evidence suggests that women in academia are hindered by conscious and unconscious biases, and often feel excluded from formal and informal opportunities for research collaboration.
 In addition to ensuring fairness and helping redress gender imbalance in the academic workforce, increasing women's access to collaboration could help scientific progress by drawing on more of the available human capital.
@@
 Contrary to our expectations, we found no evidence that homophily is driven mostly by senior academics, and no evidence that homophily is strongest in fields where women are in the minority.
-Interestingly, homophily was negatively correlated with journal impact factor (standardised by research discipline), as predicted if mixed-gender teams produce better research.
+Interestingly, homophily was negatively correlated with journal impact factor, as predicted if mixed-gender teams produce better research.
 \vspace{5mm} \par\noindent \textbf{Keywords:} Gender bias, Homophily, Scientific collaboration, Text mining, Women in STEM.
@@
 We chose the name by analogy with the Wahlund effect in population genetics [@Wahlund_1928], whereby spatial differences in allele frequencies create an excess of homozygotes across the whole population, even if alleles are randomly assorted within each subpopulation (original idea in BERGSTROM REF).
+
+![The Wahlund effect can make it appear as if authors prefer to publish with same-gendered colleagues, even if no such preference exists.
+Here, coloured circles represent male and female authors, and coauthors are linked with lines.
+Across the whole set of ten papers, there is an apparent excess of same-gender collaborations.
+Specifically, there are six same-gender papers and only four mixed-gender papers, which is fewer than the $10\times2\times0.5\times0.5 = 5$ mixed-gender papers we would expect under the null hypothesis that authors assort randomly with respect to gender.
+However, within each subset, there is no evidence that authors prefer to publish with same-gendered individuals.
+The Wahlund effect will tend to inflate the frequency of same-sex coauthors whenever the data is composed of two or more disconnected subsets of literature with different author gender ratios; these subsets could be research disciplines, older versus newer papers, or papers from authors in different countries.
+\label{wahlund_plot}](../figures/Figure 1 - Wahlund figure.pdf)
 
@@
 As expected, $\alpha'$ was usually almost identical to $\alpha$ (Figure S1), but $\alpha$ was downwardly biased relative to $\alpha'$ for small datasets (Figure S2).
-Additionally, the correlation between $\alpha'$ and sample size was negligible ($R^2 < 0.01$), suggesting that our calculation og $\alpha'$ effectively removed the dependence of $\alpha$ on sample size.
+Additionally, the correlation between $\alpha'$ and sample size was negligible ($R^2 < 0.01$), suggesting that our calculation of $\alpha'$ effectively removed the dependence of $\alpha$ on sample size.
 We therefore use the adjusted statistic $\alpha'$ in all analyses.
@@
 
+![Of the `r n.journals` journals for which we had adequate data in 2015-2016, `r n.sig.homo` showed statistically significant evidence of homophily (denoted by $\alpha' > 0$), and `r n.sig.hetero` showed statistically significant evidence of heterophily ($\alpha' < 0$), after adjusting p-values using Benjamini-Hochberg false discovery rate correction.
+The white area shows the number of journals for which homophily was significantly stronger than expected under the null hypothesis (p < 0.05), while the blue area shows all the remainder.
+Patterns were similar whether $\alpha'$ was calculated for all authors, for first authors only, or for last authors only.
+\label{alpha_histograms}](../figures/figure2.pdf)
+
 $\alpha'$ was significantly higher in the sample from 2015-16 relative to 2005-6, but the difference in means was small (Figure S3; Effect of the fixed factor 'Time period' in a linear mixed model of the data for all author positions: Cohen's $d$ = $0.09{\pm}0.04$, $t_{948}$ = 2.51, p = 0.012).
@@
 
+![There is a weakly positive, non-linear relationship between the gender ratio of authors publishing in a journal, and the coefficient of homophily ($\alpha'$).
+Specifically, journals with 50% women authors or higher tended to have more same-sex coauthorships than did journals with predominantly men authors.
+This relationship held whether $\alpha'$ was calculated for all authors, first authors only, or last authors only.
+A negative value on the x-axis denotes an excess of men authors, a positive value denotes an excess of women authors, and zero denotes gender parity.
+The lines were fitted using generalised additive models with the smoothing parameter $k$ set to 3.
+\label{alpha_gender_ratio}](../figures/figure3.pdf)
+
 ## Relationship between journal impact factor and gender homophily
 We observed a noisy but statistically significant linear relationship between discipline-standardised journal impact factor and $\alpha'$, such that journals with a high impact factor for their discipline had weaker gender homophily than did journals with a low impact factor for their discipline (\autoref{impact_factor}; linear regression: $R^2$ = `r R2`, $t_{1415}$ = -8.0, p < 0.0001).
+
+![Journal impact factor (expressed relative to the average for the discipline) is negatively correlated with $\alpha'$.
+The relationship is noisy ($R^2$ = `r R2`), but the results suggest that journals with strong homophily tend to have lower impact factors than journals with weak homophily in the same discipline.
+\label{impact_factor}](../figures/figure4.pdf)
 
@@
 In a few parameter spaces (shown in red; \autoref{simulation}), $\alpha$ was quite high, and overlapped with the values that we estimated (\autoref{alpha_histograms}).
+
+![When there is a difference in gender ratio between early-career and established researchers, and collaboration is non-random with respect to career stage, the null expectation for $\alpha$ deviates from zero.
+An excess of collaborations between career stages gives the appearance of gender heterophily (lower rows, blue areas), while an excess of within-career stage collaborations produced apparent gender homophily (upper rows, red areas).
+However, the conditions required for strong gender homophily are quite restrictive, making it unlikely that this issue explains all of the homophily observed in Figure 2.
+\label{simulation}](../figures/figure5.pdf)
 
@@
 <div id="refs"></div>
-
-\newpage
-# Figures
-![The Wahlund effect can make it appear as if authors prefer to publish with same-gendered colleagues, even if no such preference exists.
-Here, coloured circles represent male and female authors, and coauthors are linked with lines.
-Across the whole set of ten papers, there is an apparent excess of same-gender collaborations.
-Specifically, there are six same-gender papers and only four mixed-gender papers, which is fewer than the $10\times2\times0.5\times0.5 = 5$ mixed-gender papers we would expect under the null hypothesis that authors assort randomly with respect to gender.
-However, within each subset, there is no evidence that authors prefer to publish with same-gendered individuals.
-The Wahlund effect will tend to inflate the frequency of same-sex coauthors whenever the data is composed of two or more disconnected subsets of literature with different author gender ratios; these subsets could be research disciplines, older versus newer papers, or papers from authors in different countries.
-\label{wahlund_plot}](../figures/Figure 1 - Wahlund figure.pdf)
-
-\newpage
-![Of the `r n.journals` journals for which we had adequate data in 2015-2016, `r n.sig.homo` showed statistically significant evidence of homophily (denoted by $\alpha' > 0$), and `r n.sig.hetero` showed statistically significant evidence of heterophily ($\alpha' < 0$), after adjusting p-values using Benjamini-Hochberg false discovery rate correction.
-The white area shows the number of journals for which homophily was significantly stronger than expected under the null hypothesis (p < 0.05), while the blue area shows all the remainder.
-Patterns were similar whether $\alpha'$ was calculated for all authors, for first authors only, or for last authors only.
-\label{alpha_histograms}](../figures/figure2.pdf)
-
-\newpage
-![There is a weakly positive, non-linear relationship between the gender ratio of authors publishing in a journal, and the coefficient of homophily ($\alpha'$).
-Specifically, journals with 50% women authors or higher tended to have more same-sex coauthorships than did journals with predominantly men authors.
-This relationship held whether $\alpha'$ was calculated for all authors, first authors only, or last authors only.
-A negative value on the x-axis denotes an excess of men authors, a positive value denotes an excess of women authors, and zero denotes gender parity.
-The lines were fitted using generalised additive models with the smoothing parameter $k$ set to 3.
-\label{alpha_gender_ratio}](../figures/figure3.pdf)
-
-\newpage
-![Journal impact factor (expressed relative to the average for the discipline) is negatively correlated with $\alpha'$.
-The relationship is noisy ($R^2$ = `r R2`), but the results suggest that journals with strong homophily tend to have lower impact factors than journals with weak homophily in the same discipline.
-\label{impact_factor}](../figures/figure4.pdf)
-
-\newpage
-![When there is a difference in gender ratio between early-career and established researchers, and collaboration is non-random with respect to career stage, the null expectation for $\alpha$ deviates from zero.
-An excess of collaborations between career stages gives the appearance of gender heterophily (lower rows, blue areas), while an excess of within-career stage collaborations produced apparent gender homophily (upper rows, red areas).
-However, the conditions required for strong gender homophily are quite restrictive, making it unlikely that this issue explains all of the homophily observed in Figure 2.
-\label{simulation}](../figures/figure5.pdf)

### 2018-05-27 Gender homophilly project.Rmd
Some writing

@@
     number_sections: false  # set to true if you want to number sections at each table header
-    theme: lumen # Luke's favourite fonts and colours
+    theme: lumen # nice fonts and colours
     code_folding: hide # Add some buttons to show/hide the code
@@
 ```{r warning=FALSE, message=FALSE, results='hide'}
-library(dplyr) # 'data wrangling' 
+library(dplyr) # data handling 
 library(stringr) # string manipulation
@@
 library(pander) # nice Markdown tables
-library(kableExtra) # scrollable table
+library(kableExtra) # scrollable html tables
 library(tidyr) # spread and gather functions
@@
 library(ggbeeswarm) # beeswarm plots
-library(Cairo) # for exporting the alpha characters via ggsave
+library(Cairo) # for exporting the character 'alpha' via ggsave
 
@@
 Importantly, this function discards all papers where we do not know the gender of every single author.
-This is conservative but greatly simplifies the test -- we'd have to be careful how we included papers where some author's genders were unknown.
-Our failures to identify author gender are almost certainly randomly distributed across papers with respect to gender homophily, this omission makes our tests less sensitive but should not bias their outcomes.
+This is conservative but greatly simplifies the test -- we'd have to be careful how we incorporated information from papers where some author's genders were unknown.
+Our failures to identify author gender are almost certainly randomly distributed across papers with respect to gender homophily, and so this omission makes our tests less sensitive but should not bias their outcomes.
 
@@
 convert.dates <- function(dates) {
-  dates <- as.numeric(as.Date(dates, "%d_%m_%Y")) - as.numeric(as.Date("20_08_2016","%d_%m_%Y")) # Convert the date to days before the present
+  dates <- as.numeric(as.Date(dates, "%d_%m_%Y")) - 
+    as.numeric(as.Date("20_08_2016","%d_%m_%Y")) # Convert the date to days before the present
   return(dates / 365.25)  # Convert from days to years
@@
 
-retrieve.authorship.data <- function(journals, minimum.n, start.of.time.period, end.of.time.period, file.name, over.write = FALSE){
-
-  if(over.write | !(file.name %in% list.files())){ # Only do the test if over.write = TRUE or there is no datafile in the working directory
+retrieve.authorship.data <- function(journals, 
+                                     minimum.n, 
+                                     start.of.time.period, 
+                                     end.of.time.period, 
+                                     file.name, 
+                                     over.write = FALSE){
+
+  # Only do the test if over.write = TRUE or there is no datafile in the working directory
+  if(over.write | !(file.name %in% list.files())){
 
@@
 
-    # Loop over all the journals, try to get the data, and if there is enough, store the data we need as a list entry in 'count.data'
+    # Loop over all the journals, try to get the data, and if 
+    # there is enough, store the data we need as a list entry in 'count.data'
     for(i in 1:length(journals)){
@@
       # MFU means the first author was male, second female, and third was of unknown gender. 
-      # Genders were only recorded as M or F if the person's given name is associated with one gender >95% of the time.
+      # Genders were only recorded as M or F if the given name is associated with one gender >=95% of the time.
       focal <- pubmed_sqlite %>% 
@@
       # For example for papers from the last one year in the dataset, set start.of.time.period = 1 and end.of.time.period = 0
-      # For papers from between 5 and 6 yeras ago, set start.of.time.period = 6 and end.of.time.period = 5
+      # For papers from between 5 and 6 years ago, set start.of.time.period = 6 and end.of.time.period = 5
       focal <- focal %>% 
@@
                                lastF = substr(focal, n.char, n.char) == "F")
-
-          rm(focal) # Explicitly discard 'focal', in case this helps to reduce memory footprint
+          rm(focal) # Explicitly discard 'focal', to reduce memory footprint
 
@@
 There is a good reason to test whether $\alpha$ is significantly greater (or lower) than expected under the null, rather than simply checking whether the 95% CIs on $\alpha$ overlap zero.
-The reason is that the null expectation for $\alpha$ is not centred zero for small datasets.
+The reason is that the null expectation for $\alpha$ is not centred on zero for small datasets.
 This issue was summed up clearly by Prof.
@@
 ### Add additional metadata to the $\alpha$ results
-Here, we assign extra information to each journal - research discipline (see Holman _et al._ in prep), ISI impact factor (where available), and the country that publishes the journal (where known; information taken from PubMed).
+Here, we assign extra information to each journal - research discipline (see Holman _et al._ 2018), ISI impact factor (where available), and the country that publishes the journal (where known; information taken from PubMed).
 ```{r message=FALSE, warning=FALSE}
@@
 merge.and.clean <- function(homophily.filepath, journal.data){
-  left_join(read.csv(homophily.filepath, stringsAsFactors = FALSE), journal.data, by = "journal") %>% 
+  left_join(read.csv(homophily.filepath, stringsAsFactors = FALSE), 
+            journal.data, 
+            by = "journal") %>% 
     select(discipline, 
@@
   filter(Age == "New") %>%
-  ggplot(aes(x = log10(n.useable.papers), y = alpha.adjusted - observed.alpha)) + 
+  ggplot(aes(x = log10(n.useable.papers), 
+             y = alpha.adjusted - observed.alpha)) + 
   geom_point(alpha = 0.3, size = 0.5) + 
   facet_wrap(~Subset) +
-  ylab("\u03B1' - \u03B1") + xlab("Log10 sample size\n(number of papers)") + 
+  ylab("\u03B1' - \u03B1") + 
+  xlab("Log10 sample size\n(number of papers)") + 
   theme_minimal()
@@
             alpha.new = alpha.adjusted[1],
-            SE.old = (upper.95.CI2[2] - lower.95.CI2[2]) / 2 * 1.96,
-            SE.new = (upper.95.CI2[1] - lower.95.CI2[1]) / 2 * 1.96,
+            SE.old = (upper.95.CI2[2] - lower.95.CI2[2]) / (2 * 1.96),
+            SE.new = (upper.95.CI2[1] - lower.95.CI2[1]) / (2 * 1.96),
             lower883.old = alpha.old - SE.old * 1.190,
@@
 **Figure S`r supp.figure.counter`**: Histogram showing the distribution of differences in $\alpha'$ between the 2015-16 and 2005-6 samples, where positive numbers indicate an increase in $\alpha'$ with time.
-The mean is slightly greater than 0, indicating a mild increase in average $\alpha'$ with time.
+The mean is slightly positive (namely `r with(old.vs.new, mean(alpha.new - alpha.old))`), indicating a mild increase in average $\alpha'$ with time.
 
@@
 ```{r}
-summary(lmer(scale(alpha.adjusted) ~ Age + (1|journal) + (1|discipline), # Note that the response variable gets scaled, to make the effect size interpretable as Cohen's d
-             weights = 1 / ((upper.95.CI2 - lower.95.CI2)/2*1.96), # Weight each observation of alpha by its precision
+# Note that the response variable gets scaled, to make the effect size interpretable as Cohen's d
+summary(lmer(scale(alpha.adjusted) ~ Age + (1|journal) + (1|discipline), 
+             weights = 1 / ((upper.95.CI2 - lower.95.CI2)/(2*1.96)), # Weight each observation of alpha by 1/SE
              data = master.dataset %>% 
@@
 summary(lmer(scale(alpha.adjusted) ~ Subset + (1|journal) + (1|discipline), 
-             weights = 1 / ((upper.95.CI2 - lower.95.CI2)/2*1.96),
+             weights = 1 / ((upper.95.CI2 - lower.95.CI2)/(2*1.96)),
              data = first.vs.last.data))
@@
   geom_point(alpha = 0.4, size = 0.7) + 
-  stat_smooth(method = "gam", formula = y ~ s(x, k = 3), colour = "tomato", fill = "skyblue") + 
-  xlab("Deviation from gender parity for all authors in this journal") + 
+  stat_smooth(method = "gam", 
+              formula = y ~ s(x, k = 3), 
+              colour = "tomato", fill = "skyblue") + 
+  xlab("% women authors (realtive to gender parity)") + 
   ylab("Coefficient of homophily (\u03B1')") +
@@
   theme(legend.position = "none") 
-ggsave(plot = gender.ratio, filename = "figures/figure3.pdf", device = cairo_pdf, width=7, height=5)
+ggsave(plot = gender.ratio, filename = "figures/figure3.pdf", 
+       device = cairo_pdf, width=7, height=5)
 gender.ratio

### 2018-05-27 manuscript/Manuscript_Claire.Rmd
Some writing

@@
 ---
-title: "Do researchers preferentially collaborate with same-gendered colleagues?"
+title: "Do researchers preferentially collaborate with colleagues of the same gender?"
 author:
@@
   \fancyhead[LO,LE]{\textsl{\leftmark}} 
-  \rhead[]{Conference gender ratios} 
+  \rhead[]{Gender and coauthorship} 
   \usepackage[noabbrev,capitalise]{cleveref}
@@
 abstract: Evidence suggests that women in academia are hindered by conscious and unconscious biases, and often feel excluded from formal and informal opportunities for research collaboration.
-In addition to ensuring fairness and helping redress gender imbalance in the academic workforce, increasing women's access to collaboration could help scientific progress by drawing on more of the available human capital.
+In addition to ensuring fairness and helping to redress gender imbalance in the academic workforce, increasing women's access to collaboration could help scientific progress by drawing on more of the available human capital.
 Here, we test whether researchers preferentially collaborate with same-gendered colleagues, using more stringent methods and a larger dataset than in past work.
@@
 
-\maketitle\unmarkedfntext{*School of BioSciences, The University of Melbourne, Victoria, Australia}
+\maketitle\unmarkedfntext{*School of BioSciences, The University of Melbourne, Victoria, Australia.}
 \maketitle\unmarkedfntext{§Organismal and Evolutionary Biology Research Programme,
-Faculty of Biological and Environmental Sciences, University of Helsinki, Finland}
+Faculty of Biological and Environmental Sciences, University of Helsinki, Finland.}
+
 \newpage
+# Introduction
 
-# Introduction
+- Remember dataset S1 issue (put in ESM?)
+- re-run model on line 912 and 931 (weights issue)
+- check figure legend line 900. Also check the 'significant' area hasn't changed loads
+- proportional alpha?
+
 Women are substantially underrepresented in many branches of the workforce in science, technology, engineering, mathematics, and medicine (STEMM), and face additional challenges and inequities relative to men [e.g.
-@Lariviere_2013; @West_2013; @Elsevier_report; @Holman_2018].
-On average, women occupy more junior positions [@Wutte_2007; @Reuben_2014] with lower salaries [@Trower_2002; @Umbach_2007], receive less grant money [@Hosek_2005; @OER_2008], are promoted more slowly [@Zuckerman_1987; @Rosenfeld_1991; @Long_1993; @Hopkins_2013], and are allocated fewer resources [@ODorchai_2009] and less research funding [@Feldt_1986, @Stack_2004, @Lariviere_2011] than men.
-Experimental studies have shown that people regard women's research achievements less favourably than identical achievements by men [@Moss_2012; @Knobloch_2013].
+@Shaw_2012; @Lariviere_2013; @West_2013; @Elsevier_report; @Holman_2018].
+On average, women occupy more junior positions [@Wutte_2007; @Reuben_2014] with lower salaries [@Trower_2002; @Umbach_2007], receive less grant money [@Hosek_2005; @OER_2008], are promoted more slowly [@Zuckerman_1987; @Rosenfeld_1991; @Long_1993; @Hopkins_2013], and are allocated fewer resources [@ODorchai_2009] and less research funding [@Feldt_1986; @Stack_2004; @Lariviere_2011] than men.
+Experimental studies suggest that women's research achievements are regarded less favourably than identical achievements by men [@Moss_2012; @Knobloch_2013].
 
-Because publishing, networking and collaboration are all instrumental to scientific productivity and thus, academic career advancement [@Lee_2005; @Wuchty_2007; @Abramo_2009; @Lariviere_2015], dozens of studies have tested for gender differences in these areas [reviewed in Table S1 of @Holman_2018; see also @Bozeman_2011; @Abramo_2013; @Badar_2013; @Long_1992].
-For example, studies have shown that relative to men, women tend to be less involved in international collaboration [@Lewison_2001; @Webster_2001; @Bozeman_2004; @Lariviere_2011; @Abramo_2013], collaborate less with researchers from their own university department [@Webster_2001], have less prestigious collaborations [@Long_1990], and fewer collaborations in general [@Fuchs_2001].
-These gender differences in collaboration practice presumably have multiple concurrent causes, which might include implicit and explicit gender bias [REFS], differential family obligations [@Reskin_1978; @Long_1990; @Wright_2003], concerns relating to sexual harassment [REFS], lower participation of women in international conferences [@Lewison_2001; Eisen 2014; @Martin_2014], and reduced receipt of research travel funds by women [@Bozeman_2004].
+Because publishing, networking and collaboration are instrumental to scientific productivity and academic career advancement [@Lee_2005; @Wuchty_2007; @Abramo_2009; @Lariviere_2015], dozens of studies have tested for gender differences in these areas [e.g.
+@Long_1992; @Bozeman_2011; @Abramo_2013; @Badar_2013; references in Table S1 of @Holman_2018].
+For example, studies have found that women tend to be less involved in international collaboration [@Lewison_2001; @Webster_2001; @Bozeman_2004; @Lariviere_2011; @Abramo_2013], collaborate less within their own university departments [@Webster_2001], have less prestigious collaborations [@Long_1990], and fewer collaborations in total [@Fuchs_2001].
+These gender differences in collaboration practice presumably have multiple concurrent causes, which might include implicit and explicit gender bias [@Moss_2012], differential family obligations [@Reskin_1978; @Long_1990; @Wright_2003], concerns relating to sexual harassment [@jagsi_2016], and unequal access to conferences [@Martin_2014] and travel funds [@Bozeman_2004].
 
@@
 For reasons such as these, several studies have tested for gender differences in collaboration frequency or pattern by examining the author lists of published research.
-To our knowledge, every published study of this question has concluded that men co-publish with other men, and women with women, more often than expected if collaborators assort randomly with respect to gender [@Ferber_1980; @McDowell_1992; @Ghiasi_2015; @Crow_2015;@Fahmy_2017; @Zettler_2017; @Jadidi_2017; @Teele_2017; @Araujo_2017a; @Araujo_2017b].
+To our knowledge, every published study of this question suggests that men co-publish with other men, and women with women, more often than expected if collaborators assort randomly with respect to gender [@Ferber_1980; @McDowell_1992; @Ghiasi_2015; @Crow_2015; @Fahmy_2017; @Zettler_2017; @Jadidi_2017; @Teele_2017; @Araujo_2017a; @Araujo_2017b].
 This pattern of assortative publishing has been termed 'gender homophily'.
 
-However, we believe that these prior studies of gender homophily were hindered by a largely unacknowledged statistical issue that we term the Wahlund effect (\autoref{wahlund_plot}), meaning that the true extent of homophily is difficult to ascertain from current data.
+However, we believe that prior studies of gender homophily were hindered by a largely unacknowledged statistical issue that we term the Wahlund effect (\autoref{wahlund_plot}), by analogy with a conceptually similar effect in population genetics [@Wahlund_1928].
+The Wahlund effect makes it deceptively hard to measure gender homophily.
 Essentially, whenever coauthorship data are sampled from two or more discrete sets of literature, which vary in the author gender ratio and which are largely not connected by collaboration, the number of same-gendered coauthors will be inflated.
@@
 For example, a sample of literature containing bioinformatics and cell biology papers will probably contain an excess of mostly-male and mostly-female author lists, simply because researchers preferentially work with colleagues from the same discipline, and because the author gender ratio is more male-biased in bioinformatics than in cell biology [@Holman_2018].
-We chose the name by analogy with the Wahlund effect in population genetics [@Wahlund_1928], whereby spatial differences in allele frequencies create an excess of homozygotes across the whole population, even if alleles are randomly assorted within each subpopulation (original idea in BERGSTROM REF).
 
@@
 # Methods
-## Obtaining author lists and assigning gender
+## The dataset
 We used the dataset of PubMed author lists from Holman et al.
 [-@Holman_2018].
-Briefly, that dataset was created by downloading every single article indexed on PubMed, from the oldest articles up to the present (which was August 2016), and attempting to assign a gender to each author based on their given name, using the web service genderize.io.
-Each journal was assigned to one of 107 scientific disciplines (e.g.
-_Nephrology_ or _Cell Biology_), using PubMed's journal categorisations where available.
-Because the present study focuses on co-authorship, all single-author papers were removed from the dataset.
-To simplify the statistical analysis, we also discarded all papers for which we could not determine the gender of every author with at least 95% certainty.
+Briefly, that dataset was created by downloading every article indexed on PubMed, from the oldest articles up to the present, and attempting to assign a gender to each author based on their given name.
+Each journal was assigned to one of 107 scientific disciplines, using PubMed's journal categorisations where available.
+Because the present study focuses on co-authorship, all single-author papers were discarded.
+We also discarded all papers for which we could not determine the gender of every author with ${\ge}95\%$ certainty, in order to simplify the statistical analysis.
 To mitigate Wahlund effects caused by variation in the gender ratio of researchers over time (see below), we also discarded all papers except those that were published either 0-1 or 10-11 years before the PubMed data were collected (i.e.
-$20^{th}$ August 2016).
+20^th^ August 2016).
 Lastly, we excluded journals with fewer than 50 suitable papers.
-This left us with a dataset from August 2015 - August 2016 containing `r sample.size$n2015[3]` papers with `r sample.size$n2015[4]` authorships, and a dataset from August 2005 - August 2006 containing `r sample.size$n2005[3]` papers and `r sample.size$n2005[4]` authorships.
-These papers came from `r sample.size$n2015[2]` and `r sample.size$n2005[2]` journals respectively, which were grouped into `r sample.size$n2015[1]` and `r sample.size$n2005[1]` research disciplines.
+
+The final sample sizes were `r sample.size$n2015[3]` papers and `r sample.size$n2015[4]` authorships from August 2015 - August 2016, and `r sample.size$n2005[3]` papers and `r sample.size$n2005[4]` authorships from August 2005 - August 2006.
+Respectively, these papers came from `r sample.size$n2015[2]` and `r sample.size$n2005[2]` journals, which were grouped into `r sample.size$n2015[1]` and `r sample.size$n2005[1]` research disciplines.
 There was a median of `r sample.size$n2015[5]` (`r sample.size$n2005[5]`) papers per journal, `r sample.size$n2015[6]` (`r sample.size$n2005[6]`) authors per journal, and `r sample.size$n2015[7]` (`r sample.size$n2005[7]`) authors per paper (the first number is for 2015-6, and the bracketed number for 2005-6).
@@
 To estimate $\alpha$ for a particular subset of the scientific literature, we estimated $p$ as the average proportion of men's co-authors who are men (averaged across all papers with at least one man author), and $q$ as the average proportion of women's co-authors who are men (averaged across all papers with at least one woman author).
-To estimate the 95% confidence intervals on $\alpha$ for a given set of $n$ papers, we sampled $n$ papers with replacement 1000 times, estimated $\alpha$ on each sample, and recorded the 95% quantiles of the 1000 resulting estimates.
+To estimate the 95% confidence intervals on $\alpha$ for a given set of $n$ papers, we sampled $n$ papers with replacement 1000 times, estimated $\alpha$ on each sample, and recorded the 95% quantiles of the resulting 1000 estimates.
 
@@
 We did not calculate $\alpha$ for other authorship positions (e.g.
-second or third authors) because this would necessitate reducing the dataset to papers with a sufficiently long authorship list, complicating comparison with other authorship positions that were calculated using a different sample.
+second or third authors) because this would necessitate culling the dataset to include only papers with a sufficiently long author list, complicating interpretation of the results.
 
@@
 To borrow Prof.
-Bergstrom's example, consider a small research specialty comprising just two men and two women researchers, who have together produced six two-author papers: one paper in each of the six possible two-author combinations.
+Bergstrom's example, consider a small research specialty comprising just two men and two women researchers, who have together produced six two-author papers: one in each of the six possible two-author combinations.
 For these six papers, $\alpha = -\frac{1}{3}$, even though same- and opposite-gendered coauthors were selected in equal proportion to their frequency in the pool of possible collaborators.
-To control for the fact that the null expectation for $\alpha$ is not necessarily zero (particularly for very small datasets), we devised an adjusted version of the coefficient of homophily, which we term $\alpha'$.
+To control for the fact that the null expectation for $\alpha$ is not zero for small datasets, we devised an adjusted version of the coefficient of homophily, which we term $\alpha'$.
 Every time we calculated $\alpha$ for a set of papers, we also determined the expected value of $\alpha$ under the null hypothesis that authors assort randomly with respect to gender.
@@
 We also used the null-simulated $\alpha$ values to calculate a two-tailed p-value for the observed value of $\alpha$; the p-value was defined as the proportion of null simulations for which $|\alpha_{null}| > |\alpha_{obs}|$.
-We applied false discovery rate correction to each set of p-values to account for multiple testing following Benjamini and Hochberg [-@Benjamini_1995].
+We applied false discovery rate (FDR) correction to each set of p-values to account for multiple testing [@Benjamini_1995].
 
@@
 Additionally, the correlation between $\alpha'$ and sample size was negligible ($R^2 < 0.01$), suggesting that our calculation of $\alpha'$ effectively removed the dependence of $\alpha$ on sample size.
-We therefore use the adjusted statistic $\alpha'$ in all analyses.
+We therefore used $\alpha'$ in all analyses.
 
-## Minimising the Wahlund effect: discipline and time period
-As discussed in the Introduction and shown in \autoref{wahlund_plot}, the Wahlund effect can give the appearance of gender homophily (reflected in our study as $\alpha' > 0$) whenever one lumps together different 'subpopulations' of academic papers that differ in their author gender ratios.
-
+## Minimising the Wahlund effect: research discipline and time period
 To minimise bias in $\alpha'$ due to the Wahlund effect, we restricted each set of papers to a single research specialty to the greatest extent allowed by our data.
@@
 The difference in $\alpha'$ between highly multidisciplinary and more specialised journals, e.g.
-_PLoS ONE_ versus _PLoS Computational Biology_, gives a rough estimate of the extent to which multidisciplinarity can inflate $\alpha'$.
+_PLoS ONE_ versus _PLoS Computational Biology_, gives a rough estimate of the extent to which multidisciplinarity inflates $\alpha'$.
 
 As well as varying between disciplines, the gender ratio of authors has changed markedly over time [@Holman_2018].
-Because the gender ratio was more male-biased in the past, $\alpha'$ would be inflated if we calculated it for a sample of papers published over a long time frame.
+Because the gender ratio was more male-biased in the past, $\alpha'$ would be inflated if we calculated it for a sample of papers published over a long enough time frame.
 To minimise this effect, we only sampled papers from two one-year periods (namely 2005-6 and 2015-16).
-The median change per year in % (fe)male authors across journals is < 0.5% [@Holman_2018], and so restricting our dataset to a single year should prevent temporal changes in gender ratio from noticeably inflating our estimates of $\alpha'$.
+The median change per year in % (fe)male authors across journals is below 0.5% [@Holman_2018], and so restricting our dataset to a single year should prevent temporal changes in gender ratio from noticeably affecting our estimates of $\alpha'$.
 
@@
 Thus, journals with a positive standardised impact factor have a higher mean number of citations than the average for journals in their discipline.
-We then used Spearman rank correlation to test whether our estimates of gender homophily covaried with journal impact factor.
+We then used Spearman rank correlation to test whether $\alpha'$ was correlated with impact factor across journals.
 
 ## Statistical analysis
-Previous authors (e.g.
-@Bonham_2017) have hypothesised that senior scientists preferentially recruit staff and students of the same gender, and/or that junior researchers preferentially select same-gendered mentors.
+Previous authors [e.g.
+@Bonham_2017] have hypothesised that senior scientists preferentially recruit staff and students of the same gender, and/or that junior researchers preferentially select same-gendered mentors.
 In the majority of disciplines, authorship conventions mean that the first-listed author is often an early-career researcher, while the author listed last is more likely to be a senior researcher leading a research team [@Wren_2007].
@@
 
-## Theoretical expectations for $\alpha$ when there is a gender gap between career stages
+## Theoretical expectations for $\alpha$ when the gender ratio differs between career stages
 In many STEMM subjects, the gender ratio is more skewed among established researchers relative to early-career researchers [@Shaw_2012; @Holman_2018].
 We hypothesised that this skew could potentially create Wahlund effects and 'reverse' Wahlund effects.
-For example, imagine that the majority of collaborations are between students and professors, and that the gender ratio is skewed in opposite directions at each career stage: we will then see an excess of mixed-gender coauthorships (heterophily, $\alpha < 0$), even if gender has no causal effect on the selection of coauthors.
-Similarly, a hypothetical field in which students work only with students, and professors with professors, will have apparent gender homophily ($\alpha > 0$).
+For example, imagine that the majority of collaborations are between students and professors, and that the gender ratio differs between career stages: we will then see an excess of mixed-gender coauthorships (heterophily, $\alpha < 0$), even if gender has no causal effect on the selection of coauthors.
+Similarly, a hypothetical field in which students work only with students, and professors with professors, would have apparent gender homophily ($\alpha > 0$).
 
-Because we can think of no tractable way to control for this issue using our dataset (which contains no information on career stage), we instead decided to derive the theoretical expectations for $\alpha$ when there is a difference in gender ratio across career stages.
-For simplicity, we assume there are only two career stages, though the results are likely to be general.
-Under the null model that gender has no causal effect on collaboration, we calculated $\alpha$ for various combinations of the four free parameters: the gender ratios for early- and late-career researchers, and the relative frequency of collaborations between early-early, early-late, and late-late collaborations.
-We then used the theoretical expectations for $\alpha$ to properly qualify our main conclusions (see Results).
-The R code used to derive the theoretical expectations can be found in the Online Supplementary Material.
+Because we can think of no tractable method of controlling for this issue using our dataset, which contains no information on career stage.
+Therefore, we instead decided to derive the theoretical expectations for $\alpha$ when there is a difference in gender ratio across career stages.
+For simplicity, our calculations assume there are only two career stages, though the conclusions are likely to generalise to a more realistic multi-tiered career ladder.
+Under the null model that gender has no causal effect on collaboration, we calculated $\alpha$ for various combinations of the four free parameters, i.e.
+the gender ratios for early- and late-career researchers, and the relative frequency of collaborations between early-early, early-late, and late-late collaborations.
+We then used the theoretical expectations for $\alpha$ to qualify our main conclusions (see Results).
+Annotated R code used to derive the theoretical expectations can be found in the Online Supplementary Material.
 
 ## Data availability and reproducibility
-The Online Supplementary Material is an HTML document prepared with R Markdown, which contains R scripts used to produce all data, figures and tables.
-The input data from Holman et al.
-[-@Holman_2018] is archived at https://osf.io/bt9ya/.
+The Online Supplementary Material contains R scripts used to produce all data, figures and tables. The input data from Holman et al. [-@Holman_2018] is archived at https://osf.io/bt9ya/.
 
@@
 \autoref{alpha_histograms} shows the distribution of $\alpha'$ estimates in 2015-2016 across all journals for which we recovered sufficient data, when $\alpha'$ was calculated for all authors, first authors only, or last authors only.
-The great majority of journals had $\alpha' > 0$, and for many of these the corrected two-tailed p-values suggested that $\alpha'$ was significantly greater than zero (1469/2077 journals were significant in 2015-16, and 404/1192 in 2005-6).
+The great majority of journals had $\alpha' > 0$, and for many of these the FDR-corrected p-values suggested that $\alpha'$ was significantly greater than zero (1469/2077 journals were significant in 2015-16, and 404/1192 in 2005-6).
 Only 2/2077 journals had statistically significantly heterophily (i.e.
 $\alpha' < 0$) in 2015-16, and 1/1192 in 2005-6.
-The remaining 606 or 787 journals (in 2015 and 2005 respectively) had a value of $\alpha'$ not significantly different from zero, such that we could not reject the null hypothesis of random assortment with respect to gender.
+The remaining 606 or 787 journals (in 2015 and 2005 respectively) had a value of $\alpha'$ not significantly different from zero, consistent with the null hypothesis of random assortment with respect to gender.
 
@@
 
-$\alpha'$ was significantly higher in the sample from 2015-16 relative to 2005-6, but the difference in means was small (Figure S3; Effect of the fixed factor 'Time period' in a linear mixed model of the data for all author positions: Cohen's $d$ = $0.09{\pm}0.04$, $t_{948}$ = 2.51, p = 0.012).
+$\alpha'$ was significantly higher in the literature sample from 2015-16 relative to 2005-6, but the difference in means was small (Figure S3; Effect of the fixed factor 'Time period' in a linear mixed model of the data for all author positions: Cohen's $d$ = $0.09{\pm}0.04$, $t_{948}$ = 2.51, p = 0.012).
 
@@
 This suggests that the gender of the first author was a slightly stronger predictor of the remaining authors' genders than the gender of the last author, i.e.
-the opposite of our prediction based on the hypothesis that senior scientists are causally responsible for homophily.
+the opposite of what is predicted if senior scientists are causally responsible for homophily.
 
@@
 mean $\alpha'$ was $0.12{\pm}0.02$ for Urology journals, and $0.03{\pm}0.01$ for Veterinary Medicine journals; \autoref{alpha_histograms}, Table S3).
-However, there was little evidence for strong differences in $\alpha'$ between disciplines: the random factor 'Discipline' explained at most 1% of the variance in $\alpha'$ in the two linear mixed models described in the previous section (see \autoref{alpha_histograms} and mixed models in Online Supplementary Material).
+However, there was no evidence for consistent differences in $\alpha'$ between disciplines: the random factor 'Discipline' explained at most 1% of the variance in $\alpha'$ in the two linear mixed models described in the previous section (see \autoref{alpha_histograms} and mixed models in Online Supplementary Material).
 This implies that the processes causing positive $\alpha'$ are similarly strong in all the disciplines we examined.
@@
 ## Relationship between gender homophily and gender ratio
-We next tested whether researchers are more or less likely to seek out same-gendered colleagues in strongly gender-biased disciplines (e.g.
+We next tested whether researchers are more or less likely to publish with same-gendered colleagues in strongly gender-biased disciplines (e.g.
 Surgery or Nursing), relative to disciplines with a comparatively gender-balanced workforce (e.g.
 Psychiatry).
-We found a positive, non-linear relationship between the overall gender ratio of all authors publishing in a particular journal [as estimated by @Holman_2018], and the estimated value of $\alpha'$ for all authors and for first authors (\autoref{alpha_gender_ratio}).
-Journals with a balanced or female-biased author gender ratio tended to have higher $\alpha'$ than journals with a male-biased author gender ratio (GAM smooth terms p < 0.001, see Online Supplementary Material).
+We found a positive, non-linear relationship between the overall gender ratio of all authors publishing in a particular journal [as estimated in @Holman_2018], and the estimated value of $\alpha'$ for all authors and for first authors (\autoref{alpha_gender_ratio}).
+Journals with a balanced or female-biased author gender ratio tended to have higher $\alpha'$ than journals with a male-biased author gender ratio (GAM smooth terms p < 0.001; Online Supplementary Material).
 The relationship was not statistically significant when $\alpha'$ was calculated for last authors (GAM, p = 0.142), though the trend appeared similar (\autoref{alpha_gender_ratio}).
@@
 ## Relationship between journal impact factor and gender homophily
-We observed a noisy but statistically significant linear relationship between discipline-standardised journal impact factor and $\alpha'$, such that journals with a high impact factor for their discipline had weaker gender homophily than did journals with a low impact factor for their discipline (\autoref{impact_factor}; linear regression: $R^2$ = `r R2`, $t_{1415}$ = -8.0, p < 0.0001).
+We observed a noisy but statistically significant linear relationship between standardised journal impact factor and $\alpha'$, such that journals with a high impact factor for their discipline had weaker gender homophily than did journals with a low impact factor for their discipline (\autoref{impact_factor}; linear regression: $R^2$ = `r R2`, $t_{1415}$ = -8.0, p < 0.0001).
 
@@
 ## Analysis accounting for differences in author gender ratio between countries
-When we restricted the analysis to only include authors with affiliations from a single country, we found statistically significant homophily for 72 of the 325 journal-country combinations tested (64 unique journals and 18 unique countries), and no significant evidence of heterophily (Figures S5-S6).
+When we restricted the analysis to only include authors with affiliations from a single country, we observed statistically significant homophily for 72 of the 325 journal-country combinations tested (64 unique journals and 18 unique countries), and no significant evidence of heterophily (Figures S5-S6).
 Additionally, the values of $\alpha'$ calculated for each journal-country combination were mostly very similar to the $\alpha'$ values calculated for the journal as a whole (i.e.
 when pooling papers from different countries); the average difference in $\alpha'$ was 0.002 (Figure S7).
-These results suggest that our findings of widespread homophily in the main analysis cannot solely be driven by a Wahlund effect resulting from gender differences between countries.
+These results suggest that our findings of widespread homophily in the main analysis were not driven by a Wahlund effect resulting from gender differences between countries.
 
-## Theoretical expectations for $\alpha$ when there is a gender gap between career stages
-As shown in \autoref{simulation}, we confirmed that $\alpha$ is expecte to be non-zero, even if collaborators are selected in a gender-blind fashion, if there is a gender gap between career stages.
-The deviation of $\alpha$ from zero depends on the relative frequencies of collaboration within and between career stages.
-When >50% of collaborations were between early and established researchers, we observed gender heterophily ($\alpha < 0$).
-Conversely, when >50% of collaborations occured within career stages, we observed gender homophily ($\alpha > 0$).
+## Theoretical expectations for $\alpha$ when the gender ratio differs between career stages
+As shown in \autoref{simulation}, $\alpha$ is expected to be non-zero, even if collaborators are randomly selected with respect to gender, provided that there is a gender gap between career stages.
+The extent to which $\alpha$ deviates from zero depends on the relative frequencies of collaboration within and between career stages.
+When >50% of collaborations were between early and established researchers, we expect gender heterophily ($\alpha < 0$).
+Conversely, when >50% of collaborations occured within career stages, we expect gender homophily ($\alpha > 0$).
 In a few parameter spaces (shown in red; \autoref{simulation}), $\alpha$ was quite high, and overlapped with the values that we estimated (\autoref{alpha_histograms}).
@@
 
-\autoref{simulation} suggests that our main conclusions are likely to be robust to this career stage issue.
-If A) the gender ratio is highly skewed across career stages (e.g.
-50% vs 5% women among early- vs late-career researchers), and B) collaborations between early and established researchers are very rare (e.g.
-1% of the total), then $\alpha$ can approach the levels shown in \autoref{alpha_histograms} even if collaborators are selected randomly with respect to gender.
-The first of these conditions can be rejected for the majority of research fields [@Shaw_2012; @Holman_2018], and the second seems improbable because it is very common for early-career researchers to co-publish with an established mentor.
-However, one can get $\alpha > 0$ for plausible combinations of parameters, e.g.
-a moderate gender gap coupled with a moderate excess of within career stage collaboration; thus, this effect may explain some of the observed homophily.
+Despite this overlap, \autoref{simulation} suggests that our main conclusions (and those of other studies of gender homophily) are probably robust to this career stage issue.
+We only expect strongly positive $\alpha$ when A) the gender ratio is highly skewed across career stages (e.g.
+a 10-fold difference), and B) collaborations between early and established researchers are very rare (e.g.
+<10% of the total).
+Both of these conditions are unlikely: the gender gap across careers stages is generally less than 10-fold [@Shaw_2012; @Holman_2018], and it is very common for early-career researchers to co-publish with an established mentor [@macaluso_2016].
+However, one can get $\alpha > 0$ for realistic combinations of parameters, e.g.
+a moderate shortage of women in senior positions coupled with a moderate excess of within-career stage collaboration, suggesting this effect might contribute to some of the observed homophily.
 
-Additionally, if there is a large gender gap between career stages and collaboration between early-career and established researchers are commonplace, then the baseline expectation for $\alpha$ is less than zero.
-Therefore, for disciplines where these two conditions are met, our results will under-estimate the extent to which researchers preferentially select same-gendered collaborators.
+Lastly, we note that if there is a gender gap between career stages and coauthorships between early-career and established researchers comprise ${\ge}50\%$ of the total, then the baseline expectation for $\alpha$ is actually less than zero (blue areas in \autoref{simulation}).
+Therefore, our results might under-estimate the extent to which researchers preferentially select same-gendered collaborators in some cases.
 
 # Discussion
-We found evidence that researchers publish with same-gendered coauthors more often than expected under random assortment, even after implementing stringent controls for 'spurious' homophily resulting from Wahlund effects (\autoref{wahlund_plot}).
-Our data therfore reaffirm and refine the results of earlier studies finding evidence of gender homophily [@Ferber_1980; @McDowell_1992; @Bentley_2003; @Crow_2015; @Ghiasi_2015; @Araujo_2017a; @Fahmy_2017; @Jadidi_2017; @Teele_2017; @Zettler_2017].
-The excess of same-gender coauthorships was quite strong: many journals had $\alpha' > 0.1$, indicating that the gender ratio of the average man's coauthors is skewed towards men by >10%, relative to that of the average woman.
-Few journals had $\alpha'$ values below zero, and almost no journals showed statistically significant gender heterophily after controlling for multiple testing.
-The results therefore suggest that it is comparatively rare for researchers to preferentially select opposite-gendered co-authors.
+We found evidence that researchers publish with same-gendered coauthors more often than expected under random assortment, even after implementing stringent controls for 'spurious' gender homophily resulting from Wahlund effects (\autoref{wahlund_plot}).
+Our study therefore reaffirms several earlier results and establishs their generality across the life sciences [@Ferber_1980; @McDowell_1992; @Bentley_2003; @Crow_2015; @Ghiasi_2015; @Araujo_2017a; @Fahmy_2017; @Jadidi_2017; @Teele_2017; @Zettler_2017].
+Relatively few journals had $\alpha'$ values below zero, and almost no journals showed statistically significant gender heterophily after controlling for multiple testing.
+The excess of same-gender coauthorships was quite strong: many journals had $\alpha' > 0.1$, indicating that the gender ratio of men's and women's coauthors differs by >10% in absolute terms.
+In relative terms, this effect appears even more striking: for example, if the proportion of male coauthors is 80% for men and 70% for women ($\alpha' = 0.1$), the results imply that women publish with women 50% more often than men do.
 
-An important limitation of our study (and all other similar studies) is that we cannot reliably ascertain what caused the excess of same-gender coauthorships.
-As well as conscious or unconcious choice by researchers, there could be uncontrolled-for Wahlund effects.
-We suspect the contribution of these to be minor, since we found positive $\alpha'$ after controlling for three obvious sources of Wahlund effect, but it is not possible to be sure.
+An important limitation of our study is that we cannot reliably ascertain the cause(s) of the observed excess of same-gender coauthorships.
+As well as conscious or unconcious selection of collaborators based on gender, our results could be partly or wholly explained by uncontrolled Wahlund effects.
+We suspect the contribution of these to be minor, for four reasons: we found positive $\alpha'$ after controlling for three obvious sources of Wahlund effect; there was no inflation of $\alpha'$ in highly multidisciplinary journals; restricting the data by country yielded similar estimates of $\alpha'$; and we showed that differences in gender ratio between career stages are unlikely to fully explain our results.
+On balance, we believe the data suggest that researchers preferentially select same-gendered collaborators, the magnitude of this effect is uncertain.
 
@@
 We hypothesised that these gender-imbalanced disciplines might show the strongest gender homophily, e.g.
-because one needs to exercise stronger choice (with respect to gender) if one enjoys working with same-gendered collaborators, if one's field has a large gender gap.
-By contrast, researchers in fields without a gender gap will have many same-gender collaborators just by collaborating randomly with respect to gender.
+because being in the minority might increase one's drive to seek out same-gendered colleagues.
 Contrary to this hypothesis, we found no evidence that gender homophily is restricted to particular disciplines: $\alpha'$ was similarly high across the board \autoref{alpha_histograms}.
-Moreover, $\alpha'$ was _lower_ for journals with a male-biased author gender ratio relative to those with an even gener ratio, suggesting that women are _less_ likely to preferentially seek out women colleagues in male-biased research disciplines.
-Interestingly though, $\alpha'$ was highest for journals with a female-biased author gender ratio.
-One interpretation of this result is that men are more likely to preferentially select male collaborators in fields where men are a minority, relative to the homophily displayed by women in fields where women are the minority.
-However, this latter result is only tentatively supported by our data, since there are so few journals on PubMed with mostly-female authors \autoref{alpha_gender_ratio}.
+Moreover, $\alpha'$ was lower for journals with a male-biased author gender ratio relative to those with an even gender ratio.
+Interestingly, $\alpha'$ was highest in journals with a female-biased author gender ratio.
+One interpretation of this result is that men are more likely to preferentially seek out male collaborators in fields where men are a minority, relative to the homophily displayed by women in fields where women are a minority.
+However, this latter result is only tentatively supported since there are so few journals on PubMed in which most authors are women \autoref{alpha_gender_ratio}.
 
 We also found that gender homophily was marginally stronger in 2015-2016, relative to 2005-2006.
-This suggests that recent increases in the number of women working in STEMM, as well as recent efforts to recruit, retain, and celebrate women in STEMM, have co-occurred with a rising frequency of same-gender collaborations.
-Although this trend might reflect a change in the gender preferences of researchers seeking collaborators, there are other (probably more likely) explanations that could explain this result.
-Specifically, we speculate that this trend results from the increasing number of women working in senior positions in STEMM over the past decade [e.g.
+Although this trend might reflect a change in the gender preferences of researchers seeking collaborators over time, there are alternative (and probably more likely) explanations for this result.
+For example, this trend might result from the increasing number of women working in senior positions in STEMM over the past decade [e.g.
 @Long_2015; @Bendels_2018; @McKenzie_2017].
-Because many co-authorships involve a junior and a senior researcher (e.g.
-a student and professor), and the gender ratio is typically more male-biased among senior staff than junior staff [@Shaw_2012; @Holman_2018], there will be an excess of opposite-gender collaborations relative to the expectation when there is no difference in gender ratio across career stages.
-As the gender ratio equalises across career stages, this effect will weaken, causing $\alpha'$ to rise.
+As shown in \autoref{simulation}, if enough coauthorships are between junior and senior researchers, a large gender gap between career stages can give the appearance of heterophily.
+As this gender gap between career stages lessens, the observed values of $\alpha$ can increase.
 
-We found that journals with a high impact factor (relative to other journals in the same research discipline) tended to have weaker gender homophily than did low-impact journals \autoref{impact_factor}.
-If one accepts the contentious claim that papers published in high-impact journals tend to be of higher quality than those in low-impact journals [e.g.
-@Garfield_2006], then this result provides non-experimental support for the hypothesis that mixed-gender teams produce better research than single-gender teams [@Campbell_2013].
+Our study begs two questions: what causes gender homophily in science, and are our results cause for concern?
+These questions are closely related.
+For example, some of the homophily we observed might be caused by women seeking to avoid harassment or bullying from men [REF], which is clearly very concerning.
+Additionally, Sheltzer and Smith [-@sheltzer_2014] concluded that 'elite' male academics (i.e.
+recipients of major honours) have a higher proportion of male students and postdocs than non-elite male academics.
+This finding could contribute to the homophily we observed, and is cause for concern because the results might reflect discrimination against women by established scientists [@Moss_2012], or avoidance by women of elite research groups (e.g.
+due to 'imposter syndrome' or the perception that these groups are a 'boy's club').
+We also found a little evidence that gender homophily is deterimental to research quality, in that high-impact journals tended to have weaker homophily.
+Assuming that papers published in high-impact journals are of higher average quality [which is contentious; @Garfield_2006], then this result provides non-experimental support for the hypothesis that mixed-gender teams produce better research than single-gender teams [@Campbell_2013].
+
+Homophily might also have more benign causes.
+Collaboration is often especially enjoyable when working with like-minded people, who might be same-gendered more often than not.
+We also suppose that some people -- women especially -- consciously choose to preferentially collaborate with women whenever multiple qualified candidates are available, in order to help close the gender gap in the workforce.
+Related to this, Ghiasi et al.
+[-@Ghiasi_2015] wrote that women who mostly work with men are 'compliant [in reproducing] male-dominated scientific structures', although we think this charge is unfair.
+We believe that everyone (not only women) should take steps to ensure that they are not overlooking or excluding female colleagues, particularly since men are more likely to occupy stable, senior jobs that afford them the luxury of choice.
+
+----------
 
@@
 family commitments) represent the principal restrictions on their international mobility.
-As a result women are less likely to originate collaborations with foreign colleagues [@Lemoine_1992; @Lewison_2001; @ Webster_2001; @ Lariviere_2011].
+As a result women are less likely to originate collaborations with foreign colleagues [@Lemoine_1992; @Lewison_2001; @Webster_2001; @Lariviere_2011].
 Lastly, men and women think differently on average (REF) and are more likely to initiate collaborations based on similarity of thoughts to progress their scientific output (REF).
@@
 Because women make up for as little as 30% of the STEMM scientist community [@Holman_2018], they are less likely to be picked as coauthors, and, as a result, gender homophily creates further gender imbalance and disadvantages for women.
-Men tend to publish more in high impact journals as first author [@West_2013], and generate more citations than papers lead by first author women [@Long_1992; @Lariviere_2013].
+Men tend to publish more in high impact journals as first author [@Holman_2018], and generate more citations than papers lead by first author women [@Long_1992; @Lariviere_2013].
 As a consequence, gender homophily is also likely to contribute to lower publication rates for women [@McDowell_1992].

### 2018-05-27 Gender homophilly project.Rmd
Tweaks

@@
 
-```{r}
+```{r warning=FALSE}
 simulate <- function(params){
@@
 
-  # given the gender ratio at career stages a and b, find the proportion of collaborations that are man-man, woman-woman, or mixed gender
+  # given the gender ratio at career stages a and b, find the 
+  # proportion of collaborations that are man-man, woman-woman, or mixed gender
   man.man <- function(GR_a, GR_b) (1-GR_a) * (1-GR_b)
@@
 
-  # Find p and q, which are used to calcualte alpha
-
+  # Find p and q, which are used to calculate alpha:
   # p is proportion of men coauthoring with men
@@
          prop.student.student = prop.student.student * 100) %>%
-  filter(!duplicated(paste(GR.stage1,GR.stage2,prop.student.prof,prop.student.student)))
+  filter(!duplicated(paste(GR.stage1, GR.stage2, prop.student.prof, prop.student.student)))
 
 simulation.figure <- results %>%
-  ggplot(aes(GR.stage1, GR.stage2, z = alpha, fill = alpha)) + 
+  ggplot(aes(GR.stage1, GR.stage2, z = round(alpha,2), fill = alpha)) + 
   facet_grid(prop.student.prof ~ prop.student.student) + 
   geom_tile() + 
-  #stat_contour(binwidth = 0.1, colour = "grey20", size = 0.3) + 
+  geom_abline(colour = "grey20", size = 0.3) +
+  stat_contour(breaks = c(-5:-1, 1:5)/10, colour = "grey20", size = 0.3) + 
   scale_fill_distiller(name = "\u03B1",

### 2018-05-27 manuscript/Manuscript_Claire.Rmd
Tweaks

@@
 Here, coloured circles represent male and female authors, and coauthors are linked with lines.
-Across the whole set of ten papers, there is an apparent excess of same-gender collaborations.
-Specifically, there are six same-gender papers and only four mixed-gender papers, which is fewer than the $10\times2\times0.5\times0.5 = 5$ mixed-gender papers we would expect under the null hypothesis that authors assort randomly with respect to gender.
+Across the whole set of ten papers, there is an apparent excess of same-gender collaborations: there are six same-gender papers and only four mixed-gender papers, which is fewer than the $10\times2\times0.5\times0.5 = 5$ mixed-gender papers we would expect under the null hypothesis that authors assort randomly with respect to gender.
 However, within each subset, there is no evidence that authors prefer to publish with same-gendered individuals.
-The Wahlund effect will tend to inflate the frequency of same-sex coauthors whenever the data is composed of two or more disconnected subsets of literature with different author gender ratios; these subsets could be research disciplines, older versus newer papers, or papers from authors in different countries.
+The Wahlund effect will tend to inflate the frequency of same-sex coauthorships whenever the data is composed of two or more disconnected subsets of literature with different author gender ratios; these subsets could be research disciplines, older versus newer papers, or papers from authors in different countries.
 \label{wahlund_plot}](../figures/Figure 1 - Wahlund figure.pdf)
@@
 In many STEMM subjects, the gender ratio is more skewed among established researchers relative to early-career researchers [@Shaw_2012; @Holman_2018].
-We hypothesised that this skew could potentially create Wahlund effects and 'reverse' Wahlund effects.
+We hypothesised that this skew could potentially create both Wahlund effects and 'reverse' Wahlund effects.
 For example, imagine that the majority of collaborations are between students and professors, and that the gender ratio differs between career stages: we will then see an excess of mixed-gender coauthorships (heterophily, $\alpha < 0$), even if gender has no causal effect on the selection of coauthors.
@@
 
-Because we can think of no tractable method of controlling for this issue using our dataset, which contains no information on career stage.
-Therefore, we instead decided to derive the theoretical expectations for $\alpha$ when there is a difference in gender ratio across career stages.
-For simplicity, our calculations assume there are only two career stages, though the conclusions are likely to generalise to a more realistic multi-tiered career ladder.
+We can think of no tractable method of controlling for this issue using our dataset, which contains no information on career stage.
+Therefore, we instead decided to derive the theoretical expectations for $\alpha$ when there is a difference in gender ratio across career stages, in order to determine whether this effect could appreciably affect our conclusions.
+For simplicity, our calculations assume there are only two career stages, though the conclusions are likely to generalise to a realistic, multi-tiered career ladder.
 Under the null model that gender has no causal effect on collaboration, we calculated $\alpha$ for various combinations of the four free parameters, i.e.
@@
 We then used the theoretical expectations for $\alpha$ to qualify our main conclusions (see Results).
-Annotated R code used to derive the theoretical expectations can be found in the Online Supplementary Material.
+The Online Supplementary Material gives annotated R code used to derive the theoretical expectations.
 
@@
 However, the conditions required for strong gender homophily are quite restrictive, making it unlikely that this issue explains all of the homophily observed in Figure 2.
+Contour lines denote increments of 0.1.
 \label{simulation}](../figures/figure5.pdf)
@@
 As shown in \autoref{simulation}, if enough coauthorships are between junior and senior researchers, a large gender gap between career stages can give the appearance of heterophily.
-As this gender gap between career stages lessens, the observed values of $\alpha$ can increase.
+As this gender gap between career stages lessens, the observed values of $\alpha$ may increase.
 
@@
 These questions are closely related.
-For example, some of the homophily we observed might be caused by women seeking to avoid harassment or bullying from men [REF], which is clearly very concerning.
+For example, some of the homophily we observed might be caused by women seeking to avoid harassment or bullying from men [@jagsi_2016], which is clearly very concerning.
 Additionally, Sheltzer and Smith [-@sheltzer_2014] concluded that 'elite' male academics (i.e.
@@
 This finding could contribute to the homophily we observed, and is cause for concern because the results might reflect discrimination against women by established scientists [@Moss_2012], or avoidance by women of elite research groups (e.g.
-due to 'imposter syndrome' or the perception that these groups are a 'boy's club').
+due to 'imposter syndrome' or the perception that these groups are hostile to women).
 We also found a little evidence that gender homophily is deterimental to research quality, in that high-impact journals tended to have weaker homophily.

### 2018-05-27 manuscript/Manuscript_Claire.Rmd
Writing

@@
 Interestingly, homophily was negatively correlated with journal impact factor, as predicted if mixed-gender teams produce better research.
+We discuss potential causes of gender homophily in science.
 \vspace{5mm} \par\noindent \textbf{Keywords:} Gender bias, Homophily, Scientific collaboration, Text mining, Women in STEM.
@@
 For example, studies have found that women tend to be less involved in international collaboration [@Lewison_2001; @Webster_2001; @Bozeman_2004; @Lariviere_2011; @Abramo_2013], collaborate less within their own university departments [@Webster_2001], have less prestigious collaborations [@Long_1990], and fewer collaborations in total [@Fuchs_2001].
-These gender differences in collaboration practice presumably have multiple concurrent causes, which might include implicit and explicit gender bias [@Moss_2012], differential family obligations [@Reskin_1978; @Long_1990; @Wright_2003], concerns relating to sexual harassment [@jagsi_2016], and unequal access to conferences [@Martin_2014] and travel funds [@Bozeman_2004].
+These gender differences in collaboration practice presumably have multiple causes, which might include implicit and explicit gender bias [@Moss_2012], differential family obligations [@Reskin_1978; @Long_1990; @Wright_2003], concerns relating to sexual harassment [@jagsi_2016], and unequal access to conferences [@Martin_2014] and travel funds [@Bozeman_2004].
 
 A high, steadily increasing proportion of research papers is written by more than one author [@West_2013], making collaboration a key predictor of publication output, and thus of career prospects [@Tower_2007; @Jordan_2008; @Cohn_2014].
-Additionally, empirical studies suggest that mixed-gender [@Bear_2011; @Campbell_2013] or otherwise diverse [@Hong_2004] teams can produce better results on collaborative tasks than less diverse teams.
+Additionally, empirical studies imply that mixed-gender [@Bear_2011; @Campbell_2013] or otherwise diverse [@Hong_2004] teams produce better results on collaborative tasks than less diverse teams.
 For reasons such as these, several studies have tested for gender differences in collaboration frequency or pattern by examining the author lists of published research.
@@
 However, we believe that prior studies of gender homophily were hindered by a largely unacknowledged statistical issue that we term the Wahlund effect (\autoref{wahlund_plot}), by analogy with a conceptually similar effect in population genetics [@Wahlund_1928].
-The Wahlund effect makes it deceptively hard to measure gender homophily.
+The Wahlund effect makes it deceptively difficult to infer gender-based preferences simply by counting same- and mixed-gender coauthorships.
 Essentially, whenever coauthorship data are sampled from two or more discrete sets of literature, which vary in the author gender ratio and which are largely not connected by collaboration, the number of same-gendered coauthors will be inflated.
-This can give the impression that authors preferentially publish with same-gendered colleagues if no gender preferences exist, or even if researchers actually prefer opposite-gendered colleagues.
+This can give the impression that authors preferentially publish with same-gendered colleagues if no gender preferences exist, or even if opposite-gendered colleagues are prefered.
 For example, a sample of literature containing bioinformatics and cell biology papers will probably contain an excess of mostly-male and mostly-female author lists, simply because researchers preferentially work with colleagues from the same discipline, and because the author gender ratio is more male-biased in bioinformatics than in cell biology [@Holman_2018].
@@
 Here, coloured circles represent male and female authors, and coauthors are linked with lines.
-Across the whole set of ten papers, there is an apparent excess of same-gender collaborations: there are six same-gender papers and only four mixed-gender papers, which is fewer than the $10\times2\times0.5\times0.5 = 5$ mixed-gender papers we would expect under the null hypothesis that authors assort randomly with respect to gender.
-However, within each subset, there is no evidence that authors prefer to publish with same-gendered individuals.
+Across the whole set of ten papers, there is an apparent excess of same-gender collaborations: there are six same-gender papers and only four mixed-gender papers, which is fewer than the $10\times2\times0.5\times0.5 = 5$ mixed-gender papers expected under the null hypothesis that authors assort randomly with respect to gender.
+However, within each subset, there is no evidence that authors prefer to publish with same-gendered individuals (if anything, this small dataset suggests gender heterophily).
 The Wahlund effect will tend to inflate the frequency of same-sex coauthorships whenever the data is composed of two or more disconnected subsets of literature with different author gender ratios; these subsets could be research disciplines, older versus newer papers, or papers from authors in different countries.
@@
 
-In the present study, we test whether academics tend to co-publish with same-gendered collaborators, while controlling for the Wahlund effect as strictly as possible.
+In the present study, we test whether researchers working in the life sciences tend to co-publish with same-gendered collaborators, while controlling for the Wahlund effect as strictly as possible.
 Our study uses a recently-published dataset describing the gender of 35.5m authors from 9.15m articles indexed on PubMed [@Holman_2018].
@@
 [-@Holman_2018].
-Briefly, that dataset was created by downloading every article indexed on PubMed, from the oldest articles up to the present, and attempting to assign a gender to each author based on their given name.
+Briefly, that dataset was created by downloading every article indexed on PubMed and attempting to assign a gender to each author based on their given name.
 Each journal was assigned to one of 107 scientific disciplines, using PubMed's journal categorisations where available.
@@
 ## Calculating $\alpha$, the coefficient of homophily
-Inspired by Bergstrom et al.
-(http://www.eigenfactor.org/gender/assortativity/measuring_homophily.pdf), we defined the coefficient of homophily as $\alpha = p - q$, where $p$ is the probability that a randomly-chosen co-author of a _male_ author is a man and $q$ is the probability that a randomly-chosen co-author of a _female_ author is a man.
+Following Bergstrom et al.
+[-@bergstrom_2016], we defined the coefficient of homophily as $\alpha = p - q$, where $p$ is the probability that a randomly-chosen co-author of a _male_ author is a man and $q$ is the probability that a randomly-chosen co-author of a _female_ author is a man.
 Therefore, $\alpha > 0$ suggests that same-gender authors publish together more often than expected if authors assort randomly with respect to gender (homophily), $\alpha < 0$ suggests that opposite-gender authors publish together more often than expected (heterophily), and $\alpha = 0$ suggests random assortment with respect to gender.
+Like the Wahlund effect, $\alpha$ is borrowed from population genetics; for a set of 2-author papers, it is equivalent to the coefficient of inbreeding [@wright_1949].
+Mathematical work illustrates that $\alpha$ is closely related to network-based alternative methods for quantifying homophily [@wang_2016].
 
@@
 Previous authors [e.g.
-@Bonham_2017] have hypothesised that senior scientists preferentially recruit staff and students of the same gender, and/or that junior researchers preferentially select same-gendered mentors.
-In the majority of disciplines, authorship conventions mean that the first-listed author is often an early-career researcher, while the author listed last is more likely to be a senior researcher leading a research team [@Wren_2007].
+@sheltzer_2014; @Bonham_2017] have hypothesised that senior scientists preferentially recruit staff and students of the same gender, and/or that junior researchers preferentially select same-gendered mentors.
+In the majority of PubMed-indexed disciplines, authorship conventions mean that the first-listed author is often an early-career researcher, while the author listed last is more likely to be a senior researcher leading a research team [@Wren_2007].
 Assuming that senior researchers are the main drivers of homophily and that there are enough papers with three or more authors, we predict that the last author's gender will be the strongest predictor of the remaining authors' genders (i.e.
@@
 We hypothesised that this skew could potentially create both Wahlund effects and 'reverse' Wahlund effects.
-For example, imagine that the majority of collaborations are between students and professors, and that the gender ratio differs between career stages: we will then see an excess of mixed-gender coauthorships (heterophily, $\alpha < 0$), even if gender has no causal effect on the selection of coauthors.
+For example, imagine that the majority of collaborations are between students and professors, and that the gender ratio differs between career stages: we will then see an excess of mixed-gender coauthorships (heterophily, $\alpha < 0$), even if gender has no direct, causal effect on coauthor selection.
 Similarly, a hypothetical field in which students work only with students, and professors with professors, would have apparent gender homophily ($\alpha > 0$).
@@
 We can think of no tractable method of controlling for this issue using our dataset, which contains no information on career stage.
-Therefore, we instead decided to derive the theoretical expectations for $\alpha$ when there is a difference in gender ratio across career stages, in order to determine whether this effect could appreciably affect our conclusions.
-For simplicity, our calculations assume there are only two career stages, though the conclusions are likely to generalise to a realistic, multi-tiered career ladder.
+Therefore, we instead decided to derive the theoretical expectations for $\alpha$ when there is a difference in gender ratio across career stages, in order to determine if and how this effect should change our inferences.
+For simplicity, our calculations assume there are only two career stages, though the conclusions probably generalise to a realistic, multi-tier career ladder.
 Under the null model that gender has no causal effect on collaboration, we calculated $\alpha$ for various combinations of the four free parameters, i.e.
@@
 
-$\alpha'$ was significantly higher in the literature sample from 2015-16 relative to 2005-6, but the difference in means was small (Figure S3; Effect of the fixed factor 'Time period' in a linear mixed model of the data for all author positions: Cohen's $d$ = $0.09{\pm}0.04$, $t_{948}$ = 2.51, p = 0.012).
+$\alpha'$ was significantly higher in the literature sample from 2015-16 relative to 2005-6, though the difference in means was small (Figure S3; Effect of the fixed factor 'Time period' in a linear mixed model of the data for all author positions: Cohen's $d$ = $0.09{\pm}0.04$, $t_{948}$ = 2.51, p = 0.012).
 
@@
 ## Theoretical expectations for $\alpha$ when the gender ratio differs between career stages
-As shown in \autoref{simulation}, $\alpha$ is expected to be non-zero, even if collaborators are randomly selected with respect to gender, provided that there is a gender gap between career stages.
+As shown in \autoref{simulation}, we predict that $\alpha$ is expected to be non-zero, even if collaborators are randomly selected with respect to gender, provided that there is a gender gap between career stages.
 The extent to which $\alpha$ deviates from zero depends on the relative frequencies of collaboration within and between career stages.
@@
 We only expect strongly positive $\alpha$ when A) the gender ratio is highly skewed across career stages (e.g.
-a 10-fold difference), and B) collaborations between early and established researchers are very rare (e.g.
+a 5-fold difference), and B) collaborations between early and established researchers are very rare (e.g.
 <10% of the total).
-Both of these conditions are unlikely: the gender gap across careers stages is generally less than 10-fold [@Shaw_2012; @Holman_2018], and it is very common for early-career researchers to co-publish with an established mentor [@macaluso_2016].
+Both of these conditions are untrue for most fields: the gender gap across careers stages is generally less pronounced [e.g.
+@Shaw_2012; @Holman_2018], and it is very common for early-career researchers to co-publish with an established mentor [@macaluso_2016].
 However, one can get $\alpha > 0$ for realistic combinations of parameters, e.g.
@@
 
-Lastly, we note that if there is a gender gap between career stages and coauthorships between early-career and established researchers comprise ${\ge}50\%$ of the total, then the baseline expectation for $\alpha$ is actually less than zero (blue areas in \autoref{simulation}).
+Lastly, we note that if there is a gender gap between career stages and coauthorships between early-career and established researchers comprise >50% of the total, then the baseline expectation for $\alpha$ is actually less than zero (blue areas in \autoref{simulation}).
 Therefore, our results might under-estimate the extent to which researchers preferentially select same-gendered collaborators in some cases.
@@
 # Discussion
-We found evidence that researchers publish with same-gendered coauthors more often than expected under random assortment, even after implementing stringent controls for 'spurious' gender homophily resulting from Wahlund effects (\autoref{wahlund_plot}).
-Our study therefore reaffirms several earlier results and establishs their generality across the life sciences [@Ferber_1980; @McDowell_1992; @Bentley_2003; @Crow_2015; @Ghiasi_2015; @Araujo_2017a; @Fahmy_2017; @Jadidi_2017; @Teele_2017; @Zettler_2017].
+We found evidence that researchers publish with same-gendered coauthors more often than expected under random assortment, even after implementing stringent controls for Wahlund effects (\autoref{wahlund_plot}).
+Our study therefore reaffirms earlier studies' conclusions [e.g.
+@Ferber_1980; @McDowell_1992; @Bentley_2003; @Crow_2015; @Ghiasi_2015; @Araujo_2017a; @Fahmy_2017; @Jadidi_2017; @Teele_2017; @Zettler_2017] and establishes their generality across the life sciences.
 Relatively few journals had $\alpha'$ values below zero, and almost no journals showed statistically significant gender heterophily after controlling for multiple testing.
@@
 As well as conscious or unconcious selection of collaborators based on gender, our results could be partly or wholly explained by uncontrolled Wahlund effects.
-We suspect the contribution of these to be minor, for four reasons: we found positive $\alpha'$ after controlling for three obvious sources of Wahlund effect; there was no inflation of $\alpha'$ in highly multidisciplinary journals; restricting the data by country yielded similar estimates of $\alpha'$; and we showed that differences in gender ratio between career stages are unlikely to fully explain our results.
-On balance, we believe the data suggest that researchers preferentially select same-gendered collaborators, the magnitude of this effect is uncertain.
+However, we suspect the contribution of these to be minor, for four reasons: we found positive $\alpha'$ after controlling for three obvious sources of Wahlund effect; there was no inflation of $\alpha'$ in highly multidisciplinary journals; restricting the data by country yielded similar estimates of $\alpha'$; and we showed that differences in gender ratio between career stages are unlikely to fully explain our results.
+On balance, we believe the data provide good evidence that researchers preferentially select same-gendered collaborators, although the effect size of this preference is uncertain.
 
@@
 One interpretation of this result is that men are more likely to preferentially seek out male collaborators in fields where men are a minority, relative to the homophily displayed by women in fields where women are a minority.
-However, this latter result is only tentatively supported since there are so few journals on PubMed in which most authors are women \autoref{alpha_gender_ratio}.
+However, this latter result is only tentatively supported since there are so few journals in which most authors are women \autoref{alpha_gender_ratio}.
 
@@
 These questions are closely related.
-For example, some of the homophily we observed might be caused by women seeking to avoid harassment or bullying from men [@jagsi_2016], which is clearly very concerning.
-Additionally, Sheltzer and Smith [-@sheltzer_2014] concluded that 'elite' male academics (i.e.
-recipients of major honours) have a higher proportion of male students and postdocs than non-elite male academics.
-This finding could contribute to the homophily we observed, and is cause for concern because the results might reflect discrimination against women by established scientists [@Moss_2012], or avoidance by women of elite research groups (e.g.
-due to 'imposter syndrome' or the perception that these groups are hostile to women).
-We also found a little evidence that gender homophily is deterimental to research quality, in that high-impact journals tended to have weaker homophily.
+For example, some of the homophily we observed might be caused by women seeking to avoid harassment, sexism or microaggressions from men [@jagsi_2016], which is clearly very concerning.
+Additionally, Sheltzer and Smith [-@sheltzer_2014] concluded that 'elite' male academics (defined as recipients of major honours) have a higher proportion of male students and postdocs than non-elite male academics.
+This finding could contribute to the homophily we observed, and is cause for concern since Sheltzer and Smith [-@sheltzer_2014]'s results might reflect discrimination against women [@Moss_2012], or avoidance by women of elite research groups (e.g.
+due to 'imposter syndrome' or the perception that these groups are sexist).
+We also found a little evidence that gender homophily is detrimental to research quality, in that high-impact journals tended to have weaker homophily.
 Assuming that papers published in high-impact journals are of higher average quality [which is contentious; @Garfield_2006], then this result provides non-experimental support for the hypothesis that mixed-gender teams produce better research than single-gender teams [@Campbell_2013].
@@
 Homophily might also have more benign causes.
-Collaboration is often especially enjoyable when working with like-minded people, who might be same-gendered more often than not.
-We also suppose that some people -- women especially -- consciously choose to preferentially collaborate with women whenever multiple qualified candidates are available, in order to help close the gender gap in the workforce.
+Collaboration is often most enjoyable and productive when working with like-minded people, who might be same-gendered more often than not.
+We also suppose that some people consciously choose to preferentially collaborate with women in order to help close the gender gap in the workforce; this would create homophily if this motivation applies to more women than men.
 Related to this, Ghiasi et al.
-[-@Ghiasi_2015] wrote that women who mostly work with men are 'compliant [in reproducing] male-dominated scientific structures', although we think this charge is unfair.
-We believe that everyone (not only women) should take steps to ensure that they are not overlooking or excluding female colleagues, particularly since men are more likely to occupy stable, senior jobs that afford them the luxury of choice.
+[-@Ghiasi_2015] concluded that women in engineering are 'compliant [in reproducing] male-dominated scientific structures' because they do not collaborate often enough with other women (coauthorships involving two women were about 30% more common than expected under the null).
+In contrast, we believe that everyone (not just women) should work to ensure that they are not overlooking or excluding female colleagues, or devoting less effort to collaborations with them.
+Men are more often well placed to help in this regard, given their overrepresentation in jobs that allow the hiring of staff.
 

### 2018-05-29 Gender_homophily_analysis.Rmd
Finalise markdown

--- 2018-05-27 Gender homophilly project.Rmd
@@
 ---
-title: "Do researchers preferentially collaborate with same-gendered colleagues?"
+title: "Do researchers preferentially collaborate with colleagues of the same gender?"
 subtitle: Online Supplementary Material
@@
 library(pander) # nice Markdown tables
-library(kableExtra) # scrollable html tables
 library(tidyr) # spread and gather functions
@@
 # Using a database allows quick access to the whole PubMed dataset without loading the whole thing to memory.
+# You can download the pubmed_gender_data.sqlite3 database here: https://osf.io/bt9ya/
 pubmed_db <- src_sqlite("~/Desktop/Gender Pubmed mining project/data for analysis/pubmed_gender_data.sqlite3", create = F)
@@
 ```{r echo=FALSE, results='hide'}
-library(knitr) # set up cacheing to save time when re-building the html document - chunks are not re-evaluated if the code has not changed
+library(knitr) 
 opts_chunk$set(warning=FALSE, message=FALSE) # Suppress warnings in the HTML output
@@
 ```
-`r opts_chunk$set(cache=TRUE)`
 
@@
   # Only do the test if over.write = TRUE or there is no datafile in the working directory
-  if(over.write | !(file.name %in% list.files())){
+  if(over.write | !file.exists(file.name)){
 
@@
 
-### Define a function to calculate Bergstrom et al.'s 'coefficient of homophily', $\alpha$
+### Define a function to calculate Bergstrom et al.'s "coefficient of homophily", $\alpha$
 These two functions estimate $\alpha$ for a set of papers' author lists.
@@
   # Only do the test if over.write = TRUE or there is no datafile in the working directory
-  if(over.write | !(output.file %in% list.files())){ 
+  if(over.write | !file.exists(output.file)){ 
     journals <- names(count.data)
@@
                          end.of.time.period = 0, 
-                         file.name = "last.year.rds", 
+                         file.name = "data/last.year.rds", 
                          over.write = FALSE)
@@
                          end.of.time.period = 10, 
-                         file.name = "older.papers.rds", 
+                         file.name = "data/older.papers.rds", 
                          over.write = FALSE)
@@
 
-  count.data <- readRDS(paste(filename, ".rds", sep = "")) 
+  count.data <- readRDS(paste("data/", filename, ".rds", sep = "")) 
   data.frame(Quantity = c("Number of disciplines", 
@@
 
+if(!file.exists("manuscript/sample.size.table.csv")){
   sample.size.table <- make.sample.size.table("older.papers") %>% 
@@
   write.csv(sample.size.table, file = "manuscript/sample.size.table.csv", row.names = FALSE)
-
-sample.size.table %>% pander(split.cell = 40, split.table = Inf)
+}
+
+read.csv("manuscript/sample.size.table.csv") %>% pander(split.cell = 40, split.table = Inf)
 ```
@@
 # Calculate alpha across all authorship positions for all suitable NEW papers
-gender.grouping.test.many.journals(readRDS("last.year.rds"),
+gender.grouping.test.many.journals(readRDS("data/last.year.rds"),
                                    "all",
-                                   "homophily.results.all.csv",
+                                   "data/homophily.results.all.csv",
                                    nBoots = 1000, 
@@
 # Calculate alpha for first authors for all suitable NEW papers
-gender.grouping.test.many.journals(readRDS("last.year.rds"),
+gender.grouping.test.many.journals(readRDS("data/last.year.rds"),
                                    "first",
-                                   "homophily.results.first.authors.csv",
+                                   "data/homophily.results.first.authors.csv",
                                    nBoots = 1000, 
@@
 # Calculate alpha for last authors for all suitable NEW papers
-gender.grouping.test.many.journals(readRDS("last.year.rds"),
+gender.grouping.test.many.journals(readRDS("data/last.year.rds"),
                                    "last",
-                                   "homophily.results.last.authors.csv",
+                                   "data/homophily.results.last.authors.csv",
                                    nBoots = 1000, 
@@
 # Calculate alpha across all authorship positions for all suitable OLD papers
-gender.grouping.test.many.journals(readRDS("older.papers.rds"),
+gender.grouping.test.many.journals(readRDS("data/older.papers.rds"),
                                    "all",
-                                   "older.homophily.results.all.csv",
+                                   "data/older.homophily.results.all.csv",
                                    nBoots = 1000, 
@@
 # Calculate alpha for first authors for all suitable OLD papers
-gender.grouping.test.many.journals(readRDS("older.papers.rds"),
+gender.grouping.test.many.journals(readRDS("data/older.papers.rds"),
                                    "first",
-                                   "older.homophily.results.first.authors.csv",
+                                   "data/older.homophily.results.first.authors.csv",
                                    nBoots = 1000, 
@@
 # Calculate alpha for last authors for all suitable OLD papers
-gender.grouping.test.many.journals(readRDS("older.papers.rds"),
+gender.grouping.test.many.journals(readRDS("data/older.papers.rds"),
                                    "last",
-                                   "older.homophily.results.last.authors.csv",
+                                   "data/older.homophily.results.last.authors.csv",
                                    nBoots = 1000, 
@@
 # Load up all the homophily results, and merge them with the journal-specific information
-gender.homophily.all <- merge.and.clean("homophily.results.all.csv", journal.data)
-gender.homophily.first <- merge.and.clean("homophily.results.first.authors.csv", journal.data)
-gender.homophily.last <- merge.and.clean("homophily.results.last.authors.csv", journal.data)
-
-gender.homophily.all.old <- merge.and.clean("older.homophily.results.all.csv", journal.data)
-gender.homophily.first.old <- merge.and.clean("older.homophily.results.first.authors.csv", journal.data)
-gender.homophily.last.old <- merge.and.clean("older.homophily.results.last.authors.csv", journal.data)
+gender.homophily.all <- merge.and.clean("data/homophily.results.all.csv", journal.data)
+gender.homophily.first <- merge.and.clean("data/homophily.results.first.authors.csv", journal.data)
+gender.homophily.last <- merge.and.clean("data/homophily.results.last.authors.csv", journal.data)
+
+gender.homophily.all.old <- merge.and.clean("data/older.homophily.results.all.csv", journal.data)
+gender.homophily.first.old <- merge.and.clean("data/older.homophily.results.first.authors.csv", journal.data)
+gender.homophily.last.old <- merge.and.clean("data/older.homophily.results.last.authors.csv", journal.data)
 
@@
   rename(unadjusted.p.value = two.tail.p.value)
-
-# Save and export a neat and tidy version for the supplementary material
-neat <- master.dataset
-names(neat) <- gsub("[.]", " ", names(neat))
-names(neat) <- gsub("Subset", "Author position", names(neat))
-names(neat) <- paste(toupper(substring(names(neat), 1, 1)), substring(names(neat), 2), sep="")
-names(neat) <- gsub("N", "n", names(neat))
-names(neat) <- gsub("2", "", names(neat))
-for(i in which(sapply(neat, is.numeric))) neat[,i] <- round(neat[,i], 3)
-write.csv(neat, file = "Supplementary dataset S1.csv", row.names = FALSE)
-rm(neat)
 ```
@@
 
-if(!("country.data.csv" %in% list.files())){ # Write the data, assuming it's not already been done
-  country.data <- gender.grouping.test.by.journal.and.country(country.data)
-  write.csv(country.data, file = "country.data.csv", row.names = FALSE)
-}
-
-country.data <- read.csv("country.data.csv", stringsAsFactors = FALSE) %>%
+if(!file.exists("data/country.data.csv")){ # Write the data, assuming it's not already been done
+  country.data <- gender.grouping.test.by.journal.and.country()
+  write.csv(country.data, file = "data/country.data.csv", row.names = FALSE)
+}
+
+country.data <- read.csv("data/country.data.csv", stringsAsFactors = FALSE) %>%
   mutate(p.adjusted = p.adjust(two.tail.p.value, method = "BH")) %>%
-  left_join(journal.data %>% select(journal, discipline), by = "journal") # %>% 
-  # left_join(master.dataset %>% 
-  #             filter(Age == "New" & Subset == "All authors") %>%
-  #             select(journal, alpha.adjusted) %>%
-  #             rename(journal.alpha = alpha.adjusted),
-  #           by = "journal")
+  left_join(journal.data %>% select(journal, discipline), by = "journal") 
 ```
@@
 ## Results
+
+### Create S1 Data, which holds all the $\alpha'$ estimates
+
+**S1 Data**: This spreadsheet shows the $\alpha'$ values calculated for each journal, in the 2005 and 2015 samples, and for each type of author (all authors, first authors, and last authors).
+The tables gives the impact factor of each journal, the sample size, $\alpha$ and $\alpha'$ and their 95% CIs, and the p-value from a 2-tailed test evaluating the null hypothesis that $\alpha$ is zero (both raw and FDR-corrected p-values are shown).
+
+```{r}
+# Save and export a neat and tidy version for the supplementary material
+neat <- master.dataset
+names(neat) <- gsub("[.]", " ", names(neat))
+names(neat) <- gsub("Subset", "Author position", names(neat))
+names(neat) <- paste(toupper(substring(names(neat), 1, 1)), substring(names(neat), 2), sep="")
+names(neat) <- gsub("N", "n", names(neat))
+names(neat) <- gsub("2", "", names(neat))
+for(i in which(sapply(neat, is.numeric))) neat[,i] <- round(neat[,i], 3)
+write.csv(neat, file = "data/S1 data.csv", row.names = FALSE)
+rm(neat)
+```
 
@@
 
-**Table S`r supp.table.counter`**: The table shows similar information to Figure S`r supp.figure.counter`, namely the distribution of the $\alpha'$ values across journals, split by the research discipline.
+**Table S`r supp.table.counter`**: The table shows similar information to Figure S`r figure.counter`, namely the distribution of the $\alpha'$ values across journals, split by the research discipline.
 The gender ratio column shows the percentage of women authors in the sample used to calculate $\alpha'$, across all authorship positions.
@@
 mean.inflation.due.to.country <- country.data %>%
+  left_join(master.dataset %>% 
+              filter(Age == "New", Subset == "All authors") %>% 
+              select(journal, alpha.adjusted) %>% 
+              rename(journal.alpha = alpha.adjusted), 
+            by = "journal") %>%
   group_by(journal) %>%
@@
 ggsave(simulation.figure, file = "figures/figure5.pdf", 
-       height = 5.6, width = 5)#, device = cairo_pdf)
+       height = 5.6, width = 5, device = cairo_pdf)
 ```

### 2018-05-29 manuscript/Manuscript_Claire.Rmd
Finalise markdown

@@
 author:
-- Luke Holman* and Claire Morandin&sect;
-- '*\textit{luke.holman@unimelb.edu.au} \vspace{5mm}'
+- Luke Holman* and Claire Morandin\textsuperscript{§}
+- '*\textit{luke.holman@unimelb.edu.au}'
+- \textsuperscript{§}\textit{claire.morandin@helsinki.fi} \vspace{5mm}
+bibliography: references.bib
 output:
@@
   html_document: default
-bibliography: references.bib
+  word_document: default
 fontsize: 12pt
-header-includes: 
-  \usepackage{microtype} 
-  \usepackage[all]{hypcap} 
-  \usepackage{amsmath} 
-  \usepackage{booktabs}
-  \usepackage{caption} 
-  \usepackage[labelfont=bf]{caption} 
-  \usepackage{titling} 
-  \pretitle{\begin{flushleft}\LARGE}
-  \posttitle{\par\end{flushleft}\vskip 0.5em} 
-  \preauthor{\begin{flushleft}\large \lineskip 0.5em} 
-  \postauthor{\par\end{flushleft}} 
-  \predate{\begin{flushleft}\large} 
-  \postdate{\par\end{flushleft}}
-  \usepackage{fancyhdr} 
-  \pagestyle{fancy} 
-  \fancyhead[LO,LE]{\textsl{\leftmark}} 
-  \rhead[]{Gender and coauthorship} 
-  \usepackage[noabbrev,capitalise]{cleveref}
-  \usepackage{titlefoot} 
-  \usepackage{amssymb} 
-  \usepackage{rotating}
+header-includes: \usepackage{microtype} \usepackage[all]{hypcap} \usepackage{amsmath}
+  \usepackage{booktabs} \usepackage{caption} \usepackage[labelfont=bf]{caption} \usepackage{titling}
+  \pretitle{\begin{flushleft}\LARGE} \posttitle{\par\end{flushleft}\vskip 0.5em} \preauthor{\begin{flushleft}\large
+  \lineskip 0.5em} \postauthor{\par\end{flushleft}} \predate{\begin{flushleft}\large}
+  \postdate{\par\end{flushleft}} \usepackage{fancyhdr} \pagestyle{fancy} \fancyhead[LO,LE]{\textsl{\leftmark}}
+  \rhead[]{Gender and coauthorship} \usepackage[noabbrev,capitalise]{cleveref} \usepackage{titlefoot}
+  \usepackage{amssymb} \usepackage{rotating}
 csl: the-american-naturalist.csl
-abstract: Evidence suggests that women in academia are hindered by conscious and unconscious biases, and often feel excluded from formal and informal opportunities for research collaboration.
-In addition to ensuring fairness and helping to redress gender imbalance in the academic workforce, increasing women's access to collaboration could help scientific progress by drawing on more of the available human capital.
-Here, we test whether researchers preferentially collaborate with same-gendered colleagues, using more stringent methods and a larger dataset than in past work.
-Our results reaffirm that researchers co-publish with colleagues of the same gender, and show that this 'gender homophily' is slightly stronger today than it was 10 years ago.
-Contrary to our expectations, we found no evidence that homophily is driven mostly by senior academics, and no evidence that homophily is strongest in fields where women are in the minority.
-Interestingly, homophily was negatively correlated with journal impact factor, as predicted if mixed-gender teams produce better research.
-We discuss potential causes of gender homophily in science.
-\vspace{5mm} \par\noindent \textbf{Keywords:} Gender bias, Homophily, Scientific collaboration, Text mining, Women in STEM.
+abstract: Evidence suggests that women in academia are hindered by conscious and unconscious
+  biases, and often feel excluded from formal and informal opportunities for research
+  collaboration. In addition to ensuring fairness and helping to redress gender imbalance
+  in the academic workforce, increasing women's access to collaboration could help
+  scientific progress by drawing on more of the available human capital. Here, we
+  test whether researchers preferentially collaborate with same-gendered colleagues,
+  using more stringent methods and a larger dataset than in past work. Our results
+  reaffirm that researchers co-publish with colleagues of the same gender, and show
+  that this 'gender homophily' is slightly stronger today than it was 10 years ago.
+  Contrary to our expectations, we found no evidence that homophily is driven mostly
+  by senior academics, and no evidence that homophily is strongest in fields where
+  women are in the minority. Interestingly, homophily was negatively correlated with
+  journal impact factor, as predicted if mixed-gender teams produce better research.
+  We discuss potential causes of gender homophily in science. \vspace{5mm} \par\noindent
+  \textbf{Keywords:} Gender bias, Homophily, Scientific collaboration, Text mining,
+  Women in STEM. 
 ---
@@
 \maketitle\unmarkedfntext{*School of BioSciences, The University of Melbourne, Victoria, Australia.}
-\maketitle\unmarkedfntext{§Organismal and Evolutionary Biology Research Programme,
+\maketitle\unmarkedfntext{\textsuperscript{§}Organismal and Evolutionary Biology Research Programme,
 Faculty of Biological and Environmental Sciences, University of Helsinki, Finland.}
@@
 
-Women are substantially underrepresented in many branches of the workforce in science, technology, engineering, mathematics, and medicine (STEMM), and face additional challenges and inequities relative to men [e.g.
+Women are highly underrepresented in many branches of science, technology, engineering, mathematics, and medicine (STEMM), and face additional challenges and inequities relative to men [e.g.
 @Shaw_2012; @Lariviere_2013; @West_2013; @Elsevier_report; @Holman_2018].
@@
 [-@Holman_2018].
-Briefly, that dataset was created by downloading every article indexed on PubMed and attempting to assign a gender to each author based on their given name.
-Each journal was assigned to one of 107 scientific disciplines, using PubMed's journal categorisations where available.
+Briefly, that dataset was created by downloading every article indexed on PubMed and attempting to infer each author's gender from their given name.
+Each journal was assigned to one of 107 scientific disciplines, using PubMed's journal categorisations in the interests of objectivity.
 Because the present study focuses on co-authorship, all single-author papers were discarded.
@@
 [-@bergstrom_2016], we defined the coefficient of homophily as $\alpha = p - q$, where $p$ is the probability that a randomly-chosen co-author of a _male_ author is a man and $q$ is the probability that a randomly-chosen co-author of a _female_ author is a man.
-Therefore, $\alpha > 0$ suggests that same-gender authors publish together more often than expected if authors assort randomly with respect to gender (homophily), $\alpha < 0$ suggests that opposite-gender authors publish together more often than expected (heterophily), and $\alpha = 0$ suggests random assortment with respect to gender.
-Like the Wahlund effect, $\alpha$ is borrowed from population genetics; for a set of 2-author papers, it is equivalent to the coefficient of inbreeding [@wright_1949].
+Therefore, $\alpha > 0$ suggests that same-gender authors publish together more often than expected under the null hypothesis that authors assort randomly with respect to gender (homophily), $\alpha < 0$ suggests that opposite-gender authors publish together more often than expected (heterophily), and $\alpha = 0$ suggests random assortment with respect to gender.
+Like the Wahlund effect, $\alpha$ is borrowed from population genetics; for a set of 2-author papers, it is equivalent to Wright's coefficient of inbreeding [@wright_1949].
 Mathematical work illustrates that $\alpha$ is closely related to network-based alternative methods for quantifying homophily [@wang_2016].
@@
 
-Our test assumes that the expected value of $\alpha$ is zero if authors randomly assort, but for small datasets this assumption is not always true (as pointed out by Carl T.
-Bergstrom in a blog post, http://www.eigenfactor.org/gender/assortativity/note_to_eisen.rtf).
+Our test assumes that the expected value of $\alpha$ is zero if authors randomly assort, but for small datasets this assumption is not always true (as pointed out by Carl Bergstrom in a blog post, http://www.eigenfactor.org/gender/assortativity/note_to_eisen.rtf).
 To borrow Prof.
@@
 For these six papers, $\alpha = -\frac{1}{3}$, even though same- and opposite-gendered coauthors were selected in equal proportion to their frequency in the pool of possible collaborators.
+
 To control for the fact that the null expectation for $\alpha$ is not zero for small datasets, we devised an adjusted version of the coefficient of homophily, which we term $\alpha'$.
@@
 To minimise bias in $\alpha'$ due to the Wahlund effect, we restricted each set of papers to a single research specialty to the greatest extent allowed by our data.
-Specifically, we only calculated $\alpha'$ for individual journals, since papers from the same journal typically focus on one or a few closely related topics.
+Specifically, we only calculated $\alpha'$ for individual journals, since papers from the same journal typically focus on closely related topics.
 Although some journals, e.g.
@@
 The difference in $\alpha'$ between highly multidisciplinary and more specialised journals, e.g.
-_PLoS ONE_ versus _PLoS Computational Biology_, gives a rough estimate of the extent to which multidisciplinarity inflates $\alpha'$.
+_PLoS ONE_ versus _PLoS Computational Biology_, gives an estimate of the extent to which multidisciplinarity inflates $\alpha'$.
 
@@
 Therefore, a dataset containing a mix of papers from teams of authors based in these two countries would contain an excess of same-sex coauthorships, even if collaboration were random with respect to gender within each country.
-
 To address this issue, we also analysed every combination of journal and author country of affiliation for which we had enough data (i.e.
@@
 We hypothesised that this skew could potentially create both Wahlund effects and 'reverse' Wahlund effects.
-For example, imagine that the majority of collaborations are between students and professors, and that the gender ratio differs between career stages: we will then see an excess of mixed-gender coauthorships (heterophily, $\alpha < 0$), even if gender has no direct, causal effect on coauthor selection.
+For example, imagine that the majority of collaborations are between students and professors, and that the gender ratio differs between career stages: we will then see an excess of mixed-gender coauthorships (heterophily, $\alpha < 0$), even if gender has no direct, causal effect.
 Similarly, a hypothetical field in which students work only with students, and professors with professors, would have apparent gender homophily ($\alpha > 0$).
@@
 We can think of no tractable method of controlling for this issue using our dataset, which contains no information on career stage.
-Therefore, we instead decided to derive the theoretical expectations for $\alpha$ when there is a difference in gender ratio across career stages, in order to determine if and how this effect should change our inferences.
-For simplicity, our calculations assume there are only two career stages, though the conclusions probably generalise to a realistic, multi-tier career ladder.
+Therefore, we instead decided to derive the theoretical expectations for $\alpha$ when there is a difference in gender ratio across career stages, in order to determine if and how this effect should alter our inferences.
+For simplicity, our calculations assume there are only two career stages, though we intuit that the general conclusions would also apply to a multi-tier career ladder.
 Under the null model that gender has no causal effect on collaboration, we calculated $\alpha$ for various combinations of the four free parameters, i.e.
@@
 ## Data availability and reproducibility
-The Online Supplementary Material contains R scripts used to produce all data, figures and tables. The input data from Holman et al. [-@Holman_2018] is archived at https://osf.io/bt9ya/.
+The Online Supplementary Material contains R scripts used to produce all results, figures and tables.
+The input data from Holman et al.
+[-@Holman_2018] is archived at https://osf.io/bt9ya/ along with the code used to obtain it.
 
@@
 
-$\alpha'$ was significantly higher in the literature sample from 2015-16 relative to 2005-6, though the difference in means was small (Figure S3; Effect of the fixed factor 'Time period' in a linear mixed model of the data for all author positions: Cohen's $d$ = $0.09{\pm}0.04$, $t_{948}$ = 2.51, p = 0.012).
+$\alpha'$ was significantly higher in the literature sample from 2015-16 relative to 2005-6, though the difference in means was small (Figure S3; Effect of the fixed factor 'Time period' in a linear mixed model of the data for all author positions: Cohen's $d$ = $0.091{\pm}0.04$, $t_{953}$ = 2.42, p = 0.016).
 
-When comparing pairs of $\alpha'$ values estimated for the first and last authors for the same journals, we found that $\alpha'$ tended to be higher for first authors than for last authors (Figure S4; Effect of the fixed factor 'Authorship position' in a linear mixed model: Cohen's $d$ = $0.07{\pm}0.02$, $t_{1988}$ = 4.48, p < 0.0001).
+When comparing pairs of $\alpha'$ values estimated for the first and last authors for the same journals, we found that $\alpha'$ tended to be higher for first authors than for last authors (Figure S4; Effect of the fixed factor 'Authorship position' in a linear mixed model: Cohen's $d$ = $0.065{\pm}0.02$, $t_{2024}$ = 4.28, p < 0.0001).
 This suggests that the gender of the first author was a slightly stronger predictor of the remaining authors' genders than the gender of the last author, i.e.
@@
 All disciplines had a positive average $\alpha'$, although homophily appeared somewhat stronger in some disciplines than others (e.g.
-mean $\alpha'$ was $0.12{\pm}0.02$ for Urology journals, and $0.03{\pm}0.01$ for Veterinary Medicine journals; \autoref{alpha_histograms}, Table S3).
-However, there was no evidence for consistent differences in $\alpha'$ between disciplines: the random factor 'Discipline' explained at most 1% of the variance in $\alpha'$ in the two linear mixed models described in the previous section (see \autoref{alpha_histograms} and mixed models in Online Supplementary Material).
-This implies that the processes causing positive $\alpha'$ are similarly strong in all the disciplines we examined.
+mean $\alpha'$ was $0.12{\pm}0.02$ for Urology journals and $0.03{\pm}0.01$ for Veterinary Medicine journals; \autoref{alpha_histograms}, Table S3).
+However, there was no evidence for consistent differences in $\alpha'$ between disciplines: the random factor 'Discipline' explained around 1% of the variance in $\alpha'$ in the two linear mixed models described in the previous section (see \autoref{alpha_histograms} and mixed models in Online Supplementary Material).
+Thus, we cannot reject the null hypothesis that the processes causing positive $\alpha'$ are similarly strong in all the disciplines we examined.
 
 There was no indication that journals publishing on a wide range of topics have higher $\alpha'$ values than more specialised journals, due to the Wahlund effect.
-For example, the journal category 'Multidisciplinary' - which includes journals like _PLoS ONE_, _Nature_, _Science_, and _PNAS_ - did not have notably elevated $\alpha'$ (\autoref{alpha_histograms}).
-This result suggests that our estimates of homophily, and estimates from earlier studies, are not notably inflated by the presence of disparate research topics (with variable author gender ratios) being published within individual journals.
+For example, the journal category 'Multidisciplinary' -- which includes journals like _PLoS ONE_, _Nature_, _Science_, and _PNAS_ -- did not have notably elevated $\alpha'$ (\autoref{alpha_histograms}).
+This result suggests that our estimates of homophily, and estimates from some earlier studies, are not notably inflated by the presence of disparate research topics (with variable author gender ratios) being published within individual journals.
 
@@
 ## Analysis accounting for differences in author gender ratio between countries
-When we restricted the analysis to only include authors with affiliations from a single country, we observed statistically significant homophily for 72 of the 325 journal-country combinations tested (64 unique journals and 18 unique countries), and no significant evidence of heterophily (Figures S5-S6).
+When we restricted the analysis by country, we observed statistically significant homophily for 72 of the 325 journal-country combinations tested (64 unique journals and 18 unique countries), and no significant evidence of heterophily (Figures S5-S6).
 Additionally, the values of $\alpha'$ calculated for each journal-country combination were mostly very similar to the $\alpha'$ values calculated for the journal as a whole (i.e.
-when pooling papers from different countries); the average difference in $\alpha'$ was 0.002 (Figure S7).
-These results suggest that our findings of widespread homophily in the main analysis were not driven by a Wahlund effect resulting from gender differences between countries.
+when pooling papers from different countries); on average, $\alpha'$ was only inflated in the main (non-country specific) analysis by 0.002 (Figure S7).
+These results suggest that our findings of widespread homophily in the main analysis were not driven solely by a Wahlund effect resulting from gender differences between countries.
 
@@
 Contour lines denote increments of 0.1.
-\label{simulation}](../figures/figure5.pdf)
+\label{simulation}](../figures/figure5 - Simulation.pdf)
 
@@
 # Discussion
-We found evidence that researchers publish with same-gendered coauthors more often than expected under random assortment, even after implementing stringent controls for Wahlund effects (\autoref{wahlund_plot}).
+We found evidence that researchers preferentially publish with same-gendered coauthors, even after implementing stringent controls for Wahlund effects (\autoref{wahlund_plot}).
 Our study therefore reaffirms earlier studies' conclusions [e.g.
@@
 Relatively few journals had $\alpha'$ values below zero, and almost no journals showed statistically significant gender heterophily after controlling for multiple testing.
-The excess of same-gender coauthorships was quite strong: many journals had $\alpha' > 0.1$, indicating that the gender ratio of men's and women's coauthors differs by >10% in absolute terms.
-In relative terms, this effect appears even more striking: for example, if the proportion of male coauthors is 80% for men and 70% for women ($\alpha' = 0.1$), the results imply that women publish with women 50% more often than men do.
+The excess of same-gender coauthorships was quite large: many journals had $\alpha' > 0.1$, indicating that the gender ratio of men's and women's coauthors differs by >10% in absolute terms.
+In relative terms, the observed homophily is even more striking: for example, if the proportion of male coauthors is 80% for men and 70% for women ($\alpha' = 0.1$), the results imply that women publish with women 50% more frequently than men do.
 
@@
 because being in the minority might increase one's drive to seek out same-gendered colleagues.
-Contrary to this hypothesis, we found no evidence that gender homophily is restricted to particular disciplines: $\alpha'$ was similarly high across the board \autoref{alpha_histograms}.
+Contrary to this hypothesis, we found no evidence that gender homophily is restricted to particular disciplines: $\alpha'$ was similarly high across the board (\autoref{alpha_histograms}).
 Moreover, $\alpha'$ was lower for journals with a male-biased author gender ratio relative to those with an even gender ratio.
@@
 One interpretation of this result is that men are more likely to preferentially seek out male collaborators in fields where men are a minority, relative to the homophily displayed by women in fields where women are a minority.
-However, this latter result is only tentatively supported since there are so few journals in which most authors are women \autoref{alpha_gender_ratio}.
+However, this latter result is only tentatively supported since there are so few journals in which most authors are women (see \autoref{alpha_gender_ratio}).
 
-We also found that gender homophily was marginally stronger in 2015-2016, relative to 2005-2006.
-Although this trend might reflect a change in the gender preferences of researchers seeking collaborators over time, there are alternative (and probably more likely) explanations for this result.
+We also found that gender homophily was marginally stronger in 2015-2016 relative to 2005-2006.
+Although this trend might reflect a change in the gender preferences of researchers seeking collaborators over time, there are alternative (and perhaps more likely) explanations.
 For example, this trend might result from the increasing number of women working in senior positions in STEMM over the past decade [e.g.
@@
 These questions are closely related.
-For example, some of the homophily we observed might be caused by women seeking to avoid harassment, sexism or microaggressions from men [@jagsi_2016], which is clearly very concerning.
+For example, some of the homophily we observed might be caused by women seeking to avoid harassment, sexism or microaggressions from men [@jagsi_2016], which would clearly be concerning.
 Additionally, Sheltzer and Smith [-@sheltzer_2014] concluded that 'elite' male academics (defined as recipients of major honours) have a higher proportion of male students and postdocs than non-elite male academics.
-This finding could contribute to the homophily we observed, and is cause for concern since Sheltzer and Smith [-@sheltzer_2014]'s results might reflect discrimination against women [@Moss_2012], or avoidance by women of elite research groups (e.g.
+This finding could contribute to the homophily we observed, and is cause for concern since Sheltzer and Smith [-@sheltzer_2014]'s results might reflect discrimination against women during hiring [@Moss_2012], or avoidance by women of elite research groups (e.g.
 due to 'imposter syndrome' or the perception that these groups are sexist).
 We also found a little evidence that gender homophily is detrimental to research quality, in that high-impact journals tended to have weaker homophily.
-Assuming that papers published in high-impact journals are of higher average quality [which is contentious; @Garfield_2006], then this result provides non-experimental support for the hypothesis that mixed-gender teams produce better research than single-gender teams [@Campbell_2013].
+Assuming that papers published in high-impact journals are of higher average quality [which is contentious; @Garfield_2006], this result provides non-experimental support for the hypothesis that mixed-gender teams produce better research than single-gender teams [@Campbell_2013].
 
@@
 Collaboration is often most enjoyable and productive when working with like-minded people, who might be same-gendered more often than not.
-We also suppose that some people consciously choose to preferentially collaborate with women in order to help close the gender gap in the workforce; this would create homophily if this motivation applies to more women than men.
+We also suppose that some people consciously choose to preferentially collaborate with women in order to help close the gender gap in the workforce; this would create homophily if women do this more than men.
+In support of this interpretation, women appear more likely than men to promote the work of female collagues by inviting them to give talks [@nittrouer_2018; @debarre_2018].
 Related to this, Ghiasi et al.
-[-@Ghiasi_2015] concluded that women in engineering are 'compliant [in reproducing] male-dominated scientific structures' because they do not collaborate often enough with other women (coauthorships involving two women were about 30% more common than expected under the null).
-In contrast, we believe that everyone (not just women) should work to ensure that they are not overlooking or excluding female colleagues, or devoting less effort to collaborations with them.
-Men are more often well placed to help in this regard, given their overrepresentation in jobs that allow the hiring of staff.
+[-@Ghiasi_2015] concluded that women in engineering are "compliant [in reproducing] male-dominated scientific structures" because they do not collaborate often enough with other women; their data suggest that coauthorships between two women are about 30% more frequent than expected under the null (see their Figure 7).
+In contrast, we propose that it behoves men as well as women to ensure that they are not overlooking or excluding female colleagues.
+Given that women work disproportionately often in junior positions with little opportunity to be selective, one could argue that the responsibilty to address gender imblances in science rests primarily with men.
 
@@
 
-Although the proportion of women in science has significantly increased over the last decades, women still tend to leave their academic careers behind considerably more often than their male counterparts [@Debarre_2018].
-Lack of scientific recognition [@Lincoln_2012], lack of research funding [@Bozeman_2004], lack of invitation to speak at scientific events [@Isbell_2012; @Schroeder_2013; @Klein_2017], discrimination during hiring [@MossRacusin_2017; @Reuben_2014] and discrimination when writing a recommendation letter [@Trix_2003; @Schmader_2007], are only some of the small disadvantages adding up to drive women out of science.
+Although the proportion of women in science has significantly increased over the last decades, women still tend to leave their academic careers behind considerably more often than their male counterparts [REF].
+Lack of scientific recognition [@Lincoln_2012], lack of research funding [@Bozeman_2004], lack of invitation to speak at scientific events [@Isbell_2012; @Klein_2017; @debarre_2018], discrimination during hiring [@MossRacusin_2017; @Reuben_2014] and discrimination when writing a recommendation letter [@Trix_2003; @Schmader_2007], are only some of the small disadvantages adding up to drive women out of science.
 Whether their decision to leave academia is deliberate or not, it will negatively impact science overall, as mentoring, networking, and supporting younger women scientists are especially important for women in academia.

### 2018-05-29 manuscript/Manuscript_Claire.Rmd
Minor writing

@@
 # Introduction
-
-- Remember dataset S1 issue (put in ESM?)
-- re-run model on line 912 and 931 (weights issue)
-- check figure legend line 900. Also check the 'significant' area hasn't changed loads
-- proportional alpha?
 
@@
 The excess of same-gender coauthorships was quite large: many journals had $\alpha' > 0.1$, indicating that the gender ratio of men's and women's coauthors differs by >10% in absolute terms.
-In relative terms, the observed homophily is even more striking: for example, if the proportion of male coauthors is 80% for men and 70% for women ($\alpha' = 0.1$), the results imply that women publish with women 50% more frequently than men do.
+In relative terms, our findings are even more striking: for example, if men have 20% female coauthors and women have 30% (i.e.
+$\alpha' = 0.1$ for a field with a male-biased gender ratio), then women publish with women 50% more often than men do.
 
-An important limitation of our study is that we cannot reliably ascertain the cause(s) of the observed excess of same-gender coauthorships.
-As well as conscious or unconcious selection of collaborators based on gender, our results could be partly or wholly explained by uncontrolled Wahlund effects.
+An important limitation of our study is that we cannot reliably determine the cause(s) of the observed excess of same-gender coauthorships.
+As well as conscious or unconcious selection of collaborators based on gender, our results could be partly explained by uncontrolled Wahlund effects.
 However, we suspect the contribution of these to be minor, for four reasons: we found positive $\alpha'$ after controlling for three obvious sources of Wahlund effect; there was no inflation of $\alpha'$ in highly multidisciplinary journals; restricting the data by country yielded similar estimates of $\alpha'$; and we showed that differences in gender ratio between career stages are unlikely to fully explain our results.
-On balance, we believe the data provide good evidence that researchers preferentially select same-gendered collaborators, although the effect size of this preference is uncertain.
+On balance, we believe the data provide good evidence that researchers preferentially select same-gendered collaborators, although the strength of this preference is uncertain.
 
-Some disciplines exhibit especially skewed gender ratios; e.g.
-most authors in physics and computer science are men, while most authors in nursing are women [@Holman_2018].
-We hypothesised that these gender-imbalanced disciplines might show the strongest gender homophily, e.g.
-because being in the minority might increase one's drive to seek out same-gendered colleagues.
+We hypothesised that disciplines with a strongly skewed gender ratio might show the strongest gender homophily, e.g.
+because being in the minority might increase motivation to seek out same-gendered colleagues.
 Contrary to this hypothesis, we found no evidence that gender homophily is restricted to particular disciplines: $\alpha'$ was similarly high across the board (\autoref{alpha_histograms}).
@@
 Interestingly, $\alpha'$ was highest in journals with a female-biased author gender ratio.
-One interpretation of this result is that men are more likely to preferentially seek out male collaborators in fields where men are a minority, relative to the homophily displayed by women in fields where women are a minority.
-However, this latter result is only tentatively supported since there are so few journals in which most authors are women (see \autoref{alpha_gender_ratio}).
+This may suggest that men are more likely to preferentially seek out male collaborators in fields where men are a minority, relative to the homophily displayed by women in fields where women are a minority.
+However, this latter result is only tentatively supported since there are few journals in which most authors are women (\autoref{alpha_gender_ratio}).
 
@@
 These questions are closely related.
-For example, some of the homophily we observed might be caused by women seeking to avoid harassment, sexism or microaggressions from men [@jagsi_2016], which would clearly be concerning.
+For example, some of the homophily we observed might be caused by women seeking to avoid harassment or sexism from men [@jagsi_2016], which would clearly be concerning.
 Additionally, Sheltzer and Smith [-@sheltzer_2014] concluded that 'elite' male academics (defined as recipients of major honours) have a higher proportion of male students and postdocs than non-elite male academics.
@@
 [-@Ghiasi_2015] concluded that women in engineering are "compliant [in reproducing] male-dominated scientific structures" because they do not collaborate often enough with other women; their data suggest that coauthorships between two women are about 30% more frequent than expected under the null (see their Figure 7).
-In contrast, we propose that it behoves men as well as women to ensure that they are not overlooking or excluding female colleagues.
-Given that women work disproportionately often in junior positions with little opportunity to be selective, one could argue that the responsibilty to address gender imblances in science rests primarily with men.
+In contrast, we propose that men as well as women should ensure that they are not inadvertently overlooking or excluding female colleagues, particularly since men are disproportionately represented among senior researchers [@Shaw_2012].
 

### 2018-06-03 manuscript/Manuscript_Claire.Rmd
Luke writing

@@
 
-Women are highly underrepresented in many branches of science, technology, engineering, mathematics, and medicine (STEMM), and face additional challenges and inequities relative to men [e.g.
+Women are severely underrepresented in many branches of science, technology, engineering, mathematics, and medicine (STEMM), and face additional challenges and inequities relative to men [e.g.
 @Shaw_2012; @Lariviere_2013; @West_2013; @Elsevier_report; @Holman_2018].
-On average, women occupy more junior positions [@Wutte_2007; @Reuben_2014] with lower salaries [@Trower_2002; @Umbach_2007], receive less grant money [@Hosek_2005; @OER_2008], are promoted more slowly [@Zuckerman_1987; @Rosenfeld_1991; @Long_1993; @Hopkins_2013], and are allocated fewer resources [@ODorchai_2009] and less research funding [@Feldt_1986; @Stack_2004; @Lariviere_2011] than men.
-Experimental studies suggest that women's research achievements are regarded less favourably than identical achievements by men [@Moss_2012; @Knobloch_2013].
+On average, women occupy more junior positions [@Wutte_2007; @Reuben_2014] with lower salaries [@Trower_2002; @Umbach_2007], receive less grant money [@Hosek_2005; @OER_2008], are promoted more slowly [@Zuckerman_1987; @Rosenfeld_1991; @Long_1993; @Hopkins_2013], and are allocated fewer resources [@ODorchai_2009] and less research funding [@Feldt_1986; @Stack_2004; @Lariviere_2011].
+Experimental evidence suggests that bias against women plays a major role in generating these differences [@Moss_2012; @Knobloch_2013].
 
@@
 For example, studies have found that women tend to be less involved in international collaboration [@Lewison_2001; @Webster_2001; @Bozeman_2004; @Lariviere_2011; @Abramo_2013], collaborate less within their own university departments [@Webster_2001], have less prestigious collaborations [@Long_1990], and fewer collaborations in total [@Fuchs_2001].
-These gender differences in collaboration practice presumably have multiple causes, which might include implicit and explicit gender bias [@Moss_2012], differential family obligations [@Reskin_1978; @Long_1990; @Wright_2003], concerns relating to sexual harassment [@jagsi_2016], and unequal access to conferences [@Martin_2014] and travel funds [@Bozeman_2004].
+These gender differences in collaboration practice presumably have multiple causes, which might include implicit and explicit gender bias [@Moss_2012], differential family obligations [@Reskin_1978; @Long_1990; @Wright_2003], gender differences in confidence or self-esteem [@bleidorn2016age], concerns relating to sexual harassment [@jagsi_2016], and unequal access to conferences [@Martin_2014] and travel funds [@Bozeman_2004].
 
-A high, steadily increasing proportion of research papers is written by more than one author [@West_2013], making collaboration a key predictor of publication output, and thus of career prospects [@Tower_2007; @Jordan_2008; @Cohn_2014].
-Additionally, empirical studies imply that mixed-gender [@Bear_2011; @Campbell_2013] or otherwise diverse [@Hong_2004] teams produce better results on collaborative tasks than less diverse teams.
-For reasons such as these, several studies have tested for gender differences in collaboration frequency or pattern by examining the author lists of published research.
-To our knowledge, every published study of this question suggests that men co-publish with other men, and women with women, more often than expected if collaborators assort randomly with respect to gender [@Ferber_1980; @McDowell_1992; @Ghiasi_2015; @Crow_2015; @Fahmy_2017; @Zettler_2017; @Jadidi_2017; @Teele_2017; @Araujo_2017a; @Araujo_2017b].
-This pattern of assortative publishing has been termed 'gender homophily'.
+A high, steadily increasing proportion of research papers is written by more than one author [@West_2013], making collaboration a key predictor of publication output, and thus of career prospects [@Tower_2007; @Jordan_2008].
+Additionally, empirical studies imply that mixed-gender or otherwise diverse teams produce better results on collaborative tasks than less diverse teams [@reagans2001; @Hong_2004; @Bear_2011; @Campbell_2013].
+For reasons such as these, several studies have examined the author lists of published research articles in order to test for gender differences in collaboration frequency or pattern.
+To our knowledge, all of these studies imply that men co-publish with other men, and women with women, more often than expected if collaborators assort randomly with respect to gender [@Ferber_1980; @McDowell_1992; @Ghiasi_2015; @Crow_2015; @Fahmy_2017; @Zettler_2017; @Jadidi_2017; @Teele_2017; @Araujo_2017a; @Araujo_2017b].
+This pattern of assortative publishing is often termed 'gender homophily'.
 
-However, we believe that prior studies of gender homophily were hindered by a largely unacknowledged statistical issue that we term the Wahlund effect (\autoref{wahlund_plot}), by analogy with a conceptually similar effect in population genetics [@Wahlund_1928].
-The Wahlund effect makes it deceptively difficult to infer gender-based preferences simply by counting same- and mixed-gender coauthorships.
+However, we believe that prior studies of gender homophily were hindered by a largely unacknowledged statistical issue that we name the Wahlund effect (\autoref{wahlund_plot}), by analogy with the conceptually similar Whalund effect in population genetics [@Wahlund_1928].
+The Wahlund effect makes it deceptively difficult to infer gender-based preferences simply by counting the number of same- and mixed-gender coauthorships.
 Essentially, whenever coauthorship data are sampled from two or more discrete sets of literature, which vary in the author gender ratio and which are largely not connected by collaboration, the number of same-gendered coauthors will be inflated.
-This can give the impression that authors preferentially publish with same-gendered colleagues if no gender preferences exist, or even if opposite-gendered colleagues are prefered.
+This can give the impression that authors preferentially publish with same-gendered colleagues if no gender preferences exist, or even if opposite-gendered colleagues are preferred.
 For example, a sample of literature containing bioinformatics and cell biology papers will probably contain an excess of mostly-male and mostly-female author lists, simply because researchers preferentially work with colleagues from the same discipline, and because the author gender ratio is more male-biased in bioinformatics than in cell biology [@Holman_2018].
@@
 
-In the present study, we test whether researchers working in the life sciences tend to co-publish with same-gendered collaborators, while controlling for the Wahlund effect as strictly as possible.
-Our study uses a recently-published dataset describing the gender of 35.5m authors from 9.15m articles indexed on PubMed [@Holman_2018].
+In the present study, we test whether life sciences researchers tend to co-publish with same-gendered colleagues, while controlling for the Wahlund effect as strictly as possible.
+We use a recently-published dataset describing the gender of 35.5m authors from 9.15m articles indexed on PubMed [@Holman_2018].
 Holman et al.
 [-@Holman_2018] reported large differences in the gender ratio of authors across research disciplines, journals, countries, and across the years 2002-2016.
-We therefore tested for gender homophily after restricting the analysis to particular journals (i.e.
+We therefore tested for gender homophily after restricting each analysis to particular journals (i.e.
 research specialties), time periods, and countries.
@@
 For example, this trend might result from the increasing number of women working in senior positions in STEMM over the past decade [e.g.
-@Long_2015; @Bendels_2018; @McKenzie_2017].
+@Shaw_2012; @Long_2015; @McKenzie_2017; @Bendels_2018].
 As shown in \autoref{simulation}, if enough coauthorships are between junior and senior researchers, a large gender gap between career stages can give the appearance of heterophily.
@@
 We also found a little evidence that gender homophily is detrimental to research quality, in that high-impact journals tended to have weaker homophily.
-Assuming that papers published in high-impact journals are of higher average quality [which is contentious; @Garfield_2006], this result provides non-experimental support for the hypothesis that mixed-gender teams produce better research than single-gender teams [@Campbell_2013].
+Assuming that papers published in high-impact journals are of higher average quality [which is contentious; @Garfield_2006], this result provides non-experimental support for the hypothesis that mixed-gender teams produce better research than single-gender teams [@reagans2001; @Hong_2004; @Bear_2011; @Campbell_2013].
 

### 2018-06-03 manuscript/Manuscript_Claire.Rmd
Merge in Claire's updates

@@
   \rhead[]{Gender and coauthorship} \usepackage[noabbrev,capitalise]{cleveref} \usepackage{titlefoot}
-  \usepackage{amssymb} \usepackage{rotating}
+  \usepackage{amssymb} \usepackage{rotating} \usepackage{lineno} \linenumbers
 csl: the-american-naturalist.csl
@@
   by senior academics, and no evidence that homophily is strongest in fields where
-  women are in the minority. Interestingly, homophily was negatively correlated with
-  journal impact factor, as predicted if mixed-gender teams produce better research.
+  women are in the minority.
+Interestingly, journals with a high impact factor for their discipline tended to have comparatively low homophily, as predicted if mixed-gender teams produce better research.
   We discuss potential causes of gender homophily in science. \vspace{5mm} \par\noindent
@@
 @Shaw_2012; @Lariviere_2013; @West_2013; @Elsevier_report; @Holman_2018].
-On average, women occupy more junior positions [@Wutte_2007; @Reuben_2014] with lower salaries [@Trower_2002; @Umbach_2007], receive less grant money [@Hosek_2005; @OER_2008], are promoted more slowly [@Zuckerman_1987; @Rosenfeld_1991; @Long_1993; @Hopkins_2013], and are allocated fewer resources [@ODorchai_2009] and less research funding [@Feldt_1986; @Stack_2004; @Lariviere_2011].
+On average, women occupy more junior positions [@Wutte_2007; @Reuben_2014] with lower salaries [@Trower_2002; @Umbach_2007], receive less grant money [@Hosek_2005; @pohlhaus_2011], are promoted more slowly [@Zuckerman_1987; @Rosenfeld_1991; @Long_1993; @Hopkins_2013], and are allocated fewer resources [@ODorchai_2009] and less research funding [@Feldt_1986; @Stack_2004; @Lariviere_2011].
 Experimental evidence suggests that bias against women plays a major role in generating these differences [@Moss_2012; @Knobloch_2013].
@@
 A high, steadily increasing proportion of research papers is written by more than one author [@West_2013], making collaboration a key predictor of publication output, and thus of career prospects [@Tower_2007; @Jordan_2008].
-Additionally, empirical studies imply that mixed-gender or otherwise diverse teams produce better results on collaborative tasks than less diverse teams [@reagans2001; @Hong_2004; @Bear_2011; @Campbell_2013].
+Additionally, empirical studies imply that mixed-gender or otherwise diverse teams produce better results on collaborative tasks than less diverse teams [@Britton_2000; @Reagans_2001; @Hong_2004; @Whittington_2008; @Bear_2011; @Herrera_2012; @Campbell_2013].
 For reasons such as these, several studies have examined the author lists of published research articles in order to test for gender differences in collaboration frequency or pattern.
@@
 [-@Holman_2018] reported large differences in the gender ratio of authors across research disciplines, journals, countries, and across the years 2002-2016.
-We therefore tested for gender homophily after restricting each analysis to particular journals (i.e.
+We therefore tested for gender homophily while restricting our analysis to particular journals (i.e.
 research specialties), time periods, and countries.
+We quantified gender assortment using a metric called $\alpha'$ [@bergstrom_2016], which is positive when same-gender authors publish together more often than expected (gender homohily), negative when opposite-gender authors publish together more often than expected (heterophily), and equal to zero when authors assort randomly with respect to gender (see Methods).
+
+# Results
+## Gender homophily by discipline, time period, and authorship position
+\autoref{alpha_histograms} shows the distribution of $\alpha'$ estimates in 2015-2016 across all journals for which we recovered sufficient data, when $\alpha'$ was calculated for all authors, first authors only, or last authors only.
+The great majority of journals had $\alpha' > 0$, and for many of these the FDR-corrected p-values suggested that $\alpha'$ was significantly greater than zero (1469/2077 journals were significant in 2015-16, and 404/1192 in 2005-6).
+Only 2/2077 journals had statistically significantly heterophily (i.e.
+$\alpha' < 0$) in 2015-16, and 1/1192 in 2005-6 (S2 Table).
+The remaining 606 or 787 journals (in 2015 and 2005 respectively) had a value of $\alpha'$ not significantly different from zero, consistent with the null hypothesis of random assortment with respect to gender.
+
+![Of the `r n.journals` journals for which we had adequate data in 2015-2016, `r n.sig.homo` showed statistically significant evidence of gender homophily (denoted by $\alpha' > 0$), and `r n.sig.hetero` showed statistically significant evidence of heterophily ($\alpha' < 0$), after false discovery rate correction.
+The white area shows the number of journals for which homophily was significantly stronger than expected under the null hypothesis (p < 0.05), while the blue area shows all the remainder.
+Patterns were similar whether $\alpha'$ was calculated for all authors, for first authors only, or for last authors only.
+\label{alpha_histograms}](../figures/figure2.pdf)
+
+$\alpha'$ was significantly higher in the literature sample from 2015-16 relative to 2005-6, though the difference in means was small (S1 Fig; Effect of the fixed factor 'Time period' in a linear mixed model of the data for all author positions: Cohen's $d$ = $0.091{\pm}0.04$, $t_{953}$ = 2.42, p = 0.016).
+
+When comparing pairs of $\alpha'$ values estimated for the first and last authors for the same journals, we found that $\alpha'$ tended to be higher for first authors than for last authors (S2 Fig; Effect of the fixed factor 'Authorship position' in a linear mixed model: Cohen's $d$ = $0.065{\pm}0.02$, $t_{2024}$ = 4.28, p < 0.0001).
+This suggests that the gender of the first author was a slightly stronger predictor of the remaining authors' genders than the gender of the last author, i.e.
+the opposite of what is predicted if senior scientists are causally responsible for homophily.
+
+## Variance in homophily between disciplines
+\autoref{alpha_histograms} illustrates the variance in journal homophily values ($\alpha'$) across scientific disciplines.
+All disciplines had a positive average $\alpha'$, although homophily appeared somewhat stronger in some disciplines than others (e.g.
+mean $\alpha'$ was $0.12{\pm}0.02$ for Urology journals and $0.03{\pm}0.01$ for Veterinary Medicine journals; \autoref{alpha_histograms}, S3 Table).
+However, there was no evidence for consistent differences in $\alpha'$ between disciplines: the random factor 'Discipline' explained around 1% of the variance in $\alpha'$ in the two linear mixed models described in the previous section (see \autoref{alpha_histograms} and mixed models in Online Supplementary Material).
+Thus, we cannot reject the null hypothesis that the processes causing positive $\alpha'$ are similarly strong in all the disciplines we examined.
+
+There was no indication that journals publishing on a wide range of topics have higher $\alpha'$ values than more specialised journals, due to the Wahlund effect.
+For example, the journal category 'Multidisciplinary' -- which includes journals like _PLoS ONE_, _Nature_, _Science_, and _PNAS_ -- did not have notably elevated $\alpha'$ (\autoref{alpha_histograms}).
+This result suggests that our estimates of homophily, and estimates from some earlier studies, are not notably inflated by the presence of disparate research topics (with variable author gender ratios) being published within individual journals.
+
+## Relationship between gender homophily and gender ratio
+We next tested whether researchers are more or less likely to publish with same-gendered colleagues in strongly gender-biased disciplines (e.g.
+Surgery or Nursing), relative to disciplines with a comparatively gender-balanced workforce (e.g.
+Psychiatry).
+We found a positive, non-linear relationship between the overall gender ratio of all authors publishing in a particular journal [as estimated in @Holman_2018], and the estimated value of $\alpha'$ for all authors and for first authors (\autoref{alpha_gender_ratio}).
+Journals with a balanced or female-biased author gender ratio tended to have higher $\alpha'$ than journals with a male-biased author gender ratio (GAM smooth terms p < 0.001; Online Supplementary Material).
+The relationship was not statistically significant when $\alpha'$ was calculated for last authors (GAM, p = 0.142), though the trend appeared similar (\autoref{alpha_gender_ratio}).
+
+![There is a weakly positive, non-linear relationship between the gender ratio of authors publishing in a journal, and the coefficient of homophily ($\alpha'$).
+Specifically, journals with 50% women authors or higher tended to have more same-sex coauthorships than did journals with predominantly men authors.
+This relationship held whether $\alpha'$ was calculated for all authors, first authors only, or last authors only.
+A negative value on the x-axis denotes an excess of men authors, a positive value denotes an excess of women authors, and zero denotes gender parity.
+The lines were fitted using generalised additive models with the smoothing parameter $k$ set to 3.
+\label{alpha_gender_ratio}](../figures/figure3.pdf)
+
+## Relationship between journal impact factor and gender homophily
+We observed a noisy but statistically significant linear relationship between standardised journal impact factor and $\alpha'$, such that journals with a high impact factor for their discipline had weaker gender homophily than did journals with a low impact factor for their discipline (\autoref{impact_factor}; linear regression: $R^2$ = `r R2`, $t_{1415}$ = -8.0, p < 0.0001).
+
+![Journal impact factor (expressed relative to the average for the discipline) is negatively correlated with $\alpha'$.
+The relationship is noisy ($R^2$ = `r R2`), but the results suggest that journals with strong homophily tend to have lower impact factors than journals with weak homophily in the same discipline.
+\label{impact_factor}](../figures/figure4.pdf)
+
+## Analysis accounting for differences in author gender ratio between countries
+When we restricted the analysis by country, we observed statistically significant homophily for 72 of the 325 journal-country combinations tested (64 unique journals and 18 unique countries), and no significant evidence of heterophily (Figures S3-S4).
+Additionally, the values of $\alpha'$ calculated for each journal-country combination were mostly very similar to the $\alpha'$ values calculated for the journal as a whole (i.e.
+when pooling papers from different countries); on average, $\alpha'$ was only inflated in the main (non-country specific) analysis by 0.002 (S5 Fig).
+These results suggest that our findings of widespread homophily in the main analysis were not driven solely by a Wahlund effect resulting from gender differences between countries.
+
+## Theoretical expectations for $\alpha$ when the gender ratio differs between career stages
+As shown in \autoref{simulation}, we predict that $\alpha$ is expected to be non-zero, even if collaborators are randomly selected with respect to gender, provided that there is a gender gap between career stages.
+The extent to which $\alpha$ deviates from zero depends on the relative frequencies of collaboration within and between career stages.
+When >50% of collaborations were between early and established researchers, we expect gender heterophily ($\alpha < 0$).
+Conversely, when >50% of collaborations occured within career stages, we expect gender homophily ($\alpha > 0$).
+In a few parameter spaces (shown in red; \autoref{simulation}), $\alpha$ was quite high, and overlapped with the values that we estimated (\autoref{alpha_histograms}).
+
+![When there is a difference in gender ratio between early-career and established researchers, and collaboration is non-random with respect to career stage, the null expectation for $\alpha$ deviates from zero.
+An excess of collaborations between career stages gives the appearance of gender heterophily (lower rows, blue areas), while an excess of within-career stage collaborations produced apparent gender homophily (upper rows, red areas).
+However, the conditions required for strong gender homophily are quite restrictive, making it unlikely that this issue explains all of the homophily observed in Figure 2.
+Contour lines denote increments of 0.1.
+\label{simulation}](../figures/figure5 - Simulation.pdf)
+
+Despite this overlap, \autoref{simulation} suggests that our main conclusions (and those of other studies of gender homophily) are probably robust to this career stage issue.
+We only expect strongly positive $\alpha$ when A) the gender ratio is highly skewed across career stages (e.g.
+a 5-fold difference), and B) collaborations between early and established researchers are very rare (e.g.
+<10% of the total).
+Both of these conditions are untrue for most fields: the gender gap across careers stages is generally less pronounced [e.g.
+@Shaw_2012; @Holman_2018], and it is very common for early-career researchers to co-publish with an established mentor [@macaluso_2016].
+However, one can get $\alpha > 0$ for realistic combinations of parameters, e.g.
+a moderate shortage of women in senior positions coupled with a moderate excess of within-career stage collaboration, suggesting this effect might contribute to some of the observed homophily.
+
+Lastly, we note that if there is a gender gap between career stages and coauthorships between early-career and established researchers comprise >50% of the total, then the baseline expectation for $\alpha$ is actually less than zero (blue areas in \autoref{simulation}).
+Therefore, our results might under-estimate the extent to which researchers preferentially select same-gendered collaborators in some cases.
+
+# Discussion
+We found evidence that researchers preferentially publish with same-gendered coauthors, even after implementing stringent controls for Wahlund effects (\autoref{wahlund_plot}).
+Our study therefore reaffirms earlier studies' conclusions [e.g.
+@Ferber_1980; @McDowell_1992; @Bentley_2003; @Crow_2015; @Ghiasi_2015; @Araujo_2017a; @Fahmy_2017; @Jadidi_2017; @Teele_2017; @Zettler_2017] and establishes their generality across the life sciences.
+Relatively few journals had $\alpha'$ values below zero, and almost no journals showed statistically significant gender heterophily after controlling for multiple testing.
+The excess of same-gender coauthorships was quite large: many journals had $\alpha' > 0.1$, indicating that the gender ratio of men's and women's coauthors differs by >10% in absolute terms.
+In relative terms, our findings are even more striking: for example, if men have 20% female coauthors and women have 30% (i.e.
+$\alpha' = 0.1$ for a field with a male-biased gender ratio), then women publish with women 50% more often than men do.
+
+An important limitation of our study is that we cannot reliably determine the cause(s) of the observed excess of same-gender coauthorships.
+As well as conscious or unconcious selection of collaborators based on gender, our results could be partly explained by uncontrolled Wahlund effects.
+However, we suspect the contribution of these to be minor, for four reasons: we found positive $\alpha'$ after controlling for three obvious sources of Wahlund effect; there was no inflation of $\alpha'$ in highly multidisciplinary journals; restricting the data by country yielded similar estimates of $\alpha'$; and we showed that differences in gender ratio between career stages are unlikely to fully explain our results.
+On balance, we believe the data provide good evidence that researchers preferentially select same-gendered collaborators, although the strength of this preference is uncertain.
+
+We hypothesised that disciplines with a strongly skewed gender ratio might show the strongest gender homophily, e.g.
+because being in the minority might increase motivation to seek out same-gendered colleagues.
+Contrary to this hypothesis, we found no evidence that gender homophily is restricted to particular disciplines: $\alpha'$ was similarly high across the board (\autoref{alpha_histograms}).
+Moreover, $\alpha'$ was lower for journals with a male-biased author gender ratio relative to those with an even gender ratio.
+Interestingly, $\alpha'$ was highest in journals with a female-biased author gender ratio.
+This may suggest that men are more likely to preferentially seek out male collaborators in fields where men are a minority, relative to the homophily displayed by women in fields where women are a minority.
+However, this latter result is only tentatively supported since there are few journals in which most authors are women (\autoref{alpha_gender_ratio}).
+
+We also found that gender homophily was marginally stronger in 2015-2016 relative to 2005-2006.
+Although this trend might reflect a change in the gender preferences of researchers seeking collaborators over time, there are alternative (and perhaps more likely) explanations.
+For example, this trend might result from the increasing number of women working in senior positions in STEMM over the past decade [e.g.
+@Long_2015; @Bendels_2018; @McKenzie_2017].
+As shown in \autoref{simulation}, if enough coauthorships are between junior and senior researchers, a large gender gap between career stages can give the appearance of heterophily.
+As this gender gap between career stages lessens, the observed values of $\alpha$ may increase.
+
+Our study begs two questions: what causes gender homophily in science, and are our results cause for concern?
+These questions are closely related.
+For example, some of the homophily we observed might be caused by women seeking to avoid harassment or sexism from men [@jagsi_2016], which would clearly be concerning.
+Additionally, Sheltzer and Smith [-@sheltzer_2014] concluded that 'elite' male academics (defined as recipients of major honours) have a higher proportion of male students and postdocs than non-elite male academics.
+This finding could contribute to the homophily we observed, and is cause for concern since Sheltzer and Smith [-@sheltzer_2014]'s results might reflect discrimination against women during hiring [@Moss_2012], or avoidance by women of elite research groups (e.g.
+due to 'imposter syndrome' or the perception that these groups are sexist).
+We also found a little evidence that gender homophily is detrimental to research quality, in that high-impact journals tended to have weaker homophily.
+Assuming that papers published in high-impact journals are of higher average quality [which is contentious; @Garfield_2006], this result provides non-experimental support for the hypothesis that mixed-gender teams produce better research than single-gender teams [@Britton_2000; @Reagans_2001; @Hong_2004; @Whittington_2008; @Bear_2011; @Herrera_2012; @Campbell_2013].
+
+Homophily might also have more benign causes.
+Collaboration is often most enjoyable and productive when working with like-minded people, who might be same-gendered more often than not.
+We also suppose that some people consciously choose to preferentially collaborate with women in order to help close the gender gap in the workforce; this would create homophily if women do this more than men.
+In support of this interpretation, women appear more likely than men to promote the work of female collagues by inviting them to give talks [@nittrouer_2018; @debarre_2018].
+Related to this, Ghiasi et al.
+[-@Ghiasi_2015] concluded that women in engineering are "compliant [in reproducing] male-dominated scientific structures" because they do not collaborate often enough with other women; their data suggest that coauthorships between two women are about 30% more frequent than expected under random assortment (see their Figure 7).
+In contrast, we propose that men as well as women should ensure that they are not inadvertently overlooking or excluding female colleagues, particularly since men are disproportionately represented among senior researchers [@Shaw_2012].
 
@@
 Lastly, we excluded journals with fewer than 50 suitable papers.
-
-The final sample sizes were `r sample.size$n2015[3]` papers and `r sample.size$n2015[4]` authorships from August 2015 - August 2016, and `r sample.size$n2005[3]` papers and `r sample.size$n2005[4]` authorships from August 2005 - August 2006.
-Respectively, these papers came from `r sample.size$n2015[2]` and `r sample.size$n2005[2]` journals, which were grouped into `r sample.size$n2015[1]` and `r sample.size$n2005[1]` research disciplines.
-There was a median of `r sample.size$n2015[5]` (`r sample.size$n2005[5]`) papers per journal, `r sample.size$n2015[6]` (`r sample.size$n2005[6]`) authors per journal, and `r sample.size$n2015[7]` (`r sample.size$n2005[7]`) authors per paper (the first number is for 2015-6, and the bracketed number for 2005-6).
+Detailed sample size information is given in S1 Table.
 
@@
 [-@bergstrom_2016], we defined the coefficient of homophily as $\alpha = p - q$, where $p$ is the probability that a randomly-chosen co-author of a _male_ author is a man and $q$ is the probability that a randomly-chosen co-author of a _female_ author is a man.
-Therefore, $\alpha > 0$ suggests that same-gender authors publish together more often than expected under the null hypothesis that authors assort randomly with respect to gender (homophily), $\alpha < 0$ suggests that opposite-gender authors publish together more often than expected (heterophily), and $\alpha = 0$ suggests random assortment with respect to gender.
 Like the Wahlund effect, $\alpha$ is borrowed from population genetics; for a set of 2-author papers, it is equivalent to Wright's coefficient of inbreeding [@wright_1949].
-Mathematical work illustrates that $\alpha$ is closely related to network-based alternative methods for quantifying homophily [@wang_2016].
+Mathematical work illustrates that $\alpha$ is closely related to alternative network-based methods for quantifying homophily [@wang_2016].
 
@@
 
-As expected, $\alpha'$ was usually almost identical to $\alpha$ (Figure S1), but $\alpha$ was downwardly biased relative to $\alpha'$ for small datasets (Figure S2).
+As expected, $\alpha'$ was usually almost identical to $\alpha$ (S6 Fig), but $\alpha$ was downwardly biased relative to $\alpha'$ for small datasets (S7 Fig).
 Additionally, the correlation between $\alpha'$ and sample size was negligible ($R^2 < 0.01$), suggesting that our calculation of $\alpha'$ effectively removed the dependence of $\alpha$ on sample size.
@@
 
-# Results
-## Gender homophily by discipline, time period, and authorship position
-\autoref{alpha_histograms} shows the distribution of $\alpha'$ estimates in 2015-2016 across all journals for which we recovered sufficient data, when $\alpha'$ was calculated for all authors, first authors only, or last authors only.
-The great majority of journals had $\alpha' > 0$, and for many of these the FDR-corrected p-values suggested that $\alpha'$ was significantly greater than zero (1469/2077 journals were significant in 2015-16, and 404/1192 in 2005-6).
-Only 2/2077 journals had statistically significantly heterophily (i.e.
-$\alpha' < 0$) in 2015-16, and 1/1192 in 2005-6.
-The remaining 606 or 787 journals (in 2015 and 2005 respectively) had a value of $\alpha'$ not significantly different from zero, consistent with the null hypothesis of random assortment with respect to gender.
-
-![Of the `r n.journals` journals for which we had adequate data in 2015-2016, `r n.sig.homo` showed statistically significant evidence of homophily (denoted by $\alpha' > 0$), and `r n.sig.hetero` showed statistically significant evidence of heterophily ($\alpha' < 0$), after adjusting p-values using Benjamini-Hochberg false discovery rate correction.
-The white area shows the number of journals for which homophily was significantly stronger than expected under the null hypothesis (p < 0.05), while the blue area shows all the remainder.
-Patterns were similar whether $\alpha'$ was calculated for all authors, for first authors only, or for last authors only.
-\label{alpha_histograms}](../figures/figure2.pdf)
-
-$\alpha'$ was significantly higher in the literature sample from 2015-16 relative to 2005-6, though the difference in means was small (Figure S3; Effect of the fixed factor 'Time period' in a linear mixed model of the data for all author positions: Cohen's $d$ = $0.091{\pm}0.04$, $t_{953}$ = 2.42, p = 0.016).
-
-When comparing pairs of $\alpha'$ values estimated for the first and last authors for the same journals, we found that $\alpha'$ tended to be higher for first authors than for last authors (Figure S4; Effect of the fixed factor 'Authorship position' in a linear mixed model: Cohen's $d$ = $0.065{\pm}0.02$, $t_{2024}$ = 4.28, p < 0.0001).
-This suggests that the gender of the first author was a slightly stronger predictor of the remaining authors' genders than the gender of the last author, i.e.
-the opposite of what is predicted if senior scientists are causally responsible for homophily.
-
-## Variance in homophily between disciplines
-\autoref{alpha_histograms} illustrates the variance in journal homophily values ($\alpha'$) across scientific disciplines.
-All disciplines had a positive average $\alpha'$, although homophily appeared somewhat stronger in some disciplines than others (e.g.
-mean $\alpha'$ was $0.12{\pm}0.02$ for Urology journals and $0.03{\pm}0.01$ for Veterinary Medicine journals; \autoref{alpha_histograms}, Table S3).
-However, there was no evidence for consistent differences in $\alpha'$ between disciplines: the random factor 'Discipline' explained around 1% of the variance in $\alpha'$ in the two linear mixed models described in the previous section (see \autoref{alpha_histograms} and mixed models in Online Supplementary Material).
-Thus, we cannot reject the null hypothesis that the processes causing positive $\alpha'$ are similarly strong in all the disciplines we examined.
-
-There was no indication that journals publishing on a wide range of topics have higher $\alpha'$ values than more specialised journals, due to the Wahlund effect.
-For example, the journal category 'Multidisciplinary' -- which includes journals like _PLoS ONE_, _Nature_, _Science_, and _PNAS_ -- did not have notably elevated $\alpha'$ (\autoref{alpha_histograms}).
-This result suggests that our estimates of homophily, and estimates from some earlier studies, are not notably inflated by the presence of disparate research topics (with variable author gender ratios) being published within individual journals.
-
-## Relationship between gender homophily and gender ratio
-We next tested whether researchers are more or less likely to publish with same-gendered colleagues in strongly gender-biased disciplines (e.g.
-Surgery or Nursing), relative to disciplines with a comparatively gender-balanced workforce (e.g.
-Psychiatry).
-We found a positive, non-linear relationship between the overall gender ratio of all authors publishing in a particular journal [as estimated in @Holman_2018], and the estimated value of $\alpha'$ for all authors and for first authors (\autoref{alpha_gender_ratio}).
-Journals with a balanced or female-biased author gender ratio tended to have higher $\alpha'$ than journals with a male-biased author gender ratio (GAM smooth terms p < 0.001; Online Supplementary Material).
-The relationship was not statistically significant when $\alpha'$ was calculated for last authors (GAM, p = 0.142), though the trend appeared similar (\autoref{alpha_gender_ratio}).
-
-![There is a weakly positive, non-linear relationship between the gender ratio of authors publishing in a journal, and the coefficient of homophily ($\alpha'$).
-Specifically, journals with 50% women authors or higher tended to have more same-sex coauthorships than did journals with predominantly men authors.
-This relationship held whether $\alpha'$ was calculated for all authors, first authors only, or last authors only.
-A negative value on the x-axis denotes an excess of men authors, a positive value denotes an excess of women authors, and zero denotes gender parity.
-The lines were fitted using generalised additive models with the smoothing parameter $k$ set to 3.
-\label{alpha_gender_ratio}](../figures/figure3.pdf)
-
-## Relationship between journal impact factor and gender homophily
-We observed a noisy but statistically significant linear relationship between standardised journal impact factor and $\alpha'$, such that journals with a high impact factor for their discipline had weaker gender homophily than did journals with a low impact factor for their discipline (\autoref{impact_factor}; linear regression: $R^2$ = `r R2`, $t_{1415}$ = -8.0, p < 0.0001).
-
-![Journal impact factor (expressed relative to the average for the discipline) is negatively correlated with $\alpha'$.
-The relationship is noisy ($R^2$ = `r R2`), but the results suggest that journals with strong homophily tend to have lower impact factors than journals with weak homophily in the same discipline.
-\label{impact_factor}](../figures/figure4.pdf)
-
-## Analysis accounting for differences in author gender ratio between countries
-When we restricted the analysis by country, we observed statistically significant homophily for 72 of the 325 journal-country combinations tested (64 unique journals and 18 unique countries), and no significant evidence of heterophily (Figures S5-S6).
-Additionally, the values of $\alpha'$ calculated for each journal-country combination were mostly very similar to the $\alpha'$ values calculated for the journal as a whole (i.e.
-when pooling papers from different countries); on average, $\alpha'$ was only inflated in the main (non-country specific) analysis by 0.002 (Figure S7).
-These results suggest that our findings of widespread homophily in the main analysis were not driven solely by a Wahlund effect resulting from gender differences between countries.
-
-## Theoretical expectations for $\alpha$ when the gender ratio differs between career stages
-As shown in \autoref{simulation}, we predict that $\alpha$ is expected to be non-zero, even if collaborators are randomly selected with respect to gender, provided that there is a gender gap between career stages.
-The extent to which $\alpha$ deviates from zero depends on the relative frequencies of collaboration within and between career stages.
-When >50% of collaborations were between early and established researchers, we expect gender heterophily ($\alpha < 0$).
-Conversely, when >50% of collaborations occured within career stages, we expect gender homophily ($\alpha > 0$).
-In a few parameter spaces (shown in red; \autoref{simulation}), $\alpha$ was quite high, and overlapped with the values that we estimated (\autoref{alpha_histograms}).
-
-![When there is a difference in gender ratio between early-career and established researchers, and collaboration is non-random with respect to career stage, the null expectation for $\alpha$ deviates from zero.
-An excess of collaborations between career stages gives the appearance of gender heterophily (lower rows, blue areas), while an excess of within-career stage collaborations produced apparent gender homophily (upper rows, red areas).
-However, the conditions required for strong gender homophily are quite restrictive, making it unlikely that this issue explains all of the homophily observed in Figure 2.
-Contour lines denote increments of 0.1.
-\label{simulation}](../figures/figure5 - Simulation.pdf)
-
-Despite this overlap, \autoref{simulation} suggests that our main conclusions (and those of other studies of gender homophily) are probably robust to this career stage issue.
-We only expect strongly positive $\alpha$ when A) the gender ratio is highly skewed across career stages (e.g.
-a 5-fold difference), and B) collaborations between early and established researchers are very rare (e.g.
-<10% of the total).
-Both of these conditions are untrue for most fields: the gender gap across careers stages is generally less pronounced [e.g.
-@Shaw_2012; @Holman_2018], and it is very common for early-career researchers to co-publish with an established mentor [@macaluso_2016].
-However, one can get $\alpha > 0$ for realistic combinations of parameters, e.g.
-a moderate shortage of women in senior positions coupled with a moderate excess of within-career stage collaboration, suggesting this effect might contribute to some of the observed homophily.
-
-Lastly, we note that if there is a gender gap between career stages and coauthorships between early-career and established researchers comprise >50% of the total, then the baseline expectation for $\alpha$ is actually less than zero (blue areas in \autoref{simulation}).
-Therefore, our results might under-estimate the extent to which researchers preferentially select same-gendered collaborators in some cases.
-
-# Discussion
-We found evidence that researchers preferentially publish with same-gendered coauthors, even after implementing stringent controls for Wahlund effects (\autoref{wahlund_plot}).
-Our study therefore reaffirms earlier studies' conclusions [e.g.
-@Ferber_1980; @McDowell_1992; @Bentley_2003; @Crow_2015; @Ghiasi_2015; @Araujo_2017a; @Fahmy_2017; @Jadidi_2017; @Teele_2017; @Zettler_2017] and establishes their generality across the life sciences.
-Relatively few journals had $\alpha'$ values below zero, and almost no journals showed statistically significant gender heterophily after controlling for multiple testing.
-The excess of same-gender coauthorships was quite large: many journals had $\alpha' > 0.1$, indicating that the gender ratio of men's and women's coauthors differs by >10% in absolute terms.
-In relative terms, our findings are even more striking: for example, if men have 20% female coauthors and women have 30% (i.e.
-$\alpha' = 0.1$ for a field with a male-biased gender ratio), then women publish with women 50% more often than men do.
-
-An important limitation of our study is that we cannot reliably determine the cause(s) of the observed excess of same-gender coauthorships.
-As well as conscious or unconcious selection of collaborators based on gender, our results could be partly explained by uncontrolled Wahlund effects.
-However, we suspect the contribution of these to be minor, for four reasons: we found positive $\alpha'$ after controlling for three obvious sources of Wahlund effect; there was no inflation of $\alpha'$ in highly multidisciplinary journals; restricting the data by country yielded similar estimates of $\alpha'$; and we showed that differences in gender ratio between career stages are unlikely to fully explain our results.
-On balance, we believe the data provide good evidence that researchers preferentially select same-gendered collaborators, although the strength of this preference is uncertain.
-
-We hypothesised that disciplines with a strongly skewed gender ratio might show the strongest gender homophily, e.g.
-because being in the minority might increase motivation to seek out same-gendered colleagues.
-Contrary to this hypothesis, we found no evidence that gender homophily is restricted to particular disciplines: $\alpha'$ was similarly high across the board (\autoref{alpha_histograms}).
-Moreover, $\alpha'$ was lower for journals with a male-biased author gender ratio relative to those with an even gender ratio.
-Interestingly, $\alpha'$ was highest in journals with a female-biased author gender ratio.
-This may suggest that men are more likely to preferentially seek out male collaborators in fields where men are a minority, relative to the homophily displayed by women in fields where women are a minority.
-However, this latter result is only tentatively supported since there are few journals in which most authors are women (\autoref{alpha_gender_ratio}).
-
-We also found that gender homophily was marginally stronger in 2015-2016 relative to 2005-2006.
-Although this trend might reflect a change in the gender preferences of researchers seeking collaborators over time, there are alternative (and perhaps more likely) explanations.
-For example, this trend might result from the increasing number of women working in senior positions in STEMM over the past decade [e.g.
-@Shaw_2012; @Long_2015; @McKenzie_2017; @Bendels_2018].
-As shown in \autoref{simulation}, if enough coauthorships are between junior and senior researchers, a large gender gap between career stages can give the appearance of heterophily.
-As this gender gap between career stages lessens, the observed values of $\alpha$ may increase.
-
-Our study begs two questions: what causes gender homophily in science, and are our results cause for concern?
-These questions are closely related.
-For example, some of the homophily we observed might be caused by women seeking to avoid harassment or sexism from men [@jagsi_2016], which would clearly be concerning.
-Additionally, Sheltzer and Smith [-@sheltzer_2014] concluded that 'elite' male academics (defined as recipients of major honours) have a higher proportion of male students and postdocs than non-elite male academics.
-This finding could contribute to the homophily we observed, and is cause for concern since Sheltzer and Smith [-@sheltzer_2014]'s results might reflect discrimination against women during hiring [@Moss_2012], or avoidance by women of elite research groups (e.g.
-due to 'imposter syndrome' or the perception that these groups are sexist).
-We also found a little evidence that gender homophily is detrimental to research quality, in that high-impact journals tended to have weaker homophily.
-Assuming that papers published in high-impact journals are of higher average quality [which is contentious; @Garfield_2006], this result provides non-experimental support for the hypothesis that mixed-gender teams produce better research than single-gender teams [@reagans2001; @Hong_2004; @Bear_2011; @Campbell_2013].
-
-Homophily might also have more benign causes.
-Collaboration is often most enjoyable and productive when working with like-minded people, who might be same-gendered more often than not.
-We also suppose that some people consciously choose to preferentially collaborate with women in order to help close the gender gap in the workforce; this would create homophily if women do this more than men.
-In support of this interpretation, women appear more likely than men to promote the work of female collagues by inviting them to give talks [@nittrouer_2018; @debarre_2018].
-Related to this, Ghiasi et al.
-[-@Ghiasi_2015] concluded that women in engineering are "compliant [in reproducing] male-dominated scientific structures" because they do not collaborate often enough with other women; their data suggest that coauthorships between two women are about 30% more frequent than expected under the null (see their Figure 7).
-In contrast, we propose that men as well as women should ensure that they are not inadvertently overlooking or excluding female colleagues, particularly since men are disproportionately represented among senior researchers [@Shaw_2012].
-
-----------
-
-Why does gender homophily exist in science?
-Collaboration often leads to co-authorship, publications, citations and additional professional recognition, therefore first authors should always be considering collaborators that are most likely to enhance their productivity and in the long term also their academic success, rather than pick co-authors according to their gender.
-First of all, it is likely that decades ago, the probability to meet a female colleague working on the same area of expertise was slimmer than now.
-Even though the number of women in STEMM is increasing, i.e.
-it is not likely to be an obstacle to collaboration anymore; we cannot forgo that female representation in science is still not equal to that of males.
-Second, the most basic source of homophily is space, scientists are more likely to collaborate with colleagues closer to them in geographic location (@McPherson_2011).
-In the case of women, personal considerations (e.g.
-family commitments) represent the principal restrictions on their international mobility.
-As a result women are less likely to originate collaborations with foreign colleagues [@Lemoine_1992; @Lewison_2001; @Webster_2001; @Lariviere_2011].
-Lastly, men and women think differently on average (REF) and are more likely to initiate collaborations based on similarity of thoughts to progress their scientific output (REF).
-
-Academic publishing is one of the most (if not the most) important tool for a researcher to display their work and is frequently used as an indicator of success for being promoted and hired as a faculty member.
-Our finding that researchers are more likely to publish with same-gendered co-authors implies that co-authorship listing is not a random process, but instead creates additional gender imbalance.
-Homophily could make it harder for women to find collaborators in disciplines where women are in the minority [i.e.
-almost all STEMM fields; @Holman_2018].
-Because women make up for as little as 30% of the STEMM scientist community [@Holman_2018], they are less likely to be picked as coauthors, and, as a result, gender homophily creates further gender imbalance and disadvantages for women.
-Men tend to publish more in high impact journals as first author [@Holman_2018], and generate more citations than papers lead by first author women [@Long_1992; @Lariviere_2013].
-As a consequence, gender homophily is also likely to contribute to lower publication rates for women [@McDowell_1992].
-
-Although the proportion of women in science has significantly increased over the last decades, women still tend to leave their academic careers behind considerably more often than their male counterparts [REF].
-Lack of scientific recognition [@Lincoln_2012], lack of research funding [@Bozeman_2004], lack of invitation to speak at scientific events [@Isbell_2012; @Klein_2017; @debarre_2018], discrimination during hiring [@MossRacusin_2017; @Reuben_2014] and discrimination when writing a recommendation letter [@Trix_2003; @Schmader_2007], are only some of the small disadvantages adding up to drive women out of science.
-Whether their decision to leave academia is deliberate or not, it will negatively impact science overall, as mentoring, networking, and supporting younger women scientists are especially important for women in academia.
-
-Altogether, our findings suggest that researchers preferably collaborate with same-gendered colleagues in STEMM research.
-Despites using a more stringent analysis method (e.g.
-Wahlund effects) as well as the largest dataset to date, our results were in accord with longstanding and recent research, while adding new details, reasoning and insight for the current situation.
-Gender homophily in science not only reinforces the less favorable status of women in science, but also creates long term problems as described above.
-It is paramount to explore the causes and implement measures to readdress the gender disparity.
-We advocate for the introduction and implementation of gender policies in order to support opposite sex collaboration, to give men and women a chance to collaborate across gender, fields, and to promote scientific and innovative excellence.
-
 # Acknowledgements
-We are grateful to XXXX. CM was supported by the Academy of Finland (284666 to the Centre of Excellence in Biological Interactions).
+CM was supported by the Academy of Finland (284666 to the Centre of Excellence in Biological Interactions).
 
@@
 <div id="refs"></div>
+
+\newpage
+# Supporting information
+S1 Fig.
+There is a very strong correlation between the values of $\alpha$ and $\alpha'$ calculated for each journal, though in a handful of cases the difference is considerable.
+The deviation between $\alpha$ and $\alpha'$ is greatest for journals for which there is a small sample size (see S2 Fig).
+
+S2 Fig.
+For journals for which we recovered a small number of papers (<100), the unadjusted metric $\alpha$ was downwardly biased.
+This fits our expectations: because authors cannot be their own co-authors, small datasets will tend to produce negative estimates of $\alpha$ even if authors assort randomly with respect to gender (see main text).
+This suggests that $\alpha'$ is a more useful measure of homophily and heterophily, especially for small samples.
+
+S3 Fig.
+Histogram showing the distribution of differences in $\alpha'$ between the 2015-16 and 2005-6 samples, where positive numbers indicate an increase in $\alpha'$ with time.
+The mean is slightly positive (namely 0.004), indicating a mild increase in average $\alpha'$ with time.
+
+S4 Fig.
+Histogram showing the difference between $\alpha'$ calculated for first and last authors.
+Positive values mean that $\alpha'$ was higher when calculated for first authors, and negative values mean $\alpha'$ was higher when calculated for last authors.
+The mean is very slightly higher than zero, indicating that $\alpha'$ tends to be higher for first authors.
+
+S5 Fig.
+Histogram of $\alpha'$ for 325 unique combinations of journal and country, using data from August 2015 - August 2016.
+The white areas denote combinations for which $\alpha'$ differs significantly from zero (p < 0.05, following false discovery rate correction).
+
+S6 Fig. Plot showing the 68 combinations of journal and author country of affiliation for which $\alpha'$ is significantly higher than expected.
+
+S7 Fig.
+Histogram showing the estimated degree to which $\alpha'$ is inflated by inter-country differences in author gender ratio, across the 285 journals for which we had adequate data after restricting the analysis by country.
+The average inflation in $\alpha'$ is negligible, suggesting that Wahlund effects resulting from inter-country differences have a neglible effect on our estimates of gender homophily.
+
+S1 Table.
+The $\alpha'$ values calculated for each journal, in the 2005 and 2015 samples, for each type of author (i.e.
+all authors, first authors, and last authors).
+The tables gives the impact factor of each journal, the sample size, $\alpha$ and $\alpha'$ and their 95% CIs, and the p-value from a 2-tailed test evaluating the null hypothesis that $\alpha$ is zero (both raw and FDR-corrected p-values are shown).
+
+S2 Table.
+Number of journals showing statistically significant homophily or heterophily, in two one-year periods.
+The significance threshold was p < 0.05, and p-values were adjused using Benjamini-Hochberg false discovery rate correction.
+Note that the power of our test is lower for the 2005-2006 data because fewer papers were recovered per journal: thus, it is not meaningful to compare the % significant journals (i.e.
+11% vs 24%) between the two time periods.
+
+S3 Table.
+The table shows similar information to S3 Fig, namely the distribution of the $\alpha'$ values across journals, split by the research discipline.
+The gender ratio column shows the percentage of women authors in the sample used to calculate $\alpha'$, across all authorship positions.
+In the last two columns, the numbers outside parentheses give the number of journals that deviate statistically significantly from zero, while the numbers inside parentheses give the number that remain significant after false discovery rate correction.

### 2018-06-03 Gender_homophily_analysis.Rmd
Plos references

@@
               colour = "tomato", fill = "skyblue") + 
-  xlab("% women authors (realtive to gender parity)") + 
+  xlab("% women authors (relative to gender parity)") + 
   ylab("Coefficient of homophily (\u03B1')") +

### 2018-06-03 manuscript/Manuscript_Claire.Rmd
Plos references

@@
   \usepackage{amssymb} \usepackage{rotating} \usepackage{lineno} \linenumbers
-csl: the-american-naturalist.csl
+csl: plos.csl
 abstract: Evidence suggests that women in academia are hindered by conscious and unconscious
@@
 \autoref{alpha_histograms} shows the distribution of $\alpha'$ estimates in 2015-2016 across all journals for which we recovered sufficient data, when $\alpha'$ was calculated for all authors, first authors only, or last authors only.
-The great majority of journals had $\alpha' > 0$, and for many of these the FDR-corrected p-values suggested that $\alpha'$ was significantly greater than zero (1469/2077 journals were significant in 2015-16, and 404/1192 in 2005-6).
+The great majority of journals had $\alpha' > 0$, and for many of these the FDR-corrected p-values suggested that $\alpha'$ was significantly greater than zero (1469/2077 journals were significant in 2015-16, and 404/1192 in 2005-6; S1 Data).
 Only 2/2077 journals had statistically significantly heterophily (i.e.
@@
 
-S1 Table.
-The $\alpha'$ values calculated for each journal, in the 2005 and 2015 samples, for each type of author (i.e.
-all authors, first authors, and last authors).
-The tables gives the impact factor of each journal, the sample size, $\alpha$ and $\alpha'$ and their 95% CIs, and the p-value from a 2-tailed test evaluating the null hypothesis that $\alpha$ is zero (both raw and FDR-corrected p-values are shown).
+S1 Table. Sample sizes for the two datasets, which comprise papers published in the timeframes August 2005 - August 2006, and August 2015 - August 2016.
 
@@
 In the last two columns, the numbers outside parentheses give the number of journals that deviate statistically significantly from zero, while the numbers inside parentheses give the number that remain significant after false discovery rate correction.
+
+S1 Data: This spreadsheet shows the $\alpha$ values calculated for each journal, in the 2005 and 2015 samples, and for each type of author (all authors, first authors, and last authors).
+The tables gives the impact factor of each journal, the sample size, $\alpha$ and $\alpha'$ and their 95% CIs, and the p-value from a 2-tailed test evaluating the null hypothesis that $\alpha$ is zero (both raw and FDR-corrected p-values are shown).

### 2018-06-05 Gender_homophily_analysis.Rmd
Many changes after Devi comments

@@
 library(ggbeeswarm) # beeswarm plots
+library(brms) # Bayesian models for meta-analysis
 library(Cairo) # for exporting the character 'alpha' via ggsave
@@
 opts_chunk$set(warning=FALSE, message=FALSE) # Suppress warnings in the HTML output
-
-figure.counter <- 2  # Counters for numbering figures and tables
-supp.figure.counter <- 1
-table.counter <- 1
-supp.table.counter <- 1
 ```
@@
                                      end.of.time.period, 
+                                     restrict.by.n.authors = NULL,
                                      file.name, 
@@
       # Genders were only recorded as M or F if the given name is associated with one gender >=95% of the time.
+
       focal <- pubmed_sqlite %>% 
@@
         filter(nchar(gender.95) > 1)       # only keep papers with >1 author
+
+      # Potentially further restrict to 2-author papers, 3-author papers... or 5 or more author papers
+      if(!is.null(restrict.by.n.authors)) {
+        if(restrict.by.n.authors != 5) focal <- focal %>% filter(nchar(gender.95) == restrict.by.n.authors)
+        else if(restrict.by.n.authors == 5) focal <- focal %>% filter(nchar(gender.95) >= 5)
+      }
 
@@
 ### Define a function to calculate Bergstrom et al.'s "coefficient of homophily", $\alpha$
+
 These two functions estimate $\alpha$ for a set of papers' author lists.
@@
 
-**Table S`r supp.table.counter`**: Sample sizes for the two datasets, which comprise papers published in the timeframes August 2005 - August 2006, and August 2015 - August 2016.
-
-```{r echo = FALSE, results = "hide"}
-supp.table.counter <- supp.table.counter + 1
-```
+**Table S1**: Sample sizes for the two datasets, which comprise papers published in the timeframes August 2005 - August 2006, and August 2015 - August 2016.
 
@@
                          file.name = "data/older.papers.rds", 
+                         over.write = FALSE)
+
+# Get all the data we need out of the database, and save as an RDS file - for just the 2-author papers in the last 12 months
+retrieve.authorship.data(unique.journals, 
+                         minimum.n = minimum.n, 
+                         start.of.time.period = 1, 
+                         end.of.time.period = 0, 
+                         restrict.by.n.authors = 2,
+                         file.name = "data/last.year_2author.rds", 
+                         over.write = FALSE)
+
+# Get all the data we need out of the database, and save as an RDS file - for just the 3-author papers in the last 12 months
+retrieve.authorship.data(unique.journals, 
+                         minimum.n = minimum.n, 
+                         start.of.time.period = 1, 
+                         end.of.time.period = 0, 
+                         restrict.by.n.authors = 3,
+                         file.name = "data/last.year_3author.rds", 
+                         over.write = FALSE)
+
+# Get all the data we need out of the database, and save as an RDS file - for just the 4-author papers in the last 12 months
+retrieve.authorship.data(unique.journals, 
+                         minimum.n = minimum.n, 
+                         start.of.time.period = 1, 
+                         end.of.time.period = 0, 
+                         restrict.by.n.authors = 4,
+                         file.name = "data/last.year_4author.rds", 
+                         over.write = FALSE)
+
+# Get all the data we need out of the database, and save as an RDS file - for just the 5-plus-author papers in the last 12 months
+retrieve.authorship.data(unique.journals, 
+                         minimum.n = minimum.n, 
+                         start.of.time.period = 1, 
+                         end.of.time.period = 0, 
+                         restrict.by.n.authors = 5, # NB this actually means "5 or more"
+                         file.name = "data/last.year_5author.rds", 
                          over.write = FALSE)
@@
 
+# Calculate alpha across all authorship positions for all suitable NEW papers with TWO authors
+gender.grouping.test.many.journals(readRDS("data/last.year_2author.rds"),
+                                   "all",
+                                   "data/homophily.results.2author.csv",
+                                   nBoots = 1000, 
+                                   nSimulations = 1000, 
+                                   number.of.cores = 8,
+                                   over.write = FALSE)
+
+# Calculate alpha across all authorship positions for all suitable NEW papers with THREE authors
+gender.grouping.test.many.journals(readRDS("data/last.year_3author.rds"),
+                                   "all",
+                                   "data/homophily.results.3author.csv",
+                                   nBoots = 1000, 
+                                   nSimulations = 1000, 
+                                   number.of.cores = 8,
+                                   over.write = FALSE)
+
+# Calculate alpha across all authorship positions for all suitable NEW papers with FOUR authors
+gender.grouping.test.many.journals(readRDS("data/last.year_4author.rds"),
+                                   "all",
+                                   "data/homophily.results.4author.csv",
+                                   nBoots = 1000, 
+                                   nSimulations = 1000, 
+                                   number.of.cores = 8,
+                                   over.write = FALSE)
+
+# Calculate alpha across all authorship positions for all suitable NEW papers with FIVE authors
+gender.grouping.test.many.journals(readRDS("data/last.year_5author.rds"),
+                                   "all",
+                                   "data/homophily.results.5author.csv",
+                                   nBoots = 1000, 
+                                   nSimulations = 1000, 
+                                   number.of.cores = 8,
+                                   over.write = FALSE)
 ```
@@
 ```{r}
-master.dataset %>%
+fig_s7 <- master.dataset %>%
   ggplot(aes(x = observed.alpha, y = alpha.adjusted)) + 
@@
   theme_minimal()
+ggsave(fig_s7, file = "figures/S7 Fig.pdf", height = 5, width = 5, device = cairo_pdf)
+fig_s7
 ```
 <br></br> 
-**Figure S`r supp.figure.counter`**: There is a very strong correlation between the values of $\alpha$ and $\alpha'$ calculated for each journal, though in a handful of cases the difference is considerable.
-The deviation between $\alpha$ and $\alpha'$ is greatest for journals for which we have a small sample size (see next figure).
-
-```{r echo=FALSE, results='hide'}
-supp.figure.counter <- supp.figure.counter + 1
-```
+**Figure S7**: There is a very strong correlation between the values of $\alpha$ and $\alpha'$ calculated for each journal, though in a handful of cases the difference is considerable.
+The deviation between $\alpha$ and $\alpha'$ is greatest for journals for which there is a small sample size (see S8 Fig).
 
@@
 ```{r}
-master.dataset %>%
+fig_s8 <- master.dataset %>%
   filter(Age == "New") %>%
@@
   ylab("\u03B1' - \u03B1") + 
-  xlab("Log10 sample size\n(number of papers)") + 
+  xlab("Log10 sample size (number of papers)") + 
   theme_minimal()
+ggsave(fig_s8, file = "figures/S8 Fig.pdf", height = 5, width = 5, device = cairo_pdf)
 ```
 <br></br> 
-**Figure S`r supp.figure.counter`**: For journals for which we recovered a small number of papers (<100), the unadjusted metric $\alpha$ was downwardly biased.
+**Figure S8**: For journals for which we recovered a small number of papers (<100), the unadjusted metric $\alpha$ was downwardly biased.
 This fits our expectations: because authors cannot be their own co-authors, small datasets will tend to produce negative estimates of $\alpha$ even if authors assort randomly with respect to gender (see main text).
@@
 
-```{r echo=FALSE, results='hide'}
-supp.figure.counter <- supp.figure.counter + 1
-```
-
 ## Results
@@
 
-**S1 Data**: This spreadsheet shows the $\alpha'$ values calculated for each journal, in the 2005 and 2015 samples, and for each type of author (all authors, first authors, and last authors).
+**S1 Data**: This file gives the $\alpha'$ values calculated for each journal, in the 2005 and 2015 samples, and for each type of author (all authors, first authors, and last authors).
 The tables gives the impact factor of each journal, the sample size, $\alpha$ and $\alpha'$ and their 95% CIs, and the p-value from a 2-tailed test evaluating the null hypothesis that $\alpha$ is zero (both raw and FDR-corrected p-values are shown).
@@
 for(i in which(sapply(neat, is.numeric))) neat[,i] <- round(neat[,i], 3)
-write.csv(neat, file = "data/S1 data.csv", row.names = FALSE)
+write.csv(neat, file = "supplement/S1 data.csv", row.names = FALSE)
 rm(neat)
 ```
+
+### Create S2 Data, showing the number of papers with $n$ authors per journal
+
+**S2 Data**: This file gives the number and percentage of paper that have 1, 2, 3, 4, or ${\ge}5$ authors for each journal in the dataset of Holman et al.
+(2018) _PLoS Biology_.
+Note that the sample sizes include papers for which the gender of one or more authors was not determined by Holman et al.
+
+```{r}
+if(!file.exists("supplement/S2 data.csv")){
+  freq_multiauthor_by_journal <- pubmed_sqlite %>% 
+  select(journal, gender.95) %>% 
+  collect(n=Inf) %>% 
+  group_by(journal) %>% 
+  summarise(n1 = sum(nchar(gender.95) == 1), # Count the number of n author papers
+            n2 = sum(nchar(gender.95) == 2), 
+            n3 = sum(nchar(gender.95) == 3), 
+            n4 = sum(nchar(gender.95) == 4), 
+            `n5+` = sum(nchar(gender.95) >= 5)) %>%
+  gather(author_number, count, n1, n2, n3, n4, `n5+`) %>%
+  mutate(author_number = gsub("n", "", author_number)) %>%
+  arrange(journal, author_number) %>% 
+  left_join(journals_sqlite %>% 
+              select(short.title, Discipline) %>% 
+              collect(), by = c("journal" = "short.title")) %>%
+  distinct() %>% 
+  split(.$journal) %>% # Add a % column
+  purrr::map(function(x) {
+    x$percent <- 100 * x$count/sum(x$count)
+    x}) %>% 
+  do.call("rbind", .) %>%
+  mutate(percent = round(percent, 1))
+write.csv(freq_multiauthor_by_journal, file = "supplement/S2 data.csv", row.names = FALSE)
+}
+```
+
+### Create S3 Data, showing the number of papers with $n$ authors per discipline
+
+**S3 Data**: This file gives the number and percentage of paper that have 1, 2, 3, 4, or ${\ge}5$ authors for each discipline in the dataset of Holman et al.
+(2018) _PLoS Biology_.
+Note that the sample sizes include papers for which the gender of one or more authors was not determined by Holman et al.
+
+```{r}
+freq_multiauthor_by_discipline <- read.csv("supplement/S2 data.csv", stringsAsFactors = FALSE) %>%
+  group_by(Discipline, author_number) %>%
+  summarise(count = sum(count)) %>% 
+  split(.$Discipline) %>% 
+  purrr::map(function(x) {
+    x$percent <- 100 * x$count/sum(x$count)
+    x}) %>% 
+  do.call("rbind", .) %>%
+  mutate(percent = round(percent, 1))
+write.csv(freq_multiauthor_by_discipline, file = "supplement/S3 data.csv", row.names = FALSE)
+```
+
+### Multi-author papers are common in most disciplines
+
+```{r fig.height=14, fig.width = 11}
+big.plot <- freq_multiauthor_by_discipline %>%
+  ggplot(aes(author_number, percent, group = 1)) + 
+  geom_line() + 
+  scale_y_continuous(breaks = c(0, 40, 80)) +
+  facet_wrap(~Discipline, ncol = 8, labeller = label_wrap_gen(width = 25)) +
+  theme_bw() + 
+  theme(strip.text = element_text(size = 7.2), panel.grid.major.x = element_blank()) + 
+  xlab("Number of authors") + 
+  ylab("% of papers")
+
+ggsave(big.plot, file = "figures/S1 Fig.pdf", height = 11, width = 10)
+big.plot
+```
+<br></br>
+**Figure S1**: Plot showing the percentage of papers that have 1, 2, 3, 4, or ${\ge}5$ authors for each discipline in the dataset of Holman et al.
+(2018).
+This information can also be found in S3 Data.
 
@@
 
-**Table S`r supp.table.counter`**: Number of journals showing statistically significant homophily or heterophily, in two one-year periods.
+**Table S2**: Number of journals showing statistically significant homophily or heterophily, in two one-year periods.
 The significance threshold was p < 0.05, and p-values were adjused using Benjamini-Hochberg false discovery rate correction.
@@
 11% vs 24%) between the two time periods.
-
-```{r echo = FALSE, results = "hide"}
-supp.table.counter <- supp.table.counter + 1
-```
 
@@
 
-old.vs.new %>%
+fig_s2 <- old.vs.new %>%
   ggplot(aes(x = alpha.new - alpha.old, fill = Significant)) + 
@@
   theme_minimal()
+ggsave(fig_s2, file = "figures/S2 Fig.pdf", width = 5, height = 5, device = cairo_pdf)
 ```
 <br></br>
-**Figure S`r supp.figure.counter`**: Histogram showing the distribution of differences in $\alpha'$ between the 2015-16 and 2005-6 samples, where positive numbers indicate an increase in $\alpha'$ with time.
-The mean is slightly positive (namely `r with(old.vs.new, mean(alpha.new - alpha.old))`), indicating a mild increase in average $\alpha'$ with time.
-
-```{r echo=FALSE, results='hide'}
-supp.figure.counter <- supp.figure.counter + 1
-```
+**Figure S2**: Histogram showing the distribution of differences in $\alpha'$ between the 2015-16 and 2005-6 samples, where positive numbers indicate an increase in $\alpha'$ with time.
+The mean is slightly positive (i.e.
+0.004), indicating a mild increase in average $\alpha'$ with time.
 
@@
 
+ggsave(first.vs.last.plot, file = "figures/S3 Fig.pdf", width = 5, height = 5, device = cairo_pdf)
 first.vs.last.plot
@@
 <br></br>
-**Figure S`r supp.figure.counter`**: Histogram showing the difference between $\alpha'$ calculated for first and last authors.
+**Figure S3**: Histogram showing the difference between $\alpha'$ calculated for first and last authors.
 Positive values mean that $\alpha'$ was higher when calculated for first authors, and negative values mean $\alpha'$ was higher when calculated for last authors.
-The mean is very slightly higher than zero, indicated that $\alpha'$ tends to be higher for first authors.
-
-```{r echo=FALSE, results='hide'}
-supp.figure.counter <- supp.figure.counter + 1
-```
+The mean is very slightly higher than zero, indicating that $\alpha'$ tends to be higher for first authors.
 
@@
 
-ggsave(grid.arrange(figure2A + xlab(NULL), figure2B, nrow = 1, bottom = "Coefficient of homophily (\u03B1')"), filename = "figures/figure2.pdf", height = 8, width = 7, device = cairo_pdf)
+ggsave(grid.arrange(figure2A + xlab(NULL), figure2B, nrow = 1, bottom = "Coefficient of homophily (\u03B1')"), filename = "figures/Fig2.pdf", height = 8, width = 7, device = cairo_pdf)
 ```
 <br></br>
-**Figure `r figure.counter`**: Of the `r n.journals` journals for which we had adequate data in 2015-2016, `r n.sig.homo` showed statistically significant evidence of homophily (denoted by $\alpha' > 0$), and `r n.sig.hetero` showed statistically significant evidence of heterophily ($\alpha' < 0$), after adjusting p-values using Benjamini-Hochberg false discovery rate correction.
+**Figure 2**: Of the `r n.journals` journals for which we had adequate data in 2015-2016, `r n.sig.homo` showed statistically significant evidence of homophily (denoted by $\alpha' > 0$), and `r n.sig.hetero` showed statistically significant evidence of heterophily ($\alpha' < 0$), after adjusting p-values using Benjamini-Hochberg false discovery rate correction.
 In the left panels, the white area shows the number of journals for which homophily was significantly stronger than expected under the null hypothesis (p < 0.05), while the blue area shows the remainder.
@@
 
-```{r echo=FALSE, results='hide'}
-figure.counter <- figure.counter + 1
-```
-
 #### Table showing the difference in $\alpha$ between disciplines
 
-**Table S`r supp.table.counter`**: The table shows similar information to Figure S`r figure.counter`, namely the distribution of the $\alpha'$ values across journals, split by the research discipline.
+**S4 Data**: The table shows the distribution of the $\alpha'$ values across journals, split by the research discipline.
 The gender ratio column shows the percentage of women authors in the sample used to calculate $\alpha'$, across all authorship positions.
@@
 
-```{r echo=FALSE, results='hide'}
-supp.table.counter <- supp.table.counter + 1
-```
-
-```{r}
-disc.summary %>% filter(`n journals` > 9) %>% 
-  pander(split.cell = 40, split.table = Inf)
-```
-
-### Gender homophily covaries with author gender ratio
+```{r}
+s4_data <- disc.summary %>% filter(`n journals` > 9)
+write.csv(s4_data, file = "supplement/S4 data.csv", row.names = FALSE)
+
+s4_data %>% pander(split.cell = 40, split.table = Inf)
+```
+
+### Relationship between gender homophily and number of authors
+
+```{r}
+# Load and tidy the alpha values calculated using only papers with 2,3,4, or 5+ authors
+alpha_by_author_number <- lapply(2:5, function(i){
+  read.csv(paste("data/homophily.results.", i, "author.csv", sep = ""), 
+           stringsAsFactors = F) %>% mutate(authors = i)}) %>%
+  do.call("rbind", .) %>% 
+  select(journal, authors, alpha.adjusted, n.useable.papers, lower.95.CI2, upper.95.CI2) %>% 
+  mutate(SE = (upper.95.CI2 - alpha.adjusted)/1.96) %>% 
+  left_join(journals_sqlite %>% 
+              select(short.title, Discipline, IF) %>% 
+              collect(), 
+            by = c("journal" = "short.title")) %>% 
+  arrange(journal, authors, Discipline) %>%
+  mutate(authors = factor(replace(authors, authors == 5, "5+"), 
+                          levels = c("2", "3", "4", "5+")))
+
+if(!file.exists("data/brms.output.txt")){
+  # Run a meta-regression model using brms, which runs a Bayesian model using Stan
+meta_analysis.brms <- brm(alpha.adjusted | se(SE) ~ authors + (authors | journal), 
+                 data = alpha_by_author_number, 
+                 iter = 6000, warmup = 4000, chains = 4, cores = 4, seed = 1) %>% 
+  add_ic(ic = "R2") # The Bayesian r-squared is 0.44
+writeLines(capture.output(meta_analysis.brms), con = "data/brms.output.txt")
+
+# Get the predicted values for each group mean
+predictions <- meta.brms %>% 
+  fitted(newdata = data.frame(authors=levels(alpha_by_author_number$authors), 
+                               SE = mean(alpha_by_author_number$SE)), 
+          re_formula = NA) %>% 
+  as.data.frame() %>% 
+  mutate(authors=levels(alpha_by_author_number$authors)) %>% 
+  rename(low =  `2.5%ile`, up = `97.5%ile`)
+
+write.csv(predictions, file = "data/brms.predictions.csv", row.names = FALSE)
+}
+
+# Make a plot
+predictions <- read.csv("data/brms.predictions.csv", stringsAsFactors = FALSE)
+author_number_figure <- alpha_by_author_number %>% 
+  mutate(authors = as.factor(authors), Precision = 1/SE) %>%
+  ggplot(aes(authors)) + 
+  geom_violin(alpha=0.5, aes(y=alpha.adjusted, colour = authors)) + 
+  geom_jitter(aes(y=alpha.adjusted, colour = authors), size = 0.7, alpha = 0.5) + 
+  # geom_line(aes(group = journal, y=alpha.adjusted), alpha = 0.1, size=.7) + 
+  geom_errorbar(data = predictions, aes(authors, ymin = low, ymax = up), width = 0.05, size = 0.6) +
+  geom_line(data = predictions, aes(authors, Estimate, group = 1.2), size = 0.6) + 
+  geom_point(data = predictions, aes(authors, Estimate), size = 2, pch = 21, stroke = 1, fill = "white") +
+  scale_color_brewer(palette = "Dark2") + 
+  scale_fill_brewer(palette = "Dark2") + 
+  theme_bw(13) + 
+  theme(legend.position = "none", panel.grid = element_blank()) +
+  xlab("Number of authors listed on the paper") + 
+  ylab("Coefficient of homophily (\u03B1')")
+ggsave(author_number_figure, file = "figures/Fig3.pdf", height = 6, width = 6.5, device = cairo_pdf)
+author_number_figure
+```
+<br></br>
+**Figure 3**: The coefficient of homophily ($\alpha'$) was slightly less positive when calculated for two-author papers only, relative to papers with longer author lists.
+The individual points, whose distribution is summarised by the violin plots, correspond to individual journals.
+The larger white points show the mean for each group (and its 95% CIs), as calculated by a Bayesian meta-regression model accounting for repeated measures of $\alpha'$ within journals, as well as the precision with which $\alpha'$ was estimated.
+
+### Summary output of the `brms` model used to derive the means in Figure 3
+```{r}
+cat(readLines('data/brms.output.txt'), sep = '\n')
+```
+
+### Relationship between gender homophily and gender ratio
 Journals with a male-biased author gender ratio showed less author gender homophily, while journals with an even or female-biased author gender ratio showed more homophily.
@@
   theme(legend.position = "none") 
-ggsave(plot = gender.ratio, filename = "figures/figure3.pdf", 
+ggsave(plot = gender.ratio, filename = "figures/Fig4.pdf", 
        device = cairo_pdf, width=7, height=5)
@@
 <br></br>
-**Figure `r figure.counter`**: There is a weakly positive, non-linear relationship between the gender ratio of authors publishing in a journal, and the coefficient of homophily ($\alpha'$).
+**Figure 4**: There is a weakly positive, non-linear relationship between the gender ratio of authors publishing in a journal, and the coefficient of homophily ($\alpha'$).
 Specifically, journals with 50% women authors or higher tended to have more same-sex coauthorships than did journals with predominantly men authors.
@@
 
-```{r echo=FALSE, results='hide'}
-figure.counter <- figure.counter + 1
-```
-
-Here are the results of three generalised additive models (GAMs) fit to the data in the three panels of Figure `r figure.counter - 1`:
+Here are the results of three generalised additive models (GAMs) fit to the data in the three panels of Figure 4:
 ```{r}
@@
 
-figure4 <- gender.homophily.all %>% 
+figure5 <- gender.homophily.all %>% 
   filter(!is.na(standardised.IF)) %>% 
@@
   theme_minimal()
-ggsave(figure4, file = "figures/figure4.pdf", height = 5, width = 5, device = cairo_pdf)
-saveRDS(R2, "manuscript/R2_for_figure4.RData")
-figure4
+ggsave(figure5, file = "figures/Fig5.pdf", height = 5, width = 5, device = cairo_pdf)
+saveRDS(R2, "manuscript/R2_for_figure5.RData")
+figure5
 ```
 <br></br>
-**Figure `r figure.counter`**: Journal impact factor (expressed relative to the average for the discipline) is negatively correlated with $\alpha'$.
+**Figure 5**: Journal impact factor (expressed relative to the average for the discipline) is negatively correlated with $\alpha'$.
 The relationship is weak ($R^2$ = `r R2`), but the results suggest that journals with strong homophily tend to have lower impact factors than journals with weak homophily in the same discipline.
-
-```{r echo=FALSE, results='hide'}
-figure.counter <- figure.counter + 1
-```
 
@@
   theme(legend.position = "top")
+ggsave(country.histo, file = "figures/S4 Fig.pdf", width = 5, height = 5, device = cairo_pdf)
 country.histo
@@
 <br></br>
-**Figure S`r supp.figure.counter`**: Histogram of $\alpha'$ for 325 unique combinations of journal and country, using data from August 2015 - August 2016.
+**Figure S4**: Histogram of $\alpha'$ for 325 unique combinations of journal and country, using data from August 2015 - August 2016.
 The white areas denote combinations for which $\alpha'$ differs significantly from zero (p < 0.05, following false discovery rate correction).
-
-```{r echo=FALSE, results='hide'}
-supp.figure.counter <- supp.figure.counter + 1
-```
 
@@
 x <- nrow(country.data[country.data$p.adjusted < 0.05, ])
-country.data %>% 
+fig_s5 <- country.data %>% 
   filter(p.adjusted < 0.05) %>%
@@
   theme(legend.position = "none")
+ggsave(fig_s5, file = "figures/S5 Fig.pdf", width = 7, height = 9, device = cairo_pdf)
+fig_s5
 ```
 <br></br>
-**Fig S`r supp.figure.counter`**: Plot showing the `r nrow(country.data[country.data$p.adjusted < 0.05, ])` combinations of journal and author country of affiliation for which $\alpha'$ is significantly higher than expected.
-
-```{r echo=FALSE, results='hide'}
-supp.figure.counter <- supp.figure.counter + 1
-```
+**Figure S5**: Plot showing the 68 combinations of journal and author country of affiliation for which $\alpha'$ is significantly higher than expected.
 
@@
 
-avg <- mean(mean.inflation.due.to.country$mean.inflation.due.to.country)
-
-txt <- data.frame(x = avg + 0.035, y = 56, label = paste("Mean =", round(avg,3)))
-
-mean.inflation.due.to.country %>%
+median <- median(mean.inflation.due.to.country$mean.inflation.due.to.country)
+
+txt <- data.frame(x = median + 0.06, y = 66, label = paste("Median =", round(median,3)))
+
+mean.inflation.due.to.country <- mean.inflation.due.to.country %>%
   ggplot(aes(mean.inflation.due.to.country)) + 
   geom_histogram(binwidth = 0.02, colour = "black", fill = "skyblue") + 
-  geom_vline(xintercept = avg, linetype = 2) + 
+  geom_vline(xintercept = median, linetype = 2) + 
   geom_text(data=txt, aes(x=x, y=y, label=label)) + 
   theme_minimal() + 
-  xlab("Estimated inflation in \u03B1' due to differences in gender ratio between countries") + 
+  xlab("Estimated inflation in \u03B1' due to\ndifferences in gender ratio between countries") + 
   ylab("Frequency")
+
+ggsave(mean.inflation.due.to.country, file = "figures/S6 Fig.pdf", width = 5, height = 5, device = cairo_pdf)
+mean.inflation.due.to.country
 ```
 <br></br>
-**Fig S`r supp.figure.counter`**: Histogram showing the estimated degree to which $\alpha'$ is inflated by inter-country differences in author gender ratio, across the `r length(unique(country.data$journal))` journals for which we had adequate data after restricting the analysis by country.
-The average inflation in $\alpha'$ is negligible, suggesting that Wahlund effects resulting from inter-country differences have neglible effect on our estimates of gender homophily.
-
-```{r echo=FALSE, results='hide'}
-supp.figure.counter <- supp.figure.counter + 1
-```
+**Fig S6**: Histogram showing the estimated degree to which $\alpha'$ is inflated by inter-country differences in author gender ratio, across the 285 journals for which we had adequate data after restricting the analysis by country.
+The average inflation in $\alpha'$ is negligible, suggesting that Wahlund effects resulting from inter-country differences have a neglible effect on our estimates of gender homophily.
 
@@
 simulation.figure
-ggsave(simulation.figure, file = "figures/figure5.pdf", 
+ggsave(simulation.figure, file = "figures/Fig6.pdf", 
        height = 5.6, width = 5, device = cairo_pdf)

### 2018-06-05 manuscript/Manuscript_Claire.Rmd
Many changes after Devi comments

@@
 ---
-title: "Do researchers preferentially collaborate with colleagues of the same gender?"
+title: "Researchers preferentially collaborate with same-gendered colleagues across the life sciences"
 author:
@@
   Contrary to our expectations, we found no evidence that homophily is driven mostly
-  by senior academics, and no evidence that homophily is strongest in fields where
-  women are in the minority.
-Interestingly, journals with a high impact factor for their discipline tended to have comparatively low homophily, as predicted if mixed-gender teams produce better research.
-  We discuss potential causes of gender homophily in science. \vspace{5mm} \par\noindent
+  by senior academics, and no evidence that homophily is stronger in fields where
+  women are in the minority. Interestingly, journals with a high impact factor for their 
+  discipline tended to have comparatively low homophily, as predicted if mixed-gender teams 
+  produce better research. We discuss potential causes of gender homophily in science. 
+  \vspace{5mm} \par\noindent
   \textbf{Keywords:} Gender bias, Homophily, Scientific collaboration, Text mining,
@@
 # Load some little files with numbers to be inserted into the MS
-sample.size <- read.csv("sample.size.table.csv", stringsAsFactors = FALSE)
-sample.size[,2] <- prettyNum(sample.size[,2], big.mark = ",")
-sample.size[,3] <- prettyNum(sample.size[,3], big.mark = ",")
-names(sample.size)[2:3] <- c("n2005", "n2015")
+# sample.size <- read.csv("sample.size.table.csv", stringsAsFactors = FALSE)
+# sample.size[,2] <- prettyNum(sample.size[,2], big.mark = ",")
+# sample.size[,3] <- prettyNum(sample.size[,3], big.mark = ",")
+# names(sample.size)[2:3] <- c("n2005", "n2015")
 
@@
 
-R2 <- readRDS("R2_for_figure4.RData") # For Fig 4
+R2 <- readRDS("R2_for_figure5.RData") # For Fig 4
 ```
@@
 Because publishing, networking and collaboration are instrumental to scientific productivity and academic career advancement [@Lee_2005; @Wuchty_2007; @Abramo_2009; @Lariviere_2015], dozens of studies have tested for gender differences in these areas [e.g.
-@Long_1992; @Bozeman_2011; @Abramo_2013; @Badar_2013; references in Table S1 of @Holman_2018].
-For example, studies have found that women tend to be less involved in international collaboration [@Lewison_2001; @Webster_2001; @Bozeman_2004; @Lariviere_2011; @Abramo_2013], collaborate less within their own university departments [@Webster_2001], have less prestigious collaborations [@Long_1990], and fewer collaborations in total [@Fuchs_2001].
-These gender differences in collaboration practice presumably have multiple causes, which might include implicit and explicit gender bias [@Moss_2012], differential family obligations [@Reskin_1978; @Long_1990; @Wright_2003], gender differences in confidence or self-esteem [@bleidorn2016age], concerns relating to sexual harassment [@jagsi_2016], and unequal access to conferences [@Martin_2014] and travel funds [@Bozeman_2004].
+@Long_1992; @Bozeman_2011; @Abramo_2013; @Badar_2013; studies listed in Table S1 of @Holman_2018].
+For example, studies have concluded that women tend to be less involved in international collaboration [@Lewison_2001; @Webster_2001; @Bozeman_2004; @Lariviere_2011; @Abramo_2013], collaborate less within their own university departments [@Webster_2001], have less prestigious collaborations [@Long_1990], and fewer collaborations in total [@Fuchs_2001].
+These gender differences in collaboration practice presumably have multiple causes, which might include implicit and explicit gender bias [@Moss_2012], differential family obligations [@Reskin_1978; @Long_1990; @Wright_2003], gender differences in confidence or self-esteem [@bleidorn2016age], concerns relating to sexual harassment [@jagsi_2016], and unequal access to conferences [@Martin_2014] or travel funds [@Bozeman_2004].
 
 A high, steadily increasing proportion of research papers is written by more than one author [@West_2013], making collaboration a key predictor of publication output, and thus of career prospects [@Tower_2007; @Jordan_2008].
-Additionally, empirical studies imply that mixed-gender or otherwise diverse teams produce better results on collaborative tasks than less diverse teams [@Britton_2000; @Reagans_2001; @Hong_2004; @Whittington_2008; @Bear_2011; @Herrera_2012; @Campbell_2013].
-For reasons such as these, several studies have examined the author lists of published research articles in order to test for gender differences in collaboration frequency or pattern.
-To our knowledge, all of these studies imply that men co-publish with other men, and women with women, more often than expected if collaborators assort randomly with respect to gender [@Ferber_1980; @McDowell_1992; @Ghiasi_2015; @Crow_2015; @Fahmy_2017; @Zettler_2017; @Jadidi_2017; @Teele_2017; @Araujo_2017a; @Araujo_2017b].
+Additionally, empirical studies imply that mixed-gender or otherwise diverse teams produce better outputs on collaborative tasks than less diverse teams [@Britton_2000; @Reagans_2001; @Hong_2004; @Whittington_2008; @Bear_2011; @Herrera_2012; @Campbell_2013].
+For reasons such as these, multiple studies have examined the author lists of published research articles in order to test for gender differences in collaboration frequency or pattern.
+To our knowledge, all such studies imply that men co-publish with other men, and women with women, more often than expected if collaborators assort randomly with respect to gender [@Ferber_1980; @McDowell_1992; @Ghiasi_2015; @Crow_2015; @Fahmy_2017; @Zettler_2017; @Jadidi_2017; @Teele_2017; @Araujo_2017a; @Araujo_2017b].
 This pattern of assortative publishing is often termed 'gender homophily'.
@@
 The Wahlund effect will tend to inflate the frequency of same-sex coauthorships whenever the data is composed of two or more disconnected subsets of literature with different author gender ratios; these subsets could be research disciplines, older versus newer papers, or papers from authors in different countries.
-\label{wahlund_plot}](../figures/Figure 1 - Wahlund figure.pdf)
+\label{wahlund_plot}](../figures/Fig1.pdf)
 
@@
 The remaining 606 or 787 journals (in 2015 and 2005 respectively) had a value of $\alpha'$ not significantly different from zero, consistent with the null hypothesis of random assortment with respect to gender.
+We also confirmed that the majority of papers had multiple authors, in most journals (S2 Data) and disciplines (S3 Data, S1 Fig).
 
@@
 Patterns were similar whether $\alpha'$ was calculated for all authors, for first authors only, or for last authors only.
-\label{alpha_histograms}](../figures/figure2.pdf)
+\label{alpha_histograms}](../figures/Fig2.pdf)
 
-$\alpha'$ was significantly higher in the literature sample from 2015-16 relative to 2005-6, though the difference in means was small (S1 Fig; Effect of the fixed factor 'Time period' in a linear mixed model of the data for all author positions: Cohen's $d$ = $0.091{\pm}0.04$, $t_{953}$ = 2.42, p = 0.016).
+$\alpha'$ was significantly higher in the literature sample from 2015-16 relative to 2005-6, though the difference in means was small (S2 Fig; Effect of the fixed factor 'Time period' in a linear mixed model of the data for all author positions: Cohen's $d$ = $0.091{\pm}0.04$, $t_{953}$ = 2.42, p = 0.016).
 
-When comparing pairs of $\alpha'$ values estimated for the first and last authors for the same journals, we found that $\alpha'$ tended to be higher for first authors than for last authors (S2 Fig; Effect of the fixed factor 'Authorship position' in a linear mixed model: Cohen's $d$ = $0.065{\pm}0.02$, $t_{2024}$ = 4.28, p < 0.0001).
+When comparing pairs of $\alpha'$ values estimated for the first and last authors for the same journals, we found that $\alpha'$ tended to be higher for first authors than for last authors (S3 Fig; Effect of the fixed factor 'Authorship position' in a linear mixed model: Cohen's $d$ = $0.065{\pm}0.02$, $t_{2024}$ = 4.28, p < 0.0001).
 This suggests that the gender of the first author was a slightly stronger predictor of the remaining authors' genders than the gender of the last author, i.e.
@@
 All disciplines had a positive average $\alpha'$, although homophily appeared somewhat stronger in some disciplines than others (e.g.
-mean $\alpha'$ was $0.12{\pm}0.02$ for Urology journals and $0.03{\pm}0.01$ for Veterinary Medicine journals; \autoref{alpha_histograms}, S3 Table).
+mean $\alpha'$ was $0.12{\pm}0.02$ for Urology journals and $0.03{\pm}0.01$ for Veterinary Medicine journals; \autoref{alpha_histograms}, S4 Data).
 However, there was no evidence for consistent differences in $\alpha'$ between disciplines: the random factor 'Discipline' explained around 1% of the variance in $\alpha'$ in the two linear mixed models described in the previous section (see \autoref{alpha_histograms} and mixed models in Online Supplementary Material).
@@
 This result suggests that our estimates of homophily, and estimates from some earlier studies, are not notably inflated by the presence of disparate research topics (with variable author gender ratios) being published within individual journals.
+
+## Relationship between gender homophily and number of authors
+Papers with two authors had significantly lower (but still positive) $\alpha'$ values than papers with more than two authors, on average (\autoref{author_number}; model output in Online Supplmentary Material).
+Papers with 3, 4 or 5+ authors had essentially identical average $\alpha'$ values.
+The variance in $\alpha'$ across journals was also a little higher for 2-authors papers compared to the remainder (\autoref{author_number}), though part of this variance is due to the reduced sample size (in terms of number of authors) for the 2-author papers.
+
+![The coefficient of homophily ($\alpha'$) was slightly less positive when calculated for two-author papers only, relative to papers with longer author lists.
+The individual points, whose distribution is summarised by the violin plots, correspond to individual journals.
+The larger white points show the mean for each group (and its 95% CIs), as calculated by a Bayesian meta-regression model accounting for repeated measures of $\alpha'$ within journals, as well as the precision with which $\alpha'$ was estimated.
+\label{author_number}](../figures/Fig3.pdf)
 
@@
 The lines were fitted using generalised additive models with the smoothing parameter $k$ set to 3.
-\label{alpha_gender_ratio}](../figures/figure3.pdf)
+\label{alpha_gender_ratio}](../figures/Fig4.pdf)
 
@@
 The relationship is noisy ($R^2$ = `r R2`), but the results suggest that journals with strong homophily tend to have lower impact factors than journals with weak homophily in the same discipline.
-\label{impact_factor}](../figures/figure4.pdf)
+\label{impact_factor}](../figures/Fig5.pdf)
 
 ## Analysis accounting for differences in author gender ratio between countries
-When we restricted the analysis by country, we observed statistically significant homophily for 72 of the 325 journal-country combinations tested (64 unique journals and 18 unique countries), and no significant evidence of heterophily (Figures S3-S4).
+When we restricted the analysis by country, we observed statistically significant homophily for 72 of the 325 journal-country combinations tested (64 unique journals and 18 unique countries), and no significant evidence of heterophily (S4-S5 Fig).
 Additionally, the values of $\alpha'$ calculated for each journal-country combination were mostly very similar to the $\alpha'$ values calculated for the journal as a whole (i.e.
-when pooling papers from different countries); on average, $\alpha'$ was only inflated in the main (non-country specific) analysis by 0.002 (S5 Fig).
+when pooling papers from different countries); on average, $\alpha'$ was only inflated in the main (non-country specific) analysis by _c_.
+0.002 (S6 Fig).
 These results suggest that our findings of widespread homophily in the main analysis were not driven solely by a Wahlund effect resulting from gender differences between countries.
@@
 Contour lines denote increments of 0.1.
-\label{simulation}](../figures/figure5 - Simulation.pdf)
+\label{simulation}](../figures/Fig6_inkscape.pdf)
 
@@
 
+We also calculated $\alpha$ for papers with 2, 3, 4 or ${\ge}5$ authors, for all journals that had at least 50 suitable papers from 2015-2016 with the specified author list length.
+
 Our test assumes that the expected value of $\alpha$ is zero if authors randomly assort, but for small datasets this assumption is not always true (as pointed out by Carl Bergstrom in a blog post, http://www.eigenfactor.org/gender/assortativity/note_to_eisen.rtf).
@@
 
-As expected, $\alpha'$ was usually almost identical to $\alpha$ (S6 Fig), but $\alpha$ was downwardly biased relative to $\alpha'$ for small datasets (S7 Fig).
+As expected, $\alpha'$ was usually almost identical to $\alpha$ (S7 Fig), but $\alpha$ was downwardly biased relative to $\alpha'$ for small datasets (S8 Fig).
 Additionally, the correlation between $\alpha'$ and sample size was negligible ($R^2 < 0.01$), suggesting that our calculation of $\alpha'$ effectively removed the dependence of $\alpha$ on sample size.
@@
 Therefore, a dataset containing a mix of papers from teams of authors based in these two countries would contain an excess of same-sex coauthorships, even if collaboration were random with respect to gender within each country.
+
 To address this issue, we also analysed every combination of journal and author country of affiliation for which we had enough data (i.e.
@@
 
+To model the relationship between $\alpha'$ and the number of authors on the paper, we used a meta-regression model implemented in the R package `brms` [@burkner_2016].
+The model incorporated the standard error associated with easch estimate of $\alpha'$, had author number as a fixed effect, and journal as a random intercept (to control for repeated measures of each journal).
+We also fit a random slope of author number within journal, thereby allowing the response to author number to vary between journals.
+We used the default (weak) priors.
+The full output of this model can be viewed in the Online Supplementary Material.
+
 ## Theoretical expectations for $\alpha$ when the gender ratio differs between career stages
@@
 ## Data availability and reproducibility
-The Online Supplementary Material contains R scripts used to produce all results, figures and tables.
+The Online Supplementary Material contains R scripts used to produce all results, figures and tables; it can be viewed online at https://lukeholman.github.io/genderHomophily/.
 The input data from Holman et al.
-[-@Holman_2018] is archived at https://osf.io/bt9ya/ along with the code used to obtain it.
+[-@Holman_2018] is archived at https://osf.io/bt9ya/.
 
@@
 # Supporting information
-S1 Fig.
-There is a very strong correlation between the values of $\alpha$ and $\alpha'$ calculated for each journal, though in a handful of cases the difference is considerable.
-The deviation between $\alpha$ and $\alpha'$ is greatest for journals for which there is a small sample size (see S2 Fig).
+
+## Supplementary figures
+S1 Fig. Plot showing the percentage of papers that have 1, 2, 3, 4, or ${\ge}5$ authors for each discipline in the dataset of Holman et al. (2018). This information can also be found in S3 Data.
 
 S2 Fig.
+Histogram showing the distribution of differences in $\alpha'$ between the 2015-16 and 2005-6 samples, where positive numbers indicate an increase in $\alpha'$ with time.
+The mean is slightly positive (i.e.
+0.004), indicating a mild increase in average $\alpha'$ with time.
+
+S3 Fig.
+Histogram showing the difference between $\alpha'$ calculated for first and last authors.
+Positive values mean that $\alpha'$ was higher when calculated for first authors, and negative values mean $\alpha'$ was higher when calculated for last authors.
+The mean is very slightly higher than zero, indicating that $\alpha'$ tends to be higher for first authors.
+
+S4 Fig.
+Histogram of $\alpha'$ for 325 unique combinations of journal and country, using data from August 2015 - August 2016.
+The white areas denote combinations for which $\alpha'$ differs significantly from zero (p < 0.05, following false discovery rate correction).
+
+S5 Fig. Plot showing the 68 combinations of journal and author country of affiliation for which $\alpha'$ is significantly higher than expected.
+
+S6 Fig.
+Histogram showing the estimated degree to which $\alpha'$ is inflated by inter-country differences in author gender ratio, across the 285 journals for which we had adequate data after restricting the analysis by country.
+The average inflation in $\alpha'$ is negligible, suggesting that Wahlund effects resulting from inter-country differences have a neglible effect on our estimates of gender homophily.
+
+S7 Fig.
+There is a very strong correlation between the values of $\alpha$ and $\alpha'$ calculated for each journal, though in a handful of cases the difference is considerable.
+The deviation between $\alpha$ and $\alpha'$ is greatest for journals for which there is a small sample size (see S8 Fig).
+
+S8 Fig.
 For journals for which we recovered a small number of papers (<100), the unadjusted metric $\alpha$ was downwardly biased.
@@
 
-S3 Fig.
-Histogram showing the distribution of differences in $\alpha'$ between the 2015-16 and 2005-6 samples, where positive numbers indicate an increase in $\alpha'$ with time.
-The mean is slightly positive (namely 0.004), indicating a mild increase in average $\alpha'$ with time.
-
-S4 Fig.
-Histogram showing the difference between $\alpha'$ calculated for first and last authors.
-Positive values mean that $\alpha'$ was higher when calculated for first authors, and negative values mean $\alpha'$ was higher when calculated for last authors.
-The mean is very slightly higher than zero, indicating that $\alpha'$ tends to be higher for first authors.
-
-S5 Fig.
-Histogram of $\alpha'$ for 325 unique combinations of journal and country, using data from August 2015 - August 2016.
-The white areas denote combinations for which $\alpha'$ differs significantly from zero (p < 0.05, following false discovery rate correction).
-
-S6 Fig. Plot showing the 68 combinations of journal and author country of affiliation for which $\alpha'$ is significantly higher than expected.
-
-S7 Fig.
-Histogram showing the estimated degree to which $\alpha'$ is inflated by inter-country differences in author gender ratio, across the 285 journals for which we had adequate data after restricting the analysis by country.
-The average inflation in $\alpha'$ is negligible, suggesting that Wahlund effects resulting from inter-country differences have a neglible effect on our estimates of gender homophily.
-
+## Supplementary tables
 S1 Table. Sample sizes for the two datasets, which comprise papers published in the timeframes August 2005 - August 2006, and August 2015 - August 2016.
@@
 
-S3 Table.
-The table shows similar information to S3 Fig, namely the distribution of the $\alpha'$ values across journals, split by the research discipline.
+## Supplementary datasets
+S1 Data: This spreadsheet shows the $\alpha$ values calculated for each journal, in the 2005 and 2015 samples, and for each type of author (all authors, first authors, and last authors).
+The tables gives the impact factor of each journal, the sample size, $\alpha$ and $\alpha'$ and their 95% CIs, and the p-value from a 2-tailed test evaluating the null hypothesis that $\alpha$ is zero (both raw and FDR-corrected p-values are shown).
+
+S2 Data: This file gives the number and percentage of paper that have 1, 2, 3, 4, or ${\ge}5$ authors for each _journal_ in the dataset of Holman et al.
+(2018) _PLoS Biology_.
+Note that the sample sizes include papers for which the gender of one or more authors was not determined by Holman et al.
+
+S3 Data: This file gives the number and percentage of paper that have 1, 2, 3, 4, or ${\ge}5$ authors for each _discipline_ in the dataset of Holman et al.
+(2018) _PLoS Biology_.
+Note that the sample sizes include papers for which the gender of one or more authors was not determined by Holman et al.
+
+S4 Data.
+The table shows the distribution of the $\alpha'$ values across journals, split by the research discipline.
 The gender ratio column shows the percentage of women authors in the sample used to calculate $\alpha'$, across all authorship positions.
 In the last two columns, the numbers outside parentheses give the number of journals that deviate statistically significantly from zero, while the numbers inside parentheses give the number that remain significant after false discovery rate correction.
-
-S1 Data: This spreadsheet shows the $\alpha$ values calculated for each journal, in the 2005 and 2015 samples, and for each type of author (all authors, first authors, and last authors).
-The tables gives the impact factor of each journal, the sample size, $\alpha$ and $\alpha'$ and their 95% CIs, and the p-value from a 2-tailed test evaluating the null hypothesis that $\alpha$ is zero (both raw and FDR-corrected p-values are shown).

### 2018-06-06 Gender_homophily_analysis.Rmd
Tweak discussion

@@
 ### Multi-author papers are common in most disciplines
+
+This figure illustrates that single-author papers make up a small minority of the total, for the majority of disciplines we examined.
+Indeed for many disciplines, the majority of papers have at least five authors.
 
@@
 ### Summary output of the `brms` model used to derive the means in Figure 3
+Note that the 95% credible intervals for the population-level intercept do not overlap zero, which indicates "significant" positive $\alpha'$ among 2-author papers.
+Furthermore, $\alpha'$ is significantly higher for 3-, 4-, and 5-plus-author papers than for 2-author papers.
+The R^2 of the model is 0.44, and the random effects appear to explain non-trivial amounts of variation.
 ```{r}

### 2018-06-06 manuscript/Manuscript_Claire.Rmd
Acknowledgements
Tweak discussion

@@
 In relative terms, our findings are even more striking: for example, if men have 20% female coauthors and women have 30% (i.e.
-$\alpha' = 0.1$ for a field with a male-biased gender ratio), then women publish with women 50% more often than men do.
+$\alpha' = 0.1$ in a field with a male-biased gender ratio), then women publish with women 50% more often than men do.
 
 An important limitation of our study is that we cannot reliably determine the cause(s) of the observed excess of same-gender coauthorships.
-As well as conscious or unconcious selection of collaborators based on gender, our results could be partly explained by uncontrolled Wahlund effects.
+As well as the obvious interpretation -- conscious or unconcious selection of same-gender collaborators by men, women, or both -- our results could be partly explained by uncontrolled Wahlund effects.
 However, we suspect the contribution of these to be minor, for four reasons: we found positive $\alpha'$ after controlling for three obvious sources of Wahlund effect; there was no inflation of $\alpha'$ in highly multidisciplinary journals; restricting the data by country yielded similar estimates of $\alpha'$; and we showed that differences in gender ratio between career stages are unlikely to fully explain our results.
-On balance, we believe the data provide good evidence that researchers preferentially select same-gendered collaborators, although the strength of this preference is uncertain.
+On balance, we believe the data suggest that it is likely that some researchers do preferentially select same-gendered collaborators, although the effect size of this preference is difficult to measure.
 
@@
 We also found that gender homophily was marginally stronger in 2015-2016 relative to 2005-2006.
-Although this trend might reflect a change in the gender preferences of researchers seeking collaborators over time, there are alternative (and perhaps more likely) explanations.
+Although this trend might reflect a change in the gender preferences of researchers seeking collaborators, there are alternative (and perhaps more likely) explanations.
 For example, this trend might result from the increasing number of women working in senior positions in STEMM over the past decade [e.g.
@@
 As shown in \autoref{simulation}, if enough coauthorships are between junior and senior researchers, a large gender gap between career stages can give the appearance of heterophily.
-As this gender gap between career stages lessens, the observed values of $\alpha$ may increase.
+As this gender gap between career stages lessens, the observed values of $\alpha'$ may increase.
+
+Regarding our finding of weaker homophily among 2-author papers, we suspect that many 2-author teams comprise a student/postdoc and a senior staff member, making these teams especially likely to be mixed-gender, due to the elevated gender gap among senior researchers [@Shaw_2012; @Holman_2018].
+Assuming this interpretation is correct, this result suggests that our overall $\alpha'$ may underestimate the strength of peoples' preferences for same-gendered collaborators; essentially, women seeking a senior collaborator could be constrained to work mostly with men, meaning that people's ideal and realised gender preferences would be mismatched.
+On a related note, Ghiasi et al.
+[-@Ghiasi_2015] argue that women in engineering are "compliant [in reproducing] male-dominated scientific structures" because they do not collaborate often enough with other women; for reference, their Figure 7 suggests that coauthorships between women are roughly 30% more frequent than expected under random assortment.
+By contrast, we feel that it is not helpful to recommend that women collaborate primarily with other women, e.g.
+because this constrains women's options and may be counter-productive (particularly in fields like engineering, where 90% of professors are men [@Shaw_2012]).
+Instead, we suggest that researchers of both genders can help to close the gender gap in science.
+In the context of collaboration, one way to do this is to undertake self-examination to ensure that one is not inadvertently overlooking or excluding female potential students and colleagues.
+One should also take care to treat male and female collaborators equally, e.g.
+in terms of training and mentoring, allocation of work, and how one frames or promotes the collaboration (e.g.
+in conference presentations or on a website); evidence suggests that unconscious bias causes people to undervalue women's research achievements [@Moss_2012], and to assign menial or under-valued tasks to women and more prestigious and employable tasks (e.g.
+writing and project leadership) to men [@macaluso_2016].
 
@@
 This finding could contribute to the homophily we observed, and is cause for concern since Sheltzer and Smith [-@sheltzer_2014]'s results might reflect discrimination against women during hiring [@Moss_2012], or avoidance by women of elite research groups (e.g.
-due to 'imposter syndrome' or the perception that these groups are sexist).
+due to 'imposter syndrome' or a perception that some groups are sexist).
 We also found a little evidence that gender homophily is detrimental to research quality, in that high-impact journals tended to have weaker homophily.
-Assuming that papers published in high-impact journals are of higher average quality [which is contentious; @Garfield_2006], this result provides non-experimental support for the hypothesis that mixed-gender teams produce better research than single-gender teams [@Britton_2000; @Reagans_2001; @Hong_2004; @Whittington_2008; @Bear_2011; @Herrera_2012; @Campbell_2013].
+Assuming that papers published in high-impact journals are of higher average quality [which is contentious; @Garfield_2006], our results provide non-experimental support for the hypothesis that mixed-gender teams produce better research than single-gender teams [@Britton_2000; @Reagans_2001; @Hong_2004; @Whittington_2008; @Bear_2011; @Herrera_2012; @Campbell_2013].
+Another issue is that if many collaborations are between established researchers, there will be an excess of male-male collaborations in fields where women in senior positions are rare; some of the observed homophily might therefore reflect the elevated gender gap among senior researchers.
 
-Homophily might also have more benign causes.
+On the other hand, homophily might have more benign causes.
 Collaboration is often most enjoyable and productive when working with like-minded people, who might be same-gendered more often than not.
 We also suppose that some people consciously choose to preferentially collaborate with women in order to help close the gender gap in the workforce; this would create homophily if women do this more than men.
-In support of this interpretation, women appear more likely than men to promote the work of female collagues by inviting them to give talks [@nittrouer_2018; @debarre_2018].
-Related to this, Ghiasi et al.
-[-@Ghiasi_2015] concluded that women in engineering are "compliant [in reproducing] male-dominated scientific structures" because they do not collaborate often enough with other women; their data suggest that coauthorships between two women are about 30% more frequent than expected under random assortment (see their Figure 7).
-In contrast, we propose that men as well as women should ensure that they are not inadvertently overlooking or excluding female colleagues, particularly since men are disproportionately represented among senior researchers [@Shaw_2012].
+In support of this interpretation, women appear more likely than men to promote the work of female colleagues by inviting them to give talks [@nittrouer_2018; @debarre_2018].
+Given that many collaborative research projects unfortunately involve a gendered division of labour [@macaluso_2016], working with a same-gendered colleague may provide exposure to new parts of the research process, and (especially for the minority gender) a welcome change of pace.
 
@@
 # Acknowledgements
-CM was supported by the Academy of Finland (284666 to the Centre of Excellence in Biological Interactions).
+CM was supported by the Academy of Finland (284666 to the Centre of Excellence in Biological Interactions). We thank Devi Stuart-Fox and Dominique Potvin for helpful discussion.
 

### 2018-06-06 Gender_homophily_analysis.Rmd
Fix fig 3

@@
   scale_fill_brewer(palette = "Dark2") + 
-  theme_bw(13) + 
-  theme(legend.position = "none", panel.grid = element_blank()) +
+  theme_minimal(13) + 
+  theme(legend.position = "none") +
   xlab("Number of authors listed on the paper") + 

### 2018-06-07 Gender_homophily_analysis.Rmd
Small tweaks

@@
 ```{r fig.width=6, fig.height=5}
-model <- summary(lm(alpha.adjusted ~ standardised.IF, data = master.dataset %>% 
+model <- summary(lm(alpha.adjusted ~ scale(standardised.IF), data = master.dataset %>% 
                       filter(Age == "New", 

### 2018-06-13 manuscript/Manuscript_Claire.Rmd
Small tweaks
Version for bioarXiv

@@
 ---
-title: "Researchers preferentially collaborate with same-gendered colleagues across the life sciences"
+title: Researchers preferentially collaborate with same-gendered colleagues across
+  the life sciences
 author:
@@
 output:
-  pdf_document: default
+  pdf_document: 
+    keep_tex: true
   html_document: default
@@
   using more stringent methods and a larger dataset than in past work. Our results
-  reaffirm that researchers co-publish with colleagues of the same gender, and show
-  that this 'gender homophily' is slightly stronger today than it was 10 years ago.
-  Contrary to our expectations, we found no evidence that homophily is driven mostly
-  by senior academics, and no evidence that homophily is stronger in fields where
-  women are in the minority. Interestingly, journals with a high impact factor for their 
-  discipline tended to have comparatively low homophily, as predicted if mixed-gender teams 
-  produce better research. We discuss potential causes of gender homophily in science. 
+  reaffirm that researchers preferentially co-publish with colleagues of the same
+  gender, and show that this 'gender homophily' is slightly stronger today than it
+  was 10 years ago. Contrary to our expectations, we found no evidence that homophily
+  is driven mostly by senior academics, and no evidence that homophily is stronger
+  in fields where women are in the minority. Interestingly, journals with a high impact
+  factor for their discipline tended to have comparatively low homophily, as predicted
+  if mixed-gender teams produce better research. We discuss some potential causes of gender
+  homophily in science.
   \vspace{5mm} \par\noindent
-  \textbf{Keywords:} Gender bias, Homophily, Scientific collaboration, Text mining,
-  Women in STEM. 
+  \textbf{Keywords:} Gender bias, Homophily, Research collaboration, Text mining, Women in STEM.
 ---
@@
 
-Because publishing, networking and collaboration are instrumental to scientific productivity and academic career advancement [@Lee_2005; @Wuchty_2007; @Abramo_2009; @Lariviere_2015], dozens of studies have tested for gender differences in these areas [e.g.
+Writing papers, networking, and collaboration are all instrumental to research productivity and academic career advancement [@Lee_2005; @Wuchty_2007; @Abramo_2009; @Lariviere_2015], and dozens of studies have tested for gender differences in these areas [e.g.
 @Long_1992; @Bozeman_2011; @Abramo_2013; @Badar_2013; studies listed in Table S1 of @Holman_2018].
@@
 For reasons such as these, multiple studies have examined the author lists of published research articles in order to test for gender differences in collaboration frequency or pattern.
-To our knowledge, all such studies imply that men co-publish with other men, and women with women, more often than expected if collaborators assort randomly with respect to gender [@Ferber_1980; @McDowell_1992; @Ghiasi_2015; @Crow_2015; @Fahmy_2017; @Zettler_2017; @Jadidi_2017; @Teele_2017; @Araujo_2017a; @Araujo_2017b].
-This pattern of assortative publishing is often termed 'gender homophily'.
+To our knowledge, most or all such studies imply that men co-publish with men, and women with women, more often than expected if collaborators assort randomly with respect to gender [@Ferber_1980; @McDowell_1992; @Ghiasi_2015; @Crow_2015; @Fahmy_2017; @Zettler_2017; @Jadidi_2017; @Teele_2017; @Araujo_2017a; @Araujo_2017b].
+This pattern of assortative publishing has often been termed 'gender homophily'.
 
@@
 Essentially, whenever coauthorship data are sampled from two or more discrete sets of literature, which vary in the author gender ratio and which are largely not connected by collaboration, the number of same-gendered coauthors will be inflated.
-This can give the impression that authors preferentially publish with same-gendered colleagues if no gender preferences exist, or even if opposite-gendered colleagues are preferred.
-For example, a sample of literature containing bioinformatics and cell biology papers will probably contain an excess of mostly-male and mostly-female author lists, simply because researchers preferentially work with colleagues from the same discipline, and because the author gender ratio is more male-biased in bioinformatics than in cell biology [@Holman_2018].
+This can give the impression that authors preferentially publish with same-gendered colleagues even if no gender preferences exist, or if the true preference is for opposite-gendered colleagues ('gender heterophily').
+For example, a sample of literature containing bioinformatics and cell biology papers will probably contain an excess of mostly-male and mostly-female author lists, simply because researchers usually collaborate within their own discipline, and because the author gender ratio is more male-biased in bioinformatics than in cell biology [@Holman_2018].
 
@@
 The Wahlund effect will tend to inflate the frequency of same-sex coauthorships whenever the data is composed of two or more disconnected subsets of literature with different author gender ratios; these subsets could be research disciplines, older versus newer papers, or papers from authors in different countries.
+The example countries and disciplines were selected based on [@Holman_2018].
 \label{wahlund_plot}](../figures/Fig1.pdf)
@@
 \autoref{alpha_histograms} shows the distribution of $\alpha'$ estimates in 2015-2016 across all journals for which we recovered sufficient data, when $\alpha'$ was calculated for all authors, first authors only, or last authors only.
-The great majority of journals had $\alpha' > 0$, and for many of these the FDR-corrected p-values suggested that $\alpha'$ was significantly greater than zero (1469/2077 journals were significant in 2015-16, and 404/1192 in 2005-6; S1 Data).
+Most journals had positive values of $\alpha'$ (77-92%, depending on time period and author type; S1 Data), and for many of these the FDR-corrected p-values suggested that $\alpha'$ was significantly greater than zero (1469/2077 journals were significant in 2015-16, and 404/1192 in 2005-6; S1 Data).
 Only 2/2077 journals had statistically significantly heterophily (i.e.
@@
 ![Of the `r n.journals` journals for which we had adequate data in 2015-2016, `r n.sig.homo` showed statistically significant evidence of gender homophily (denoted by $\alpha' > 0$), and `r n.sig.hetero` showed statistically significant evidence of heterophily ($\alpha' < 0$), after false discovery rate correction.
-The white area shows the number of journals for which homophily was significantly stronger than expected under the null hypothesis (p < 0.05), while the blue area shows all the remainder.
+The white area shows the number of journals for which homophily was significantly stronger than expected under the null hypothesis (corrected p < 0.05), while the blue area shows all the remainder.
 Patterns were similar whether $\alpha'$ was calculated for all authors, for first authors only, or for last authors only.
@@
 \autoref{alpha_histograms} illustrates the variance in journal homophily values ($\alpha'$) across scientific disciplines.
-All disciplines had a positive average $\alpha'$, although homophily appeared somewhat stronger in some disciplines than others (e.g.
+All disciplines had positive mean $\alpha'$, although homophily appeared somewhat stronger in some disciplines than others (e.g.
 mean $\alpha'$ was $0.12{\pm}0.02$ for Urology journals and $0.03{\pm}0.01$ for Veterinary Medicine journals; \autoref{alpha_histograms}, S4 Data).
-However, there was no evidence for consistent differences in $\alpha'$ between disciplines: the random factor 'Discipline' explained around 1% of the variance in $\alpha'$ in the two linear mixed models described in the previous section (see \autoref{alpha_histograms} and mixed models in Online Supplementary Material).
-Thus, we cannot reject the null hypothesis that the processes causing positive $\alpha'$ are similarly strong in all the disciplines we examined.
+However, there was no formal evidence for consistent differences in $\alpha'$ between disciplines: the random factor 'Discipline' explained around 1% of the variance in $\alpha'$ in the two linear mixed models described in the previous section (see \autoref{alpha_histograms} and mixed models in Online Supplementary Material).
+Thus, the processes responsible for producing positive $\alpha'$ values appear to be similarly strong in all the disciplines we examined.
 
@@
 ## Relationship between gender homophily and number of authors
-Papers with two authors had significantly lower (but still positive) $\alpha'$ values than papers with more than two authors, on average (\autoref{author_number}; model output in Online Supplmentary Material).
+Papers with two authors had significantly lower (but still positive) $\alpha'$ values relative to papers with more than two authors, on average (\autoref{author_number}; statistical results in Online Supplmentary Material).
 Papers with 3, 4 or 5+ authors had essentially identical average $\alpha'$ values.
@@
 We observed a noisy but statistically significant linear relationship between standardised journal impact factor and $\alpha'$, such that journals with a high impact factor for their discipline had weaker gender homophily than did journals with a low impact factor for their discipline (\autoref{impact_factor}; linear regression: $R^2$ = `r R2`, $t_{1415}$ = -8.0, p < 0.0001).
+The slope of the regression was $-0.012{\pm}0.0015$, indicating that increasing the discipline-standardised impact factor by one standard deviation is associated with a reduction in $\alpha'$ of 0.012.
 
@@
 ## Analysis accounting for differences in author gender ratio between countries
-When we restricted the analysis by country, we observed statistically significant homophily for 72 of the 325 journal-country combinations tested (64 unique journals and 18 unique countries), and no significant evidence of heterophily (S4-S5 Fig).
-Additionally, the values of $\alpha'$ calculated for each journal-country combination were mostly very similar to the $\alpha'$ values calculated for the journal as a whole (i.e.
-when pooling papers from different countries); on average, $\alpha'$ was only inflated in the main (non-country specific) analysis by _c_.
-0.002 (S6 Fig).
+When we restricted the analysis by country, we observed statistically significant homophily for 72 of the 325 journal-country combinations tested (64 unique journals and 18 unique countries), and no significant heterophily (S4-S5 Fig).
+Additionally, the values of $\alpha'$ calculated for each journal-country combination were only very slightly lower than the $\alpha'$ values calculated for the journal as a whole (i.e.
+when pooling papers from different countries, as was done to make \autoref{alpha_histograms}): on average, the difference in $\alpha'$ was only 0.002 (S6 Fig).
 These results suggest that our findings of widespread homophily in the main analysis were not driven solely by a Wahlund effect resulting from gender differences between countries.
@@
 However, one can get $\alpha > 0$ for realistic combinations of parameters, e.g.
-a moderate shortage of women in senior positions coupled with a moderate excess of within-career stage collaboration, suggesting this effect might contribute to some of the observed homophily.
+a moderate shortage of women in senior positions coupled with a moderate excess of within-career stage collaboration, suggesting this effect might contribute to some of the observed homophily (in this and previous studies).
 
@@
 In relative terms, our findings are even more striking: for example, if men have 20% female coauthors and women have 30% (i.e.
-$\alpha' = 0.1$ in a field with a male-biased gender ratio), then women publish with women 50% more often than men do.
+$\alpha' = 0.1$ in a field with a typical gender ratio [@Holman_2018]), then women publish with women 50% more often than men do.
 
@@
 However, we suspect the contribution of these to be minor, for four reasons: we found positive $\alpha'$ after controlling for three obvious sources of Wahlund effect; there was no inflation of $\alpha'$ in highly multidisciplinary journals; restricting the data by country yielded similar estimates of $\alpha'$; and we showed that differences in gender ratio between career stages are unlikely to fully explain our results.
-On balance, we believe the data suggest that it is likely that some researchers do preferentially select same-gendered collaborators, although the effect size of this preference is difficult to measure.
+On balance, we believe the data suggest that it is likely that some researchers do preferentially select same-gendered collaborators, although the frequency and strength of this preference is difficult to ascertain.
 
@@
 Contrary to this hypothesis, we found no evidence that gender homophily is restricted to particular disciplines: $\alpha'$ was similarly high across the board (\autoref{alpha_histograms}).
-Moreover, $\alpha'$ was lower for journals with a male-biased author gender ratio relative to those with an even gender ratio.
-Interestingly, $\alpha'$ was highest in journals with a female-biased author gender ratio.
+Interestingly, gender homophily was weakest for journals with a male-biased author gender ratio, and strongest in journals with a female-biased author gender ratio.
 This may suggest that men are more likely to preferentially seek out male collaborators in fields where men are a minority, relative to the homophily displayed by women in fields where women are a minority.
-However, this latter result is only tentatively supported since there are few journals in which most authors are women (\autoref{alpha_gender_ratio}).
+However, this latter result is only tentatively supported since our sample contains few journals in which most authors are women (\autoref{alpha_gender_ratio}).
 
@@
 Regarding our finding of weaker homophily among 2-author papers, we suspect that many 2-author teams comprise a student/postdoc and a senior staff member, making these teams especially likely to be mixed-gender, due to the elevated gender gap among senior researchers [@Shaw_2012; @Holman_2018].
-Assuming this interpretation is correct, this result suggests that our overall $\alpha'$ may underestimate the strength of peoples' preferences for same-gendered collaborators; essentially, women seeking a senior collaborator could be constrained to work mostly with men, meaning that people's ideal and realised gender preferences would be mismatched.
+Assuming this interpretation is correct, this result suggests that our reported $\alpha'$ values may underestimate the strength of peoples' preferences for same-gendered collaborators; essentially, women seeking a senior collaborator could be constrained to work mostly with men, meaning that people's ideal and realised gender preferences would be mismatched.
 On a related note, Ghiasi et al.
-[-@Ghiasi_2015] argue that women in engineering are "compliant [in reproducing] male-dominated scientific structures" because they do not collaborate often enough with other women; for reference, their Figure 7 suggests that coauthorships between women are roughly 30% more frequent than expected under random assortment.
+[-@Ghiasi_2015] argue that women in engineering are "compliant [in reproducing] male-dominated scientific structures" because they do not collaborate often enough with other women (their Figure 7 suggests that coauthorships between women are 30% more frequent than expected under random assortment).
 By contrast, we feel that it is not helpful to recommend that women collaborate primarily with other women, e.g.
 because this constrains women's options and may be counter-productive (particularly in fields like engineering, where 90% of professors are men [@Shaw_2012]).
-Instead, we suggest that researchers of both genders can help to close the gender gap in science.
+Instead, we suggest that researchers of both genders can help to close the gender gap in STEMM.
 In the context of collaboration, one way to do this is to undertake self-examination to ensure that one is not inadvertently overlooking or excluding female potential students and colleagues.
@@
 in terms of training and mentoring, allocation of work, and how one frames or promotes the collaboration (e.g.
-in conference presentations or on a website); evidence suggests that unconscious bias causes people to undervalue women's research achievements [@Moss_2012], and to assign menial or under-valued tasks to women and more prestigious and employable tasks (e.g.
-writing and project leadership) to men [@macaluso_2016].
+in conference presentations or on a website); evidence suggests that unconscious bias causes people to undervalue women's research achievements [@Moss_2012], and possibly to assign menial or under-valued tasks to women and more prestigious tasks to men [@macaluso_2016].
 
@@
 This finding could contribute to the homophily we observed, and is cause for concern since Sheltzer and Smith [-@sheltzer_2014]'s results might reflect discrimination against women during hiring [@Moss_2012], or avoidance by women of elite research groups (e.g.
-due to 'imposter syndrome' or a perception that some groups are sexist).
+due to gender differences in confidence, or a perception that some groups are sexist).
 We also found a little evidence that gender homophily is detrimental to research quality, in that high-impact journals tended to have weaker homophily.

### 2018-06-13 manuscript/Manuscript_Claire.tex
Version for bioarXiv

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Researchers preferentially collaborate with same-gendered colleagues across the life sciences},
            pdfauthor={Luke Holman* and Claire Morandin; *;},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{Researchers preferentially collaborate with same-gendered colleagues
across the life sciences}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{Luke Holman* and Claire Morandin\textsuperscript{§} \\ *\textit{luke.holman@unimelb.edu.au} \\ \textsuperscript{§}\textit{claire.morandin@helsinki.fi} \vspace{5mm}}
  \preauthor{\centering\large\emph}
  \postauthor{\par}
  \date{}
  \predate{}\postdate{}

\usepackage{microtype} \usepackage[all]{hypcap} \usepackage{amsmath}
\usepackage{booktabs} \usepackage{caption}
\usepackage[labelfont=bf]{caption} \usepackage{titling}
\pretitle{\begin{flushleft}\LARGE}
\posttitle{\par\end{flushleft}\vskip 0.5em}
\preauthor{\begin{flushleft}\large \lineskip 0.5em}
\postauthor{\par\end{flushleft}} \predate{\begin{flushleft}\large}
\postdate{\par\end{flushleft}} \usepackage{fancyhdr} \pagestyle{fancy}
\fancyhead[LO,LE]{\textsl{\leftmark}} \rhead[]{Gender and coauthorship}
\usepackage[noabbrev,capitalise]{cleveref} \usepackage{titlefoot}
\usepackage{amssymb} \usepackage{rotating} \usepackage{lineno}
\linenumbers

\begin{document}
\maketitle
\begin{abstract}
Evidence suggests that women in academia are hindered by conscious and
unconscious biases, and often feel excluded from formal and informal
opportunities for research collaboration. In addition to ensuring
fairness and helping to redress gender imbalance in the academic
workforce, increasing women's access to collaboration could help
scientific progress by drawing on more of the available human capital.
Here, we test whether researchers preferentially collaborate with
same-gendered colleagues, using more stringent methods and a larger
dataset than in past work. Our results reaffirm that researchers
preferentially co-publish with colleagues of the same gender, and show
that this `gender homophily' is slightly stronger today than it was 10
years ago. Contrary to our expectations, we found no evidence that
homophily is driven mostly by senior academics, and no evidence that
homophily is stronger in fields where women are in the minority.
Interestingly, journals with a high impact factor for their discipline
tended to have comparatively low homophily, as predicted if mixed-gender
teams produce better research. We discuss some potential causes of
gender homophily in science. \vspace{5mm}
\par\noindent \textbf{Keywords:} Gender bias, Homophily, Research
collaboration, Text mining, Women in STEM.
\end{abstract}

\maketitle

\unmarkedfntext{*School of BioSciences, The University of Melbourne, Victoria, Australia.}
\maketitle\unmarkedfntext{\textsuperscript{§}Organismal and Evolutionary Biology Research Programme,
Faculty of Biological and Environmental Sciences, University of Helsinki, Finland.}

\newpage

\section{Introduction}\label{introduction}

Women are severely underrepresented in many branches of science,
technology, engineering, mathematics, and medicine (STEMM), and face
additional challenges and inequities relative to men {[}1--5{]}. On
average, women occupy more junior positions {[}6,7{]} with lower
salaries {[}8,9{]}, receive less grant money {[}10,11{]}, are promoted
more slowly {[}12--15{]}, and are allocated fewer resources {[}16{]} and
less research funding {[}17--19{]}. Experimental evidence suggests that
bias against women plays a major role in generating these differences
{[}20,21{]}.

Writing papers, networking, and collaboration are all instrumental to
research productivity and academic career advancement {[}22--25{]}, and
dozens of studies have tested for gender differences in these areas
{[}5,26--29{]}. For example, studies have concluded that women tend to
be less involved in international collaboration {[}19,28,30--32{]},
collaborate less within their own university departments {[}31{]}, have
less prestigious collaborations {[}33{]}, and fewer collaborations in
total {[}34{]}. These gender differences in collaboration practice
presumably have multiple causes, which might include implicit and
explicit gender bias {[}20{]}, differential family obligations
{[}33,35,36{]}, gender differences in confidence or self-esteem
{[}37{]}, concerns relating to sexual harassment {[}38{]}, and unequal
access to conferences {[}39{]} or travel funds {[}32{]}.

A high, steadily increasing proportion of research papers is written by
more than one author {[}3{]}, making collaboration a key predictor of
publication output, and thus of career prospects {[}40,41{]}.
Additionally, empirical studies imply that mixed-gender or otherwise
diverse teams produce better outputs on collaborative tasks than less
diverse teams {[}42--48{]}. For reasons such as these, multiple studies
have examined the author lists of published research articles in order
to test for gender differences in collaboration frequency or pattern. To
our knowledge, most or all such studies imply that men co-publish with
men, and women with women, more often than expected if collaborators
assort randomly with respect to gender {[}49--58{]}. This pattern of
assortative publishing has often been termed `gender homophily'.

However, we believe that prior studies of gender homophily were hindered
by a largely unacknowledged statistical issue that we name the Wahlund
effect (\autoref{wahlund_plot}), by analogy with the conceptually
similar Whalund effect in population genetics {[}59{]}. The Wahlund
effect makes it deceptively difficult to infer gender-based preferences
simply by counting the number of same- and mixed-gender coauthorships.
Essentially, whenever coauthorship data are sampled from two or more
discrete sets of literature, which vary in the author gender ratio and
which are largely not connected by collaboration, the number of
same-gendered coauthors will be inflated. This can give the impression
that authors preferentially publish with same-gendered colleagues even
if no gender preferences exist, or if the true preference is for
opposite-gendered colleagues (`gender heterophily'). For example, a
sample of literature containing bioinformatics and cell biology papers
will probably contain an excess of mostly-male and mostly-female author
lists, simply because researchers usually collaborate within their own
discipline, and because the author gender ratio is more male-biased in
bioinformatics than in cell biology {[}5{]}.

\begin{figure}[htbp]
\centering
\includegraphics{../figures/Fig1.pdf}
\caption{The Wahlund effect can make it appear as if authors prefer to
publish with same-gendered colleagues, even if no such preference
exists. Here, coloured circles represent male and female authors, and
coauthors are linked with lines. Across the whole set of ten papers,
there is an apparent excess of same-gender collaborations: there are six
same-gender papers and only four mixed-gender papers, which is fewer
than the \(10\times2\times0.5\times0.5 = 5\) mixed-gender papers
expected under the null hypothesis that authors assort randomly with
respect to gender. However, within each subset, there is no evidence
that authors prefer to publish with same-gendered individuals (if
anything, this small dataset suggests gender heterophily). The Wahlund
effect will tend to inflate the frequency of same-sex coauthorships
whenever the data is composed of two or more disconnected subsets of
literature with different author gender ratios; these subsets could be
research disciplines, older versus newer papers, or papers from authors
in different countries. The example countries and disciplines were
selected based on {[}5{]}. \label{wahlund_plot}}
\end{figure}

In the present study, we test whether life sciences researchers tend to
co-publish with same-gendered colleagues, while controlling for the
Wahlund effect as strictly as possible. We use a recently-published
dataset describing the gender of 35.5m authors from 9.15m articles
indexed on PubMed {[}5{]}. Holman et al. {[}5{]} reported large
differences in the gender ratio of authors across research disciplines,
journals, countries, and across the years 2002-2016. We therefore tested
for gender homophily while restricting our analysis to particular
journals (i.e.~research specialties), time periods, and countries. We
quantified gender assortment using a metric called \(\alpha'\) {[}60{]},
which is positive when same-gender authors publish together more often
than expected (gender homohily), negative when opposite-gender authors
publish together more often than expected (heterophily), and equal to
zero when authors assort randomly with respect to gender (see Methods).

\section{Results}\label{results}

\subsection{Gender homophily by discipline, time period, and authorship
position}\label{gender-homophily-by-discipline-time-period-and-authorship-position}

\autoref{alpha_histograms} shows the distribution of \(\alpha'\)
estimates in 2015-2016 across all journals for which we recovered
sufficient data, when \(\alpha'\) was calculated for all authors, first
authors only, or last authors only. Most journals had positive values of
\(\alpha'\) (77-92\%, depending on time period and author type; S1
Data), and for many of these the FDR-corrected p-values suggested that
\(\alpha'\) was significantly greater than zero (1469/2077 journals were
significant in 2015-16, and 404/1192 in 2005-6; S1 Data). Only 2/2077
journals had statistically significantly heterophily (i.e.
\(\alpha' < 0\)) in 2015-16, and 1/1192 in 2005-6 (S2 Table). The
remaining 606 or 787 journals (in 2015 and 2005 respectively) had a
value of \(\alpha'\) not significantly different from zero, consistent
with the null hypothesis of random assortment with respect to gender. We
also confirmed that the majority of papers had multiple authors, in most
journals (S2 Data) and disciplines (S3 Data, S1 Fig).

\begin{figure}[htbp]
\centering
\includegraphics{../figures/Fig2.pdf}
\caption{Of the 2116 journals for which we had adequate data in
2015-2016, 825 showed statistically significant evidence of gender
homophily (denoted by \(\alpha' > 0\)), and 1 showed statistically
significant evidence of heterophily (\(\alpha' < 0\)), after false
discovery rate correction. The white area shows the number of journals
for which homophily was significantly stronger than expected under the
null hypothesis (corrected p \textless{} 0.05), while the blue area
shows all the remainder. Patterns were similar whether \(\alpha'\) was
calculated for all authors, for first authors only, or for last authors
only. \label{alpha_histograms}}
\end{figure}

\(\alpha'\) was significantly higher in the literature sample from
2015-16 relative to 2005-6, though the difference in means was small (S2
Fig; Effect of the fixed factor `Time period' in a linear mixed model of
the data for all author positions: Cohen's \(d\) = \(0.091{\pm}0.04\),
\(t_{953}\) = 2.42, p = 0.016).

When comparing pairs of \(\alpha'\) values estimated for the first and
last authors for the same journals, we found that \(\alpha'\) tended to
be higher for first authors than for last authors (S3 Fig; Effect of the
fixed factor `Authorship position' in a linear mixed model: Cohen's
\(d\) = \(0.065{\pm}0.02\), \(t_{2024}\) = 4.28, p \textless{} 0.0001).
This suggests that the gender of the first author was a slightly
stronger predictor of the remaining authors' genders than the gender of
the last author, i.e.~the opposite of what is predicted if senior
scientists are causally responsible for homophily.

\subsection{Variance in homophily between
disciplines}\label{variance-in-homophily-between-disciplines}

\autoref{alpha_histograms} illustrates the variance in journal homophily
values (\(\alpha'\)) across scientific disciplines. All disciplines had
positive mean \(\alpha'\), although homophily appeared somewhat stronger
in some disciplines than others (e.g.~mean \(\alpha'\) was
\(0.12{\pm}0.02\) for Urology journals and \(0.03{\pm}0.01\) for
Veterinary Medicine journals; \autoref{alpha_histograms}, S4 Data).
However, there was no formal evidence for consistent differences in
\(\alpha'\) between disciplines: the random factor `Discipline'
explained around 1\% of the variance in \(\alpha'\) in the two linear
mixed models described in the previous section (see
\autoref{alpha_histograms} and mixed models in Online Supplementary
Material). Thus, the processes responsible for producing positive
\(\alpha'\) values appear to be similarly strong in all the disciplines
we examined.

There was no indication that journals publishing on a wide range of
topics have higher \(\alpha'\) values than more specialised journals,
due to the Wahlund effect. For example, the journal category
`Multidisciplinary' -- which includes journals like \emph{PLoS ONE},
\emph{Nature}, \emph{Science}, and \emph{PNAS} -- did not have notably
elevated \(\alpha'\) (\autoref{alpha_histograms}). This result suggests
that our estimates of homophily, and estimates from some earlier
studies, are not notably inflated by the presence of disparate research
topics (with variable author gender ratios) being published within
individual journals.

\subsection{Relationship between gender homophily and number of
authors}\label{relationship-between-gender-homophily-and-number-of-authors}

Papers with two authors had significantly lower (but still positive)
\(\alpha'\) values relative to papers with more than two authors, on
average (\autoref{author_number}; statistical results in Online
Supplmentary Material). Papers with 3, 4 or 5+ authors had essentially
identical average \(\alpha'\) values. The variance in \(\alpha'\) across
journals was also a little higher for 2-authors papers compared to the
remainder (\autoref{author_number}), though part of this variance is due
to the reduced sample size (in terms of number of authors) for the
2-author papers.

\begin{figure}[htbp]
\centering
\includegraphics{../figures/Fig3.pdf}
\caption{The coefficient of homophily (\(\alpha'\)) was slightly less
positive when calculated for two-author papers only, relative to papers
with longer author lists. The individual points, whose distribution is
summarised by the violin plots, correspond to individual journals. The
larger white points show the mean for each group (and its 95\% CIs), as
calculated by a Bayesian meta-regression model accounting for repeated
measures of \(\alpha'\) within journals, as well as the precision with
which \(\alpha'\) was estimated. \label{author_number}}
\end{figure}

\subsection{Relationship between gender homophily and gender
ratio}\label{relationship-between-gender-homophily-and-gender-ratio}

We next tested whether researchers are more or less likely to publish
with same-gendered colleagues in strongly gender-biased disciplines
(e.g.~Surgery or Nursing), relative to disciplines with a comparatively
gender-balanced workforce (e.g.~Psychiatry). We found a positive,
non-linear relationship between the overall gender ratio of all authors
publishing in a particular journal {[}5{]}, and the estimated value of
\(\alpha'\) for all authors and for first authors
(\autoref{alpha_gender_ratio}). Journals with a balanced or
female-biased author gender ratio tended to have higher \(\alpha'\) than
journals with a male-biased author gender ratio (GAM smooth terms p
\textless{} 0.001; Online Supplementary Material). The relationship was
not statistically significant when \(\alpha'\) was calculated for last
authors (GAM, p = 0.142), though the trend appeared similar
(\autoref{alpha_gender_ratio}).

\begin{figure}[htbp]
\centering
\includegraphics{../figures/Fig4.pdf}
\caption{There is a weakly positive, non-linear relationship between the
gender ratio of authors publishing in a journal, and the coefficient of
homophily (\(\alpha'\)). Specifically, journals with 50\% women authors
or higher tended to have more same-sex coauthorships than did journals
with predominantly men authors. This relationship held whether
\(\alpha'\) was calculated for all authors, first authors only, or last
authors only. A negative value on the x-axis denotes an excess of men
authors, a positive value denotes an excess of women authors, and zero
denotes gender parity. The lines were fitted using generalised additive
models with the smoothing parameter \(k\) set to 3.
\label{alpha_gender_ratio}}
\end{figure}

\subsection{Relationship between journal impact factor and gender
homophily}\label{relationship-between-journal-impact-factor-and-gender-homophily}

We observed a noisy but statistically significant linear relationship
between standardised journal impact factor and \(\alpha'\), such that
journals with a high impact factor for their discipline had weaker
gender homophily than did journals with a low impact factor for their
discipline (\autoref{impact_factor}; linear regression: \(R^2\) = 0.043,
\(t_{1415}\) = -8.0, p \textless{} 0.0001). The slope of the regression
was \(-0.012{\pm}0.0015\), indicating that increasing the
discipline-standardised impact factor by one standard deviation is
associated with a reduction in \(\alpha'\) of 0.012.

\begin{figure}[htbp]
\centering
\includegraphics{../figures/Fig5.pdf}
\caption{Journal impact factor (expressed relative to the average for
the discipline) is negatively correlated with \(\alpha'\). The
relationship is noisy (\(R^2\) = 0.043), but the results suggest that
journals with strong homophily tend to have lower impact factors than
journals with weak homophily in the same discipline.
\label{impact_factor}}
\end{figure}

\subsection{Analysis accounting for differences in author gender ratio
between
countries}\label{analysis-accounting-for-differences-in-author-gender-ratio-between-countries}

When we restricted the analysis by country, we observed statistically
significant homophily for 72 of the 325 journal-country combinations
tested (64 unique journals and 18 unique countries), and no significant
heterophily (S4-S5 Fig). Additionally, the values of \(\alpha'\)
calculated for each journal-country combination were only very slightly
lower than the \(\alpha'\) values calculated for the journal as a whole
(i.e.~when pooling papers from different countries, as was done to make
\autoref{alpha_histograms}): on average, the difference in \(\alpha'\)
was only 0.002 (S6 Fig). These results suggest that our findings of
widespread homophily in the main analysis were not driven solely by a
Wahlund effect resulting from gender differences between countries.

\subsection{\texorpdfstring{Theoretical expectations for \(\alpha\) when
the gender ratio differs between career
stages}{Theoretical expectations for \textbackslash{}alpha when the gender ratio differs between career stages}}\label{theoretical-expectations-for-alpha-when-the-gender-ratio-differs-between-career-stages}

As shown in \autoref{simulation}, we predict that \(\alpha\) is expected
to be non-zero, even if collaborators are randomly selected with respect
to gender, provided that there is a gender gap between career stages.
The extent to which \(\alpha\) deviates from zero depends on the
relative frequencies of collaboration within and between career stages.
When \textgreater{}50\% of collaborations were between early and
established researchers, we expect gender heterophily (\(\alpha < 0\)).
Conversely, when \textgreater{}50\% of collaborations occured within
career stages, we expect gender homophily (\(\alpha > 0\)). In a few
parameter spaces (shown in red; \autoref{simulation}), \(\alpha\) was
quite high, and overlapped with the values that we estimated
(\autoref{alpha_histograms}).

\begin{figure}[htbp]
\centering
\includegraphics{../figures/Fig6_inkscape.pdf}
\caption{When there is a difference in gender ratio between early-career
and established researchers, and collaboration is non-random with
respect to career stage, the null expectation for \(\alpha\) deviates
from zero. An excess of collaborations between career stages gives the
appearance of gender heterophily (lower rows, blue areas), while an
excess of within-career stage collaborations produced apparent gender
homophily (upper rows, red areas). However, the conditions required for
strong gender homophily are quite restrictive, making it unlikely that
this issue explains all of the homophily observed in Figure 2. Contour
lines denote increments of 0.1. \label{simulation}}
\end{figure}

Despite this overlap, \autoref{simulation} suggests that our main
conclusions (and those of other studies of gender homophily) are
probably robust to this career stage issue. We only expect strongly
positive \(\alpha\) when A) the gender ratio is highly skewed across
career stages (e.g.~a 5-fold difference), and B) collaborations between
early and established researchers are very rare (e.g. \textless{}10\% of
the total). Both of these conditions are untrue for most fields: the
gender gap across careers stages is generally less pronounced {[}1,5{]},
and it is very common for early-career researchers to co-publish with an
established mentor {[}61{]}. However, one can get \(\alpha > 0\) for
realistic combinations of parameters, e.g.~a moderate shortage of women
in senior positions coupled with a moderate excess of within-career
stage collaboration, suggesting this effect might contribute to some of
the observed homophily (in this and previous studies).

Lastly, we note that if there is a gender gap between career stages and
coauthorships between early-career and established researchers comprise
\textgreater{}50\% of the total, then the baseline expectation for
\(\alpha\) is actually less than zero (blue areas in
\autoref{simulation}). Therefore, our results might under-estimate the
extent to which researchers preferentially select same-gendered
collaborators in some cases.

\section{Discussion}\label{discussion}

We found evidence that researchers preferentially publish with
same-gendered coauthors, even after implementing stringent controls for
Wahlund effects (\autoref{wahlund_plot}). Our study therefore reaffirms
earlier studies' conclusions {[}49--57,62{]} and establishes their
generality across the life sciences. Relatively few journals had
\(\alpha'\) values below zero, and almost no journals showed
statistically significant gender heterophily after controlling for
multiple testing. The excess of same-gender coauthorships was quite
large: many journals had \(\alpha' > 0.1\), indicating that the gender
ratio of men's and women's coauthors differs by \textgreater{}10\% in
absolute terms. In relative terms, our findings are even more striking:
for example, if men have 20\% female coauthors and women have 30\% (i.e.
\(\alpha' = 0.1\) in a field with a typical gender ratio {[}5{]}), then
women publish with women 50\% more often than men do.

An important limitation of our study is that we cannot reliably
determine the cause(s) of the observed excess of same-gender
coauthorships. As well as the obvious interpretation -- conscious or
unconcious selection of same-gender collaborators by men, women, or both
-- our results could be partly explained by uncontrolled Wahlund
effects. However, we suspect the contribution of these to be minor, for
four reasons: we found positive \(\alpha'\) after controlling for three
obvious sources of Wahlund effect; there was no inflation of \(\alpha'\)
in highly multidisciplinary journals; restricting the data by country
yielded similar estimates of \(\alpha'\); and we showed that differences
in gender ratio between career stages are unlikely to fully explain our
results. On balance, we believe the data suggest that it is likely that
some researchers do preferentially select same-gendered collaborators,
although the frequency and strength of this preference is difficult to
ascertain.

We hypothesised that disciplines with a strongly skewed gender ratio
might show the strongest gender homophily, e.g.~because being in the
minority might increase motivation to seek out same-gendered colleagues.
Contrary to this hypothesis, we found no evidence that gender homophily
is restricted to particular disciplines: \(\alpha'\) was similarly high
across the board (\autoref{alpha_histograms}). Interestingly, gender
homophily was weakest for journals with a male-biased author gender
ratio, and strongest in journals with a female-biased author gender
ratio. This may suggest that men are more likely to preferentially seek
out male collaborators in fields where men are a minority, relative to
the homophily displayed by women in fields where women are a minority.
However, this latter result is only tentatively supported since our
sample contains few journals in which most authors are women
(\autoref{alpha_gender_ratio}).

We also found that gender homophily was marginally stronger in 2015-2016
relative to 2005-2006. Although this trend might reflect a change in the
gender preferences of researchers seeking collaborators, there are
alternative (and perhaps more likely) explanations. For example, this
trend might result from the increasing number of women working in senior
positions in STEMM over the past decade {[}63--65{]}. As shown in
\autoref{simulation}, if enough coauthorships are between junior and
senior researchers, a large gender gap between career stages can give
the appearance of heterophily. As this gender gap between career stages
lessens, the observed values of \(\alpha'\) may increase.

Regarding our finding of weaker homophily among 2-author papers, we
suspect that many 2-author teams comprise a student/postdoc and a senior
staff member, making these teams especially likely to be mixed-gender,
due to the elevated gender gap among senior researchers {[}1,5{]}.
Assuming this interpretation is correct, this result suggests that our
reported \(\alpha'\) values may underestimate the strength of peoples'
preferences for same-gendered collaborators; essentially, women seeking
a senior collaborator could be constrained to work mostly with men,
meaning that people's ideal and realised gender preferences would be
mismatched. On a related note, Ghiasi et al. {[}51{]} argue that women
in engineering are ``compliant {[}in reproducing{]} male-dominated
scientific structures'' because they do not collaborate often enough
with other women (their Figure 7 suggests that coauthorships between
women are 30\% more frequent than expected under random assortment). By
contrast, we feel that it is not helpful to recommend that women
collaborate primarily with other women, e.g.~because this constrains
women's options and may be counter-productive (particularly in fields
like engineering, where 90\% of professors are men {[}1{]}). Instead, we
suggest that researchers of both genders can help to close the gender
gap in STEMM. In the context of collaboration, one way to do this is to
undertake self-examination to ensure that one is not inadvertently
overlooking or excluding female potential students and colleagues. One
should also take care to treat male and female collaborators equally,
e.g.~in terms of training and mentoring, allocation of work, and how one
frames or promotes the collaboration (e.g.~in conference presentations
or on a website); evidence suggests that unconscious bias causes people
to undervalue women's research achievements {[}20{]}, and possibly to
assign menial or under-valued tasks to women and more prestigious tasks
to men {[}61{]}.

Our study begs two questions: what causes gender homophily in science,
and are our results cause for concern? These questions are closely
related. For example, some of the homophily we observed might be caused
by women seeking to avoid harassment or sexism from men {[}38{]}, which
would clearly be concerning. Additionally, Sheltzer and Smith {[}66{]}
concluded that `elite' male academics (defined as recipients of major
honours) have a higher proportion of male students and postdocs than
non-elite male academics. This finding could contribute to the homophily
we observed, and is cause for concern since Sheltzer and Smith
{[}66{]}'s results might reflect discrimination against women during
hiring {[}20{]}, or avoidance by women of elite research groups
(e.g.~due to gender differences in confidence, or a perception that some
groups are sexist). We also found a little evidence that gender
homophily is detrimental to research quality, in that high-impact
journals tended to have weaker homophily. Assuming that papers published
in high-impact journals are of higher average quality {[}67{]}, our
results provide non-experimental support for the hypothesis that
mixed-gender teams produce better research than single-gender teams
{[}42--48{]}. Another issue is that if many collaborations are between
established researchers, there will be an excess of male-male
collaborations in fields where women in senior positions are rare; some
of the observed homophily might therefore reflect the elevated gender
gap among senior researchers.

On the other hand, homophily might have more benign causes.
Collaboration is often most enjoyable and productive when working with
like-minded people, who might be same-gendered more often than not. We
also suppose that some people consciously choose to preferentially
collaborate with women in order to help close the gender gap in the
workforce; this would create homophily if women do this more than men.
In support of this interpretation, women appear more likely than men to
promote the work of female colleagues by inviting them to give talks
{[}68,69{]}. Given that many collaborative research projects
unfortunately involve a gendered division of labour {[}61{]}, working
with a same-gendered colleague may provide exposure to new parts of the
research process, and (especially for the minority gender) a welcome
change of pace.

\section{Methods}\label{methods}

\subsection{The dataset}\label{the-dataset}

We used the dataset of PubMed author lists from Holman et al. {[}5{]}.
Briefly, that dataset was created by downloading every article indexed
on PubMed and attempting to infer each author's gender from their given
name. Each journal was assigned to one of 107 scientific disciplines,
using PubMed's journal categorisations in the interests of objectivity.
Because the present study focuses on co-authorship, all single-author
papers were discarded. We also discarded all papers for which we could
not determine the gender of every author with \({\ge}95\%\) certainty,
in order to simplify the statistical analysis. To mitigate Wahlund
effects caused by variation in the gender ratio of researchers over time
(see below), we also discarded all papers except those that were
published either 0-1 or 10-11 years before the PubMed data were
collected (i.e.~20\textsuperscript{th} August 2016). Lastly, we excluded
journals with fewer than 50 suitable papers. Detailed sample size
information is given in S1 Table.

\subsection{\texorpdfstring{Calculating \(\alpha\), the coefficient of
homophily}{Calculating \textbackslash{}alpha, the coefficient of homophily}}\label{calculating-alpha-the-coefficient-of-homophily}

Following Bergstrom et al. {[}60{]}, we defined the coefficient of
homophily as \(\alpha = p - q\), where \(p\) is the probability that a
randomly-chosen co-author of a \emph{male} author is a man and \(q\) is
the probability that a randomly-chosen co-author of a \emph{female}
author is a man. Like the Wahlund effect, \(\alpha\) is borrowed from
population genetics; for a set of 2-author papers, it is equivalent to
Wright's coefficient of inbreeding {[}70{]}. Mathematical work
illustrates that \(\alpha\) is closely related to alternative
network-based methods for quantifying homophily {[}71{]}.

To estimate \(\alpha\) for a particular subset of the scientific
literature, we estimated \(p\) as the average proportion of men's
co-authors who are men (averaged across all papers with at least one man
author), and \(q\) as the average proportion of women's co-authors who
are men (averaged across all papers with at least one woman author). To
estimate the 95\% confidence intervals on \(\alpha\) for a given set of
\(n\) papers, we sampled \(n\) papers with replacement 1000 times,
estimated \(\alpha\) on each sample, and recorded the 95\% quantiles of
the resulting 1000 estimates.

As well as calculating \(\alpha\) for all authors, we calculated
\(\alpha\) for first or last authors only. \(\alpha\) was again defined
as \(p - q\), but this time \(p\) was estimated as the average
proportion of male co-authors on papers with a male first (or last)
author, and \(q\) was estimated as the average proportion of male
co-authors on papers with female first (or last) authors. We did not
calculate \(\alpha\) for other authorship positions (e.g.~second or
third authors) because this would necessitate culling the dataset to
include only papers with a sufficiently long author list, complicating
interpretation of the results.

We also calculated \(\alpha\) for papers with 2, 3, 4 or \({\ge}5\)
authors, for all journals that had at least 50 suitable papers from
2015-2016 with the specified author list length.

Our test assumes that the expected value of \(\alpha\) is zero if
authors randomly assort, but for small datasets this assumption is not
always true (as pointed out by Carl Bergstrom in a blog post,
\url{http://www.eigenfactor.org/gender/assortativity/note_to_eisen.rtf}).
To borrow Prof.~Bergstrom's example, consider a small research specialty
comprising just two men and two women researchers, who have together
produced six two-author papers: one in each of the six possible
two-author combinations. For these six papers,
\(\alpha = -\frac{1}{3}\), even though same- and opposite-gendered
coauthors were selected in equal proportion to their frequency in the
pool of possible collaborators.

To control for the fact that the null expectation for \(\alpha\) is not
zero for small datasets, we devised an adjusted version of the
coefficient of homophily, which we term \(\alpha'\). Every time we
calculated \(\alpha\) for a set of papers, we also determined the
expected value of \(\alpha\) under the null hypothesis that authors
assort randomly with respect to gender. This was accomplished by
randomly permuting authors across papers 1000 times, recalculating
\(\alpha\), and taking the median. We then calculated \(\alpha'\) by
subtracting the null expectation for \(\alpha\) from the observed value.
We also used the null-simulated \(\alpha\) values to calculate a
two-tailed p-value for the observed value of \(\alpha\); the p-value was
defined as the proportion of null simulations for which
\(|\alpha_{null}| > |\alpha_{obs}|\). We applied false discovery rate
(FDR) correction to each set of p-values to account for multiple testing
{[}72{]}.

As expected, \(\alpha'\) was usually almost identical to \(\alpha\) (S7
Fig), but \(\alpha\) was downwardly biased relative to \(\alpha'\) for
small datasets (S8 Fig). Additionally, the correlation between
\(\alpha'\) and sample size was negligible (\(R^2 < 0.01\)), suggesting
that our calculation of \(\alpha'\) effectively removed the dependence
of \(\alpha\) on sample size. We therefore used \(\alpha'\) in all
analyses.

\subsection{Minimising the Wahlund effect: research discipline and time
period}\label{minimising-the-wahlund-effect-research-discipline-and-time-period}

To minimise bias in \(\alpha'\) due to the Wahlund effect, we restricted
each set of papers to a single research specialty to the greatest extent
allowed by our data. Specifically, we only calculated \(\alpha'\) for
individual journals, since papers from the same journal typically focus
on closely related topics. Although some journals, e.g. \emph{PLoS ONE},
publish research from diverse disciplines with very different author
gender ratios {[}5{]}, calculating \(\alpha'\) for these highly
multidisciplinary journals is still useful as a contrast. The difference
in \(\alpha'\) between highly multidisciplinary and more specialised
journals, e.g. \emph{PLoS ONE} versus \emph{PLoS Computational Biology},
gives an estimate of the extent to which multidisciplinarity inflates
\(\alpha'\).

As well as varying between disciplines, the gender ratio of authors has
changed markedly over time {[}5{]}. Because the gender ratio was more
male-biased in the past, \(\alpha'\) would be inflated if we calculated
it for a sample of papers published over a long enough time frame. To
minimise this effect, we only sampled papers from two one-year periods
(namely 2005-6 and 2015-16). The median change per year in \% (fe)male
authors across journals is below 0.5\% {[}5{]}, and so restricting our
dataset to a single year should prevent temporal changes in gender ratio
from noticeably affecting our estimates of \(\alpha'\).

\subsection{Minimising the Wahlund effect: author country of
affiliation}\label{minimising-the-wahlund-effect-author-country-of-affiliation}

A Wahlund effect could arise even if one calculates \(\alpha'\) for a
single discipline and time period, because of variation in the gender
ratio of researchers from different countries. For example, Holman et
al. {[}5{]} showed that PubMed-indexed authors based in Serbia are more
than twice as likely to be women as are authors based in Japan.
Therefore, a dataset containing a mix of papers from teams of authors
based in these two countries would contain an excess of same-sex
coauthorships, even if collaboration were random with respect to gender
within each country.

To address this issue, we also analysed every combination of journal and
author country of affiliation for which we had enough data (i.e.~50 or
more papers published in 2015-16). For simplicity, we restricted the
dataset to only include papers for which Holman et al. {[}5{]} had
identified the country of affiliation for all authors on the paper, and
all authors shared the same country of affiliation. Restricting the
dataset in this fashion produced enough data to measure \(\alpha'\) for
325 combinations of journal and country (median: 70 papers and 273
authors per combination).

\subsection{Calculating standardised journal impact
factor}\label{calculating-standardised-journal-impact-factor}

We obtained the 3-year impact factor for each journal from Clarivate
Analytics. To account for large differences in impact factor between
disciplines, we took the the residuals from a model with \(Log_{10}\)
impact factor as the response and the research discipline of the journal
as a random effect. Thus, journals with a positive standardised impact
factor have a higher mean number of citations than the average for
journals in their discipline. We then used Spearman rank correlation to
test whether \(\alpha'\) was correlated with impact factor across
journals.

\subsection{Statistical analysis}\label{statistical-analysis}

Previous authors {[}66,73{]} have hypothesised that senior scientists
preferentially recruit staff and students of the same gender, and/or
that junior researchers preferentially select same-gendered mentors. In
the majority of PubMed-indexed disciplines, authorship conventions mean
that the first-listed author is often an early-career researcher, while
the author listed last is more likely to be a senior researcher leading
a research team {[}74{]}. Assuming that senior researchers are the main
drivers of homophily and that there are enough papers with three or more
authors, we predict that the last author's gender will be the strongest
predictor of the remaining authors' genders (i.e.~the gender of the last
author will be more salient than that of the first author, or any other
authorship position). This is because the first author's gender would
simply be an imperfect correlate of the true causal effect, while the
last author's gender would be the causal effect itself.

To test whether \(\alpha'\) for last authors tends to be higher than
\(\alpha'\) for first authors for any given dataset, we used a linear
mixed model implemented in the \texttt{lme4} and \texttt{lmerTest}
packages for R, with \emph{authorship position} (first or last) as a
fixed factor, and \emph{journal} and \emph{research discipline} as
crossed random effects. The response variable was \(\alpha'\), and we
weighted each observation by the inverse of the standard error from our
estimate of \(\alpha'\), meaning that more accurate measurements of
\(\alpha'\) had more influence on the results. We used a similar model
to test for a difference in \(\alpha'\) between the 2005-6 and the
2015-16 datasets, with two differences: we fit year range as a two-level
fixed factor (instead of authorship position), and we used \(\alpha'\)
estimated for all authors (not first/last authors) as the response
variable.

The relationship between the gender ratio of authors publishing in a
journal and its \(\alpha'\) value appeared nonlinear (see Results). We
therefore fit a generalised additive model with thin plate regression
spline smoothing, implemented using the \texttt{mgcv} package for R.

To model the relationship between \(\alpha'\) and the number of authors
on the paper, we used a meta-regression model implemented in the R
package \texttt{brms} {[}75{]}. The model incorporated the standard
error associated with easch estimate of \(\alpha'\), had author number
as a fixed effect, and journal as a random intercept (to control for
repeated measures of each journal). We also fit a random slope of author
number within journal, thereby allowing the response to author number to
vary between journals. We used the default (weak) priors. The full
output of this model can be viewed in the Online Supplementary Material.

\subsection{\texorpdfstring{Theoretical expectations for \(\alpha\) when
the gender ratio differs between career
stages}{Theoretical expectations for \textbackslash{}alpha when the gender ratio differs between career stages}}\label{theoretical-expectations-for-alpha-when-the-gender-ratio-differs-between-career-stages-1}

In many STEMM subjects, the gender ratio is more skewed among
established researchers relative to early-career researchers {[}1,5{]}.
We hypothesised that this skew could potentially create both Wahlund
effects and `reverse' Wahlund effects. For example, imagine that the
majority of collaborations are between students and professors, and that
the gender ratio differs between career stages: we will then see an
excess of mixed-gender coauthorships (heterophily, \(\alpha < 0\)), even
if gender has no direct, causal effect. Similarly, a hypothetical field
in which students work only with students, and professors with
professors, would have apparent gender homophily (\(\alpha > 0\)).

We can think of no tractable method of controlling for this issue using
our dataset, which contains no information on career stage. Therefore,
we instead decided to derive the theoretical expectations for \(\alpha\)
when there is a difference in gender ratio across career stages, in
order to determine if and how this effect should alter our inferences.
For simplicity, our calculations assume there are only two career
stages, though we intuit that the general conclusions would also apply
to a multi-tier career ladder. Under the null model that gender has no
causal effect on collaboration, we calculated \(\alpha\) for various
combinations of the four free parameters, i.e.~the gender ratios for
early- and late-career researchers, and the relative frequency of
collaborations between early-early, early-late, and late-late
collaborations. We then used the theoretical expectations for \(\alpha\)
to qualify our main conclusions (see Results). The Online Supplementary
Material gives annotated R code used to derive the theoretical
expectations.

\subsection{Data availability and
reproducibility}\label{data-availability-and-reproducibility}

The Online Supplementary Material contains R scripts used to produce all
results, figures and tables; it can be viewed online at
\url{https://lukeholman.github.io/genderHomophily/}. The input data from
Holman et al. {[}5{]} is archived at \url{https://osf.io/bt9ya/}.

\section{Acknowledgements}\label{acknowledgements}

CM was supported by the Academy of Finland (284666 to the Centre of
Excellence in Biological Interactions). We thank Devi Stuart-Fox and
Dominique Potvin for helpful discussion.

\section{References}\label{references}

\hypertarget{refs}{}
\hypertarget{ref-Shaw_2012}{}
1. Shaw AK, Stanton DE. Leaks in the pipeline: separating demographic
inertia from ongoing gender differences in academia. Proceedings of the
Royal Society of London B. 2012;272: 3736--3741.

\hypertarget{ref-Lariviere_2013}{}
2. Larivière V, Ni C, Gingras Y, Cronin B, Sugimoto CR. Bibliometrics:
global gender disparities in science. Nature. 2013;504: 211--213.

\hypertarget{ref-West_2013}{}
3. West JD, Jacquet J, King MM, Correll SJ, Bergstrom CT. The role of
gender in scholarly authorship. PLoS ONE. 2013;8: e66212.

\hypertarget{ref-Elsevier_report}{}
4. Elsevier Report. Gender in the global research landscape.
elseviercom/research-intelligence/resource-library/gender-report. 2017;

\hypertarget{ref-Holman_2018}{}
5. Holman L, Stuart Fox D, Hauser CE. The gender gap in science: How
long until women are equally represented? PLoS Biology. 2018;16:
e2004956.

\hypertarget{ref-Wutte_2007}{}
6. Wutte M. Closing the gender gap. Nature. 2007;448: NJ101--NJ102.

\hypertarget{ref-Reuben_2014}{}
7. Reuben E, Sapienza P, Zingales L. How stereotypes impair women's
careers in science. Proceedings of the National Academy of Sciences.
2014;111: 4403--4408.

\hypertarget{ref-Trower_2002}{}
8. Trower CA, Chait RP. Faculty diversity: Why women and minorities are
underrepresented in the professoriate, and fresh ideas to induce needed
reform. Harvard Magazine. 2002;104: 33--37.

\hypertarget{ref-Umbach_2007}{}
9. Umbach PD. Gender equity in the academic labor market: An analysis of
academic disciplines. Research in Higher Education. 2007;48: 169--192.

\hypertarget{ref-Hosek_2005}{}
10. Hosek S, Cox AG, Ghosh-Dastidar B, Kofner A, Ramphal N, Scott J, et
al. Gender differences in major federal external grant programs. RAND
Corporation. 2005;

\hypertarget{ref-pohlhaus_2011}{}
11. Pohlhaus JR, Jiang H, Wagner RM, Schaffer WT, Pinn VW. Sex
differences in application, success, and funding rates for NIH
extramural programs. Academic Medicine. 2011;86: 759.

\hypertarget{ref-Zuckerman_1987}{}
12. Zuckerman H. Persistence and change in the careers of men and women
scientists and engineers. National Academy Press. 1987; 127--156.

\hypertarget{ref-Rosenfeld_1991}{}
13. Rosenfeld RA. Outcome analysis of academic careers. Review prepared
for the Office of Scientific and Engineering Personnel, National
Research Council. 1991;

\hypertarget{ref-Long_1993}{}
14. Long JS, Paul DA, Robert M. Rank advancement in academic careers:
Sex differences and the effects of productivity. American Sociological
Review. 1993; 703--722.

\hypertarget{ref-Hopkins_2013}{}
15. Hopkins AL, Jawitz JW, McCarty C, Goldman A, Basu NB. Disparities in
publication patterns by gender, race and ethnicity based on a survey of
a random sample of authors. Scientometrics. 2013;96: 515--534.

\hypertarget{ref-ODorchai_2009}{}
16. O'Dorchai S, Meulders D, Crippa F, Margherita A. She figures
2009--Statistics and indicators on gender equality in science.
Publications Office of the European Union. 2009;

\hypertarget{ref-Feldt_1986}{}
17. Feldt B. The faculty cohort study: School of medicine. Ann Arbor,
MI: Office of Affirmative Action. 1986;

\hypertarget{ref-Stack_2004}{}
18. Stack S. Gender, children and research productivity. Scientometrics.
2004;45: 891--920.

\hypertarget{ref-Lariviere_2011}{}
19. Larivière V, Vignola-Gagné E, Villeneuve C, Gélinas P, Gingras Y.
Sex differences in research funding, productivity and impact: an
analysis of Québec university professors. Scientometrics. 2011;87:
483--498.

\hypertarget{ref-Moss_2012}{}
20. Moss-Racusin CA, Dovidio JF, Brescoll VL, Graham MJ, Handelsman J.
Science faculty's subtle gender biases favor male students. Proceedings
of the National Academy of Sciences. 2012;109: 16474--16479.

\hypertarget{ref-Knobloch_2013}{}
21. Knobloch-Westerwick S, Glynn CJ, Huge M. Science faculty's subtle
gender biases favor male students. Science Communication. 2013;35:
603--625.

\hypertarget{ref-Lee_2005}{}
22. Lee S, Bozeman B. The impact of research collaboration on scientific
productivity. Social Studies of Science. 2005;35: 673--702.

\hypertarget{ref-Wuchty_2007}{}
23. Wuchty S, Jones BF, Uzzi B. The increasing dominance of teams in
production of knowledge. Science. 2007;316: 1036--1039.

\hypertarget{ref-Abramo_2009}{}
24. Abramo G, D'Angelo CA, Di Costa F. Research collaboration and
productivity: is there correlation? Higher Education. 2009;57: 155--171.

\hypertarget{ref-Lariviere_2015}{}
25. Larivière V, Gingras Y, Sugimoto CR, Tsou A. Team size matters:
Collaboration and scientific impact since 1900. Journal of the
Association for Information Science and Technology. 2015;66: 1323--1332.

\hypertarget{ref-Long_1992}{}
26. Long JS. Measures of sex differences in scientific productivity.
Social Forces. 1992;71: 159--178.

\hypertarget{ref-Bozeman_2011}{}
27. Bozeman B, Monica G. How do men and women differ in research
collaborations? An analysis of the collaborative motives and strategies
of academic researchers. Research Policy. 2011;40: 1393--1402.

\hypertarget{ref-Abramo_2013}{}
28. Abramo G, D'Angelo CA, Di Costa F. Gender differences in research
collaboration. Journal of Informetrics. 2013;7: 811--822.

\hypertarget{ref-Badar_2013}{}
29. Badar K, Hite JM, Badir YF. Examining the relationship of
co-authorship network centrality and gender on academic research
performance: The case of chemistry researchers in pakistan.
Scientometrics. 2013;94: 755--775.

\hypertarget{ref-Lewison_2001}{}
30. Lewison G. The quantity and quality of female researchers: A
bibliometric study of Iceland. Scientometrics. 2001;52: 29--43.

\hypertarget{ref-Webster_2001}{}
31. Webster BM. Polish women in science: A bibliometric analysis of
Polish science and its publications. Research Evaluation. 2001;10:
185--194.

\hypertarget{ref-Bozeman_2004}{}
32. Bozeman B, Corley E. Scientists' collaboration strategies:
implications for scientific and technical human capital. Research
Policy. 2004;33: 599--616.

\hypertarget{ref-Long_1990}{}
33. Long JS. The origins of sex differences in science. Social Forces.
1990;68: 1297--1316.

\hypertarget{ref-Fuchs_2001}{}
34. Fuchs S, Von Stebut J, Allmendinger J. Gender, science, and
scientific organizations in Germany. Minerva. 2001;39: 175--201.

\hypertarget{ref-Reskin_1978}{}
35. Reskin BF. Scientific productivity, sex, and location in the
institution of science. American Journal of Sociology. 1978;83:
1235--1243.

\hypertarget{ref-Wright_2003}{}
36. Wright AL, Schwindt LA, Bassford TL, Reyna VF, Shisslak PAS
Catherine M amd Germain, Reed KL. Gender differences in academic
advancement: Patterns, causes, and potential solutions in one U.S.
college of medicine. Social Forces. 2003;68: 1297--1316.

\hypertarget{ref-bleidorn2016age}{}
37. Bleidorn W, Arslan RC, Denissen JJ, Rentfrow PJ, Gebauer JE, Potter
J, et al. Age and gender differences in self-esteem -- a cross-cultural
window. Journal of Personality and Social Psychology. 2016;111: 396.

\hypertarget{ref-jagsi_2016}{}
38. Jagsi R, Griffith KA, Jones R, Perumalswami CR, Ubel P, Stewart A.
Sexual harassment and discrimination experiences of academic medical
faculty. JAMA. 2016;315: 2120--2121.

\hypertarget{ref-Martin_2014}{}
39. Martin JL. Ten simple rules to achieve conference speaker gender
balance. PLoS computational biology. 2014;10: e1003903.

\hypertarget{ref-Tower_2007}{}
40. Tower G, Julie P, Brenda R. A multidisciplinary study of
gender-based research productivity in the world's best journals. Journal
of Diversity Management. 2007;2: 23--32.

\hypertarget{ref-Jordan_2008}{}
41. Jordan CE, Clark SJ, Vann CE. Do gender differences exist in the
publication productivity of accounting faculty?. Journal of Applied
Business Research. 2008;24: 77--85.

\hypertarget{ref-Britton_2000}{}
42. Britton DM. The epistemology of the gendered organization. Gender
and Society. 2000;14: 418--434.

\hypertarget{ref-Reagans_2001}{}
43. Reagans R, Zuckerman EW. Networks, diversity, and productivity: The
social capital of corporate R\&D teams. Organization Science. 2001;12:
502--517.

\hypertarget{ref-Hong_2004}{}
44. Hong L, Page SE. Groups of diverse problem solvers can outperform
groups of high-ability problem solvers. Proceedings of the National
Academy of Sciences. 2004;101: 16385--16389.

\hypertarget{ref-Whittington_2008}{}
45. Whittington KB, Smith-Doerr L. Women inventors in context:
Disparities in patenting across academia and industry. Gender \&
Society. 2008;22: 194--218.

\hypertarget{ref-Bear_2011}{}
46. Bear JB, Woolley AW. The role of gender in team collaboration and
performance. Interdisciplinary Science Reviews. 2011;36: 46--153.

\hypertarget{ref-Herrera_2012}{}
47. Herrera R, Duncan PA, Green MT, Skaggs SL. The effect of gender on
leadership and culture. Global Business and Organizational Excellence.
2012;31: 37--48.

\hypertarget{ref-Campbell_2013}{}
48. Campbell LG, Mehtani S, Dozier ME, Rinehart J. Gender-heterogeneous
working groups produce higher quality science. PloS ONE. 2013; e79147.

\hypertarget{ref-Ferber_1980}{}
49. Ferber MA, Teiman M. Are women economists at a disadvantage in
publishing journal articles? Eastern Economic Journal. 1980;6:
1189--193.

\hypertarget{ref-McDowell_1992}{}
50. McDowell JM, Smith JK. The effect of gender-sorting on propensity to
coauthor: Implications for academic promotion. Economic Inquiry.
1992;30: 68--82.

\hypertarget{ref-Ghiasi_2015}{}
51. Ghiasi G, Larivière V, Sugimoto CR. On the compliance of women
engineers with a gendered scientific system. PloS ONE. 2015;10:
e0145931.

\hypertarget{ref-Crow_2015}{}
52. Crow MS, Smykla JO. An examination of author characteristics in
national and regional criminology and criminal justice journals,
2008-2010: Are female scholars changing the nature of publishing in
criminology and criminal justice? American Journal of Criminal Justice.
2015;40: 441--455.

\hypertarget{ref-Fahmy_2017}{}
53. Fahmy C, Young JT. Gender inequality and knowledge production in
criminology and criminal justice. Journal of Criminal Justice Education.
2017;28: 285--305.

\hypertarget{ref-Zettler_2017}{}
54. Zettler HR, Stephanie M Cardwell, Jessica M C. The gendering effects
of co-authorship in criminology \& criminal justice research. Criminal
Justice Studies. 2017;30: 30--44.

\hypertarget{ref-Jadidi_2017}{}
55. Jadidi M, Karimi F, Lietz H, Wagner C. Gender disparities in
science? Dropout, productivity, collaborations and success of male and
female computer scientists. Advances in Complex Systems. 2017; 1750011.

\hypertarget{ref-Teele_2017}{}
56. Teele DL, Kathleen T. Gender in the journals: Publication patterns
in political science. PS: Political Science \& Politics. 2017;50:
433--447.

\hypertarget{ref-Araujo_2017a}{}
57. Araújo T, Elsa F. The specific shapes of gender imbalance in
scientific authorships: a network approach. Journal of Informetrics.
2017;11: 88--102.

\hypertarget{ref-Araujo_2017b}{}
58. Araújo T, Elsa F. Big Missing Data: are scientific memes inherited
differently from gendered authorship? arXiv preprint arXiv. 2017;
1706.05156.

\hypertarget{ref-Wahlund_1928}{}
59. Wahlund S. Zusammensetzung von populationen und
korrelationserscheinungen vom standpunkt der vererbungslehre aus
betrachtet. Hereditas. 1928;11: 65--106.

\hypertarget{ref-bergstrom_2016}{}
60. Bergstrom T, Bergstrom C, King M, Jacquet J, West J, Correll S. A
note on measuring gender homophily among scholarly authors.
http://eigenfactororg/gender/ assortativity/measuring\_homophilypdf.
2016;

\hypertarget{ref-macaluso_2016}{}
61. Macaluso B, Larivière V, Sugimoto T, Sugimoto CR. Is science built
on the shoulders of women? A study of gender differences in
contributorship. Academic Medicine. 2016;91: 1136--1142.

\hypertarget{ref-Bentley_2003}{}
62. Bentley JT, Adamson R. Gender differences in the careers of academic
scientists and engineers: A literature review. Special Report. 2003;

\hypertarget{ref-Long_2015}{}
63. Long MT, Leszczynski A, Thompson KD, Wasan SK, Calderwood AH. Female
authorship in major academic gastroenterology journals: A look over 20
years. Gastrointestinal Endoscopy. 2015;81: 1440--1447.

\hypertarget{ref-Bendels_2018}{}
64. Bendels MH, Bauer J, Schöffel N, Groneberg DA. The gender gap in
schizophrenia research. Schizophrenia Research. 2018;193: 445--446.

\hypertarget{ref-McKenzie_2017}{}
65. McKenzie K, Ramonas M, Patlas M, Katz DS. Assessing the gap in
female authorship in the journal emergency radiology: Trends over a
20-year period. Emergency Radiology. 2017;24: 641--644.

\hypertarget{ref-sheltzer_2014}{}
66. Sheltzer JM, Smith JC. Elite male faculty in the life sciences
employ fewer women. Proceedings of the National Academy of Sciences.
2014;111: 10107--10112.

\hypertarget{ref-Garfield_2006}{}
67. Garfield E. The history and meaning of the journal impact factor.
JAMA. 2006;295: 90--93.

\hypertarget{ref-nittrouer_2018}{}
68. Nittrouer CL, Hebl MR, Ashburn-Nardo L, Trump-Steele RC, Lane DM,
Valian V. Gender disparities in colloquium speakers at top universities.
Proceedings of the National Academy of Sciences. 2018;115: 104--108.

\hypertarget{ref-debarre_2018}{}
69. Débarre F, Rode N, Ugelvig L. Gender equity at scientific events.
Evolution Letters. 2018;in press: doi:10.1002/evl3.49.

\hypertarget{ref-wright_1949}{}
70. Wright S. The genetical structure of populations. Annals of Human
Genetics. 1949;15: 323--354.

\hypertarget{ref-wang_2016}{}
71. Wang YS, Erosheva EA. On the relationship between set-based and
network-based measures of gender homophily in scholarly publications.
arXiv preprint arXiv:161009026. 2016;

\hypertarget{ref-Benjamini_1995}{}
72. Benjamini Y, Hochberg Y. Controlling the false discovery rate: a
practical and powerful approach to multiple testing. Journal of the
Royal Statistical Society: Series B. 1995; 289--300.

\hypertarget{ref-Bonham_2017}{}
73. Bonham KS, Stefan MI. Women are underrepresented in computational
biology: An analysis of the scholarly literature in biology, computer
science and computational biology. PLoS Computational Biology. 2017;13:
e1005134.

\hypertarget{ref-Wren_2007}{}
74. Wren JD, Kozak KZ, Johnson KR, Deakyne SJ, Schilling LM, Dellavalle
RP. The write position: A survey of perceived contributions to papers
based on byline position and number of authors. EMBO reports. 2007;8:
988--991.

\hypertarget{ref-burkner_2016}{}
75. Bürkner P-C. brms: An R package for Bayesian multilevel models using
Stan. Journal of Statistical Software. 2016;80: 1--28.

\newpage

\section{Supporting information}\label{supporting-information}

\subsection{Supplementary figures}\label{supplementary-figures}

S1 Fig. Plot showing the percentage of papers that have 1, 2, 3, 4, or
\({\ge}5\) authors for each discipline in the dataset of Holman et al.
(2018). This information can also be found in S3 Data.

S2 Fig. Histogram showing the distribution of differences in \(\alpha'\)
between the 2015-16 and 2005-6 samples, where positive numbers indicate
an increase in \(\alpha'\) with time. The mean is slightly positive
(i.e.~0.004), indicating a mild increase in average \(\alpha'\) with
time.

S3 Fig. Histogram showing the difference between \(\alpha'\) calculated
for first and last authors. Positive values mean that \(\alpha'\) was
higher when calculated for first authors, and negative values mean
\(\alpha'\) was higher when calculated for last authors. The mean is
very slightly higher than zero, indicating that \(\alpha'\) tends to be
higher for first authors.

S4 Fig. Histogram of \(\alpha'\) for 325 unique combinations of journal
and country, using data from August 2015 - August 2016. The white areas
denote combinations for which \(\alpha'\) differs significantly from
zero (p \textless{} 0.05, following false discovery rate correction).

S5 Fig. Plot showing the 68 combinations of journal and author country
of affiliation for which \(\alpha'\) is significantly higher than
expected.

S6 Fig. Histogram showing the estimated degree to which \(\alpha'\) is
inflated by inter-country differences in author gender ratio, across the
285 journals for which we had adequate data after restricting the
analysis by country. The average inflation in \(\alpha'\) is negligible,
suggesting that Wahlund effects resulting from inter-country differences
have a neglible effect on our estimates of gender homophily.

S7 Fig. There is a very strong correlation between the values of
\(\alpha\) and \(\alpha'\) calculated for each journal, though in a
handful of cases the difference is considerable. The deviation between
\(\alpha\) and \(\alpha'\) is greatest for journals for which there is a
small sample size (see S8 Fig).

S8 Fig. For journals for which we recovered a small number of papers
(\textless{}100), the unadjusted metric \(\alpha\) was downwardly
biased. This fits our expectations: because authors cannot be their own
co-authors, small datasets will tend to produce negative estimates of
\(\alpha\) even if authors assort randomly with respect to gender (see
main text). This suggests that \(\alpha'\) is a more useful measure of
homophily and heterophily, especially for small samples.

\subsection{Supplementary tables}\label{supplementary-tables}

S1 Table. Sample sizes for the two datasets, which comprise papers
published in the timeframes August 2005 - August 2006, and August 2015 -
August 2016.

S2 Table. Number of journals showing statistically significant homophily
or heterophily, in two one-year periods. The significance threshold was
p \textless{} 0.05, and p-values were adjused using Benjamini-Hochberg
false discovery rate correction. Note that the power of our test is
lower for the 2005-2006 data because fewer papers were recovered per
journal: thus, it is not meaningful to compare the \% significant
journals (i.e.~11\% vs 24\%) between the two time periods.

\subsection{Supplementary datasets}\label{supplementary-datasets}

S1 Data: This spreadsheet shows the \(\alpha\) values calculated for
each journal, in the 2005 and 2015 samples, and for each type of author
(all authors, first authors, and last authors). The tables gives the
impact factor of each journal, the sample size, \(\alpha\) and
\(\alpha'\) and their 95\% CIs, and the p-value from a 2-tailed test
evaluating the null hypothesis that \(\alpha\) is zero (both raw and
FDR-corrected p-values are shown).

S2 Data: This file gives the number and percentage of paper that have 1,
2, 3, 4, or \({\ge}5\) authors for each \emph{journal} in the dataset of
Holman et al. (2018) \emph{PLoS Biology}. Note that the sample sizes
include papers for which the gender of one or more authors was not
determined by Holman et al.

S3 Data: This file gives the number and percentage of paper that have 1,
2, 3, 4, or \({\ge}5\) authors for each \emph{discipline} in the dataset
of Holman et al. (2018) \emph{PLoS Biology}. Note that the sample sizes
include papers for which the gender of one or more authors was not
determined by Holman et al.

S4 Data. The table shows the distribution of the \(\alpha'\) values
across journals, split by the research discipline. The gender ratio
column shows the percentage of women authors in the sample used to
calculate \(\alpha'\), across all authorship positions. In the last two
columns, the numbers outside parentheses give the number of journals
that deviate statistically significantly from zero, while the numbers
inside parentheses give the number that remain significant after false
discovery rate correction.

\end{document}

### 2018-10-24 manuscript/Response_letter.Rmd
Luke's changes after PLOS ONE review

---
title: "Response to PLOS ONE decision letter"
author: "Luke Holman and Claire Morandin"
output:
  pdf_document: 
    keep_tex: false
fontsize: 12pt
header-includes:
  - \usepackage{microtype}
---

## Comments from the Editor

Generally speaking, this manuscript brings a relevant contribution and is technically sound, but some additional effort is needed in terms of framing and result interpretation. In addition to comments made by Reviewer 1 (especially the one related to the meaning of ‘preference’), please address the following issues:

1.- The motivation and contributions of this manuscript against the state of the art should be stated in a clearer way in the introduction. The article proposes a methodology to control the "Wahlund effect" in gender studies applied to scientific collaboration. What are the aspects of scientific collaboration potentially producing such an effect and, consequently, being the focus of this article? (e.g. different career stages and countries, heavily gender-biased disciplines ..).

**Firstly, many thanks for your feedback and your attention to our manuscript, and this resubmission.**

**As noted by Reviewer 1, we need to be quite careful speculating as to the causes of 'gender homophily', because we have no direct data on these causes. The main purpose of this paper was to establish where homophily really exists as has been claimed (and measure it precisely in many fields of STEM), or whether it disappears once one accounts for statistical artefacts (what we call the Wahlund effect). There are many reasons why same-gender collaborations might be disproportionately common (besides artefacts), and we discussed many of these in the paper (a few in the Introdution, and several more in the Discussion). The most obvious one is that people prefer to work with others who are similar to themselves, and/or they prefer to avoid opposite-sex collaborators. It's also not clear whether this hypothetical preference is predominantly expressed by men, by women, or both. We could get into this more if you like, but we note that reviewer 1 already thought our cautious treatment was a little too much, and they disliked  the paragraph in which we speculated about the individual-level processes that drive the population-level patterns of gender assortment that we observed. We are quite happy with the balance between speculation and caution as it is, but we are open to specific suggestions.**

2.- Please, motivate the need to explore the correlation between $\alpha'$ and the number of authors.

**We have now expanded this section to provide some motivation, which was previously left until the Discussion. Basically, we are trying to extract information about the processes that cause non-random assortment by gender, and responding to a colleague's comments on an earlier draft which specifically requested this figure. Here is the revised section of the Results:**

***One possible explanation for this finding is that 2-authors papers might be more likely to have an author list that is evenly split between career stages (e.g. a postgraduate student and their supervisor), increasing the chance that the authors are mixed gender (see Figure 6). The result also suggests that the processes responsible for gender homophily are similar in small (e.g. 3 author) and larger (5+ authors) collaborations (and across disciplines where small versus large collaborations are the norm).***

3.- The article reports a statistically significant linear relationship between standardized journal impact factor and $\alpha'$. However, there is a lot of noise and the slope of the corresponding regression is approx. 0.012. This looks like a rather weak relationship. Does it really worth being stressed so much?

**We don't think that we stress this result much. The impact factor result is only discussed in two very cautious sentences in the Discussion (cautious words shown in bold):**

We also **found a little evidence** that gender homophily is detrimental to research quality, in that high-impact journals **tended to have** weaker homophily. **Assuming** that papers published in high-impact journals are of higher average quality **(which is contentious; [67])**, our results **provide non-experimental support for the hypothesis** that mixed-gender teams produce better research than single-gender teams [42–48].

**Please let us know if any of that still seems excessive upon a second reading. Also, a slope of -0.012 is actually quite a lot (perhaps you thought the slope was a correlation coefficient?). The mean $\alpha'$ across journals is about 0.07, and so this slope means that increasing the journal impact factor by one standard deviation above the mean decreases $\alpha'$ from about 0.070 to 0.058; in proportional terms, this is 0.058 / 0.07 = 83% of the mean. The $R^2$ is 4%, meaning that if we know the journal impact factor, we can explain about 4% of the variance in $\alpha'$, which is not huge, but is perhaps bigger than one would expect given the large number of unmeasured covariates that also affect the two variables being correlated here. At any rate, we are very upfront in saying the relationship is noisy, we give all the statistics and confidence intervals needed to interpret the result, and we are careful to remind readers that correlation is not causality.**

4.- What does ‘significant heterogeneity’ mean in line 141? I guess they are located above the 95% CI, but please specify.

**It says ‘significant heterophily’ (not heterogeneity), meaning that the estimate of $\alpha$ is negative and its 95% CIs are also negative. We explained how we defined statistical significant in the Methods (page 15); essentially, we define $\alpha$ values that we are >95% certain are non-zero as statistically significant.**

## Comments from Reviewer 1

This study investigates correlations among the genders of authors of papers, using authorship data extracted from the Pubmed database. Prior studies have already shown that there is homophily among authors, i.e. that publications include more authors of the same gender than expected under random assortment of authors. However, homophily may be the result of hidden subdivision of the data, for instance if different subfields have different male:female ratios. The authors call this a "Wahlund effect", by analogy with the effect of population subdivision on homozygozity in population genetics.

The study is overall well-written and well-referenced, as far as I can tell. I would also like to congratulate the authors for providing the entire R code behind their study, in an amazingly well-commented file. This is still too rare and I surely hope that this will become a standard.

That said, I have a few comments on the manuscript. They are presented below as they appear in the manuscript, and I highlight just below the most important ones.

**You have our sincere thanks for your time and the very useful feedback, as well as these compliments.**

### Main comments

1) It is an issue common to all these "big data" studies, but the limitation should be better highlighted in the study. You are considering the lists of authors of articles in different journals, but you do not have information on individuals themselves. On of the main issues to identify the production of individuals is the presence of homonyms, and different ways to identify an author in a paper (e.g., F. Lastname vs Firstname Lastname).
Consequently, the results are about the correlation among genders of authors of papers, but not directly about individual preference.
For instance, you may have some individuals who are over-represented in your dataset (very prolific authors -- or powerful heads of institutions adding their names everywhere), and their presence on different papers/journals may be counted as independent units -- while they are not.
I am aware that the problem cannot be solved with ISI or Pubmed data, but it really needs to be acknowledged.

**We did not knowingly make any claims that pertain to individuals: we are well aware that because our study focuses on population-level patterns of assortment by gender, we can only speculate about the individual-level processes that produce the population-level patterns. Therefore, the fact that we cannot (or rather, did not try to) identify specific individuals does not compromise the paper's results and conclusions. We agree that some uses of the word 'preference' in our manuscript could be misleading in this respect, and so in the revision, we have replaced the word in key place with more neutral/agnostic terms (e.g. 'non-random assortment').**

I do not think that the authors can conclude about individual preferences. As explained above, the study is not at the scale of individuals, and I fear that the over-representation vs under-representation of some authors may play a part in the results. However, it is entirely fair to conclude about list of authors being more same-gendered than expected by chance.

**See previous comment; we agree that we cannot make firm individual-level inferences. However, we believe that the population-level patterns almost certainly result from individuals' decisions and choices (even though these choices are constrained, e.g. by the shortage of women in supervisory roles, minor taboos against opposite-sex collaboration in some countries, etc), and so we believe it is important to speculate about the causal processes (while making it clear what is speculation and what is demonstrable fact). In the revision, we also take extra care to avoid possible sources of reader confusion, given that this is a contentious topic and the paper is quite statistics-heavy. For example, we point out that individual people with all-male collaborators are not necessarily picking their collaborators by sex, since thousands of people would have all-male collaborators even in a world with totally random assortment (especially in mostly-male fields).**

In addition, I'd advise against the use of "preference/preferentially", because it may imply a choice. The correlation can however be a by-product of other factors, and the data are not about actual preference (while it would for instance if the authors had been surveyed).

**Agreed: this word can denote active/conscious choice of collaborators based on gender, which is only one possible mechanism that might produce homophily. This word does not properly capture our intended meaning (i.e. "greater frequency than expected by chance"), and so we have re-written the manuscript with this in mind.**

2) It is almost impossible to properly understand the part on theoretical explanations with the information provided in the manuscript. While it is nice to provide the R code, the reader should be able to have a rather precise idea of what you did without having to dive into the R code. The corresponding subsection in the Methods should be expanded and much more detailed.

**We have now expanded the methods, and greatly expanded the supplementary material associated with Figure 6. See our response to this point in the response letter below.**

3) An important motivation for the study was to control for a Wahlund effect in this kind of study. It would be great to actually illustrate this feature by providing the estimate for alpha' when all journals are pooled together (otherwise, this seems to be only a replication of previous studies).

**Firstly, we took the reviewer's recommendation and calculated $\alpha$ for our entire PubMed dataset (all journals and all 15 years; n = >3 million papers, >16 million authors), using the same methods as were used for a single journal. We found that $\alpha = 0.126$, which is almost twice as high as the median $\alpha$ for individual journals ($\alpha = 0.070$, shown in top left of Figure 2). This suggests that indeed, $\alpha$ can be inflated if one lumps together data from very different research disciplines and time periods, as shown in Figure 1. However, $\alpha$ remained positive even though we controled for the Wahlund effect far more stictly than in all previous studies, suggesting that those studies are likely to be broadly correct (though they might over-estimate the strength of homophily). We now discuss this new result in the first paragraph of the Results section.**

**Secondly, our study is not a simple replication of previous work, whatever the value of $\alpha'$ pooled across all the journals. To our knowledge it is the only study to explain that statistical artefacts can theoretically produce 'spurious' gender homophily (and probably did, in some of the previous studies). Additionally, ours is the largest study of gender homophily, and the broadest in the set of research disciplines it covers. Also, we selected PLOS ONE partly because its editorial policy does not penalise replication studies or papers with non-novel results.**

### Comments as they appear in the manuscript
(I am also mentioning typos since PLOS One papers are just put in the PLOS format but not properly copy-edited)

#### Title and Abstract
Flags raised while reading:
  - Meaning of preferentially: actual preference (active behavior) or other factors also playing a role? (see point 1) above)

**See our response above**

- "High impact factor" is this a proxy for quality? (the point is addressed later in the discussion)

**No; we agree journal impact factor is a poor proxy for quality. A LaTeX error in the original submission prevented part of the Discussion from appearing: it was supposed to read "Assuming that papers published in high-impact journals are of higher average quality (which is contentious; [16])...". Originally, the "contentious" part did not format properly.**

#### Introduction
- line 62 please spell out "m" (not sure it's international notation since m is meters)

**We now write 'million' instead of m.**

- line 77 please spell out FDR the first time you use it (since the methods are currently at the end of the manuscript)

**Done**

- line 89 Author order differs across disciplines, e.g. it is alphabetical in maths. Please specify that you focus on life-sciences journals.

**We already specify that PubMed focuses on life sciences journals, as the reviewer notes elsewhere. Also, the presence of some alphabetically-ordered papers in the set would not alter the fact that the average seniority of last authors is greater than that of first authors, in most/all of the journals we studied.**

#### Methods
- line 265 Please specify how gender was assessed (even though this may already be in [5]). In particular, was it done automatically or manually? There may be issues with automatic gender assignment for authors of different geographic origins.

**We prefer to leave the full details to our previous paper, which has extensive descriptions, two different quality controls, and the full R code that was used to programmatically identify gender using authors' first names. Given our sample size (36 million authors), it hopefully goes without saying that automatic methods were used, but we now point this out explicitly in the Methods. Our earlier paper has extensive discussion of the geographic origin issue, and success rates for each of >100 different countries (almost all of which are over 70%, with China and Korea being the only significant exceptions among the major research-producing nations).**

- line 271 Maybe it is clearer to say what you kept rather than what you discarded (avoid double negations)

**Fixed, thanks.**

- lines 296-303 Please better explain where the effect comes from: you cannot be your own co-author. If there are m men and f women, a man can be a co-author with (m-1) men while women can be a co-author with m men. If you could be your own co-author, there would be 10 possibilities in your 2-author examples, and alpha would be 0.

**Changed as suggested, thanks.**

- line 322 Now it's PLOS (not PLoS anymore)

**Changed throughout**

- line 326 cf. point 3) above

**We addressed this by calculating $\alpha$ across the whole of PubMed; see above.**

- line 352 log (no capital L)

**Done**

- line 362 Why would senior authors be the main drivers? It could as well be junior authors who chose their advisor...

**Actually, the purpose of the first author vs last author analysis was to see if we can get some clues as to whether homophily is driven mostly by senior or junior scientists. We discussed both possibilities. I think you were misled because the two sentences beginning on line 362 use the word "Assume" rhetorically; the sentences basically say, "Assuming senior researchers drive the results, we predict X. This is because of Y." We do not start from the assumption that senior researchers are responsible for homophily. We also began that paragraph by pointing out that the 'senior researchers drive homophily' hypothesis is already common in the literature (to our knowledge it has not been effectively tested before).**

- lines 397-407: Where do your data come from? actual proportions of women among early vs late career researchers?

**There is no data - it's a theoretical model, as explained in the section referenced here (e.g. the title says "Theoretical expectations"). We investigated all combinations of values from 0% - 60% women, which spans >95% of journals on PubMed (see Holman et al. 2018).**

#### Results
  - Figure 2:
  -- please use transparent shade so that the curve behind remains visible when it is lower than the curve in the front.

  **This is a stacked density plot, so there is no curve behind the other one (the blue is stacked over the white). We have clarified this in the figure legend.**

-- Right-hand side figure: what do the dots correspond to? a single journal?

**Yes -- we now state this in the figure legend.**

  - Do you identify individual authors/ Control for the fact that some are more prolific than others?

**No, but this would not affect the veracity of our findings. Let's imagine that 90% of all papers in the world had one very prolific individual in their author lists. Let's assume this author is a man, though the argument would be the same if they were a woman. The existence of this prolific man would increase the proportion of male co-authors for both sexes, since most people have co-published with him. However, it would not affect our test statistic $\alpha'$, because the proportion of male coauthors would be elevated equally for men and women (and so $\alpha'$ would not change). You might ask, what if the super-author has preference for one gender of coauthors? But then, $\alpha'$ would be working as intended: it would correctly record that homophilous (or heterophilous) co-authorships were more common than expected under random assortment by gender. This highlights a feature of $\alpha'$ noted earlier by the reviewer: it is a population-level measure, so we cannot make statements about individuals (e.g. we don't know if everyone is a little bit homophilous, or if there are a few prolific and very homophilous people).

**Additionally, the basic version of $\alpha$, proposed in a short pre-print by Bergstrom et al., would be downwardly-biased (basically, because our hypothetical super-author cannot co-author papers with himself). We developed $\alpha'$ in order to get rid of this issue, though in practice the two versions of alpha were close to identical (see Figure S8), except for a few very small subsets of the data, so the basic $\alpha$ is fine most of the time.**

  - Theoretical expectations: it is not possible to understand what has been done in this section. The methods section about this part is too limited; the reader is asked to go and check an annotated R code; we should be able to have at least an idea of what has been done. As a result, it is not really possible to fully understand Figure 6, and in particular, the differences between columns in the figure. Also, we'd need details about the choices made regarding the list of authors, such as the number of authors.

**We have expanded the methods in the paper so that this part is clear without reading the supplement, as well as expanding the supplement to make it legible to people who have never seen R code. We also added much clearer variable names to the R code, and added extra comments to the code to make it easier to follow. For your convenience, here is the revised paragraph of the methods describing this theoretical exercise:**

***We can think of no tractable method of controlling for this issue using our dataset, which contains no information on career stage. Therefore, we instead decided to derive theoretical expectations for $\alpha$ when there is a difference in gender ratio across career stages, in order to determine if and how this effect should alter our inferences. For simplicity, our calculations assume there are only two career stages ('early-career' and 'established'), though we expect that the general conclusions would also apply to a multi-tier career ladder. Under the null model that gender has no causal effect on collaboration, we calculated $\alpha$ for various combinations of the four free parameters in our simple model. These parameters are: the gender ratio among early-career researchers (x-axis of Figure 6), the gender ratio among established researchers (y-axis of Figure 6), the frequency of within- versus between career stage collaborator pairs (rows in Figure 6), and lastly the frequency of within-stage collaborations that are between two early-career researchers as opposed to two late-career reserachers (columns in Figure 6). When these four parameters are specified, one can easily calculate the relative frequencies of collaborator pairs that involve two men, two women, or a man and a woman. In short, if we have specified the frequency of women at both career stages, as well as the frequency of the three possible types of collaboration with respect to career stage (early-early, early-established, and established-established), then we can calculate the frequency of collaborators pairs comprising two women, or a woman and a man, and thus find $\alpha$ (see the Online Supplementary Material for the annotated R code).***

#### Discussion
- line 173 : "We found evidence that researchers preferentially publish with same-gendered coauthors" I do not think that this conclusion is adequate given your data. You found evidence that there is a correlation between the genders of authors, but your data does not provide evidence regarding the behavior of individual authors.
Even if there were also such a correlation at the individual level, your data does not allow you to identify the cause of this correlation -- a preference (implying some choice) or a by-product of other constraints.

**We originally used the word 'preference' a bit incautiously, and now use terms like 'assortment' that are more agnostic about the mechanism. We do think individual-level preferences are likely to be partly responsible, however, and so we discuss preference in a few places. This seems like an uncontroversial hypothesis to explain our results.**

- line 176: "Across the life sciences" note that Pubmed is a more limited database than e.g. Web of Science, even for biology... so some disciplines are not covered, the main focus is on biomedical sciences.

**We are aware that PubMed (and even Web of Science, which is not open-access) does not cover the entire life sciences literature, but we think the statement is accurate as written (since 'across' does not imply 'all'). PubMed does span essentially all major fields of the life sciences, even if its coverage of fields and journals is not 100%. See Figures 1 and 2 in Holman et al. 2018, and the accompanying web app, which lists data for >6000 journals indexed by PubMed (these include essentially all of chemistry, biology, psychology, biophysics, etc in addition to things related to medicine and health).**

- line 181: "Typical gender ratio" Please give a range (it seems low in the example)

**To support this statement, we cited Holman et al. 2018, which draws on data from the author lists of >10 million research papers, and shows that many fields have at least this great a shortage of women authors. See Figures 1 and 2 in that paper for the percentages (there are >100, for many diverse fields, so it would not be helpful to give a one-size-fits-all estimate without 'getting into the weeds' about why we picked that particular estimate, which is not relevant for the point being made on line 181).**

- line 181: "there was no inflation of $\alpha$ in highly multidisciplinary journals" One of the aims of your study was to control for Wahlund effect. Maybe you could show the value of alpha' if you group all journals, to better motivate the need to control for the effect of "subdivision" by journal/disciple?

**As suggested, we now calculate $\alpha$ for every suitable PubMed paper in our dataset (3.7 paper, 16m authors), ignoring the journal and publication date, to see if it inflates $\alpha$ as we expected. Reassuringly, the $\alpha$ for PubMed as a whole is 0.126, which is almost double the median $\alpha$ calculated on 1-year datasets for individuals journals (i.e. our main result, shown in Figure 1), supporting our theoretical argument that ignoring the Wahlund effect does add a non-trivial amount of inflation to $\alpha$.**

  - Line 246 "our results provide non-experimental support for the hypothesis that mixed-gender teams produce better research than single-gender teams" is better said this way than in the abstract

**We feel that both phrasings are carefully chosen, and accurately reflect what we found in the study. For comparison, the Abstract equivalent of this sentence reads, "Interestingly, journals with a high impact factor for their discipline tended to have comparatively low homophily, as predicted if mixed-gender teams produce better research." Note that: 1) the word 'Interestingly' reflects our feeling that this result is an unexpected curio that requires follow up study (although there are already some experimental studies examining the effects of diversity on the success of team-based projects, which we cite), 2) the word 'tended' stresses that this is not a hard-and-fast finding, but rather a noisy correlation, and 3) the word 'predicted' stresses to the reader that this result is consistent with the theory but does not prove it (that is, we do not commit the fallacy of 'affirming the consequent').**

- line 257: "Given that many collaborative research projects unfortunately involve a gendered division of labour [61], working with a same-gendered colleague may provide exposure to new parts of the research process, and (especially for the minority gender) a welcome change of pace." Not sure I understand (/agree with) the logic behind this last sentence. Do you mean that if you collaborate with a same-gendered colleague, you may end up doing things that are different than if you collaborated with a different-gendered colleague? This sounds a bit odd to me... A different type of interaction, different level of implication, sure. But if I am a lab tech, working with a male or female PI won't really make me change the type of work that I do -- or will it?
And I am not sure I understand what is alluded to by "welcome change of pace"...

**You mention the right answer here -- our intended meaning was "a different type of interaction". For women working in a field where women are the minority (e.g. mathematics or surgery), the majority of their professional interactions will be with men, and so working with another woman would be a (perhaps welcome) novelty, given that men and women differ consistently in various traits (and that some people sometimes respond differently to male-female collaborator pairs relative to same-sex pairs, e.g. by making gender-based assumptions about who did what towards the project).**

#### Supplementary
- Kudos for providing the code! Rmarkdown is wonderful (show/hide comments is also a great feature!).

**Thanks, we agree it's an excellent tool for presenting the details of a coding-intensive analysis and maximising its transparency, reproducibility, and potential to be adapted and re-used by other researchers.**

- Could there be a pdf output as well? and a table of contents and lists of tables/figures? To avoid having to use ctrl-F to find the supplementary figures...

**If the MS is accepted, PLOS will ask us to upload all the supplementary figures and tables one by one, giving a clickable route to each one of the supplementary figures in the paper itself, and so it's not necessary to make a table of contents for our HTML supplement. As for making it into a PDF, this doesn't seem useful (since it will contain the same information, but will lose all the nice features of HTML), but we're happy to do it if asked to by the copy editors. This will simply involve opening the HTML document using a web browser, then saving it as a PDF; we can do this if asked, or leave it up to any readers who need a PDF copy (note that HTML can be opened using any web browser, and is likely to be just as future-proof as PDF).**

- Fig S2 seems to be missing.

**Thanks, we fixed a typo and the figure now shows up.**

### 2018-10-24 Gender_homophily_analysis.Rmd
Luke's changes after PLOS ONE review

@@
 ```{r}
-gender.grouping.test.many.journals <- function(count.data, authors, output.file, nBoots, nSimulations, number.of.cores, over.write = FALSE){
+gender.grouping.test.many.journals <- function(count.data, 
+                                               authors, output.file, 
+                                               nBoots, nSimulations, 
+                                               number.of.cores, over.write = FALSE){
 
@@
 ggsave(fig_s8, file = "figures/S8 Fig.pdf", height = 5, width = 5, device = cairo_pdf)
+fig_s8
 ```
@@
 ggsave(fig_s2, file = "figures/S2 Fig.pdf", width = 5, height = 5, device = cairo_pdf)
+fig_s2
 ```
@@
 Furthermore, $\alpha'$ is significantly higher for 3-, 4-, and 5-plus-author papers than for 2-author papers.
-The R^2 of the model is 0.44, and the random effects appear to explain non-trivial amounts of variation.
+The $R^2$ of the model is 0.44, and the random effects appear to explain non-trivial amounts of variation.
 ```{r}
@@
 
+### Estimating inflation in $\alpha'$ due to country effects
+
+Using the country data, we can estimate whether the Wahlund effect created by gender differences between countries produced 'spurious' positive $\alpha'$ values in our dataset.
+
+For journal $i$, we take the average of the differences $\alpha'_i - \alpha'_{i,j}$, where $\alpha'_i$ is the coefficient of homophily calculated for all papers from journal $i$, and $\alpha'_{i,j}$ is the coefficient of homophily calculated for papers from journal $i$, by authors from country $j$.
+A positive value implies that pooling authors from different countries gives a higher estimate of homophily, consistent with a Wahlund effect.
+
+Plotting a histogram of these average differences for the `r length(unique(country.data$journal))` journals shows that the mean and median difference is essentially zero, illustrating that $\alpha'$ is not strongly inflated by Wahlund effects caused by differences in author gender ratio between countries.
+
+```{r}
+mean.inflation.due.to.country <- country.data %>%
+  left_join(master.dataset %>% 
+              filter(Age == "New", Subset == "All authors") %>% 
+              select(journal, alpha.adjusted) %>% 
+              rename(journal.alpha = alpha.adjusted), 
+            by = "journal") %>%
+  group_by(journal) %>%
+  summarise(mean.inflation.due.to.country = mean(journal.alpha - alpha.adjusted))
+
+median <- median(mean.inflation.due.to.country$mean.inflation.due.to.country)
+
+txt <- data.frame(x = median + 0.06, y = 66, label = paste("Median =", round(median,3)))
+
+mean.inflation.due.to.country <- mean.inflation.due.to.country %>%
+  ggplot(aes(mean.inflation.due.to.country)) + 
+  geom_histogram(binwidth = 0.02, colour = "black", fill = "skyblue") + 
+  geom_vline(xintercept = median, linetype = 2) + 
+  geom_text(data=txt, aes(x=x, y=y, label=label)) + 
+  theme_minimal() + 
+  xlab("Estimated inflation in \u03B1' due to\ndifferences in gender ratio between countries") + 
+  ylab("Frequency")
+
+ggsave(mean.inflation.due.to.country, file = "figures/S6 Fig.pdf", width = 5, height = 5, device = cairo_pdf)
+mean.inflation.due.to.country
+```
+<br></br>
+**Fig S6**: Histogram showing the estimated degree to which $\alpha'$ is inflated by inter-country differences in author gender ratio, across the 285 journals for which we had adequate data after restricting the analysis by country.
+The average inflation in $\alpha'$ is negligible, suggesting that Wahlund effects resulting from inter-country differences have a neglible effect on our estimates of gender homophily.
+
+### Calculate $\alpha$ for the whole of PubMed
+In response to a reviewer query, we here calculate the value of $\alpha$ across all the PubMed journals in our dataset, across the whole $c.$ 15 year time period.
+After removing papers with at least one author of unknown gender as well as single-author papers, we were left with 3,709,136 papers in total.
+$\alpha$ was 0.126, which is higher than the median for individual journals, consistent with the arguments presented in Figure 1 (i.e.
+that combining gender data across fields and time periods inflates $\alpha$ via the Wahlund effect).
+
+```{r}
+author_counts_for_whole_of_Pubmed <- 
+  pubmed_sqlite %>% # Access the big SQLite database
+  select(gender.95) %>%  # Get the column of genders (to 95% certainty, based on genderize.io database)
+  filter(gender.95 != "U", # Remove papers with author lists like this (cannot search strings on the database side)
+         gender.95 != "UU",
+         gender.95 != "UUU",
+         gender.95 != "UUUU",
+         gender.95 != "UUUUU",
+         gender.95 != "F", # no single-author papers either
+         gender.95 != "M") %>%
+  collect(n = Inf) %>%  # Bring all the remaining data from the SQL database into R
+  filter(!str_detect(gender.95, "U")) %>% # Remove all papers with any unknown-gender authors
+  filter(nchar(gender.95) > 1) %>% # Remove single-author papers
+  mutate(nFemales = str_count(gender.95, "F"), # Count the men and women on each remaining paper
+         nMales   = str_count(gender.95, "M")) %>%
+  select(-gender.95)
+
+# Number of papers in this dataset: 3,709,136 
+# Number of authors on these papers: 16,100,616
+
+print(paste("The value of alpha for the whole of Pubmed is:",
+round(find.alpha(nMales = author_counts_for_whole_of_Pubmed$nMales, 
+           nFemales = author_counts_for_whole_of_Pubmed$nFemales), 3)))
+```
+
 #### Sample sizes for the journal-country dataset
@@
 
-### Estimating inflation in $\alpha'$ due to country effects
-
-Using the country data, we can estimate whether the Wahlund effect created by gender differences between countries produced 'spurious' positive $\alpha'$ values in our dataset.
-
-For journal $i$, we take the average of the differences $\alpha'_i - \alpha'_{i,j}$, where $\alpha'_i$ is the coefficient of homophily calculated for all papers from journal $i$, and $\alpha'_{i,j}$ is the coefficient of homophily calculated for papers from journal $i$, by authors from country $j$.
-A positive value implies that pooling authors from different countries gives a higher estimate of homophily, consistent with a Wahlund effect.
-
-Plotting a histogram of these average differences for the `r length(unique(country.data$journal))` journals shows that the mean and median difference is essentially zero, illustrating that $\alpha'$ is not strongly inflated by Wahlund effects caused by differences in author gender ratio between countries.
-
-```{r}
-mean.inflation.due.to.country <- country.data %>%
-  left_join(master.dataset %>% 
-              filter(Age == "New", Subset == "All authors") %>% 
-              select(journal, alpha.adjusted) %>% 
-              rename(journal.alpha = alpha.adjusted), 
-            by = "journal") %>%
-  group_by(journal) %>%
-  summarise(mean.inflation.due.to.country = mean(journal.alpha - alpha.adjusted))
-
-median <- median(mean.inflation.due.to.country$mean.inflation.due.to.country)
-
-txt <- data.frame(x = median + 0.06, y = 66, label = paste("Median =", round(median,3)))
-
-mean.inflation.due.to.country <- mean.inflation.due.to.country %>%
-  ggplot(aes(mean.inflation.due.to.country)) + 
-  geom_histogram(binwidth = 0.02, colour = "black", fill = "skyblue") + 
-  geom_vline(xintercept = median, linetype = 2) + 
-  geom_text(data=txt, aes(x=x, y=y, label=label)) + 
-  theme_minimal() + 
-  xlab("Estimated inflation in \u03B1' due to\ndifferences in gender ratio between countries") + 
-  ylab("Frequency")
-
-ggsave(mean.inflation.due.to.country, file = "figures/S6 Fig.pdf", width = 5, height = 5, device = cairo_pdf)
-mean.inflation.due.to.country
-```
-<br></br>
-**Fig S6**: Histogram showing the estimated degree to which $\alpha'$ is inflated by inter-country differences in author gender ratio, across the 285 journals for which we had adequate data after restricting the analysis by country.
-The average inflation in $\alpha'$ is negligible, suggesting that Wahlund effects resulting from inter-country differences have a neglible effect on our estimates of gender homophily.
-
 ## Theoretical expectations for $\alpha$ when there is a gender gap between career stages
@@
 This section makes Figure 6, by deriving the expected $\alpha$ under various assumptions, assuming that there is (potentially) a gender gap between career stages but no direct choice of collaborators by gender.
+The model has four parameters:
+
+  - The proportion of women among people in career stage 1 (e.g. students), termed `GR.stage1` in the code below (this parameter is plotted on the x-axis of Figure 6)
+  - The proportion of women among people in career stage 2 (e.g. professors), termed `GR.stage2` in the code below (this parameter is plotted on the y-axis of Figure 6)
+  - The proportion of co-authorships that are between two people at career stages 1 and 2, termed `prop_between_career_stages` in the code below (this parameter is plotted in the columns of Figure 6)
+  - The proportion of within-career stage co-authorships that are between two people at career stage 1, termed `prop_within_stage_between_two_stage1s` in the code below (this parameter is plotted in the rows of Figure 6)
+
+Note that we do not need to define a parameter for the proportion of within-career stage co-authorships that are between two people at career stage 2; this is already known, since it is defined as `1 - prop_within_stage_between_two_stage1s`.
+
+Given these 4 parameters, we can easily calculate the proportion of the three possible types of collaborations with respect to career stage: stage1 with stage1, stage1 with stage2, and stage2 with stage2.
+See step 1 in the code.
+
+Next, we need to calculate the proportion of collaborator pairs that are male-male, female-female, and male-female.
+Since we have defined the frequencies of each collaboration type with respect to career stage, as well as the gender ratios at each career stage, we can calculate the frequency of each gender combination.
+See step 2 in the code.
+
+Finally, we calculate $\alpha = p - q$, where $p$ and $q$ are the proportions of men and women, respectively, whose coauthor is male (step 3 in the code).
 
 ```{r warning=FALSE}
-simulate <- function(params){
-
-  # Attach the vectors of paramters from the dataframe input
+run_career_stage_model <- function(params){
+
+  # Attach the vectors of parameters from the dataframe input
+  # GR = gender ratio, so GR.stage1 means gender ratio among career stage 1 ("early-career", e.g. PhD students)
+  # and GR.stage1 means gender ratio among career stage 2 ("established", e.g. professors)
   GR.stage1 <- params$GR.stage1
   GR.stage2 <- params$GR.stage2
-  prop.student.prof <- params$prop.student.prof
-  prop.student.student <- params$prop.student.student
-
-  # This is the proportion of each type of collaboration:
-  # early-early, early-late, late-late respectively
-  p1.1 <- (1 - prop.student.prof) * prop.student.student
-  p1.2 <- prop.student.prof
-  p2.2 <- (1 - prop.student.prof) * (1-prop.student.student)
-
-  # given the gender ratio at career stages a and b, find the 
+
+  # Gere are the proportions of collaborations of each of the three types:
+  # between career stages, between two stage 1 researchers, and between two stage 2 researchers
+  prop_between_career_stages <- params$prop_between_career_stages
+  prop_within_career_stages <- 1 - prop_between_career_stages
+  prop_within_stage_between_two_stage1s <- params$prop_within_stage_between_two_stage1s
+  prop_within_stage_between_two_stage2s <- 1 - prop_within_stage_between_two_stage1s
+
+  # Step 1. Calculate the proportions of each type of collaboration:
+  # early-early, early-established, established-established respectively
+  p1.1 <- prop_within_career_stages * prop_within_stage_between_two_stage1s
+  p1.2 <- prop_between_career_stages
+  p2.2 <- prop_within_career_stages * prop_within_stage_between_two_stage2s
+
+  # Step 2
+  # Define helper functions to calculate the proportion of MM, FF, and MF collaborations.
+  # Given the gender ratio at career stages a and b, termed GR_a, GR_b, find the 
   # proportion of collaborations that are man-man, woman-woman, or mixed gender
-  man.man <- function(GR_a, GR_b) (1-GR_a) * (1-GR_b)
+  man.man <- function(GR_a, GR_b) (1 - GR_a) * (1 - GR_b)
   woman.woman <- function(GR_a, GR_b) GR_a * GR_b
-  mixed <- function(GR_a, GR_b) (1-GR_a) * GR_b + (1-GR_b) * GR_a
-
-  # Find the frequencies of the 3 kinds of collaboration
+  mixed <- function(GR_a, GR_b) (1 - GR_a) * GR_b + (1 - GR_b) * GR_a
+
+  # Find the frequencies of the 3 kinds of co-author pairs w.r.t. gender, i.e. MM, FF, and MF
   man.man <- p1.1 * man.man(GR.stage1, GR.stage1) +
@@
 
-  # Find p and q, which are used to calculate alpha:
+  # Step 3: Find p and q, which are used to calculate alpha, as p - q:
   # p is proportion of men coauthoring with men
-  p <- 1 - mixed / (2 * man.man + mixed)
+  p <- 1 - (mixed / (2 * man.man + mixed))
   # q is proportion of women coauthoring with men
@@
 
+  # Return the calculated value of alpha, as well as the four parameters used to calculate it
   data.frame(alpha = p - q, 
@@
              GR.stage2,
-             prop.student.prof,
-             prop.student.student)
-}
-
-results <- expand.grid(
+             prop_between_career_stages,
+             prop_within_stage_between_two_stage1s)
+}
+
+# Now run the model on a large parameter space
+results <- expand.grid( # Define the parameter space for the model
   GR.stage1 = seq(0.01, 0.7, length = 201),
   GR.stage2 = seq(0.01, 0.7, length = 201),
-  prop.student.prof = c(0.01, 0.25, 0.5, 0.75, .99), 
-  prop.student.student = c(0.01, 0.25, 0.5, 0.75, .99)) %>% 
-  filter(GR.stage1 >= GR.stage2) %>%
-  simulate() %>% 
-  mutate(GR.stage1 = round(GR.stage1 * 100),
+  prop_between_career_stages = c(0.01, 0.25, 0.5, 0.75, .99), 
+  prop_within_stage_between_two_stage1s = c(0.01, 0.25, 0.5, 0.75, .99)) %>% 
+  filter(GR.stage1 >= GR.stage2) %>% # We assume the % women is equal or lower in career 1 than in career stage 2
+  run_career_stage_model() %>%  # run the model
+  mutate(GR.stage1 = round(GR.stage1 * 100), # Convert proportions to percentages
          GR.stage2 = round(GR.stage2 * 100),
-         prop.student.prof = prop.student.prof * 100,
-         prop.student.student = prop.student.student * 100) %>%
-  filter(!duplicated(paste(GR.stage1, GR.stage2, prop.student.prof, prop.student.student)))
-
-simulation.figure <- results %>%
+         prop_between_career_stages = prop_between_career_stages * 100,
+         prop_within_stage_between_two_stage1s = prop_within_stage_between_two_stage1s * 100) %>%
+  filter(!duplicated(paste(GR.stage1, GR.stage2, prop_between_career_stages, prop_within_stage_between_two_stage1s)))
+
+# Plot the results to make figure 6
+# NB that I edited the top- and right-side labels in Inkscape afterwards, as it was tricky to do it entirely in R.
+model_figure <- results %>%
   ggplot(aes(GR.stage1, GR.stage2, z = round(alpha,2), fill = alpha)) + 
-  facet_grid(prop.student.prof ~ prop.student.student) + 
+  facet_grid(prop_between_career_stages ~ prop_within_stage_between_two_stage1s) + 
   geom_tile() + 
@@
 
-simulation.figure
-ggsave(simulation.figure, file = "figures/Fig6.pdf", 
+model_figure
+ggsave(model_figure, file = "figures/Fig6.pdf", 
        height = 5.6, width = 5, device = cairo_pdf)

### 2018-10-24 manuscript/Manuscript_Claire.Rmd
Tweak abstract
Luke's changes after PLOS ONE review

@@
 csl: plos.csl
-abstract: Evidence suggests that women in academia are hindered by conscious and unconscious
-  biases, and often feel excluded from formal and informal opportunities for research
-  collaboration. In addition to ensuring fairness and helping to redress gender imbalance
-  in the academic workforce, increasing women's access to collaboration could help
-  scientific progress by drawing on more of the available human capital. Here, we
-  test whether researchers preferentially collaborate with same-gendered colleagues,
-  using more stringent methods and a larger dataset than in past work. Our results
-  reaffirm that researchers preferentially co-publish with colleagues of the same
-  gender, and show that this 'gender homophily' is slightly stronger today than it
-  was 10 years ago. Contrary to our expectations, we found no evidence that homophily
-  is driven mostly by senior academics, and no evidence that homophily is stronger
-  in fields where women are in the minority. Interestingly, journals with a high impact
-  factor for their discipline tended to have comparatively low homophily, as predicted
-  if mixed-gender teams produce better research. We discuss some potential causes of gender
-  homophily in science.
+abstract: Evidence suggests that women in academia are hindered by conscious and unconscious biases, and often feel excluded from formal and informal opportunities for research collaboration.
+In addition to ensuring fairness and helping to redress gender imbalance in the academic workforce, increasing women's access to collaboration could help scientific progress by drawing on more of the available human capital.
+Here, we test whether researchers preferentially collaborate with same-gendered colleagues, using more stringent methods and a larger dataset than in past work.
+Our results reaffirm that researchers preferentially co-publish with colleagues of the same gender, and show that this 'gender homophily' is slightly stronger today than it was 10 years ago.
+Contrary to our expectations, we found no evidence that homophily is driven mostly by senior academics, and no evidence that homophily is stronger in fields where women are in the minority.
+Interestingly, journals with a high impact factor for their discipline tended to have comparatively low homophily, as predicted if mixed-gender teams produce better research.
+We discuss some potential causes of gender homophily in science.
   \vspace{5mm} \par\noindent
@@
 
-However, we believe that prior studies of gender homophily were hindered by a largely unacknowledged statistical issue that we name the Wahlund effect (\autoref{wahlund_plot}), by analogy with the conceptually similar Whalund effect in population genetics [@Wahlund_1928].
+However, we believe that earlier studies of gender homophily were hindered by a largely unacknowledged statistical issue that we will refer to as the Wahlund effect (\autoref{wahlund_plot}), by analogy with the conceptually similar Whalund effect in population genetics [@Wahlund_1928].
 The Wahlund effect makes it deceptively difficult to infer gender-based preferences simply by counting the number of same- and mixed-gender coauthorships.
@@
 This can give the impression that authors preferentially publish with same-gendered colleagues even if no gender preferences exist, or if the true preference is for opposite-gendered colleagues ('gender heterophily').
-For example, a sample of literature containing bioinformatics and cell biology papers will probably contain an excess of mostly-male and mostly-female author lists, simply because researchers usually collaborate within their own discipline, and because the author gender ratio is more male-biased in bioinformatics than in cell biology [@Holman_2018].
+For example, a sample of literature containing a mixture of bioinformatics and cell biology papers will probably contain an excess of mostly-male and mostly-female author lists, simply because researchers usually collaborate within their own discipline, and because the author gender ratio is more male-biased in bioinformatics than in cell biology [@Holman_2018].
 
-![The Wahlund effect can make it appear as if authors prefer to publish with same-gendered colleagues, even if no such preference exists.
+![The Wahlund effect can make it appear as if authors publish with same-gendered colleagues disproportionately often, even if collaboration is completely random with respect to gender.
 Here, coloured circles represent male and female authors, and coauthors are linked with lines.
-Across the whole set of ten papers, there is an apparent excess of same-gender collaborations: there are six same-gender papers and only four mixed-gender papers, which is fewer than the $10\times2\times0.5\times0.5 = 5$ mixed-gender papers expected under the null hypothesis that authors assort randomly with respect to gender.
+Across the whole set of ten papers, there is an apparent excess of same-gender collaborations: there are six same-gender papers and only four mixed-gender papers, which is fewer than the $10\times2\times0.5\times0.5 = 5$ mixed-gender papers expected under the null hypothesis that authors assort randomly.
 However, within each subset, there is no evidence that authors prefer to publish with same-gendered individuals (if anything, this small dataset suggests gender heterophily).
-The Wahlund effect will tend to inflate the frequency of same-sex coauthorships whenever the data is composed of two or more disconnected subsets of literature with different author gender ratios; these subsets could be research disciplines, older versus newer papers, or papers from authors in different countries.
-The example countries and disciplines were selected based on [@Holman_2018].
+The Wahlund effect will tend to inflate the frequency of same-gender coauthorships whenever the data is composed of two or more disconnected subsets of literature with different author gender ratios; these subsets could be research disciplines, older versus newer papers, or papers from authors in different countries.
+The example countries and disciplines were selected based on data in [@Holman_2018].
 \label{wahlund_plot}](../figures/Fig1.pdf)
@@
 In the present study, we test whether life sciences researchers tend to co-publish with same-gendered colleagues, while controlling for the Wahlund effect as strictly as possible.
-We use a recently-published dataset describing the gender of 35.5m authors from 9.15m articles indexed on PubMed [@Holman_2018].
+We use a recently-published dataset describing the gender of 35.5 million authors from 9.15 million articles indexed on PubMed [@Holman_2018].
 Holman et al.
@@
 research specialties), time periods, and countries.
-We quantified gender assortment using a metric called $\alpha'$ [@bergstrom_2016], which is positive when same-gender authors publish together more often than expected (gender homohily), negative when opposite-gender authors publish together more often than expected (heterophily), and equal to zero when authors assort randomly with respect to gender (see Methods).
+We quantified gender assortment using a metric called $\alpha'$ [@bergstrom_2016], which is positive when same-gender authors publish together more often than expected (gender homophily), negative when opposite-gender authors publish together more often than expected (heterophily), and equal to zero when authors assort randomly with respect to gender (see Methods).
 
@@
 \autoref{alpha_histograms} shows the distribution of $\alpha'$ estimates in 2015-2016 across all journals for which we recovered sufficient data, when $\alpha'$ was calculated for all authors, first authors only, or last authors only.
-Most journals had positive values of $\alpha'$ (77-92%, depending on time period and author type; S1 Data), and for many of these the FDR-corrected p-values suggested that $\alpha'$ was significantly greater than zero (1469/2077 journals were significant in 2015-16, and 404/1192 in 2005-6; S1 Data).
-Only 2/2077 journals had statistically significantly heterophily (i.e.
+Most journals had positive values of $\alpha'$ (77-92%, depending on time period and author type; S1 Data), and for many of these the false discovery rate (FDR)-corrected p-values suggested that $\alpha'$ was significantly greater than zero (1469/2077 journals were significant in 2015-16, and 404/1192 in 2005-6; S1 Data).
+Only 2/2077 journals had statistically significant heterophily (i.e.
 $\alpha' < 0$) in 2015-16, and 1/1192 in 2005-6 (S2 Table).
 The remaining 606 or 787 journals (in 2015 and 2005 respectively) had a value of $\alpha'$ not significantly different from zero, consistent with the null hypothesis of random assortment with respect to gender.
-We also confirmed that the majority of papers had multiple authors, in most journals (S2 Data) and disciplines (S3 Data, S1 Fig).
+We also confirmed that in most journals (S2 Data) and most research disciplines (S3 Data, S1 Fig), the majority of papers had multiple authors.
+To verify the argument put forward in Figure 1, we also calculated $\alpha$ across all two-author papers in our entire 15-year PubMed dataset (excluding single-author papers, and papers where at least one author's gender was unknown; n = >3 million papers, >16 million authors).
+Across the whole dataset, $\alpha$ was 0.126, which is almost double the median value for individual journals (Figure 2; $\alpha = 0.070$ for 'All authors'), suggesting that lumping together papers from different fields and different time periods can indeed inflate $\alpha$.
 
 ![Of the `r n.journals` journals for which we had adequate data in 2015-2016, `r n.sig.homo` showed statistically significant evidence of gender homophily (denoted by $\alpha' > 0$), and `r n.sig.hetero` showed statistically significant evidence of heterophily ($\alpha' < 0$), after false discovery rate correction.
-The white area shows the number of journals for which homophily was significantly stronger than expected under the null hypothesis (corrected p < 0.05), while the blue area shows all the remainder.
+In the stacked desnity plot, the white area shows the number of journals for which homophily was significantly stronger than expected under the null hypothesis (corrected p < 0.05), while the blue area shows all the remainder.
 Patterns were similar whether $\alpha'$ was calculated for all authors, for first authors only, or for last authors only.
-\label{alpha_histograms}](../figures/Fig2.pdf)
+Points in the right panel show $\alpha'$ for individual journals.\label{alpha_histograms}](../figures/Fig2.pdf)
 
@@
 \autoref{alpha_histograms} illustrates the variance in journal homophily values ($\alpha'$) across scientific disciplines.
-All disciplines had positive mean $\alpha'$, although homophily appeared somewhat stronger in some disciplines than others (e.g.
+All disciplines had positive mean $\alpha'$ (averaged over journals), although homophily appeared somewhat stronger in some disciplines than others (e.g.
 mean $\alpha'$ was $0.12{\pm}0.02$ for Urology journals and $0.03{\pm}0.01$ for Veterinary Medicine journals; \autoref{alpha_histograms}, S4 Data).
 However, there was no formal evidence for consistent differences in $\alpha'$ between disciplines: the random factor 'Discipline' explained around 1% of the variance in $\alpha'$ in the two linear mixed models described in the previous section (see \autoref{alpha_histograms} and mixed models in Online Supplementary Material).
-Thus, the processes responsible for producing positive $\alpha'$ values appear to be similarly strong in all the disciplines we examined.
+Thus, whatever processes are responsible for producing positive $\alpha'$ values appear to be similarly strong in all the disciplines we examined.
 
 There was no indication that journals publishing on a wide range of topics have higher $\alpha'$ values than more specialised journals, due to the Wahlund effect.
-For example, the journal category 'Multidisciplinary' -- which includes journals like _PLoS ONE_, _Nature_, _Science_, and _PNAS_ -- did not have notably elevated $\alpha'$ (\autoref{alpha_histograms}).
-This result suggests that our estimates of homophily, and estimates from some earlier studies, are not notably inflated by the presence of disparate research topics (with variable author gender ratios) being published within individual journals.
+For example, the journal category 'Multidisciplinary' -- which includes journals like _PLOS ONE_, _Nature_, _Science_, and _PNAS_ -- did not have markedly elevated $\alpha'$ (\autoref{alpha_histograms}).
+This result suggests that our estimates of homophily, and estimates from some of the earlier studies listed in the Introduction, are probably not inflated by the presence of disparate research topics (with variable author gender ratios) being published within individual journals.
 
@@
 The variance in $\alpha'$ across journals was also a little higher for 2-authors papers compared to the remainder (\autoref{author_number}), though part of this variance is due to the reduced sample size (in terms of number of authors) for the 2-author papers.
+One possible explanation for this finding is that 2-authors papers might be more likely to have an author list that is evenly split between career stages (e.g.
+a postgraduate student and their supervisor), increasing the chance that the authors are mixed gender (see Figure 6).
+The result also suggests that the processes responsible for gender homophily are similar in small (e.g.
+3 author) and larger (5+ authors) collaborations (and across disciplines where small versus large collaborations are the norm).
 
@@
 We found a positive, non-linear relationship between the overall gender ratio of all authors publishing in a particular journal [as estimated in @Holman_2018], and the estimated value of $\alpha'$ for all authors and for first authors (\autoref{alpha_gender_ratio}).
-Journals with a balanced or female-biased author gender ratio tended to have higher $\alpha'$ than journals with a male-biased author gender ratio (GAM smooth terms p < 0.001; Online Supplementary Material).
+Journals with a balanced or female-biased author gender ratio tended to have higher $\alpha'$ (i.e.
+stronger homophily) than journals with a male-biased author gender ratio (GAM smooth terms p < 0.001; Online Supplementary Material).
 The relationship was not statistically significant when $\alpha'$ was calculated for last authors (GAM, p = 0.142), though the trend appeared similar (\autoref{alpha_gender_ratio}).
@@
 This relationship held whether $\alpha'$ was calculated for all authors, first authors only, or last authors only.
-A negative value on the x-axis denotes an excess of men authors, a positive value denotes an excess of women authors, and zero denotes gender parity.
+A negative value on the x-axis denotes an excess of men authors, a positive value denotes an excess of women authors, and zero denotes gender parity (i.e.
+equal numbers of men and women).
 The lines were fitted using generalised additive models with the smoothing parameter $k$ set to 3.
@@
 ## Theoretical expectations for $\alpha$ when the gender ratio differs between career stages
+Given that we cannot identify individual researchers or their career stages, we used a simple model to derive the theoretical expectations for $\alpha$ when the gender ratio differs between career stages.
 As shown in \autoref{simulation}, we predict that $\alpha$ is expected to be non-zero, even if collaborators are randomly selected with respect to gender, provided that there is a gender gap between career stages.
-The extent to which $\alpha$ deviates from zero depends on the relative frequencies of collaboration within and between career stages.
-When >50% of collaborations were between early and established researchers, we expect gender heterophily ($\alpha < 0$).
-Conversely, when >50% of collaborations occured within career stages, we expect gender homophily ($\alpha > 0$).
+The extent to which $\alpha$ deviates from zero depends on the relative frequencies of collaboration within and between career stages (rows and columns in Figure 6), and the size of the gender gap between stages (x- and y-axes in Figure 6).
+When >50% of coauthor pairs comprise one early-career and one established researcher, we expect gender heterophily ($\alpha < 0$) whenever the gender ratio differs between career stages.
+Conversely, when >50% of collaborations are between people at the same career stage, we expect gender homophily ($\alpha > 0$).
 In a few parameter spaces (shown in red; \autoref{simulation}), $\alpha$ was quite high, and overlapped with the values that we estimated (\autoref{alpha_histograms}).
@@
 <10% of the total).
-Both of these conditions are untrue for most fields: the gender gap across careers stages is generally less pronounced [e.g.
+Both of these conditions seem unlikely to be true for most fields: the gender gap across careers stages is generally less pronounced [e.g.
 @Shaw_2012; @Holman_2018], and it is very common for early-career researchers to co-publish with an established mentor [@macaluso_2016].
 However, one can get $\alpha > 0$ for realistic combinations of parameters, e.g.
-a moderate shortage of women in senior positions coupled with a moderate excess of within-career stage collaboration, suggesting this effect might contribute to some of the observed homophily (in this and previous studies).
+a moderate shortage of women in senior positions coupled with a moderate excess of within-career stage collaboration, suggesting this effect might contribute to some of the homophily observed by this and previous studies.
 
 Lastly, we note that if there is a gender gap between career stages and coauthorships between early-career and established researchers comprise >50% of the total, then the baseline expectation for $\alpha$ is actually less than zero (blue areas in \autoref{simulation}).
-Therefore, our results might under-estimate the extent to which researchers preferentially select same-gendered collaborators in some cases.
+Therefore, it is possible that researchers preferentially assort with same-gendered collaborators even more strongly than implied by other results, at least for certain journals or research disciplines.
 
 # Discussion
-We found evidence that researchers preferentially publish with same-gendered coauthors, even after implementing stringent controls for Wahlund effects (\autoref{wahlund_plot}).
+We found evidence that researchers work with same-gendered coauthors more often than expected under the null model, even after implementing stringent controls for Wahlund effects (\autoref{wahlund_plot}).
 Our study therefore reaffirms earlier studies' conclusions [e.g.
-@Ferber_1980; @McDowell_1992; @Bentley_2003; @Crow_2015; @Ghiasi_2015; @Araujo_2017a; @Fahmy_2017; @Jadidi_2017; @Teele_2017; @Zettler_2017] and establishes their generality across the life sciences.
+@Ferber_1980; @McDowell_1992; @Bentley_2003; @Crow_2015; @Ghiasi_2015; @Araujo_2017a; @Fahmy_2017; @Jadidi_2017; @Teele_2017; @Zettler_2017] using stricter methodology, and generalises their results across the life sciences.
 Relatively few journals had $\alpha'$ values below zero, and almost no journals showed statistically significant gender heterophily after controlling for multiple testing.
@@
 An important limitation of our study is that we cannot reliably determine the cause(s) of the observed excess of same-gender coauthorships.
-As well as the obvious interpretation -- conscious or unconcious selection of same-gender collaborators by men, women, or both -- our results could be partly explained by uncontrolled Wahlund effects.
-However, we suspect the contribution of these to be minor, for four reasons: we found positive $\alpha'$ after controlling for three obvious sources of Wahlund effect; there was no inflation of $\alpha'$ in highly multidisciplinary journals; restricting the data by country yielded similar estimates of $\alpha'$; and we showed that differences in gender ratio between career stages are unlikely to fully explain our results.
-On balance, we believe the data suggest that it is likely that some researchers do preferentially select same-gendered collaborators, although the frequency and strength of this preference is difficult to ascertain.
+As well as the obvious interpretation -- conscious or unconcious selection of same-gender collaborators by men, by women, or both -- our results could be partly explained by uncontrolled Wahlund effects.
+However, we suspect the contribution of these uncontrolled artefacts to be minor, for four reasons: we found positive $\alpha'$ after controlling for three obvious sources of Wahlund effect; there was no inflation of $\alpha'$ in highly multidisciplinary journals relative to specialised journals; restricting the data by country yielded similar estimates of $\alpha'$; and we used a simple model to show that differences in gender ratio between career stages are unlikely to fully explain our results.
+On balance, we believe the data suggest that it is likely that some researchers preferentially select same-gendered collaborators, although it is difficult to ascertain what proportion of people show such a preference, or how much the strength of the preference varies between individuals.
+We also note that even in a world in which collaboration was completely random with respect to gender, a high proportion of individual researchers would have entirely same-gendered collaborators by chance alone (especially in gender-biased disciplines); thus, individuals who only have same-gendered co-authors are not necessarily doing anything differently from people with gender-balanced co-authors.
 
 We hypothesised that disciplines with a strongly skewed gender ratio might show the strongest gender homophily, e.g.
-because being in the minority might increase motivation to seek out same-gendered colleagues.
+because being in the minority might increase people's motivation to seek out same-gendered colleagues.
 Contrary to this hypothesis, we found no evidence that gender homophily is restricted to particular disciplines: $\alpha'$ was similarly high across the board (\autoref{alpha_histograms}).
 Interestingly, gender homophily was weakest for journals with a male-biased author gender ratio, and strongest in journals with a female-biased author gender ratio.
-This may suggest that men are more likely to preferentially seek out male collaborators in fields where men are a minority, relative to the homophily displayed by women in fields where women are a minority.
-However, this latter result is only tentatively supported since our sample contains few journals in which most authors are women (\autoref{alpha_gender_ratio}).
+One possible reason is that men are more likely to preferentially seek out male collaborators in fields where men are a minority, relative to the homophily displayed by women in fields where women are a minority.
+However, this latter result only has tentative statistical support since our sample contains few journals in which most authors are women (\autoref{alpha_gender_ratio}).
 
@@
 
-Regarding our finding of weaker homophily among 2-author papers, we suspect that many 2-author teams comprise a student/postdoc and a senior staff member, making these teams especially likely to be mixed-gender, due to the elevated gender gap among senior researchers [@Shaw_2012; @Holman_2018].
+Regarding our finding of weaker homophily among 2-author papers, we suspect that many 2-author teams comprise a student/postdoc and a senior staff member, making these teams especially likely to be mixed-gender, due to the greater shortage of women among senior researchers [@Shaw_2012; @Holman_2018].
 Assuming this interpretation is correct, this result suggests that our reported $\alpha'$ values may underestimate the strength of peoples' preferences for same-gendered collaborators; essentially, women seeking a senior collaborator could be constrained to work mostly with men, meaning that people's ideal and realised gender preferences would be mismatched.
 On a related note, Ghiasi et al.
-[-@Ghiasi_2015] argue that women in engineering are "compliant [in reproducing] male-dominated scientific structures" because they do not collaborate often enough with other women (their Figure 7 suggests that coauthorships between women are 30% more frequent than expected under random assortment).
-By contrast, we feel that it is not helpful to recommend that women collaborate primarily with other women, e.g.
-because this constrains women's options and may be counter-productive (particularly in fields like engineering, where 90% of professors are men [@Shaw_2012]).
+[-@Ghiasi_2015] argue that women in engineering are "compliant [in reproducing] male-dominated scientific structures" because they do not collaborate often enough with other women (for reference, their Figure 7 implies that coauthorships involving two women are 30% more frequent than expected under random assortment).
+By contrast, we feel that it may be counter-productive to recommend that women collaborate primarily with other women, e.g.
+because this constrains women's options (particularly in fields like engineering, where 90% of professors are men [@Shaw_2012]).
 Instead, we suggest that researchers of both genders can help to close the gender gap in STEMM.
-In the context of collaboration, one way to do this is to undertake self-examination to ensure that one is not inadvertently overlooking or excluding female potential students and colleagues.
+In the context of collaboration, one way to do this is to undertake self-examination to ensure that one is not inadvertently overlooking or excluding women among potential students and colleagues.
 One should also take care to treat male and female collaborators equally, e.g.
-in terms of training and mentoring, allocation of work, and how one frames or promotes the collaboration (e.g.
-in conference presentations or on a website); evidence suggests that unconscious bias causes people to undervalue women's research achievements [@Moss_2012], and possibly to assign menial or under-valued tasks to women and more prestigious tasks to men [@macaluso_2016].
+in terms of training and mentoring, allocation of work, how one frames or promotes the collaboration (e.g.
+in conference presentations or on a website), and in terms of the credit afforded to collaborators' contributions.
+Experimental work suggests that unconscious bias causes people to undervalue women's research achievements [@Moss_2012], and a study of author contribution statements found observational evidence that menial or under-valued tasks are assigned to women while more prestigious tasks are assigned to men [@macaluso_2016].
 
@@
 Additionally, Sheltzer and Smith [-@sheltzer_2014] concluded that 'elite' male academics (defined as recipients of major honours) have a higher proportion of male students and postdocs than non-elite male academics.
-This finding could contribute to the homophily we observed, and is cause for concern since Sheltzer and Smith [-@sheltzer_2014]'s results might reflect discrimination against women during hiring [@Moss_2012], or avoidance by women of elite research groups (e.g.
+This finding could contribute to the homophily we observed, and is cause for concern since the results might reflect discrimination against women during hiring [@Moss_2012], or avoidance by women of elite research groups (e.g.
 due to gender differences in confidence, or a perception that some groups are sexist).
 We also found a little evidence that gender homophily is detrimental to research quality, in that high-impact journals tended to have weaker homophily.
-Assuming that papers published in high-impact journals are of higher average quality [which is contentious; @Garfield_2006], our results provide non-experimental support for the hypothesis that mixed-gender teams produce better research than single-gender teams [@Britton_2000; @Reagans_2001; @Hong_2004; @Whittington_2008; @Bear_2011; @Herrera_2012; @Campbell_2013].
+Assuming that papers published in high-impact journals are of higher average quality (which is contentious; [@Garfield_2006]), our results provide non-experimental support for the hypothesis that mixed-gender teams produce better research than single-gender teams [@Britton_2000; @Reagans_2001; @Hong_2004; @Whittington_2008; @Bear_2011; @Herrera_2012; @Campbell_2013].
 Another issue is that if many collaborations are between established researchers, there will be an excess of male-male collaborations in fields where women in senior positions are rare; some of the observed homophily might therefore reflect the elevated gender gap among senior researchers.
@@
 On the other hand, homophily might have more benign causes.
-Collaboration is often most enjoyable and productive when working with like-minded people, who might be same-gendered more often than not.
+Collaboration is often most enjoyable and productive when working with like-minded people, who might tend to be same-gendered more often than not.
 We also suppose that some people consciously choose to preferentially collaborate with women in order to help close the gender gap in the workforce; this would create homophily if women do this more than men.
-In support of this interpretation, women appear more likely than men to promote the work of female colleagues by inviting them to give talks [@nittrouer_2018; @debarre_2018].
+In support of this interpretation, there is some evidence that women are more likely than men to promote the work of female colleagues by inviting them to give talks [@nittrouer_2018; @debarre_2018].
 Given that many collaborative research projects unfortunately involve a gendered division of labour [@macaluso_2016], working with a same-gendered colleague may provide exposure to new parts of the research process, and (especially for the minority gender) a welcome change of pace.
@@
 [-@Holman_2018].
-Briefly, that dataset was created by downloading every article indexed on PubMed and attempting to infer each author's gender from their given name.
+Briefly, that dataset was created by downloading every article indexed on PubMed and attempting to infer each author's gender from their given name using computational methods.
 Each journal was assigned to one of 107 scientific disciplines, using PubMed's journal categorisations in the interests of objectivity.
@@
 We also discarded all papers for which we could not determine the gender of every author with ${\ge}95\%$ certainty, in order to simplify the statistical analysis.
-To mitigate Wahlund effects caused by variation in the gender ratio of researchers over time (see below), we also discarded all papers except those that were published either 0-1 or 10-11 years before the PubMed data were collected (i.e.
+To mitigate Wahlund effects caused by variation in the gender ratio of researchers over time (see below), we only kept papers with publicaiton dates falling in two one-year time periods, namely 0-1 or 10-11 years prior to the collection date of the PubMed data (i.e.
 20^th^ August 2016).
@@
 
-Our test assumes that the expected value of $\alpha$ is zero if authors randomly assort, but for small datasets this assumption is not always true (as pointed out by Carl Bergstrom in a blog post, http://www.eigenfactor.org/gender/assortativity/note_to_eisen.rtf).
-To borrow Prof.
-Bergstrom's example, consider a small research specialty comprising just two men and two women researchers, who have together produced six two-author papers: one in each of the six possible two-author combinations.
-For these six papers, $\alpha = -\frac{1}{3}$, even though same- and opposite-gendered coauthors were selected in equal proportion to their frequency in the pool of possible collaborators.
+Our test assumes that the expected value of $\alpha$ is zero if authors randomly assort, but for small datasets this assumption is not always true.
+Essentially, this issue arises because a person cannot be their own co-author.
+In a small dataset comprising $m$ men and $f$ women, a man can co-author with $m - 1$ men while women can co-author with $m$ men.
+Thus, the null expectation for $alpha$ is a negative number -- potentially a large one if $m$ and $f$ are very small.
 
@@
 Although some journals, e.g.
-_PLoS ONE_, publish research from diverse disciplines with very different author gender ratios [@Holman_2018], calculating $\alpha'$ for these highly multidisciplinary journals is still useful as a contrast.
+_PLOS ONE_, publish research from diverse disciplines with very different author gender ratios [@Holman_2018], calculating $\alpha'$ for these highly multidisciplinary journals is still useful as a contrast.
 The difference in $\alpha'$ between highly multidisciplinary and more specialised journals, e.g.
-_PLoS ONE_ versus _PLoS Computational Biology_, gives an estimate of the extent to which multidisciplinarity inflates $\alpha'$.
+_PLOS ONE_ versus _PLOS Computational Biology_, gives an estimate of the extent to which multidisciplinarity within journals inflates $\alpha'$.
 
@@
 For example, Holman et al.
-[-@Holman_2018] showed that PubMed-indexed authors based in Serbia are more than twice as likely to be women as are authors based in Japan.
+[-@Holman_2018] found that PubMed-indexed authors based in Serbia are more than twice as likely to be women as are authors based in Japan.
 Therefore, a dataset containing a mix of papers from teams of authors based in these two countries would contain an excess of same-sex coauthorships, even if collaboration were random with respect to gender within each country.
@@
 ## Calculating standardised journal impact factor
-We obtained the 3-year impact factor for each journal from Clarivate Analytics.
-To account for large differences in impact factor between disciplines, we took the the residuals from a model with $Log_{10}$ impact factor as the response and the research discipline of the journal as a random effect.
+We obtained the 3-year impact factor for each journal from Clarivate Analytics (formerly ISI).
+To account for large differences in impact factor between disciplines, we took the the residuals from a model with $log_{10}$ impact factor as the response and the research discipline of the journal as a random effect.
 Thus, journals with a positive standardised impact factor have a higher mean number of citations than the average for journals in their discipline.
@@
 ## Theoretical expectations for $\alpha$ when the gender ratio differs between career stages
-In many STEMM subjects, the gender ratio is more skewed among established researchers relative to early-career researchers [@Shaw_2012; @Holman_2018].
-We hypothesised that this skew could potentially create both Wahlund effects and 'reverse' Wahlund effects.
-For example, imagine that the majority of collaborations are between students and professors, and that the gender ratio differs between career stages: we will then see an excess of mixed-gender coauthorships (heterophily, $\alpha < 0$), even if gender has no direct, causal effect.
+In most STEMM subjects, the gender ratio is more skewed among established researchers relative to early-career researchers, due both to women leaving STEMM careers at greater rates (the 'leaky pipeline'), and to historical shortages of women studying STEMM ('demographic inertia') [@Shaw_2012; @Holman_2018].
+We hypothesised that this difference in gender ratio between career stages could potentially create both Wahlund effects and 'reverse' Wahlund effects.
+For example, imagine that the majority of collaborations in a particular field are between students and professors, and that the gender ratio differs between career stages: we would then see an excess of mixed-gender coauthorships (heterophily, $\alpha < 0$), even if gender has no direct, causal effect.
 Similarly, a hypothetical field in which students work only with students, and professors with professors, would have apparent gender homophily ($\alpha > 0$).
@@
 We can think of no tractable method of controlling for this issue using our dataset, which contains no information on career stage.
-Therefore, we instead decided to derive the theoretical expectations for $\alpha$ when there is a difference in gender ratio across career stages, in order to determine if and how this effect should alter our inferences.
-For simplicity, our calculations assume there are only two career stages, though we intuit that the general conclusions would also apply to a multi-tier career ladder.
-Under the null model that gender has no causal effect on collaboration, we calculated $\alpha$ for various combinations of the four free parameters, i.e.
-the gender ratios for early- and late-career researchers, and the relative frequency of collaborations between early-early, early-late, and late-late collaborations.
-We then used the theoretical expectations for $\alpha$ to qualify our main conclusions (see Results).
-The Online Supplementary Material gives annotated R code used to derive the theoretical expectations.
+Therefore, we instead decided to derive theoretical expectations for $\alpha$ when there is a difference in gender ratio across career stages, in order to determine if and how this effect should alter our inferences.
+For simplicity, our calculations assume there are only two career stages ('early-career' and 'established'), though we expect that the general conclusions would also apply to a multi-tier career ladder.
+Under the null model that gender has no causal effect on collaboration, we calculated $\alpha$ for various combinations of the four free parameters in our simple model.
+These parameters are: the gender ratio among early-career researchers (x-axis of Figure 6), the gender ratio among established researchers (y-axis of Figure 6), the frequency of within- versus between career stage collaborator pairs (rows in Figure 6), and lastly the frequency of within-stage collaborations that are between two early-career researchers as opposed to two late-career reserachers (columns in Figure 6).
+When these four parameters are specified, one can easily calculate the relative frequencies of collaborator pairs that involve two men, two women, or a man and a woman.
+In short, if we have specified the frequency of women at both career stages, as well as the frequency of the three possible types of collaboration with respect to career stage (early-early, early-established, and established-established), then we can calculate the frequency of collaborators pairs comprising two women, or a woman and a man, and thus find $\alpha$ (see the Online Supplementary Material for the annotated R code).
 
@@
 For journals for which we recovered a small number of papers (<100), the unadjusted metric $\alpha$ was downwardly biased.
-This fits our expectations: because authors cannot be their own co-authors, small datasets will tend to produce negative estimates of $\alpha$ even if authors assort randomly with respect to gender (see main text).
-This suggests that $\alpha'$ is a more useful measure of homophily and heterophily, especially for small samples.
+This fits our expectations: because researchers cannot be their own co-authors, small datasets will tend to produce negative estimates of $\alpha$ even if authors assort randomly with respect to gender (see main text).
+This suggests that $\alpha'$ is a better measure of homophily and heterophily, though the improvement is trivial in large enough samples.
 
@@
 
-S2 Data: This file gives the number and percentage of paper that have 1, 2, 3, 4, or ${\ge}5$ authors for each _journal_ in the dataset of Holman et al.
-(2018) _PLoS Biology_.
+S2 Data: This file gives the number and percentage of papers that have 1, 2, 3, 4, or ${\ge}5$ authors for each _journal_ in the dataset of Holman et al.
+(2018) _PLOS Biology_.
 Note that the sample sizes include papers for which the gender of one or more authors was not determined by Holman et al.
 
-S3 Data: This file gives the number and percentage of paper that have 1, 2, 3, 4, or ${\ge}5$ authors for each _discipline_ in the dataset of Holman et al.
-(2018) _PLoS Biology_.
+S3 Data: This file gives the number and percentage of papers that have 1, 2, 3, 4, or ${\ge}5$ authors for each _discipline_ in the dataset of Holman et al.
+(2018) _PLOS Biology_.
 Note that the sample sizes include papers for which the gender of one or more authors was not determined by Holman et al.

### 2018-10-29 manuscript/Response_letter.Rmd
Resubmission to PLOS ONE

@@
 **Please let us know if any of that still seems excessive upon a second reading.
-Also, a slope of -0.012 is actually quite a lot (perhaps you thought the slope was a correlation coefficient?).
+Also, a regression slope of -0.012 is actually quite a lot, for this dataset.
 The mean $\alpha'$ across journals is about 0.07, and so this slope means that increasing the journal impact factor by one standard deviation above the mean decreases $\alpha'$ from about 0.070 to 0.058; in proportional terms, this is 0.058 / 0.07 = 83% of the mean.
@@
 
-**It says ‘significant heterophily’ (not heterogeneity), meaning that the estimate of $\alpha$ is negative and its 95% CIs are also negative.
-We explained how we defined statistical significant in the Methods (page 15); essentially, we define $\alpha$ values that we are >95% certain are non-zero as statistically significant.**
+**Line 141 (in the original draft) actually says ‘significant heterophily’, where heterophily is our term for an excess of opposite-sex coauthor pairs relative to the randome expection (defined in the Introduction).
+We define 'significant' heterophily as the case where our measure of assortment by gender, $\alpha$, is negative and has 95% confidence limits that lie entirely below zero.
+We explain this definition of statistical significance in the Methods (page 15).**
 
@@
 Given our sample size (36 million authors), it hopefully goes without saying that automatic methods were used, but we now point this out explicitly in the Methods.
-Our earlier paper has extensive discussion of the geographic origin issue, and success rates for each of >100 different countries (almost all of which are over 70%, with China and Korea being the only significant exceptions among the major research-producing nations).**
+Our earlier paper has extensive discussion of the geographic origin issue, and success rates for each of >100 different countries (almost all of which are over 70%, with China and Korea being significant exceptions among the major research-producing nations).**
 
@@
 
-**Actually, the purpose of the first author vs last author analysis was to see if we can get some clues as to whether homophily is driven mostly by senior or junior scientists.
-We discussed both possibilities.
+**Indeed, the purpose of the first author vs last author analysis was to see if we can get some clues as to whether homophily is driven mostly by senior or junior scientists: we outlined both possibilities in this section.
 I think you were misled because the two sentences beginning on line 362 use the word "Assume" rhetorically; the sentences basically say, "Assuming senior researchers drive the results, we predict X.
 This is because of Y." We do not start from the assumption that senior researchers are responsible for homophily.
-We also began that paragraph by pointing out that the 'senior researchers drive homophily' hypothesis is already common in the literature (to our knowledge it has not been effectively tested before).**
+We also began that paragraph by pointing out that the 'senior researchers drive homophily' hypothesis is already predominant in the literature (though to our knowledge it has not been effectively tested before).**
 
@@
 However, it would not affect our test statistic $\alpha'$, because the proportion of male coauthors would be elevated equally for men and women (and so $\alpha'$ would not change).
-You might ask, what if the super-author has preference for one gender of coauthors?
+You might ask, what if the super-author has preference for one gender of coauthors -- wouldn't that skew the results somehow?
 But then, $\alpha'$ would be working as intended: it would correctly record that homophilous (or heterophilous) co-authorships were more common than expected under random assortment by gender.
-This highlights a feature of $\alpha'$ noted earlier by the reviewer: it is a population-level measure, so we cannot make statements about individuals (e.g.
-we don't know if everyone is a little bit homophilous, or if there are a few prolific and very homophilous people).
+This highlights a feature of $\alpha'$ noted earlier by the reviewer: it is a population-level measure, so we cannot make statements about individuals, e.g.
+we don't know if everyone is a little bit homophilous, or if there are a few prolific and very homophilous people (though the former is probably closest to the truth, since super-prolific authors with atypical levels of homophily are presumably the exception not the rule).**
 
@@
 **We originally used the word 'preference' a bit incautiously, and now use terms like 'assortment' that are more agnostic about the mechanism.
-We do think individual-level preferences are likely to be partly responsible, however, and so we discuss preference in a few places.
-This seems like an uncontroversial hypothesis to explain our results.**
+We do think individual-level preferences are likely to be partly responsible, however, and so we discuss preference in a few places, while spelling out that it is one possible interpretation for the results.**
 

### 2018-10-29 manuscript/Manuscript_Claire.Rmd
Resubmission to PLOS ONE

@@
 ---
-title: Researchers preferentially collaborate with same-gendered colleagues across
-  the life sciences
+title: Researchers collaborate with same-gendered colleagues more often than expected across the life sciences
 author:
@@
 In addition to ensuring fairness and helping to redress gender imbalance in the academic workforce, increasing women's access to collaboration could help scientific progress by drawing on more of the available human capital.
-Here, we test whether researchers preferentially collaborate with same-gendered colleagues, using more stringent methods and a larger dataset than in past work.
-Our results reaffirm that researchers preferentially co-publish with colleagues of the same gender, and show that this 'gender homophily' is slightly stronger today than it was 10 years ago.
+Here, we test whether researchers tend to collaborate with same-gendered colleagues, using more stringent methods and a larger dataset than in past work.
+Our results reaffirm that researchers co-publish with colleagues of the same gender more often than expected by chance, and show that this 'gender homophily' is slightly stronger today than it was 10 years ago.
 Contrary to our expectations, we found no evidence that homophily is driven mostly by senior academics, and no evidence that homophily is stronger in fields where women are in the minority.
@@
 knitr::opts_chunk$set(echo = FALSE)
-
-# Load some little files with numbers to be inserted into the MS
-# sample.size <- read.csv("sample.size.table.csv", stringsAsFactors = FALSE)
-# sample.size[,2] <- prettyNum(sample.size[,2], big.mark = ",")
-# sample.size[,3] <- prettyNum(sample.size[,3], big.mark = ",")
-# names(sample.size)[2:3] <- c("n2005", "n2015")
 

### 2018-10-29 manuscript/Manuscript_Claire.tex
Resubmission to PLOS ONE

@@
 \hypersetup{unicode=true,
-            pdftitle={Researchers preferentially collaborate with same-gendered colleagues across the life sciences},
+            pdftitle={Researchers collaborate with same-gendered colleagues more often than expected across the life sciences},
             pdfauthor={Luke Holman* and Claire Morandin; *;},
@@
 \setlength{\droptitle}{-2em}
-  \title{Researchers preferentially collaborate with same-gendered colleagues
-across the life sciences}
+
+  \title{Researchers collaborate with same-gendered colleagues more often than
+expected across the life sciences}
     \pretitle{\vspace{\droptitle}\centering\huge}
@@
 scientific progress by drawing on more of the available human capital.
-Here, we test whether researchers preferentially collaborate with
-same-gendered colleagues, using more stringent methods and a larger
-dataset than in past work. Our results reaffirm that researchers
-preferentially co-publish with colleagues of the same gender, and show
-that this `gender homophily' is slightly stronger today than it was 10
-years ago. Contrary to our expectations, we found no evidence that
+Here, we test whether researchers tend to collaborate with same-gendered
+colleagues, using more stringent methods and a larger dataset than in
+past work. Our results reaffirm that researchers co-publish with
+colleagues of the same gender more often than expected by chance, and
+show that this `gender homophily' is slightly stronger today than it was
+10 years ago. Contrary to our expectations, we found no evidence that
 homophily is driven mostly by senior academics, and no evidence that
@@
 gender homophily in science. \vspace{5mm}
-\par\noindent \textbf{Keywords:} Gender bias, Homophily, Research
+
+\par
+
+\noindent \textbf{Keywords:} Gender bias, Homophily, Research
 collaboration, Text mining, Women in STEM.
@@
 
-\maketitle
-
-\unmarkedfntext{*School of BioSciences, The University of Melbourne, Victoria, Australia.}
+\maketitle\unmarkedfntext{*School of BioSciences, The University of Melbourne, Victoria, Australia.}
 \maketitle\unmarkedfntext{\textsuperscript{§}Organismal and Evolutionary Biology Research Programme,
@@
 
-\section{Introduction}\label{introduction}
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
@@
 
-However, we believe that prior studies of gender homophily were hindered
-by a largely unacknowledged statistical issue that we name the Wahlund
-effect (\autoref{wahlund_plot}), by analogy with the conceptually
-similar Whalund effect in population genetics {[}59{]}. The Wahlund
-effect makes it deceptively difficult to infer gender-based preferences
-simply by counting the number of same- and mixed-gender coauthorships.
-Essentially, whenever coauthorship data are sampled from two or more
-discrete sets of literature, which vary in the author gender ratio and
-which are largely not connected by collaboration, the number of
-same-gendered coauthors will be inflated. This can give the impression
-that authors preferentially publish with same-gendered colleagues even
-if no gender preferences exist, or if the true preference is for
-opposite-gendered colleagues (`gender heterophily'). For example, a
-sample of literature containing bioinformatics and cell biology papers
-will probably contain an excess of mostly-male and mostly-female author
-lists, simply because researchers usually collaborate within their own
-discipline, and because the author gender ratio is more male-biased in
-bioinformatics than in cell biology {[}5{]}.
-
-\begin{figure}[htbp]
+However, we believe that earlier studies of gender homophily were
+hindered by a largely unacknowledged statistical issue that we will
+refer to as the Wahlund effect (\autoref{wahlund_plot}), by analogy with
+the conceptually similar Whalund effect in population genetics {[}59{]}.
+The Wahlund effect makes it deceptively difficult to infer gender-based
+preferences simply by counting the number of same- and mixed-gender
+coauthorships. Essentially, whenever coauthorship data are sampled from
+two or more discrete sets of literature, which vary in the author gender
+ratio and which are largely not connected by collaboration, the number
+of same-gendered coauthors will be inflated. This can give the
+impression that authors preferentially publish with same-gendered
+colleagues even if no gender preferences exist, or if the true
+preference is for opposite-gendered colleagues (`gender heterophily').
+For example, a sample of literature containing a mixture of
+bioinformatics and cell biology papers will probably contain an excess
+of mostly-male and mostly-female author lists, simply because
+researchers usually collaborate within their own discipline, and because
+the author gender ratio is more male-biased in bioinformatics than in
+cell biology {[}5{]}.
+
+\begin{figure}
 \centering
 \includegraphics{../figures/Fig1.pdf}
-\caption{The Wahlund effect can make it appear as if authors prefer to
-publish with same-gendered colleagues, even if no such preference
-exists. Here, coloured circles represent male and female authors, and
-coauthors are linked with lines. Across the whole set of ten papers,
-there is an apparent excess of same-gender collaborations: there are six
-same-gender papers and only four mixed-gender papers, which is fewer
-than the \(10\times2\times0.5\times0.5 = 5\) mixed-gender papers
-expected under the null hypothesis that authors assort randomly with
-respect to gender. However, within each subset, there is no evidence
-that authors prefer to publish with same-gendered individuals (if
-anything, this small dataset suggests gender heterophily). The Wahlund
-effect will tend to inflate the frequency of same-sex coauthorships
-whenever the data is composed of two or more disconnected subsets of
-literature with different author gender ratios; these subsets could be
-research disciplines, older versus newer papers, or papers from authors
-in different countries. The example countries and disciplines were
-selected based on {[}5{]}. \label{wahlund_plot}}
+\caption{The Wahlund effect can make it appear as if authors publish
+with same-gendered colleagues disproportionately often, even if
+collaboration is completely random with respect to gender. Here,
+coloured circles represent male and female authors, and coauthors are
+linked with lines. Across the whole set of ten papers, there is an
+apparent excess of same-gender collaborations: there are six same-gender
+papers and only four mixed-gender papers, which is fewer than the
+\(10\times2\times0.5\times0.5 = 5\) mixed-gender papers expected under
+the null hypothesis that authors assort randomly. However, within each
+subset, there is no evidence that authors prefer to publish with
+same-gendered individuals (if anything, this small dataset suggests
+gender heterophily). The Wahlund effect will tend to inflate the
+frequency of same-gender coauthorships whenever the data is composed of
+two or more disconnected subsets of literature with different author
+gender ratios; these subsets could be research disciplines, older versus
+newer papers, or papers from authors in different countries. The example
+countries and disciplines were selected based on data in {[}5{]}.
+\label{wahlund_plot}}
 \end{figure}
@@
 Wahlund effect as strictly as possible. We use a recently-published
-dataset describing the gender of 35.5m authors from 9.15m articles
-indexed on PubMed {[}5{]}. Holman et al. {[}5{]} reported large
+dataset describing the gender of 35.5 million authors from 9.15 million
+articles indexed on PubMed {[}5{]}. Holman et al. {[}5{]} reported large
 differences in the gender ratio of authors across research disciplines,
@@
 which is positive when same-gender authors publish together more often
-than expected (gender homohily), negative when opposite-gender authors
+than expected (gender homophily), negative when opposite-gender authors
 publish together more often than expected (heterophily), and equal to
@@
 
-\section{Results}\label{results}
-
+\hypertarget{results}{%
+\section{Results}\label{results}}
+
+\hypertarget{gender-homophily-by-discipline-time-period-and-authorship-position}{%
 \subsection{Gender homophily by discipline, time period, and authorship
-position}\label{gender-homophily-by-discipline-time-period-and-authorship-position}
+position}\label{gender-homophily-by-discipline-time-period-and-authorship-position}}
 
@@
 \(\alpha'\) (77-92\%, depending on time period and author type; S1
-Data), and for many of these the FDR-corrected p-values suggested that
-\(\alpha'\) was significantly greater than zero (1469/2077 journals were
-significant in 2015-16, and 404/1192 in 2005-6; S1 Data). Only 2/2077
-journals had statistically significantly heterophily (i.e.
-\(\alpha' < 0\)) in 2015-16, and 1/1192 in 2005-6 (S2 Table). The
+Data), and for many of these the false discovery rate (FDR)-corrected
+p-values suggested that \(\alpha'\) was significantly greater than zero
+(1469/2077 journals were significant in 2015-16, and 404/1192 in 2005-6;
+S1 Data). Only 2/2077 journals had statistically significant heterophily
+(i.e. \(\alpha' < 0\)) in 2015-16, and 1/1192 in 2005-6 (S2 Table). The
 remaining 606 or 787 journals (in 2015 and 2005 respectively) had a
@@
 with the null hypothesis of random assortment with respect to gender. We
-also confirmed that the majority of papers had multiple authors, in most
-journals (S2 Data) and disciplines (S3 Data, S1 Fig).
-
-\begin{figure}[htbp]
+also confirmed that in most journals (S2 Data) and most research
+disciplines (S3 Data, S1 Fig), the majority of papers had multiple
+authors. To verify the argument put forward in Figure 1, we also
+calculated \(\alpha\) across all two-author papers in our entire 15-year
+PubMed dataset (excluding single-author papers, and papers where at
+least one author's gender was unknown; n = \textgreater{}3 million
+papers, \textgreater{}16 million authors). Across the whole dataset,
+\(\alpha\) was 0.126, which is almost double the median value for
+individual journals (Figure 2; \(\alpha = 0.070\) for `All authors'),
+suggesting that lumping together papers from different fields and
+different time periods can indeed inflate \(\alpha\).
+
+\begin{figure}
 \centering
@@
 significant evidence of heterophily (\(\alpha' < 0\)), after false
-discovery rate correction. The white area shows the number of journals
-for which homophily was significantly stronger than expected under the
-null hypothesis (corrected p \textless{} 0.05), while the blue area
-shows all the remainder. Patterns were similar whether \(\alpha'\) was
-calculated for all authors, for first authors only, or for last authors
-only. \label{alpha_histograms}}
+discovery rate correction. In the stacked desnity plot, the white area
+shows the number of journals for which homophily was significantly
+stronger than expected under the null hypothesis (corrected p
+\textless{} 0.05), while the blue area shows all the remainder. Patterns
+were similar whether \(\alpha'\) was calculated for all authors, for
+first authors only, or for last authors only. Points in the right panel
+show \(\alpha'\) for individual journals.\label{alpha_histograms}}
 \end{figure}
@@
 
+\hypertarget{variance-in-homophily-between-disciplines}{%
 \subsection{Variance in homophily between
-disciplines}\label{variance-in-homophily-between-disciplines}
+disciplines}\label{variance-in-homophily-between-disciplines}}
 
@@
 values (\(\alpha'\)) across scientific disciplines. All disciplines had
-positive mean \(\alpha'\), although homophily appeared somewhat stronger
-in some disciplines than others (e.g.~mean \(\alpha'\) was
-\(0.12{\pm}0.02\) for Urology journals and \(0.03{\pm}0.01\) for
-Veterinary Medicine journals; \autoref{alpha_histograms}, S4 Data).
-However, there was no formal evidence for consistent differences in
-\(\alpha'\) between disciplines: the random factor `Discipline'
-explained around 1\% of the variance in \(\alpha'\) in the two linear
-mixed models described in the previous section (see
-\autoref{alpha_histograms} and mixed models in Online Supplementary
-Material). Thus, the processes responsible for producing positive
-\(\alpha'\) values appear to be similarly strong in all the disciplines
-we examined.
+positive mean \(\alpha'\) (averaged over journals), although homophily
+appeared somewhat stronger in some disciplines than others (e.g.~mean
+\(\alpha'\) was \(0.12{\pm}0.02\) for Urology journals and
+\(0.03{\pm}0.01\) for Veterinary Medicine journals;
+\autoref{alpha_histograms}, S4 Data). However, there was no formal
+evidence for consistent differences in \(\alpha'\) between disciplines:
+the random factor `Discipline' explained around 1\% of the variance in
+\(\alpha'\) in the two linear mixed models described in the previous
+section (see \autoref{alpha_histograms} and mixed models in Online
+Supplementary Material). Thus, whatever processes are responsible for
+producing positive \(\alpha'\) values appear to be similarly strong in
+all the disciplines we examined.
 
@@
 due to the Wahlund effect. For example, the journal category
-`Multidisciplinary' -- which includes journals like \emph{PLoS ONE},
-\emph{Nature}, \emph{Science}, and \emph{PNAS} -- did not have notably
+`Multidisciplinary' -- which includes journals like \emph{PLOS ONE},
+\emph{Nature}, \emph{Science}, and \emph{PNAS} -- did not have markedly
 elevated \(\alpha'\) (\autoref{alpha_histograms}). This result suggests
-that our estimates of homophily, and estimates from some earlier
-studies, are not notably inflated by the presence of disparate research
-topics (with variable author gender ratios) being published within
-individual journals.
-
+that our estimates of homophily, and estimates from some of the earlier
+studies listed in the Introduction, are probably not inflated by the
+presence of disparate research topics (with variable author gender
+ratios) being published within individual journals.
+
+\hypertarget{relationship-between-gender-homophily-and-number-of-authors}{%
 \subsection{Relationship between gender homophily and number of
-authors}\label{relationship-between-gender-homophily-and-number-of-authors}
+authors}\label{relationship-between-gender-homophily-and-number-of-authors}}
 
@@
 to the reduced sample size (in terms of number of authors) for the
-2-author papers.
-
-\begin{figure}[htbp]
+2-author papers. One possible explanation for this finding is that
+2-authors papers might be more likely to have an author list that is
+evenly split between career stages (e.g.~a postgraduate student and
+their supervisor), increasing the chance that the authors are mixed
+gender (see Figure 6). The result also suggests that the processes
+responsible for gender homophily are similar in small (e.g.~3 author)
+and larger (5+ authors) collaborations (and across disciplines where
+small versus large collaborations are the norm).
+
+\begin{figure}
 \centering
@@
 
+\hypertarget{relationship-between-gender-homophily-and-gender-ratio}{%
 \subsection{Relationship between gender homophily and gender
-ratio}\label{relationship-between-gender-homophily-and-gender-ratio}
+ratio}\label{relationship-between-gender-homophily-and-gender-ratio}}
 
@@
 (\autoref{alpha_gender_ratio}). Journals with a balanced or
-female-biased author gender ratio tended to have higher \(\alpha'\) than
-journals with a male-biased author gender ratio (GAM smooth terms p
-\textless{} 0.001; Online Supplementary Material). The relationship was
-not statistically significant when \(\alpha'\) was calculated for last
-authors (GAM, p = 0.142), though the trend appeared similar
-(\autoref{alpha_gender_ratio}).
-
-\begin{figure}[htbp]
+female-biased author gender ratio tended to have higher \(\alpha'\)
+(i.e.~stronger homophily) than journals with a male-biased author gender
+ratio (GAM smooth terms p \textless{} 0.001; Online Supplementary
+Material). The relationship was not statistically significant when
+\(\alpha'\) was calculated for last authors (GAM, p = 0.142), though the
+trend appeared similar (\autoref{alpha_gender_ratio}).
+
+\begin{figure}
 \centering
@@
 authors, a positive value denotes an excess of women authors, and zero
-denotes gender parity. The lines were fitted using generalised additive
-models with the smoothing parameter \(k\) set to 3.
-\label{alpha_gender_ratio}}
+denotes gender parity (i.e.~equal numbers of men and women). The lines
+were fitted using generalised additive models with the smoothing
+parameter \(k\) set to 3. \label{alpha_gender_ratio}}
 \end{figure}
 
+\hypertarget{relationship-between-journal-impact-factor-and-gender-homophily}{%
 \subsection{Relationship between journal impact factor and gender
-homophily}\label{relationship-between-journal-impact-factor-and-gender-homophily}
+homophily}\label{relationship-between-journal-impact-factor-and-gender-homophily}}
 
@@
 
-\begin{figure}[htbp]
+\begin{figure}
 \centering
@@
 
+\hypertarget{analysis-accounting-for-differences-in-author-gender-ratio-between-countries}{%
 \subsection{Analysis accounting for differences in author gender ratio
 between
-countries}\label{analysis-accounting-for-differences-in-author-gender-ratio-between-countries}
+countries}\label{analysis-accounting-for-differences-in-author-gender-ratio-between-countries}}
 
@@
 
+\hypertarget{theoretical-expectations-for-alpha-when-the-gender-ratio-differs-between-career-stages}{%
 \subsection{\texorpdfstring{Theoretical expectations for \(\alpha\) when
 the gender ratio differs between career
-stages}{Theoretical expectations for \textbackslash{}alpha when the gender ratio differs between career stages}}\label{theoretical-expectations-for-alpha-when-the-gender-ratio-differs-between-career-stages}
-
-As shown in \autoref{simulation}, we predict that \(\alpha\) is expected
-to be non-zero, even if collaborators are randomly selected with respect
-to gender, provided that there is a gender gap between career stages.
-The extent to which \(\alpha\) deviates from zero depends on the
-relative frequencies of collaboration within and between career stages.
-When \textgreater{}50\% of collaborations were between early and
-established researchers, we expect gender heterophily (\(\alpha < 0\)).
-Conversely, when \textgreater{}50\% of collaborations occured within
-career stages, we expect gender homophily (\(\alpha > 0\)). In a few
-parameter spaces (shown in red; \autoref{simulation}), \(\alpha\) was
-quite high, and overlapped with the values that we estimated
-(\autoref{alpha_histograms}).
-
-\begin{figure}[htbp]
+stages}{Theoretical expectations for \textbackslash{}alpha when the gender ratio differs between career stages}}\label{theoretical-expectations-for-alpha-when-the-gender-ratio-differs-between-career-stages}}
+
+Given that we cannot identify individual researchers or their career
+stages, we used a simple model to derive the theoretical expectations
+for \(\alpha\) when the gender ratio differs between career stages. As
+shown in \autoref{simulation}, we predict that \(\alpha\) is expected to
+be non-zero, even if collaborators are randomly selected with respect to
+gender, provided that there is a gender gap between career stages. The
+extent to which \(\alpha\) deviates from zero depends on the relative
+frequencies of collaboration within and between career stages (rows and
+columns in Figure 6), and the size of the gender gap between stages (x-
+and y-axes in Figure 6). When \textgreater{}50\% of coauthor pairs
+comprise one early-career and one established researcher, we expect
+gender heterophily (\(\alpha < 0\)) whenever the gender ratio differs
+between career stages. Conversely, when \textgreater{}50\% of
+collaborations are between people at the same career stage, we expect
+gender homophily (\(\alpha > 0\)). In a few parameter spaces (shown in
+red; \autoref{simulation}), \(\alpha\) was quite high, and overlapped
+with the values that we estimated (\autoref{alpha_histograms}).
+
+\begin{figure}
 \centering
@@
 early and established researchers are very rare (e.g. \textless{}10\% of
-the total). Both of these conditions are untrue for most fields: the
-gender gap across careers stages is generally less pronounced {[}1,5{]},
-and it is very common for early-career researchers to co-publish with an
-established mentor {[}61{]}. However, one can get \(\alpha > 0\) for
-realistic combinations of parameters, e.g.~a moderate shortage of women
-in senior positions coupled with a moderate excess of within-career
-stage collaboration, suggesting this effect might contribute to some of
-the observed homophily (in this and previous studies).
+the total). Both of these conditions seem unlikely to be true for most
+fields: the gender gap across careers stages is generally less
+pronounced {[}1,5{]}, and it is very common for early-career researchers
+to co-publish with an established mentor {[}61{]}. However, one can get
+\(\alpha > 0\) for realistic combinations of parameters, e.g.~a moderate
+shortage of women in senior positions coupled with a moderate excess of
+within-career stage collaboration, suggesting this effect might
+contribute to some of the homophily observed by this and previous
+studies.
 
@@
 \(\alpha\) is actually less than zero (blue areas in
-\autoref{simulation}). Therefore, our results might under-estimate the
-extent to which researchers preferentially select same-gendered
-collaborators in some cases.
-
-\section{Discussion}\label{discussion}
-
-We found evidence that researchers preferentially publish with
-same-gendered coauthors, even after implementing stringent controls for
-Wahlund effects (\autoref{wahlund_plot}). Our study therefore reaffirms
-earlier studies' conclusions {[}49--57,62{]} and establishes their
-generality across the life sciences. Relatively few journals had
-\(\alpha'\) values below zero, and almost no journals showed
-statistically significant gender heterophily after controlling for
-multiple testing. The excess of same-gender coauthorships was quite
-large: many journals had \(\alpha' > 0.1\), indicating that the gender
-ratio of men's and women's coauthors differs by \textgreater{}10\% in
-absolute terms. In relative terms, our findings are even more striking:
-for example, if men have 20\% female coauthors and women have 30\% (i.e.
-\(\alpha' = 0.1\) in a field with a typical gender ratio {[}5{]}), then
-women publish with women 50\% more often than men do.
+\autoref{simulation}). Therefore, it is possible that researchers
+preferentially assort with same-gendered collaborators even more
+strongly than implied by other results, at least for certain journals or
+research disciplines.
+
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
+
+We found evidence that researchers work with same-gendered coauthors
+more often than expected under the null model, even after implementing
+stringent controls for Wahlund effects (\autoref{wahlund_plot}). Our
+study therefore reaffirms earlier studies' conclusions {[}49--57,62{]}
+using stricter methodology, and generalises their results across the
+life sciences. Relatively few journals had \(\alpha'\) values below
+zero, and almost no journals showed statistically significant gender
+heterophily after controlling for multiple testing. The excess of
+same-gender coauthorships was quite large: many journals had
+\(\alpha' > 0.1\), indicating that the gender ratio of men's and women's
+coauthors differs by \textgreater{}10\% in absolute terms. In relative
+terms, our findings are even more striking: for example, if men have
+20\% female coauthors and women have 30\% (i.e. \(\alpha' = 0.1\) in a
+field with a typical gender ratio {[}5{]}), then women publish with
+women 50\% more often than men do.
 
@@
 coauthorships. As well as the obvious interpretation -- conscious or
-unconcious selection of same-gender collaborators by men, women, or both
--- our results could be partly explained by uncontrolled Wahlund
-effects. However, we suspect the contribution of these to be minor, for
-four reasons: we found positive \(\alpha'\) after controlling for three
-obvious sources of Wahlund effect; there was no inflation of \(\alpha'\)
-in highly multidisciplinary journals; restricting the data by country
-yielded similar estimates of \(\alpha'\); and we showed that differences
-in gender ratio between career stages are unlikely to fully explain our
-results. On balance, we believe the data suggest that it is likely that
-some researchers do preferentially select same-gendered collaborators,
-although the frequency and strength of this preference is difficult to
-ascertain.
+unconcious selection of same-gender collaborators by men, by women, or
+both -- our results could be partly explained by uncontrolled Wahlund
+effects. However, we suspect the contribution of these uncontrolled
+artefacts to be minor, for four reasons: we found positive \(\alpha'\)
+after controlling for three obvious sources of Wahlund effect; there was
+no inflation of \(\alpha'\) in highly multidisciplinary journals
+relative to specialised journals; restricting the data by country
+yielded similar estimates of \(\alpha'\); and we used a simple model to
+show that differences in gender ratio between career stages are unlikely
+to fully explain our results. On balance, we believe the data suggest
+that it is likely that some researchers preferentially select
+same-gendered collaborators, although it is difficult to ascertain what
+proportion of people show such a preference, or how much the strength of
+the preference varies between individuals. We also note that even in a
+world in which collaboration was completely random with respect to
+gender, a high proportion of individual researchers would have entirely
+same-gendered collaborators by chance alone (especially in gender-biased
+disciplines); thus, individuals who only have same-gendered co-authors
+are not necessarily doing anything differently from people with
+gender-balanced co-authors.
 
@@
 might show the strongest gender homophily, e.g.~because being in the
-minority might increase motivation to seek out same-gendered colleagues.
-Contrary to this hypothesis, we found no evidence that gender homophily
-is restricted to particular disciplines: \(\alpha'\) was similarly high
-across the board (\autoref{alpha_histograms}). Interestingly, gender
-homophily was weakest for journals with a male-biased author gender
-ratio, and strongest in journals with a female-biased author gender
-ratio. This may suggest that men are more likely to preferentially seek
-out male collaborators in fields where men are a minority, relative to
-the homophily displayed by women in fields where women are a minority.
-However, this latter result is only tentatively supported since our
-sample contains few journals in which most authors are women
-(\autoref{alpha_gender_ratio}).
+minority might increase people's motivation to seek out same-gendered
+colleagues. Contrary to this hypothesis, we found no evidence that
+gender homophily is restricted to particular disciplines: \(\alpha'\)
+was similarly high across the board (\autoref{alpha_histograms}).
+Interestingly, gender homophily was weakest for journals with a
+male-biased author gender ratio, and strongest in journals with a
+female-biased author gender ratio. One possible reason is that men are
+more likely to preferentially seek out male collaborators in fields
+where men are a minority, relative to the homophily displayed by women
+in fields where women are a minority. However, this latter result only
+has tentative statistical support since our sample contains few journals
+in which most authors are women (\autoref{alpha_gender_ratio}).
 
@@
 staff member, making these teams especially likely to be mixed-gender,
-due to the elevated gender gap among senior researchers {[}1,5{]}.
+due to the greater shortage of women among senior researchers {[}1,5{]}.
 Assuming this interpretation is correct, this result suggests that our
@@
 scientific structures'' because they do not collaborate often enough
-with other women (their Figure 7 suggests that coauthorships between
-women are 30\% more frequent than expected under random assortment). By
-contrast, we feel that it is not helpful to recommend that women
-collaborate primarily with other women, e.g.~because this constrains
-women's options and may be counter-productive (particularly in fields
-like engineering, where 90\% of professors are men {[}1{]}). Instead, we
-suggest that researchers of both genders can help to close the gender
-gap in STEMM. In the context of collaboration, one way to do this is to
-undertake self-examination to ensure that one is not inadvertently
-overlooking or excluding female potential students and colleagues. One
-should also take care to treat male and female collaborators equally,
-e.g.~in terms of training and mentoring, allocation of work, and how one
-frames or promotes the collaboration (e.g.~in conference presentations
-or on a website); evidence suggests that unconscious bias causes people
-to undervalue women's research achievements {[}20{]}, and possibly to
-assign menial or under-valued tasks to women and more prestigious tasks
-to men {[}61{]}.
+with other women (for reference, their Figure 7 implies that
+coauthorships involving two women are 30\% more frequent than expected
+under random assortment). By contrast, we feel that it may be
+counter-productive to recommend that women collaborate primarily with
+other women, e.g.~because this constrains women's options (particularly
+in fields like engineering, where 90\% of professors are men {[}1{]}).
+Instead, we suggest that researchers of both genders can help to close
+the gender gap in STEMM. In the context of collaboration, one way to do
+this is to undertake self-examination to ensure that one is not
+inadvertently overlooking or excluding women among potential students
+and colleagues. One should also take care to treat male and female
+collaborators equally, e.g.~in terms of training and mentoring,
+allocation of work, how one frames or promotes the collaboration
+(e.g.~in conference presentations or on a website), and in terms of the
+credit afforded to collaborators' contributions. Experimental work
+suggests that unconscious bias causes people to undervalue women's
+research achievements {[}20{]}, and a study of author contribution
+statements found observational evidence that menial or under-valued
+tasks are assigned to women while more prestigious tasks are assigned to
+men {[}61{]}.
 
@@
 non-elite male academics. This finding could contribute to the homophily
-we observed, and is cause for concern since Sheltzer and Smith
-{[}66{]}'s results might reflect discrimination against women during
-hiring {[}20{]}, or avoidance by women of elite research groups
-(e.g.~due to gender differences in confidence, or a perception that some
-groups are sexist). We also found a little evidence that gender
-homophily is detrimental to research quality, in that high-impact
-journals tended to have weaker homophily. Assuming that papers published
-in high-impact journals are of higher average quality {[}67{]}, our
-results provide non-experimental support for the hypothesis that
-mixed-gender teams produce better research than single-gender teams
-{[}42--48{]}. Another issue is that if many collaborations are between
-established researchers, there will be an excess of male-male
-collaborations in fields where women in senior positions are rare; some
-of the observed homophily might therefore reflect the elevated gender
-gap among senior researchers.
+we observed, and is cause for concern since the results might reflect
+discrimination against women during hiring {[}20{]}, or avoidance by
+women of elite research groups (e.g.~due to gender differences in
+confidence, or a perception that some groups are sexist). We also found
+a little evidence that gender homophily is detrimental to research
+quality, in that high-impact journals tended to have weaker homophily.
+Assuming that papers published in high-impact journals are of higher
+average quality (which is contentious; {[}67{]}), our results provide
+non-experimental support for the hypothesis that mixed-gender teams
+produce better research than single-gender teams {[}42--48{]}. Another
+issue is that if many collaborations are between established
+researchers, there will be an excess of male-male collaborations in
+fields where women in senior positions are rare; some of the observed
+homophily might therefore reflect the elevated gender gap among senior
+researchers.
 
@@
 Collaboration is often most enjoyable and productive when working with
-like-minded people, who might be same-gendered more often than not. We
-also suppose that some people consciously choose to preferentially
-collaborate with women in order to help close the gender gap in the
-workforce; this would create homophily if women do this more than men.
-In support of this interpretation, women appear more likely than men to
-promote the work of female colleagues by inviting them to give talks
-{[}68,69{]}. Given that many collaborative research projects
-unfortunately involve a gendered division of labour {[}61{]}, working
-with a same-gendered colleague may provide exposure to new parts of the
-research process, and (especially for the minority gender) a welcome
-change of pace.
-
-\section{Methods}\label{methods}
-
-\subsection{The dataset}\label{the-dataset}
+like-minded people, who might tend to be same-gendered more often than
+not. We also suppose that some people consciously choose to
+preferentially collaborate with women in order to help close the gender
+gap in the workforce; this would create homophily if women do this more
+than men. In support of this interpretation, there is some evidence that
+women are more likely than men to promote the work of female colleagues
+by inviting them to give talks {[}68,69{]}. Given that many
+collaborative research projects unfortunately involve a gendered
+division of labour {[}61{]}, working with a same-gendered colleague may
+provide exposure to new parts of the research process, and (especially
+for the minority gender) a welcome change of pace.
+
+\hypertarget{methods}{%
+\section{Methods}\label{methods}}
+
+\hypertarget{the-dataset}{%
+\subsection{The dataset}\label{the-dataset}}
 
@@
 on PubMed and attempting to infer each author's gender from their given
-name. Each journal was assigned to one of 107 scientific disciplines,
-using PubMed's journal categorisations in the interests of objectivity.
-Because the present study focuses on co-authorship, all single-author
-papers were discarded. We also discarded all papers for which we could
-not determine the gender of every author with \({\ge}95\%\) certainty,
-in order to simplify the statistical analysis. To mitigate Wahlund
-effects caused by variation in the gender ratio of researchers over time
-(see below), we also discarded all papers except those that were
-published either 0-1 or 10-11 years before the PubMed data were
-collected (i.e.~20\textsuperscript{th} August 2016). Lastly, we excluded
+name using computational methods. Each journal was assigned to one of
+107 scientific disciplines, using PubMed's journal categorisations in
+the interests of objectivity. Because the present study focuses on
+co-authorship, all single-author papers were discarded. We also
+discarded all papers for which we could not determine the gender of
+every author with \({\ge}95\%\) certainty, in order to simplify the
+statistical analysis. To mitigate Wahlund effects caused by variation in
+the gender ratio of researchers over time (see below), we only kept
+papers with publicaiton dates falling in two one-year time periods,
+namely 0-1 or 10-11 years prior to the collection date of the PubMed
+data (i.e.~20\textsuperscript{th} August 2016). Lastly, we excluded
 journals with fewer than 50 suitable papers. Detailed sample size
@@
 
+\hypertarget{calculating-alpha-the-coefficient-of-homophily}{%
 \subsection{\texorpdfstring{Calculating \(\alpha\), the coefficient of
-homophily}{Calculating \textbackslash{}alpha, the coefficient of homophily}}\label{calculating-alpha-the-coefficient-of-homophily}
+homophily}{Calculating \textbackslash{}alpha, the coefficient of homophily}}\label{calculating-alpha-the-coefficient-of-homophily}}
 
@@
 authors randomly assort, but for small datasets this assumption is not
-always true (as pointed out by Carl Bergstrom in a blog post,
-\url{http://www.eigenfactor.org/gender/assortativity/note_to_eisen.rtf}).
-To borrow Prof.~Bergstrom's example, consider a small research specialty
-comprising just two men and two women researchers, who have together
-produced six two-author papers: one in each of the six possible
-two-author combinations. For these six papers,
-\(\alpha = -\frac{1}{3}\), even though same- and opposite-gendered
-coauthors were selected in equal proportion to their frequency in the
-pool of possible collaborators.
+always true. Essentially, this issue arises because a person cannot be
+their own co-author. In a small dataset comprising \(m\) men and \(f\)
+women, a man can co-author with \(m - 1\) men while women can co-author
+with \(m\) men. Thus, the null expectation for \(alpha\) is a negative
+number -- potentially a large one if \(m\) and \(f\) are very small.
 
@@
 
+\hypertarget{minimising-the-wahlund-effect-research-discipline-and-time-period}{%
 \subsection{Minimising the Wahlund effect: research discipline and time
-period}\label{minimising-the-wahlund-effect-research-discipline-and-time-period}
+period}\label{minimising-the-wahlund-effect-research-discipline-and-time-period}}
 
@@
 individual journals, since papers from the same journal typically focus
-on closely related topics. Although some journals, e.g. \emph{PLoS ONE},
+on closely related topics. Although some journals, e.g. \emph{PLOS ONE},
 publish research from diverse disciplines with very different author
@@
 in \(\alpha'\) between highly multidisciplinary and more specialised
-journals, e.g. \emph{PLoS ONE} versus \emph{PLoS Computational Biology},
-gives an estimate of the extent to which multidisciplinarity inflates
-\(\alpha'\).
+journals, e.g. \emph{PLOS ONE} versus \emph{PLOS Computational Biology},
+gives an estimate of the extent to which multidisciplinarity within
+journals inflates \(\alpha'\).
 
@@
 
+\hypertarget{minimising-the-wahlund-effect-author-country-of-affiliation}{%
 \subsection{Minimising the Wahlund effect: author country of
-affiliation}\label{minimising-the-wahlund-effect-author-country-of-affiliation}
+affiliation}\label{minimising-the-wahlund-effect-author-country-of-affiliation}}
 
@@
 ratio of researchers from different countries. For example, Holman et
-al. {[}5{]} showed that PubMed-indexed authors based in Serbia are more
+al. {[}5{]} found that PubMed-indexed authors based in Serbia are more
 than twice as likely to be women as are authors based in Japan.
@@
 
+\hypertarget{calculating-standardised-journal-impact-factor}{%
 \subsection{Calculating standardised journal impact
-factor}\label{calculating-standardised-journal-impact-factor}
+factor}\label{calculating-standardised-journal-impact-factor}}
 
 We obtained the 3-year impact factor for each journal from Clarivate
-Analytics. To account for large differences in impact factor between
-disciplines, we took the the residuals from a model with \(Log_{10}\)
-impact factor as the response and the research discipline of the journal
-as a random effect. Thus, journals with a positive standardised impact
-factor have a higher mean number of citations than the average for
-journals in their discipline. We then used Spearman rank correlation to
-test whether \(\alpha'\) was correlated with impact factor across
-journals.
-
-\subsection{Statistical analysis}\label{statistical-analysis}
+Analytics (formerly ISI). To account for large differences in impact
+factor between disciplines, we took the the residuals from a model with
+\(log_{10}\) impact factor as the response and the research discipline
+of the journal as a random effect. Thus, journals with a positive
+standardised impact factor have a higher mean number of citations than
+the average for journals in their discipline. We then used Spearman rank
+correlation to test whether \(\alpha'\) was correlated with impact
+factor across journals.
+
+\hypertarget{statistical-analysis}{%
+\subsection{Statistical analysis}\label{statistical-analysis}}
 
@@
 
+\hypertarget{theoretical-expectations-for-alpha-when-the-gender-ratio-differs-between-career-stages-1}{%
 \subsection{\texorpdfstring{Theoretical expectations for \(\alpha\) when
 the gender ratio differs between career
-stages}{Theoretical expectations for \textbackslash{}alpha when the gender ratio differs between career stages}}\label{theoretical-expectations-for-alpha-when-the-gender-ratio-differs-between-career-stages-1}
-
-In many STEMM subjects, the gender ratio is more skewed among
-established researchers relative to early-career researchers {[}1,5{]}.
-We hypothesised that this skew could potentially create both Wahlund
+stages}{Theoretical expectations for \textbackslash{}alpha when the gender ratio differs between career stages}}\label{theoretical-expectations-for-alpha-when-the-gender-ratio-differs-between-career-stages-1}}
+
+In most STEMM subjects, the gender ratio is more skewed among
+established researchers relative to early-career researchers, due both
+to women leaving STEMM careers at greater rates (the `leaky pipeline'),
+and to historical shortages of women studying STEMM (`demographic
+inertia') {[}1,5{]}. We hypothesised that this difference in gender
+ratio between career stages could potentially create both Wahlund
 effects and `reverse' Wahlund effects. For example, imagine that the
-majority of collaborations are between students and professors, and that
-the gender ratio differs between career stages: we will then see an
-excess of mixed-gender coauthorships (heterophily, \(\alpha < 0\)), even
-if gender has no direct, causal effect. Similarly, a hypothetical field
-in which students work only with students, and professors with
-professors, would have apparent gender homophily (\(\alpha > 0\)).
+majority of collaborations in a particular field are between students
+and professors, and that the gender ratio differs between career stages:
+we would then see an excess of mixed-gender coauthorships (heterophily,
+\(\alpha < 0\)), even if gender has no direct, causal effect. Similarly,
+a hypothetical field in which students work only with students, and
+professors with professors, would have apparent gender homophily
+(\(\alpha > 0\)).
 
@@
 our dataset, which contains no information on career stage. Therefore,
-we instead decided to derive the theoretical expectations for \(\alpha\)
+we instead decided to derive theoretical expectations for \(\alpha\)
 when there is a difference in gender ratio across career stages, in
 order to determine if and how this effect should alter our inferences.
-For simplicity, our calculations assume there are only two career
-stages, though we intuit that the general conclusions would also apply
-to a multi-tier career ladder. Under the null model that gender has no
-causal effect on collaboration, we calculated \(\alpha\) for various
-combinations of the four free parameters, i.e.~the gender ratios for
-early- and late-career researchers, and the relative frequency of
-collaborations between early-early, early-late, and late-late
-collaborations. We then used the theoretical expectations for \(\alpha\)
-to qualify our main conclusions (see Results). The Online Supplementary
-Material gives annotated R code used to derive the theoretical
-expectations.
-
+For simplicity, our calculations assume there are only two career stages
+(`early-career' and `established'), though we expect that the general
+conclusions would also apply to a multi-tier career ladder. Under the
+null model that gender has no causal effect on collaboration, we
+calculated \(\alpha\) for various combinations of the four free
+parameters in our simple model. These parameters are: the gender ratio
+among early-career researchers (x-axis of Figure 6), the gender ratio
+among established researchers (y-axis of Figure 6), the frequency of
+within- versus between career stage collaborator pairs (rows in Figure
+6), and lastly the frequency of within-stage collaborations that are
+between two early-career researchers as opposed to two late-career
+reserachers (columns in Figure 6). When these four parameters are
+specified, one can easily calculate the relative frequencies of
+collaborator pairs that involve two men, two women, or a man and a
+woman. In short, if we have specified the frequency of women at both
+career stages, as well as the frequency of the three possible types of
+collaboration with respect to career stage (early-early,
+early-established, and established-established), then we can calculate
+the frequency of collaborators pairs comprising two women, or a woman
+and a man, and thus find \(\alpha\) (see the Online Supplementary
+Material for the annotated R code).
+
+\hypertarget{data-availability-and-reproducibility}{%
 \subsection{Data availability and
-reproducibility}\label{data-availability-and-reproducibility}
+reproducibility}\label{data-availability-and-reproducibility}}
 
@@
 
-\section{Acknowledgements}\label{acknowledgements}
+\hypertarget{acknowledgements}{%
+\section{Acknowledgements}\label{acknowledgements}}
 
@@
 
-\section{References}\label{references}
+\hypertarget{references}{%
+\section{References}\label{references}}
 
 \hypertarget{refs}{}
-\hypertarget{ref-Shaw_2012}{}
+\leavevmode\hypertarget{ref-Shaw_2012}{}%
 1. Shaw AK, Stanton DE. Leaks in the pipeline: separating demographic
@@
 
-\hypertarget{ref-Lariviere_2013}{}
+\leavevmode\hypertarget{ref-Lariviere_2013}{}%
 2. Larivière V, Ni C, Gingras Y, Cronin B, Sugimoto CR. Bibliometrics:
@@
 
-\hypertarget{ref-West_2013}{}
+\leavevmode\hypertarget{ref-West_2013}{}%
 3. West JD, Jacquet J, King MM, Correll SJ, Bergstrom CT. The role of
@@
 
-\hypertarget{ref-Elsevier_report}{}
+\leavevmode\hypertarget{ref-Elsevier_report}{}%
 4. Elsevier Report. Gender in the global research landscape.
@@
 
-\hypertarget{ref-Holman_2018}{}
+\leavevmode\hypertarget{ref-Holman_2018}{}%
 5. Holman L, Stuart Fox D, Hauser CE. The gender gap in science: How
@@
 
-\hypertarget{ref-Wutte_2007}{}
+\leavevmode\hypertarget{ref-Wutte_2007}{}%
 6. Wutte M. Closing the gender gap. Nature. 2007;448: NJ101--NJ102.
 
-\hypertarget{ref-Reuben_2014}{}
+\leavevmode\hypertarget{ref-Reuben_2014}{}%
 7. Reuben E, Sapienza P, Zingales L. How stereotypes impair women's
@@
 
-\hypertarget{ref-Trower_2002}{}
+\leavevmode\hypertarget{ref-Trower_2002}{}%
 8. Trower CA, Chait RP. Faculty diversity: Why women and minorities are
@@
 
-\hypertarget{ref-Umbach_2007}{}
+\leavevmode\hypertarget{ref-Umbach_2007}{}%
 9. Umbach PD. Gender equity in the academic labor market: An analysis of
@@
 
-\hypertarget{ref-Hosek_2005}{}
+\leavevmode\hypertarget{ref-Hosek_2005}{}%
 10. Hosek S, Cox AG, Ghosh-Dastidar B, Kofner A, Ramphal N, Scott J, et
@@
 
-\hypertarget{ref-pohlhaus_2011}{}
+\leavevmode\hypertarget{ref-pohlhaus_2011}{}%
 11. Pohlhaus JR, Jiang H, Wagner RM, Schaffer WT, Pinn VW. Sex
@@
 
-\hypertarget{ref-Zuckerman_1987}{}
+\leavevmode\hypertarget{ref-Zuckerman_1987}{}%
 12. Zuckerman H. Persistence and change in the careers of men and women
@@
 
-\hypertarget{ref-Rosenfeld_1991}{}
+\leavevmode\hypertarget{ref-Rosenfeld_1991}{}%
 13. Rosenfeld RA. Outcome analysis of academic careers. Review prepared
@@
 
-\hypertarget{ref-Long_1993}{}
+\leavevmode\hypertarget{ref-Long_1993}{}%
 14. Long JS, Paul DA, Robert M. Rank advancement in academic careers:
@@
 
-\hypertarget{ref-Hopkins_2013}{}
+\leavevmode\hypertarget{ref-Hopkins_2013}{}%
 15. Hopkins AL, Jawitz JW, McCarty C, Goldman A, Basu NB. Disparities in
@@
 
-\hypertarget{ref-ODorchai_2009}{}
+\leavevmode\hypertarget{ref-ODorchai_2009}{}%
 16. O'Dorchai S, Meulders D, Crippa F, Margherita A. She figures
@@
 
-\hypertarget{ref-Feldt_1986}{}
+\leavevmode\hypertarget{ref-Feldt_1986}{}%
 17. Feldt B. The faculty cohort study: School of medicine. Ann Arbor,
@@
 
-\hypertarget{ref-Stack_2004}{}
+\leavevmode\hypertarget{ref-Stack_2004}{}%
 18. Stack S. Gender, children and research productivity. Scientometrics.
@@
 
-\hypertarget{ref-Lariviere_2011}{}
+\leavevmode\hypertarget{ref-Lariviere_2011}{}%
 19. Larivière V, Vignola-Gagné E, Villeneuve C, Gélinas P, Gingras Y.
@@
 
-\hypertarget{ref-Moss_2012}{}
+\leavevmode\hypertarget{ref-Moss_2012}{}%
 20. Moss-Racusin CA, Dovidio JF, Brescoll VL, Graham MJ, Handelsman J.
@@
 
-\hypertarget{ref-Knobloch_2013}{}
+\leavevmode\hypertarget{ref-Knobloch_2013}{}%
 21. Knobloch-Westerwick S, Glynn CJ, Huge M. Science faculty's subtle
@@
 
-\hypertarget{ref-Lee_2005}{}
+\leavevmode\hypertarget{ref-Lee_2005}{}%
 22. Lee S, Bozeman B. The impact of research collaboration on scientific
@@
 
-\hypertarget{ref-Wuchty_2007}{}
+\leavevmode\hypertarget{ref-Wuchty_2007}{}%
 23. Wuchty S, Jones BF, Uzzi B. The increasing dominance of teams in
@@
 
-\hypertarget{ref-Abramo_2009}{}
+\leavevmode\hypertarget{ref-Abramo_2009}{}%
 24. Abramo G, D'Angelo CA, Di Costa F. Research collaboration and
@@
 
-\hypertarget{ref-Lariviere_2015}{}
+\leavevmode\hypertarget{ref-Lariviere_2015}{}%
 25. Larivière V, Gingras Y, Sugimoto CR, Tsou A. Team size matters:
@@
 
-\hypertarget{ref-Long_1992}{}
+\leavevmode\hypertarget{ref-Long_1992}{}%
 26. Long JS. Measures of sex differences in scientific productivity.
@@
 
-\hypertarget{ref-Bozeman_2011}{}
+\leavevmode\hypertarget{ref-Bozeman_2011}{}%
 27. Bozeman B, Monica G. How do men and women differ in research
@@
 
-\hypertarget{ref-Abramo_2013}{}
+\leavevmode\hypertarget{ref-Abramo_2013}{}%
 28. Abramo G, D'Angelo CA, Di Costa F. Gender differences in research
@@
 
-\hypertarget{ref-Badar_2013}{}
+\leavevmode\hypertarget{ref-Badar_2013}{}%
 29. Badar K, Hite JM, Badir YF. Examining the relationship of
@@
 
-\hypertarget{ref-Lewison_2001}{}
+\leavevmode\hypertarget{ref-Lewison_2001}{}%
 30. Lewison G. The quantity and quality of female researchers: A
@@
 
-\hypertarget{ref-Webster_2001}{}
+\leavevmode\hypertarget{ref-Webster_2001}{}%
 31. Webster BM. Polish women in science: A bibliometric analysis of
@@
 
-\hypertarget{ref-Bozeman_2004}{}
+\leavevmode\hypertarget{ref-Bozeman_2004}{}%
 32. Bozeman B, Corley E. Scientists' collaboration strategies:
@@
 
-\hypertarget{ref-Long_1990}{}
+\leavevmode\hypertarget{ref-Long_1990}{}%
 33. Long JS. The origins of sex differences in science. Social Forces.
@@
 
-\hypertarget{ref-Fuchs_2001}{}
+\leavevmode\hypertarget{ref-Fuchs_2001}{}%
 34. Fuchs S, Von Stebut J, Allmendinger J. Gender, science, and
@@
 
-\hypertarget{ref-Reskin_1978}{}
+\leavevmode\hypertarget{ref-Reskin_1978}{}%
 35. Reskin BF. Scientific productivity, sex, and location in the
@@
 
-\hypertarget{ref-Wright_2003}{}
+\leavevmode\hypertarget{ref-Wright_2003}{}%
 36. Wright AL, Schwindt LA, Bassford TL, Reyna VF, Shisslak PAS
@@
 
-\hypertarget{ref-bleidorn2016age}{}
+\leavevmode\hypertarget{ref-bleidorn2016age}{}%
 37. Bleidorn W, Arslan RC, Denissen JJ, Rentfrow PJ, Gebauer JE, Potter
@@
 
-\hypertarget{ref-jagsi_2016}{}
+\leavevmode\hypertarget{ref-jagsi_2016}{}%
 38. Jagsi R, Griffith KA, Jones R, Perumalswami CR, Ubel P, Stewart A.
@@
 
-\hypertarget{ref-Martin_2014}{}
+\leavevmode\hypertarget{ref-Martin_2014}{}%
 39. Martin JL. Ten simple rules to achieve conference speaker gender
@@
 
-\hypertarget{ref-Tower_2007}{}
+\leavevmode\hypertarget{ref-Tower_2007}{}%
 40. Tower G, Julie P, Brenda R. A multidisciplinary study of
@@
 
-\hypertarget{ref-Jordan_2008}{}
+\leavevmode\hypertarget{ref-Jordan_2008}{}%
 41. Jordan CE, Clark SJ, Vann CE. Do gender differences exist in the
@@
 
-\hypertarget{ref-Britton_2000}{}
+\leavevmode\hypertarget{ref-Britton_2000}{}%
 42. Britton DM. The epistemology of the gendered organization. Gender
@@
 
-\hypertarget{ref-Reagans_2001}{}
+\leavevmode\hypertarget{ref-Reagans_2001}{}%
 43. Reagans R, Zuckerman EW. Networks, diversity, and productivity: The
@@
 
-\hypertarget{ref-Hong_2004}{}
+\leavevmode\hypertarget{ref-Hong_2004}{}%
 44. Hong L, Page SE. Groups of diverse problem solvers can outperform
@@
 
-\hypertarget{ref-Whittington_2008}{}
+\leavevmode\hypertarget{ref-Whittington_2008}{}%
 45. Whittington KB, Smith-Doerr L. Women inventors in context:
@@
 
-\hypertarget{ref-Bear_2011}{}
+\leavevmode\hypertarget{ref-Bear_2011}{}%
 46. Bear JB, Woolley AW. The role of gender in team collaboration and
@@
 
-\hypertarget{ref-Herrera_2012}{}
+\leavevmode\hypertarget{ref-Herrera_2012}{}%
 47. Herrera R, Duncan PA, Green MT, Skaggs SL. The effect of gender on
@@
 
-\hypertarget{ref-Campbell_2013}{}
+\leavevmode\hypertarget{ref-Campbell_2013}{}%
 48. Campbell LG, Mehtani S, Dozier ME, Rinehart J. Gender-heterogeneous
@@
 
-\hypertarget{ref-Ferber_1980}{}
+\leavevmode\hypertarget{ref-Ferber_1980}{}%
 49. Ferber MA, Teiman M. Are women economists at a disadvantage in
@@
 
-\hypertarget{ref-McDowell_1992}{}
+\leavevmode\hypertarget{ref-McDowell_1992}{}%
 50. McDowell JM, Smith JK. The effect of gender-sorting on propensity to
@@
 
-\hypertarget{ref-Ghiasi_2015}{}
+\leavevmode\hypertarget{ref-Ghiasi_2015}{}%
 51. Ghiasi G, Larivière V, Sugimoto CR. On the compliance of women
@@
 
-\hypertarget{ref-Crow_2015}{}
+\leavevmode\hypertarget{ref-Crow_2015}{}%
 52. Crow MS, Smykla JO. An examination of author characteristics in
@@
 
-\hypertarget{ref-Fahmy_2017}{}
+\leavevmode\hypertarget{ref-Fahmy_2017}{}%
 53. Fahmy C, Young JT. Gender inequality and knowledge production in
@@
 
-\hypertarget{ref-Zettler_2017}{}
+\leavevmode\hypertarget{ref-Zettler_2017}{}%
 54. Zettler HR, Stephanie M Cardwell, Jessica M C. The gendering effects
@@
 
-\hypertarget{ref-Jadidi_2017}{}
+\leavevmode\hypertarget{ref-Jadidi_2017}{}%
 55. Jadidi M, Karimi F, Lietz H, Wagner C. Gender disparities in
@@
 
-\hypertarget{ref-Teele_2017}{}
+\leavevmode\hypertarget{ref-Teele_2017}{}%
 56. Teele DL, Kathleen T. Gender in the journals: Publication patterns
@@
 
-\hypertarget{ref-Araujo_2017a}{}
+\leavevmode\hypertarget{ref-Araujo_2017a}{}%
 57. Araújo T, Elsa F. The specific shapes of gender imbalance in
@@
 
-\hypertarget{ref-Araujo_2017b}{}
+\leavevmode\hypertarget{ref-Araujo_2017b}{}%
 58. Araújo T, Elsa F. Big Missing Data: are scientific memes inherited
@@
 
-\hypertarget{ref-Wahlund_1928}{}
+\leavevmode\hypertarget{ref-Wahlund_1928}{}%
 59. Wahlund S. Zusammensetzung von populationen und
@@
 
-\hypertarget{ref-bergstrom_2016}{}
+\leavevmode\hypertarget{ref-bergstrom_2016}{}%
 60. Bergstrom T, Bergstrom C, King M, Jacquet J, West J, Correll S. A
-note on measuring gender homophily among scholarly authors.
-http://eigenfactororg/gender/ assortativity/measuring\_homophilypdf.
-2016;
-
-\hypertarget{ref-macaluso_2016}{}
+note on measuring gender homophily among scholarly authors. 2016;
+
+\leavevmode\hypertarget{ref-macaluso_2016}{}%
 61. Macaluso B, Larivière V, Sugimoto T, Sugimoto CR. Is science built
@@
 
-\hypertarget{ref-Bentley_2003}{}
+\leavevmode\hypertarget{ref-Bentley_2003}{}%
 62. Bentley JT, Adamson R. Gender differences in the careers of academic
@@
 
-\hypertarget{ref-Long_2015}{}
+\leavevmode\hypertarget{ref-Long_2015}{}%
 63. Long MT, Leszczynski A, Thompson KD, Wasan SK, Calderwood AH. Female
@@
 
-\hypertarget{ref-Bendels_2018}{}
+\leavevmode\hypertarget{ref-Bendels_2018}{}%
 64. Bendels MH, Bauer J, Schöffel N, Groneberg DA. The gender gap in
@@
 
-\hypertarget{ref-McKenzie_2017}{}
+\leavevmode\hypertarget{ref-McKenzie_2017}{}%
 65. McKenzie K, Ramonas M, Patlas M, Katz DS. Assessing the gap in
@@
 
-\hypertarget{ref-sheltzer_2014}{}
+\leavevmode\hypertarget{ref-sheltzer_2014}{}%
 66. Sheltzer JM, Smith JC. Elite male faculty in the life sciences
@@
 
-\hypertarget{ref-Garfield_2006}{}
+\leavevmode\hypertarget{ref-Garfield_2006}{}%
 67. Garfield E. The history and meaning of the journal impact factor.
@@
 
-\hypertarget{ref-nittrouer_2018}{}
+\leavevmode\hypertarget{ref-nittrouer_2018}{}%
 68. Nittrouer CL, Hebl MR, Ashburn-Nardo L, Trump-Steele RC, Lane DM,
@@
 
-\hypertarget{ref-debarre_2018}{}
+\leavevmode\hypertarget{ref-debarre_2018}{}%
 69. Débarre F, Rode N, Ugelvig L. Gender equity at scientific events.
@@
 
-\hypertarget{ref-wright_1949}{}
+\leavevmode\hypertarget{ref-wright_1949}{}%
 70. Wright S. The genetical structure of populations. Annals of Human
@@
 
-\hypertarget{ref-wang_2016}{}
+\leavevmode\hypertarget{ref-wang_2016}{}%
 71. Wang YS, Erosheva EA. On the relationship between set-based and
@@
 
-\hypertarget{ref-Benjamini_1995}{}
+\leavevmode\hypertarget{ref-Benjamini_1995}{}%
 72. Benjamini Y, Hochberg Y. Controlling the false discovery rate: a
@@
 
-\hypertarget{ref-Bonham_2017}{}
+\leavevmode\hypertarget{ref-Bonham_2017}{}%
 73. Bonham KS, Stefan MI. Women are underrepresented in computational
@@
 
-\hypertarget{ref-Wren_2007}{}
+\leavevmode\hypertarget{ref-Wren_2007}{}%
 74. Wren JD, Kozak KZ, Johnson KR, Deakyne SJ, Schilling LM, Dellavalle
@@
 
-\hypertarget{ref-burkner_2016}{}
+\leavevmode\hypertarget{ref-burkner_2016}{}%
 75. Bürkner P-C. brms: An R package for Bayesian multilevel models using
@@
 
-\section{Supporting information}\label{supporting-information}
-
-\subsection{Supplementary figures}\label{supplementary-figures}
+\hypertarget{supporting-information}{%
+\section{Supporting information}\label{supporting-information}}
+
+\hypertarget{supplementary-figures}{%
+\subsection{Supplementary figures}\label{supplementary-figures}}
 
@@
 (\textless{}100), the unadjusted metric \(\alpha\) was downwardly
-biased. This fits our expectations: because authors cannot be their own
-co-authors, small datasets will tend to produce negative estimates of
-\(\alpha\) even if authors assort randomly with respect to gender (see
-main text). This suggests that \(\alpha'\) is a more useful measure of
-homophily and heterophily, especially for small samples.
-
-\subsection{Supplementary tables}\label{supplementary-tables}
+biased. This fits our expectations: because researchers cannot be their
+own co-authors, small datasets will tend to produce negative estimates
+of \(\alpha\) even if authors assort randomly with respect to gender
+(see main text). This suggests that \(\alpha'\) is a better measure of
+homophily and heterophily, though the improvement is trivial in large
+enough samples.
+
+\hypertarget{supplementary-tables}{%
+\subsection{Supplementary tables}\label{supplementary-tables}}
 
@@
 
-\subsection{Supplementary datasets}\label{supplementary-datasets}
+\hypertarget{supplementary-datasets}{%
+\subsection{Supplementary datasets}\label{supplementary-datasets}}
 
@@
 
-S2 Data: This file gives the number and percentage of paper that have 1,
-2, 3, 4, or \({\ge}5\) authors for each \emph{journal} in the dataset of
-Holman et al. (2018) \emph{PLoS Biology}. Note that the sample sizes
+S2 Data: This file gives the number and percentage of papers that have
+1, 2, 3, 4, or \({\ge}5\) authors for each \emph{journal} in the dataset
+of Holman et al. (2018) \emph{PLOS Biology}. Note that the sample sizes
 include papers for which the gender of one or more authors was not
@@
 
-S3 Data: This file gives the number and percentage of paper that have 1,
-2, 3, 4, or \({\ge}5\) authors for each \emph{discipline} in the dataset
-of Holman et al. (2018) \emph{PLoS Biology}. Note that the sample sizes
-include papers for which the gender of one or more authors was not
-determined by Holman et al.
+S3 Data: This file gives the number and percentage of papers that have
+1, 2, 3, 4, or \({\ge}5\) authors for each \emph{discipline} in the
+dataset of Holman et al. (2018) \emph{PLOS Biology}. Note that the
+sample sizes include papers for which the gender of one or more authors
+was not determined by Holman et al.
 

### 2018-11-01 manuscript/Response_letter.Rmd
Ready for resubmission

@@
 Therefore, the fact that we cannot (or rather, did not try to) identify specific individuals does not compromise the paper's results and conclusions.
-We agree that some uses of the word 'preference' in our manuscript could be misleading in this respect, and so in the revision, we have replaced the word in key place with more neutral/agnostic terms (e.g.
-'non-random assortment').**
+We agree that some uses of the word 'preference' in our manuscript could be misleading in this respect, and so in the revision, we have replaced the word in key places (e.g.
+the title of the paper) with more neutral/agnostic terms (e.g.
+'assortment').**
 
@@
 Given our sample size (36 million authors), it hopefully goes without saying that automatic methods were used, but we now point this out explicitly in the Methods.
-Our earlier paper has extensive discussion of the geographic origin issue, and success rates for each of >100 different countries (almost all of which are over 70%, with China and Korea being significant exceptions among the major research-producing nations).**
+Our earlier paper has extensive discussion of the geographic origin issue, and the proportion of authors to whom we could reliably assign a gender for each of >100 different countries (almost all countries had a gendering success rate of 70-90%, with China and Korea being the exceptions among the major research-producing nations).**
 
@@
 the title says "Theoretical expectations").
-We investigated all combinations of values from 0% - 60% women, which spans >95% of journals on PubMed (see Holman et al.
+We have re-written this section to make it even more obvious that there is switch from discussing data to discussing a model.
+For illustrative purposes, we investigated all combinations of values from 0% - 60% women, which spans >95% of journals on PubMed (see Holman et al.
 2018).**
@@
 
-  **This is a stacked density plot, so there is no curve behind the other one (the blue is stacked over the white). We have clarified this in the figure legend.**
+  **This is a stacked density plot, so there is no curve behind the other one (the blue is stacked on top of the white). We have clarified this in the figure legend.**
 
@@
 This highlights a feature of $\alpha'$ noted earlier by the reviewer: it is a population-level measure, so we cannot make statements about individuals, e.g.
-we don't know if everyone is a little bit homophilous, or if there are a few prolific and very homophilous people (though the former is probably closest to the truth, since super-prolific authors with atypical levels of homophily are presumably the exception not the rule).**
+we don't know if everyone is a little bit homophilous, or if there are a few prolific and very homophilous people (though the former is probably closest to the truth, since hyper-prolific authors with atypical levels of homophily are presumably the exception not the rule).**
 

### 2018-11-01 manuscript/Manuscript_Claire.Rmd
Ready for resubmission

@@
 We also confirmed that in most journals (S2 Data) and most research disciplines (S3 Data, S1 Fig), the majority of papers had multiple authors.
-To verify the argument put forward in Figure 1, we also calculated $\alpha$ across all two-author papers in our entire 15-year PubMed dataset (excluding single-author papers, and papers where at least one author's gender was unknown; n = >3 million papers, >16 million authors).
-Across the whole dataset, $\alpha$ was 0.126, which is almost double the median value for individual journals (Figure 2; $\alpha = 0.070$ for 'All authors'), suggesting that lumping together papers from different fields and different time periods can indeed inflate $\alpha$.
 
 ![Of the `r n.journals` journals for which we had adequate data in 2015-2016, `r n.sig.homo` showed statistically significant evidence of gender homophily (denoted by $\alpha' > 0$), and `r n.sig.hetero` showed statistically significant evidence of heterophily ($\alpha' < 0$), after false discovery rate correction.
-In the stacked desnity plot, the white area shows the number of journals for which homophily was significantly stronger than expected under the null hypothesis (corrected p < 0.05), while the blue area shows all the remainder.
+In the stacked density plot, the white area shows the number of journals for which homophily was significantly stronger than expected under the null hypothesis (corrected p < 0.05), while the blue area shows all the remainder.
 Patterns were similar whether $\alpha'$ was calculated for all authors, for first authors only, or for last authors only.
@@
 
+Nevertheless, when we calculated $\alpha$ across all non-single-author papers in our entire 15-year PubMed dataset (as before, excluding papers where at least one author's gender was unknown; n = >3 million papers, >16 million authors), we found that $\alpha$ was 0.126.
+This figure is almost double the median value of $\alpha'$ for individual journals (Figure 2; $\alpha' = 0.070$ for 'All authors'), suggesting that lumping together papers from different fields and different time periods can indeed produce spurious evidence for gender homophily as outlined in Figure 1.
+
 ## Relationship between gender homophily and number of authors
 Papers with two authors had significantly lower (but still positive) $\alpha'$ values relative to papers with more than two authors, on average (\autoref{author_number}; statistical results in Online Supplmentary Material).
-Papers with 3, 4 or 5+ authors had essentially identical average $\alpha'$ values.
+Papers with 3, 4 or $\ge5$ authors had essentially identical average $\alpha'$ values.
 The variance in $\alpha'$ across journals was also a little higher for 2-authors papers compared to the remainder (\autoref{author_number}), though part of this variance is due to the reduced sample size (in terms of number of authors) for the 2-author papers.
@@
 ## Theoretical expectations for $\alpha$ when the gender ratio differs between career stages
-Given that we cannot identify individual researchers or their career stages, we used a simple model to derive the theoretical expectations for $\alpha$ when the gender ratio differs between career stages.
+Given that we cannot identify individual researchers or their career stages, we used a simple model to derive the theoretical expectations for $\alpha$ when the gender ratio differs between career stages (see Methods).
 As shown in \autoref{simulation}, we predict that $\alpha$ is expected to be non-zero, even if collaborators are randomly selected with respect to gender, provided that there is a gender gap between career stages.
@@
 
-![When there is a difference in gender ratio between early-career and established researchers, and collaboration is non-random with respect to career stage, the null expectation for $\alpha$ deviates from zero.
-An excess of collaborations between career stages gives the appearance of gender heterophily (lower rows, blue areas), while an excess of within-career stage collaborations produced apparent gender homophily (upper rows, red areas).
-However, the conditions required for strong gender homophily are quite restrictive, making it unlikely that this issue explains all of the homophily observed in Figure 2.
+![When the gender ratio of early-career researchers is not equal to the gender ratio among established researchers, the null expectation for $\alpha$ is not necessarily zero.
+Specifically, if most collaborations occur between career stages, there will be an excess of mixed-gender collaborations ($\alpha < 0$, blue areas), while if most collaborator pairs comprise two people at the same career stage, there will be an excess of same-gender collaborations ($\alpha > 0$, red areas).
+However, the conditions required for strong gender homophily (i.e.
+the red areas) are quite restrictive, making it unlikely that this issue can fully explain the homophily observed in our study.
+Additionally, in research disciplines where between-career stage collaboration is common and there is a shortage of women among established researchers (i.e.
+the blue areas), our study will underestimate the strength of gender homophily.
 Contour lines denote increments of 0.1.
@@
 Lastly, we note that if there is a gender gap between career stages and coauthorships between early-career and established researchers comprise >50% of the total, then the baseline expectation for $\alpha$ is actually less than zero (blue areas in \autoref{simulation}).
-Therefore, it is possible that researchers preferentially assort with same-gendered collaborators even more strongly than implied by other results, at least for certain journals or research disciplines.
+Therefore, it is possible that researchers preferentially assort with same-gendered collaborators even more strongly than implied by our results, at least for certain journals or research disciplines.
 
@@
 On a related note, Ghiasi et al.
-[-@Ghiasi_2015] argue that women in engineering are "compliant [in reproducing] male-dominated scientific structures" because they do not collaborate often enough with other women (for reference, their Figure 7 implies that coauthorships involving two women are 30% more frequent than expected under random assortment).
+[-@Ghiasi_2015] argue that women in engineering are "compliant [in reproducing] male-dominated scientific structures" because they do not collaborate often enough with other women (for reference, Figure 7 in [-@Ghiasi_2015] implies that coauthorships involving two women are $c$.
+30% more frequent than expected under random assortment).
 By contrast, we feel that it may be counter-productive to recommend that women collaborate primarily with other women, e.g.
@@
 Our study begs two questions: what causes gender homophily in science, and are our results cause for concern?
-These questions are closely related.
+We beleive that the answers to these questions are closely related.
 For example, some of the homophily we observed might be caused by women seeking to avoid harassment or sexism from men [@jagsi_2016], which would clearly be concerning.
@@
 ## Theoretical expectations for $\alpha$ when the gender ratio differs between career stages
-In most STEMM subjects, the gender ratio is more skewed among established researchers relative to early-career researchers, due both to women leaving STEMM careers at greater rates (the 'leaky pipeline'), and to historical shortages of women studying STEMM ('demographic inertia') [@Shaw_2012; @Holman_2018].
+In most STEMM disciplines, the gender ratio is more skewed among established researchers relative to early-career researchers, due both to women leaving STEMM careers at greater rates (the 'leaky pipeline'), and to historical shortages of women studying STEMM subjects at university ('demographic inertia') [@Shaw_2012; @Holman_2018].
 We hypothesised that this difference in gender ratio between career stages could potentially create both Wahlund effects and 'reverse' Wahlund effects.

### 2018-11-01 manuscript/Manuscript_Claire.tex
Ready for resubmission

@@
 gender homophily in science. \vspace{5mm}
-
-\par
-
-\noindent \textbf{Keywords:} Gender bias, Homophily, Research
+\par\noindent \textbf{Keywords:} Gender bias, Homophily, Research
 collaboration, Text mining, Women in STEM.
@@
 
-\maketitle\unmarkedfntext{*School of BioSciences, The University of Melbourne, Victoria, Australia.}
+\maketitle
+
+\unmarkedfntext{*School of BioSciences, The University of Melbourne, Victoria, Australia.}
 \maketitle\unmarkedfntext{\textsuperscript{§}Organismal and Evolutionary Biology Research Programme,
@@
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+\section{Introduction}\label{introduction}
 
@@
 
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-\hypertarget{gender-homophily-by-discipline-time-period-and-authorship-position}{%
+\section{Results}\label{results}
+
 \subsection{Gender homophily by discipline, time period, and authorship
-position}\label{gender-homophily-by-discipline-time-period-and-authorship-position}}
+position}\label{gender-homophily-by-discipline-time-period-and-authorship-position}
 
@@
 disciplines (S3 Data, S1 Fig), the majority of papers had multiple
-authors. To verify the argument put forward in Figure 1, we also
-calculated \(\alpha\) across all two-author papers in our entire 15-year
-PubMed dataset (excluding single-author papers, and papers where at
-least one author's gender was unknown; n = \textgreater{}3 million
-papers, \textgreater{}16 million authors). Across the whole dataset,
-\(\alpha\) was 0.126, which is almost double the median value for
-individual journals (Figure 2; \(\alpha = 0.070\) for `All authors'),
-suggesting that lumping together papers from different fields and
-different time periods can indeed inflate \(\alpha\).
+authors.
 
@@
 significant evidence of heterophily (\(\alpha' < 0\)), after false
-discovery rate correction. In the stacked desnity plot, the white area
+discovery rate correction. In the stacked density plot, the white area
 shows the number of journals for which homophily was significantly
@@
 
-\hypertarget{variance-in-homophily-between-disciplines}{%
 \subsection{Variance in homophily between
-disciplines}\label{variance-in-homophily-between-disciplines}}
+disciplines}\label{variance-in-homophily-between-disciplines}
 
@@
 
-\hypertarget{relationship-between-gender-homophily-and-number-of-authors}{%
+Nevertheless, when we calculated \(\alpha\) across all non-single-author
+papers in our entire 15-year PubMed dataset (as before, excluding papers
+where at least one author's gender was unknown; n = \textgreater{}3
+million papers, \textgreater{}16 million authors), we found that
+\(\alpha\) was 0.126. This figure is almost double the median value of
+\(\alpha'\) for individual journals (Figure 2; \(\alpha' = 0.070\) for
+`All authors'), suggesting that lumping together papers from different
+fields and different time periods can indeed produce spurious evidence
+for gender homophily as outlined in Figure 1.
+
 \subsection{Relationship between gender homophily and number of
-authors}\label{relationship-between-gender-homophily-and-number-of-authors}}
+authors}\label{relationship-between-gender-homophily-and-number-of-authors}
 
@@
 average (\autoref{author_number}; statistical results in Online
-Supplmentary Material). Papers with 3, 4 or 5+ authors had essentially
-identical average \(\alpha'\) values. The variance in \(\alpha'\) across
-journals was also a little higher for 2-authors papers compared to the
-remainder (\autoref{author_number}), though part of this variance is due
-to the reduced sample size (in terms of number of authors) for the
-2-author papers. One possible explanation for this finding is that
-2-authors papers might be more likely to have an author list that is
-evenly split between career stages (e.g.~a postgraduate student and
-their supervisor), increasing the chance that the authors are mixed
-gender (see Figure 6). The result also suggests that the processes
-responsible for gender homophily are similar in small (e.g.~3 author)
-and larger (5+ authors) collaborations (and across disciplines where
-small versus large collaborations are the norm).
+Supplmentary Material). Papers with 3, 4 or \(\ge5\) authors had
+essentially identical average \(\alpha'\) values. The variance in
+\(\alpha'\) across journals was also a little higher for 2-authors
+papers compared to the remainder (\autoref{author_number}), though part
+of this variance is due to the reduced sample size (in terms of number
+of authors) for the 2-author papers. One possible explanation for this
+finding is that 2-authors papers might be more likely to have an author
+list that is evenly split between career stages (e.g.~a postgraduate
+student and their supervisor), increasing the chance that the authors
+are mixed gender (see Figure 6). The result also suggests that the
+processes responsible for gender homophily are similar in small (e.g.~3
+author) and larger (5+ authors) collaborations (and across disciplines
+where small versus large collaborations are the norm).
 
@@
 
-\hypertarget{relationship-between-gender-homophily-and-gender-ratio}{%
 \subsection{Relationship between gender homophily and gender
-ratio}\label{relationship-between-gender-homophily-and-gender-ratio}}
+ratio}\label{relationship-between-gender-homophily-and-gender-ratio}
 
@@
 
-\hypertarget{relationship-between-journal-impact-factor-and-gender-homophily}{%
 \subsection{Relationship between journal impact factor and gender
-homophily}\label{relationship-between-journal-impact-factor-and-gender-homophily}}
+homophily}\label{relationship-between-journal-impact-factor-and-gender-homophily}
 
@@
 
-\hypertarget{analysis-accounting-for-differences-in-author-gender-ratio-between-countries}{%
 \subsection{Analysis accounting for differences in author gender ratio
 between
-countries}\label{analysis-accounting-for-differences-in-author-gender-ratio-between-countries}}
+countries}\label{analysis-accounting-for-differences-in-author-gender-ratio-between-countries}
 
@@
 
-\hypertarget{theoretical-expectations-for-alpha-when-the-gender-ratio-differs-between-career-stages}{%
 \subsection{\texorpdfstring{Theoretical expectations for \(\alpha\) when
 the gender ratio differs between career
-stages}{Theoretical expectations for \textbackslash{}alpha when the gender ratio differs between career stages}}\label{theoretical-expectations-for-alpha-when-the-gender-ratio-differs-between-career-stages}}
+stages}{Theoretical expectations for \textbackslash{}alpha when the gender ratio differs between career stages}}\label{theoretical-expectations-for-alpha-when-the-gender-ratio-differs-between-career-stages}
 
@@
 stages, we used a simple model to derive the theoretical expectations
-for \(\alpha\) when the gender ratio differs between career stages. As
-shown in \autoref{simulation}, we predict that \(\alpha\) is expected to
-be non-zero, even if collaborators are randomly selected with respect to
-gender, provided that there is a gender gap between career stages. The
-extent to which \(\alpha\) deviates from zero depends on the relative
-frequencies of collaboration within and between career stages (rows and
-columns in Figure 6), and the size of the gender gap between stages (x-
-and y-axes in Figure 6). When \textgreater{}50\% of coauthor pairs
-comprise one early-career and one established researcher, we expect
-gender heterophily (\(\alpha < 0\)) whenever the gender ratio differs
-between career stages. Conversely, when \textgreater{}50\% of
+for \(\alpha\) when the gender ratio differs between career stages (see
+Methods). As shown in \autoref{simulation}, we predict that \(\alpha\)
+is expected to be non-zero, even if collaborators are randomly selected
+with respect to gender, provided that there is a gender gap between
+career stages. The extent to which \(\alpha\) deviates from zero depends
+on the relative frequencies of collaboration within and between career
+stages (rows and columns in Figure 6), and the size of the gender gap
+between stages (x- and y-axes in Figure 6). When \textgreater{}50\% of
+coauthor pairs comprise one early-career and one established researcher,
+we expect gender heterophily (\(\alpha < 0\)) whenever the gender ratio
+differs between career stages. Conversely, when \textgreater{}50\% of
 collaborations are between people at the same career stage, we expect
@@
 \includegraphics{../figures/Fig6_inkscape.pdf}
-\caption{When there is a difference in gender ratio between early-career
-and established researchers, and collaboration is non-random with
-respect to career stage, the null expectation for \(\alpha\) deviates
-from zero. An excess of collaborations between career stages gives the
-appearance of gender heterophily (lower rows, blue areas), while an
-excess of within-career stage collaborations produced apparent gender
-homophily (upper rows, red areas). However, the conditions required for
-strong gender homophily are quite restrictive, making it unlikely that
-this issue explains all of the homophily observed in Figure 2. Contour
-lines denote increments of 0.1. \label{simulation}}
+\caption{When the gender ratio of early-career researchers is not equal
+to the gender ratio among established researchers, the null expectation
+for \(\alpha\) is not necessarily zero. Specifically, if most
+collaborations occur between career stages, there will be an excess of
+mixed-gender collaborations (\(\alpha < 0\), blue areas), while if most
+collaborator pairs comprise two people at the same career stage, there
+will be an excess of same-gender collaborations (\(\alpha > 0\), red
+areas). However, the conditions required for strong gender homophily
+(i.e.~the red areas) are quite restrictive, making it unlikely that this
+issue can fully explain the homophily observed in our study.
+Additionally, in research disciplines where between-career stage
+collaboration is common and there is a shortage of women among
+established researchers (i.e.~the blue areas), our study will
+underestimate the strength of gender homophily. Contour lines denote
+increments of 0.1. \label{simulation}}
 \end{figure}
@@
 preferentially assort with same-gendered collaborators even more
-strongly than implied by other results, at least for certain journals or
+strongly than implied by our results, at least for certain journals or
 research disciplines.
 
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
+\section{Discussion}\label{discussion}
 
@@
 scientific structures'' because they do not collaborate often enough
-with other women (for reference, their Figure 7 implies that
-coauthorships involving two women are 30\% more frequent than expected
-under random assortment). By contrast, we feel that it may be
+with other women (for reference, Figure 7 in {[}51{]} implies that
+coauthorships involving two women are \(c\). 30\% more frequent than
+expected under random assortment). By contrast, we feel that it may be
 counter-productive to recommend that women collaborate primarily with
@@
 Our study begs two questions: what causes gender homophily in science,
-and are our results cause for concern? These questions are closely
-related. For example, some of the homophily we observed might be caused
-by women seeking to avoid harassment or sexism from men {[}38{]}, which
-would clearly be concerning. Additionally, Sheltzer and Smith {[}66{]}
-concluded that `elite' male academics (defined as recipients of major
-honours) have a higher proportion of male students and postdocs than
-non-elite male academics. This finding could contribute to the homophily
-we observed, and is cause for concern since the results might reflect
-discrimination against women during hiring {[}20{]}, or avoidance by
-women of elite research groups (e.g.~due to gender differences in
-confidence, or a perception that some groups are sexist). We also found
-a little evidence that gender homophily is detrimental to research
-quality, in that high-impact journals tended to have weaker homophily.
-Assuming that papers published in high-impact journals are of higher
-average quality (which is contentious; {[}67{]}), our results provide
-non-experimental support for the hypothesis that mixed-gender teams
-produce better research than single-gender teams {[}42--48{]}. Another
-issue is that if many collaborations are between established
-researchers, there will be an excess of male-male collaborations in
-fields where women in senior positions are rare; some of the observed
-homophily might therefore reflect the elevated gender gap among senior
-researchers.
+and are our results cause for concern? We beleive that the answers to
+these questions are closely related. For example, some of the homophily
+we observed might be caused by women seeking to avoid harassment or
+sexism from men {[}38{]}, which would clearly be concerning.
+Additionally, Sheltzer and Smith {[}66{]} concluded that `elite' male
+academics (defined as recipients of major honours) have a higher
+proportion of male students and postdocs than non-elite male academics.
+This finding could contribute to the homophily we observed, and is cause
+for concern since the results might reflect discrimination against women
+during hiring {[}20{]}, or avoidance by women of elite research groups
+(e.g.~due to gender differences in confidence, or a perception that some
+groups are sexist). We also found a little evidence that gender
+homophily is detrimental to research quality, in that high-impact
+journals tended to have weaker homophily. Assuming that papers published
+in high-impact journals are of higher average quality (which is
+contentious; {[}67{]}), our results provide non-experimental support for
+the hypothesis that mixed-gender teams produce better research than
+single-gender teams {[}42--48{]}. Another issue is that if many
+collaborations are between established researchers, there will be an
+excess of male-male collaborations in fields where women in senior
+positions are rare; some of the observed homophily might therefore
+reflect the elevated gender gap among senior researchers.
 
@@
 
-\hypertarget{methods}{%
-\section{Methods}\label{methods}}
-
-\hypertarget{the-dataset}{%
-\subsection{The dataset}\label{the-dataset}}
+\section{Methods}\label{methods}
+
+\subsection{The dataset}\label{the-dataset}
 
@@
 
-\hypertarget{calculating-alpha-the-coefficient-of-homophily}{%
 \subsection{\texorpdfstring{Calculating \(\alpha\), the coefficient of
-homophily}{Calculating \textbackslash{}alpha, the coefficient of homophily}}\label{calculating-alpha-the-coefficient-of-homophily}}
+homophily}{Calculating \textbackslash{}alpha, the coefficient of homophily}}\label{calculating-alpha-the-coefficient-of-homophily}
 
@@
 
-\hypertarget{minimising-the-wahlund-effect-research-discipline-and-time-period}{%
 \subsection{Minimising the Wahlund effect: research discipline and time
-period}\label{minimising-the-wahlund-effect-research-discipline-and-time-period}}
+period}\label{minimising-the-wahlund-effect-research-discipline-and-time-period}
 
@@
 
-\hypertarget{minimising-the-wahlund-effect-author-country-of-affiliation}{%
 \subsection{Minimising the Wahlund effect: author country of
-affiliation}\label{minimising-the-wahlund-effect-author-country-of-affiliation}}
+affiliation}\label{minimising-the-wahlund-effect-author-country-of-affiliation}
 
@@
 
-\hypertarget{calculating-standardised-journal-impact-factor}{%
 \subsection{Calculating standardised journal impact
-factor}\label{calculating-standardised-journal-impact-factor}}
+factor}\label{calculating-standardised-journal-impact-factor}
 
@@
 
-\hypertarget{statistical-analysis}{%
-\subsection{Statistical analysis}\label{statistical-analysis}}
+\subsection{Statistical analysis}\label{statistical-analysis}
 
@@
 
-\hypertarget{theoretical-expectations-for-alpha-when-the-gender-ratio-differs-between-career-stages-1}{%
 \subsection{\texorpdfstring{Theoretical expectations for \(\alpha\) when
 the gender ratio differs between career
-stages}{Theoretical expectations for \textbackslash{}alpha when the gender ratio differs between career stages}}\label{theoretical-expectations-for-alpha-when-the-gender-ratio-differs-between-career-stages-1}}
-
-In most STEMM subjects, the gender ratio is more skewed among
+stages}{Theoretical expectations for \textbackslash{}alpha when the gender ratio differs between career stages}}\label{theoretical-expectations-for-alpha-when-the-gender-ratio-differs-between-career-stages-1}
+
+In most STEMM disciplines, the gender ratio is more skewed among
 established researchers relative to early-career researchers, due both
 to women leaving STEMM careers at greater rates (the `leaky pipeline'),
-and to historical shortages of women studying STEMM (`demographic
-inertia') {[}1,5{]}. We hypothesised that this difference in gender
-ratio between career stages could potentially create both Wahlund
-effects and `reverse' Wahlund effects. For example, imagine that the
-majority of collaborations in a particular field are between students
-and professors, and that the gender ratio differs between career stages:
-we would then see an excess of mixed-gender coauthorships (heterophily,
-\(\alpha < 0\)), even if gender has no direct, causal effect. Similarly,
-a hypothetical field in which students work only with students, and
-professors with professors, would have apparent gender homophily
-(\(\alpha > 0\)).
+and to historical shortages of women studying STEMM subjects at
+university (`demographic inertia') {[}1,5{]}. We hypothesised that this
+difference in gender ratio between career stages could potentially
+create both Wahlund effects and `reverse' Wahlund effects. For example,
+imagine that the majority of collaborations in a particular field are
+between students and professors, and that the gender ratio differs
+between career stages: we would then see an excess of mixed-gender
+coauthorships (heterophily, \(\alpha < 0\)), even if gender has no
+direct, causal effect. Similarly, a hypothetical field in which students
+work only with students, and professors with professors, would have
+apparent gender homophily (\(\alpha > 0\)).
 
@@
 
-\hypertarget{data-availability-and-reproducibility}{%
 \subsection{Data availability and
-reproducibility}\label{data-availability-and-reproducibility}}
+reproducibility}\label{data-availability-and-reproducibility}
 
@@
 
-\hypertarget{acknowledgements}{%
-\section{Acknowledgements}\label{acknowledgements}}
+\section{Acknowledgements}\label{acknowledgements}
 
@@
 
-\hypertarget{references}{%
-\section{References}\label{references}}
+\section{References}\label{references}
 
 \hypertarget{refs}{}
-\leavevmode\hypertarget{ref-Shaw_2012}{}%
+\hypertarget{ref-Shaw_2012}{}
 1. Shaw AK, Stanton DE. Leaks in the pipeline: separating demographic
@@
 
-\leavevmode\hypertarget{ref-Lariviere_2013}{}%
+\hypertarget{ref-Lariviere_2013}{}
 2. Larivière V, Ni C, Gingras Y, Cronin B, Sugimoto CR. Bibliometrics:
@@
 
-\leavevmode\hypertarget{ref-West_2013}{}%
+\hypertarget{ref-West_2013}{}
 3. West JD, Jacquet J, King MM, Correll SJ, Bergstrom CT. The role of
@@
 
-\leavevmode\hypertarget{ref-Elsevier_report}{}%
+\hypertarget{ref-Elsevier_report}{}
 4. Elsevier Report. Gender in the global research landscape.
@@
 
-\leavevmode\hypertarget{ref-Holman_2018}{}%
+\hypertarget{ref-Holman_2018}{}
 5. Holman L, Stuart Fox D, Hauser CE. The gender gap in science: How
@@
 
-\leavevmode\hypertarget{ref-Wutte_2007}{}%
+\hypertarget{ref-Wutte_2007}{}
 6. Wutte M. Closing the gender gap. Nature. 2007;448: NJ101--NJ102.
 
-\leavevmode\hypertarget{ref-Reuben_2014}{}%
+\hypertarget{ref-Reuben_2014}{}
 7. Reuben E, Sapienza P, Zingales L. How stereotypes impair women's
@@
 
-\leavevmode\hypertarget{ref-Trower_2002}{}%
+\hypertarget{ref-Trower_2002}{}
 8. Trower CA, Chait RP. Faculty diversity: Why women and minorities are
@@
 
-\leavevmode\hypertarget{ref-Umbach_2007}{}%
+\hypertarget{ref-Umbach_2007}{}
 9. Umbach PD. Gender equity in the academic labor market: An analysis of
@@
 
-\leavevmode\hypertarget{ref-Hosek_2005}{}%
+\hypertarget{ref-Hosek_2005}{}
 10. Hosek S, Cox AG, Ghosh-Dastidar B, Kofner A, Ramphal N, Scott J, et
@@
 
-\leavevmode\hypertarget{ref-pohlhaus_2011}{}%
+\hypertarget{ref-pohlhaus_2011}{}
 11. Pohlhaus JR, Jiang H, Wagner RM, Schaffer WT, Pinn VW. Sex
@@
 
-\leavevmode\hypertarget{ref-Zuckerman_1987}{}%
+\hypertarget{ref-Zuckerman_1987}{}
 12. Zuckerman H. Persistence and change in the careers of men and women
@@
 
-\leavevmode\hypertarget{ref-Rosenfeld_1991}{}%
+\hypertarget{ref-Rosenfeld_1991}{}
 13. Rosenfeld RA. Outcome analysis of academic careers. Review prepared
@@
 
-\leavevmode\hypertarget{ref-Long_1993}{}%
+\hypertarget{ref-Long_1993}{}
 14. Long JS, Paul DA, Robert M. Rank advancement in academic careers:
@@
 
-\leavevmode\hypertarget{ref-Hopkins_2013}{}%
+\hypertarget{ref-Hopkins_2013}{}
 15. Hopkins AL, Jawitz JW, McCarty C, Goldman A, Basu NB. Disparities in
@@
 
-\leavevmode\hypertarget{ref-ODorchai_2009}{}%
+\hypertarget{ref-ODorchai_2009}{}
 16. O'Dorchai S, Meulders D, Crippa F, Margherita A. She figures
@@
 
-\leavevmode\hypertarget{ref-Feldt_1986}{}%
+\hypertarget{ref-Feldt_1986}{}
 17. Feldt B. The faculty cohort study: School of medicine. Ann Arbor,
@@
 
-\leavevmode\hypertarget{ref-Stack_2004}{}%
+\hypertarget{ref-Stack_2004}{}
 18. Stack S. Gender, children and research productivity. Scientometrics.
@@
 
-\leavevmode\hypertarget{ref-Lariviere_2011}{}%
+\hypertarget{ref-Lariviere_2011}{}
 19. Larivière V, Vignola-Gagné E, Villeneuve C, Gélinas P, Gingras Y.
@@
 
-\leavevmode\hypertarget{ref-Moss_2012}{}%
+\hypertarget{ref-Moss_2012}{}
 20. Moss-Racusin CA, Dovidio JF, Brescoll VL, Graham MJ, Handelsman J.
@@
 
-\leavevmode\hypertarget{ref-Knobloch_2013}{}%
+\hypertarget{ref-Knobloch_2013}{}
 21. Knobloch-Westerwick S, Glynn CJ, Huge M. Science faculty's subtle
@@
 
-\leavevmode\hypertarget{ref-Lee_2005}{}%
+\hypertarget{ref-Lee_2005}{}
 22. Lee S, Bozeman B. The impact of research collaboration on scientific
@@
 
-\leavevmode\hypertarget{ref-Wuchty_2007}{}%
+\hypertarget{ref-Wuchty_2007}{}
 23. Wuchty S, Jones BF, Uzzi B. The increasing dominance of teams in
@@
 
-\leavevmode\hypertarget{ref-Abramo_2009}{}%
+\hypertarget{ref-Abramo_2009}{}
 24. Abramo G, D'Angelo CA, Di Costa F. Research collaboration and
@@
 
-\leavevmode\hypertarget{ref-Lariviere_2015}{}%
+\hypertarget{ref-Lariviere_2015}{}
 25. Larivière V, Gingras Y, Sugimoto CR, Tsou A. Team size matters:
@@
 
-\leavevmode\hypertarget{ref-Long_1992}{}%
+\hypertarget{ref-Long_1992}{}
 26. Long JS. Measures of sex differences in scientific productivity.
@@
 
-\leavevmode\hypertarget{ref-Bozeman_2011}{}%
+\hypertarget{ref-Bozeman_2011}{}
 27. Bozeman B, Monica G. How do men and women differ in research
@@
 
-\leavevmode\hypertarget{ref-Abramo_2013}{}%
+\hypertarget{ref-Abramo_2013}{}
 28. Abramo G, D'Angelo CA, Di Costa F. Gender differences in research
@@
 
-\leavevmode\hypertarget{ref-Badar_2013}{}%
+\hypertarget{ref-Badar_2013}{}
 29. Badar K, Hite JM, Badir YF. Examining the relationship of
@@
 
-\leavevmode\hypertarget{ref-Lewison_2001}{}%
+\hypertarget{ref-Lewison_2001}{}
 30. Lewison G. The quantity and quality of female researchers: A
@@
 
-\leavevmode\hypertarget{ref-Webster_2001}{}%
+\hypertarget{ref-Webster_2001}{}
 31. Webster BM. Polish women in science: A bibliometric analysis of
@@
 
-\leavevmode\hypertarget{ref-Bozeman_2004}{}%
+\hypertarget{ref-Bozeman_2004}{}
 32. Bozeman B, Corley E. Scientists' collaboration strategies:
@@
 
-\leavevmode\hypertarget{ref-Long_1990}{}%
+\hypertarget{ref-Long_1990}{}
 33. Long JS. The origins of sex differences in science. Social Forces.
@@
 
-\leavevmode\hypertarget{ref-Fuchs_2001}{}%
+\hypertarget{ref-Fuchs_2001}{}
 34. Fuchs S, Von Stebut J, Allmendinger J. Gender, science, and
@@
 
-\leavevmode\hypertarget{ref-Reskin_1978}{}%
+\hypertarget{ref-Reskin_1978}{}
 35. Reskin BF. Scientific productivity, sex, and location in the
@@
 
-\leavevmode\hypertarget{ref-Wright_2003}{}%
+\hypertarget{ref-Wright_2003}{}
 36. Wright AL, Schwindt LA, Bassford TL, Reyna VF, Shisslak PAS
@@
 
-\leavevmode\hypertarget{ref-bleidorn2016age}{}%
+\hypertarget{ref-bleidorn2016age}{}
 37. Bleidorn W, Arslan RC, Denissen JJ, Rentfrow PJ, Gebauer JE, Potter
@@
 
-\leavevmode\hypertarget{ref-jagsi_2016}{}%
+\hypertarget{ref-jagsi_2016}{}
 38. Jagsi R, Griffith KA, Jones R, Perumalswami CR, Ubel P, Stewart A.
@@
 
-\leavevmode\hypertarget{ref-Martin_2014}{}%
+\hypertarget{ref-Martin_2014}{}
 39. Martin JL. Ten simple rules to achieve conference speaker gender
@@
 
-\leavevmode\hypertarget{ref-Tower_2007}{}%
+\hypertarget{ref-Tower_2007}{}
 40. Tower G, Julie P, Brenda R. A multidisciplinary study of
@@
 
-\leavevmode\hypertarget{ref-Jordan_2008}{}%
+\hypertarget{ref-Jordan_2008}{}
 41. Jordan CE, Clark SJ, Vann CE. Do gender differences exist in the
@@
 
-\leavevmode\hypertarget{ref-Britton_2000}{}%
+\hypertarget{ref-Britton_2000}{}
 42. Britton DM. The epistemology of the gendered organization. Gender
@@
 
-\leavevmode\hypertarget{ref-Reagans_2001}{}%
+\hypertarget{ref-Reagans_2001}{}
 43. Reagans R, Zuckerman EW. Networks, diversity, and productivity: The
@@
 
-\leavevmode\hypertarget{ref-Hong_2004}{}%
+\hypertarget{ref-Hong_2004}{}
 44. Hong L, Page SE. Groups of diverse problem solvers can outperform
@@
 
-\leavevmode\hypertarget{ref-Whittington_2008}{}%
+\hypertarget{ref-Whittington_2008}{}
 45. Whittington KB, Smith-Doerr L. Women inventors in context:
@@
 
-\leavevmode\hypertarget{ref-Bear_2011}{}%
+\hypertarget{ref-Bear_2011}{}
 46. Bear JB, Woolley AW. The role of gender in team collaboration and
@@
 
-\leavevmode\hypertarget{ref-Herrera_2012}{}%
+\hypertarget{ref-Herrera_2012}{}
 47. Herrera R, Duncan PA, Green MT, Skaggs SL. The effect of gender on
@@
 
-\leavevmode\hypertarget{ref-Campbell_2013}{}%
+\hypertarget{ref-Campbell_2013}{}
 48. Campbell LG, Mehtani S, Dozier ME, Rinehart J. Gender-heterogeneous
@@
 
-\leavevmode\hypertarget{ref-Ferber_1980}{}%
+\hypertarget{ref-Ferber_1980}{}
 49. Ferber MA, Teiman M. Are women economists at a disadvantage in
@@
 
-\leavevmode\hypertarget{ref-McDowell_1992}{}%
+\hypertarget{ref-McDowell_1992}{}
 50. McDowell JM, Smith JK. The effect of gender-sorting on propensity to
@@
 
-\leavevmode\hypertarget{ref-Ghiasi_2015}{}%
+\hypertarget{ref-Ghiasi_2015}{}
 51. Ghiasi G, Larivière V, Sugimoto CR. On the compliance of women
@@
 
-\leavevmode\hypertarget{ref-Crow_2015}{}%
+\hypertarget{ref-Crow_2015}{}
 52. Crow MS, Smykla JO. An examination of author characteristics in
@@
 
-\leavevmode\hypertarget{ref-Fahmy_2017}{}%
+\hypertarget{ref-Fahmy_2017}{}
 53. Fahmy C, Young JT. Gender inequality and knowledge production in
@@
 
-\leavevmode\hypertarget{ref-Zettler_2017}{}%
+\hypertarget{ref-Zettler_2017}{}
 54. Zettler HR, Stephanie M Cardwell, Jessica M C. The gendering effects
@@
 
-\leavevmode\hypertarget{ref-Jadidi_2017}{}%
+\hypertarget{ref-Jadidi_2017}{}
 55. Jadidi M, Karimi F, Lietz H, Wagner C. Gender disparities in
@@
 
-\leavevmode\hypertarget{ref-Teele_2017}{}%
+\hypertarget{ref-Teele_2017}{}
 56. Teele DL, Kathleen T. Gender in the journals: Publication patterns
@@
 
-\leavevmode\hypertarget{ref-Araujo_2017a}{}%
+\hypertarget{ref-Araujo_2017a}{}
 57. Araújo T, Elsa F. The specific shapes of gender imbalance in
@@
 
-\leavevmode\hypertarget{ref-Araujo_2017b}{}%
+\hypertarget{ref-Araujo_2017b}{}
 58. Araújo T, Elsa F. Big Missing Data: are scientific memes inherited
@@
 
-\leavevmode\hypertarget{ref-Wahlund_1928}{}%
+\hypertarget{ref-Wahlund_1928}{}
 59. Wahlund S. Zusammensetzung von populationen und
@@
 
-\leavevmode\hypertarget{ref-bergstrom_2016}{}%
+\hypertarget{ref-bergstrom_2016}{}
 60. Bergstrom T, Bergstrom C, King M, Jacquet J, West J, Correll S. A
-note on measuring gender homophily among scholarly authors. 2016;
-
-\leavevmode\hypertarget{ref-macaluso_2016}{}%
+note on measuring gender homophily among scholarly authors.
+http://eigenfactororg/gender/ assortativity/measuring\_homophilypdf.
+2016;
+
+\hypertarget{ref-macaluso_2016}{}
 61. Macaluso B, Larivière V, Sugimoto T, Sugimoto CR. Is science built
@@
 
-\leavevmode\hypertarget{ref-Bentley_2003}{}%
+\hypertarget{ref-Bentley_2003}{}
 62. Bentley JT, Adamson R. Gender differences in the careers of academic
@@
 
-\leavevmode\hypertarget{ref-Long_2015}{}%
+\hypertarget{ref-Long_2015}{}
 63. Long MT, Leszczynski A, Thompson KD, Wasan SK, Calderwood AH. Female
@@
 
-\leavevmode\hypertarget{ref-Bendels_2018}{}%
+\hypertarget{ref-Bendels_2018}{}
 64. Bendels MH, Bauer J, Schöffel N, Groneberg DA. The gender gap in
@@
 
-\leavevmode\hypertarget{ref-McKenzie_2017}{}%
+\hypertarget{ref-McKenzie_2017}{}
 65. McKenzie K, Ramonas M, Patlas M, Katz DS. Assessing the gap in
@@
 
-\leavevmode\hypertarget{ref-sheltzer_2014}{}%
+\hypertarget{ref-sheltzer_2014}{}
 66. Sheltzer JM, Smith JC. Elite male faculty in the life sciences
@@
 
-\leavevmode\hypertarget{ref-Garfield_2006}{}%
+\hypertarget{ref-Garfield_2006}{}
 67. Garfield E. The history and meaning of the journal impact factor.
@@
 
-\leavevmode\hypertarget{ref-nittrouer_2018}{}%
+\hypertarget{ref-nittrouer_2018}{}
 68. Nittrouer CL, Hebl MR, Ashburn-Nardo L, Trump-Steele RC, Lane DM,
@@
 
-\leavevmode\hypertarget{ref-debarre_2018}{}%
+\hypertarget{ref-debarre_2018}{}
 69. Débarre F, Rode N, Ugelvig L. Gender equity at scientific events.
@@
 
-\leavevmode\hypertarget{ref-wright_1949}{}%
+\hypertarget{ref-wright_1949}{}
 70. Wright S. The genetical structure of populations. Annals of Human
@@
 
-\leavevmode\hypertarget{ref-wang_2016}{}%
+\hypertarget{ref-wang_2016}{}
 71. Wang YS, Erosheva EA. On the relationship between set-based and
@@
 
-\leavevmode\hypertarget{ref-Benjamini_1995}{}%
+\hypertarget{ref-Benjamini_1995}{}
 72. Benjamini Y, Hochberg Y. Controlling the false discovery rate: a
@@
 
-\leavevmode\hypertarget{ref-Bonham_2017}{}%
+\hypertarget{ref-Bonham_2017}{}
 73. Bonham KS, Stefan MI. Women are underrepresented in computational
@@
 
-\leavevmode\hypertarget{ref-Wren_2007}{}%
+\hypertarget{ref-Wren_2007}{}
 74. Wren JD, Kozak KZ, Johnson KR, Deakyne SJ, Schilling LM, Dellavalle
@@
 
-\leavevmode\hypertarget{ref-burkner_2016}{}%
+\hypertarget{ref-burkner_2016}{}
 75. Bürkner P-C. brms: An R package for Bayesian multilevel models using
@@
 
-\hypertarget{supporting-information}{%
-\section{Supporting information}\label{supporting-information}}
-
-\hypertarget{supplementary-figures}{%
-\subsection{Supplementary figures}\label{supplementary-figures}}
+\section{Supporting information}\label{supporting-information}
+
+\subsection{Supplementary figures}\label{supplementary-figures}
 
@@
 
-\hypertarget{supplementary-tables}{%
-\subsection{Supplementary tables}\label{supplementary-tables}}
+\subsection{Supplementary tables}\label{supplementary-tables}
 
@@
 
-\hypertarget{supplementary-datasets}{%
-\subsection{Supplementary datasets}\label{supplementary-datasets}}
+\subsection{Supplementary datasets}\label{supplementary-datasets}
 

### 2018-11-01 manuscript/Manuscript_Claire.Rmd
Ready for resubmission - typos fixed

@@
 ## Relationship between gender homophily and number of authors
-Papers with two authors had significantly lower (but still positive) $\alpha'$ values relative to papers with more than two authors, on average (\autoref{author_number}; statistical results in Online Supplmentary Material).
+Papers with two authors had significantly lower (but still positive) $\alpha'$ values relative to papers with more than two authors, on average (\autoref{author_number}; statistical results in Online Supplementary Material).
 Papers with 3, 4 or $\ge5$ authors had essentially identical average $\alpha'$ values.
@@
 An important limitation of our study is that we cannot reliably determine the cause(s) of the observed excess of same-gender coauthorships.
-As well as the obvious interpretation -- conscious or unconcious selection of same-gender collaborators by men, by women, or both -- our results could be partly explained by uncontrolled Wahlund effects.
+As well as the obvious interpretation -- conscious or unconscious selection of same-gender collaborators by men, by women, or both -- our results could be partly explained by uncontrolled Wahlund effects.
 However, we suspect the contribution of these uncontrolled artefacts to be minor, for four reasons: we found positive $\alpha'$ after controlling for three obvious sources of Wahlund effect; there was no inflation of $\alpha'$ in highly multidisciplinary journals relative to specialised journals; restricting the data by country yielded similar estimates of $\alpha'$; and we used a simple model to show that differences in gender ratio between career stages are unlikely to fully explain our results.
@@
 Our study begs two questions: what causes gender homophily in science, and are our results cause for concern?
-We beleive that the answers to these questions are closely related.
+We believe that the answers to these questions are closely related.
 For example, some of the homophily we observed might be caused by women seeking to avoid harassment or sexism from men [@jagsi_2016], which would clearly be concerning.
@@
 We also discarded all papers for which we could not determine the gender of every author with ${\ge}95\%$ certainty, in order to simplify the statistical analysis.
-To mitigate Wahlund effects caused by variation in the gender ratio of researchers over time (see below), we only kept papers with publicaiton dates falling in two one-year time periods, namely 0-1 or 10-11 years prior to the collection date of the PubMed data (i.e.
+To mitigate Wahlund effects caused by variation in the gender ratio of researchers over time (see below), we only kept papers with publication dates falling in two one-year time periods, namely 0-1 or 10-11 years prior to the collection date of the PubMed data (i.e.
 20^th^ August 2016).
@@
 To model the relationship between $\alpha'$ and the number of authors on the paper, we used a meta-regression model implemented in the R package `brms` [@burkner_2016].
-The model incorporated the standard error associated with easch estimate of $\alpha'$, had author number as a fixed effect, and journal as a random intercept (to control for repeated measures of each journal).
+The model incorporated the standard error associated with each estimate of $\alpha'$, had author number as a fixed effect, and journal as a random intercept (to control for repeated measures of each journal).
 We also fit a random slope of author number within journal, thereby allowing the response to author number to vary between journals.
@@
 Under the null model that gender has no causal effect on collaboration, we calculated $\alpha$ for various combinations of the four free parameters in our simple model.
-These parameters are: the gender ratio among early-career researchers (x-axis of Figure 6), the gender ratio among established researchers (y-axis of Figure 6), the frequency of within- versus between career stage collaborator pairs (rows in Figure 6), and lastly the frequency of within-stage collaborations that are between two early-career researchers as opposed to two late-career reserachers (columns in Figure 6).
+These parameters are: the gender ratio among early-career researchers (x-axis of Figure 6), the gender ratio among established researchers (y-axis of Figure 6), the frequency of within- versus between career stage collaborator pairs (rows in Figure 6), and lastly the frequency of within-stage collaborations that are between two early-career researchers as opposed to two late-career researchers (columns in Figure 6).
 When these four parameters are specified, one can easily calculate the relative frequencies of collaborator pairs that involve two men, two women, or a man and a woman.
@@
 Histogram showing the estimated degree to which $\alpha'$ is inflated by inter-country differences in author gender ratio, across the 285 journals for which we had adequate data after restricting the analysis by country.
-The average inflation in $\alpha'$ is negligible, suggesting that Wahlund effects resulting from inter-country differences have a neglible effect on our estimates of gender homophily.
+The average inflation in $\alpha'$ is negligible, suggesting that Wahlund effects resulting from inter-country differences have a negligible effect on our estimates of gender homophily.
 
@@
 Number of journals showing statistically significant homophily or heterophily, in two one-year periods.
-The significance threshold was p < 0.05, and p-values were adjused using Benjamini-Hochberg false discovery rate correction.
+The significance threshold was p < 0.05, and p-values were adjusted using Benjamini-Hochberg false discovery rate correction.
 Note that the power of our test is lower for the 2005-2006 data because fewer papers were recovered per journal: thus, it is not meaningful to compare the % significant journals (i.e.

### 2018-11-01 manuscript/Manuscript_Claire.tex
Ready for resubmission - typos fixed

@@
 average (\autoref{author_number}; statistical results in Online
-Supplmentary Material). Papers with 3, 4 or \(\ge5\) authors had
+Supplementary Material). Papers with 3, 4 or \(\ge5\) authors had
 essentially identical average \(\alpha'\) values. The variance in
@@
 coauthorships. As well as the obvious interpretation -- conscious or
-unconcious selection of same-gender collaborators by men, by women, or
+unconscious selection of same-gender collaborators by men, by women, or
 both -- our results could be partly explained by uncontrolled Wahlund
@@
 Our study begs two questions: what causes gender homophily in science,
-and are our results cause for concern? We beleive that the answers to
+and are our results cause for concern? We believe that the answers to
 these questions are closely related. For example, some of the homophily
@@
 the gender ratio of researchers over time (see below), we only kept
-papers with publicaiton dates falling in two one-year time periods,
+papers with publication dates falling in two one-year time periods,
 namely 0-1 or 10-11 years prior to the collection date of the PubMed
@@
 package \texttt{brms} {[}75{]}. The model incorporated the standard
-error associated with easch estimate of \(\alpha'\), had author number
-as a fixed effect, and journal as a random intercept (to control for
+error associated with each estimate of \(\alpha'\), had author number as
+a fixed effect, and journal as a random intercept (to control for
 repeated measures of each journal). We also fit a random slope of author
@@
 between two early-career researchers as opposed to two late-career
-reserachers (columns in Figure 6). When these four parameters are
+researchers (columns in Figure 6). When these four parameters are
 specified, one can easily calculate the relative frequencies of
@@
 suggesting that Wahlund effects resulting from inter-country differences
-have a neglible effect on our estimates of gender homophily.
+have a negligible effect on our estimates of gender homophily.
 
@@
 or heterophily, in two one-year periods. The significance threshold was
-p \textless{} 0.05, and p-values were adjused using Benjamini-Hochberg
+p \textless{} 0.05, and p-values were adjusted using Benjamini-Hochberg
 false discovery rate correction. Note that the power of our test is

### 2019-02-28 manuscript/homophily_manuscript.Rmd
Ready for second resubmission

--- 2018-11-01 manuscript/Manuscript_Claire.Rmd
@@
 
-However, we believe that earlier studies of gender homophily were hindered by a largely unacknowledged statistical issue that we will refer to as the Wahlund effect (\autoref{wahlund_plot}), by analogy with the conceptually similar Whalund effect in population genetics [@Wahlund_1928].
+However, we believe that earlier studies of gender homophily were hindered by a largely unacknowledged statistical issue that we will refer to as the Wahlund effect (\autoref{wahlund_plot}), by analogy with the conceptually similar Wahlund effect in population genetics [@Wahlund_1928].
 The Wahlund effect makes it deceptively difficult to infer gender-based preferences simply by counting the number of same- and mixed-gender coauthorships.
@@
 Papers with 3, 4 or $\ge5$ authors had essentially identical average $\alpha'$ values.
-The variance in $\alpha'$ across journals was also a little higher for 2-authors papers compared to the remainder (\autoref{author_number}), though part of this variance is due to the reduced sample size (in terms of number of authors) for the 2-author papers.
+The variance in $\alpha'$ across journals was also a little higher for 2-author papers compared to the remainder (\autoref{author_number}), though part of this variance is due to the reduced sample size (in terms of number of authors) for the 2-author papers.
 One possible explanation for this finding is that 2-authors papers might be more likely to have an author list that is evenly split between career stages (e.g.
@@
 The result also suggests that the processes responsible for gender homophily are similar in small (e.g.
-3 author) and larger (5+ authors) collaborations (and across disciplines where small versus large collaborations are the norm).
+3-author) and larger (5+ author) collaborations (and across disciplines where small versus large collaborations are the norm).
 
@@
 In a small dataset comprising $m$ men and $f$ women, a man can co-author with $m - 1$ men while women can co-author with $m$ men.
-Thus, the null expectation for $alpha$ is a negative number -- potentially a large one if $m$ and $f$ are very small.
+Thus, the null expectation for $\alpha$ is a negative number -- potentially a large one if $m$ and $f$ are very small.
 
@@
 # Acknowledgements
-CM was supported by the Academy of Finland (284666 to the Centre of Excellence in Biological Interactions). We thank Devi Stuart-Fox and Dominique Potvin for helpful discussion.
+We thank Devi Stuart-Fox and Dominique Potvin for helpful discussion.
 

### 2019-03-04 manuscript/homophily_manuscript.Rmd

@@
 For example, studies have concluded that women tend to be less involved in international collaboration [@Lewison_2001; @Webster_2001; @Bozeman_2004; @Lariviere_2011; @Abramo_2013], collaborate less within their own university departments [@Webster_2001], have less prestigious collaborations [@Long_1990], and fewer collaborations in total [@Fuchs_2001].
-These gender differences in collaboration practice presumably have multiple causes, which might include implicit and explicit gender bias [@Moss_2012], differential family obligations [@Reskin_1978; @Long_1990; @Wright_2003], gender differences in confidence or self-esteem [@bleidorn2016age], concerns relating to sexual harassment [@jagsi_2016], and unequal access to conferences [@Martin_2014] or travel funds [@Bozeman_2004].
+These gender differences in collaboration presumably have multiple causes, which might include implicit and explicit gender bias [@Moss_2012], differential family obligations [@Reskin_1978; @Long_1990; @Wright_2003], gender differences in confidence or self-esteem [@bleidorn2016age], concerns relating to sexual harassment [@jagsi_2016], and unequal access to conferences [@Martin_2014] or travel funds [@Bozeman_2004].
 
@@
 To our knowledge, most or all such studies imply that men co-publish with men, and women with women, more often than expected if collaborators assort randomly with respect to gender [@Ferber_1980; @McDowell_1992; @Ghiasi_2015; @Crow_2015; @Fahmy_2017; @Zettler_2017; @Jadidi_2017; @Teele_2017; @Araujo_2017a; @Araujo_2017b].
-This pattern of assortative publishing has often been termed 'gender homophily'.
+This non-random assortment is often termed 'gender homophily'.
 
-However, we believe that earlier studies of gender homophily were hindered by a largely unacknowledged statistical issue that we will refer to as the Wahlund effect (\autoref{wahlund_plot}), by analogy with the conceptually similar Wahlund effect in population genetics [@Wahlund_1928].
+We believe that most or all earlier studies of gender homophily were hindered by a largely unacknowledged statistical issue that we will refer to as the Wahlund effect (\autoref{wahlund_plot}), by analogy with the conceptually similar Wahlund effect in population genetics [@Wahlund_1928].
 The Wahlund effect makes it deceptively difficult to infer gender-based preferences simply by counting the number of same- and mixed-gender coauthorships.
@@
 For example, the journal category 'Multidisciplinary' -- which includes journals like _PLOS ONE_, _Nature_, _Science_, and _PNAS_ -- did not have markedly elevated $\alpha'$ (\autoref{alpha_histograms}).
-This result suggests that our estimates of homophily, and estimates from some of the earlier studies listed in the Introduction, are probably not inflated by the presence of disparate research topics (with variable author gender ratios) being published within individual journals.
+This result suggests that our estimates of homophily, and estimates from some of the earlier studies listed in the Introduction, are probably not markedly inflated by the presence of disparate research topics (with variable author gender ratios) being published within individual journals.
 
@@
 On balance, we believe the data suggest that it is likely that some researchers preferentially select same-gendered collaborators, although it is difficult to ascertain what proportion of people show such a preference, or how much the strength of the preference varies between individuals.
-We also note that even in a world in which collaboration was completely random with respect to gender, a high proportion of individual researchers would have entirely same-gendered collaborators by chance alone (especially in gender-biased disciplines); thus, individuals who only have same-gendered co-authors are not necessarily doing anything differently from people with gender-balanced co-authors.
+We also note that even in a world in which everyone selected their collaborators at random with respect to gender, a high proportion of individual researchers would have entirely same-gendered collaborators by chance alone (especially in gender-biased disciplines); thus, individuals who only have same-gendered co-authors are not necessarily doing anything differently from people with gender-balanced co-authors.
 

### 2019-03-04 manuscript/homophily_manuscript.tex

--- 2018-10-29 manuscript/Manuscript_Claire.tex
@@
 refer to as the Wahlund effect (\autoref{wahlund_plot}), by analogy with
-the conceptually similar Whalund effect in population genetics {[}59{]}.
+the conceptually similar Wahlund effect in population genetics {[}59{]}.
 The Wahlund effect makes it deceptively difficult to infer gender-based
@@
 
+Nevertheless, when we calculated \(\alpha\) across all non-single-author
+papers in our entire 15-year PubMed dataset (as before, excluding papers
+where at least one author's gender was unknown; n = \textgreater{}3
+million papers, \textgreater{}16 million authors), we found that
+\(\alpha\) was 0.126. This figure is almost double the median value of
+\(\alpha'\) for individual journals (Figure 2; \(\alpha' = 0.070\) for
+`All authors'), suggesting that lumping together papers from different
+fields and different time periods can indeed produce spurious evidence
+for gender homophily as outlined in Figure 1.
+
 \hypertarget{relationship-between-gender-homophily-and-number-of-authors}{%
@@
 average (\autoref{author_number}; statistical results in Online
-Supplmentary Material). Papers with 3, 4 or 5+ authors had essentially
-identical average \(\alpha'\) values. The variance in \(\alpha'\) across
-journals was also a little higher for 2-authors papers compared to the
-remainder (\autoref{author_number}), though part of this variance is due
-to the reduced sample size (in terms of number of authors) for the
-2-author papers. One possible explanation for this finding is that
-2-authors papers might be more likely to have an author list that is
-evenly split between career stages (e.g.~a postgraduate student and
-their supervisor), increasing the chance that the authors are mixed
-gender (see Figure 6). The result also suggests that the processes
-responsible for gender homophily are similar in small (e.g.~3 author)
-and larger (5+ authors) collaborations (and across disciplines where
-small versus large collaborations are the norm).
+Supplementary Material). Papers with 3, 4 or \(\ge5\) authors had
+essentially identical average \(\alpha'\) values. The variance in
+\(\alpha'\) across journals was also a little higher for 2-author papers
+compared to the remainder (\autoref{author_number}), though part of this
+variance is due to the reduced sample size (in terms of number of
+authors) for the 2-author papers. One possible explanation for this
+finding is that 2-authors papers might be more likely to have an author
+list that is evenly split between career stages (e.g.~a postgraduate
+student and their supervisor), increasing the chance that the authors
+are mixed gender (see Figure 6). The result also suggests that the
+processes responsible for gender homophily are similar in small
+(e.g.~3-author) and larger (5+ author) collaborations (and across
+disciplines where small versus large collaborations are the norm).
 
@@
 preferentially assort with same-gendered collaborators even more
-strongly than implied by other results, at least for certain journals or
+strongly than implied by our results, at least for certain journals or
 research disciplines.
@@
 Our study begs two questions: what causes gender homophily in science,
-and are our results cause for concern? These questions are closely
-related. For example, some of the homophily we observed might be caused
-by women seeking to avoid harassment or sexism from men {[}38{]}, which
-would clearly be concerning. Additionally, Sheltzer and Smith {[}66{]}
-concluded that `elite' male academics (defined as recipients of major
-honours) have a higher proportion of male students and postdocs than
-non-elite male academics. This finding could contribute to the homophily
-we observed, and is cause for concern since the results might reflect
-discrimination against women during hiring {[}20{]}, or avoidance by
-women of elite research groups (e.g.~due to gender differences in
-confidence, or a perception that some groups are sexist). We also found
-a little evidence that gender homophily is detrimental to research
-quality, in that high-impact journals tended to have weaker homophily.
-Assuming that papers published in high-impact journals are of higher
-average quality (which is contentious; {[}67{]}), our results provide
-non-experimental support for the hypothesis that mixed-gender teams
-produce better research than single-gender teams {[}42--48{]}. Another
-issue is that if many collaborations are between established
-researchers, there will be an excess of male-male collaborations in
-fields where women in senior positions are rare; some of the observed
-homophily might therefore reflect the elevated gender gap among senior
-researchers.
+and are our results cause for concern? We believe that the answers to
+these questions are closely related. For example, some of the homophily
+we observed might be caused by women seeking to avoid harassment or
+sexism from men {[}38{]}, which would clearly be concerning.
+Additionally, Sheltzer and Smith {[}66{]} concluded that `elite' male
+academics (defined as recipients of major honours) have a higher
+proportion of male students and postdocs than non-elite male academics.
+This finding could contribute to the homophily we observed, and is cause
+for concern since the results might reflect discrimination against women
+during hiring {[}20{]}, or avoidance by women of elite research groups
+(e.g.~due to gender differences in confidence, or a perception that some
+groups are sexist). We also found a little evidence that gender
+homophily is detrimental to research quality, in that high-impact
+journals tended to have weaker homophily. Assuming that papers published
+in high-impact journals are of higher average quality (which is
+contentious; {[}67{]}), our results provide non-experimental support for
+the hypothesis that mixed-gender teams produce better research than
+single-gender teams {[}42--48{]}. Another issue is that if many
+collaborations are between established researchers, there will be an
+excess of male-male collaborations in fields where women in senior
+positions are rare; some of the observed homophily might therefore
+reflect the elevated gender gap among senior researchers.
 
@@
 women, a man can co-author with \(m - 1\) men while women can co-author
-with \(m\) men. Thus, the null expectation for \(alpha\) is a negative
+with \(m\) men. Thus, the null expectation for \(\alpha\) is a negative
 number -- potentially a large one if \(m\) and \(f\) are very small.
@@
 
-In most STEMM subjects, the gender ratio is more skewed among
+In most STEMM disciplines, the gender ratio is more skewed among
 established researchers relative to early-career researchers, due both
 to women leaving STEMM careers at greater rates (the `leaky pipeline'),
-and to historical shortages of women studying STEMM (`demographic
-inertia') {[}1,5{]}. We hypothesised that this difference in gender
-ratio between career stages could potentially create both Wahlund
-effects and `reverse' Wahlund effects. For example, imagine that the
-majority of collaborations in a particular field are between students
-and professors, and that the gender ratio differs between career stages:
-we would then see an excess of mixed-gender coauthorships (heterophily,
-\(\alpha < 0\)), even if gender has no direct, causal effect. Similarly,
-a hypothetical field in which students work only with students, and
-professors with professors, would have apparent gender homophily
-(\(\alpha > 0\)).
+and to historical shortages of women studying STEMM subjects at
+university (`demographic inertia') {[}1,5{]}. We hypothesised that this
+difference in gender ratio between career stages could potentially
+create both Wahlund effects and `reverse' Wahlund effects. For example,
+imagine that the majority of collaborations in a particular field are
+between students and professors, and that the gender ratio differs
+between career stages: we would then see an excess of mixed-gender
+coauthorships (heterophily, \(\alpha < 0\)), even if gender has no
+direct, causal effect. Similarly, a hypothetical field in which students
+work only with students, and professors with professors, would have
+apparent gender homophily (\(\alpha > 0\)).
 
@@
 
-CM was supported by the Academy of Finland (284666 to the Centre of
-Excellence in Biological Interactions). We thank Devi Stuart-Fox and
-Dominique Potvin for helpful discussion.
+We thank Devi Stuart-Fox and Dominique Potvin for helpful discussion.
 

### 2019-03-04 manuscript/homophily_manuscript.tex
Ready for plos resubmission

@@
 
-\usepackage{microtype} \usepackage[all]{hypcap} \usepackage{amsmath}
-\usepackage{booktabs} \usepackage{caption}
-\usepackage[labelfont=bf]{caption} \usepackage{titling}
+\usepackage{microtype} \usepackage[all]{hypcap} \usepackage{amsmath} \usepackage{booktabs} \usepackage{caption} \usepackage[labelfont=bf]{caption} \usepackage{titling}
+
 \pretitle{\begin{flushleft}\LARGE}
@@
 \postauthor{\par\end{flushleft}} \predate{\begin{flushleft}\large}
-\postdate{\par\end{flushleft}} \usepackage{fancyhdr} \pagestyle{fancy}
-\fancyhead[LO,LE]{\textsl{\leftmark}} \rhead[]{Gender and coauthorship}
-\usepackage[noabbrev,capitalise]{cleveref} \usepackage{titlefoot}
-\usepackage{amssymb} \usepackage{rotating} \usepackage{lineno}
+\postdate{\par\end{flushleft}}
+
+\usepackage{fancyhdr}
+
+\pagestyle{fancy} \fancyhead[LO,LE]{\textsl{\leftmark}}
+\rhead[]{Gender and coauthorship}
+
+\usepackage[noabbrev,capitalise]{cleveref} \usepackage{titlefoot} \usepackage{amssymb} \usepackage{rotating} \usepackage{lineno}
+
 \linenumbers
@@
 less prestigious collaborations {[}33{]}, and fewer collaborations in
-total {[}34{]}. These gender differences in collaboration practice
-presumably have multiple causes, which might include implicit and
-explicit gender bias {[}20{]}, differential family obligations
-{[}33,35,36{]}, gender differences in confidence or self-esteem
-{[}37{]}, concerns relating to sexual harassment {[}38{]}, and unequal
-access to conferences {[}39{]} or travel funds {[}32{]}.
+total {[}34{]}. These gender differences in collaboration presumably
+have multiple causes, which might include implicit and explicit gender
+bias {[}20{]}, differential family obligations {[}33,35,36{]}, gender
+differences in confidence or self-esteem {[}37{]}, concerns relating to
+sexual harassment {[}38{]}, and unequal access to conferences {[}39{]}
+or travel funds {[}32{]}.
 
@@
 men, and women with women, more often than expected if collaborators
-assort randomly with respect to gender {[}49--58{]}. This pattern of
-assortative publishing has often been termed `gender homophily'.
-
-However, we believe that earlier studies of gender homophily were
+assort randomly with respect to gender {[}49--58{]}. This non-random
+assortment is often termed `gender homophily'.
+
+We believe that most or all earlier studies of gender homophily were
 hindered by a largely unacknowledged statistical issue that we will
@@
 that our estimates of homophily, and estimates from some of the earlier
-studies listed in the Introduction, are probably not inflated by the
-presence of disparate research topics (with variable author gender
-ratios) being published within individual journals.
+studies listed in the Introduction, are probably not markedly inflated
+by the presence of disparate research topics (with variable author
+gender ratios) being published within individual journals.
 
@@
 the preference varies between individuals. We also note that even in a
-world in which collaboration was completely random with respect to
-gender, a high proportion of individual researchers would have entirely
-same-gendered collaborators by chance alone (especially in gender-biased
-disciplines); thus, individuals who only have same-gendered co-authors
-are not necessarily doing anything differently from people with
-gender-balanced co-authors.
+world in which everyone selected their collaborators at random with
+respect to gender, a high proportion of individual researchers would
+have entirely same-gendered collaborators by chance alone (especially in
+gender-biased disciplines); thus, individuals who only have
+same-gendered co-authors are not necessarily doing anything differently
+from people with gender-balanced co-authors.
 

### 2019-04-11 Gender_homophily_analysis.Rmd
Response to 3rd plos review

@@
 ```{r}
-fig_s7 <- master.dataset %>%
+fig_s8 <- master.dataset %>%
   ggplot(aes(x = observed.alpha, y = alpha.adjusted)) + 
@@
   theme_minimal()
-ggsave(fig_s7, file = "figures/S7 Fig.pdf", height = 5, width = 5, device = cairo_pdf)
-fig_s7
+ggsave(fig_s8, file = "figures/S8 Fig.pdf", height = 5, width = 5, device = cairo_pdf)
+fig_s8
 ```
 <br></br> 
-**Figure S7**: There is a very strong correlation between the values of $\alpha$ and $\alpha'$ calculated for each journal, though in a handful of cases the difference is considerable.
-The deviation between $\alpha$ and $\alpha'$ is greatest for journals for which there is a small sample size (see S8 Fig).
+**S8 Figure**: There is a very strong correlation between the values of $\alpha$ and $\alpha'$ calculated for each journal, though in a handful of cases the difference is considerable.
+The deviation between $\alpha$ and $\alpha'$ is greatest for journals for which there is a small sample size (see S9 Fig).
 
@@
 ```{r}
-fig_s8 <- master.dataset %>%
+fig_s9 <- master.dataset %>%
   filter(Age == "New") %>%
@@
   theme_minimal()
-ggsave(fig_s8, file = "figures/S8 Fig.pdf", height = 5, width = 5, device = cairo_pdf)
-fig_s8
+ggsave(fig_s9, file = "figures/S9 Fig.pdf", height = 5, width = 5, device = cairo_pdf)
+fig_s9
 ```
 <br></br> 
-**Figure S8**: For journals for which we recovered a small number of papers (<100), the unadjusted metric $\alpha$ was downwardly biased.
+**S9 Figure**: For journals for which we recovered a small number of papers (<100), the unadjusted metric $\alpha$ was downwardly biased.
 This fits our expectations: because authors cannot be their own co-authors, small datasets will tend to produce negative estimates of $\alpha$ even if authors assort randomly with respect to gender (see main text).
@@
 <br></br>
-**Figure S1**: Plot showing the percentage of papers that have 1, 2, 3, 4, or ${\ge}5$ authors for each discipline in the dataset of Holman et al.
+**S1 Figure**: Plot showing the percentage of papers that have 1, 2, 3, 4, or ${\ge}5$ authors for each discipline in the dataset of Holman et al.
 (2018).
@@
 <br></br>
-**Figure S2**: Histogram showing the distribution of differences in $\alpha'$ between the 2015-16 and 2005-6 samples, where positive numbers indicate an increase in $\alpha'$ with time.
+**S2 Figure**: Histogram showing the distribution of differences in $\alpha'$ between the 2015-16 and 2005-6 samples, where positive numbers indicate an increase in $\alpha'$ with time.
 The mean is slightly positive (i.e.
@@
 <br></br>
-**Figure S3**: Histogram showing the difference between $\alpha'$ calculated for first and last authors.
+**S3 Figure**: Histogram showing the difference between $\alpha'$ calculated for first and last authors.
 Positive values mean that $\alpha'$ was higher when calculated for first authors, and negative values mean $\alpha'$ was higher when calculated for last authors.
@@
   theme(legend.position = "none") 
-ggsave(plot = gender.ratio, filename = "figures/Fig4.pdf", 
+ggsave(plot = gender.ratio, filename = "figures/S4 Fig.pdf", 
        device = cairo_pdf, width=7, height=5)
@@
 <br></br>
-**Figure 4**: There is a weakly positive, non-linear relationship between the gender ratio of authors publishing in a journal, and the coefficient of homophily ($\alpha'$).
+**S4 Figure**: There is a weakly positive, non-linear relationship between the gender ratio of authors publishing in a journal, and the coefficient of homophily ($\alpha'$).
 Specifically, journals with 50% women authors or higher tended to have more same-sex coauthorships than did journals with predominantly men authors.
@@
 
-figure5 <- gender.homophily.all %>% 
+figure4 <- gender.homophily.all %>% 
   filter(!is.na(standardised.IF)) %>% 
@@
   theme_minimal()
-ggsave(figure5, file = "figures/Fig5.pdf", height = 5, width = 5, device = cairo_pdf)
-saveRDS(R2, "manuscript/R2_for_figure5.RData")
-figure5
+ggsave(figure4, file = "figures/Fig4.pdf", height = 5, width = 5, device = cairo_pdf)
+saveRDS(R2, "manuscript/R2_for_figure4.RData")
+figure4
 ```
 <br></br>
-**Figure 5**: Journal impact factor (expressed relative to the average for the discipline) is negatively correlated with $\alpha'$.
+**Figure 4**: Journal impact factor (expressed relative to the average for the discipline) is negatively correlated with $\alpha'$.
 The relationship is weak ($R^2$ = `r R2`), but the results suggest that journals with strong homophily tend to have lower impact factors than journals with weak homophily in the same discipline.
+
+#### Spearman correlation for the relationship in Figure 4
+
+```{r}
+with(master.dataset %>% 
+       filter(Age == "New", 
+              Subset == "All authors", 
+              !is.na(standardised.IF)), cor.test(alpha.adjusted, scale(standardised.IF), method = "spearman"))
+```
 
@@
   theme(legend.position = "top")
-ggsave(country.histo, file = "figures/S4 Fig.pdf", width = 5, height = 5, device = cairo_pdf)
+ggsave(country.histo, file = "figures/S5 Fig.pdf", width = 5, height = 5, device = cairo_pdf)
 country.histo
@@
 <br></br>
-**Figure S4**: Histogram of $\alpha'$ for 325 unique combinations of journal and country, using data from August 2015 - August 2016.
+**S5 Figure**: Histogram of $\alpha'$ for 325 unique combinations of journal and country, using data from August 2015 - August 2016.
 The white areas denote combinations for which $\alpha'$ differs significantly from zero (p < 0.05, following false discovery rate correction).
@@
 
-ggsave(mean.inflation.due.to.country, file = "figures/S6 Fig.pdf", width = 5, height = 5, device = cairo_pdf)
+ggsave(mean.inflation.due.to.country, file = "figures/S7 Fig.pdf", width = 5, height = 5, device = cairo_pdf)
 mean.inflation.due.to.country
@@
 <br></br>
-**Fig S6**: Histogram showing the estimated degree to which $\alpha'$ is inflated by inter-country differences in author gender ratio, across the 285 journals for which we had adequate data after restricting the analysis by country.
+**S7 Figure**: Histogram showing the estimated degree to which $\alpha'$ is inflated by inter-country differences in author gender ratio, across the 285 journals for which we had adequate data after restricting the analysis by country.
 The average inflation in $\alpha'$ is negligible, suggesting that Wahlund effects resulting from inter-country differences have a neglible effect on our estimates of gender homophily.
@@
 x <- nrow(country.data[country.data$p.adjusted < 0.05, ])
-fig_s5 <- country.data %>% 
+fig_s6 <- country.data %>% 
   filter(p.adjusted < 0.05) %>%
@@
   theme(legend.position = "none")
-ggsave(fig_s5, file = "figures/S5 Fig.pdf", width = 7, height = 9, device = cairo_pdf)
-fig_s5
+ggsave(fig_s6, file = "figures/S6 Fig.pdf", width = 7, height = 9, device = cairo_pdf)
+fig_s6
 ```
 <br></br>
-**Figure S5**: Plot showing the 68 combinations of journal and author country of affiliation for which $\alpha'$ is significantly higher than expected.
+**S6 Figure**: Plot showing the 68 combinations of journal and author country of affiliation for which $\alpha'$ is significantly higher than expected.
 
@@
 
-This section makes Figure 6, by deriving the expected $\alpha$ under various assumptions, assuming that there is (potentially) a gender gap between career stages but no direct choice of collaborators by gender.
+This section makes Figure 5, by deriving the expected $\alpha$ under various assumptions, assuming that there is (potentially) a gender gap between career stages but no direct choice of collaborators by gender.
 The model has four parameters:
 
-  - The proportion of women among people in career stage 1 (e.g. students), termed `GR.stage1` in the code below (this parameter is plotted on the x-axis of Figure 6)
-  - The proportion of women among people in career stage 2 (e.g. professors), termed `GR.stage2` in the code below (this parameter is plotted on the y-axis of Figure 6)
-  - The proportion of co-authorships that are between two people at career stages 1 and 2, termed `prop_between_career_stages` in the code below (this parameter is plotted in the columns of Figure 6)
-  - The proportion of within-career stage co-authorships that are between two people at career stage 1, termed `prop_within_stage_between_two_stage1s` in the code below (this parameter is plotted in the rows of Figure 6)
+  - The proportion of women among people in career stage 1 (e.g. students), termed `GR.stage1` in the code below (this parameter is plotted on the x-axis of Figure 5)
+  - The proportion of women among people in career stage 2 (e.g. professors), termed `GR.stage2` in the code below (this parameter is plotted on the y-axis of Figure 5)
+  - The proportion of co-authorships that are between two people at career stages 1 and 2, termed `prop_between_career_stages` in the code below (this parameter is plotted in the columns of Figure 5)
+  - The proportion of within-career stage co-authorships that are between two people at career stage 1, termed `prop_within_stage_between_two_stage1s` in the code below (this parameter is plotted in the rows of Figure 5)
 
@@
 
-# Plot the results to make figure 6
+# Plot the results to make figure 5
 # NB that I edited the top- and right-side labels in Inkscape afterwards, as it was tricky to do it entirely in R.
@@
 model_figure
-ggsave(model_figure, file = "figures/Fig6.pdf", 
+ggsave(model_figure, file = "figures/Fig5.pdf", 
        height = 5.6, width = 5, device = cairo_pdf)
 ```
+<br></br>
+**Figure 5**: When the gender ratio of early-career researchers is not equal to the gender ratio among established researchers, the null expectation for $\alpha$ is not necessarily zero.
+Specifically, if most collaborations occur between career stages, there will be an excess of mixed-gender collaborations ($\alpha < 0$, blue areas), while if most collaborator pairs comprise two people at the same career stage, there will be an excess of same-gender collaborations ($\alpha > 0$, red areas).
+However, the conditions required for strong gender homophily (i.e.
+the red areas) are quite restrictive, making it unlikely that this issue can fully explain the homophily observed in our study.
+Additionally, in research disciplines where between-career stage collaboration is common and there is a shortage of women among established researchers (i.e.
+the blue areas), our study will underestimate the strength of gender homophily.
+Contour lines denote increments of 0.1.
 

### 2019-04-11 manuscript/homophily_manuscript.Rmd
Response to 3rd plos review

@@
 
-R2 <- readRDS("R2_for_figure5.RData") # For Fig 4
+R2 <- readRDS("R2_for_figure4.RData") # R^2 For Fig 4
 ```
@@
 
-We believe that most or all earlier studies of gender homophily were hindered by a largely unacknowledged statistical issue that we will refer to as the Wahlund effect (\autoref{wahlund_plot}), by analogy with the conceptually similar Wahlund effect in population genetics [@Wahlund_1928].
-The Wahlund effect makes it deceptively difficult to infer gender-based preferences simply by counting the number of same- and mixed-gender coauthorships.
-Essentially, whenever coauthorship data are sampled from two or more discrete sets of literature, which vary in the author gender ratio and which are largely not connected by collaboration, the number of same-gendered coauthors will be inflated.
+We believe that most or all earlier studies of gender homophily were hindered by a largely unacknowledged statistical issue that we will refer to as the Wahlund effect (Fig 1), by analogy with the conceptually similar Wahlund effect in population genetics [@Wahlund_1928].
+The Wahlund effect makes it deceptively difficult to test for gender-based co-author choice simply by counting the relative number of same- and mixed-gender coauthorships.
+Essentially, the Wahlund effect means that whenever coauthorship data are sampled from two or more discrete sets of literature, which vary in the author gender ratio and which are largely unconnected by collaboration, the number of same-gendered coauthors will be inflated.
 This can give the impression that authors preferentially publish with same-gendered colleagues even if no gender preferences exist, or if the true preference is for opposite-gendered colleagues ('gender heterophily').
@@
 
-![The Wahlund effect can make it appear as if authors publish with same-gendered colleagues disproportionately often, even if collaboration is completely random with respect to gender.
+\vspace{2cm}
+**Fig 1**: The Wahlund effect can make it appear as if authors publish with same-gendered colleagues disproportionately often, even if collaboration is completely random with respect to gender.
 Here, coloured circles represent male and female authors, and coauthors are linked with lines.
@@
 However, within each subset, there is no evidence that authors prefer to publish with same-gendered individuals (if anything, this small dataset suggests gender heterophily).
-The Wahlund effect will tend to inflate the frequency of same-gender coauthorships whenever the data is composed of two or more disconnected subsets of literature with different author gender ratios; these subsets could be research disciplines, older versus newer papers, or papers from authors in different countries.
+The Wahlund effect will tend to inflate the frequency of same-gender coauthorships whenever the data is composed of two or more disconnected subsets of literature with different author gender ratios; these subsets could be research disciplines, older versus newer papers, or papers from authors based in different countries.
 The example countries and disciplines were selected based on data in [@Holman_2018].
-\label{wahlund_plot}](../figures/Fig1.pdf)
+\vspace{2cm}
 
-In the present study, we test whether life sciences researchers tend to co-publish with same-gendered colleagues, while controlling for the Wahlund effect as strictly as possible.
+In the present study, we test whether life sciences researchers tend to co-publish with same-gendered colleagues, while controlling for the Wahlund effect as strictly as we are able.
 We use a recently-published dataset describing the gender of 35.5 million authors from 9.15 million articles indexed on PubMed [@Holman_2018].
@@
 [-@Holman_2018] reported large differences in the gender ratio of authors across research disciplines, journals, countries, and across the years 2002-2016.
-We therefore tested for gender homophily while restricting our analysis to particular journals (i.e.
-research specialties), time periods, and countries.
+We therefore tested for gender homophily while restricting our analysis to particular journals (a proxy for research specialties), time periods, and countries.
 We quantified gender assortment using a metric called $\alpha'$ [@bergstrom_2016], which is positive when same-gender authors publish together more often than expected (gender homophily), negative when opposite-gender authors publish together more often than expected (heterophily), and equal to zero when authors assort randomly with respect to gender (see Methods).
@@
 ## Gender homophily by discipline, time period, and authorship position
-\autoref{alpha_histograms} shows the distribution of $\alpha'$ estimates in 2015-2016 across all journals for which we recovered sufficient data, when $\alpha'$ was calculated for all authors, first authors only, or last authors only.
+Fig 2 shows the distribution of $\alpha'$ estimates in 2015-2016 across all journals for which we recovered sufficient data, when $\alpha'$ was calculated for all authors, first authors only, or last authors only.
 Most journals had positive values of $\alpha'$ (77-92%, depending on time period and author type; S1 Data), and for many of these the false discovery rate (FDR)-corrected p-values suggested that $\alpha'$ was significantly greater than zero (1469/2077 journals were significant in 2015-16, and 404/1192 in 2005-6; S1 Data).
@@
 
-![Of the `r n.journals` journals for which we had adequate data in 2015-2016, `r n.sig.homo` showed statistically significant evidence of gender homophily (denoted by $\alpha' > 0$), and `r n.sig.hetero` showed statistically significant evidence of heterophily ($\alpha' < 0$), after false discovery rate correction.
+\vspace{2cm}
+**Fig 2**: Of the `r n.journals` journals for which we had adequate data in 2015-2016, `r n.sig.homo` showed statistically significant evidence of gender homophily (denoted by $\alpha' > 0$), and `r n.sig.hetero` showed statistically significant evidence of heterophily ($\alpha' < 0$), after false discovery rate correction.
 In the stacked density plot, the white area shows the number of journals for which homophily was significantly stronger than expected under the null hypothesis (corrected p < 0.05), while the blue area shows all the remainder.
 Patterns were similar whether $\alpha'$ was calculated for all authors, for first authors only, or for last authors only.
-Points in the right panel show $\alpha'$ for individual journals.\label{alpha_histograms}](../figures/Fig2.pdf)
+Points in the right panel show $\alpha'$ for individual journals.
+\vspace{2cm}
 
@@
 ## Variance in homophily between disciplines
-\autoref{alpha_histograms} illustrates the variance in journal homophily values ($\alpha'$) across scientific disciplines.
+Fig 2 illustrates the variance in journal homophily values ($\alpha'$) across scientific disciplines.
 All disciplines had positive mean $\alpha'$ (averaged over journals), although homophily appeared somewhat stronger in some disciplines than others (e.g.
-mean $\alpha'$ was $0.12{\pm}0.02$ for Urology journals and $0.03{\pm}0.01$ for Veterinary Medicine journals; \autoref{alpha_histograms}, S4 Data).
-However, there was no formal evidence for consistent differences in $\alpha'$ between disciplines: the random factor 'Discipline' explained around 1% of the variance in $\alpha'$ in the two linear mixed models described in the previous section (see \autoref{alpha_histograms} and mixed models in Online Supplementary Material).
-Thus, whatever processes are responsible for producing positive $\alpha'$ values appear to be similarly strong in all the disciplines we examined.
+mean $\alpha'$ was $0.12{\pm}0.02$ for Urology journals and $0.03{\pm}0.01$ for Veterinary Medicine journals; Fig 2, S4 Data).
+However, there was no formal evidence for consistent differences in $\alpha'$ between disciplines: the random factor 'Discipline' explained around 1% of the variance in $\alpha'$ in the two linear mixed models described in the previous section (see Fig 2 and mixed models in Online Supplementary Material).
+Thus, the causal mechanisms underlying the observed positive $\alpha'$ values appear to be similarly strong in all the disciplines we examined.
 
-There was no indication that journals publishing on a wide range of topics have higher $\alpha'$ values than more specialised journals, due to the Wahlund effect.
-For example, the journal category 'Multidisciplinary' -- which includes journals like _PLOS ONE_, _Nature_, _Science_, and _PNAS_ -- did not have markedly elevated $\alpha'$ (\autoref{alpha_histograms}).
-This result suggests that our estimates of homophily, and estimates from some of the earlier studies listed in the Introduction, are probably not markedly inflated by the presence of disparate research topics (with variable author gender ratios) being published within individual journals.
+There was no indication that journals publishing on a wide range of topics have higher $\alpha'$ values than more specialised journals due to the Wahlund effect (Fig 1).
+For example, the journal category 'Multidisciplinary' -- which includes general interest journals like _PLOS ONE_, _Nature_, _Science_, and _PNAS_ -- did not have markedly elevated $\alpha'$ (Fig 2).
+This result suggests that our estimates of homophily, and estimates from some of the earlier studies of homophily listed in the Introduction, are probably not markedly inflated by the presence of disparate research topics (with variable author gender ratios) being published within individual journals.
 
 Nevertheless, when we calculated $\alpha$ across all non-single-author papers in our entire 15-year PubMed dataset (as before, excluding papers where at least one author's gender was unknown; n = >3 million papers, >16 million authors), we found that $\alpha$ was 0.126.
-This figure is almost double the median value of $\alpha'$ for individual journals (Figure 2; $\alpha' = 0.070$ for 'All authors'), suggesting that lumping together papers from different fields and different time periods can indeed produce spurious evidence for gender homophily as outlined in Figure 1.
+This figure is almost double the median value of $\alpha'$ for individual journals (Fig 2; $\alpha' = 0.070$ for 'All authors'), suggesting that lumping together papers from different fields and different time periods can indeed produce spurious evidence for gender homophily as outlined in Fig 1.
 
 ## Relationship between gender homophily and number of authors
-Papers with two authors had significantly lower (but still positive) $\alpha'$ values relative to papers with more than two authors, on average (\autoref{author_number}; statistical results in Online Supplementary Material).
-Papers with 3, 4 or $\ge5$ authors had essentially identical average $\alpha'$ values.
-The variance in $\alpha'$ across journals was also a little higher for 2-author papers compared to the remainder (\autoref{author_number}), though part of this variance is due to the reduced sample size (in terms of number of authors) for the 2-author papers.
-One possible explanation for this finding is that 2-authors papers might be more likely to have an author list that is evenly split between career stages (e.g.
-a postgraduate student and their supervisor), increasing the chance that the authors are mixed gender (see Figure 6).
-The result also suggests that the processes responsible for gender homophily are similar in small (e.g.
-3-author) and larger (5+ author) collaborations (and across disciplines where small versus large collaborations are the norm).
+Papers with two authors had significantly lower (but still positive) $\alpha'$ values relative to papers with more than two authors, while papers with 3, 4 or $\ge5$ authors had essentially identical average $\alpha'$ values (Fig 3).
+Specifically, the posterior estimate of mean $\alpha'$ was 0.014 (95% CIs: 0.002 - 0.026) for 2-author papers and 0.065 (95% CIs: 0.056 - 0.074) for 3-author papers (and roughly the same for 4- and $\ge5$-author papers; Fig 3).
+One possible explanation for this finding is that 2-authors papers are more likely to have an author list that is evenly split between career stages (e.g.
+a postgraduate student and their supervisor), increasing the chance that the authors are mixed gender (see section 'Theoretical expectations for $\alpha$ when the gender ratio differs between career stages').
+The result also suggests that the causal mechanisms responsible for gender homophily are similar in small (e.g.
+3-author) and larger ($\ge5$ author) collaborations (and across disciplines where small versus large collaborations are the norm).
 
-![The coefficient of homophily ($\alpha'$) was slightly less positive when calculated for two-author papers only, relative to papers with longer author lists.
+\vspace{2cm}
+**Fig 3**: The coefficient of homophily ($\alpha'$) was slightly less positive when calculated for two-author papers only, relative to papers with longer author lists.
 The individual points, whose distribution is summarised by the violin plots, correspond to individual journals.
 The larger white points show the mean for each group (and its 95% CIs), as calculated by a Bayesian meta-regression model accounting for repeated measures of $\alpha'$ within journals, as well as the precision with which $\alpha'$ was estimated.
-\label{author_number}](../figures/Fig3.pdf)
+\vspace{2cm}
 
@@
 Psychiatry).
-We found a positive, non-linear relationship between the overall gender ratio of all authors publishing in a particular journal [as estimated in @Holman_2018], and the estimated value of $\alpha'$ for all authors and for first authors (\autoref{alpha_gender_ratio}).
+We found a positive, non-linear relationship between the gender ratio across all the authors publishing in a particular journal [as estimated in @Holman_2018], and the estimated value of $\alpha'$ for all authors and for first authors, but not last authors (S4 Fig).
 Journals with a balanced or female-biased author gender ratio tended to have higher $\alpha'$ (i.e.
-stronger homophily) than journals with a male-biased author gender ratio (GAM smooth terms p < 0.001; Online Supplementary Material).
-The relationship was not statistically significant when $\alpha'$ was calculated for last authors (GAM, p = 0.142), though the trend appeared similar (\autoref{alpha_gender_ratio}).
-
-![There is a weakly positive, non-linear relationship between the gender ratio of authors publishing in a journal, and the coefficient of homophily ($\alpha'$).
-Specifically, journals with 50% women authors or higher tended to have more same-sex coauthorships than did journals with predominantly men authors.
-This relationship held whether $\alpha'$ was calculated for all authors, first authors only, or last authors only.
-A negative value on the x-axis denotes an excess of men authors, a positive value denotes an excess of women authors, and zero denotes gender parity (i.e.
-equal numbers of men and women).
-The lines were fitted using generalised additive models with the smoothing parameter $k$ set to 3.
-\label{alpha_gender_ratio}](../figures/Fig4.pdf)
+stronger homophily) than journals with a male-biased author gender ratio (GAM smooth term p = 0.0002 for all-author homophily, p < 0.0001 for first-author homophily, and p = 0.13 for last-author homophily).
 
 ## Relationship between journal impact factor and gender homophily
-We observed a noisy but statistically significant linear relationship between standardised journal impact factor and $\alpha'$, such that journals with a high impact factor for their discipline had weaker gender homophily than did journals with a low impact factor for their discipline (\autoref{impact_factor}; linear regression: $R^2$ = `r R2`, $t_{1415}$ = -8.0, p < 0.0001).
+We observed a noisy but statistically significant linear relationship between standardised journal impact factor and $\alpha'$, such that journals with a high impact factor for their discipline had weaker gender homophily than did journals with a low impact factor for their discipline (Fig 4; linear regression: $R^2$ = `r R2`, $t_{1415}$ = -8.0, p < 0.0001).
 The slope of the regression was $-0.012{\pm}0.0015$, indicating that increasing the discipline-standardised impact factor by one standard deviation is associated with a reduction in $\alpha'$ of 0.012.
+The Spearman correlation coefficient was -0.19 (p < 0.0001).
 
-![Journal impact factor (expressed relative to the average for the discipline) is negatively correlated with $\alpha'$.
-The relationship is noisy ($R^2$ = `r R2`), but the results suggest that journals with strong homophily tend to have lower impact factors than journals with weak homophily in the same discipline.
-\label{impact_factor}](../figures/Fig5.pdf)
+\vspace{2cm}
+**Fig 4**: Journal impact factor (expressed relative to the average for the discipline) is negatively correlated with $\alpha'$.
+The relationship is noisy ($R^2$ = `r R2`, Spearman correlation = -0.19), but the results indicate that journals with strong gender homophily tend to have lower impact factors than journals from the same discipline that have weaker homophily.
+\vspace{2cm}
 
 ## Analysis accounting for differences in author gender ratio between countries
-When we restricted the analysis by country, we observed statistically significant homophily for 72 of the 325 journal-country combinations tested (64 unique journals and 18 unique countries), and no significant heterophily (S4-S5 Fig).
+When we restricted the analysis by country, we observed statistically significant homophily for 72 of the 325 journal-country combinations tested (64 unique journals and 18 unique countries), and no significant heterophily (S5-S6 Fig).
 Additionally, the values of $\alpha'$ calculated for each journal-country combination were only very slightly lower than the $\alpha'$ values calculated for the journal as a whole (i.e.
-when pooling papers from different countries, as was done to make \autoref{alpha_histograms}): on average, the difference in $\alpha'$ was only 0.002 (S6 Fig).
+when pooling papers from different countries, as was done to make Fig 2): on average, the difference in $\alpha'$ was only 0.002 (S7 Fig).
 These results suggest that our findings of widespread homophily in the main analysis were not driven solely by a Wahlund effect resulting from gender differences between countries.
@@
 Given that we cannot identify individual researchers or their career stages, we used a simple model to derive the theoretical expectations for $\alpha$ when the gender ratio differs between career stages (see Methods).
-As shown in \autoref{simulation}, we predict that $\alpha$ is expected to be non-zero, even if collaborators are randomly selected with respect to gender, provided that there is a gender gap between career stages.
-The extent to which $\alpha$ deviates from zero depends on the relative frequencies of collaboration within and between career stages (rows and columns in Figure 6), and the size of the gender gap between stages (x- and y-axes in Figure 6).
+As shown in Fig 5, we predict that $\alpha$ is expected to be non-zero, even if collaborators are randomly selected with respect to gender, provided that there is a gender gap between career stages.
+The extent to which $\alpha$ deviates from zero depends on the relative frequencies of collaboration within and between career stages (rows and columns in Fig 5), and the size of the gender gap between stages ($x$- and $y$-axes in Fig 5).
 When >50% of coauthor pairs comprise one early-career and one established researcher, we expect gender heterophily ($\alpha < 0$) whenever the gender ratio differs between career stages.
 Conversely, when >50% of collaborations are between people at the same career stage, we expect gender homophily ($\alpha > 0$).
-In a few parameter spaces (shown in red; \autoref{simulation}), $\alpha$ was quite high, and overlapped with the values that we estimated (\autoref{alpha_histograms}).
+In a few parameter spaces (shown in red; Fig 5), $\alpha$ was quite high, and overlapped with the values that we estimated (Fig 2).
 
-![When the gender ratio of early-career researchers is not equal to the gender ratio among established researchers, the null expectation for $\alpha$ is not necessarily zero.
+\vspace{2cm}
+**Fig 5**: When the gender ratio of early-career researchers is not equal to the gender ratio among established researchers, the null expectation for $\alpha$ is not necessarily zero.
 Specifically, if most collaborations occur between career stages, there will be an excess of mixed-gender collaborations ($\alpha < 0$, blue areas), while if most collaborator pairs comprise two people at the same career stage, there will be an excess of same-gender collaborations ($\alpha > 0$, red areas).
@@
 the blue areas), our study will underestimate the strength of gender homophily.
-Contour lines denote increments of 0.1.
-\label{simulation}](../figures/Fig6_inkscape.pdf)
+Contour lines mark increments of 0.1.
+\vspace{2cm}
 
-Despite this overlap, \autoref{simulation} suggests that our main conclusions (and those of other studies of gender homophily) are probably robust to this career stage issue.
+Despite this overlap, Fig 5 suggests that our main conclusions (and those of other studies of gender homophily) are probably robust to this career stage issue.
 We only expect strongly positive $\alpha$ when A) the gender ratio is highly skewed across career stages (e.g.
@@
 
-Lastly, we note that if there is a gender gap between career stages and coauthorships between early-career and established researchers comprise >50% of the total, then the baseline expectation for $\alpha$ is actually less than zero (blue areas in \autoref{simulation}).
+Lastly, we note that if there is a gender gap between career stages and coauthorships between early-career and established researchers comprise >50% of the total, then the baseline expectation for $\alpha$ is actually less than zero (blue areas in Fig 5).
 Therefore, it is possible that researchers preferentially assort with same-gendered collaborators even more strongly than implied by our results, at least for certain journals or research disciplines.
@@
 # Discussion
-We found evidence that researchers work with same-gendered coauthors more often than expected under the null model, even after implementing stringent controls for Wahlund effects (\autoref{wahlund_plot}).
+We found evidence that researchers work with same-gendered coauthors more often than expected under the null model, even after implementing stringent controls for Wahlund effects of the kind illustrated in Fig 1.
 Our study therefore reaffirms earlier studies' conclusions [e.g.
@@
 An important limitation of our study is that we cannot reliably determine the cause(s) of the observed excess of same-gender coauthorships.
-As well as the obvious interpretation -- conscious or unconscious selection of same-gender collaborators by men, by women, or both -- our results could be partly explained by uncontrolled Wahlund effects.
-However, we suspect the contribution of these uncontrolled artefacts to be minor, for four reasons: we found positive $\alpha'$ after controlling for three obvious sources of Wahlund effect; there was no inflation of $\alpha'$ in highly multidisciplinary journals relative to specialised journals; restricting the data by country yielded similar estimates of $\alpha'$; and we used a simple model to show that differences in gender ratio between career stages are unlikely to fully explain our results.
-On balance, we believe the data suggest that it is likely that some researchers preferentially select same-gendered collaborators, although it is difficult to ascertain what proportion of people show such a preference, or how much the strength of the preference varies between individuals.
+As well as the obvious interpretation -- conscious or unconscious selection of same-gendered collaborators by men, women, or both genders -- our results could be partly explained by uncontrolled Wahlund effects.
+However, we suspect the contribution of these uncontrolled artefacts to be minor, for four reasons: we found positive $\alpha'$ after controlling for three obvious sources of Wahlund effect; there was no inflation of $\alpha'$ in highly multidisciplinary journals relative to specialised journals; restricting the data by country yielded similar estimates of $\alpha'$; and our modelling work suggested that differences in gender ratio between career stages are unlikely to fully explain our results.
+On balance, we believe the data suggest that it is likely that some researchers preferentially select same-gendered collaborators, although it is difficult to ascertain what proportion of people show such a preference, or how much the strength of the preference varies between individual researchers.
 We also note that even in a world in which everyone selected their collaborators at random with respect to gender, a high proportion of individual researchers would have entirely same-gendered collaborators by chance alone (especially in gender-biased disciplines); thus, individuals who only have same-gendered co-authors are not necessarily doing anything differently from people with gender-balanced co-authors.
@@
 We hypothesised that disciplines with a strongly skewed gender ratio might show the strongest gender homophily, e.g.
-because being in the minority might increase people's motivation to seek out same-gendered colleagues.
-Contrary to this hypothesis, we found no evidence that gender homophily is restricted to particular disciplines: $\alpha'$ was similarly high across the board (\autoref{alpha_histograms}).
+because being in the minority might increase one's motivation to seek out same-gendered colleagues.
+Contrary to this hypothesis, we found no evidence that gender homophily is restricted to particular disciplines: $\alpha'$ was similarly high across the board (Fig 2).
 Interestingly, gender homophily was weakest for journals with a male-biased author gender ratio, and strongest in journals with a female-biased author gender ratio.
 One possible reason is that men are more likely to preferentially seek out male collaborators in fields where men are a minority, relative to the homophily displayed by women in fields where women are a minority.
-However, this latter result only has tentative statistical support since our sample contains few journals in which most authors are women (\autoref{alpha_gender_ratio}).
+However, this latter result only has tentative statistical support since our sample contains few journals in which most authors are women (S4 Fig).
 
@@
 @Long_2015; @Bendels_2018; @McKenzie_2017].
-As shown in \autoref{simulation}, if enough coauthorships are between junior and senior researchers, a large gender gap between career stages can give the appearance of heterophily.
+As shown in Fig 5, if enough coauthorships are between junior and senior researchers, a large gender gap between career stages can give the appearance of heterophily.
 As this gender gap between career stages lessens, the observed values of $\alpha'$ may increase.
@@
 By contrast, we feel that it may be counter-productive to recommend that women collaborate primarily with other women, e.g.
-because this constrains women's options (particularly in fields like engineering, where 90% of professors are men [@Shaw_2012]).
+because this constrains women's options (particularly in fields like the one studied by Ghiasi et al.
+-- engineering -- where 90% of professors are men [@Shaw_2012]).
 Instead, we suggest that researchers of both genders can help to close the gender gap in STEMM.
@@
 One should also take care to treat male and female collaborators equally, e.g.
-in terms of training and mentoring, allocation of work, how one frames or promotes the collaboration (e.g.
-in conference presentations or on a website), and in terms of the credit afforded to collaborators' contributions.
-Experimental work suggests that unconscious bias causes people to undervalue women's research achievements [@Moss_2012], and a study of author contribution statements found observational evidence that menial or under-valued tasks are assigned to women while more prestigious tasks are assigned to men [@macaluso_2016].
+in terms of training and mentoring, allocation of work, and how one descibes the collaboration to other people (e.g.
+in conference presentations, on the lab website, or in the 'Author contributions' section of a paper).
+Experimental work suggests that unconscious bias causes people to undervalue women's research achievements [@Moss_2012], and a study of author contribution statements found observational evidence that menial or under-valued tasks are more often assigned to women while more prestigious tasks are assigned to men [@macaluso_2016].
 
@@
 We believe that the answers to these questions are closely related.
-For example, some of the homophily we observed might be caused by women seeking to avoid harassment or sexism from men [@jagsi_2016], which would clearly be concerning.
+For example, some of the homophily we observed might be caused by women seeking to avoid harassment or sexism from men [@jagsi_2016], which would clearly be very concerning.
 Additionally, Sheltzer and Smith [-@sheltzer_2014] concluded that 'elite' male academics (defined as recipients of major honours) have a higher proportion of male students and postdocs than non-elite male academics.
@@
 due to gender differences in confidence, or a perception that some groups are sexist).
-We also found a little evidence that gender homophily is detrimental to research quality, in that high-impact journals tended to have weaker homophily.
+We also found a little evidence that gender homophily is detrimental to research quality, in that high-impact journals tended to have weaker homophily (though the relationship was very noisy).
 Assuming that papers published in high-impact journals are of higher average quality (which is contentious; [@Garfield_2006]), our results provide non-experimental support for the hypothesis that mixed-gender teams produce better research than single-gender teams [@Britton_2000; @Reagans_2001; @Hong_2004; @Whittington_2008; @Bear_2011; @Herrera_2012; @Campbell_2013].
@@
 Collaboration is often most enjoyable and productive when working with like-minded people, who might tend to be same-gendered more often than not.
-We also suppose that some people consciously choose to preferentially collaborate with women in order to help close the gender gap in the workforce; this would create homophily if women do this more than men.
+We also suppose that some people consciously choose to preferentially collaborate with women in order to help close the gender gap in the workforce; this would create homophily if women adopt this strategy more often than men.
 In support of this interpretation, there is some evidence that women are more likely than men to promote the work of female colleagues by inviting them to give talks [@nittrouer_2018; @debarre_2018].
-Given that many collaborative research projects unfortunately involve a gendered division of labour [@macaluso_2016], working with a same-gendered colleague may provide exposure to new parts of the research process, and (especially for the minority gender) a welcome change of pace.
+Given that many collaborative research projects unfortunately involve a gendered division of labour [@macaluso_2016], working with a same-gendered colleague may provide exposure to new parts of the research process.
 
@@
 [-@Holman_2018].
-Briefly, that dataset was created by downloading every article indexed on PubMed and attempting to infer each author's gender from their given name using computational methods.
+Briefly, that dataset was created by downloading every article indexed on PubMed and attempting to infer gender from each authors' given name using computational methods.
 Each journal was assigned to one of 107 scientific disciplines, using PubMed's journal categorisations in the interests of objectivity.
@@
 Essentially, this issue arises because a person cannot be their own co-author.
-In a small dataset comprising $m$ men and $f$ women, a man can co-author with $m - 1$ men while women can co-author with $m$ men.
+In a small dataset comprising $m$ men and $f$ women authors, a man can co-author with $m - 1$ men while a woman can co-author with $m$ men.
 Thus, the null expectation for $\alpha$ is a negative number -- potentially a large one if $m$ and $f$ are very small.
@@
 
-As expected, $\alpha'$ was usually almost identical to $\alpha$ (S7 Fig), but $\alpha$ was downwardly biased relative to $\alpha'$ for small datasets (S8 Fig).
+As expected, $\alpha'$ was usually almost identical to $\alpha$ (S8 Fig), but $\alpha$ was downwardly biased relative to $\alpha'$ for the smallest datasets (S9 Fig).
 Additionally, the correlation between $\alpha'$ and sample size was negligible ($R^2 < 0.01$), suggesting that our calculation of $\alpha'$ effectively removed the dependence of $\alpha$ on sample size.
@@
 For example, Holman et al.
-[-@Holman_2018] found that PubMed-indexed authors based in Serbia are more than twice as likely to be women as are authors based in Japan.
+[-@Holman_2018] found that PubMed-indexed authors based in Serbia are far more likely to be women than are authors based in Japan.
 Therefore, a dataset containing a mix of papers from teams of authors based in these two countries would contain an excess of same-sex coauthorships, even if collaboration were random with respect to gender within each country.
@@
 Thus, journals with a positive standardised impact factor have a higher mean number of citations than the average for journals in their discipline.
-We then used Spearman rank correlation to test whether $\alpha'$ was correlated with impact factor across journals.
+We then used linear regression and Spearman rank correlation to examine the relationship between $\alpha'$ and impact factor across journals.
 
@@
 
-The relationship between the gender ratio of authors publishing in a journal and its $\alpha'$ value appeared nonlinear (see Results).
+The relationship between the gender ratio of authors publishing in a journal and its $\alpha'$ value appeared nonlinear (S4 Fig).
 We therefore fit a generalised additive model with thin plate regression spline smoothing, implemented using the ``mgcv`` package for R.
@@
 We hypothesised that this difference in gender ratio between career stages could potentially create both Wahlund effects and 'reverse' Wahlund effects.
-For example, imagine that the majority of collaborations in a particular field are between students and professors, and that the gender ratio differs between career stages: we would then see an excess of mixed-gender coauthorships (heterophily, $\alpha < 0$), even if gender has no direct, causal effect.
+For example, imagine that the majority of collaborations in a particular field are between students and professors, and that the gender ratio differs between career stages: we would then see an excess of mixed-gender coauthorships (heterophily, $\alpha < 0$), even if gender has no direct, causal effect on the choice of coauthors.
 Similarly, a hypothetical field in which students work only with students, and professors with professors, would have apparent gender homophily ($\alpha > 0$).
 
-We can think of no tractable method of controlling for this issue using our dataset, which contains no information on career stage.
+We can think of no tractable method of controlling for this issue using our dataset, which contains no reliable information on career stage.
 Therefore, we instead decided to derive theoretical expectations for $\alpha$ when there is a difference in gender ratio across career stages, in order to determine if and how this effect should alter our inferences.
@@
 Under the null model that gender has no causal effect on collaboration, we calculated $\alpha$ for various combinations of the four free parameters in our simple model.
-These parameters are: the gender ratio among early-career researchers (x-axis of Figure 6), the gender ratio among established researchers (y-axis of Figure 6), the frequency of within- versus between career stage collaborator pairs (rows in Figure 6), and lastly the frequency of within-stage collaborations that are between two early-career researchers as opposed to two late-career researchers (columns in Figure 6).
+These parameters are: the gender ratio among early-career researchers ($x$-axis of Fig 5), the gender ratio among established researchers ($y$-axis of Fig 5), the frequency of within- versus between career stage collaborator pairs (rows in Fig 5), and lastly the frequency of within-stage collaborations that are between two early-career researchers as opposed to two late-career researchers (columns in Fig 5).
 When these four parameters are specified, one can easily calculate the relative frequencies of collaborator pairs that involve two men, two women, or a man and a woman.
-In short, if we have specified the frequency of women at both career stages, as well as the frequency of the three possible types of collaboration with respect to career stage (early-early, early-established, and established-established), then we can calculate the frequency of collaborators pairs comprising two women, or a woman and a man, and thus find $\alpha$ (see the Online Supplementary Material for the annotated R code).
+In short, if we have specified the frequency of women at both career stages, as well as the frequency of the three possible types of collaboration with respect to career stage (early-early, early-established, and established-established), then we can calculate the frequency of collaborators pairs comprising two women, or a woman and a man, and thus calcualte $\alpha$ (see the Online Supplementary Material for the annotated R code).
 
-## Data availability and reproducibility
-The Online Supplementary Material contains R scripts used to produce all results, figures and tables; it can be viewed online at https://lukeholman.github.io/genderHomophily/.
+## Data and code availability
+The R scripts used to produce all results, figures and tables can be viewed online at https://lukeholman.github.io/genderHomophily/.
 The input data from Holman et al.
@@
 # Acknowledgements
-We thank Devi Stuart-Fox and Dominique Potvin for helpful discussion.
+We thank Devi Stuart-Fox and Dominique Potvin for their very helpful comments on the manuscript.
 
@@
 ## Supplementary figures
-S1 Fig. Plot showing the percentage of papers that have 1, 2, 3, 4, or ${\ge}5$ authors for each discipline in the dataset of Holman et al. (2018). This information can also be found in S3 Data.
+**S1 Fig**: Plot showing the percentage of papers that have 1, 2, 3, 4, or ${\ge}5$ authors for each discipline in the dataset of Holman et al. (2018). This information can also be found in S3 Data.
 
-S2 Fig.
-Histogram showing the distribution of differences in $\alpha'$ between the 2015-16 and 2005-6 samples, where positive numbers indicate an increase in $\alpha'$ with time.
+**S2 Fig**: Histogram showing the distribution of differences in $\alpha'$ between the 2015-16 and 2005-6 samples, where positive numbers indicate an increase in $\alpha'$ with time.
 The mean is slightly positive (i.e.
@@
 
-S3 Fig.
-Histogram showing the difference between $\alpha'$ calculated for first and last authors.
+**S3 Fig**: Histogram showing the difference between $\alpha'$ calculated for first and last authors.
 Positive values mean that $\alpha'$ was higher when calculated for first authors, and negative values mean $\alpha'$ was higher when calculated for last authors.
@@
 
-S4 Fig.
-Histogram of $\alpha'$ for 325 unique combinations of journal and country, using data from August 2015 - August 2016.
+**S4 Fig**: There is a weakly positive, non-linear relationship between the gender ratio of authors publishing in a journal, and the coefficient of homophily ($\alpha'$).
+Specifically, journals with 50% women authors or higher tended to have more same-sex coauthorships than did journals in which most authors are men.
+This relationship held whether $\alpha'$ was calculated for all authors or first authors only, but not for last authors only.
+A negative value on the x-axis denotes an excess of men authors, a positive value denotes an excess of women authors, and zero denotes gender parity (i.e.
+equal numbers of male and female authors).
+The lines were fitted using generalised additive models with the smoothing parameter $k$ set to 3.
+
+**S5 Fig**: Histogram of $\alpha'$ for 325 unique combinations of journal and country, using data from August 2015 - August 2016.
 The white areas denote combinations for which $\alpha'$ differs significantly from zero (p < 0.05, following false discovery rate correction).
 
-S5 Fig. Plot showing the 68 combinations of journal and author country of affiliation for which $\alpha'$ is significantly higher than expected.
+**S6 Fig**: Plot showing the 68 combinations of journal and author country of affiliation for which $\alpha'$ is significantly higher than expected.
 
-S6 Fig.
-Histogram showing the estimated degree to which $\alpha'$ is inflated by inter-country differences in author gender ratio, across the 285 journals for which we had adequate data after restricting the analysis by country.
+**S7 Fig**: Histogram showing the estimated degree to which $\alpha'$ is inflated by inter-country differences in author gender ratio, across the 285 journals for which we had adequate data after restricting the analysis by country.
 The average inflation in $\alpha'$ is negligible, suggesting that Wahlund effects resulting from inter-country differences have a negligible effect on our estimates of gender homophily.
 
-S7 Fig.
-There is a very strong correlation between the values of $\alpha$ and $\alpha'$ calculated for each journal, though in a handful of cases the difference is considerable.
-The deviation between $\alpha$ and $\alpha'$ is greatest for journals for which there is a small sample size (see S8 Fig).
+**S8 Fig**: There is a very strong correlation between the values of $\alpha$ and $\alpha'$ calculated for each journal, though in a handful of cases the difference is considerable.
+The deviation between $\alpha$ and $\alpha'$ is greatest for journals for which there is a small sample size (see S9 Fig).
 
-S8 Fig.
-For journals for which we recovered a small number of papers (<100), the unadjusted metric $\alpha$ was downwardly biased.
+**S9 Fig**: For journals for which we recovered a small number of papers (<100), the unadjusted metric $\alpha$ was downwardly biased.
 This fits our expectations: because researchers cannot be their own co-authors, small datasets will tend to produce negative estimates of $\alpha$ even if authors assort randomly with respect to gender (see main text).
@@
 ## Supplementary tables
-S1 Table. Sample sizes for the two datasets, which comprise papers published in the timeframes August 2005 - August 2006, and August 2015 - August 2016.
+**S1 Table**: Sample sizes for the two datasets, which comprise papers published in the timeframes August 2005 - August 2006, and August 2015 - August 2016.
 
-S2 Table.
-Number of journals showing statistically significant homophily or heterophily, in two one-year periods.
+**S2 Table**: Number of journals showing statistically significant homophily or heterophily, in two one-year periods.
 The significance threshold was p < 0.05, and p-values were adjusted using Benjamini-Hochberg false discovery rate correction.
@@
 ## Supplementary datasets
-S1 Data: This spreadsheet shows the $\alpha$ values calculated for each journal, in the 2005 and 2015 samples, and for each type of author (all authors, first authors, and last authors).
+**S1 Data**: This spreadsheet shows the $\alpha$ values calculated for each journal, in the 2005 and 2015 samples, and for each type of author (all authors, first authors, and last authors).
 The tables gives the impact factor of each journal, the sample size, $\alpha$ and $\alpha'$ and their 95% CIs, and the p-value from a 2-tailed test evaluating the null hypothesis that $\alpha$ is zero (both raw and FDR-corrected p-values are shown).
 
-S2 Data: This file gives the number and percentage of papers that have 1, 2, 3, 4, or ${\ge}5$ authors for each _journal_ in the dataset of Holman et al.
+**S2 Data**: This file gives the number and percentage of papers that have 1, 2, 3, 4, or ${\ge}5$ authors for each _journal_ in the dataset of Holman et al.
 (2018) _PLOS Biology_.
@@
 
-S3 Data: This file gives the number and percentage of papers that have 1, 2, 3, 4, or ${\ge}5$ authors for each _discipline_ in the dataset of Holman et al.
+**S3 Data**: This file gives the number and percentage of papers that have 1, 2, 3, 4, or ${\ge}5$ authors for each _discipline_ in the dataset of Holman et al.
 (2018) _PLOS Biology_.
@@
 
-S4 Data.
-The table shows the distribution of the $\alpha'$ values across journals, split by the research discipline.
+**S4 Data**: The table shows the distribution of the $\alpha'$ values across journals, split by the research discipline.
 The gender ratio column shows the percentage of women authors in the sample used to calculate $\alpha'$, across all authorship positions.

### 2019-04-11 manuscript/homophily_manuscript.tex
Response to 3rd plos review

@@
 hindered by a largely unacknowledged statistical issue that we will
-refer to as the Wahlund effect (\autoref{wahlund_plot}), by analogy with
-the conceptually similar Wahlund effect in population genetics {[}59{]}.
-The Wahlund effect makes it deceptively difficult to infer gender-based
-preferences simply by counting the number of same- and mixed-gender
-coauthorships. Essentially, whenever coauthorship data are sampled from
-two or more discrete sets of literature, which vary in the author gender
-ratio and which are largely not connected by collaboration, the number
-of same-gendered coauthors will be inflated. This can give the
-impression that authors preferentially publish with same-gendered
-colleagues even if no gender preferences exist, or if the true
-preference is for opposite-gendered colleagues (`gender heterophily').
-For example, a sample of literature containing a mixture of
-bioinformatics and cell biology papers will probably contain an excess
-of mostly-male and mostly-female author lists, simply because
+refer to as the Wahlund effect (Fig 1), by analogy with the conceptually
+similar Wahlund effect in population genetics {[}59{]}. The Wahlund
+effect makes it deceptively difficult to test for gender-based co-author
+choice simply by counting the relative number of same- and mixed-gender
+coauthorships. Essentially, the Wahlund effect means that whenever
+coauthorship data are sampled from two or more discrete sets of
+literature, which vary in the author gender ratio and which are largely
+unconnected by collaboration, the number of same-gendered coauthors will
+be inflated. This can give the impression that authors preferentially
+publish with same-gendered colleagues even if no gender preferences
+exist, or if the true preference is for opposite-gendered colleagues
+(`gender heterophily'). For example, a sample of literature containing a
+mixture of bioinformatics and cell biology papers will probably contain
+an excess of mostly-male and mostly-female author lists, simply because
 researchers usually collaborate within their own discipline, and because
@@
 
-\begin{figure}
-\centering
-\includegraphics{../figures/Fig1.pdf}
-\caption{The Wahlund effect can make it appear as if authors publish
-with same-gendered colleagues disproportionately often, even if
+\vspace{2cm}
+
+\textbf{Fig 1}: The Wahlund effect can make it appear as if authors
+publish with same-gendered colleagues disproportionately often, even if
 collaboration is completely random with respect to gender. Here,
@@
 gender ratios; these subsets could be research disciplines, older versus
-newer papers, or papers from authors in different countries. The example
-countries and disciplines were selected based on data in {[}5{]}.
-\label{wahlund_plot}}
-\end{figure}
+newer papers, or papers from authors based in different countries. The
+example countries and disciplines were selected based on data in
+{[}5{]}. \vspace{2cm}
 
@@
 co-publish with same-gendered colleagues, while controlling for the
-Wahlund effect as strictly as possible. We use a recently-published
+Wahlund effect as strictly as we are able. We use a recently-published
 dataset describing the gender of 35.5 million authors from 9.15 million
@@
 for gender homophily while restricting our analysis to particular
-journals (i.e.~research specialties), time periods, and countries. We
-quantified gender assortment using a metric called \(\alpha'\) {[}60{]},
-which is positive when same-gender authors publish together more often
-than expected (gender homophily), negative when opposite-gender authors
-publish together more often than expected (heterophily), and equal to
-zero when authors assort randomly with respect to gender (see Methods).
+journals (a proxy for research specialties), time periods, and
+countries. We quantified gender assortment using a metric called
+\(\alpha'\) {[}60{]}, which is positive when same-gender authors publish
+together more often than expected (gender homophily), negative when
+opposite-gender authors publish together more often than expected
+(heterophily), and equal to zero when authors assort randomly with
+respect to gender (see Methods).
 
@@
 
-\autoref{alpha_histograms} shows the distribution of \(\alpha'\)
-estimates in 2015-2016 across all journals for which we recovered
-sufficient data, when \(\alpha'\) was calculated for all authors, first
-authors only, or last authors only. Most journals had positive values of
-\(\alpha'\) (77-92\%, depending on time period and author type; S1
-Data), and for many of these the false discovery rate (FDR)-corrected
-p-values suggested that \(\alpha'\) was significantly greater than zero
-(1469/2077 journals were significant in 2015-16, and 404/1192 in 2005-6;
-S1 Data). Only 2/2077 journals had statistically significant heterophily
-(i.e. \(\alpha' < 0\)) in 2015-16, and 1/1192 in 2005-6 (S2 Table). The
+Fig 2 shows the distribution of \(\alpha'\) estimates in 2015-2016
+across all journals for which we recovered sufficient data, when
+\(\alpha'\) was calculated for all authors, first authors only, or last
+authors only. Most journals had positive values of \(\alpha'\) (77-92\%,
+depending on time period and author type; S1 Data), and for many of
+these the false discovery rate (FDR)-corrected p-values suggested that
+\(\alpha'\) was significantly greater than zero (1469/2077 journals were
+significant in 2015-16, and 404/1192 in 2005-6; S1 Data). Only 2/2077
+journals had statistically significant heterophily (i.e.
+\(\alpha' < 0\)) in 2015-16, and 1/1192 in 2005-6 (S2 Table). The
 remaining 606 or 787 journals (in 2015 and 2005 respectively) had a
@@
 
-\begin{figure}
-\centering
-\includegraphics{../figures/Fig2.pdf}
-\caption{Of the 2116 journals for which we had adequate data in
+\vspace{2cm}
+
+\textbf{Fig 2}: Of the 2116 journals for which we had adequate data in
 2015-2016, 825 showed statistically significant evidence of gender
@@
 first authors only, or for last authors only. Points in the right panel
-show \(\alpha'\) for individual journals.\label{alpha_histograms}}
-\end{figure}
+show \(\alpha'\) for individual journals. \vspace{2cm}
 
@@
 
-\autoref{alpha_histograms} illustrates the variance in journal homophily
-values (\(\alpha'\)) across scientific disciplines. All disciplines had
-positive mean \(\alpha'\) (averaged over journals), although homophily
-appeared somewhat stronger in some disciplines than others (e.g.~mean
-\(\alpha'\) was \(0.12{\pm}0.02\) for Urology journals and
-\(0.03{\pm}0.01\) for Veterinary Medicine journals;
-\autoref{alpha_histograms}, S4 Data). However, there was no formal
-evidence for consistent differences in \(\alpha'\) between disciplines:
-the random factor `Discipline' explained around 1\% of the variance in
-\(\alpha'\) in the two linear mixed models described in the previous
-section (see \autoref{alpha_histograms} and mixed models in Online
-Supplementary Material). Thus, whatever processes are responsible for
-producing positive \(\alpha'\) values appear to be similarly strong in
-all the disciplines we examined.
+Fig 2 illustrates the variance in journal homophily values (\(\alpha'\))
+across scientific disciplines. All disciplines had positive mean
+\(\alpha'\) (averaged over journals), although homophily appeared
+somewhat stronger in some disciplines than others (e.g.~mean \(\alpha'\)
+was \(0.12{\pm}0.02\) for Urology journals and \(0.03{\pm}0.01\) for
+Veterinary Medicine journals; Fig 2, S4 Data). However, there was no
+formal evidence for consistent differences in \(\alpha'\) between
+disciplines: the random factor `Discipline' explained around 1\% of the
+variance in \(\alpha'\) in the two linear mixed models described in the
+previous section (see Fig 2 and mixed models in Online Supplementary
+Material). Thus, the causal mechanisms underlying the observed positive
+\(\alpha'\) values appear to be similarly strong in all the disciplines
+we examined.
 
 There was no indication that journals publishing on a wide range of
-topics have higher \(\alpha'\) values than more specialised journals,
-due to the Wahlund effect. For example, the journal category
-`Multidisciplinary' -- which includes journals like \emph{PLOS ONE},
-\emph{Nature}, \emph{Science}, and \emph{PNAS} -- did not have markedly
-elevated \(\alpha'\) (\autoref{alpha_histograms}). This result suggests
+topics have higher \(\alpha'\) values than more specialised journals due
+to the Wahlund effect (Fig 1). For example, the journal category
+`Multidisciplinary' -- which includes general interest journals like
+\emph{PLOS ONE}, \emph{Nature}, \emph{Science}, and \emph{PNAS} -- did
+not have markedly elevated \(\alpha'\) (Fig 2). This result suggests
 that our estimates of homophily, and estimates from some of the earlier
-studies listed in the Introduction, are probably not markedly inflated
-by the presence of disparate research topics (with variable author
-gender ratios) being published within individual journals.
+studies of homophily listed in the Introduction, are probably not
+markedly inflated by the presence of disparate research topics (with
+variable author gender ratios) being published within individual
+journals.
 
@@
 \(\alpha\) was 0.126. This figure is almost double the median value of
-\(\alpha'\) for individual journals (Figure 2; \(\alpha' = 0.070\) for
-`All authors'), suggesting that lumping together papers from different
-fields and different time periods can indeed produce spurious evidence
-for gender homophily as outlined in Figure 1.
+\(\alpha'\) for individual journals (Fig 2; \(\alpha' = 0.070\) for `All
+authors'), suggesting that lumping together papers from different fields
+and different time periods can indeed produce spurious evidence for
+gender homophily as outlined in Fig 1.
 
@@
 Papers with two authors had significantly lower (but still positive)
-\(\alpha'\) values relative to papers with more than two authors, on
-average (\autoref{author_number}; statistical results in Online
-Supplementary Material). Papers with 3, 4 or \(\ge5\) authors had
-essentially identical average \(\alpha'\) values. The variance in
-\(\alpha'\) across journals was also a little higher for 2-author papers
-compared to the remainder (\autoref{author_number}), though part of this
-variance is due to the reduced sample size (in terms of number of
-authors) for the 2-author papers. One possible explanation for this
-finding is that 2-authors papers might be more likely to have an author
-list that is evenly split between career stages (e.g.~a postgraduate
-student and their supervisor), increasing the chance that the authors
-are mixed gender (see Figure 6). The result also suggests that the
-processes responsible for gender homophily are similar in small
-(e.g.~3-author) and larger (5+ author) collaborations (and across
-disciplines where small versus large collaborations are the norm).
-
-\begin{figure}
-\centering
-\includegraphics{../figures/Fig3.pdf}
-\caption{The coefficient of homophily (\(\alpha'\)) was slightly less
-positive when calculated for two-author papers only, relative to papers
-with longer author lists. The individual points, whose distribution is
-summarised by the violin plots, correspond to individual journals. The
-larger white points show the mean for each group (and its 95\% CIs), as
-calculated by a Bayesian meta-regression model accounting for repeated
-measures of \(\alpha'\) within journals, as well as the precision with
-which \(\alpha'\) was estimated. \label{author_number}}
-\end{figure}
+\(\alpha'\) values relative to papers with more than two authors, while
+papers with 3, 4 or \(\ge5\) authors had essentially identical average
+\(\alpha'\) values (Fig 3). Specifically, the posterior estimate of mean
+\(\alpha'\) was 0.014 (95\% CIs: 0.002 - 0.026) for 2-author papers and
+0.065 (95\% CIs: 0.056 - 0.074) for 3-author papers (and roughly the
+same for 4- and \(\ge5\)-author papers; Fig 3). One possible explanation
+for this finding is that 2-authors papers are more likely to have an
+author list that is evenly split between career stages (e.g.~a
+postgraduate student and their supervisor), increasing the chance that
+the authors are mixed gender (see section `Theoretical expectations for
+\(\alpha\) when the gender ratio differs between career stages'). The
+result also suggests that the causal mechanisms responsible for gender
+homophily are similar in small (e.g.~3-author) and larger (\(\ge5\)
+author) collaborations (and across disciplines where small versus large
+collaborations are the norm).
+
+\vspace{2cm}
+
+\textbf{Fig 3}: The coefficient of homophily (\(\alpha'\)) was slightly
+less positive when calculated for two-author papers only, relative to
+papers with longer author lists. The individual points, whose
+distribution is summarised by the violin plots, correspond to individual
+journals. The larger white points show the mean for each group (and its
+95\% CIs), as calculated by a Bayesian meta-regression model accounting
+for repeated measures of \(\alpha'\) within journals, as well as the
+precision with which \(\alpha'\) was estimated. \vspace{2cm}
 
@@
 gender-balanced workforce (e.g.~Psychiatry). We found a positive,
-non-linear relationship between the overall gender ratio of all authors
+non-linear relationship between the gender ratio across all the authors
 publishing in a particular journal {[}5{]}, and the estimated value of
-\(\alpha'\) for all authors and for first authors
-(\autoref{alpha_gender_ratio}). Journals with a balanced or
-female-biased author gender ratio tended to have higher \(\alpha'\)
-(i.e.~stronger homophily) than journals with a male-biased author gender
-ratio (GAM smooth terms p \textless{} 0.001; Online Supplementary
-Material). The relationship was not statistically significant when
-\(\alpha'\) was calculated for last authors (GAM, p = 0.142), though the
-trend appeared similar (\autoref{alpha_gender_ratio}).
-
-\begin{figure}
-\centering
-\includegraphics{../figures/Fig4.pdf}
-\caption{There is a weakly positive, non-linear relationship between the
-gender ratio of authors publishing in a journal, and the coefficient of
-homophily (\(\alpha'\)). Specifically, journals with 50\% women authors
-or higher tended to have more same-sex coauthorships than did journals
-with predominantly men authors. This relationship held whether
-\(\alpha'\) was calculated for all authors, first authors only, or last
-authors only. A negative value on the x-axis denotes an excess of men
-authors, a positive value denotes an excess of women authors, and zero
-denotes gender parity (i.e.~equal numbers of men and women). The lines
-were fitted using generalised additive models with the smoothing
-parameter \(k\) set to 3. \label{alpha_gender_ratio}}
-\end{figure}
+\(\alpha'\) for all authors and for first authors, but not last authors
+(S4 Fig). Journals with a balanced or female-biased author gender ratio
+tended to have higher \(\alpha'\) (i.e.~stronger homophily) than
+journals with a male-biased author gender ratio (GAM smooth term p =
+0.0002 for all-author homophily, p \textless{} 0.0001 for first-author
+homophily, and p = 0.13 for last-author homophily).
 
@@
 gender homophily than did journals with a low impact factor for their
-discipline (\autoref{impact_factor}; linear regression: \(R^2\) = 0.043,
-\(t_{1415}\) = -8.0, p \textless{} 0.0001). The slope of the regression
-was \(-0.012{\pm}0.0015\), indicating that increasing the
+discipline (Fig 4; linear regression: \(R^2\) = 0.043, \(t_{1415}\) =
+-8.0, p \textless{} 0.0001). The slope of the regression was
+\(-0.012{\pm}0.0015\), indicating that increasing the
 discipline-standardised impact factor by one standard deviation is
-associated with a reduction in \(\alpha'\) of 0.012.
-
-\begin{figure}
-\centering
-\includegraphics{../figures/Fig5.pdf}
-\caption{Journal impact factor (expressed relative to the average for
-the discipline) is negatively correlated with \(\alpha'\). The
-relationship is noisy (\(R^2\) = 0.043), but the results suggest that
-journals with strong homophily tend to have lower impact factors than
-journals with weak homophily in the same discipline.
-\label{impact_factor}}
-\end{figure}
+associated with a reduction in \(\alpha'\) of 0.012. The Spearman
+correlation coefficient was -0.19 (p \textless{} 0.0001).
+
+\vspace{2cm}
+
+\textbf{Fig 4}: Journal impact factor (expressed relative to the average
+for the discipline) is negatively correlated with \(\alpha'\). The
+relationship is noisy (\(R^2\) = 0.043, Spearman correlation = -0.19),
+but the results indicate that journals with strong gender homophily tend
+to have lower impact factors than journals from the same discipline that
+have weaker homophily. \vspace{2cm}
 
@@
 tested (64 unique journals and 18 unique countries), and no significant
-heterophily (S4-S5 Fig). Additionally, the values of \(\alpha'\)
+heterophily (S5-S6 Fig). Additionally, the values of \(\alpha'\)
 calculated for each journal-country combination were only very slightly
@@
 (i.e.~when pooling papers from different countries, as was done to make
-\autoref{alpha_histograms}): on average, the difference in \(\alpha'\)
-was only 0.002 (S6 Fig). These results suggest that our findings of
-widespread homophily in the main analysis were not driven solely by a
-Wahlund effect resulting from gender differences between countries.
+Fig 2): on average, the difference in \(\alpha'\) was only 0.002 (S7
+Fig). These results suggest that our findings of widespread homophily in
+the main analysis were not driven solely by a Wahlund effect resulting
+from gender differences between countries.
 
@@
 for \(\alpha\) when the gender ratio differs between career stages (see
-Methods). As shown in \autoref{simulation}, we predict that \(\alpha\)
-is expected to be non-zero, even if collaborators are randomly selected
-with respect to gender, provided that there is a gender gap between
-career stages. The extent to which \(\alpha\) deviates from zero depends
-on the relative frequencies of collaboration within and between career
-stages (rows and columns in Figure 6), and the size of the gender gap
-between stages (x- and y-axes in Figure 6). When \textgreater{}50\% of
-coauthor pairs comprise one early-career and one established researcher,
-we expect gender heterophily (\(\alpha < 0\)) whenever the gender ratio
-differs between career stages. Conversely, when \textgreater{}50\% of
+Methods). As shown in Fig 5, we predict that \(\alpha\) is expected to
+be non-zero, even if collaborators are randomly selected with respect to
+gender, provided that there is a gender gap between career stages. The
+extent to which \(\alpha\) deviates from zero depends on the relative
+frequencies of collaboration within and between career stages (rows and
+columns in Fig 5), and the size of the gender gap between stages (\(x\)-
+and \(y\)-axes in Fig 5). When \textgreater{}50\% of coauthor pairs
+comprise one early-career and one established researcher, we expect
+gender heterophily (\(\alpha < 0\)) whenever the gender ratio differs
+between career stages. Conversely, when \textgreater{}50\% of
 collaborations are between people at the same career stage, we expect
 gender homophily (\(\alpha > 0\)). In a few parameter spaces (shown in
-red; \autoref{simulation}), \(\alpha\) was quite high, and overlapped
-with the values that we estimated (\autoref{alpha_histograms}).
-
-\begin{figure}
-\centering
-\includegraphics{../figures/Fig6_inkscape.pdf}
-\caption{When the gender ratio of early-career researchers is not equal
-to the gender ratio among established researchers, the null expectation
-for \(\alpha\) is not necessarily zero. Specifically, if most
-collaborations occur between career stages, there will be an excess of
-mixed-gender collaborations (\(\alpha < 0\), blue areas), while if most
-collaborator pairs comprise two people at the same career stage, there
-will be an excess of same-gender collaborations (\(\alpha > 0\), red
-areas). However, the conditions required for strong gender homophily
+red; Fig 5), \(\alpha\) was quite high, and overlapped with the values
+that we estimated (Fig 2).
+
+\vspace{2cm}
+
+\textbf{Fig 5}: When the gender ratio of early-career researchers is not
+equal to the gender ratio among established researchers, the null
+expectation for \(\alpha\) is not necessarily zero. Specifically, if
+most collaborations occur between career stages, there will be an excess
+of mixed-gender collaborations (\(\alpha < 0\), blue areas), while if
+most collaborator pairs comprise two people at the same career stage,
+there will be an excess of same-gender collaborations (\(\alpha > 0\),
+red areas). However, the conditions required for strong gender homophily
 (i.e.~the red areas) are quite restrictive, making it unlikely that this
@@
 established researchers (i.e.~the blue areas), our study will
-underestimate the strength of gender homophily. Contour lines denote
-increments of 0.1. \label{simulation}}
-\end{figure}
-
-Despite this overlap, \autoref{simulation} suggests that our main
-conclusions (and those of other studies of gender homophily) are
-probably robust to this career stage issue. We only expect strongly
-positive \(\alpha\) when A) the gender ratio is highly skewed across
-career stages (e.g.~a 5-fold difference), and B) collaborations between
-early and established researchers are very rare (e.g. \textless{}10\% of
-the total). Both of these conditions seem unlikely to be true for most
-fields: the gender gap across careers stages is generally less
-pronounced {[}1,5{]}, and it is very common for early-career researchers
-to co-publish with an established mentor {[}61{]}. However, one can get
-\(\alpha > 0\) for realistic combinations of parameters, e.g.~a moderate
-shortage of women in senior positions coupled with a moderate excess of
-within-career stage collaboration, suggesting this effect might
-contribute to some of the homophily observed by this and previous
-studies.
+underestimate the strength of gender homophily. Contour lines mark
+increments of 0.1. \vspace{2cm}
+
+Despite this overlap, Fig 5 suggests that our main conclusions (and
+those of other studies of gender homophily) are probably robust to this
+career stage issue. We only expect strongly positive \(\alpha\) when A)
+the gender ratio is highly skewed across career stages (e.g.~a 5-fold
+difference), and B) collaborations between early and established
+researchers are very rare (e.g. \textless{}10\% of the total). Both of
+these conditions seem unlikely to be true for most fields: the gender
+gap across careers stages is generally less pronounced {[}1,5{]}, and it
+is very common for early-career researchers to co-publish with an
+established mentor {[}61{]}. However, one can get \(\alpha > 0\) for
+realistic combinations of parameters, e.g.~a moderate shortage of women
+in senior positions coupled with a moderate excess of within-career
+stage collaboration, suggesting this effect might contribute to some of
+the homophily observed by this and previous studies.
 
@@
 \textgreater{}50\% of the total, then the baseline expectation for
-\(\alpha\) is actually less than zero (blue areas in
-\autoref{simulation}). Therefore, it is possible that researchers
-preferentially assort with same-gendered collaborators even more
-strongly than implied by our results, at least for certain journals or
-research disciplines.
+\(\alpha\) is actually less than zero (blue areas in Fig 5). Therefore,
+it is possible that researchers preferentially assort with same-gendered
+collaborators even more strongly than implied by our results, at least
+for certain journals or research disciplines.
 
@@
 more often than expected under the null model, even after implementing
-stringent controls for Wahlund effects (\autoref{wahlund_plot}). Our
-study therefore reaffirms earlier studies' conclusions {[}49--57,62{]}
-using stricter methodology, and generalises their results across the
-life sciences. Relatively few journals had \(\alpha'\) values below
-zero, and almost no journals showed statistically significant gender
-heterophily after controlling for multiple testing. The excess of
-same-gender coauthorships was quite large: many journals had
-\(\alpha' > 0.1\), indicating that the gender ratio of men's and women's
-coauthors differs by \textgreater{}10\% in absolute terms. In relative
-terms, our findings are even more striking: for example, if men have
-20\% female coauthors and women have 30\% (i.e. \(\alpha' = 0.1\) in a
-field with a typical gender ratio {[}5{]}), then women publish with
-women 50\% more often than men do.
+stringent controls for Wahlund effects of the kind illustrated in Fig 1.
+Our study therefore reaffirms earlier studies' conclusions
+{[}49--57,62{]} using stricter methodology, and generalises their
+results across the life sciences. Relatively few journals had
+\(\alpha'\) values below zero, and almost no journals showed
+statistically significant gender heterophily after controlling for
+multiple testing. The excess of same-gender coauthorships was quite
+large: many journals had \(\alpha' > 0.1\), indicating that the gender
+ratio of men's and women's coauthors differs by \textgreater{}10\% in
+absolute terms. In relative terms, our findings are even more striking:
+for example, if men have 20\% female coauthors and women have 30\% (i.e.
+\(\alpha' = 0.1\) in a field with a typical gender ratio {[}5{]}), then
+women publish with women 50\% more often than men do.
 
@@
 coauthorships. As well as the obvious interpretation -- conscious or
-unconscious selection of same-gender collaborators by men, by women, or
-both -- our results could be partly explained by uncontrolled Wahlund
-effects. However, we suspect the contribution of these uncontrolled
-artefacts to be minor, for four reasons: we found positive \(\alpha'\)
-after controlling for three obvious sources of Wahlund effect; there was
-no inflation of \(\alpha'\) in highly multidisciplinary journals
-relative to specialised journals; restricting the data by country
-yielded similar estimates of \(\alpha'\); and we used a simple model to
-show that differences in gender ratio between career stages are unlikely
-to fully explain our results. On balance, we believe the data suggest
-that it is likely that some researchers preferentially select
-same-gendered collaborators, although it is difficult to ascertain what
-proportion of people show such a preference, or how much the strength of
-the preference varies between individuals. We also note that even in a
-world in which everyone selected their collaborators at random with
-respect to gender, a high proportion of individual researchers would
-have entirely same-gendered collaborators by chance alone (especially in
-gender-biased disciplines); thus, individuals who only have
-same-gendered co-authors are not necessarily doing anything differently
-from people with gender-balanced co-authors.
+unconscious selection of same-gendered collaborators by men, women, or
+both genders -- our results could be partly explained by uncontrolled
+Wahlund effects. However, we suspect the contribution of these
+uncontrolled artefacts to be minor, for four reasons: we found positive
+\(\alpha'\) after controlling for three obvious sources of Wahlund
+effect; there was no inflation of \(\alpha'\) in highly
+multidisciplinary journals relative to specialised journals; restricting
+the data by country yielded similar estimates of \(\alpha'\); and our
+modelling work suggested that differences in gender ratio between career
+stages are unlikely to fully explain our results. On balance, we believe
+the data suggest that it is likely that some researchers preferentially
+select same-gendered collaborators, although it is difficult to
+ascertain what proportion of people show such a preference, or how much
+the strength of the preference varies between individual researchers. We
+also note that even in a world in which everyone selected their
+collaborators at random with respect to gender, a high proportion of
+individual researchers would have entirely same-gendered collaborators
+by chance alone (especially in gender-biased disciplines); thus,
+individuals who only have same-gendered co-authors are not necessarily
+doing anything differently from people with gender-balanced co-authors.
 
@@
 might show the strongest gender homophily, e.g.~because being in the
-minority might increase people's motivation to seek out same-gendered
+minority might increase one's motivation to seek out same-gendered
 colleagues. Contrary to this hypothesis, we found no evidence that
 gender homophily is restricted to particular disciplines: \(\alpha'\)
-was similarly high across the board (\autoref{alpha_histograms}).
-Interestingly, gender homophily was weakest for journals with a
-male-biased author gender ratio, and strongest in journals with a
-female-biased author gender ratio. One possible reason is that men are
-more likely to preferentially seek out male collaborators in fields
-where men are a minority, relative to the homophily displayed by women
-in fields where women are a minority. However, this latter result only
-has tentative statistical support since our sample contains few journals
-in which most authors are women (\autoref{alpha_gender_ratio}).
+was similarly high across the board (Fig 2). Interestingly, gender
+homophily was weakest for journals with a male-biased author gender
+ratio, and strongest in journals with a female-biased author gender
+ratio. One possible reason is that men are more likely to preferentially
+seek out male collaborators in fields where men are a minority, relative
+to the homophily displayed by women in fields where women are a
+minority. However, this latter result only has tentative statistical
+support since our sample contains few journals in which most authors are
+women (S4 Fig).
 
@@
 trend might result from the increasing number of women working in senior
-positions in STEMM over the past decade {[}63--65{]}. As shown in
-\autoref{simulation}, if enough coauthorships are between junior and
-senior researchers, a large gender gap between career stages can give
-the appearance of heterophily. As this gender gap between career stages
-lessens, the observed values of \(\alpha'\) may increase.
+positions in STEMM over the past decade {[}63--65{]}. As shown in Fig 5,
+if enough coauthorships are between junior and senior researchers, a
+large gender gap between career stages can give the appearance of
+heterophily. As this gender gap between career stages lessens, the
+observed values of \(\alpha'\) may increase.
 
@@
 other women, e.g.~because this constrains women's options (particularly
-in fields like engineering, where 90\% of professors are men {[}1{]}).
-Instead, we suggest that researchers of both genders can help to close
-the gender gap in STEMM. In the context of collaboration, one way to do
-this is to undertake self-examination to ensure that one is not
-inadvertently overlooking or excluding women among potential students
-and colleagues. One should also take care to treat male and female
-collaborators equally, e.g.~in terms of training and mentoring,
-allocation of work, how one frames or promotes the collaboration
-(e.g.~in conference presentations or on a website), and in terms of the
-credit afforded to collaborators' contributions. Experimental work
-suggests that unconscious bias causes people to undervalue women's
-research achievements {[}20{]}, and a study of author contribution
-statements found observational evidence that menial or under-valued
-tasks are assigned to women while more prestigious tasks are assigned to
-men {[}61{]}.
+in fields like the one studied by Ghiasi et al. -- engineering -- where
+90\% of professors are men {[}1{]}). Instead, we suggest that
+researchers of both genders can help to close the gender gap in STEMM.
+In the context of collaboration, one way to do this is to undertake
+self-examination to ensure that one is not inadvertently overlooking or
+excluding women among potential students and colleagues. One should also
+take care to treat male and female collaborators equally, e.g.~in terms
+of training and mentoring, allocation of work, and how one descibes the
+collaboration to other people (e.g.~in conference presentations, on the
+lab website, or in the `Author contributions' section of a paper).
+Experimental work suggests that unconscious bias causes people to
+undervalue women's research achievements {[}20{]}, and a study of author
+contribution statements found observational evidence that menial or
+under-valued tasks are more often assigned to women while more
+prestigious tasks are assigned to men {[}61{]}.
 
@@
 we observed might be caused by women seeking to avoid harassment or
-sexism from men {[}38{]}, which would clearly be concerning.
+sexism from men {[}38{]}, which would clearly be very concerning.
 Additionally, Sheltzer and Smith {[}66{]} concluded that `elite' male
@@
 homophily is detrimental to research quality, in that high-impact
-journals tended to have weaker homophily. Assuming that papers published
-in high-impact journals are of higher average quality (which is
-contentious; {[}67{]}), our results provide non-experimental support for
-the hypothesis that mixed-gender teams produce better research than
-single-gender teams {[}42--48{]}. Another issue is that if many
-collaborations are between established researchers, there will be an
-excess of male-male collaborations in fields where women in senior
-positions are rare; some of the observed homophily might therefore
-reflect the elevated gender gap among senior researchers.
+journals tended to have weaker homophily (though the relationship was
+very noisy). Assuming that papers published in high-impact journals are
+of higher average quality (which is contentious; {[}67{]}), our results
+provide non-experimental support for the hypothesis that mixed-gender
+teams produce better research than single-gender teams {[}42--48{]}.
+Another issue is that if many collaborations are between established
+researchers, there will be an excess of male-male collaborations in
+fields where women in senior positions are rare; some of the observed
+homophily might therefore reflect the elevated gender gap among senior
+researchers.
 
@@
 preferentially collaborate with women in order to help close the gender
-gap in the workforce; this would create homophily if women do this more
-than men. In support of this interpretation, there is some evidence that
-women are more likely than men to promote the work of female colleagues
-by inviting them to give talks {[}68,69{]}. Given that many
-collaborative research projects unfortunately involve a gendered
-division of labour {[}61{]}, working with a same-gendered colleague may
-provide exposure to new parts of the research process, and (especially
-for the minority gender) a welcome change of pace.
+gap in the workforce; this would create homophily if women adopt this
+strategy more often than men. In support of this interpretation, there
+is some evidence that women are more likely than men to promote the work
+of female colleagues by inviting them to give talks {[}68,69{]}. Given
+that many collaborative research projects unfortunately involve a
+gendered division of labour {[}61{]}, working with a same-gendered
+colleague may provide exposure to new parts of the research process.
 
@@
 Briefly, that dataset was created by downloading every article indexed
-on PubMed and attempting to infer each author's gender from their given
-name using computational methods. Each journal was assigned to one of
-107 scientific disciplines, using PubMed's journal categorisations in
-the interests of objectivity. Because the present study focuses on
+on PubMed and attempting to infer gender from each authors' given name
+using computational methods. Each journal was assigned to one of 107
+scientific disciplines, using PubMed's journal categorisations in the
+interests of objectivity. Because the present study focuses on
 co-authorship, all single-author papers were discarded. We also
@@
 their own co-author. In a small dataset comprising \(m\) men and \(f\)
-women, a man can co-author with \(m - 1\) men while women can co-author
-with \(m\) men. Thus, the null expectation for \(\alpha\) is a negative
-number -- potentially a large one if \(m\) and \(f\) are very small.
+women authors, a man can co-author with \(m - 1\) men while a woman can
+co-author with \(m\) men. Thus, the null expectation for \(\alpha\) is a
+negative number -- potentially a large one if \(m\) and \(f\) are very
+small.
 
@@
 
-As expected, \(\alpha'\) was usually almost identical to \(\alpha\) (S7
+As expected, \(\alpha'\) was usually almost identical to \(\alpha\) (S8
 Fig), but \(\alpha\) was downwardly biased relative to \(\alpha'\) for
-small datasets (S8 Fig). Additionally, the correlation between
+the smallest datasets (S9 Fig). Additionally, the correlation between
 \(\alpha'\) and sample size was negligible (\(R^2 < 0.01\)), suggesting
@@
 ratio of researchers from different countries. For example, Holman et
-al. {[}5{]} found that PubMed-indexed authors based in Serbia are more
-than twice as likely to be women as are authors based in Japan.
-Therefore, a dataset containing a mix of papers from teams of authors
-based in these two countries would contain an excess of same-sex
-coauthorships, even if collaboration were random with respect to gender
-within each country.
+al. {[}5{]} found that PubMed-indexed authors based in Serbia are far
+more likely to be women than are authors based in Japan. Therefore, a
+dataset containing a mix of papers from teams of authors based in these
+two countries would contain an excess of same-sex coauthorships, even if
+collaboration were random with respect to gender within each country.
 
@@
 standardised impact factor have a higher mean number of citations than
-the average for journals in their discipline. We then used Spearman rank
-correlation to test whether \(\alpha'\) was correlated with impact
-factor across journals.
+the average for journals in their discipline. We then used linear
+regression and Spearman rank correlation to examine the relationship
+between \(\alpha'\) and impact factor across journals.
 
@@
 The relationship between the gender ratio of authors publishing in a
-journal and its \(\alpha'\) value appeared nonlinear (see Results). We
+journal and its \(\alpha'\) value appeared nonlinear (S4 Fig). We
 therefore fit a generalised additive model with thin plate regression
@@
 coauthorships (heterophily, \(\alpha < 0\)), even if gender has no
-direct, causal effect. Similarly, a hypothetical field in which students
-work only with students, and professors with professors, would have
-apparent gender homophily (\(\alpha > 0\)).
+direct, causal effect on the choice of coauthors. Similarly, a
+hypothetical field in which students work only with students, and
+professors with professors, would have apparent gender homophily
+(\(\alpha > 0\)).
 
 We can think of no tractable method of controlling for this issue using
-our dataset, which contains no information on career stage. Therefore,
-we instead decided to derive theoretical expectations for \(\alpha\)
-when there is a difference in gender ratio across career stages, in
-order to determine if and how this effect should alter our inferences.
-For simplicity, our calculations assume there are only two career stages
-(`early-career' and `established'), though we expect that the general
-conclusions would also apply to a multi-tier career ladder. Under the
-null model that gender has no causal effect on collaboration, we
-calculated \(\alpha\) for various combinations of the four free
+our dataset, which contains no reliable information on career stage.
+Therefore, we instead decided to derive theoretical expectations for
+\(\alpha\) when there is a difference in gender ratio across career
+stages, in order to determine if and how this effect should alter our
+inferences. For simplicity, our calculations assume there are only two
+career stages (`early-career' and `established'), though we expect that
+the general conclusions would also apply to a multi-tier career ladder.
+Under the null model that gender has no causal effect on collaboration,
+we calculated \(\alpha\) for various combinations of the four free
 parameters in our simple model. These parameters are: the gender ratio
-among early-career researchers (x-axis of Figure 6), the gender ratio
-among established researchers (y-axis of Figure 6), the frequency of
-within- versus between career stage collaborator pairs (rows in Figure
-6), and lastly the frequency of within-stage collaborations that are
-between two early-career researchers as opposed to two late-career
-researchers (columns in Figure 6). When these four parameters are
-specified, one can easily calculate the relative frequencies of
-collaborator pairs that involve two men, two women, or a man and a
-woman. In short, if we have specified the frequency of women at both
-career stages, as well as the frequency of the three possible types of
-collaboration with respect to career stage (early-early,
-early-established, and established-established), then we can calculate
-the frequency of collaborators pairs comprising two women, or a woman
-and a man, and thus find \(\alpha\) (see the Online Supplementary
-Material for the annotated R code).
-
-\hypertarget{data-availability-and-reproducibility}{%
-\subsection{Data availability and
-reproducibility}\label{data-availability-and-reproducibility}}
-
-The Online Supplementary Material contains R scripts used to produce all
-results, figures and tables; it can be viewed online at
-\url{https://lukeholman.github.io/genderHomophily/}. The input data from
-Holman et al. {[}5{]} is archived at \url{https://osf.io/bt9ya/}.
+among early-career researchers (\(x\)-axis of Fig 5), the gender ratio
+among established researchers (\(y\)-axis of Fig 5), the frequency of
+within- versus between career stage collaborator pairs (rows in Fig 5),
+and lastly the frequency of within-stage collaborations that are between
+two early-career researchers as opposed to two late-career researchers
+(columns in Fig 5). When these four parameters are specified, one can
+easily calculate the relative frequencies of collaborator pairs that
+involve two men, two women, or a man and a woman. In short, if we have
+specified the frequency of women at both career stages, as well as the
+frequency of the three possible types of collaboration with respect to
+career stage (early-early, early-established, and
+established-established), then we can calculate the frequency of
+collaborators pairs comprising two women, or a woman and a man, and thus
+calcualte \(\alpha\) (see the Online Supplementary Material for the
+annotated R code).
+
+\hypertarget{data-and-code-availability}{%
+\subsection{Data and code
+availability}\label{data-and-code-availability}}
+
+The R scripts used to produce all results, figures and tables can be
+viewed online at \url{https://lukeholman.github.io/genderHomophily/}.
+The input data from Holman et al. {[}5{]} is archived at
+\url{https://osf.io/bt9ya/}.
 
@@
 
-We thank Devi Stuart-Fox and Dominique Potvin for helpful discussion.
+We thank Devi Stuart-Fox and Dominique Potvin for their very helpful
+comments on the manuscript.
 
@@
 
-S1 Fig. Plot showing the percentage of papers that have 1, 2, 3, 4, or
-\({\ge}5\) authors for each discipline in the dataset of Holman et al.
-(2018). This information can also be found in S3 Data.
-
-S2 Fig. Histogram showing the distribution of differences in \(\alpha'\)
-between the 2015-16 and 2005-6 samples, where positive numbers indicate
-an increase in \(\alpha'\) with time. The mean is slightly positive
-(i.e.~0.004), indicating a mild increase in average \(\alpha'\) with
-time.
-
-S3 Fig. Histogram showing the difference between \(\alpha'\) calculated
-for first and last authors. Positive values mean that \(\alpha'\) was
-higher when calculated for first authors, and negative values mean
-\(\alpha'\) was higher when calculated for last authors. The mean is
-very slightly higher than zero, indicating that \(\alpha'\) tends to be
-higher for first authors.
-
-S4 Fig. Histogram of \(\alpha'\) for 325 unique combinations of journal
-and country, using data from August 2015 - August 2016. The white areas
-denote combinations for which \(\alpha'\) differs significantly from
-zero (p \textless{} 0.05, following false discovery rate correction).
-
-S5 Fig. Plot showing the 68 combinations of journal and author country
-of affiliation for which \(\alpha'\) is significantly higher than
-expected.
-
-S6 Fig. Histogram showing the estimated degree to which \(\alpha'\) is
-inflated by inter-country differences in author gender ratio, across the
-285 journals for which we had adequate data after restricting the
-analysis by country. The average inflation in \(\alpha'\) is negligible,
-suggesting that Wahlund effects resulting from inter-country differences
-have a negligible effect on our estimates of gender homophily.
-
-S7 Fig. There is a very strong correlation between the values of
-\(\alpha\) and \(\alpha'\) calculated for each journal, though in a
+\textbf{S1 Fig}: Plot showing the percentage of papers that have 1, 2,
+3, 4, or \({\ge}5\) authors for each discipline in the dataset of Holman
+et al. (2018). This information can also be found in S3 Data.
+
+\textbf{S2 Fig}: Histogram showing the distribution of differences in
+\(\alpha'\) between the 2015-16 and 2005-6 samples, where positive
+numbers indicate an increase in \(\alpha'\) with time. The mean is
+slightly positive (i.e.~0.004), indicating a mild increase in average
+\(\alpha'\) with time.
+
+\textbf{S3 Fig}: Histogram showing the difference between \(\alpha'\)
+calculated for first and last authors. Positive values mean that
+\(\alpha'\) was higher when calculated for first authors, and negative
+values mean \(\alpha'\) was higher when calculated for last authors. The
+mean is very slightly higher than zero, indicating that \(\alpha'\)
+tends to be higher for first authors.
+
+\textbf{S4 Fig}: There is a weakly positive, non-linear relationship
+between the gender ratio of authors publishing in a journal, and the
+coefficient of homophily (\(\alpha'\)). Specifically, journals with 50\%
+women authors or higher tended to have more same-sex coauthorships than
+did journals in which most authors are men. This relationship held
+whether \(\alpha'\) was calculated for all authors or first authors
+only, but not for last authors only. A negative value on the x-axis
+denotes an excess of men authors, a positive value denotes an excess of
+women authors, and zero denotes gender parity (i.e.~equal numbers of
+male and female authors). The lines were fitted using generalised
+additive models with the smoothing parameter \(k\) set to 3.
+
+\textbf{S5 Fig}: Histogram of \(\alpha'\) for 325 unique combinations of
+journal and country, using data from August 2015 - August 2016. The
+white areas denote combinations for which \(\alpha'\) differs
+significantly from zero (p \textless{} 0.05, following false discovery
+rate correction).
+
+\textbf{S6 Fig}: Plot showing the 68 combinations of journal and author
+country of affiliation for which \(\alpha'\) is significantly higher
+than expected.
+
+\textbf{S7 Fig}: Histogram showing the estimated degree to which
+\(\alpha'\) is inflated by inter-country differences in author gender
+ratio, across the 285 journals for which we had adequate data after
+restricting the analysis by country. The average inflation in
+\(\alpha'\) is negligible, suggesting that Wahlund effects resulting
+from inter-country differences have a negligible effect on our estimates
+of gender homophily.
+
+\textbf{S8 Fig}: There is a very strong correlation between the values
+of \(\alpha\) and \(\alpha'\) calculated for each journal, though in a
 handful of cases the difference is considerable. The deviation between
 \(\alpha\) and \(\alpha'\) is greatest for journals for which there is a
-small sample size (see S8 Fig).
-
-S8 Fig. For journals for which we recovered a small number of papers
-(\textless{}100), the unadjusted metric \(\alpha\) was downwardly
+small sample size (see S9 Fig).
+
+\textbf{S9 Fig}: For journals for which we recovered a small number of
+papers (\textless{}100), the unadjusted metric \(\alpha\) was downwardly
 biased. This fits our expectations: because researchers cannot be their
@@
 
-S1 Table. Sample sizes for the two datasets, which comprise papers
-published in the timeframes August 2005 - August 2006, and August 2015 -
-August 2016.
-
-S2 Table. Number of journals showing statistically significant homophily
-or heterophily, in two one-year periods. The significance threshold was
-p \textless{} 0.05, and p-values were adjusted using Benjamini-Hochberg
-false discovery rate correction. Note that the power of our test is
-lower for the 2005-2006 data because fewer papers were recovered per
-journal: thus, it is not meaningful to compare the \% significant
-journals (i.e.~11\% vs 24\%) between the two time periods.
+\textbf{S1 Table}: Sample sizes for the two datasets, which comprise
+papers published in the timeframes August 2005 - August 2006, and August
+2015 - August 2016.
+
+\textbf{S2 Table}: Number of journals showing statistically significant
+homophily or heterophily, in two one-year periods. The significance
+threshold was p \textless{} 0.05, and p-values were adjusted using
+Benjamini-Hochberg false discovery rate correction. Note that the power
+of our test is lower for the 2005-2006 data because fewer papers were
+recovered per journal: thus, it is not meaningful to compare the \%
+significant journals (i.e.~11\% vs 24\%) between the two time periods.
 
@@
 
-S1 Data: This spreadsheet shows the \(\alpha\) values calculated for
-each journal, in the 2005 and 2015 samples, and for each type of author
-(all authors, first authors, and last authors). The tables gives the
-impact factor of each journal, the sample size, \(\alpha\) and
-\(\alpha'\) and their 95\% CIs, and the p-value from a 2-tailed test
-evaluating the null hypothesis that \(\alpha\) is zero (both raw and
-FDR-corrected p-values are shown).
-
-S2 Data: This file gives the number and percentage of papers that have
-1, 2, 3, 4, or \({\ge}5\) authors for each \emph{journal} in the dataset
-of Holman et al. (2018) \emph{PLOS Biology}. Note that the sample sizes
-include papers for which the gender of one or more authors was not
-determined by Holman et al.
-
-S3 Data: This file gives the number and percentage of papers that have
-1, 2, 3, 4, or \({\ge}5\) authors for each \emph{discipline} in the
-dataset of Holman et al. (2018) \emph{PLOS Biology}. Note that the
+\textbf{S1 Data}: This spreadsheet shows the \(\alpha\) values
+calculated for each journal, in the 2005 and 2015 samples, and for each
+type of author (all authors, first authors, and last authors). The
+tables gives the impact factor of each journal, the sample size,
+\(\alpha\) and \(\alpha'\) and their 95\% CIs, and the p-value from a
+2-tailed test evaluating the null hypothesis that \(\alpha\) is zero
+(both raw and FDR-corrected p-values are shown).
+
+\textbf{S2 Data}: This file gives the number and percentage of papers
+that have 1, 2, 3, 4, or \({\ge}5\) authors for each \emph{journal} in
+the dataset of Holman et al. (2018) \emph{PLOS Biology}. Note that the
 sample sizes include papers for which the gender of one or more authors
@@
 
-S4 Data. The table shows the distribution of the \(\alpha'\) values
-across journals, split by the research discipline. The gender ratio
-column shows the percentage of women authors in the sample used to
+\textbf{S3 Data}: This file gives the number and percentage of papers
+that have 1, 2, 3, 4, or \({\ge}5\) authors for each \emph{discipline}
+in the dataset of Holman et al. (2018) \emph{PLOS Biology}. Note that
+the sample sizes include papers for which the gender of one or more
+authors was not determined by Holman et al.
+
+\textbf{S4 Data}: The table shows the distribution of the \(\alpha'\)
+values across journals, split by the research discipline. The gender
+ratio column shows the percentage of women authors in the sample used to
 calculate \(\alpha'\), across all authorship positions. In the last two
